### 2020-10-17 distribution_theory/C1.tex
added current summaries

\section{Distributions}
%\begin{recap}
%    A \emph{topological field} $\KK$ is a field endowed with a topology such that addition, multiplication, and inversion are continuous.
%    
%    A \emph{topological vector space} $V$ over a topological field $\KK$ is a vector space endowed with a topology such that vector addition and scalar multiplication are continuous. 
%    
%    The \emph{topological dual space} $V^*$ of $V$ is the set of all continuous linear maps from $V$ to $\KK$. 
%\end{recap}
\subsection{Test functions and distributions}
\begin{definition}
    Let $X \subseteq \RR^n$ be open, then we define the set of \emph{test functions} on $X$ as
    \[
    \Cal D(X) \ceq C^\infty_0(X) = \qty{f \colon X \to \CC \mid f \text{ is smooth with compact support}}. 
    \]
\end{definition}

\begin{definition} \label{def:test_convergence}
    Let $(\phi_m) \subseteq \Cal D(X)$. We say that $(\phi_m) \to 0$ in $\Cal D(X)$ if
    \begin{enumerate}
        \item  there exists a compact $K \subseteq X$ such that $\supp\phi_m \subseteq K$ for all $m$;
        \item $\pt^\alpha \phi_m \to 0$ uniformly for each multi-index $\alpha$.
    \end{enumerate}
\end{definition}

Note that, for any $\phi, \psi \in \Cal D(X)$ and any multi-index $\alpha$ we have
\[
\int_X \phi\cdot  \pt^\alpha \psi \dd{x} = (-1)^{\abs{\alpha}} \int_X \psi \cdot \pt^\alpha \phi \dd{x}, 
\]
which follows from partial integration and the fact that all boundary terms vanish since $\phi$ and $\psi$ have compact support.

Also, by Taylor's theorem, for any $\phi \in \Cal D(X)$, $x, h \in X$ and $N \in \NN$ we have
\[
\phi(x + h) = \sum_{\abs{\alpha} \leq N} \frac{h^\alpha}{\alpha!} \pt^\alpha \phi(x) + R_N(x, h) \quad\text{where } R_N(x, h) = o(\abs{h}^N) \text{ uniformly in $x$}. 
\]

\begin{definition}
    A \emph{distribution} on $X$ is a linear map $u \colon \Cal D(X) \to \CC$ if for every compact set $K \subseteq X$ there exist constants $C, N$ such that for all $\phi \in \Cal D(X)$ with $\supp\phi\subseteq K$ we have
    \begin{equation} \label{eq:seminorm_cond}
        \abs{u(\phi)} \leq C \sum_{\abs{\alpha} \leq N} \sup \abs{\pt^\alpha \phi}.
    \end{equation}

Condition \ref{eq:seminorm_cond} is called the \emph{seminorm condition}. If, in the seminorm condition, the same $N$ can be used for every compact set $K \subseteq X$, then the least such $N$ is called the \emph{order} of $u$, written $\ord(u)$.

The set of all distributions in $X$ is denoted $\Cal D'(X)$. 
\end{definition}

If $u \in \DD'(X)$ and $\phi \in \DD(X)$, then instead of $u(\phi)$ we usually write $\ang{u, \phi}$.

\begin{recap}
    A function $f \colon X \to \CC$ is called \emph{locally integrable} if $\int_K \abs{f} \dd{x} < \infty$ for all compact $K \subseteq X$.

    The set of locally integrable functions on $X$ is denoted $\locint(X)$. 
\end{recap}

\begin{example}
    Let $M \in \NN$ and let $f_\alpha \in \locint(X)$ for all $\abs{\alpha} \leq M$. Define the linear map $T \colon \DD(X) \to \CC$ by
    \[
    \ang{T, \phi} \ceq \sum_{\abs{\alpha} \leq M} \int_X f_\alpha \cdot \pt^\alpha \phi \dd{x}. 
    \]
    It is trivial that $T$ is linear, and we verify that $T$ is a distribution as follows: take $\phi \in \Cal D(X)$ with $\supp\phi \subseteq K$. Then we have
    \begin{align*}
        \abs{\ang{T, \phi}} &\leq \sum_{\abs{\alpha} \leq M} \int_K \abs{f_\alpha} \cdot \abs{\pt^\alpha \phi} \dd{x} \\
        &\leq \sum_{\abs{\alpha} \leq M} \sup \abs{\pt^\alpha \phi} \cdot \int_K \abs{f_\alpha} \dd{x} \\
        &\leq \qty(\max_\alpha \int_K \abs{f_\alpha} \dd{x})   \sum_{\abs{\alpha} \leq M} \sup \abs{\pt^\alpha \phi}. 
    \end{align*}
Therefore, the seminorm condition is satisfied with $N = M$. From this, it also follows that $\ord(T) \leq M$. 
\end{example}

A special case of the previous example is the case $M = 0$: in this case the distribution simply becomes
\[
\ang{T_f, \phi} = \int_X f \phi \dd{x}. 
\]
Henceforth we will abuse notation: if $f \in \locint(X)$, then we will write $f$ instead of $T_f$, i.e., $\ang{f, \phi} = \int_X f\phi \dd{x}$.

\begin{lemma}
    Let $u \colon \DD(X) \to \CC$ be a linear map. Then $u$ is a distribution if and only if, for every sequence $(\phi_j) \subseteq \DD(X)$ with $\phi_j \to 0$ as in \cref{def:test_convergence}, we have $\ang{u, \phi_m} \to 0$. 
\end{lemma}

\begin{proof}
    `$\implies$' If $u$ is a distribution and $(\phi_m) \to 0$, then $\supp\phi_m \subseteq K$ for some compact $K$, and by \cref{eq:seminorm_cond} there exist $C, N$ such that
    \[
    \abs{\ang{u, \phi_m}} \leq C \sum_{\abs{\alpha} \leq N} \sup \abs{\pt^\alpha \phi_m} \to 0. 
    \]

    `$\impliedby$' Suppose there is a compact set $K$ such that \cref{eq:seminorm_cond} is not valid for any $C, N$. Let $m \in \NN$ and $C = N = m$, then there is some $\phi_m$ with $\supp(\phi_m) \subseteq K$, and 
    \[
    \abs{\ang{u, \phi_m}} > m \sum_{\abs{\alpha} \leq m} \sup \abs{\pt^\alpha \phi_m}. 
    \]
    By dividing $\phi_m$ by $\ang{u, \phi_m} \neq 0$, we may assume w.l.o.g.\ that $\ang{u, \phi_m} = 1$. 
    We now have a sequence $(\phi_m)$ such that
    \[
    \frac1m > \sum_{\abs{\alpha} \leq m} \sup \abs{\pt^\alpha \phi_m} \implies  \abs{\pt^\alpha \phi_m} < \frac1m \quad\text{ for $\abs{\alpha} \leq m$} \implies \pt^\alpha \phi_m \to 0 \text{ uniformly for all $\alpha$}. 
    \]
    Since each $\phi_m$ also satisfies $\supp\phi_m\subseteq K$, by \cref{def:test_convergence} we have that $\phi_m \to 0$, but also $\ang{u, \phi_m} \to 1$, a contradiction. 
\end{proof}

\subsection{Limits in $\DD'(X)$}
\begin{definition}
    We say that a sequence $(u_m) \subseteq \DD'(X)$ converges to $u \in \DD'(X)$ and write $u_m \to u$ if 
    \[
    \ang{u_m, \phi} \to \ang{u, \phi} \quad\text{for all $\phi \in \DD(X)$}. 
    \]
\end{definition}

\begin{mdframed}
The following theorem is non-examinable but interesting: 
\begin{theorem}
    Let $(u_m)$ be a sequence in $\DD'(X)$ such that $\lim\limits_{m\to\infty} \ang{u_m, \phi}$ exists for all $\phi \in \DD(X)$. Then the map $\ang{u, \phi} \ceq \lim\limits_{m\to\infty} \ang{u_m, \phi}$ is a distribution in $X$.
\end{theorem}

\begin{proof}
    This is a direct application of the uniform boundedness principle. 
\end{proof}
\end{mdframed}

\begin{example}
    Let $X = \RR$ and consider the sequence of functions $u_m \in \locint(\RR)$ defined by $u_m(x) = \sin(mx)$. Then, for all $\phi \in \DD(\RR)$, we have
    \[
    \ang{u_m, \phi} = \int_\RR \sin(mx) \phi(x) \dd{x} = \frac1m \int_\RR \cos(mx) \phi'(x) \dd{x} \leq \frac1m \int \abs{\phi'(x)} \dd{x} \to 0.   \]
    Therefore, it holds that $u_m \to 0$ in $\DD'(\RR)$. With our abuse of notation we write this as $\lim\limits_{m\to\infty} \sin(mx) = 0$ in $\DD'(\RR)$. 
\end{example}

\subsection{Basic operations}

\subsubsection{Differentiation and multiplication by smooth functions}
For $u \in C^\infty(X)$ and $\phi \in \DD(X)$, we have noted that
\[
\ang{\pt^\alpha u, \phi} = \int_X \pt^\alpha u \cdot \phi \dd{x} = (-1)^\abs{\alpha}  \int_X u \cdot \pt^\alpha \phi \dd{x} = \ang{u, (-1)^\abs{\alpha} \pt^\alpha \phi}. 
\]

Since the RHS makes sense for any distribution $u$, we define 
\begin{definition}
    For $f \in C^\infty(X)$, $u \in \DD'(X)$, we define $\pt^\alpha(fu)$ by
    \[
    \ang{\pt^\alpha(fu), \phi}  \ceq \ang{u, (-1)^\abs{\alpha} f \cdot  \pt^\alpha \phi}
    \]
\end{definition}

\begin{remark}
    This definition outlines a more general pattern when working with distributions: first we take some well-defined operator on the collection of smooth maps, then we rewrite it to a form that is sensible for any distribution, and then we \emph{define} that new form as the operator on distributions. This process is called \emph{extending the definition by duality}.
\end{remark}

\begin{example}
    Let $u = \delta_x$, then we have
    \[
    \ang{\pt^\alpha \delta_x, \phi} = \ang{\delta_x, (-1)^\abs{\alpha} \pt^\alpha\phi} = (-1)^\abs{\alpha} \pt^\alpha \phi(x)
    \]

    Furthermore, consider the Heaviside function $H(x) = \ind_{x \geq 0}$. We have
    \[
    \ang{H', \phi} = - \ang{H, \phi'} = -\int_\RR H(x) \phi'(x) \dd{x} = - \int_0^\infty \phi'(x) \dd{x} = \phi(0) = \ang{\delta_0, \phi}, 
    \]
    so we write $H'= \delta_0$ \emph{in the distributional sense}. 
\end{example}

\begin{lemma}
    Suppose $u' \in \DD'(\RR)$ satisfies $u' = 0$. Then $u$ is constant (i.e., $\ang{u, \phi} = \ang{c, \phi} = c \int_\RR \phi \dd{x}$ for some $c$). 
\end{lemma}

\begin{proof}
    Fix  any $\theta \in \DD(\RR)$ with $\ang{1, \theta} = 1$. Let $\phi \in \DD(\RR)$ and define
    \[
    \phi_A \ceq \phi - \ang{1, \phi}\theta, \quad \phi_B \ceq \ang{1, \phi}\theta \quad\text{such that }\phi = \phi_A + \phi_B.
    \]

    Note that $\ang{1, \phi_A} = \ang{1, \phi} - \ang{1, \phi}\ang{1, \theta} = 0$.

    We claim that the function $\Phi_A(x) \ceq \int_{-\infty}^x \phi_A(y) \dd{y}$ has compact support: since $\supp\phi_A \subseteq [a, b]$ for some $a, b \in \RR$, clearly $\Phi_A(x) = 0$ for $x < a$, while for $x > b$ we have $\Phi_A(x) = \ang{1, \phi_A} = 0$. 
    Obviously, it holds that $\Phi_A' = \phi_a$. Now we compute
    \[
    \ang{u, \phi} = \ang{u, \phi_A} + \ang{u, \phi_B} = \ang{u, \Phi_A'} + \ang{1, \phi}\ang{u, \theta} = - \ang{u', \Phi_A} + c \ang{1, \phi} = \ang{c, \phi}. 
    \]
    Since $\phi$ was chosen arbitrarily this shows that $u$ is constant. 
\end{proof}

\subsubsection{Reflection and translation} 
For $\phi \in \DD(\RR^n), h \in \RR^n$, define the \emph{translation of $\phi$ by $h$} by 
\[
(T_h\phi)(x) \ceq \phi(x - h),
\]
and the \emph{reflection of $\phi$} by $\refl\phi(x) \ceq \phi(-x)$.

Extending the definitions of translation and reflection by duality yields the following:
\begin{definition}
    For $u \in \DD'(\RR^n)$, $h \in \RR^n$, $\phi \in \DD(\RR^n)$, define
    \[
    \ang{T_hu, \phi} \ceq \ang{u, T_{-h}\phi} \quad\text{and}\quad \ang{\refl u, \phi} \ceq \ang{u, \refl{\phi}}. 
    \]
\end{definition}

\begin{lemma}
    For $u \in \DD'(\RR^n)$, define $V_h \in \DD'(\RR^n)$ for $0 \neq h \in \RR^n$ by
    \[
    V_h \ceq \frac{T_{-h}u - u}{\norm{h}}
    \]
    If $(h_j) \subseteq \RR^n$ is a sequence for which $\lim_{j \to \infty} \frac{h_j}{\norm{h_j}} = m \in S^{n-1}$, then $V_{h_j} \to \sum_i m_i \pdv{u}{x_i}$ in $\DD'(\RR^n)$. 
\end{lemma}

\begin{proof}
    By definition, we can write $\ang{V_h, \phi} = \frac{1}{\norm{h}} \ang{u, T_h\phi - \phi}$. Now Taylor's theorem tells us that
    \[
    (T_h\phi - \phi)(x) = \phi(x - h) - \phi(x) = - \sum_i h_i \pdv{\phi}{x_i}(x) + R(x, h),
    \]
    where $R(x, h) = o(\norm{h})$ in $D(\RR^n)$ (exercise sheet 1, question 2).

    By sequential continuity, we have
    \[
    \lim_{j \to \infty} \ang{V_{h_j}, \phi} = \ang{u, - \sum_i m_i \pdv{\phi}{x_i}} = \ang{\sum_i m_i \pdv{u}{x_i}, \phi}, 
    \]
    which shows that $V_{h_j} \to \sum_i m_i \pdv{u}{x_i}$ in $\DD'(\RR^n)$. 
\end{proof}

\subsubsection{Convolution}
For $\phi \in \DD(\RR^n)$, note that $(T_x \refl\phi)(y) = \refl\phi(y-x) = \phi(x-y)$. 
\begin{definition}
    For $u \in C^\infty(\RR^n), \phi \in \DD(\RR^n)$, we define the \emph{convolution} $u * \phi \colon \RR^n \to \CC$ as
    \[
    (u * \phi)(x) \ceq \int_{\RR^n} u(x-y) \phi(y) \dd{y} = \int_{\RR^n} u(y) \phi(x - y) \dd{y} = \ang{u, T_x \refl\phi}. 
    \]
    Since the RHS makes sense for any $u \in \DD'(\RR^n)$, we extend the definition this way: for $u \in \DD'(\RR^n), \phi \in \DD(\RR^n)$, we define the convolution $u * \phi$ as
    \[
    (u * \phi)(x) \ceq \ang{u, T_x\refl\phi}. 
    \]
\end{definition}

\begin{lemma}
    Let $\phi \in C^\infty(\RR^n \times \RR^m)$ and define $\Phi_x(y) \ceq \phi(x, y)$. Suppose for any $x \in \RR^n$ there exists a neighbourhood $N(x)$ and a compact $K \subseteq \RR^m$ such that $\phi(x, y)$ for all $x \in N(x), y \notin K$.

    Then $x \mapsto \ang{u, \Phi_x}$ is differentiable with
    \[
    \pt_x^\alpha \ang{u, \Phi_x} = \ang{u, \pt_x^\alpha \Phi_x}
    \]
    for any $u \in \DD'(\RR^m)$. 
\end{lemma}

\begin{proof}
    Fix $x \in \RR^n$, then by Taylor's formula we have
    \[
    \Phi_{x + h}(y) = \Phi_x(y) + \sum_{i=1}^n h_i \pdv{\phi}{x_i} (x, y)+ R(x, y, h), 
    \]
    where $\pt_y^\alpha R(x, y, h) = o(\norm{h})$, uniformly in $y$, for any multi-index $\alpha$. Furthermore, by assumption there exists a compact $K$ such that for $h$ small enough, $\supp R(x, \cdot, h) \subseteq K$. Therefore, $R(x, \cdot, h)$ is a test function for $h$ small enough.

    Combining the previous two facts shows that $R(x, \cdot, h) = o(\norm{h})$ in $\DD(\RR^m)$ as $h \to 0$.

    Let $u \in \DD'(\RR^m)$, then we find by sequential continuity that $\ang{u, R(x, \cdot, h)}$ is also $o(\norm{h})$, and therefore
    \[
    \ang{u, \Phi_{x+h}} = \ang{u, \Phi_x} + \sum_{i=1}^n h_i \ang{u, \pdv{\Phi_x}{x_i}} + o(\norm{h}). 
    \]
    This shows that $x \mapsto \ang{u, \Phi_x}$ is differentiable with 
    \[
    \pdv{x_i} \ang{u, \Phi_x}  = \ang{u, \pdv{\Phi_x}{x_i}}. 
    \]
    From this the result follows. 
\end{proof}

\begin{corollary}
    If $u \in \DD'(\RR^n)$, $\phi \in \DD(\RR^n)$, then $u * \phi$ is differentiable with $\pt^\alpha (u * \phi) = u * \pt^\alpha\phi$. 
\end{corollary}

\begin{proof}
    Apply the previous lemma with $\Phi_x(y) \ceq  \phi(x-y)$. 
\end{proof}

### 2020-10-17 inverse_problems/C1.tex
added current summaries

\begin{mdframed}
A \emph{direct problem} is a problem where given an object or \emph{cause}, we must determine the data or \emph{effect}. In an \emph{inverse problem}, we observe data and wish to recover the object.

A problem is called \emph{well-posed} if a unique solution exists that depends continuously on the data. Most inverse problems are, unfortunately, ill-posed. 
\end{mdframed}

\section{Generalised Solutions}
\begin{recap}
    \begin{enumerate}
        \item     An operator $A \in \LLL(\XXX, \YYY)$ is called \emph{bounded} if
        \[
        \norm{A}_{\LLL(\XXX, \YYY)} \ceq \sup_{u \neq 0} \frac{\norm{Au}_\YYY}{\norm{u}_\XXX} = \sup_{\norm{u}_\XXX \leq 1} < \infty. 
        \]
            It is known that a linear operator between normed spaces is continuous if and only if it is bounded. 
            \item 
        We let $\Dom(A)$, $\Nul(A)$ and  $\Ran(A)$ denote the domain, null space, and range of $A$ respectively.

        \item     We will assume $\XXX$ and $\YYY$ are Hilbert spaces, so there is an inner product $\ang{\cdot, \cdot}$ and any bounded operator $A \in \LLL(\XXX, \YYY)$ has a unique adjoint $A^* \in \LLL(\YYY, \XXX)$ which satisfies
        \[
        \ang{Au, v}_\YYY = \ang{u, A^* v}_\XXX \quad\text{for all $u \in \XXX$, $v \in \YYY$.}
        \]

        \item For any $\XXX' \subseteq \XXX$ we define the \emph{orthogonal complement} of $\XXX'$ as
        \[
        (\XXX')\P \ceq \qty{u \in \XXX \mid \ang{u, v}_\XXX = 0 \ \forall v \in \XXX'}. 
        \]
        It is known that $(\XXX')\P$ is a closed subspace of $\XXX$ and that $\XXX' \subseteq ((\XXX')\P)\P$, where equality holds if and only if $\XXX'$ is a closed subspace of $\XXX$. For a non-closed subspace $\XXX'$ we have $((\XXX')\P)\P = \clos{\XXX'}$.

        \item If $\XXX'$ is a closed subspace of $\XXX$, then for any $u \in \XXX$ there exist unique $x_u \in \XXX'$, $x_u\P \in (\XXX')\P$ such that $u = x_u + x_u\P$. The map $u \mapsto x_u$ is denoted $P_{\XXX'}$ and is called the \emph{orthogonal projection} on $\XXX'$. Properties are:
        \begin{enumerate}
            \item $P_{\XXX'}$ is bounded and self-adjoint with norm 1;
            \item $P_{\XXX'} + P_{(\XXX')\P} = I$;
            \item $P_{\XXX'}u$ minimises the distance from $u$ to $\XXX'$;
            \item $x = P_{\XXX'}u$ if and only if $x \in \XXX'$ and $u - x \in (\XXX')\P$. 
        \end{enumerate}

    \item For any $A \in \LLL(\XXX, \YYY)$ we have
    \[
    \Ran(A)\P = \Nul(A^*) \quad\text{and}\quad \Nul(A)\P = \clos{\Ran(A^*)}. 
    \]

     \end{enumerate}
\end{recap}

\begin{lemma}
    For any $A \in \LLL(\XXX, \YYY)$ we have $\clos{\Ran(A^*A)} = \clos{\Ran(A^*)}$. 
\end{lemma}

\begin{proof}
    It is trivial that $\clos{\Ran(A^*A)} \subseteq \clos{\Ran(A^*)}$.

    Now, suppose $u \in \clos{\Ran(A^*)}$ and let $\eps > 0$. Then there exists $v \in \XXX$ such that $\norm{A^*v - u} < \eps/2$. Writing $v = e + f$ with $e \in \Nul(A^*)$, $f \in \Nul(A^*)\P = \clos{\Ran(A)}$, we see that $\norm{A^*f - u} < \eps/2$.

    Since $f \in \clos{\Ran(A)}$, there exists $x \in \XXX$ such that $\norm{Ax - f} < \eps/(2\norm{A})$. We now compute
    \[
    \norm{A^* A x - u} \leq \norm{A^*A x - A^* f} + \norm{A^*f - u} < \norm{A^*} \frac{\eps}{2\norm{A}} + \frac\eps2 = \eps, 
    \]
    and conclude that $u \in \clos{\Ran(A^*A)}$. This shows that $\clos{\Ran(A)} \subseteq \clos{\Ran(A^*A)}$. 
\end{proof}

\subsection{Generalised inverses}
We consider the equation \begin{equation} \label{eq:basic_ip}
    Au = f,
\end{equation} $A \in \LLL(\XXX, \YYY)$, $f$ is known, and we wish to find $u$. 
\begin{definition}
    An element $u \in \XXX$ is called a \emph{least-squares solution} of \cref{eq:basic_ip} if $u$ is a minimiser of the function $v \mapsto \norm{Av - f}_\YYY$. It is called a \emph{minimal-norm solution} of \cref{eq:basic_ip} if it has minimal norm among all least-squares solutions. 
\end{definition}

Note that a least-squares solution may not exist. If a least-squares solution $u$ exists, then the affine subspace of all least-squares solutions is given by $u + \Nul(A)$. By writing $u = u\D + v$ for $u\D \in \Nul(A)\P$, $v \in \Nul(A)$, we find that the space of least-squares solutions is given by $u\D + \Nul(A)$, and it is now clear that $u\D$ is the unique minimum-norm solution.

\begin{theorem}
    Let $f \in \YYY$ and $A \in \LLL(\XXX, \YYY)$. Then the following are equivalent:
    \begin{enumerate}
        \item $u \in \XXX$ satisfies $Au = P_{\clos{\Ran(A)} }f$;
        \item $u$ is a least-squares solution of $\cref{eq:basic_ip}$:
        \item $u$ solves the \uline{normal equation}
        \begin{equation} \label{eq:normal}
                   A^* f = A^* A u. 
        \end{equation}
    \end{enumerate}
\end{theorem}

\begin{proof}
     ``(1) $\implies$ (2)'': We have
     \[
     \norm{Au - f}_\YYY = \norm{ P_{\clos{\Ran(A)} }f - f} = \inf_{g \in \clos{\Ran(A)}} \norm{g - f} \leq \inf_{g \in \Ran(A)} \norm{g - f} = \inf_{u \in \XXX} \norm{Au - f}. 
     \]

     ``(2) $\implies$ (3)'':  Let $u \in \XXX$ be a least-squares solution and $v \in \XXX$ arbitrary. Define the quadratic polynomial 
     \begin{align*}
     F \colon \RR \to \RR \colon \lambda \mapsto & \norm{A(u + \lambda v) - f}^2 \\
     &= \ang{Au + \lambda Av - f, Au + \lambda Av - f} \\
     &= \lambda^2 \norm{Av}^2 - 2\lambda \ang{Av, f - Au} + \norm{f - Au}^2. 
     \end{align*}
    As $u$ is a least-squares solution, we know that $F$ attains a minimum in $\lambda =0 $ and therefore that 
    \[
    0 = F'(0) = 2\ang{Av, f - Au} = 2 \ang{v, A^*(f - Au)}.
    \]
    Since $v$ is arbitrary, we must have $A^*(f - Au) = 0$, so $u$ satisfies $\cref{eq:normal}$.

    ``(3) $\implies$ (1)'':  From the normal equation we know that $A^*(f - Au) = 0$. For any $x \in \XXX$, we have
    \[
    \ang{Ax, f - Au} = \ang{x, A^*(f - Au)} = \ang{x, 0} = 0,
    \]
    so $f - Au \in \Ran(A)\P$.

    So we have $Au \in \clos{\Ran(A)}$ and $f - Au \in \Ran(A)\P = \clos{\Ran(A)}\P$, from which it follows that $Au = P_{\clos{\Ran(A)}}f$. 
\end{proof}

The following lemma gives a precise condition for when a least-squares solution exists: 
\begin{lemma}
    \Cref{eq:basic_ip} has a least-squares solution if and only if $f \in \Ran(A) \oplus \Ran(A)\P$. 
\end{lemma}

\begin{proof}
    ``$\implies$'' Suppose $u$ is a least-squares solution. Then $f - Au \in \Ran(A)\P$, so $f = Au + (f - Au) \in \Ran(A) \oplus \Ran(A)\P$.

    ``$\impliedby$'' Suppose $f = Au + g$ for some $u \in \XXX$, $g \in \Ran(A)\P = \clos{\Ran(A)}\P$. Then by the previous theorem, $Au = P_{\clos{\Ran(A)}}f$, so $u$ is a least-squares solution. 
\end{proof}

\begin{corollary}
    If $\Ran(A)$ is closed, then \cref{eq:basic_ip} always has a least-squares solution. 
\end{corollary}
In particular, this holds if $\Ran(A)$ is finite-dimensional. Therefore, if either $\XXX$ or $\YYY$ is finite-dimensional, \cref{eq:basic_ip} has a least-squares solution for any $A$.

We have already seen that if a least-squares solution $u$ exists, then the affine subspace of all least-sqares solutions is $u + \Nul(A)$, and the unique minimum-norm solution is the projection of 0 onto this affine subspace, which is the unique element of $u + \Nul(A)$ that lies in $\Nul(A)\P$.

\begin{definition}
    Let $A \in \LLL(\XXX, \YYY)$, and define
    \[
    \tilde A \ceq A \restr_{\Nul(A)\P} \colon \Nul(A)\P \to \Ran(A). 
    \]
    Clearly $\tilde A$ is bijective and we define the \emph{Moore-Penrose inverse} 
    \[
    A\D \colon \Ran(A) \oplus \Ran(A)\P \to \Nul(A)\P \colon f \mapsto \tilde A^{-1} P_{\clos{\Ran(A)}} f. 
    \]
\end{definition}

\begin{remark}
    Note that $\clos{\Ran(A) \oplus \Ran(A)\P} = \clos{\Ran(A)} \oplus \Ran(A)\P = \clos{\Ran(A)} \oplus \clos{\Ran(A)}\P = \YYY$, and therefore the operator $\tilde A$ is \emph{densely defined}, and it is defined on all of $
    \YYY$ if and only if $\Ran(A)$ is closed. 
\end{remark}

We will not prove the following theorem, but it is interesting: 
\begin{theorem} \label{thm:mp_inverse_continuous}
    The Moore-Penrose inverse $A\D$ is continuous if and only if $\Ran(A)$ is closed. 
\end{theorem}

The following characterises all important facts about the Moore-Penrose inverse: 
\begin{theorem}[Moore-Penrose equations] \label{thm:mp_equations}
    The operator $A\D$ satisfies the following equations: 
    \begin{enumerate}[(1)]
        \item $A\D A = P_{\Nul(A)\P}$; 
        \item $AA\D = P_{\clos{\Ran(A)}}\restr_{
        \Dom(A\D)}$; 
        \item $AA\D A = A$;
        \item $A\D AA\D = A\D$. 
    \end{enumerate}

Conversely, if any linear operator $B \colon \YYY \to \XXX$ satisfies (1) and (2), then $B = A\D$.  
\end{theorem}

\begin{proof}
    We will not prove (1) and (2). Point (3) and (4) follow immediately from (1) and (2) respectively. 
\end{proof}

The Moore-Penrose inverse has the important property that it maps every $f$ in its domain to the corresponding minimum-norm least-squares solution:
\begin{theorem}
    For every $f \in \Dom(A\D)$, the minimum-norm solution $u\D$ to \cref{eq:basic_ip} is given by $u\D = A\D f$. 
\end{theorem}

\begin{proof}
    Since $f \in \Ran(A) \oplus \Ran(A)\P$, we know that there exists a unique minimum-norm solution $u\D \in \Nul(A)\D$. We write
    \[
    u\D = P_{\Nul(A)\P}(u\D) = A\D A u\D = A\D P_{\clos{\Ran(A)}} f = A\D A A\D f = A\D f. 
    \]
\end{proof}

\begin{remark}
    We can also consider the normal equation $A^* f = A^* A u$ as a least-squares problem, whose minimum-norm solution is $(A^* A)\D A^* f$. It is clear that this expression must equal the minimum-norm solution $u\D$ from \cref{eq:basic_ip}. 
\end{remark}

\subsection{Compact operators}
\begin{definition}
    Let $A \in \LLL(\XXX, \YYY)$. Then $A$ is called \emph{compact} if for any bounded $B \subseteq \XXX$, the image $A(B)$ is precompact in $\YYY$. The set of compact operators in $\LLL(\XXX, \YYY)$ is denoted $\KKK(\XXX, \YYY)$. 
\end{definition}

\begin{lemma}
    Let $A \in \LLL(\XXX, \YYY)$. Then $A$ is compact if and only if, for every bounded sequence $(x_n) \subseteq X$, the sequence $(Ax_n) \subseteq Y$ has a convergent subsequence. 
\end{lemma}

\begin{theorem}
    Let $A \in \KKK(\XXX, \YYY)$ with $\dim(\Ran(A)) = \infty$. Then $A\D$ is discontinuous. 
\end{theorem}

\begin{proof}
    If $\dim\Ran(A) = \infty$, then $\XXX$ and $\Nul(A)\P$ are infinite-dimensional as well. Chose an orthonormal sequence $(x_n) \subseteq \Nul(A)\P$, then after taking a subsequence if necessary, we can assume that $f_n \ceq Ax_n$ converges. However, we have
    \[
    \norm{A\D(f_n - f_m)}^2 = \norm{A\D A (x_n - x_m)}^2 = \norm{P_{\Nul(A)\P} (x_n - x_m)}^2 = \norm{x_n -x_m}^2 = 2, 
    \]
    and in particular the sequence $(A\D f_n)$ does not converge. This shows that $A\D$ is discontinuous. 
\end{proof}

In particular, combining this with \cref{thm:mp_inverse_continuous} shows that the range of a compact operator is always open, and that not every element in $\YYY$ has a least-squares solution.

We will need the following theorem, an infinite-dimensional analogue of the spectral theorem: 
\begin{theorem}[Eigenvalue decomposition of self-adjoint compact operators] Let $\XXX$ be a Hilbert space, and $A \in \KKK(\XXX, \XXX)$ self-adjoint. Then there exists an orthonormal basis $(x_j)$ of $
   \clos{\Ran(A)}$ and a sequence of eigenvalues $\abs{\lambda_1} \geq \abs{\lambda_2} \geq \dotsb > 0$ such that for all $u \in \XXX$ we have
    \[
    Au = \sum_{j=1}^\infty \lambda_j \ang{u, x_j} x_j. 
    \]
The sequence $(\lambda_j)$ is either finite or converges to 0. 
\end{theorem}

The previous theorem gives rise to an infinite-dimensional analogue of the SVD:
\begin{theorem}
    Let $A \in \KKK(\XXX, \YYY)$. Then there exists a (not necessarily infinite) sequence $\sigma_1 \geq \sigma_2 \geq \dotsb > 0$ converging to 0, and orthonormal bases $(x_j)$, $(y_j)$ of $\Nul(A)\P$ and $\clos{\Ran(A)}$ respectively, such that
    \[
    Ax_j = \sigma_j y_j, \quad A^* y_j = \sigma_j x_j \quad\text{for all $j \in \NN$}, 
    \]
    and such that for all $u \in \XXX$ and $f \in \YYY$ we have
    \[
    Au = \sum_{j=1}^\infty \sigma_j \ang{u, x_j} y_j, \quad A^* f = \sum_{j=1}^\infty \sigma_j \ang{f, y_j} x_j.
    \]

    The sequence $\qty{(\sigma_j, x_j, y_j)}$ is called the \uline{singular value decomposition} (SVD) of $A$.
\end{theorem}

\begin{proof}
    Define $B \ceq A^*A$ and $C \ceq AA^*$, which are both compact, self-adjoint, and positive semi-definite operators. By the previous theorem, we can write
    \[
     Cf = \sum_{j=1}^\infty \sigma_j^2 \ang{f, y_j} y_j,
    \]
    where $(y_j)$ is a basis of $\clos{\Ran(AA^*)} = \clos{\Ran(A)}$ and $(\sigma_j)$ is a positive decreasing sequence converging to 0.

    Note that
    \[
    B A^* y_j = A^* A A y_j = A^* C y_j = A^* \sigma^2 y_j = \sigma_j^2 A^* y_j,
    \]
    so $A^* y_j$ is an eigenvector of $B$ with eigenvector $\sigma_j^2$.

    We show that $\qty(\frac{A^* y_j}{\sigma_j})$ is an orthonormal basis of $\Ran(A)\P$. is an orthonormal basis of $\Nul(A)\P$: their inner product is given by 
    \[
    \ang*{\frac{A^* y_j}{\sigma_j}, \frac{A^* y_k}{\sigma_k}} = \frac{1}{\sigma_j\sigma_k} \ang{y_j, C y_k} = \frac{\sigma_k}{\sigma_j} \ang{y_j, y_k} = 0, 
    \]
    and since the $(y_j)$ are a basis of $\clos{\Ran(A)} = \Nul(A^*)\P$ it is clear that the span of $(A^* y_j)$ is dense in $\clos{\Ran(A^*)} = \Nul(A)\P$.

    If we choose $x_j = \frac{A^* y_j}{\sigma_j}$, we find by construction that $A^* y_j = \sigma_j x_j$ and 
    \[
    A x_j = \frac{AA^* y_j}{\sigma_j} = \frac{C y_j}{\sigma_j} = \sigma_j y_j.
    \]

    Finally, we see that
    \[
    A u = \sum_{j=1}^\infty  \ang{u, x_j} Ax_j = \sum_{j=1}^\infty \sigma_j \ang{u, x_j} y_j \quad\text{and}\quad A^* f = \sum_{j=1}^\infty \ang{f, y_j} A^* y_j = \sum_{j=1}^\infty \sigma_j \ang{f, y_j} x_j. 
    \]
\end{proof}

\begin{theorem}
    Let $A \in \KKK(\XXX, \YYY)$ with SVD $\qty{(\sigma_j, x_j, y_j)}$ and let $f \in \Dom(A\D)$. Then
    \[
    A\D f = \sum_{j=1}^\infty \frac{1}{\sigma_j} \ang{f, y_j} x_j.
    \]
\end{theorem}

\begin{remark}
    Note that this is comparable to $A^* f = \sum_{j=1}^\infty \sigma_j \ang{f, y_j} x_j$, except that $A^*$ is a smoothing operator (since $\sigma_j \to 0$), while $A\D$ does the opposite. Furthermore, $A\D$ amplifies the right singular vectors corresponding to small singular values the most --- intuitively, the corresponding left singular vectors are vectors where $A$ doesn't ``see much''. 
\end{remark}

\begin{proof}
    Define $Bf = \sum_{j=1}^\infty \frac1{\sigma_j} \ang{f, y_j} x_j$. Then by \cref{thm:mp_equations}, we must check that $BA = P_{\Nul(A)\P}$ and $AB = P_{\clos{\Ran(A)}} \restr_{\Dom(A\P)}$.

    For the first equation, we compute
    \[
    BAu = \sum_{j=1}^\infty \frac{1}{\sigma_j} \ang*{\sum_{i=1}^\infty \sigma_i \ang{u, x_i}y_i, y_j} x_j = \sum_{j=1}^\infty \sum_{i=1}^\infty \frac{\sigma_i}{\sigma_j} \ang{u, x_i} \ang{y_i, y_j} x_j = \sum_{j=1}^\infty \ang{u, x_j} x_j . 
    \]
    Since $(x_j)$ is a basis of $\Nul(A)\P$, this proves that $BA = P_{\Nul(A)\P}$.

    For the second equation, an analogous computation gives $ABf = \sum_{i=1}^\infty \ang{f, y_i}y_i$, and since $(y_i)$ is a basis of $\clos{\Ran(A)}$, this proves that $AB = P_{\clos{\Ran(A)}}\restr_{\Dom(A\D)}$. 
\end{proof}

\begin{definition}
    Let $A \in \KKK(\XXX, \YYY)$ have SVD $\qty{(\sigma_j, x_j, y_j)}$. 
    We say that $f \in \YYY$ satisfies the \emph{Picard criterion} if 
    \[
    \sum_j \frac{\abs{\ang{f, y_j}}^2}{\sigma_j^2} < \infty. 
    \]
    Note that the expression on the left corresponds to $\norm{A\D f}^2$ if $f \in \Dom(A\D)$. 
\end{definition}

\begin{theorem}
    Let $f \in \clos{\Ran(A)}$. Then $f \in \Ran(A)$ if and only if $f$ satisfies the Picard criterion. 
\end{theorem}

\begin{proof}
    `$\implies$' Write $f = Au$, then
    \[
    \sum_j \frac{\abs{\ang{f, y_j}}^2}{\sigma_j^2} = \sum_j \frac{\abs{\ang{Au, y_j}}^2}{\sigma_j^2} = \sum_j \frac{\abs{\ang{u, A^* y_j}}^2}{\sigma_j^2} = \sum_j \abs{\ang{u, x_j}}^2 < \infty. 
    \]

    `$\impliedby$' Define $u \ceq  \sum_{j=1}^\infty \frac{1}{\sigma_j} \ang{f, y_j} x_j$ (note that by assumption this sum converges). Then 
    \[
    Au = A \sum_{j=1}^\infty \frac{1}{\sigma_j} \ang{f, y_j} x_j = \sum_{j=1}^\infty \frac{1}{\sigma_j} \ang{f, y_j} Ax_j = \sum_{j=1}^\infty \ang{f, y_j} y_j = P_{\clos{\Ran(A)}} f = f,
    \]
    so $Au = f$ which implies $f \in \Ran(A)$. 
\end{proof}

We have seen that the stability of $A\D$ depends on the speed of decay of the singular values $(\sigma_j)$. We formalise this:
\begin{definition}
    Let $A \in \KKK(\XXX, \YYY)$ have singular values $(\sigma_j)$. Then the ill-posed inverse problem $Au = f$ is called \emph{mildly ill-posed} if the $\sigma_j$ decay polynomially (i.e., $\frac{1}{\sigma_n} \leq C n^\gamma$ for some $C, \gamma$) and \emph{severely ill-posed} otherwise. 
\end{definition}

\begin{example}
    Consider the heat equation with initial conditions and boundary values:
    \[
    \begin{cases}
        v_t - v_{xx} = 0 &(x, t) \in (0, \pi) \times \RR_{>0}, \\
        v(0, t) = v(\pi, t) = 0 &t \geq 0, \\
        v(x, 0) = u(x) &x \in (0, \pi), \\
        v(x, T) = f(x) &x \in (0, \pi). 
    \end{cases}
    \]
    Then the forward problem is to determine $f$ given $u$, while the inverse problem is to determine $u$ given $f$. The solution for the foward problem is given by
    \[
    f = Au \ceq \sum_{j=1}^\infty e^{-j^2T} \ang{u, \sin(jx)}  \sin(jx), 
    \]
    and the eigenvalues are therefore $\sigma_j = e^{-j^2T}$. Since these clearly decay exponentially, this problem is severely ill-posed. 
\end{example}

### 2020-10-17 inverse_problems/C2.tex
added current summaries

\section{Classical regularisation theory}
\begin{definition}
    Let $A \in \LLL(\XXX, \YYY)$ and 
\end{definition}

### 2020-10-17 modern_statistical_methods/C1.tex
added current summaries

Classical models rely on so-called ``large $n$ asymptotics'' (where $n$ is the sample size). This course focuses on the scenario where $p$, the number of variables, is larger or about as large as $n$. In this case, the classical theory breaks down, so we need new methods.

\section{Kernel machines}
We represent data are pairs $(Y_i, x_i) \in \RR \times \RR^p$ ($i = 1, \dotsc, n$). The random variables $Y_i$ are called the \emph{responses}, and the (fixed) variables $x_i$ are called \emph{predictors}.

\begin{recap}
    Let $X = (X_1, \dotsc, X_n)\T$ be a multivariate random variable. Its distribution function is given by
    \[
    F_{X} \colon \RR^n \to [0, 1] \colon \vec x \mapsto \PP(X_1 \leq x_1, \dotsc, X_n \leq x_n). 
    \]
    Its expected value is given by
    \[
    \EE[X] \ceq (\EE[X_1], \dotsc, \EE[X_n])\T \in \RR^n. 
    \]
    Its covariance matrix is given by 
    \[
    \Var[X] = \EE[(X - \EE[X])(X - \EE[X])\T] = \EE[XX\T] - \EE[X] \EE[X]\T. 
    \]
    The matrix $\Var[X]$ is symmetric positive semidefinite and satisfies $(\Var[X])_{ij} = \Cov(X_i, X_j)$. 
\end{recap}
\begin{definition}
    In a \emph{linear model}, we assume that 
    \[
    Y_i = x_i\T \beta^0 + \eps_i \quad (i = 1, \dotsc, n). 
    \]
    where $\beta \in \RR^p$ is unknown and the multivariate random variable $\eps = (\eps_1, \dotsc, \eps_n)$ satisfies $\EE (\eps) = 0$ and $\Var(\eps) = \sigma^2 I$. 
\end{definition}

\begin{definition}
    For an estimator $\tilde\beta$ of $\beta^0$, its \emph{mean squared error} (MSE) is given by 
    \[
    \EE_{\beta^0, \sigma^2} \qty[(\tilde\beta - \beta^0)(\tilde\beta - \beta^0)\T]  = \Var(\tilde\beta) + \qty[\EE(\tilde\beta - \beta^0)]\qty[\EE(\tilde\beta - \beta^0)]\T. 
    \]
\end{definition}
Note that if $\tilde\beta$ is unbiased, the second term will disappear and the MSE is simply the variance.

\begin{recap}
    The maximum likelihood estimator (MLE) in this model is the ordinary least squares (OLS) estimator $\hat\beta^\Rm{OLS} = (X\T X)^{-1} X\T Y$, where the \emph{design matrix} $X \in \RR^{n \times p}$ is the matrix whose rows are the vectors $x_i$. This estimator only exists if $X$ has full column rank, so in particular, it is required that $p \leq n$.

    The Cramér-Rao lower bown states that, out of all unbiased estimators, the MLE has the optimal variance \emph{asymptotically} (i.e., for $n \to\infty$). 
\end{recap}

\subsection{Ridge regression}
\begin{definition}
    Let $\lambda \geq 0$, and let $\Bf 1 \in \RR^n$ be the all-ones vector. Then we define the \emph{Ridge regression} estimators
    \[
    (\hat\mu_\lambda^\Rm{R}, \hat\beta_\lambda^\Rm{R}) \ceq \argmin_{(\mu, \beta) \in \RR \times \RR^p} \qty{ \norm{Y - \mu\Bf 1 - X\beta}^2 + \lambda \norm{\beta}^2}, 
    \]
    where the used norm is the 2-norm. The parameter $\lambda$ is called the \emph{regularisation parameter}. 
\end{definition}

The parameter $\lambda$ represents a penalty for large coefficients in the design matrix. The intercept is not penalised --- this is because a shift in units should not affect the fitted values. However, $X\hat\beta$ is not invariant under scale transformations, so it is common practice to centre the columns of $X$ to have mean 0, and then scale them to have $\ell_2$-norm $\sqrt n$.

After that, we can compute $\hat\mu_\lambda^\Rm{R}$ by taking the derivative:  
\begin{align*}
    \norm{Y - \mu\Bf{1} - X\beta}^2 &=  \sum_i (Y_i - \mu - \sum_j X_{ij}\beta_j)^2. \\
    \pdv{\mu} \norm{Y- \mu\Bf{1} - X\beta}^2 &= -2 \sum_i \qty( Y_i - \mu - \sum_j X_{ij}\beta_j).
\end{align*}
Setting this derivative equal to 0 yields
\begin{align*}
    -2\sum_i \qty(Y_i - \mu - \sum_j X_{ij}\beta_j) &= 0 \\
    \sum_i Y_i - n\mu - \sum_j \beta_j (\sum_i X_{ij}) &= 0 \\
    \sum_i Y_i - n\mu &= 0 \\
    \mu &= \frac1n \sum_i Y_i = \bar Y. 
\end{align*}

Therefore we conclude $\hat\mu_\lambda^{\Rm R} = \bar Y$. 
After centering the responses (i.e.\, replacing $Y_i$ by $Y_i - \bar Y$), the problem can be reduced to 
\[
\hat\beta_\lambda^{\Rm R} = \argmin_{\beta \in \RR^p} Q(\beta) \ceq \argmin_{\beta \in \RR^p}  \qty[ \norm{Y- X\beta}^2 + \lambda \norm{\beta}^2]. 
\]

Since $Q(\beta)$ is convex quadratic, there is a unique root, and to find it we compute
\[
\nabla_\beta Q(\beta) = 2 X\T (Y - X \beta) + 2\lambda\beta = 0 \iff \beta = (X\T X + \lambda I)^{-1} X\T Y. 
\]

We conclude that $\hat\beta^\Rm{R}_\lambda = (X\T X + \lambda I)^{-1} X\T Y$. Note that, even if $X$ does not have full column rank, this estimator exists for all $\lambda > 0$.  In fact, for $\lambda$ sufficiently small, the Ridge estimator outperforms the MLE in terms of mean squared error:
\begin{theorem}
    Fix $\beta^0, \sigma^2$, and assume that $\hat\beta^\Rm{OLS}$ exists (i.e., $X$ has full column rank). For some $\lambda > 0$ sufficiently small, it holds that the MSE of $\hat\beta^\Rm{OLS}$ minus the MSE of $\hat\beta^\Rm{R}_\lambda$ is positive definite. 
\end{theorem}

\begin{proof}
    This is simply writing out the MSE's. In the end, we find that the result holds for $0 < \lambda < 2\sigma^2 / \norm{\beta^0}^2$. 
\end{proof}

\subsubsection{The SVD and PCA}
\begin{recap}
    Recall that any $X \in \RR^{n \times p}$ can be factorised as $X = UDV\T$, where $U, V$ are $n \times n$ and $p \times p$ orthogonal matrices respectively, and $D \in \RR^{n \times p}$ satisfies $D_{11} \geq\dotsb\geq D_{mm} \geq 0$ where $m \ceq \min(n, p)$, and all other entries of $D$ are 0. This is called the \emph{singular value decomposition} or SVD of $X$.

    If $n > p$, we can replace $U$ by its first $p$ columns and $D$ by its first $p$ rows to produce the so-called \emph{thin SVD} of $X$. Then $U \in \RR^{n \times p}$ has orthogo nal columns (so $U\T U = I$) and $D \in \RR^{p \times p}$ is square and diagonal. 
\end{recap}

Suppose $n \geq p$ and let $X = UDV\T$ be the thin SVD of our design matrix $X$. Then we can write the fitted values from the Ridge regression as follows: 
\begin{align*}
    X\hat\beta_\lambda^\Rm{R} &= X(X\T X + \lambda I)^{-1} X\T Y \\
    &= UDV\T (VD^2V\T + \lambda I)^{-1} VDU\T Y \\
    &= UDV\T \qty(V(D^2 + \lambda I)V\T)^{-1} VDU\T Y \\
    &= UD (D^2 + \lambda I)^{-1} DU\T Y \\
    &= U D^2 (D^2 + \lambda I)^{-1} U\T Y \\
    &= \sum_{j=1}^p \frac{D_{jj}^2}{D_{jj}^2 + \lambda} U_j U_j\T Y. 
\end{align*}

Note that for OLS ($\lambda = 0$), this is simply the projection of $Y$ onto the column space of $X$ (if $X$ has full column rank). If $\lambda > 0$, $Y$ is still projected onto the column space of $X$, but the projection is shrunk in the directions of the left singular vectors, and the lower the corresponding singular value, the higher the shrinkage.

\paragraph{Principal component analysis} Consider $v \in \RR^p$ with norm 1, then since the columns of $X$ have been centered, the sample mean of $Xv$ is 0, and the sample variance is therefore
\[
\frac1n \sum_i (Xv)_i^2 = \frac1n (Xv)\T Xv = \frac1n v\T X\T X v = \frac1n v\T VD^2V\T v. 
\]
Writing $a = V\T v$ (with $\norm{a} = 1$), we find
\[
\frac1n v\T VD^2 V\T v = \frac1n a\T D^2 a = \frac1n \sum_j a_j^2 D_{jj}^2
\]

Therefore, we see that the above is maximised if $a = \pm e_1$, or equivalently $v = \pm V_1$. Therefore, $V_1$ determines which combination of columns of $X$ has the largest variance (subject to having norm 1), and $XV_1 = D_{11} U_1$ is known as the \emph{first principal component} of $X$. Analogously, it can be shown that $D_{22} U_2, \dotsc, D_{pp} U_p$ have maximum variance $D_{jj}^2 / n$, subject to being orthonormal to all earlier principal components.

We see that Ridge regression shrinks $Y$ most in the smaller principal components of $X$. Therefore it will work well if most of the information is in the larger principal components of $X$.

\paragraph{A comment on computation}
By analogous calculations as before, one cam compute $\hat\beta_\lambda^\Rm{R} = V(D^2 + \lambda I)^{-1} DU\T Y$. Since calculating the inverse of a diagonal matrix is trivial, we see that the complexity of computing $\hat\beta_\lambda^\Rm{R}$ for any $\lambda$ lies in $O(np)$. Of course, this is after computation of the SVD of $X$, which lies in $O(np \min(n, p))$.

\subsection{$v$-fold cross validation}
Of course, we are still left with the problem of choosing $\lambda$ in ridge regression. We consider one possible way of doing so, namely $v$-fold cross validation, which is a general way of selection a good regression method from several competing methods. Here, we assume that our predictors are random, so that we have i.i.d.\ data pairs $(x_i, Y_i)$ ($i = 1, \dotsc, n$). Suppose $(x^*, Y^*)$ is a new data pair, independent of $(X, Y)$ and identically distributed. Ideally, we want to pick $\lambda$ which minimises the prediction error (averaged over $Y^*$ and $x^*$)
\[
\EE\qty[ \qty(Y^* - (x^*)\T \hat\beta_\lambda^\Rm{R}(X, Y))^2 \mid X, Y],
\]
where the dependence of $\hat\beta_\lambda^\Rm{R}$ on the training data $(X, Y)$ is made explicit by denoting it $\hat\beta_\lambda^\Rm{R}(X, Y)$.

This is impossible to minimise, but it may be possible to minimise the expected prediction error (averaged over the training data)
\begin{equation} \label{eq:averaged_prediction}
\EE\qty{ \EE\qty[ \qty(Y^* - (x^*)\T \hat\beta_\lambda^\Rm{R}(X, Y))^2 \mid X, Y]}. 
\end{equation}

This is still not possible to compute directly, but we estimate it using $v$-fold cross validation. Split the data into $v$ groups or \emph{folds} of roughly equal size $(X^{(1)}, Y^{(1)}), \dotsc, (X^{(v)}, Y^{(v)})$ and let $(X^{(-k)}, Y^{(-k)})$ denote all data except that in the $k$-th fold. Then we define 
\[
\CV(\lambda) \ceq \frac1n \sum_{i=1}^n \qty[Y_i - x_i\T\hat\beta_\lambda^\Rm{R}(X^{(-\kappa(i))}, Y^{(-\kappa(i))})]^2, 
\]
and choose the value of $\lambda$ that minimises $\CV(\lambda)$.

The function $\CV(\lambda)$ is called the \emph{out-of-sample error}, since the training data does not include $x_i$.

\begin{recap}
    The \emph{tower rule} states that for random variables $X, Y$ we have $\EE[X] = \EE[\EE[X \mid Y]]$. 
\end{recap}

Note that for each $i$, we have
\[
\EE\qty[\qty{Y_i - x_i\T\hat\beta_\lambda^\Rm{R}(X^{(-\kappa(i))}, Y^{(-\kappa(i))})}^2] = \EE\qty[ \EE\qty[\qty{Y_i - x_i\T\hat\beta_\lambda^\Rm{R}(X^{(-\kappa(i))}, Y^{(-\kappa(i))})}^2 \mid X^{-\kappa(i)}, Y^{-\kappa(i)} ]]. 
\]
This equals the expected prediction error in \cref{eq:averaged_prediction}, except that the training data $X, Y$ are replaced with a smaller data set.

We now have a bias-variance tradeoff in the size of the folds: if $v = n$ (known as ``leave-one-out'' cross-validation), the estimation will be almost unbiased, but the averaged quantities in $\CV(\lambda)$ will be highly correlated which leads to high variance. Typical choices of $v$ are 5 or 10.

Instead of finding the single best $\lambda$, we can also aim to find the best weighted combination of $\lambda$'s. For example, suppose $\lambda$ is restricted to a grid $\lambda_1 > \dotsb > \lambda_L$. Then we can use any nonnegative least-squares optimization algorithm to minimise 
\[
\frac1n \sum_{i=1}^n \qty[ Y_i - \sum_{\ell = 1}^L w_\ell x_i\T \hat\beta_{\lambda_\ell}^\Rm{R}(X^{(-\kappa(i))}, Y^{(-\kappa(i))})]^2,
\]
over all $w \in \RR^L_{\geq 0}$. This procedure is known as \emph{stacking} and often outperforms cross-validation.

\subsection{The kernel trick}
We note that 
\[
 X\T (XX\T + \lambda I) = (X\T X + \lambda I) X\T,
\]
and multiplying from the left with $(X\T X + \lambda I)^{-1}$ and from the right with $(XX\T + \lambda I)^{-1}$ gives
\[
(X\T X + \lambda I)^{-1} X\T = X\T (XX\T + \lambda I)^{-1}.
\]
Using this, we see that we can rewrite the fitted values from ridge regression as follows:
\[
X\hat\beta_\lambda^R = X(X\T X + \lambda I)^{-1} X\T Y = XX\T (XX\T + \lambda I)^{-1} Y.
\]
Two important remarks: 
\begin{enumerate}
    \item Computing the LHS of this equation takes roughly $O(np^2 + p^3)$ operations, while computing the RHS takes $O(n^2 p + n^3)$ operations (this is because in the LHS we invert an $p \times p$ matrix, while in the RHS we invert a $n \times n$ matrix). Therefore, if $p \gg n$, the RHS can be much cheaper to compute. 
    \item The LHS depends only on the matrix $K = XX\T$ (this matrix is called the \emph{kernel matrix}). Intuitively, since $K_{ij} = \ang{x_i, x_j}$, the entries of the kernel matrix show how `similar' the corresponding predictors are. 
\end{enumerate}

\begin{example}
    Suppose we have data $(Y_i, z_i)_{i = 1, \dotsc, n}$ with $z_i = (z_{i1}, \dotsc, z_{id})\T$, and we believe the following quadratic relation holds:
    \[
    Y_i = \sum_k \sqrt{2}\gamma_k z_{ik}  + \sum_{k, \ell} \theta_{k\ell} z_{ik} z_{i\ell} + \eps_i.
    \]
    To compute fitted values using ridge regression, we can rewrite this as a linear model $Y = X\beta + \eps$ where
    \[
    \beta = \mqty(\gamma_1 \\ \vdots \\ \gamma_d \\ \theta_{11} \\ \theta_{12} \\ \vdots \\ \theta_{dd}), \quad x_i = \mqty(\sqrt{2} z_{i1} \\ \vdots \\ \sqrt 2 z_{id} \\ z_{i1} z_{i1} \\ z_{i1}z_{i2} \\ \vdots \\ z_{id}z_{id}).
    \]

    In this case, we have $p = d^2 + d$ variables, which means computing $(X\T X + \lambda I)^{-1}$ takes $O(d^6)$ operations. In this case, computing $(XX\T + \lambda I)^{-1}$ is probably easier.

    We are still left with the problem of computing $K \ceq XX\T$, which can take $O(n^2 p) = O(n^2 d^2)$ operations if done naively. However, observe that
    \[
    K_{ij} = x_i\T x_j = 2 \sum_k z_{ik} z_{jk} + \sum_{k, \ell} z_{ik} z_{i\ell} z_{jk} z_{j\ell} = \qty( 1 + \sum_k z_{ik} z_{jk})^2 - 1 = (1 + z_i\T z_j) - 1.
    \]
    This quantity can be computed in $O(d)$, and therefore $K$ can be computed in $O(n^2 d)$ operations: we have a factor $d$ improvement. 
\end{example}

The general point of the previous example is that we can bypass the features $x_i$ entirely and instead think directly of $K = XX\T$ where an entry $K_{ij}$ represents similarity between the inputs of the $i$-th and $j$-th samples. This leads to the notion of a kernel in general.

\subsection{Kernels}
We will assume our inputs $x_1, \dotsc, x_n$ live in an abstract space $\XXX$. 
\begin{definition}
    A (\emph{positive-definite}) \emph{kernel} is a symmetric map $k \colon \XXX^2 \to \RR$ such that for all $n \in \NN$ and all $x_1, \dotsc, x_n \in \XXX$, the matrix $K \in \RR^{n \times n}$ with $K_{ij} = k(x_i, x_j)$ is positive semi-definite. 
\end{definition}

\begin{proposition}[Cauchy-Schwarz for kernels] \label{prop:kernel_cs}
    Let $k$ be a kernel and $x, x' \in \XXX$, then 
    \[
    k(x, x')^2 \leq k(x, x) k(x', x'). 
    \]
\end{proposition}

\begin{proof}
    The matrix $\smqty(k(x, x) & k(x, x') \\ k(x', x) & k(x', x'))$ must be positive semi-definite so its determinant must be non\-negative. 
\end{proof}

In our old models, the data points $x_i$ were vectors in $\RR^p$. Now we try to think of them as points in an abstract space with an associated \emph{feature map} $\phi \colon \XXX \to \HHH$ (with $\HHH$ an inner product space), and a kernel $k(x, x')$ gives a measure of similarity between $\phi(x)$ and $\phi(x')$. In this case, we have the following: 
\begin{proposition} \label{prop:ip_kernel}
    Let $\HHH$ be an inner product space,  $\phi \colon \XXX \to \HHH$ and define $k(x, x') \ceq \ang{\phi(x), \phi(x')}$. Then $k$ is a kernel. 
\end{proposition}

\begin{proof}
    We have, for all $x_1, \dotsc, x_n \in \XXX$ and $\alpha \in \RR^n$ that
    \[
    \alpha\T K \alpha = \sum_{i, j} K_{ij} \alpha_i \alpha_j = \sum_{i, j} \ang{\phi(x_i), \phi(x_j)} \alpha_i \alpha_j = \norm{\sum_i \alpha_i \phi(x_i)}^2 \geq 0. 
    \]
\end{proof}

The following proposition shows how to make new kernels from old:
\begin{proposition}
    Suppose $k_1, k_2, \dotsc$ are kernels. Then:
    \begin{enumerate}
        \item If $\alpha_1, \alpha_2 \geq 0$ then $\alpha_1 k_1 + \alpha_2 k_2$ is a kernel.
        \item The pointwise limit of a sequence of kernels is a kernel (if it exists). 
        \item The pointwise product $k_1 k_2$ is a kernel. 
    \end{enumerate}
\end{proposition}

\begin{proof}
    See Example Sheet 1. 
\end{proof}

\begin{example} Let us consider some examples of kernels:
    \begin{enumerate}
        \item For $\XXX = \RR^p$ we have already seen the \emph{linear kernel} $k(x, x') = x\T x'$. 
        \item For $\XXX = \RR^p$, the \emph{polynomial kernel} is defined as $k(x, x') = (1 + x\T x')^d$. This is a kernel since it is a power of a sum of two kernels. 
        \item For $\XXX = \RR^p$, the \emph{Gaussian kernel} is defined by
        \[
        k(x, x') = \exp\qty( - \frac{\norm{x - x'}_2^2}{2\sigma^2}).
        \]
        To show this is a kernel, write $k$ as the pointwise product $k_1k_2$ where
        \[
        k_1(x, x') = \exp\qty(-\frac{\norm{x}^2}{2\sigma^2}) \exp\qty(-\frac{\norm{x'}^2}{2\sigma^2}), \quad k_2(x, x') = \exp\qty(\frac{x\T x'}{\sigma^2}). 
        \]
        Clearly $k_1$ is the kernel induced by the feature map $\phi(x) = \exp(-\norm{x}^2/(2\sigma^2))$, while $k_2$ can be seen to be a kernel by using the Taylor expansion, which shows that $k_2$ is a limit of nonnegative linear combinations of kernels.

        \item For $\XXX = [0, 1]$, define the \emph{Sobolev kernel} $k(x, x') = \min(x, x')$. The proof that this is a kernel is on example sheet 1. 
        \item For $\XXX = \pow(\qty{1, \dotsc, p})$, define the \emph{Jaccard kernel}
        \[
        k(x, x') = \frac{\abs{x \cap x'}}{\abs{x \cup x'}} \quad\text{where $0/0 \ceq 1$}. 
        \]
        The proof that this is a kernel is on example sheet 1. 
    \end{enumerate}
\end{example}

By \cref{prop:ip_kernel}, we see that every feature map $\phi \colon \XXX \to \HHH$ gives rise to a kernel. In the next (important!) theorem, we will see that every kernel is in fact induced by a feature map.

\begin{theorem} \label{thm:kernel_ip}
    Let $k$ be a kernel, then there exists an inner product space $\HHH$ and a feature map $\phi \colon \XXX \to \HHH$ such that 
    \[
    k(x, x') = \ang{\phi(x), \phi(x')} \quad\text{for all $x, x' \in \XXX$}. 
    \]
\end{theorem}

\begin{proof}
    We will construct $\HHH$ and $\phi$ explicitly. First we define the function space
    \[
    \HHH = \qty{ \sum_{i=1}^n \alpha_i k(\cdot, x_i) \mid n \in \NN,  \alpha_i \in \RR, x_i \in \XXX}. 
    \]
    Let $f = \sum_{i=1}^n \alpha_i k(\cdot, x_i)$ and $g = \sum_{j=1}^n \beta_j k(\cdot, x_j')$, then the inner product on $\HHH$ is given by
    \[
    \ang{f, g} = \ang*{\sum_{i=1}^n \alpha_i k(\cdot, x_i), \sum_{j=1}^m \beta_j k(\cdot, x_j')} \ceq \sum_{i=1}^n \sum_{j=1}^m \alpha_i \beta_j k(x_i, x_j'). 
    \]
    We define $\phi \colon \XXX \to \HHH$ as $\phi(x) = k(\cdot, x)$.

    We must check that the inner product does not depend on the choice of representation of $f$ and $g$. For this, note that 
    \[
    \sum_{i=1}^n \sum_{j=1}^m \alpha_i\beta_j k(x_i, x_j') = \sum_{i=1}^n \alpha_i g(x_i) = \sum_{j=1}^m \beta_j f(x_j'),
    \]
    which holds by symmetry of the kernel. Since $\sum_i \alpha_i g(x_i)$ is independent of the representation of $g$, while $\sum_j \beta_j f(x_j')$ is independent of the representation of $f$, we conclude that the entire expression is independent of both representations.

    Secondly, we must verify that the formula $k(x, x') = \ang{\phi(x), \phi(x')}$ indeed holds. For any $f \in \HHH, x \in \XXX$ we have
    \begin{equation}
    \ang{k(\cdot, x), f} = \sum_{i=1}^n \alpha_i k(x_i, x) = f(x),
    \end{equation}
    i.e., evaluation of a function is a linear functional in $\HHH$.

    In particular, we have 
    \[
    \ang{\phi(x), \phi(x')} = \ang{k(\cdot, x), k(\cdot, x')} =  k(x, x'). 
    \]

    Finally, we must check that $\ang{\cdot, \cdot}$ is indeed an inner product. Symmetry and bilinearity are clear. Furthermore, we have
    \[
   \ang{f, f} = \sum_{i=1}^n \sum_{j=1}^n \alpha_i \alpha_j k(x_i, x_j) = \alpha\T K\alpha \geq 0
    \]
    by the fact that $k$ is a kernel. We must now only show that $f \neq 0 \implies \ang{f, f} > 0$. For this, note that $\ang{\cdot, \cdot}$ is a kernel on $\HHH$, so by \cref{prop:kernel_cs} (Cauchy-Schwarz) we have
    \[
    f(x)^2 = \ang{k(\cdot, x), f}^2 \leq \ang{k(\cdot, x), k(\cdot, x)} \ang{f, f}, 
    \]
    and therefore if $f$ is nonzero anywhere, $\ang{f, f}$ must also be nonzero. 
\end{proof}

While $\HHH$ constructed in the proof is an inner product space, it is not necessarily a Hilbert space. Let $(f_n) \subseteq \HHH$ be Cauchy, then by Cauchy-Schwarz for kernels we find
\[
f_m(x) - f_n(x) = (f_m - f_n)(x) = \ang{k(\cdot, x), f_m - f_n} \leq \sqrt{k(x, x)} \norm{f_n - f_m}. 
\]
We can do an analogous computation for $f_n - f_m$ to conclude that $\abs{f_m(x) - f_n(x)} \leq \sqrt{k(x, x)} \norm{f_n - f_m}$, and therefore, if $(f_n)$ is Cauchy, then it converges pointwise to some $f^* \colon \XXX \to \RR$. We will not prove the following theorem:
\begin{theorem}
    The inner product space $\HHH$ constructed in the proof of \cref{thm:kernel_ip} can be extended to a Hilbert space by adding all pointwise limits $f^*$ of Cauchy sequences $(f_n) \subseteq \HHH$. 
\end{theorem}

The completion of $\HHH$ is a special type of Hilbert space:
\begin{definition}
    A Hilbert space $\BBB$ of functions $f \colon \XXX \to \RR$ is called a \emph{reproducing kernel Hilbert space} (RKHS) if for all $x \in \XXX$, there exists $k_x \in \BBB$ such that 
    \[
    f(x) = \ang{k_x, f}, 
    \]
    i.e., evaluation of functions is a linear functional.

    The function $k(x, x') = \ang{k_x, k_{x'}}$ is known as the \emph{reproducing kernel} of $\BBB$ (induced by the feature map $\phi(x) = k_x$). 
\end{definition}

If we start with a kernel $k$, construct the corresponding RKHS $\BBB$, then it is easily checked that $k$ is indeed the reproducing kernel of $\BBB$.

\begin{example}[Linear kernel]
    Let $X = \RR^p$ and $k(x, x') = x\T x'$. Then we have 
    \[
    \HHH = \qty{x \mapsto \sum_{i=1}^n \alpha_i x\T x_i = x\T \qty(\sum_i \alpha_i x_i) \mid \alpha_i \in \RR, x_i \in \RR^p} = \qty{x \mapsto x\T \beta \mid \beta \in \RR^p},
    \]
    and if $f(x) = x\T\beta$, $g(x) = x\T \beta'$, then 
    \[
    \ang{f, g} = k(\beta, \beta') = \beta\T \beta' \quad\text{so } \norm{f}_\HHH = \norm{\beta}_2. 
    \]
\end{example}

### 2020-10-17 topics_statistical_theory/C1.tex
added current summaries

\section{Basic concepts}
\subsection{Parametric vs nonparametric models}
\begin{definition}
    A \emph{statistical model} is a family of possible data-generating mechanisms. If the parameter space $\Theta$ is finite-dimensional, we speak of a \emph{parametric model}.

    A model is called \emph{well-specified} if there is a $\theta_0 \in \Theta$ for which the data was generated from the distribution with parameter $\theta_0$, and otherwise it is called \emph{misspecified}. 
\end{definition}

\begin{recap}
    Let $(Y_n)$ be a sequence of random vectors and $Y$ a random vector. 
    \begin{enumerate}
        \item We say that $(Y_n)$ \emph{converges almost surely} to $Y$, notation $Y_n \toas Y$, if $\PP(Y_n \to Y) = 1$. 
        \item We say that $(Y_n)$ \emph{converges in probability} to $Y$, notation $Y_n \toprob Y$, if for every $\eps > 0$ we have $\PP(\norm{Y_n - Y} > \eps) \to 0$.
        \item We say that  $(Y_n)$ \emph{converges in distribution} to $Y$, notation $Y_n \todist Y$, if $\PP(Y_n \leq y) \to \PP(Y \leq y)$ for all $y$ where the distribution function of $Y$ is continuous.

        This is equivalent to the condition that $\EE[f(Y_n)] \to \EE[f(Y)]$ for all bounded Lipschitz functions $f$. 
    \end{enumerate}
    It is known that $Y_n \toas Y \implies Y_n \toprob Y \implies Y_n \todist Y$.

    If $(Y_n)$ is a sequence of random vectors and $(a_n)$ is a positive sequence, then we write $Y_n = O_p(a_n)$ if, for all $\eps > 0$, there exists $C > 0$ such that for sufficiently large $n$ we have
    \[
    \PP\qty(\frac{\norm{Y_n}}{a_n} > C) < \eps. 
    \]
    We write $Y_n = o_n(a_n)$ if $Y_n/a_n \toprob 0$. 
\end{recap}
In a well-specified parametric model, the maximum likelihood estimator (MLE) $\hat\theta_n$ typically satisfies $\hat\theta_n - \theta_0 \in O_p(n^{-1/2})$. On the other hand, if the model is misspecified, any inference can give very misleading results. To circumvent this problem, we consider \emph{nonparametric models}, which make much weaker assumptions. Such infinite-dimensional models are much less vulnerable to model misspecification, however we will typically pay a price in terms of a slower convergence rate than in well-specified parametric models.

\begin{example}
    Examples of nonparametric models include: 
    \begin{enumerate}
        \item Assume $X_1, \dotsc, X_n \iid F$ for some unknown distribution function $F$. 
        \item Assume $X_1, \dotsc, X_n \iid f$ for some unknown density $f$ belonging to a smoothness class.
        \item Assume $Y_i = m(x_i) + \eps_i$ ($i = 1, \dotsc, n$), where the $x_i$ are known, $m$ is unknown and belongs to some smoothness class, and the $\eps_i$ are i.i.d.\ with $\EE (\eps_i) = 0$ and $\Var(\eps_i) = \sigma^2$. 
    \end{enumerate}
\end{example}

\subsection{Estimating an arbitrary distribution function}
\begin{definition}
    Let $\Cal F$ denote the class of all distribution functions on $\RR$ and suppose $X_1, \dotsc, X_n \iid F \in \Cal F$. The \emph{empirical distribution function} $\hat{F_n}$ of $X_1, \dotsc, X_n$ is defined as
    \[
    \hat F_n(x) \ceq \frac1n \sum_{i=1}^n \ind_\qty{X_i \leq x}.
    \]
\end{definition}

\begin{recap}
    The \emph{strong law of large numbers} tells us that if $(Y_n)$ are i.i.d.\ with finite mean $\mu$, then $\bar{Y} \ceq \frac1n \sum_{i=1}^n Y_i \toas \mu$. 
\end{recap}

Note that the strong law of large numbers immediately implies that $\hat{F_n}(x)$ converges almost surely to $F(x)$ as $n \to \infty$. However, the following stronger result states that this convergence holds uniformly in $x$:

\begin{theorem}[Glivenko-Cantelli] \label{thm:GC}
    Let $X_1, X_2, \dotsc \iid F$. Then we have 
    \[
    \sup_{x \in \RR} \abs{\hat{F_n}(x) - F(x)}\toas 0. 
    \]
\end{theorem}

\begin{proof}
    See lecture notes. The main idea of the proof is to ``control'' $\hat{F_n}$ in a finite number of points $x_1, \dotsc, x_k$, and then deduce what happens between those points using the fact that distributions are increasing and right-continuous. On \href{https://en.wikipedia.org/wiki/Glivenko%E2%80%93Cantelli_theorem#Proof}{Wikipedia}, 
        a simplified proof can be found assuming that $F$ is continuous, which still encapsulates the main idea.  
\end{proof}

\begin{theorem}[Dvoretzky-Kiefer-Wolfowitz] 
    Under the conditions of \cref{thm:GC}, for every $\eps > 0$ it holds that
    \[
    \PP_F \qty(\sup_{x \in \RR} \abs{\hat{F_n}(x) - F(x)} > \eps) \leq 2e^{-2n\eps^2},
    \]
    and this is a tight bound. 
\end{theorem}

We will not prover this theorem, however, we will explore a few consequences. 
One of these consequences is the following:

\begin{corollary}[Uniform Glivenko-Cantelli theorem] Under the conditions of \cref{thm:GC}, for every $\eps > 0$, it holds that
    \[
    \sup_{F \in \Cal F} \PP_F \qty(\sup_{m \geq n} \sup_{x \in \RR} \abs{\hat{F_m}(x) - F(x)} > \eps) \to 0 \quad\text{as $n \to\infty$.}
    \]
\end{corollary}

\begin{proof}
    By a union bound, the DKW inequality, and convergence of the geometric series we have
    \begin{align*}
    \sup_{F \in \Cal F} \PP_F \qty(\sup_{m \geq n} \sup_{x \in \RR} \abs{\hat{F_m}(x) - F(x)} > \eps) &\leq \sup_{F \in \Cal F} \sum_{m = n} \PP_F \qty(\sup_{x \in \RR} \abs{\hat{F_n}(x) - F(x)} > \eps) \\
    &\leq 2 \sum_{m=n}^\infty e^{-2m\eps^2},    \end{align*}
which converges to 0 as it is the tail of a converging sum. 
\end{proof}

For another consequence, we consider the problem of finding a confidence band for $F$. Given $\alpha \in (0, 1)$, set $\eps_n \ceq \sqrt{-\frac1{2n}\log(\alpha/2)}$. Then the DKW inequality tells us that
\[
\PP_F \qty( \sup_{x \in \RR} \abs{\hat{F_n}(x) - F(x)} > \eps_n) \leq \alpha,
\]
or equivalently, that 
\[
\PP_F\qty(\hat{F_n}(x) - \eps_n \leq F(x) \leq \hat{F_n}(x) + \eps_n \ \text{for all $x \in \RR$}) \geq 1 - \alpha. 
\]

We can say even more.

\begin{recap}
    For any distribution function $F$, its \emph{quantile function} is defined as
    \[
    F^{-1} \colon (0, 1] \to \RR \cup \qty{\infty} \colon p \mapsto \inf \qty{x \in \RR \mid F(x) \geq p}. 
    \]
    When necessary, we also define $F^{-1}(0) \ceq \sup\qty{x \in \RR \mid F(x) = 0}$.

    If $U \sim U(0, 1)$ and $X \sim F$, then for any $x \in \RR$ we have
    \[
    \PP(F^{-1}(U) \leq x) = \PP(U \leq F(x)) = F(x) = \PP(X \leq x). 
    \]
    This can be written simply as $F^{-1}(U) \deq X$. 
\end{recap}

Let $U_1, \dotsc, U_n \iid U(0, 1)$ with empirical distribution function $\hat{G_n}$, and let $X_1, \dotsc, X_n \iid F$. Then, we have
\begin{align*}
\hat{G_n}(F(x)) = \frac1n \sum_{i=1}^n \ind_\qty{U_i \leq F(x)} \deq \frac1n \sum_{i=1}^n \ind_\qty{X \leq x} = \hat{F_n}(x), 
\end{align*}
where $\deq$ means equality in distribution. It follows that
\[
\sup_{x \in \RR} \abs{\hat{F_n}(x) - F(x)} \deq \sup_{x \in \RR} \abs{\hat{G_n}(F(x)) - F(x)} \leq \sup_{t \in [0, 1]} \abs{\hat{G_n}(t) - t},
\]
with equality if $F$ is continuous. We conclude that if $F$ is continuous, the distribution of $\sup_{x \in \RR} \abs{\hat{F_n}(x) - F(x)}$ does not depend on $F$.

Other generalisations of \cref{thm:GC} include Uniform Laws of Large Numbers. Let $X, X_1, \dotsc, X_n$ be i.i.d.\ on a measurable space $(\Cal X, \Cal A)$, and $\Cal G$ a class of measurable functions on $\Cal X$. We say that $\Cal G$ satisfies a ULLN if 
\[
\sup_{g \in \Cal G} \abs{\frac1n \sum_{i=1}^n g(X_i) - \EE[g(X)]} \toas 0. 
\]
In \cref{thm:GC}, we showed that $\Cal  G = \qty{ \ind_\qty{\cdot \leq x} \mid x \in \RR}$ satisfies a ULLN.

\begin{recap}
    We recall the central limit theorem: if $X_1, \dotsc, X_n$ are i.i.d.\ random variables with mean $\mu$ and variance $\sigma^2 < \infty$, then $\sqrt n \qty(\bar{X_n} - \mu) \todist N(0, \sigma^2)$.

    Dividing by $\sigma$ yields
    \[
    \frac{\sqrt n \qty(\bar{X_n} - \mu)}{\sigma} \todist N(0, 1), 
    \]
    and multiplying both sides by $n$ and writing $V_i = \sum_{j=1}^i X_j$ we obtain
    \[
    \frac{V_i - \EE V_i}{\sqrt{\Var(V_i)}} \todist N(0, 1).
    \]
\end{recap}
Another extension starts with the observation that $\sqrt n \qty(\hat{F_n}(x) - F(x)) \todist N(0, \sigma^2)$, where
\[
\sigma^2 = \Var(\ind_\qty{X \leq x}) = \EE[\ind_{X \leq x}^2] - \EE[\ind_{X \leq x}]^2 = F(x) - F(x)^2 = F(x)(1 - F(x)). 
\]
This can be strengthened by considering $(\sqrt n \qty(\hat{F_n}(x) - F(x)) \todist N(0, \sigma^2) \mid x \in \RR)$ as a stochastic process.

\subsection{Order statistics and quantiles}
\begin{definition}
    Let $X_1, \dotsc, X_n \iid F \in \Cal F$. The \emph{order statistics} are the ordered samples $X_{(1)} \leq \dotsb \leq X_{(n)}$ (where the original order is preserved in case of a tie). 
\end{definition}

The order statistics of the uniform distribution can be computed explicitly: 
\begin{proposition}
    Let $U_1, \dotsc, U_n \iid U(0, 1)$,  let $Y_1, \dotsc, Y_{n+1} \iid \Exp(1)$, and write $S_j \ceq \sum_{i=1}^j Y_j$ ($j = 1, \dotsc, n+1$). Then
    \[
    U_{(j)} \deq \frac{S_j}{S_{n+1}} \sim \Rm{Beta}(j, n - j + 1) \quad\text{for $j = 1, \dotsc, n$}. 
    \]
\end{proposition}

\begin{proof}
    See example sheet 1, question 1.  
\end{proof}
\begin{definition}
    Let $X_1, \dotsc, X_n \iid F$. Then the \emph{sample quantile function} is defined as 
    \[
    \hat F_n^{-1}(p) = \inf \qty{x \in \RR \mid \hat F_n(x) \geq p}. 
    \]
\end{definition}

\begin{proposition}
    It holds that $\hat F_n^{-1}(p) = X_{(\ceil{np})}$. 
\end{proposition}

\begin{proof}
    By definition, $\hat F_n^{-1}(p)$ is the smallest value of $x$ for which $\hat F(x)$ is larger than $p$. Note that
    \[
    \hat F(x) \geq p \iff \frac1n \sum_{i=1}^n \ind_\qty{X_i \leq x} \geq p \iff \sum_{i=1}^n \ind_\qty{X_i \leq x} \geq np \iff \sum_{i=1}^n \ind_\qty{X_i \leq x} \geq \ceil{np}.
    \]
    The smallest value of $x$ for which this occurs is the smallest value of $x$ such that exactly $\ceil{np}$ of the variables $X_1, \dotsc, X_n$ satisfy $X_i \leq x$. We conclude that $\hat F_n^{-1}(p) = X_{(\ceil{np})}$. 
\end{proof}

For $p = \frac12$ for example, this proposition tells us that $\hat F_n^{-1}(p) = X_{(\ceil{n/2})}$, the median of the data. We now explore the distribution of $X_{(\ceil{np})}$.

\begin{recap}
We recall two theorems. The first is \emph{Slutsky's theorem}:
\begin{theorem}
    Let $(Y_n)$ and $(Z_n)$ be sequences of random vectors with $Y_n \todist Y$ and $Z_n \toprob c$ for some constant $c$. If $g$ is a continuous real-valued function, then $g(Y_n, Z_n) \todist g(Y, c)$. 
\end{theorem}

The second is the \emph{delta method}:
\begin{theorem}
    Let $(Y_n)$ be a sequence of random vectors such that $\sqrt n (Y_n - \mu) \todist Z$. If $g \colon \RR^d \to \RR$ is differentiable at $\mu$, then
    \[
    \sqrt n \qty(g(Y_n) - g(\mu)) \todist g'(\mu) Z. 
    \]
\end{theorem}
\end{recap}

\begin{lemma}
    If $U_1, \dotsc, U_n \iid U(0, 1)$ and $p \in (0, 1)$, then $\sqrt n \qty(U_{\ceil{np}} - p) \todist N(0, p(1-p))$. 
\end{lemma}

\begin{proof}
    Let $Y_1, \dotsc, Y_{n+1} \iid \Exp(1)$, $V_n \ceq \sum_{i=1}^{\ceil{np}} Y_i$ and $W_n \ceq \sum_{i = \ceil{np} + 1}^{n+1} Y_i$. Then $V_n$ and $W_n$ are independent, and we have seen that $U_{\ceil{np}} \sim \frac{V_n}{V_n + W_n}$.

    Noting that $\EE V_n = \Var(V_n) = \ceil{np}$ we find
    \begin{align*}
    \sqrt{n} \qty(\frac{V_n}{n} - p) &= \frac{\sqrt{\ceil{np}}}{\sqrt n} \qty(\frac{V_n - \ceil{np}}{\sqrt{\ceil{np}}}) + \frac{\ceil{np} - np}{\sqrt n} \\
    &= \frac{\sqrt{\ceil{np}}}{\sqrt n}  \qty(\frac{V_n - \EE V_n}{\sqrt{\Var(V_n)}}) +  \frac{\ceil{np} - np}{\sqrt n}.
    \end{align*}
    Now, by the central limit theorem, the term between brackets converges to a standard $N(0, 1)$ distribution. The term $\sqrt{\ceil{np}}{\sqrt n}$ converges to $\sqrt p$ and the term $(\ceil{np} - np)/\sqrt n$ converges to 0, so by Slutsky's lemma, we find
    \[
    \sqrt{n}\qty(\frac{V_n}{n} - p) \todist \sqrt p N(0, 1) = N(0, p). 
    \]

    An analogous calculation shows that $\sqrt n \qty(\frac{W_n}{n} - (1-p)) \to N(0, 1 - p)$.

    Now we define $g \colon (0, \infty)^2 \to (0, \infty)$ by $g(x, y) \ceq x / (x + y)$, which is differentiable at $(p, 1 - p)$. Note that the distribution of $(V_n, W_n)$ is an $N(0, \smqty(p & 0 \\ 0 & q))$ distribution. By the delta method we find
    \begin{align*}
        \sqrt n \qty(U_{\ceil{np}} - p) &\deq \sqrt n\qty( g\qty(\frac{V_n}{n}, \frac{W_n}{n}) - g(p, q)) \\
        &\todist g'(p, 1-p) N\qty(0, \mqty(p & 0 \\ 0 & q)) \\
        &= N\qty(0, g'(p, 1-p) \mqty(p & 0 \\ 0 & q) g'(p, 1-p)\T) \\
        &= N(0, p(1-p)).                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        
    \end{align*}
\end{proof}

We now relate what we know about the uniform distribution to the  quantile function: 
\begin{theorem}
    Let $p \in (0, 1)$ and let $X_1, \dotsc, X_n \iid F$. Suppose that $F$ is differentiable at $\xi_p \ceq F^{-1}(p)$ with derivative $f(\xi_p)$. Then
    \[
    \sqrt n \qty(X_{(\ceil{np})} - \xi_p) \todist N\qty(0, \frac{p(1-p)}{f(\xi_p)^2}).
    \]
\end{theorem}

\begin{proof}
    Let $U_1, \dotsc, U_n \iid U(0, 1)$, then we know that $F^{-1}(U_i) \deq X_i$ and thus $F^{-1} (U_{(\ceil{np})}) \deq X_{(\ceil{np})}$. Applying the delta method with $g = F^{-1}$, together with the previous theorem yields
    \[
    \sqrt n \qty(X_{(\ceil{np})} - \xi_p) = \sqrt n \qty(F^{-1}(U_{(\ceil{np})}) - F^{-1}(p)) \todist (F^{-1})'(p) \cdot  N(0, p(1-p)).
    \]
    Noting that $(F^{-1})'(p) = \frac1{f(\xi_p)}$ yields the result. 
\end{proof}

\subsection{Concentration inequalities}
We turn our attention to concentration inequalities, with a focus on finite-sample results (instead of results that only hold for $n \to\infty$). 
\begin{definition}
    A random variable $X$ with mean 0 is called \emph{sub-Gaussian} with parameter $\sigma^2$ if
    \[
    M_X(t) = \EE(e^{tX}) \leq e^{t^2\sigma^2/2}
    \]
    for every $t \in \RR$. 
\end{definition}
Note that equality holds when $X \sim N(0, \sigma^2)$, since the MGF of an $N(\mu, \sigma^2)$ distribution is given by $t \mapsto \exp(\mu t + \sigma^2 t^2 /2)$.

\begin{proposition}We consider some characterisations of sub-Gaussianity: 
    \begin{enumerate}[(a)]
        \item     Let $X$ be sub-Gaussian with parameter $\sigma^2$. Then 
        \begin{equation} \label{eq:sg_tail_bounds}
        \max\qty{\PP(X \geq x), \PP(X \leq -x)} \leq e^{-x^2/(2\sigma^2)} \quad\text{for every $x \geq 0$}.
        \end{equation}

        \item Let $X$ be a random variable which satisfies $\EE(X) = 0$ and \cref{eq:sg_tail_bounds}. Then for every $q \in \NN$ it holds that
        \[
        \EE(X^{2q}) \leq 2 \cdot q! (2\sigma^2)^q \leq q! (2\sigma)^{2q}.
        \]

        \item If $X$ is a random variable with $\EE(X) = 0$ and $\EE(X^{2q}) \leq q! C^{2q}$ for all $q \in \NN$, then $X$ is sub-Gaussian with parameter $4C^2$. 
    \end{enumerate}

\end{proposition}

\begin{recap}
    Recall the \emph{tail bound formula} for the expectation: if $X$ is a nonnegative random variable, then 
    \[
    \EE[X] = \int_0^\infty \PP(X > x) \dd{x}. 
    \]

    Furthermore, recall that the \emph{gamma function} is defined for $z \in (0, \infty)$ by
    \[
    \Gamma(z) = \int_0^\infty x^{z-1} e^{-x} \dd{x}
    \]
    and satisfies $\Gamma(n) = (n-1)!$ for all $n \in \NN$. 
\end{recap}
\begin{proof}
    \begin{enumerate}[(a)]
        \item We first consider $\PP(X \geq x)$. 
        By Markov's inequality, we have for all $t \in \RR$ that
        \[
        \PP(X \geq x) = \PP(e^{tX} \geq e^{tx}) \leq e^{-tX} \EE(e^{tX}) \leq e^{-tx + t^2\sigma^2/2}.  
        \]
        Since the LHS is independent of $t$, we can take the infimum over $t$ on the RHS and obtain
        \[
        \PP(X \geq x) \leq \inf_{t \in \RR} e^{-tx + t^2\sigma^2/2} = e^{-x^2/(2\sigma^2)}, 
        \]
        since the infimum of $t^2\sigma^2/2 - tx$ is attained at $t = x/\sigma^2$.

        For $\PP(X \leq -x) = \PP(-X \geq x)$ we can use the fact that $-X$ is also sub-Gaussian with parameter $\sigma^2$.

        \item By the previous part, we have $\PP(\abs{X} \geq x) \leq 2 e^{-x^2/(2\sigma^2)}$. Some calculations give
        \begin{align*}
            \EE(X^{2q}) &= \int_0^\infty \PP(X^{2q} \geq x) \dd{x} = \int_0^\infty \PP(\abs{X} \geq x^{1/(2q)}) \\
            &= 2q \int_0^\infty x^{2q - 1} \PP(\abs{X}\geq x)\dd{x} \\
             &\leq 4q \int_0^\infty x^{2q - 1} e^{-x^2/(2\sigma^2)} \dd{x}.
        \end{align*}
    Now set $t = x^2/2\sigma^2$, so that $x = \sigma (2t)^{1/2}$ and thus $\dd{x} = \sigma (2t)^{-1/2} \dd{t}$. Plugging that in we get
    \begin{align*}
    \EE(X^{2q}) &\leq 4q \int_0^\infty (\sigma(2t)^{1/2})^{2q - 1} e^{-t} \sigma (2t)^{-1/2} \dd{t} = 2^{q+1} q \sigma^{2q} \int_0^\infty t^{q-1} e^{-t} \dd{t} \\
    &= 2^{q+1} q \sigma^{2q} \Gamma(q) = 2 \cdot q! (2\sigma)^q. 
    \end{align*}

\item Note that $x \mapsto e^{-tx}$ is convex for every $t \in \RR$, so $\EE(e^{-tX}) \geq e^{-t \EE(X)} \geq 1$ by Jensen's inequality.
    \end{enumerate}
\end{proof}

### 2020-10-17 modern_statistical_methods/msm_ES1.tex
added example sheet solutions for msm and topics

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[margin=3cm]{geometry}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\usepackage{mathtools, amsmath, amssymb, amsthm, enumerate, mdframed, bbm, graphicx, float, physics, xcolor, cleveref}

\hypersetup{
    colorlinks   = true, %Colours links instead of ugly boxes
    urlcolor     = blue, %Colour for external hyperlinks
    linkcolor    = blue, %Colour of internal links
    citecolor   = red %Colour of citations
}

% Definition of numbered environments.
% Usage: \begin{theorem} ... \end{theorem}
% Remark has no numbering.
\theoremstyle{plain}
\newtheorem{question}{Question}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

% Question with box around it
\newenvironment{qbox}{\begin{mdframed}\begin{question}}{\end{question}\end{mdframed}}

% A proof with "solution" instead of "proof" and no QED symbol
\newenvironment{solution}{\begin{proof}[Solution]\renewcommand\qedsymbol{}}{\end{proof}}

% Some renewed commands
\renewcommand{\vec}{\mathbf}
\renewcommand{\emptyset}{\varnothing}
\renewcommand{\epsilon}{\varepsilon}
\renewcommand{\theta}{\vartheta}
\renewcommand{\phi}{\varphi}

% Frequently used math alphabets
\newcommand{\Bb}{\mathbb}
\newcommand{\Cal}{\mathcal}
\newcommand{\Bf}{\mathbf}
\newcommand{\Rm}{\mathrm}

% Frequently used letters in the blackboard alphabet
\newcommand{\CC}{\Bb C}
\newcommand{\NN}{\Bb N}
\newcommand{\PP}{\Bb P}
\newcommand{\QQ}{\Bb Q}
\newcommand{\RR}{\Bb R}
\newcommand{\EE}{\Bb E}

% Usage: \ang{...} is equivalent to \langle ... \rangle, while \ang*{...} is equivalent to \left\langle ... \right\rangle
% For other delimiters: use \qty from the physics package (i.e., \qty(...))
\DeclarePairedDelimiter{\ang}{\langle}{\rangle}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

% Frequently used commands
\newcommand{\T}{^\top} % Matrix transpose A\T
\newcommand{\C}{^\complement} % Set complement A\C
\newcommand\ceq\coloneqq % Definitions :=
\newcommand\pow{\Cal P} % Power sets
\newcommand\eps\epsilon
\newcommand\ind{\mathbbm 1} % Blackboard 1 for indicator functions
\newcommand\restr{\mathord\restriction}
\newcommand\TODO{{\color{red} TODO: }}

% Functions that appear frequently
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\Int}{Int}
\DeclareMathOperator{\Span}{Span}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\title{Modern Statistical Methods --- Example Sheet 1} % subject
\author{Lucas Riedstra}
\date{...} % date

\begin{document}
\maketitle

\begin{qbox}
    Consider minimising the following objective involving response $Y \in \RR^n$ and design matrix $X \in \RR^{n \times p}$ over $(\mu, \beta) \in \RR \times \RR^p$: 
    \[
    \norm{Y - \mu \Bf{1} - X\beta}_2^2 + J(\beta).
    \]
    Here $J \colon \RR^p \to \RR$ is an arbitrary penalty function. Suppose $\bar{X_k} = 0$ for $k = 1, \dotsc, p$. Assuming that a minimiser $(\hat\mu, \hat\beta)$ exists, show that $\hat\mu = \bar Y$. Now take $J(\beta) = \lambda \norm{\beta}_2^2$ so we have the ridge regression objective. Show that
    \[
    \hat\beta = (X\T X + \lambda I)^{-1} X\T Y.
    \] 
    From here onwards, whenever we refer to ridge regression, we will assume $X$ has had its columns mean-centred. 
\end{qbox}

\begin{solution}
    If a minimiser $\hat\mu$ exists, then the derivative of the objective must be 0. We expand the objective function as
    \begin{align*}
    \norm{Y - \mu\Bf{1} - X\beta}_2^2 + J(\beta) &= \sum_{i=1}^n \qty(Y_i - \mu - (X\beta)_i)^2 + J(\beta) \\
    &= \sum_{i=1}^n \qty(Y_i - \mu - \sum_{j=1}^p X_{ij}\beta_j)^2. 
    \end{align*}
Now, differentiating w.r.t.\ $\mu$ yields
\begin{align*}
\pdv{\mu} \norm{Y - \mu\Bf{1} - X\beta}_2^2 + J(\beta) &= -2 \sum_{i=1}^n \qty( Y_i - \mu - \sum_{j=1}^p X_{ij}\beta_j) \\
&= -2 \qty(\sum_{i=1}^n Y_i  - n\mu - \sum_{i=1}^n \sum_{j=1}^p X_{ij}\beta_j) \\
&= -2 \qty(\sum_{i=1}^n Y_i  - n\mu - \sum_{j=1}^p \beta_j (\sum_{i=1}^p X_{ij})) \\
&= -2 \qty( \sum_{i=1}^n Y_i - n\mu), 
\end{align*}
since $\sum_{i=1}^p X_{ij} = 0$ for all $i$ by assumption. Now, setting the derivative to 0 shows that $\hat\mu = \frac1n \sum_{i=1}^n Y_i = \bar Y$ is the minimiser. Since the second derivative of the objective function is constantly $+2$, this $\hat\mu$ is indeed a minimum.

To compute $\hat\beta$, we differentiate the objective function w.r.t.\ $\beta$: 
\begin{align*}
    \nabla_\beta \norm{Y- \mu \vec 1 - X\beta}_2^2 + \lambda \norm{\beta}_2^2 &= ...
\end{align*}
\TODO Finish
\end{solution}

\begin{question}
    Consider performing ridge regression when $Y = X \beta^0 + \eps$, where $X \in \RR^{n \times p}$ has full column rank, and $\Var(\eps) = \sigma^2 I$. Let the SVD of $X$ be $UDV\T$ and write $U\T X \beta^0 = \gamma$. Show that
    \[
    \frac1n \EE\norm{X \beta^0 - X \hat\beta_\lambda^R}_2^2 = \frac1n \sum_{j=1}^p \qty(\frac\lambda{\lambda + D_{jj}^2})^2 \gamma_j^2 + \frac{\sigma^2}{n} \sum_{j=1}^p \frac{D_{jj}^4}{(\lambda + D_{jj}^2)^2}. 
    \]
\end{question}

\begin{solution}
    We write $\hat\beta \ceq \hat\beta_\lambda^R$. We write out
    \[
    X\hat\beta = UDV\T (V\T D^2 V + \lambda I)^{-1} V\T D U\T Y =  UD^2 (D^2 + \lambda I)^{-1} U\T Y = \sum_j \frac{D_{jj}^2}{\lambda + D_{jj}^2} U_j U_j\T Y. 
    \]
    Since $X\beta^0 = UDV\T \beta^0$ lies in the range of $U$, we have $X\beta^0 = \sum_j U_j U_j\T X\beta^0$. 
    \begin{align*}
        X\beta^0 - X\hat\beta &= \sum_j U_j U_j\T X \beta^0 -  \sum_j \frac{D_{jj}^2}{\lambda + D_{jj}^2} U_j U_j\T Y \\
        &= \sum_j \qty(U_j\T X \beta^0 - \frac{D_{jj}^2 U_j\T Y}{\lambda + D_{jj}^2}) U_j \\
        &= \sum_j \qty(\frac{D_{jj}^2 U_j\T (X\beta^0 - Y) + \lambda U_j\T X \beta^0}{\lambda + D_{jj}^2})
    \end{align*}
and therefore
\begin{align*}
    \frac1n \EE\norm{X\beta^0 - X\hat\beta}_2^2 &= \frac1n \sum_j \EE \qty(\frac{D_{jj}^2 U_j\T (X\beta^0 - Y) + \lambda U_j\T X \beta^0}{\lambda + D_{jj}^2})^2 \\
    &\overset{*}= \frac1n \sum_j \frac{D_{jj}^4}{(\lambda + D_{jj}^2)^2} \EE[(U_j\T (X\beta^0 - Y))^2] + \frac1n \sum_j \qty(\frac{\lambda}{\lambda + D_{jj}^2})^2  (U_j\T X \beta^0)^2 \\
&\overset{**}= \frac{\sigma^2}{n} \sum_j \frac{D_{jj}^4}{(\lambda + D_{jj}^2)^2}  + \frac1n \sum_j \qty(\frac{\lambda}{\lambda + D_{jj}^2})^2  \gamma_j^2.
\end{align*}
Note that the cross-terms in $(*)$ disappear since they are linear combinations of $\EE(X\beta^0 - Y) = 0$, while the equality $**$ holds since 
\[
\EE[(U_j\T(X \beta^0 - Y))^2] = \Var(U_j\T (X\beta^0 - Y)) = U_j\T \Var(X\beta^0 - Y) U_j = \sigma^2 U_j\T U_j = \sigma^2. 
\]
\end{solution}

\begin{question}
    Show that the ridge regression estimates can be otained by ordinary least squares regression on an augmented data set with $\sqrt\lambda I$ added to the botten of $X$, and $p$ zeroes added to the end of the response $Y$. 
\end{question}

\begin{solution}
    Let $\tilde X, \tilde Y$ be the new data set, then our ordinary least squares objective function is
    \begin{align*}
    L(\beta) & \ceq \norm{\tilde Y - \tilde X \beta}^2 \\
    &= \norm{Y - X\beta}^2 + \norm{\sqrt\lambda\beta}^2 \\
    &= \norm{Y - X\beta}^2 + \lambda \norm{\beta}^2,
    \end{align*}
which is exactly our ridge regression objective function. 
\end{solution}

\begin{question}
    In the following, assume that forming $AB$ where $A \in \RR^{a \times b}, B \in \RR^{b \times c}$ requires $O(abc)$ computational operations, and that if $M \in \RR^{d \times d}$ is invertible, then forming $M^{-1}$ requires $O(d^3)$ operations. 
    \begin{enumerate}[(a)]
        \item Suppose we wish to apply ridge regression to data $(Y, X) \in \RR^n \times \RR^{n \times p}$ with $n \gg p$. A complication is that the data is split into $m$ separate datasets of size $n/m \in \NN$, 
        \[
        Y = \mqty( Y^{(1)} \\ \vdots \\ Y^{(m)}), \quad X = \mqty( X^{(1)} \\\vdots\\X^{(m)}),
        \]
        with each dataset located on a different server. Moving large amounts of data between servers is expensive. Explain how one can produce ridge estimates $\hat\beta_\lambda$ by communicating only $O(p^2)$ numbers from each server to some central server. What is the total order of the computation time required at each server, and at the central server for your approach?

        \item Now suppose instead that $p \gg n$ and it is instead the variable sthat are split across $m$ servers, so each server has only a subset of $p/m \in \NN$ variables for each observation, and some central server stores $Y$. Explain how one can obtain the fitted values $X\hat\beta_\lambda$ communicating only $O(n^2)$ numbers from each server to the central server. What is the total order of the computation time required at each server, and at the central server for your approach? 
    \end{enumerate}
\end{question}

\begin{solution}
    \begin{enumerate}[(a)]
        \item Note that
        \[
        X\T Y = \mqty( X^{(1)\top} \dotsc X^{(m)\top}) \mqty( Y^{(1)}\\\vdots\\Y^{(m)}) = \sum_{i=1}^m X^{(i)\top} Y^{(i)}. 
        \]
        Each of the $X^{(i)\top} Y^{(i)}$ can be computed at the data server (requiring $O(np)$ computation) and produces $p$ numbers per data server.

        Analogously, we have
        \[
        X\T X = \sum_{i=1}^m X^{(i)\top} X^{(i)},
        \]
        which can also be computed at the data server (requiring $O(np^2)$ computation) and produces $p^2$ numbers per data server. In the end, each server will therefore have $O(p^2)$ numbers to send, having done $O(np^2)$ computations.

        When each server sends their numbers to the central data server, the central data server can then compute $(X\T X + \lambda I)^{-1}$ using $O(p^3)$ computations and then multiply that with $X\T Y$ which will require $O(np)$ computations.

        \item In this case, we have 
        \[
        X = \mqty( X^{(1)} \dotsb X^{(m)} ). 
        \]
        We use the alternative form of  $\hat\beta_\lambda$, namely
        \[
        X \hat\beta_\lambda^R = XX\T (XX\T + \lambda I)^{-1} Y.
        \]

        Now, we have $XX\T = \sum_{i=1}^m X^{(i)} X^{(i)\top}$, and each $X^{(i)} X^{(i)\top}$ can be computed at the data server, requiring $O(n^2 p)$ computations and producing $O(n^2)$ numbers to send.

        After sending these numbers over to the central server, the inverse $(XX\T + \lambda I)^{-1}$ can be computed in $O(n^3)$ time, then multiplied from the left with $XX\T$ in $O(n^3)$ time, and multiplied with $Y$ in $O(n^2)$ time. The total amount of computation at the central server will therefore be $O(n^3)$ as well. 
    \end{enumerate}
\end{solution}

\end{document}

### 2020-10-17 topics_statistical_theory/topics_ES1.tex
added example sheet solutions for msm and topics

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[margin=3cm]{geometry}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\usepackage{mathtools, amsmath, amssymb, amsthm, enumerate, mdframed, bbm, graphicx, float, physics, xcolor, cleveref}

\hypersetup{
    colorlinks   = true, %Colours links instead of ugly boxes
    urlcolor     = blue, %Colour for external hyperlinks
    linkcolor    = blue, %Colour of internal links
    citecolor   = red %Colour of citations
}

% Definition of numbered environments.
% Usage: \begin{theorem} ... \end{theorem}
% Remark has no numbering.
\theoremstyle{plain}
\newtheorem{question}{Question}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

% Question with box around it
\newenvironment{qbox}{\begin{mdframed}\begin{question}}{\end{question}\end{mdframed}}

% A proof with "solution" instead of "proof" and no QED symbol
\newenvironment{solution}{\begin{proof}[Solution]\renewcommand\qedsymbol{}}{\end{proof}}

% Some renewed commands
\renewcommand{\vec}{\mathbf}
\renewcommand{\emptyset}{\varnothing}
\renewcommand{\epsilon}{\varepsilon}
\renewcommand{\theta}{\vartheta}
\renewcommand{\phi}{\varphi}

% Frequently used math alphabets
\newcommand{\Bb}{\mathbb}
\newcommand{\Cal}{\mathcal}
\newcommand{\Bf}{\mathbf}
\newcommand{\Rm}{\mathrm}

% Frequently used letters in the blackboard alphabet
\newcommand{\CC}{\Bb C}
\newcommand{\NN}{\Bb N}
\newcommand{\PP}{\Bb P}
\newcommand{\QQ}{\Bb Q}
\newcommand{\RR}{\Bb R}
\newcommand{\EE}{\Bb E}

% Usage: \ang{...} is equivalent to \langle ... \rangle, while \ang*{...} is equivalent to \left\langle ... \right\rangle
% For other delimiters: use \qty from the physics package (i.e., \qty(...))
\DeclarePairedDelimiter{\ang}{\langle}{\rangle}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

% Frequently used commands
\newcommand{\T}{^\top} % Matrix transpose A\T
\newcommand{\C}{^\complement} % Set complement A\C
\newcommand\ceq\coloneqq % Definitions :=
\newcommand\pow{\Cal P} % Power sets
\newcommand\eps\epsilon
\newcommand\ind{\mathbbm 1} % Blackboard 1 for indicator functions
\newcommand\restr{\mathord\restriction}
\newcommand\TODO{{\color{red} TODO: }}
\newcommand\toprob{\overset{\Rm{p}}{\to}}
\newcommand\todist{\overset{\Rm{d}}{\to}}
\newcommand\toas{\overset{\Rm{a.s.}}{\to}}
\newcommand\deq{\overset{\Rm{d}}=}
\newcommand\iid{\overset{\Rm{iid}}{\sim}}

% Functions that appear frequently
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\Int}{Int}
\DeclareMathOperator{\Span}{Span}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator\Exp{Exp}
\DeclareMathOperator\Beta{Beta}
\DeclareMathOperator\Gam{Gamma}

\title{Topics in Statistical Theory --- Example Sheet 1} % subject
\author{Lucas Riedstra}
\date{...} % date

\begin{document}
\maketitle

\begin{question}
    Let $U_1, \dotsc, U_n \iid U(0, 1)$ and let $Y_1, \dotsc, Y_{n+1} \iid \Exp(1)$. Writing $S_j \ceq \sum_{i=1}^j Y_i$ for $j = 1, \dotsc, n+1$, show that
    \[
    U_{(j)} \deq \frac{S_j}{S_{n+1}} \sim \Beta(j, n - j + 1)
    \]
    for $j = 1, \dotsc, n$. 
\end{question}

\begin{solution}
    First we compute the distribution function of $X_{(j)}$. If $X_{(j)} \leq x$, then at least $j$ of the $X_i$ must be $\leq x$.

For $k = j, j+1, \dotsc, n$, there are $\binom{n}{k}$ ways to choose $k$ of the $X_i$'s that must be $\leq j$. Fix $x \in [0, 1]$, then we have
\begin{align*}
    \PP(X_{(j)} \leq x) &= \sum_{k=j}^n \binom{n}{k} \PP(X_1 \leq x, \dotsc, X_k \leq x, X_{k+1} > x, \dotsc, X_n > x) \\
    &= \sum_{k=j}^n \binom{n}{k} (F(x))^k \cdot (1 - F(x))^{n-k} \\
    &=  \sum_{k=j}^n \binom{n}{k} x^k (1-x)^{n-k}.
\end{align*}

Now, the density function of $X_{(k)}$ is given by the derivative of $F$. This is a hefty calculation, see for example \href{https://math.stackexchange.com/a/1179378/696417}{this StackExchange link}, but it turns out to be
\[
\dv{x} F_{(j)}(x) = f_{(j)}(x) = \frac{n!}{(j-1)!(n-j)!} x^{j-1}(1-x)^{n-j},
\]
which is also the density function of the $\Rm{Beta}(j, n-j+1)$ distribution.

Finally, define $T = S_{n+1} - S_j$, so that $S_j$ and $T$ are independent. It is known that $S_j \sim \Gam(j, 1)$, $T \sim \gamma(n - j + 1)$, and furthermore that
\[
\frac{S_j}{S_{n+1}} = \frac{S_j}{S_j + T} \deq \frac{\Gamma(j, 1)}{\Gamma(j, 1) + \Gamma(n - j + 1, 1)} \sim \Beta(j, n - j + 1). 
\]
\end{solution}

\begin{question}
    Let $X$ be a random variable with mean zero that satisfies $a \leq X \leq b$. Use convexity to show that for every $t \in \RR$, we have
    \[
    \log \EE(e^{tX}) \leq - \alpha u + \log(\beta + \alpha e^u),
    \]
    where $u \ceq t(b-a)$ and $\alpha \ceq 1 - \beta = -a/(b-a)$. Using a second-order Taylor expansion around the origin, deduce that $\log \EE(e^{tX}) \leq t^2 (b-a)^2/8$. 
\end{question}

\begin{proof}
    Let $x \in [a, b]$, then we know there exists a unique $\lambda \in [0, 1]$ such that $x = (1-\lambda) a + \lambda b$. A simple computation gives $\lambda = (x-a)/(b-a)$, $1 - \lambda = (b-x)/(b-a)$. By convexity of $t \mapsto e^{tx}$ we find
    \[
    e^{tx} \leq \frac{b-x}{b-a} e^{t a} + \frac{x-a}{b-a} e^{t b}. 
    \]

    From this we deduce that 
    \[
    \EE[e^{tX}] \leq \EE\qty[\frac{b-X}{b-a} e^{t a} + \frac{X-a}{b-a} e^{t b}] = \frac{b}{b - a} e ^{ta} + \frac{-a}{b - a} e^{tb} = \beta e^{ta} + \alpha e^{tb} = e^{-\alpha u + \log(\beta + \alpha e^u)}. 
    \]
    Since $\log$ is increasing, we can take the logarithm on both sides to conclude
    \[
    \log \EE[e^{tX}] \leq -\alpha u + \log(\beta + \alpha e^u). 
    \]

    Now, we compute the taylor polynomial of $f(u) \ceq -\alpha u + \log(\beta + \alpha e^u)$ in $u = 0$: we have
    \begin{align*}
        f(0) &= \log(\beta + \alpha) = \log(1) = 0; \\
        f'(u) &= -\alpha + \frac{\alpha e^u}{\beta + \alpha e^u}; \\
        f'(0) &= -\alpha + \frac{\alpha}{\beta + \alpha} = 0; \\
        f''(u) &= \frac{(\beta + \alpha e^u)\alpha e^u - (\alpha e^u)^2}{(\beta + \alpha e^u)^2} = \frac{\alpha e^u}{(\beta + \alpha e^u)} \qty(1 - \frac{\alpha e^u}{(\beta + \alpha e^u)})
    \end{align*}

Note that $\frac{\alpha e^u}{\beta + \alpha e^u} \in [0, 1]$ since $\alpha, \beta \geq 0$ (this holds because $a$ must be negative and $b$ must be positive due to the condition $\EE X = 0$). For $y \in [0, 1]$, the polynomial $y(1-y)$ takes values in $[0, \frac14]$. Therefore, by Taylor's theorem with remainder, we conclude
\[
\log \EE[e^{tX}] \leq \qty(\sup_{u \in \RR} \frac{f''(u)}{2} )u^2 = \frac18 u^2 = \frac{t^2(b-a)^2}{8}. 
\]
\end{proof}

\begin{question}
    Let $X_1, \dotsc, X_n$ be independent with distribution $P$ on a measurable space $(\Cal X, \Cal A)$, and let $\hat P_n$ be the empirical measure of $X_1, \dotsc, X_n$; thus $\hat P_n(A) = n^{-1} \sum_{i=1}^n \ind_{X_i \in A}$. Show that, for all $\eps > 0$ and $A \in \Cal A$, we have
    \[
    \PP\qty(\abs{\hat P_n(A) - P(A)} > \eps) \leq 2 e^{-2n\eps^2}. 
    \]
\end{question}

\begin{proof}
    Define a new distribution $Y = \ind_{X \notin A}$. Its distribution function is given by
    \[
    F_Y(y) = \begin{cases}
        0 &\text{$y < 0$}; \\
        P(A) &\text{$y \in [0, 1)$}; \\
        1 &y \geq 1. 
    \end{cases}, \quad F_Y(0) = P(A). 
    \]
    The empirical distribution function of $Y_1, \dotsc, Y_n \iid Y$ is given by
    \[
    \hat F_n(y) = \frac1n \sum_{i=1}^n \ind_{Y_i \leq y},
    \]
    and thus for $y = 0$ we have
    \[
    \hat F_n(0) = \frac1n \sum_{i=1}^n \ind_{Y_i \leq 0} = \frac1n \sum_{i=1}^n \ind_{X_i \in A} = \hat P_n(A).
    \]

    The result now follows from the DKW inequality. \TODO make more precise. 
\end{proof}
\end{document}

### 2020-10-18 modern_statistical_methods/msm_ES1.tex
updates topics summary and example sheets for msm & topics

@@
 
-To compute $\hat\beta$, we differentiate the objective function w.r.t.\ $\beta$: 
-\begin{align*}
-    \nabla_\beta \norm{Y- \mu \vec 1 - X\beta}_2^2 + \lambda \norm{\beta}_2^2 &= ...
-\end{align*}
-\TODO Finish
+The computation of $\hat\beta$ is given in the lecture notes. 
 \end{solution}
@@
 
+\begin{question}
+	Prove Proposition 4 in the notes.
+\emph{Hint: for part (ii) it may help to consider the eigendecompositions of positive semi-definite matrices $K^{(1)}$ and $K^{(2)}$ derived from kernels $k_1$ and $k_2$ in the form $K^{(1)} = PDP\T = \sum_{i=1}^n P_iP_i\T D_{ii}$ for example.
+}
+\end{question}
+
+\begin{proof}
+	Proposition 4 is the following:
+	\begin{mdframed}
+		Suppose $k_1, k_2, \dotsc$ are kernels. 
+		\begin{enumerate}[(i)]
+			\item If $\alpha_1, \alpha_2 \geq 0$ then $\alpha_1 k_1 + \alpha_2 k_2$ is a kernel. If $\lim_{k\to\infty} k(x, x') \eqqcolon k(x, x')$ exists for all $x, x' \in \Cal X$, then $k$ is a kernel. 
+			\item The pointwise product $k = k_1k_2$ is a kernel.
+		\end{enumerate}
+	\end{mdframed}
+
+Let $x_1, \dotsc, x_n \in \Cal X$ and define the positive definite matrix $K^{(m)}$ by $K^{(m)}_{i, j} = k_m(x_i, x_j)$. 
+\begin{enumerate}[(i)]
+	\item Since $(\alpha_1 k_1 + \alpha_2 k_2)(x_i, x_j) = \alpha_1 k_1(x_i, x_j) + \alpha_2 k_2(x_i, x_j)$, the kernel matrix $K_{ij} \ceq (\alpha_1 k_1 + \alpha_2 k_2)(x_i, x_j)$ equals $\alpha_1 K_1 + \alpha_2 K_2$, and is therefore clearly symmetric positive semi-definite.
+This shows that $\alpha_1 k_1 + \alpha_2 k_2$ is a kernel.
+
+	For the second part (the pointwise limit of kernels is a kernel), it suffices to show that the entrywise limit of positive semi-definite matrices is again positive semidefinite.
+But this is clear: if $K^{(m)} \to K$ entrywise, then for any $\vec x \in \RR^n$ we have
+	\[
+	\vec x\T K \vec x = \vec x\T (\lim_{m\to\infty} K^{(m)}) \vec x = \lim_{m\to\infty} \vec x\T K^{(m)}\vec x \geq 0,
+	\]
+	so $K$ is positive semi-definite.
+
+	\item Let $K^{(1)}$ and $K^{(2)}$ be positive semi-definite, then we must show that the pointwise product $K^{(1)} \circ K^{(2)}$ is positive semi-definite as well.
+Write $K^{(1)} = \sum_{\ell=1}^n \lambda_\ell P_\ell P_\ell\T$ and $K^{(2)} = \sum_{\ell=1}^n \mu_\ell Q_\ell Q_\ell\T$, where $\lambda_i, \mu_i \geq 0$ and $\qty{P_\ell}$, $\qty{Q_\ell}$ are orthonormal bases of $\RR^n$.
+
+	Then we find
+	\begin{align*}
+	(K^{(1)} \circ K^{(2)})_{ij} &= \qty(\sum_m \lambda_m P_{im} P_{jm}) \qty(\sum_\ell \mu_\ell Q_{i\ell} Q_{j\ell}) = \sum_{m, \ell} \lambda_m \mu_\ell (P_{im} P_{jm} Q_{i\ell} Q_{j\ell}) \\
+	&= \sum_{m, \ell} \lambda_m \mu_\ell \qty[(P_m \circ Q_\ell)(P_m \circ Q_\ell)\T]_{ij},
+	\end{align*}
+
+	and therefore $K^{(1)} \circ K^{(2)} = \sum_{m, \ell} \lambda_m \mu_\ell \qty[(P_m \circ Q_\ell)(P_m \circ Q_\ell)\T]$, which is a nonnegative combination of positive semi-definite matrices and therefore positive semi-definite.
+\end{enumerate}
+
+\end{proof}
 \end{document}

### 2020-10-18 topics_statistical_theory/C1.tex
updates topics summary and example sheets for msm & topics

@@
 
+\begin{recap}
+	Recall the \emph{tail bound formula} for the expectation: if $X$ is a nonnegative random variable, then 
+	\[
+	\EE[X] = \int_0^\infty \PP(X > x) \dd{x}. 
+	\]
+
+	Furthermore, recall that the \emph{gamma function} is defined for $z \in (0, \infty)$ by
+	\[
+	\Gamma(z) = \int_0^\infty x^{z-1} e^{-x} \dd{x}
+	\]
+	and satisfies $\Gamma(n) = (n-1)!$ for all $n \in \NN$.
+
+	Finally, recall the following inequality: for all $a, b \in \RR$ and $p \geq 1$
+	\[
+	(a+b)^p \leq 2^{p-1} (a^p + b^p). 
+	\]
+	This follows from the convexity of the function $x \mapsto x^p$. 
+\end{recap}
+
 \begin{proposition}We consider some characterisations of sub-Gaussianity: 
@@
 
-\begin{recap}
-    Recall the \emph{tail bound formula} for the expectation: if $X$ is a nonnegative random variable, then 
-    \[
-    \EE[X] = \int_0^\infty \PP(X > x) \dd{x}. 
-    \]
-
-    Furthermore, recall that the \emph{gamma function} is defined for $z \in (0, \infty)$ by
-    \[
-    \Gamma(z) = \int_0^\infty x^{z-1} e^{-x} \dd{x}
-    \]
-    and satisfies $\Gamma(n) = (n-1)!$ for all $n \in \NN$. 
-\end{recap}
 \begin{proof}
@@
 
-\item Note that $x \mapsto e^{-tx}$ is convex for every $t \in \RR$, so $\EE(e^{-tX}) \geq e^{-t \EE(X)} \geq 1$ by Jensen's inequality.
+\item Note that $x \mapsto e^{-tx}$ is convex for every $t \in \RR$, so $\EE(e^{-tX}) \geq e^{-t \EE(X)} = e^{0} = 1$ by Jensen's inequality.
+Let $X'$ denote an independent copy of $X$: then $X - X'$ has a symmetric distribution, so all its odd moments vanish.
+Therefore we find
+\begin{align*}
+	\EE[e^{tX}] &\leq \EE[e^{-tX'}] \EE[e^{tX}] = \EE[e^{t(X - X')}] = \EE \sum_{q=0}^\infty \qty[\frac{t^{2q} (X - X')^{2q}}{(2q)!}] \\
+	&= \sum_{q=0}^\infty \frac{t^{2q} \EE[(X - X')^{2q}]}{(2q)!} \leq \sum_{q=0}^\infty \frac{2^{2q-1} t^{2q} \qty(\EE[X^{2q}] + \EE[(X')^{2q}])}{(2q)!} \\
+	&\leq \sum_{q=0}^\infty  \frac{2^{2q - 1}t^{2q} 2 q! C^{2q}}{(2q)!} =\sum_{q=0}^\infty  \frac{(2tC)^{2q}q!}{(2q)!} = \sum_{q=0}^\infty \frac{(2tC)^{2q}}{\prod_{j=1}^q (q+j)} \\
+	&\leq \sum_{q=0}^\infty \frac{(2tC)^{2q}}{\prod_{j=1}^q (2j)} = \sum_{q=1}^\infty \frac{(2t^2C^2)^q}{q!} = e^{2t^2C^2}.
+\end{align*}
+This shows that $X$ is sub-Gaussian with parameter $4C^2$. 
     \end{enumerate}
 \end{proof}
+
+Note that the proposition is not an ``if and only if''-type theorem: suppose we start with a sub-Gaussian variable $X$ with parameter $\sigma^2$.
+Then by (b), we have $\EE[X^{2q}] \leq q!(2\sigma)^{2q}$, and (c) then implies that $X$ is sub-Gaussian with parameter $16\sigma^2$.
+
+\begin{theorem}[Hoeffding's inequality]
+	Let $X_1, \dotsc, X_n$ be independent sub-Gaussian random variables, with $X_i$ having parameter $\sigma_i^2$.
+Then $\bar X$ is sub-Gaussian with parameter $\overline{\sigma^2}$.
+In particular, we have
+	\[
+	\max\qty{\PP(\bar X \geq x), \PP(\bar X \leq -x)} \leq e^{-nx^2/(2\overline{\sigma^2})}.
+	\]
+\end{theorem}
+
+\begin{proof}
+	For $t \in \RR$, we have
+	\[
+	\EE[e^{t\bar X}] = \EE[e^{(t/n)\sum_i X_i}] = \prod_{i=1}^n \EE[e^{(t/n)X_i}] \leq \prod_{i=1}^n e^{t^2\sigma_i^2/(2n^2)} = e^{t^2\overline{\sigma^2}/(2n)},
+	\]
+	which shows $\bar X$ is sub-Gaussian with parameter $\overline{\sigma^2}/n$. Applying part (a) of the previous proposition shows the second result. 
+\end{proof}
+
+\begin{remark}
+	A direct consequence of Hoeffding's inequality is that
+	\[
+	\PP(\abs{\bar X} \geq x) \leq 2 e^{-nx^2/(2\overline{\sigma^2})}.
+	\]
+	The inequality is often stated in this weaker way. 
+\end{remark}
+
+\begin{lemma}[Hoeffding's lemma]
+	Let $X$ be a random variable with $\EE X = 0$ that satisfies $a \leq X \leq b$. Then $X$ is sub-Gaussian with parameter $(b-a)^2/4$. 
+\end{lemma}
+
+\begin{proof}
+	See Example Sheet 1, question 2. 
+\end{proof}

### 2020-10-18 topics_statistical_theory/topics_ES1.tex
updates topics summary and example sheets for msm & topics

@@
 \begin{solution}
-    First we compute the distribution function of $X_{(j)}$. If $X_{(j)} \leq x$, then at least $j$ of the $X_i$ must be $\leq x$.
+%    First we compute the distribution function of $U_{(j)}$. If $U_{(j)} \leq x$, then at least $j$ of the $U_{i}$ must be $\leq x$. 
+%
+%For $k = j, j+1, \dotsc, n$, there are $\binom{n}{k}$ ways to choose $k$ of the $U_{i}$'s that must be $\leq j$. Fix $x \in [0, 1]$, then we have
+%\begin{align*}
+%    \PP(U_{(j)} \leq x) &= \sum_{k=j}^n \binom{n}{k} \PP(X_1 \leq x, \dotsc, X_k \leq x, X_{k+1} > x, \dotsc, X_n > x) \\
+%    &= \sum_{k=j}^n \binom{n}{k} (F(x))^k \cdot (1 - F(x))^{n-k} \\
+%    &=  \sum_{k=j}^n \binom{n}{k} x^k (1-x)^{n-k}.
+%\end{align*}
+%
+%Now, the density function of $X_{(k)}$ is given by the derivative of $F$.
+This is a hefty calculation, see for example \href{https://math.stackexchange.com/a/1179378/696417}{this StackExchange link}, but it turns out to be
+%\[
+%\dv{x} F_{(j)}(x) = f_{(j)}(x) = \frac{n!}{(j-1)!(n-j)!} x^{j-1}(1-x)^{n-j},
+%\]
+%which is also the density function of the $\Rm{Beta}(j, n-j+1)$ distribution.
 
-For $k = j, j+1, \dotsc, n$, there are $\binom{n}{k}$ ways to choose $k$ of the $X_i$'s that must be $\leq j$. Fix $x \in [0, 1]$, then we have
+We compute the density function of $U_{(j)}$ as follows: let $x \in (0, 1)$, then we know that 
+\[
+f_{(j)}(x) = \dv{x} F_{(j)}(x) = \lim_{h\to 0} \frac{F_{(j)}(x+h) - F_{(j)}(x)}{h} = \lim_{h \to 0} \frac{\PP(x < U_{(j)} \leq x + h)}{h}. 
+\]
+The probability $\PP(x < U_{(j)} \leq x + h)$ is the probability that exactly $j-1$ of the $U_{i}$ are less than $x$, and that at least one of the $U_{i}$ is in $(x, x+ h]$.
+
+The probability that two or more of the $U_{i}$ lie in $(x, x+h]$ is $O(h^2)$ and therefore negligible, so we must compute the probability that exactly $j-1$ of the $U_{i}$ are smaller than $x$, one of the $U_{i}$ is in $(x, x+h]$, and the other $U_{i}$ are greater than $x+h$.
+This is easily seen to be
 \begin{align*}
-    \PP(X_{(j)} \leq x) &= \sum_{k=j}^n \binom{n}{k} \PP(X_1 \leq x, \dotsc, X_k \leq x, X_{k+1} > x, \dotsc, X_n > x) \\
-    &= \sum_{k=j}^n \binom{n}{k} (F(x))^k \cdot (1 - F(x))^{n-k} \\
-    &=  \sum_{k=j}^n \binom{n}{k} x^k (1-x)^{n-k}.
+&\binom{n}{j-1} \PP(U \leq x)^{j-1} \cdot (n - j + 1) \PP(x < U \leq x + h) \cdot \PP(U > x + h)^{n-j} \\
+&= \frac{n!}{(j-1)!(n-j+1)!} (n-j+1) x^{j-1} h (1 - x - h)^{n-j} \\
+&= \frac{n!}{(j-1)!(n-j)!} x^{j-1} (1-x-h)^{n-j} h. 
 \end{align*}
-
-Now, the density function of $X_{(k)}$ is given by the derivative of $F$.
-This is a hefty calculation, see for example \href{https://math.stackexchange.com/a/1179378/696417}{this StackExchange link}, but it turns out to be
+Therefore, we easily compute
 \[
-\dv{x} F_{(j)}(x) = f_{(j)}(x) = \frac{n!}{(j-1)!(n-j)!} x^{j-1}(1-x)^{n-j},
+f_{(j)}(x) = \lim_{h \to 0} \frac{\frac{n!}{(j-1)!(n-j)!} x^{j-1} (1-x-h)^{n-j} h}{h} = \frac{n!}{(j-1)!(n-j)!} x^{j-1} (1-x)^{n-j}. 
 \]
-which is also the density function of the $\Rm{Beta}(j, n-j+1)$ distribution.
+This is also the density function of a $\Beta(j, n-j+1)$ distribution.
 
@@
 \begin{question}
-    Let $X_1, \dotsc, X_n$ be independent with distribution $P$ on a measurable space $(\Cal X, \Cal A)$, and let $\hat P_n$ be the empirical measure of $X_1, \dotsc, X_n$; thus $\hat P_n(A) = n^{-1} \sum_{i=1}^n \ind_{X_i \in A}$.
+    Let $X_1, \dotsc, X_n$ be independent with distribution $P$ on a measurable space $(\Cal X, \Cal A)$, and let $\hat P_n$ be the empirical measure of $X_1, \dotsc, X_n$; thus $\hat P_n(A) = n^{-1} \sum_{i=1}^n \ind_{U_{i} \in A}$.
 Show that, for all $\eps > 0$ and $A \in \Cal A$, we have
@@
     \[
-    \hat F_n(0) = \frac1n \sum_{i=1}^n \ind_{Y_i \leq 0} = \frac1n \sum_{i=1}^n \ind_{X_i \in A} = \hat P_n(A).
+    \hat F_n(0) = \frac1n \sum_{i=1}^n \ind_{Y_i \leq 0} = \frac1n \sum_{i=1}^n \ind_{U_{i} \in A} = \hat P_n(A).
     \]

### 2020-10-18 distribution_theory/distribution_ES1.tex
started on Distribution sheet 1

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[margin=3cm]{geometry}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\usepackage{mathtools, amsmath, amssymb, amsthm, mathabx, enumerate, mdframed, bbm, graphicx, float, physics, xcolor, cleveref}

\hypersetup{
    colorlinks   = true, %Colours links instead of ugly boxes
    urlcolor     = blue, %Colour for external hyperlinks
    linkcolor    = blue, %Colour of internal links
    citecolor   = red %Colour of citations
}

% Definition of numbered environments.
% Usage: \begin{theorem} ... \end{theorem}
% Remark has no numbering.
\theoremstyle{plain}
\newtheorem{question}{Question}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

% Question with box around it
\newenvironment{qbox}{\begin{mdframed}\begin{question}}{\end{question}\end{mdframed}}

% A proof with "solution" instead of "proof" and no QED symbol
\newenvironment{solution}{\begin{proof}[Solution]\renewcommand\qedsymbol{}}{\end{proof}}

% Some renewed commands
\renewcommand{\vec}{\mathbf}
\renewcommand{\emptyset}{\varnothing}
\renewcommand{\epsilon}{\varepsilon}
\renewcommand{\theta}{\vartheta}
\renewcommand{\phi}{\varphi}

% Frequently used math alphabets
\newcommand{\Bb}{\mathbb}
\newcommand{\Cal}{\mathcal}
\newcommand{\Bf}{\mathbf}
\newcommand{\Rm}{\mathrm}

% Frequently used letters in the blackboard alphabet
\newcommand{\CC}{\Bb C}
\newcommand{\NN}{\Bb N}
\newcommand{\PP}{\Bb P}
\newcommand{\QQ}{\Bb Q}
\newcommand{\RR}{\Bb R}
\newcommand{\EE}{\Bb E}
\newcommand{\DD}{\Cal D}

% Usage: \ang{...} is equivalent to \langle ... \rangle, while \ang*{...} is equivalent to \left\langle ... \right\rangle
% For other delimiters: use \qty from the physics package (i.e., \qty(...))
\DeclarePairedDelimiter{\ang}{\langle}{\rangle}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

% Frequently used commands
\newcommand{\T}{^\top} % Matrix transpose A\T
\newcommand{\C}{^\complement} % Set complement A\C
\newcommand\ceq\coloneqq % Definitions :=
\newcommand\pow{\Cal P} % Power sets
\newcommand\eps\epsilon
\newcommand\ind{\mathbbm 1} % Blackboard 1 for indicator functions
\newcommand\restr{\mathord\restriction}
\newcommand\TODO{{\color{red} TODO: }}
\newcommand\pt\partial
\newcommand\locint{L^1_{\Rm{loc}}}
\newcommand\refl\widecheck

% Functions that appear frequently
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\Int}{Int}
\DeclareMathOperator{\Span}{Span}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\ord}{ord}

\newcommand\clos\overline

\title{Distribution Theory and Applications --- Example Sheet 1} % subject
\author{Lucas Riedstra}
\date{...} % date

\begin{document}
\maketitle

\begin{question}
	Construct a non-zero element of $\DD(\RR)$ that vanishes outside $(0, 1)$. Construct a non-zero of $\DD(\RR^n)$ that vanishes outside the ball $B_\eps = \qty{x \in \RR^n : \norm{x} < \eps}$. 
\end{question}

\begin{proof}
	It is well-known that the function 
	\[
	\phi \colon \RR \to \RR \colon x \mapsto \begin{cases}
		0 &\text{if $x \leq 0$}; \\
		e^{-1/x} &\text{if $x > 0$},
	\end{cases}
	\]
	is smooth and vanishes outside $(0, \infty)$. The function $\psi(x) \ceq \phi(x)\phi(1-x)$ is therefore also smooth and vanishes outside $(0, 1)$.

	Since $\psi$ vanishes outside $(0, 1)$, the function $\psi(x/\eps)$ vanishes outside $(0, \eps)$, and therefore the function $\vec x \mapsto \psi(\norm{\vec x}/\eps)$ vanishes outside $B_\eps$. 
\end{proof}

\begin{question}
	Given $\phi \in \DD(X)$, Taylor's theorem gives
	\[
	\phi(x+h) = \sum_{\abs{\alpha} \leq N} \frac{h^\alpha}{\alpha!} \pt^\alpha \phi(x) + R_N(x, h).
	\]
	Prove that $\supp(R_N)$ is contained in some fixed compact $K \subseteq X$ for $\abs{h}$ sufficiently small. Show also that $\pt^\alpha R_N = o(\abs{h}^N)$ uniformly in $x$ for each multi-index $\alpha$, i.e.\ prove 
	\[
	\lim_{\abs{h}\to 0} \frac{\sup_x \abs{\pt^\alpha R_N(x, h)}}{\abs{h}^N} = 0
	\]
	for each multi-index $\alpha$.

	\emph{Hint: you may find it convenient to use the following form of the remainder}
	\[
	R_N(x, h) = \sum_{\abs{\alpha} = N+1} \frac{h^\alpha}{\alpha!}(N+1)\int_0^1 (1-t)^N (\pt^\alpha\phi)(x + th)\dd{t},
	\]
	\emph{and note that} $(N+1)! \sum_{\abs{\alpha} = N+1} \frac{h^\alpha}{\alpha!} = (h_1 + \dotsb + h_n)^{N+1}$. 
\end{question}

\begin{solution}
	Since $\phi \in \DD(X)$, we know that $\supp\phi \subseteq \clos{B_N}$ for some $N \in \NN$. Now, suppose $\norm{h} < 1$, then
	\[
	\phi(x+h) \neq 0 \implies \norm{x + h} \leq N \implies \norm{x} \leq \norm{x + h} + \norm{h} \leq N + 1, 
	\]
	so if we define $\psi_h(x) = \phi(x+h)$ then we know that $\supp\phi_h \subseteq \clos{B_{N+1}}$.

	By Taylor's theorem, we have
	\[
	\phi(x+h) = \sum_{\abs{\alpha} \leq N} \frac{h^\alpha}{\alpha!} \pt^\alpha \phi(x) + R_N(x, h),  
	\]
	and since $\sum_{\abs{\alpha} \leq N} \frac{h^\alpha}{\alpha!} \pt^\alpha \phi(x)$ vanishes for $x \notin \clos{B_N}$, it is clear that $\supp(R_N(\cdot, h))$ must also be contained in $\clos{B_{N+1}}$ (again, for $\norm{h} \leq 1$). This shows that $\supp(R_N)$ is contained in $\clos{B_{N+1}}$ for $\abs{h}$ sufficiently small.

	Now let $\beta$ be a multi-index and define $C \ceq \frac{1}{N!} \max_{\abs{\alpha} = N+1, x \in \RR^n} \abs{(\pt^{\alpha + \beta})\phi(x)}$ (note that $C$ exists and is finite since all partial derivatives of $\phi$ have compact support), then we have
	\begin{align*}
	\abs{\pt^\beta R_N(x, h)} &= \abs{\pt^\beta \sum_{\abs{\alpha} = N+1} \frac{h^\alpha}{\alpha!}(N+1)\int_0^1 (1-t)^N (\pt^\alpha\phi)(x + th)\dd{t}} \\
	%&= \abs{\sum_{\abs{\alpha} = N+1} \frac{h^\alpha}{\alpha!}(N+1) \cdot \pt^\beta\int_0^1 (1-t)^N (\pt^\alpha\phi)(x + th)\dd{t}} \\
	&\overset\star= \abs{\sum_{\abs{\alpha} = N+1} \frac{h^\alpha}{\alpha!}(N+1) \int_0^1 (1-t)^N (\pt^{\alpha + \beta}\phi)(x + th)\dd{t}} \\
	&\leq \sum_{\abs{\alpha} = N+1} \frac{\abs{h^\alpha}}{\alpha!}(N+1) \int_0^1 (1-t)^N \abs{\qty(\pt^{\alpha + \beta}\phi)(x+th)} \dd{t} \\
	&\leq \qty[\max_{\abs{\alpha} = N+1, x \in \RR^n} \abs{(\pt^{\alpha + \beta})\phi(x)}] \sum_{\abs{\alpha} = N+1} \frac{\abs{h^\alpha}}{\alpha!} (N+1) \\
	&\leq C (N+1)! \sum_{\abs{\alpha}= N+1}\frac{\abs{h^\alpha}}{\alpha!} = C(\abs{h_1} + \dotsb + \abs{h_n})^{N+1}.
\end{align*}
Since this upper bound does not depend on $x$, we also have
\[
\sup_x \abs{\pt^\beta R_N(x, h)} \leq C(\abs{h_1} + \dotsb + \abs{h_n})^{N+1}, 
\]
and we conclude that
\[ 
\frac{\sup_x \abs{\pt^\beta R_N(x, h)}}{\norm{h}^N} \leq \frac{C(\abs{h_1} + \dotsb + \abs{h_n})^{N+1}}{\norm{h}^N} \leq \frac{C N^{N+1} \norm{h}^{N+1}}{\norm{h}^N} = CN^{N+1} \norm{h} \to 0, 
\]
and therefore that $\pt^\beta R_N(x, h) = o(\norm{h}^n)$ for all multi-indices $\beta$. 
\end{solution}
\end{document}

### 2020-10-19 inverse_problems/C1.tex
updatd msm & ip notes

@@
     \begin{enumerate}
-        \item     An operator $A \in \LLL(\XXX, \YYY)$ is called \emph{bounded} if
+        \item     An linear operator $A \colon \XX \to \YY$ is called \emph{bounded} if
         \[
-        \norm{A}_{\LLL(\XXX, \YYY)} \ceq \sup_{u \neq 0} \frac{\norm{Au}_\YYY}{\norm{u}_\XXX} = \sup_{\norm{u}_\XXX \leq 1} < \infty. 
+        \norm{A}_{\LL(\XX, \YY)} \ceq \sup_{u \neq 0} \frac{\norm{Au}_\YY}{\norm{u}_\XX} = \sup_{\norm{u}_\XX \leq 1} < \infty. 
         \]
-            It is known that a linear operator between normed spaces is continuous if and only if it is bounded. 
+            It is known that a linear operator between normed spaces is continuous if and only if it is bounded. The set of bounded linear operators from $\XX$ to $\YY$ is denotes $\LL(\XX, \YY)$. 
             \item 
@@
 
-        \item     We will assume $\XXX$ and $\YYY$ are Hilbert spaces, so there is an inner product $\ang{\cdot, \cdot}$ and any bounded operator $A \in \LLL(\XXX, \YYY)$ has a unique adjoint $A^* \in \LLL(\YYY, \XXX)$ which satisfies
+        \item     We will assume $\XX$ and $\YY$ are Hilbert spaces, so there is an inner product $\ang{\cdot, \cdot}$ and any bounded operator $A \in \LL(\XX, \YY)$ has a unique adjoint $A^* \in \LL(\YY, \XX)$ which satisfies
         \[
-        \ang{Au, v}_\YYY = \ang{u, A^* v}_\XXX \quad\text{for all $u \in \XXX$, $v \in \YYY$.}
+        \ang{Au, v}_\YY = \ang{u, A^* v}_\XX \quad\text{for all $u \in \XX$, $v \in \YY$.}
         \]
 
-        \item For any $\XXX' \subseteq \XXX$ we define the \emph{orthogonal complement} of $\XXX'$ as
+        \item For any $\XX' \subseteq \XX$ we define the \emph{orthogonal complement} of $\XX'$ as
         \[
-        (\XXX')\P \ceq \qty{u \in \XXX \mid \ang{u, v}_\XXX = 0 \ \forall v \in \XXX'}. 
+        (\XX')\P \ceq \qty{u \in \XX \mid \ang{u, v}_\XX = 0 \ \forall v \in \XX'}. 
         \]
-        It is known that $(\XXX')\P$ is a closed subspace of $\XXX$ and that $\XXX' \subseteq ((\XXX')\P)\P$, where equality holds if and only if $\XXX'$ is a closed subspace of $\XXX$.
-For a non-closed subspace $\XXX'$ we have $((\XXX')\P)\P = \clos{\XXX'}$.
+        It is known that $(\XX')\P$ is a closed subspace of $\XX$ and that $\XX' \subseteq ((\XX')\P)\P$, where equality holds if and only if $\XX'$ is a closed subspace of $\XX$.
+For a non-closed subspace $\XX'$ we have $((\XX')\P)\P = \clos{\XX'}$.
 
-        \item If $\XXX'$ is a closed subspace of $\XXX$, then for any $u \in \XXX$ there exist unique $x_u \in \XXX'$, $x_u\P \in (\XXX')\P$ such that $u = x_u + x_u\P$.
-The map $u \mapsto x_u$ is denoted $P_{\XXX'}$ and is called the \emph{orthogonal projection} on $\XXX'$.
+        \item If $\XX'$ is a closed subspace of $\XX$, then for any $u \in \XX$ there exist unique $x_u \in \XX'$, $x_u\P \in (\XX')\P$ such that $u = x_u + x_u\P$.
+The map $u \mapsto x_u$ is denoted $P_{\XX'}$ and is called the \emph{orthogonal projection} on $\XX'$.
 Properties are:
         \begin{enumerate}
-            \item $P_{\XXX'}$ is bounded and self-adjoint with norm 1;
-            \item $P_{\XXX'} + P_{(\XXX')\P} = I$;
-            \item $P_{\XXX'}u$ minimises the distance from $u$ to $\XXX'$;
-            \item $x = P_{\XXX'}u$ if and only if $x \in \XXX'$ and $u - x \in (\XXX')\P$. 
+            \item $P_{\XX'}$ is bounded and self-adjoint with norm 1;
+            \item $P_{\XX'} + P_{(\XX')\P} = I$;
+            \item $P_{\XX'}u$ minimises the distance from $u$ to $\XX'$;
+            \item $x = P_{\XX'}u$ if and only if $x \in \XX'$ and $u - x \in (\XX')\P$. 
         \end{enumerate}
 
-    \item For any $A \in \LLL(\XXX, \YYY)$ we have
+    \item For any $A \in \LL(\XX, \YY)$ we have
     \[
@@
 \begin{lemma}
-    For any $A \in \LLL(\XXX, \YYY)$ we have $\clos{\Ran(A^*A)} = \clos{\Ran(A^*)}$. 
+    For any $A \in \LL(\XX, \YY)$ we have $\clos{\Ran(A^*A)} = \clos{\Ran(A^*)}$. 
 \end{lemma}
@@
     Now, suppose $u \in \clos{\Ran(A^*)}$ and let $\eps > 0$.
-Then there exists $v \in \XXX$ such that $\norm{A^*v - u} < \eps/2$.
+Then there exists $v \in \XX$ such that $\norm{A^*v - u} < \eps/2$.
 Writing $v = e + f$ with $e \in \Nul(A^*)$, $f \in \Nul(A^*)\P = \clos{\Ran(A)}$, we see that $\norm{A^*f - u} < \eps/2$.
 
-    Since $f \in \clos{\Ran(A)}$, there exists $x \in \XXX$ such that $\norm{Ax - f} < \eps/(2\norm{A})$. We now compute
+    Since $f \in \clos{\Ran(A)}$, there exists $x \in \XX$ such that $\norm{Ax - f} < \eps/(2\norm{A})$. We now compute
     \[
@@
     Au = f,
-\end{equation} $A \in \LLL(\XXX, \YYY)$, $f$ is known, and we wish to find $u$. 
+\end{equation} $A \in \LL(\XX, \YY)$, $f$ is known, and we wish to find $u$. 
 \begin{definition}
-    An element $u \in \XXX$ is called a \emph{least-squares solution} of \cref{eq:basic_ip} if $u$ is a minimiser of the function $v \mapsto \norm{Av - f}_\YYY$.
+    An element $u \in \XX$ is called a \emph{least-squares solution} of \cref{eq:basic_ip} if $u$ is a minimiser of the function $v \mapsto \norm{Av - f}_\YY$.
 It is called a \emph{minimal-norm solution} of \cref{eq:basic_ip} if it has minimal norm among all least-squares solutions.
@@
 \begin{theorem}
-    Let $f \in \YYY$ and $A \in \LLL(\XXX, \YYY)$. Then the following are equivalent:
+    Let $f \in \YY$ and $A \in \LL(\XX, \YY)$. Then the following are equivalent:
     \begin{enumerate}
-        \item $u \in \XXX$ satisfies $Au = P_{\clos{\Ran(A)} }f$;
+        \item $u \in \XX$ satisfies $Au = P_{\clos{\Ran(A)} }f$;
         \item $u$ is a least-squares solution of $\cref{eq:basic_ip}$:
@@
      \[
-     \norm{Au - f}_\YYY = \norm{ P_{\clos{\Ran(A)} }f - f} = \inf_{g \in \clos{\Ran(A)}} \norm{g - f} \leq \inf_{g \in \Ran(A)} \norm{g - f} = \inf_{u \in \XXX} \norm{Au - f}. 
+     \norm{Au - f}_\YY = \norm{ P_{\clos{\Ran(A)} }f - f} = \inf_{g \in \clos{\Ran(A)}} \norm{g - f} \leq \inf_{g \in \Ran(A)} \norm{g - f} = \inf_{u \in \XX} \norm{Au - f}. 
      \]
 
-     ``(2) $\implies$ (3)'':  Let $u \in \XXX$ be a least-squares solution and $v \in \XXX$ arbitrary. Define the quadratic polynomial 
+     ``(2) $\implies$ (3)'':  Let $u \in \XX$ be a least-squares solution and $v \in \XX$ arbitrary. Define the quadratic polynomial 
      \begin{align*}
@@
 
-    ``(3) $\implies$ (1)'':  From the normal equation we know that $A^*(f - Au) = 0$. For any $x \in \XXX$, we have
+    ``(3) $\implies$ (1)'':  From the normal equation we know that $A^*(f - Au) = 0$. For any $x \in \XX$, we have
     \[
@@
 
-    ``$\impliedby$'' Suppose $f = Au + g$ for some $u \in \XXX$, $g \in \Ran(A)\P = \clos{\Ran(A)}\P$. Then by the previous theorem, $Au = P_{\clos{\Ran(A)}}f$, so $u$ is a least-squares solution. 
+    ``$\impliedby$'' Suppose $f = Au + g$ for some $u \in \XX$, $g \in \Ran(A)\P = \clos{\Ran(A)}\P$. Then by the previous theorem, $Au = P_{\clos{\Ran(A)}}f$, so $u$ is a least-squares solution. 
 \end{proof}
@@
 \end{corollary}
-In particular, this holds if $\Ran(A)$ is finite-dimensional. Therefore, if either $\XXX$ or $\YYY$ is finite-dimensional, \cref{eq:basic_ip} has a least-squares solution for any $A$.
+In particular, this holds if $\Ran(A)$ is finite-dimensional. Therefore, if either $\XX$ or $\YY$ is finite-dimensional, \cref{eq:basic_ip} has a least-squares solution for any $A$.
 
@@
 \begin{definition}
-    Let $A \in \LLL(\XXX, \YYY)$, and define
+    Let $A \in \LL(\XX, \YY)$, and define
     \[
@@
 \begin{remark}
-    Note that $\clos{\Ran(A) \oplus \Ran(A)\P} = \clos{\Ran(A)} \oplus \Ran(A)\P = \clos{\Ran(A)} \oplus \clos{\Ran(A)}\P = \YYY$, and therefore the operator $\tilde A$ is \emph{densely defined}, and it is defined on all of $
-    \YYY$ if and only if $\Ran(A)$ is closed. 
+    Note that $\clos{\Ran(A) \oplus \Ran(A)\P} = \clos{\Ran(A)} \oplus \Ran(A)\P = \clos{\Ran(A)} \oplus \clos{\Ran(A)}\P = \YY$, and therefore the operator $\tilde A$ is \emph{densely defined}, and it is defined on all of $
+    \YY$ if and only if $\Ran(A)$ is closed. 
 \end{remark}
@@
 
-Conversely, if any linear operator $B \colon \YYY \to \XXX$ satisfies (1) and (2), then $B = A\D$.  
+Conversely, if any linear operator $B \colon \Ran(A) \oplus \Ran(A)\P \to \Nul(A)\P$ satisfies $BA = P_{\Nul(A)\P}$ and $AB = P_{\clos{\Ran(A)}}\restr_{\Dom(A\D)}$ then $B = A\D$.  
 \end{theorem}
@@
 \begin{proof}
-    We will not prove (1) and (2). Point (3) and (4) follow immediately from (1) and (2) respectively. 
+    We have 
+    \[
+    A\D A u = \tilde A^{-1} A P_{\Nul(A)\P} u = P_{\Nul(A)\P} u, 
+    \]
+    which proves (1). Furthermore, we have
+    \[
+    AA\D f = A \tilde A^{-1} P_{\clos{\Ran(A)}}f = P_{\clos{\Ran(A)}} f,
+    \]
+    which proves (2). Finally, (3) follows from (1) and (4) follows from (2).
+
+    Now, suppose $B$ satisfies (1) and (2). First we show that $B\restr_{\Ran(A)} = \tilde A^{-1}$, then we show that $B \restr_{\Ran(A)\P} = 0$. This shows that $B = A\D$. 
+    Let $f = Au \in \Ran(A)$ with $u \in \Nul(A)\P$, then
+    \[
+    Bf = BAu = P_{\Nul(A)\P} u = u = \tilde A^{-1} f, \quad\text{so } B\restr_{\Ran(A)} = \tilde A^{-1}. 
+    \]
+    Finally, let $f \in \Ran(A)\P$, then $ABf = P_{\clos{\Ran(A)}} f = 0$, and since $Bf \in \Nul(A)\P$ this implies $Bf = 0$.
+We conclude that $B\restr_{\Ran(A)\P} = 0$, and this concludes the proof.
 \end{proof}
@@
 \begin{definition}
-    Let $A \in \LLL(\XXX, \YYY)$.
-Then $A$ is called \emph{compact} if for any bounded $B \subseteq \XXX$, the image $A(B)$ is precompact in $\YYY$.
-The set of compact operators in $\LLL(\XXX, \YYY)$ is denoted $\KKK(\XXX, \YYY)$.
+    Let $A \in \LL(\XX, \YY)$.
+Then $A$ is called \emph{compact} if for any bounded $B \subseteq \XX$, the image $A(B)$ is precompact in $\YY$.
+The set of compact operators in $\LL(\XX, \YY)$ is denoted $\KK(\XX, \YY)$.
 \end{definition}
@@
 \begin{lemma}
-    Let $A \in \LLL(\XXX, \YYY)$. Then $A$ is compact if and only if, for every bounded sequence $(x_n) \subseteq X$, the sequence $(Ax_n) \subseteq Y$ has a convergent subsequence. 
+    Let $A \in \LL(\XX, \YY)$. Then $A$ is compact if and only if, for every bounded sequence $(x_n) \subseteq X$, the sequence $(Ax_n) \subseteq Y$ has a convergent subsequence. 
 \end{lemma}
@@
 \begin{theorem}
-    Let $A \in \KKK(\XXX, \YYY)$ with $\dim(\Ran(A)) = \infty$. Then $A\D$ is discontinuous. 
+    Let $A \in \KK(\XX, \YY)$ with $\dim(\Ran(A)) = \infty$. Then $A\D$ is discontinuous. 
 \end{theorem}
@@
 \begin{proof}
-    If $\dim\Ran(A) = \infty$, then $\XXX$ and $\Nul(A)\P$ are infinite-dimensional as well.
+    If $\dim\Ran(A) = \infty$, then $\XX$ and $\Nul(A)\P$ are infinite-dimensional as well.
 Chose an orthonormal sequence $(x_n) \subseteq \Nul(A)\P$, then after taking a subsequence if necessary, we can assume that $f_n \ceq Ax_n$ converges.
@@
 
-In particular, combining this with \cref{thm:mp_inverse_continuous} shows that the range of a compact operator is always open, and that not every element in $\YYY$ has a least-squares solution.
+In particular, combining this with \cref{thm:mp_inverse_continuous} shows that the range of a compact operator is always open, and that not every element in $\YY$ has a least-squares solution.
 
 We will need the following theorem, an infinite-dimensional analogue of the spectral theorem: 
-\begin{theorem}[Eigenvalue decomposition of self-adjoint compact operators] Let $\XXX$ be a Hilbert space, and $A \in \KKK(\XXX, \XXX)$ self-adjoint.
-Then there exists an orthonormal basis $(x_j)$ of $
-   \clos{\Ran(A)}$ and a sequence of eigenvalues $\abs{\lambda_1} \geq \abs{\lambda_2} \geq \dotsb > 0$ such that for all $u \in \XXX$ we have
+\begin{theorem}[Eigenvalue decomposition of self-adjoint compact operators] Let $\XX$ be a Hilbert space, and $A \in \KK(\XX, \XX)$ self-adjoint. Then there exists an orthonormal basis $(x_j)$ of $
+   \clos{\Ran(A)}$ and a sequence of eigenvalues $\abs{\lambda_1} \geq \abs{\lambda_2} \geq \dotsb > 0$ such that for all $u \in \XX$ we have
     \[
@@
 \begin{theorem}
-    Let $A \in \KKK(\XXX, \YYY)$.
+    Let $A \in \KK(\XX, \YY)$.
 Then there exists a (not necessarily infinite) sequence $\sigma_1 \geq \sigma_2 \geq \dotsb > 0$ converging to 0, and orthonormal bases $(x_j)$, $(y_j)$ of $\Nul(A)\P$ and $\clos{\Ran(A)}$ respectively, such that
@@
     \]
-    and such that for all $u \in \XXX$ and $f \in \YYY$ we have
+    and such that for all $u \in \XX$ and $f \in \YY$ we have
     \[
@@
 \begin{theorem}
-    Let $A \in \KKK(\XXX, \YYY)$ with SVD $\qty{(\sigma_j, x_j, y_j)}$ and let $f \in \Dom(A\D)$. Then
+    Let $A \in \KK(\XX, \YY)$ with SVD $\qty{(\sigma_j, x_j, y_j)}$ and let $f \in \Dom(A\D)$. Then
     \[
@@
 \begin{definition}
-    Let $A \in \KKK(\XXX, \YYY)$ have SVD $\qty{(\sigma_j, x_j, y_j)}$. 
-    We say that $f \in \YYY$ satisfies the \emph{Picard criterion} if 
+    Let $A \in \KK(\XX, \YY)$ have SVD $\qty{(\sigma_j, x_j, y_j)}$. 
+    We say that $f \in \YY$ satisfies the \emph{Picard criterion} if 
     \[
@@
 \begin{definition}
-    Let $A \in \KKK(\XXX, \YYY)$ have singular values $(\sigma_j)$.
+    Let $A \in \KK(\XX, \YY)$ have singular values $(\sigma_j)$.
 Then the ill-posed inverse problem $Au = f$ is called \emph{mildly ill-posed} if the $\sigma_j$ decay polynomially (i.e., $\frac{1}{\sigma_n} \leq C n^\gamma$ for some $C, \gamma$) and \emph{severely ill-posed} otherwise.

### 2020-10-19 modern_statistical_methods/C1.tex
updatd msm & ip notes

@@
 \begin{definition}
-    Let $\lambda \geq 0$, and let $\Bf 1 \in \RR^n$ be the all-ones vector. Then we define the \emph{Ridge regression} estimators
+    Let $\lambda \geq 0$, and let $\Bf 1 \in \RR^n$ be the all-ones vector. Then we define the \emph{ridge regression} estimators
     \[
@@
 Note that, even if $X$ does not have full column rank, this estimator exists for all $\lambda > 0$.
-In fact, for $\lambda$ sufficiently small, the Ridge estimator outperforms the MLE in terms of mean squared error:
+In fact, for $\lambda$ sufficiently small, the ridge estimator outperforms the MLE in terms of mean squared error:
 \begin{theorem}
@@
 
-Suppose $n \geq p$ and let $X = UDV\T$ be the thin SVD of our design matrix $X$. Then we can write the fitted values from the Ridge regression as follows: 
+Suppose $n \geq p$ and let $X = UDV\T$ be the thin SVD of our design matrix $X$. Then we can write the fitted values from the ridge regression as follows: 
 \begin{align*}
@@
 
-We see that Ridge regression shrinks $Y$ most in the smaller principal components of $X$. Therefore it will work well if most of the information is in the larger principal components of $X$.
+We see that ridge regression shrinks $Y$ most in the smaller principal components of $X$. Therefore it will work well if most of the information is in the larger principal components of $X$.
 
@@
 \subsection{Kernels}
-We will assume our inputs $x_1, \dotsc, x_n$ live in an abstract space $\XXX$. 
+We will assume our inputs $x_1, \dotsc, x_n$ live in an abstract space $\XX$. 
 \begin{definition}
-    A (\emph{positive-definite}) \emph{kernel} is a symmetric map $k \colon \XXX^2 \to \RR$ such that for all $n \in \NN$ and all $x_1, \dotsc, x_n \in \XXX$, the matrix $K \in \RR^{n \times n}$ with $K_{ij} = k(x_i, x_j)$ is positive semi-definite.
+    A (\emph{positive-definite}) \emph{kernel} is a symmetric map $k \colon \XX^2 \to \RR$ such that for all $n \in \NN$ and all $x_1, \dotsc, x_n \in \XX$, the matrix $K \in \RR^{n \times n}$ with $K_{ij} = k(x_i, x_j)$ is positive semi-definite.
 \end{definition}
@@
 \begin{proposition}[Cauchy-Schwarz for kernels] \label{prop:kernel_cs}
-    Let $k$ be a kernel and $x, x' \in \XXX$, then 
+    Let $k$ be a kernel and $x, x' \in \XX$, then 
     \[
@@
 In our old models, the data points $x_i$ were vectors in $\RR^p$.
-Now we try to think of them as points in an abstract space with an associated \emph{feature map} $\phi \colon \XXX \to \HHH$ (with $\HHH$ an inner product space), and a kernel $k(x, x')$ gives a measure of similarity between $\phi(x)$ and $\phi(x')$.
+Now we try to think of them as points in an abstract space with an associated \emph{feature map} $\phi \colon \XX \to \HH$ (with $\HH$ an inner product space), and a kernel $k(x, x')$ gives a measure of similarity between $\phi(x)$ and $\phi(x')$.
 In this case, we have the following:
 \begin{proposition} \label{prop:ip_kernel}
-    Let $\HHH$ be an inner product space,  $\phi \colon \XXX \to \HHH$ and define $k(x, x') \ceq \ang{\phi(x), \phi(x')}$. Then $k$ is a kernel. 
+    Let $\HH$ be an inner product space,  $\phi \colon \XX \to \HH$ and define $k(x, x') \ceq \ang{\phi(x), \phi(x')}$. Then $k$ is a kernel. 
 \end{proposition}
@@
 \begin{proof}
-    We have, for all $x_1, \dotsc, x_n \in \XXX$ and $\alpha \in \RR^n$ that
+    We have, for all $x_1, \dotsc, x_n \in \XX$ and $\alpha \in \RR^n$ that
     \[
@@
 
+\subsubsection{Examples of kernels}
 The following proposition shows how to make new kernels from old:
@@
     \begin{enumerate}
-        \item For $\XXX = \RR^p$ we have already seen the \emph{linear kernel} $k(x, x') = x\T x'$. 
-        \item For $\XXX = \RR^p$, the \emph{polynomial kernel} is defined as $k(x, x') = (1 + x\T x')^d$. This is a kernel since it is a power of a sum of two kernels. 
-        \item For $\XXX = \RR^p$, the \emph{Gaussian kernel} is defined by
+        \item For $\XX = \RR^p$ we have already seen the \emph{linear kernel} $k(x, x') = x\T x'$. 
+        \item For $\XX = \RR^p$, the \emph{polynomial kernel} is defined as $k(x, x') = (1 + x\T x')^d$. This is a kernel since it is a power of a sum of two kernels. 
+        \item For $\XX = \RR^p$, the \emph{Gaussian kernel} is defined by
         \[
@@
 
-        \item For $\XXX = [0, 1]$, define the \emph{Sobolev kernel} $k(x, x') = \min(x, x')$. The proof that this is a kernel is on example sheet 1. 
-        \item For $\XXX = \pow(\qty{1, \dotsc, p})$, define the \emph{Jaccard kernel}
+        \item For $\XX = [0, 1]$, define the \emph{Sobolev kernel} $k(x, x') = \min(x, x')$. The proof that this is a kernel is on example sheet 1. 
+        \item For $\XX = \pow(\qty{1, \dotsc, p})$, define the \emph{Jaccard kernel}
         \[
@@
 
-By \cref{prop:ip_kernel}, we see that every feature map $\phi \colon \XXX \to \HHH$ gives rise to a kernel.
+\subsubsection{Reproducing kernel Hilbert spaces}
+By \cref{prop:ip_kernel}, we see that every feature map $\phi \colon \XX \to \HH$ gives rise to a kernel.
 In the next (important!) theorem, we will see that every kernel is in fact induced by a feature map.
@@
 \begin{theorem} \label{thm:kernel_ip} 
-    Let $k$ be a kernel, then there exists an inner product space $\HHH$ and a feature map $\phi \colon \XXX \to \HHH$ such that 
+    Let $k$ be a kernel, then there exists an inner product space $\HH$ and a feature map $\phi \colon \XX \to \HH$ such that 
     \[
-    k(x, x') = \ang{\phi(x), \phi(x')} \quad\text{for all $x, x' \in \XXX$}. 
+    k(x, x') = \ang{\phi(x), \phi(x')} \quad\text{for all $x, x' \in \XX$}. 
     \]
@@
 \begin{proof}
-    We will construct $\HHH$ and $\phi$ explicitly. First we define the function space
+    We will construct $\HH$ and $\phi$ explicitly. First we define the function space
     \[
-    \HHH = \qty{ \sum_{i=1}^n \alpha_i k(\cdot, x_i) \mid n \in \NN,  \alpha_i \in \RR, x_i \in \XXX}. 
+    \HH = \qty{ \sum_{i=1}^n \alpha_i k(\cdot, x_i) \mid n \in \NN,  \alpha_i \in \RR, x_i \in \XX}. 
     \]
-    Let $f = \sum_{i=1}^n \alpha_i k(\cdot, x_i)$ and $g = \sum_{j=1}^n \beta_j k(\cdot, x_j')$, then the inner product on $\HHH$ is given by
+    Let $f = \sum_{i=1}^n \alpha_i k(\cdot, x_i)$ and $g = \sum_{j=1}^n \beta_j k(\cdot, x_j')$, then the inner product on $\HH$ is given by
     \[
@@
     \]
-    We define $\phi \colon \XXX \to \HHH$ as $\phi(x) = k(\cdot, x)$.
+    We define $\phi \colon \XX \to \HH$ as $\phi(x) = k(\cdot, x)$.
 
@@
 
-    Secondly, we must verify that the formula $k(x, x') = \ang{\phi(x), \phi(x')}$ indeed holds. For any $f \in \HHH, x \in \XXX$ we have
+    Secondly, we must verify that the formula $k(x, x') = \ang{\phi(x), \phi(x')}$ indeed holds. For any $f \in \HH, x \in \XX$ we have
     \begin{equation}
@@
     \end{equation}
-    i.e., evaluation of a function is a linear functional in $\HHH$.
+    i.e., evaluation of a function is a linear functional in $\HH$.
 
@@
 We must now only show that $f \neq 0 \implies \ang{f, f} > 0$.
-For this, note that $\ang{\cdot, \cdot}$ is a kernel on $\HHH$, so by \cref{prop:kernel_cs} (Cauchy-Schwarz) we have
+For this, note that $\ang{\cdot, \cdot}$ is a kernel on $\HH$, so by \cref{prop:kernel_cs} (Cauchy-Schwarz) we have
     \[
@@
 
-While $\HHH$ constructed in the proof is an inner product space, it is not necessarily a Hilbert space. Let $(f_n) \subseteq \HHH$ be Cauchy, then by Cauchy-Schwarz for kernels we find
+While $\HH$ constructed in the proof is an inner product space, it is not necessarily a Hilbert space. Let $(f_n) \subseteq \HH$ be Cauchy, then by Cauchy-Schwarz for kernels we find
 \[
@@
 \]
-We can do an analogous computation for $f_n - f_m$ to conclude that $\abs{f_m(x) - f_n(x)} \leq \sqrt{k(x, x)} \norm{f_n - f_m}$, and therefore, if $(f_n)$ is Cauchy, then it converges pointwise to some $f^* \colon \XXX \to \RR$.
+We can do an analogous computation for $f_n - f_m$ to conclude that $\abs{f_m(x) - f_n(x)} \leq \sqrt{k(x, x)} \norm{f_n - f_m}$, and therefore, if $(f_n)$ is Cauchy, then it converges pointwise to some $f^* \colon \XX \to \RR$.
 We will not prove the following theorem:
-\begin{theorem}
-    The inner product space $\HHH$ constructed in the proof of \cref{thm:kernel_ip} can be extended to a Hilbert space by adding all pointwise limits $f^*$ of Cauchy sequences $(f_n) \subseteq \HHH$. 
+\begin{theorem} \label{thm:kernel_ip_hilbert}
+    The inner product space $\HH$ constructed in the proof of \cref{thm:kernel_ip} can be extended to a Hilbert space by adding all pointwise limits $f^*$ of Cauchy sequences $(f_n) \subseteq \HH$. 
 \end{theorem}
 
-The completion of $\HHH$ is a special type of Hilbert space:
+The completion of $\HH$ is a special type of Hilbert space:
 \begin{definition}
-    A Hilbert space $\BBB$ of functions $f \colon \XXX \to \RR$ is called a \emph{reproducing kernel Hilbert space} (RKHS) if for all $x \in \XXX$, there exists $k_x \in \BBB$ such that 
+    A Hilbert space $\BB$ of functions $f \colon \XX \to \RR$ is called a \emph{reproducing kernel Hilbert space} (RKHS) if for all $x \in \XX$, there exists $k_x \in \BB$ such that 
     \[
@@
 
-    The function $k(x, x') = \ang{k_x, k_{x'}}$ is known as the \emph{reproducing kernel} of $\BBB$ (induced by the feature map $\phi(x) = k_x$). 
+    The function $k(x, x') = \ang{k_x, k_{x'}}$ is known as the \emph{reproducing kernel} of $\BB$ (induced by the feature map $\phi(x) = k_x$). 
 \end{definition}
 
-If we start with a kernel $k$, construct the corresponding RKHS $\BBB$, then it is easily checked that $k$ is indeed the reproducing kernel of $\BBB$.
+If we start with a kernel $k$, construct the corresponding RKHS $\BB$, then it is easily checked that $k$ is indeed the reproducing kernel of $\BB$.
 
@@
     \[
-    \HHH = \qty{x \mapsto \sum_{i=1}^n \alpha_i x\T x_i = x\T \qty(\sum_i \alpha_i x_i) \mid \alpha_i \in \RR, x_i \in \RR^p} = \qty{x \mapsto x\T \beta \mid \beta \in \RR^p},
+    \HH = \qty{x \mapsto \sum_{i=1}^n \alpha_i x\T x_i = x\T \qty(\sum_i \alpha_i x_i) \mid \alpha_i \in \RR, x_i \in \RR^p} = \qty{x \mapsto x\T \beta \mid \beta \in \RR^p},
     \]
@@
     \[
-    \ang{f, g} = k(\beta, \beta') = \beta\T \beta' \quad\text{so } \norm{f}_\HHH = \norm{\beta}_2. 
+    \ang{f, g} = k(\beta, \beta') = \beta\T \beta' \quad\text{so } \norm{f}_\HH = \norm{\beta}_2. 
     \]
 \end{example}
+
+\subsubsection{The representer theorem}
+We started with data $(Y_i, x_i)$ in a model where our fitted values depend only on $XX\T = K$, where $K$ is the kernel matrix derived from the kernel $k(x_i, x_j) = \ang{x_i, x_j} = x_i\T x_j$.
+Suppose we change the kernel to some $\tilde k$, and replace $K$ by $\tilde K$ accordingly.
+Then by \cref{thm:kernel_ip} this is equivalent by replacing our predictors $x_i$ by $\phi(x_i)$ for some feature map $\phi$.
+
+Recall that the ridge regression objective function is given by
+\[
+\min_{\beta \in \RR^p} \sum_{i=1}^n (Y_i - x_i\T \beta)^2 + \lambda\norm{\beta}^2. 
+\]
+Let $\HH$ be the RKHS of the linear kernel, then the above is equivalent to
+\begin{equation} \label{eq:objective_ridge_rkhs}
+\min_{f \in \HH} \sum_{i=1}^n (Y_i - f(x_i))^2 + \lambda \norm{f}^2. 
+\end{equation}
+
+We will show that in general, if we change our predictors $x_i$ to $\phi(x_i)$, and update the corresponding RKHS in \cref{eq:objective_ridge_rkhs}, then \cref{eq:objective_ridge_rkhs} solves our new problem.
+Note that there are no references to the feature map $\phi$ in \cref{eq:objective_ridge_rkhs}.
+
+\begin{theorem}[Representer theorem]
+Let:
+\begin{itemize}
+	\item $(Y_i, x_i)_{i=1}^n$ be our data with $x_i \in \XX$;
+	\item $c \colon \RR^n \times \XX^n \times \RR^n \to \RR$ be an arbitrary loss function;
+	\item $J \colon \RR_{\geq 0} \to \RR$ strictly increasing;
+	\item $k$ a kernel with RKHS $\HH$ and kernel matrix $K_{ij} = k(x_i, x_j)$. 
+\end{itemize}
+Then $\hat f \in \HH$ minimises
+\begin{equation*} % \label{eq:mini_q1}
+Q_1(f) \ceq c(Y, x_1, \dotsc, x_n, f(x_1), \dotsc, f(x_n))
+\end{equation*}
+if and only if $\hat f(\cdot) = \sum_{i=1}^n \hat\alpha_i k(\cdot, x_i)$, where $\hat\alpha \in \RR^n$ minimises
+\begin{equation*} % \label{eq:mini_q2}
+Q_2(\alpha) \ceq c(Y, x_1, \dotsc, x_n, K\alpha) + J(\alpha\T K\alpha). 
+\end{equation*}
+\end{theorem}
+
+\begin{proof}
+	Let $f \in \HH$ and $U = \Span\qty{k(\cdot, x_1), \dotsc, k(\cdot, x_n)} \subseteq \HH$, then $U$ is closed since it is finite-dimensional so we can write $f = u + v$ with $u \in U, v \in U\P$.
+Note that
+	\[
+	f(x_i) = \ang{k(\cdot, x_i), u+v} = \ang{k(\cdot, x_i), u} = u(x_i) \quad(i = 1, \dotsc, n),
+	\]
+	so minimisation of $Q_1$ is equivalent to minimisation of
+	\begin{equation} \label{eq:mini_proof_representer}
+	c(Y, x_1, \dotsc, x_n, u(x_1), \dotsc, u(x_n)) + J(\norm{u}^2 + \norm{v}^2)
+	\end{equation}
+	w.r.t.\ $u \in U$, $v \in U\P$. Now, since $J$ is strictly increasing, any minimiser $(u, v)$ of  \cref{eq:mini_proof_representer} will satisfy $v = 0$.
+
+	Write $u(\cdot) = \sum_{i=1}^n \alpha_i k(\cdot, x_i) \in U$, then clearly
+	\[
+	\mqty[ u(x_1) \\ \vdots \\ u(x_n)] = \mqty[ \sum_i \alpha_i k(x_1, x_i) \\ \vdots \\ \sum_i \alpha_i k(x_n, x_i)] = K\alpha \quad\text{and}\quad\norm{u}^2 = \sum_{i, j} \alpha_i \alpha_j k(x_i, x_j) = \alpha\T K \alpha.
+	\]
+
+	This shows that minimisation of \cref{eq:mini_proof_representer} is equivalent to minimisation of
+	\[
+	c(Y, x_1, \dotsc, x_n, K\alpha) + J(\alpha\T K\alpha) = Q_2(\alpha), 
+	\]
+	which completes the proof. 
+\end{proof}
+
+\begin{example}
+	In ridge regression, the representer theorem tells us that minimising \cref{eq:objective_ridge_rkhs} is equivalent to minimising $\norm{Y - K\alpha}^2 + \lambda \alpha\T K\alpha$, and indeed it can be shown (example sheet 1) that the minimiser satisfies $K\hat\alpha = K(K+\lambda I)^{-1} Y = X\hat\beta_\lambda^R$.
+Since $\HH$ may be infinite-dimensional, this gives a way to rewrite an infinite-dimensional optimization problem to a finite-dimensional optimisation problem.
+\end{example}
+
+\subsection{Kernel ridge regression}
+We have now defined kernel ridge regression and shown how the estimator may be computed, but we have yet to assess its predictive performance. We consider the model
+\[
+Y_i = f^0(x_i) + \eps_i, \quad \EE[\eps] = 0, \Var[\eps] = \sigma^2 I, 
+\]
+where we assume $f^0 \in \HH$ where $\HH$ is an RKHS with reproducing kernel $k$.
+By scaling the equation on both sides we may assume $\norm{f^0} \leq 1$ (note that this changes $\Var[\eps]$).
+Let $K$ be the kernel matrix with eigenvalues $d_1 \geq \dotsb \geq d_n \geq 0$, and let $\hat f = \hat f_\lambda$ be the estimated regression function, so
+\[
+\hat f_\lambda = \argmin_{f \in \HH} \qty(\sum_{i=1}^n (Y_i - f(x_i))^2 + \lambda \norm{f}^2). 
+\]
+
+\begin{theorem}
+	The mean squared prediction error (MSPE) satisfies
+	\begin{align*}
+		\frac1n \EE\qty[ \sum_{i=1}^n \qty(f^0(x_i) - \hat f_\lambda(x_i))^2] &\leq \frac{\sigma^2}{n} \sum_{i=1}^n \frac{d_i^2}{(d_i + \lambda)^2} + \frac\lambda{4n} \\
+		&\leq \frac{\sigma^2}{n \lambda} \sum_{i=1}^n \min\qty(\frac{d_i}{4}, \lambda) + \frac{\lambda}{4n}. 
+	\end{align*}
+\end{theorem}
+
+\begin{proof}
+	The representer theorem tells us that $(\hat f(x_1), \dotsc, \hat f(x_n))\T = K(K + \lambda I)^{-1} Y$.
+By projecting $f^0$ onto $\Span{k(\cdot, x_1), \dotsc, k(\cdot, x_n)}$ it is easily seen that there exists $\alpha \in \RR^n$ such that
+	\[
+	(f^0(x_1), \dotsc, f^0(x_n))\T = K\alpha \quad\text{and}\quad \norm{f^0}^2 \geq \alpha\T K\alpha. 
+	\]
+
+	Write $K = UDU\T$ where $D_{ii} = d_i$ and define $\theta \ceq U\T K\alpha = DU\T \alpha$ (note $U\theta = K\alpha$ and note that $d_i = 0 \implies \theta_i =0$). Then we have
+	\begin{align*}
+		 \EE\qty[ \sum_{i=1}^n \qty(f^0(x_i) - \hat f_\lambda(x_i))^2] &= \EE \norm{K\alpha - K(K + \lambda I)^{-1} Y}^2 \\
+		 &= \EE \norm{K(K + \lambda I)^{-1} (U\theta + \eps) - U\theta}^2 \\
+		 &= \EE\norm{DU\T (UDU\T + \lambda I)^{-1} (U\theta + \eps) - \theta}^2 \\
+		 &= \EE \norm{D(D + \lambda I)^{-1} (\theta + U\T \eps) - \theta}^2 \\
+		 &= \EE \norm{\qty[D(D + \lambda I)^{-1} - I]\theta + D(D + \lambda I)^{-1} U\T \eps}^2 \\
+		 &= \norm{\qty[D(D + \lambda I)^{-1} - I]\theta}^2 + \EE \norm{D(D+ \lambda I)^{-1} U\T\eps}^2. 
+	\end{align*}
+Note that the cross-term in the final equality disappears since it is the  expectation of a linear combination of $\eps$, which is 0.
+
+The first term is a deterministic quantity, which equals
+\[
+\norm{\qty[D(D + \lambda I)^{-1} - I]\theta}^2 = \sum_{i=1}^n \qty(\qty(\frac{d_i}{d_i + \lambda} - 1)\theta)^2 = \sum_{i=1}^n \frac{\lambda^2\theta_i^2}{(d_i + \lambda)^2}. 
+\]
+Let $D^+$ be the diagonal matrix with $D_{ii}^+ = d_i^{-1}$ if $d_i \neq 0$ and 0 else. Then we have
+\[
+\sum_{i: d_i > 0} \frac{\theta_i^2}{d_i} = \theta\T D^+\theta  = \alpha\T K U D^+ U\T K\alpha = \alpha\T UDD^+ DU\T \alpha = \alpha\T UDU\T \alpha = \alpha\T K\alpha \geq 1,
+\]
+Using this, we can bound the first term by
+\[
+\sum_{i=1}^n \frac{\lambda^2\theta_i^2}{(d_i + \lambda)^2} = \sum_{i: d_i \neq 0} \frac{\theta_i^2}{d_i} \frac{d_i\lambda^2}{(d_i + \lambda)^2} \leq \max_i \frac{d_i \lambda^2}{(d_i + \lambda)^2} \leq \max_i \frac{d_i\lambda^2}{4 \lambda d_i} \leq\frac\lambda4.
+\]
+
+To compute the second term, we use the \emph{trace trick}:
+\begin{align*}
+	\EE \norm{D(D + \lambda I)^{-1} U\T \eps}^2 &= \EE\qty[ \qty(D(D + \lambda I)^{-1} U\T \eps)\T \qty(D(D + \lambda I)^{-1} U\T \eps)] \\
+	&= \EE \qty[\tr\qty[D(D + \lambda I)^{-1} U\T \eps \eps\T U (D + \lambda I)^{-1} D]] \\
+	&= \tr\qty[ D(D + \lambda I)^{-1} U\T \EE(\eps\eps\T) U (D + \lambda I)^{-1} D] \\
+	&= \sigma^2 \tr \qty[ D^2 (D + \lambda I)^{-2}] = \sigma^2 \sum_{i=1}^n \frac{d_i^2}{(d_i + \lambda)^2}.
+\end{align*}
+
+This gives our first inequality $MSPE \leq \frac{\sigma^2}{n} \sum_{i=1}^n \frac{d_i^2}{(d_i + \lambda)^2} + \frac\lambda{4n}$.
+For the second inequality, note that $\frac{d_i^2}{(d_i + \lambda)^2} \leq 1$ and also $\frac{d_i^2}{(d_i + \lambda)^2} \leq \frac{d_i^2}{4 d_i\lambda} = \frac{d_i}{4\lambda}$.
+Therefore, we can write $\frac{d_i^2}{(d_i + \lambda)^2} \leq \frac1\lambda \min(\lambda, \frac{d_i}{4})$ which proves the second inequality.
+\end{proof}

### 2020-10-20 modern_statistical_methods/msm_ES1.tex
updated notes and worked on msm exercises

@@
 \newcommand{\EE}{\Bb E}
+
+\newcommand\XX{\Cal X}
 
@@
 \end{enumerate}
+\end{proof}
 
+\begin{question}
+	Let $\XX \ceq \qty{x \in \RR^d : \norm{x} < 1}$. Show that $k(x, x') \ceq (1 - x\T x')^{-\alpha}$ defined on $\XX \times \XX$, where $\alpha > 0$, is a kernel.
+\end{question}
+
+\begin{proof}
+	Note by Cauchy-Schwarz we have for $x, x' \in \XX$ that $\abs{x\T x'} \leq \norm{x} \norm{x'} < 1$, so $k$ is well-defined.
+
+	Symmetry of $k$ is clear. Define the function 
+	\[
+	\phi \colon (-\infty, 1) \to \RR \colon x \mapsto (1-x)^{-\alpha}. 
+	\]
+	Then $\phi$ is continuous, and we can compute its Taylor series around 0. It is computed that 
+	\[
+	\phi^{(k)}(x) = (1-x)^{-\alpha - k} \prod_{j=0}^{k-1}(\alpha + j) \implies \phi^{(k)}(0) = \prod_{j=0}^{k-1} (\alpha + j). 	\]
+	Therefore, the Taylor series of $\phi$ is given by
+	\[
+	\sum_{k=0}^\infty \qty(\prod_{j=0}^{k-1} (\alpha + j)) \frac{x^k}{k!}, 
+	\]
+	and its radius of convergence is given by
+	\[
+	\lim_{k\to\infty} \frac{\frac{1}{k!}\prod_{j=0}^{k-1} (\alpha + j)}{\frac{1}{(k+1)!} \prod_{j=0}^k (\alpha + j)} = \lim_{k\to\infty} \frac{k}{\alpha + k} = 1,
+	\]
+	so $\phi$ agrees with its Taylor series on $(-1, 1)$. By Cauchy-Schwarz, for $x, x' \in \XX$ we have $x_i\T x_j \leq \norm{x_i} \norm{x_j} < 1$, so we can write for $x, x' \in \XX$
+	\[
+	k(x, x') = \phi(x\T x') = \sum_{k=0}^\infty \qty(\prod_{j=0}^{k-1} (\alpha + j)) \frac{(x\T x')^k}{k!}.
+	\]
+	By proposition 4, every term in the sum is a kernel, and therefore the limit is a kernel as well. 
 \end{proof}
+
+\begin{question}
+	Suppose we have a matrix of predictors $X \in \RR^{n \times p}$ where $p \gg n$. Explain how to obtain the fitted values of the following ridge regression using the kernel trick:
+	\begin{gather*}
+		\text{Minimise over $\beta \in \RR^p$, $\theta \in \RR^{p(p-1)/2}$, $\gamma \in \RR^p$,} \\
+		\sum_{i=1}^n \qty(Y_i - \sum_{k=1}^p X_{ik}\beta_k - \sum_{k=1}^p \sum_{j=1}^{k-1} X_{ik} X_{ij} \theta_{jk} - \sum_{k=1}^p X_{ik}^2\gamma_k)^2 + \lambda_1 \norm{\beta}^2 + \lambda_2 \norm{\theta}^2 + \lambda_3\norm{\gamma}^2.
+	\end{gather*}
+Note that we have indexed $\theta$ with two numbers for convenience. 
+\end{question}
+
+\begin{proof}
+	Our linear model has predictors
+	\[
+	\qty{X_{ik} \mid 1 \leq k \leq p} \cup \qty{X_{ij} X_{ik} \mid 1 \leq j < k \leq p} \cup \qty{X_{ik}^2 \mid 1 \leq k \leq p}
+	\]
+	for $i = 1, \dotsc, k$.
+We want to assume that $\lambda_1 = \lambda_2 = \lambda_3$: to this end, define $\xi = \frac{\lambda_1}{\lambda_2}$ and $\eta = \frac{\lambda_3}{\lambda_1}$, and replace $\theta$ by $\xi\theta$ and $\gamma$ by $\eta\gamma$.
+This means we will have to replace our predictors by
+	\[
+	Z_i \ceq 
+	(X_{ik} \mid 1 \leq k \leq p)\T \cup (\xi X_{ij} X_{ik} \mid 1 \leq j < k \leq p)\T \cup (\gamma X_{ik}^2 \mid 1 \leq k \leq p)\T
+	\]
+	We use the kernel trick: note that
+	\begin{align*}
+	\ang{Z_i, Z_j} &= \sum_k X_{ik}X_{jk} + \xi^2 \sum_{k < \ell} X_{ik} X_{i\ell} X_{jk} X_{j\ell} + \gamma^2 \sum_k X_{ik}^2 X_{jk}^2 \\
+	&= \qty(\frac{1}{\sqrt 2\xi} + \frac{\xi}{\sqrt 2}\sum_k X_{ik}X_{jk})^2 + (\gamma^2 - \frac{\xi^2}{2})\sum_k X_{ik}^2 X_{jk}^2 - \frac{1}{2\xi^2}.
+	\end{align*}
+	We have therefore rewritten $\ang{Z_i, Z_j}$ into something that can be computed in $O(p)$, which is exactly the aim of the kernel trick.
+\end{proof}
+
+\begin{question}
+	Let $\hat\alpha$ be a minimiser of $\norm{Y - K\alpha}^2 + \lambda \alpha\T K\alpha$, with $K$ being a kernel matrix as usual. Show that $K\hat\alpha = K(K + \lambda I)^{-1} Y$. 
+\end{question}
+
+\begin{solution}
+	Suppose $\hat\alpha$ minimises $Q(\alpha) \ceq \norm{Y - K\alpha}^2 + \lambda \alpha\T K\alpha$. Then we have
+	\begin{align*}
+		0 = \nabla_\alpha Q(\hat\alpha) &= -2 K (Y - K\hat\alpha) + 2\lambda K\hat\alpha \\
+		\lambda K\hat\alpha &= KY - K^2\hat\alpha \\
+		(K + \lambda) K\hat\alpha &= KY \\
+		K\hat\alpha &= (K + \lambda)^{-1} KY.
+	\end{align*}
+Since $K$ and $(K + \lambda)^{-1}$ are simultaneously diagonalisable, they commute, and therefore $K\hat\alpha = K(K + \lambda)^{-1} Y$. 
+%\begin{align*}
+%(K + \lambda)^{-1} KY &= (U(D + \lambda) U\T)^{-1} UDU\T Y \\
+%&= U(D + \lambda)^{-1} D U\T Y \\
+%&= UD (D+ \lambda)^{-1} U\T Y\\
+%&= UDU\T (U (D+ \lambda) U\T) Y\\
+%&= K(K + \lambda)^{-1} Y,
+%\end{align*}
+%which proves the claim.
+\end{solution}
+
+\begin{question}
+	...
+\end{question}
+
+\begin{proof}
+	content...
+\end{proof}
+
+\begin{question}
+	...
+\end{question}
+
+\begin{proof}
+	...
+\end{proof}
+
+\begin{question}
+	Show from first principles that the Sobolev kernel is indeed a (positive definite) kernel.
+\end{question}
+
+\begin{solution}
+	Let $\XX = [0, 1]$ and $k(x, x') = \min(x, x')$ the Sobolev kernel. We must then show that for any $x_1, \dotsc, x_n \in \XX$, the matrix $K_{ij} = \min(x_i, x_j)$ is positive semi-definite. 
+\end{solution}
 \end{document}

### 2020-10-20 topics_statistical_theory/C1.tex
updated notes and worked on msm exercises

@@
         \]
-        since the infimum of $t^2\sigma^2/2 - tx$ is attained at $t = x/\sigma^2$.
+        since the infimum of $t^2\sigma^2/2 - tx$ is attained at $t = x/\sigma^2$ (this method is called \emph{Chernoff bounding}).
 
@@
 
-\begin{theorem}[Hoeffding's inequality]
+\begin{theorem}[Hoeffding's inequality] \label{thm:hoeffding}
 	Let $X_1, \dotsc, X_n$ be independent sub-Gaussian random variables, with $X_i$ having parameter $\sigma_i^2$.
@@
 \end{proof}
+
+\begin{corollary}
+	Let $X_1, \dotsc, X_n$ be independent random variables where $\EE[X_i] = \mu_i$ and $a_i \leq X_i \leq b_i$. Then we have
+	\[
+	\PP(\bar X - \bar \mu \geq x) \leq \exp\qty(-\frac{2n^2x^2}{\sum_{i=1}^n (b_i - a_i)^2}). 
+	\]
+\end{corollary}
+
+\begin{proof}
+	By Hoeffding's lemma, $X_i - \mu_i$ is sub-Gaussian with parameter $(b_i - a_i)^2/4$ for each $i$. The result now follows from \cref{thm:hoeffding}.
+\end{proof}
+
+Note that when $X$ takes values in $[a, b]$, its variance is at most $(b-a)^2$.
+However, when $\Var(X_i) \ll (b_i - a_i)^2$, Hoeffding's inequality can be loose (for example, when $X_i \sim \Rm{Bern}(p_i)$ with $p_i$ small).
+In such circumstances, Bennett's or Bernstein's inequality may give better results.
+
+\begin{theorem}[Bennett's inequality]
+	Let $X_1, \dotsc, X_n$ be independent random variables with $\EE[X_i] = 0$, $\sigma_i^2 \ceq \Var(X_i) < \infty$, and $X_i \leq b$ for some $b > 0$.
+Define  $S \ceq \sum_{i=1}^n X_i$, $\nu \ceq \nu$ and $\phi \colon \RR \to \RR$ by $\phi(u) \ceq e^u - 1 - u = \sum_{k=2}^\infty \frac{u^k}{k!}$, then for every $t > 0$ we have
+	\[
+	\log\EE[e^{tS}] \leq \frac{n \nu }{b^2} \phi(bt). 
+	\]
+
+	Defining $h \colon (0, \infty) \to [0, \infty)$ by $h(u) \ceq (1+u)\log(1+u) - u$, we have for every $x > 0$ that
+	\[
+	\PP(\bar X \geq x) \leq \exp\qty(-\frac{n \nu}{b^2} h\qty(\frac{bx}{\nu})). 
+	\]
+\end{theorem}
+
+\begin{proof}
+	Define $g \colon \RR \to \RR$ by 
+	\[
+	g(u) \ceq \sum_{k=0}^\infty \frac{u^k}{(k+2)!} = \begin{cases}
+		\frac{\phi(u)}{u^2}&\text{if $u \neq 0$}, \\
+		\frac12 &\text{if $u = 0$}. 
+	\end{cases} 
+	\]
+	Then one can check that $g$ is increasing on $\RR$, so 
+	\[
+	e^{tX_i} - 1 - t X_i = t^2 X_i^2 g(tX_i) \leq t^2 X_i^2 g(tb) = X_i^2 \frac{\phi(bt)}{b^2},
+	\]
+	and therefore
+	\[
+	e^{tX_i} \leq 1 + tX_i + X_i^2 \frac{\phi(bt)}{b^2} \implies \EE[e^{tX_i}] \leq 1 + \EE[X_i^2] \frac{\phi(bt)}{t^2} = 1 + \Var(X_i) \frac{\phi(bt)}{b^2}. 
+	\]
+
+	Hence for $t > 0$ we have
+	\begin{align*}
+		\log \EE[e^{tS}] &= \sum_{i=1}^n \log\EE[e^{tX_i}] \leq n \cdot \frac1n\sum_{i=1}^n \log\qty(1 + \Var(X_i) \frac{\phi(bt)}{b^2}) \\
+		&\overset*\leq n \log(1 + \frac{\nu \phi(bt)}{b^2}) \overset{**}\leq  \frac{n\nu}{b^2}\phi(bt).
+	\end{align*}
+Here, $(*)$ follows from the fact that $\log$ is a concave function while $(**)$ follows from the fact that $\log(1+u) \leq u$ for all $u \geq 0$.
+This concludes the proof for the first part of the theorem.
+
+Now, we apply the method of Chernoff bounding and find
+\[
+\PP(\bar X \geq x) = \PP(S \geq nx) \leq \inf_{t > 0} e^{-ntx} \EE[e^{t S}] \leq \inf_{t > 0} e^{-ntx + n\nu \phi(bt)/b^2} = \exp\qty(-\frac{n\nu}{b^2} h\qty(\frac{bx}{\nu})),
+\]
+since once can check that the infimum is attained at $t = b^{-1} \log(1 + bx/\nu)$. 
+\end{proof}
+
+\begin{definition}
+	A random variable $X$ with $\EE X = 0$ is called \emph{sub-Gamma in the right tail} with variance factor $\sigma^2 > 0$ and scale $c > 0$ if
+	\[
+	\EE[e^{tX}] \leq \exp\qty(\frac{\sigma^2 t^2}{2(1 - ct)})
+	\]
+	for all $t \in [0, 1/c)$. 
+\end{definition}
+
+Note that this definition looks like that of sub-Gaussianity, except that $e^{tX}$ can explode as $t$ approaches $1/c$. We give some characteristics of sub-Gamma distributions:
+
+\begin{definition}
+	For any $x \in \RR$ we define $x_+ \ceq \max(x, 0)$. 
+\end{definition}
+
+\begin{proposition}
+\begin{enumerate}[(a)]
+\item Let $X$ be sub-Gamma in the right tail with variance factor $\sigma^2$ and scale $c$. Then 
+\[
+\PP(X \geq x) \leq \exp\qty(-\frac{x^2}{2(\sigma^2 + cx)})
+\]
+for all $x \geq 0$.
+
+\item Let $X$ be a random variable with $\EE X = 0$, $\EE[X^2] \leq \sigma^2$ and $\EE[(X_+)^q] \leq q!
+\sigma^2 c^{q-2}/2$ for all $q \geq 3$.
+Then $X$ is sub-Gamma in the right tail with variance factor $\sigma^2$ and scale parameter $c$.
+\end{enumerate}
+\end{proposition}
+
+\begin{proof}
+\begin{enumerate}[(a)]
+\item Again, we apply a Chernoff bound: we have
+\begin{align*}
+\PP(X \geq x) \leq \inf_{t \in [0, 1/c)} e^{-tx} \EE[e^{tX}] &\leq \inf_{t \in [0, 1/c)} \exp(-tx + \frac{\sigma^2t^2}{2(1 - ct)}) \\
+&\leq \exp(- \frac{x^2}{2(\sigma^2 + cx)}),
+\end{align*}
+where we have set $t = x/(\sigma^2 + cx) \in [0, 1/c)$ in the final step.
+
+\item Recall from the proof of Bennett's inequality that $g$ is increasing and therefore for $u \leq 0$ we have $\phi(u) = u^2 g(u) \leq u^2 g(0) = \frac{u^2}{2}$.
+Therefore, for every $u \in \RR$ we have
+\[
+\phi(u) \leq \frac{u^2}{2} + \sum_{q=3}^\infty \frac{(u_+)^q}{q!}.
+\]
+We deduce that for $t \in [0, 1/c)$ we have (note $\log(x) \leq x - 1$ for all $x$):
+\[
+\log \EE[e^{tX}] \leq \EE(e^{tX}) - 1 = \EE[\phi(tX)] \leq \EE \qty[\frac{t^2X^2}{2} + \sum_{q=3}^\infty \frac{t^q X_+^q}{q!}].
+\]
+By Fubini's theorem, since the infinite sum has only positive terms we may interchange sum and expectation to obtain
+\[
+\EE \qty[\frac{t^2X^2}{2} + \sum_{q=3}^\infty \frac{t^q \EE[X_+^q]}{q!}] =  \frac{t^2 \Var[X]}{2} + \sum_{q=3}^\infty \frac{t^q \EE[X_+^q]}{q!} \leq \frac{\sigma^2 t^2}{2} \sum_{q=2}^\infty t^{q-2} c^{q-2} = \frac{\sigma^2t^2}{2(1-ct)}.
+\]
+\end{enumerate}
+\end{proof}
+
+Following this proposition, we can prove Bernstein's inequality:
+\begin{theorem}[Bernstein's inequality]
+	Let $X_1, \dotsc, X_n$ be independent random variables with $\EE[X] = 0$, $\frac1n\sum_{i=1}^n \Var(X_i) \leq \sigma^2$ and $\frac1n \sum_{i=1}^n \EE[(X_i)_+^q] \leq q!
+\sigma^2 c^{q-2}/2$ some $\sigma, c > 0$ and for all $q \geq 3$.
+Then $S \ceq \sum_{i=1}^n X_i$ is sub-Gamma in the right tail with variance factor $n\sigma^2$ and scale parameter $c$.
+In particular we have
+	\[
+	\PP(\bar X \geq x) \leq \exp\qty(- \frac{nx^2}{2(\sigma^2 + cx)}). 
+	\]
+\end{theorem}
+
+\begin{proof}
+	We have by part (b) of the previous proposition
+	\[
+	\log\EE[e^{tS}] = \sum_{i=1}^n \log \EE[e^{tX_i}] \leq n \frac{\sigma^2 t^2}{2(1 - ct)},
+	\]
+	and the second claim follows from part (a) of the previous proposition:
+	\[
+	\PP(\bar X \geq x) = \PP(S \geq nx) \leq \exp\qty(- \frac{nx^2}{2(\sigma^2 + cx)}). 
+	\]
+\end{proof}

### 2020-10-20 inverse_problems/C1.tex
worked on msm exercise sheet and almost finished inverse problems notes up to lecture 6

@@
         \[
-        \norm{A}_{\LL(\XX, \YY)} \ceq \sup_{u \neq 0} \frac{\norm{Au}_\YY}{\norm{u}_\XX} = \sup_{\norm{u}_\XX \leq 1} < \infty. 
+        \norm{A}_{\BB(\XX, \YY)} \ceq \sup_{u \neq 0} \frac{\norm{Au}_\YY}{\norm{u}_\XX} = \sup_{\norm{u}_\XX \leq 1} < \infty. 
         \]
-            It is known that a linear operator between normed spaces is continuous if and only if it is bounded. The set of bounded linear operators from $\XX$ to $\YY$ is denotes $\LL(\XX, \YY)$. 
+            It is known that a linear operator between normed spaces is continuous if and only if it is bounded. The set of bounded linear operators from $\XX$ to $\YY$ is denotes $\BB(\XX, \YY)$. 
             \item 
@@
 
-        \item     We will assume $\XX$ and $\YY$ are Hilbert spaces, so there is an inner product $\ang{\cdot, \cdot}$ and any bounded operator $A \in \LL(\XX, \YY)$ has a unique adjoint $A^* \in \LL(\YY, \XX)$ which satisfies
+        \item     We will assume $\XX$ and $\YY$ are Hilbert spaces, so there is an inner product $\ang{\cdot, \cdot}$ and any bounded operator $A \in \BB(\XX, \YY)$ has a unique adjoint $A^* \in \BB(\YY, \XX)$ which satisfies
         \[
@@
 
-    \item For any $A \in \LL(\XX, \YY)$ we have
+    \item For any $A \in \BB(\XX, \YY)$ we have
     \[
@@
 \begin{lemma}
-    For any $A \in \LL(\XX, \YY)$ we have $\clos{\Ran(A^*A)} = \clos{\Ran(A^*)}$. 
+    For any $A \in \BB(\XX, \YY)$ we have $\clos{\Ran(A^*A)} = \clos{\Ran(A^*)}$. 
 \end{lemma}
@@
     Au = f,
-\end{equation} $A \in \LL(\XX, \YY)$, $f$ is known, and we wish to find $u$. 
+\end{equation} where $A \in \BB(\XX, \YY)$ and $f$ are known, and we wish to find $u$. 
 \begin{definition}
@@
 \begin{theorem}
-    Let $f \in \YY$ and $A \in \LL(\XX, \YY)$. Then the following are equivalent:
+    Let $f \in \YY$ and $A \in \BB(\XX, \YY)$. Then the following are equivalent:
     \begin{enumerate}
@@
 \begin{definition}
-    Let $A \in \LL(\XX, \YY)$, and define
+    Let $A \in \BB(\XX, \YY)$, and define
     \[
@@
 \begin{definition}
-    Let $A \in \LL(\XX, \YY)$.
+    Let $A \in \BB(\XX, \YY)$.
 Then $A$ is called \emph{compact} if for any bounded $B \subseteq \XX$, the image $A(B)$ is precompact in $\YY$.
-The set of compact operators in $\LL(\XX, \YY)$ is denoted $\KK(\XX, \YY)$.
+The set of compact operators in $\BB(\XX, \YY)$ is denoted $\KK(\XX, \YY)$.
 \end{definition}
@@
 \begin{lemma}
-    Let $A \in \LL(\XX, \YY)$. Then $A$ is compact if and only if, for every bounded sequence $(x_n) \subseteq X$, the sequence $(Ax_n) \subseteq Y$ has a convergent subsequence. 
+    Let $A \in \BB(\XX, \YY)$. Then $A$ is compact if and only if, for every bounded sequence $(x_n) \subseteq X$, the sequence $(Ax_n) \subseteq Y$ has a convergent subsequence. 
 \end{lemma}

### 2020-10-20 inverse_problems/C2.tex
updatd msm & ip notes
worked on msm exercise sheet and almost finished inverse problems notes up to lecture 6

@@
 \section{Classical regularisation theory}
+Let $A \in \BB(\XX, \YY)$ such that $\Ran(A)$ is not closed (this happens for example when $A$ is compact and does not have finite rank), and consider the inverse problem $Au = f$.
+Suppose we measure not $f$, but noisy data $f_\delta$ such that $\norm{f_\delta - f} \leq \delta$.
+Then since $A\D$ is discontinuous, we cannot expect that $A\D f_\delta \to A\D f$ as $\delta \to 0$.
+Therefore, we must replace $A\D$ by operators that approximate it.
+
 \begin{definition}
-    Let $A \in \LLL(\XXX, \YYY)$ and 
+	Let $A\in \BB(\XX, \YY)$. A family $(R_\alpha)_{\alpha > 0}$ of continuous operators is called a \emph{regularisation} of $A\D$ if
+	\[
+	\lim_{\alpha \to 0} R_\alpha f = A\D f \quad\text{for all $f \in \Dom(A\D)$}. 
+	\]
+	If all $R_\alpha$ are linear, then we speak of a \emph{linear regularisation} of $A\D$. 
 \end{definition}
+
+\begin{theorem}[Banach-Steinhaus]
+	Let $\XX, \YY$ be Hilbert spaces and $\qty{A_\alpha} \subseteq \BB(\XX, \YY)$ a family of pointwise bounded operators. Then $\qty{A_\alpha}$ is bounded in norm. 
+\end{theorem}
+
+\begin{corollary} \label{cor:bounded_converge}
+	Let $\XX, \YY$ be Hilbert spaces and $(A_j) \subseteq \BB(\XX, \YY)$.
+Then $(A_j)$ converges pointwise to some $A \in \BB(\XX, \YY)$ if and only if $\qty{A_j}$ is norm-bounded and converges pointwise on some dense subset $\XX' \subseteq \XX$.
+\end{corollary}
+
+\begin{theorem}
+	Let $\XX, \YY$ be Hilbert spaces, $A \in \BB(\XX, \YY)$ and $(R_\alpha)_{\alpha > 0}$ a linear regularisation.
+If $A\D$ is not continuous, $(R_\alpha)$ is not norm-bounded.
+In particular, there exists $f \in \YY$ with $\norm{R_\alpha f} \to \infty$.
+\end{theorem}
+
+\begin{proof}
+	Suppose $(R_\alpha)$ is norm-bounded.
+Let $\alpha_j \to 0$, then we know that $R_{\alpha_j} \to A\D$ pointwise on $\Dom(A\D)$.
+Since $\Dom(A\D)$ is dense in $\YY$, \cref{cor:bounded_converge} then  tells us that $A\D$ is bounded and therefore continuous, a contradiction.
+
+	By the Banach-Steinhaus theorem, if $(R_\alpha)$ is not norm-bounded, it is not pointwise bounded, so there must exist $f \in \YY$ such that $\qty{\norm{R_\alpha f}}$ is not bounded. 
+\end{proof}
+
+\begin{recap}
+	Recall that any bounded sequence in a Hilbert space has a weakly convergent subsequence. 
+\end{recap}
+\begin{theorem}
+	Let $A \in \BB(\XX, \YY)$ and $(R_\alpha)$ a linear regularisation of $A\D$.
+If $\qty{\norm{A R_\alpha}}_{\alpha > 0}$ is bounded, then $\norm{R_\alpha f} \to \infty$ as $\alpha \to 0$ for every $f \notin \Dom(A\D)$.
+\end{theorem}
+
+\begin{proof}
+	Define $u_\alpha \ceq R_\alpha f$ for $f \notin \Dom(A\D)$, and assume there exists a sequence $\alpha_k \to 0$ such that $\qty{\norm{u_{\alpha_k}}}$ is bounded.
+After taking a subsequence if necessary, we may assume that $u_{\alpha_k} \rightharpoonup u$ for some $u \in \XX$, and therefore we also have $Au_{\alpha_k} \rightharpoonup Au$.
+
+	We also have $\lim_{\alpha \to 0} A R_\alpha f = AA\D f = P_{\clos{\Ran(A)}}f$ for $f \in \Dom(A\D)$, and since we assumed $\qty{A R_\alpha}$ was norm-bounded, by \cref{cor:bounded_converge} we have $\lim_{\alpha \to 0} AR_\alpha f = P_{\clos{\Ran(A)}} f$ for all $f \in \YY$.
+
+	Since $Au_{\alpha_k}$ is convergent and has weak limit $Au$, it must also have limit $Au$, so we find $Au = P_{\clos{\Ran(A)}} f$ so $f \in \Dom(A\D)$, a contradiction. 
+\end{proof}
+
+We need some process to choose a parameter. To this end, note that we have
+\[
+\norm{R_\alpha f_\delta - A\D f} \leq \norm{R_\alpha (f_\delta - f)} + \norm{(R_\alpha - A\D) f} \\
+\leq  \norm{R_\alpha} \norm{f_\delta - f} + \norm{f} \norm{R_\alpha - A\D}. 
+\]
+The first term is called the \emph{data error} and is unbounded for $\alpha \to 0$, and the second term is called the \emph{approximation error} which does vanish for $\alpha \to 0$.
+Therefore, we want to choose $\alpha$ small enough to have a low approximation error, while keeping the data error at bay.
+We define this formally:
+
+\begin{definition}
+	A function $\alpha \colon  \RR_{> 0} \times \YY \to \RR_{> 0} \colon (\delta, f_\delta) \mapsto \alpha(\delta, f_\delta)$ is called a \emph{parameter choice rule} (PCR). We distinguish three types:
+	\begin{enumerate}
+		\item An \emph{a priori} PCR depends only on $\delta$;
+		\item An \emph{a posteriori} PCR depends on both $\delta$ and $f_\delta$;
+		\item A \emph{heuristic} PCR depends only on $f_\delta$. 
+	\end{enumerate}
+\end{definition}

### 2020-10-20 modern_statistical_methods/msm_ES1.tex
worked on msm exercise sheet and almost finished inverse problems notes up to lecture 6

@@
 \begin{solution}
-	Let $\XX = [0, 1]$ and $k(x, x') = \min(x, x')$ the Sobolev kernel. We must then show that for any $x_1, \dotsc, x_n \in \XX$, the matrix $K_{ij} = \min(x_i, x_j)$ is positive semi-definite. 
+	Let $\XX = [0, 1]$ and $k(x, x') = \min(x, x')$ the Sobolev kernel.
+We must then show that for any $x_1, \dotsc, x_n \in \XX$, the matrix $K_{ij} = \min(x_i, x_j)$ is positive semi-definite (symmetry of $k$ is clear).
+
+	To this end, assume $x_1 \leq \dotsb \leq x_n$ (this may require relabeling of the $x_i$ which does not affect positive semi-definiteness of $K$).
+Now let $\vec f_i$ denote the vector $f_{i1}, \dotsc, f_{i(i-1)} = 0$, $f_{ii}, \dotsc, f_{in} = 1$.
+	\[
+	K = \mqty[ x_1 & x_1 & x_1 & \cdots & x_1 \\
+	x_1 & x_2 & x_2 & \dotsb & x_2 \\
+	x_1 & x_2 & x_3 & \dotsb & x_3 \\
+	\vdots & \vdots & \vdots & \ddots & \vdots \\
+	x_1 & x_2 & x_3 & \cdots & x_n] = x_1 \vec f_1 \vec f_1\T + (x_2 - x_1) \vec f_2 \vec f_2\T + \dotsb + (x_n - x_{n-1}) \vec f_n \vec f_n\T. 
+	\]
+	Since this is a nonnegative combination of positive semidefinite matrices, we conclude that $K$ is positive semidefinite. 
 \end{solution}

### 2020-10-20 inverse_problems/inverse_ES1.tex
started IP ES1

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[margin=3cm]{geometry}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\usepackage{mathtools, amsmath, amssymb, amsthm, enumerate, mdframed, bbm, graphicx, float, physics, xcolor, cleveref}

\hypersetup{
    colorlinks   = true, %Colours links instead of ugly boxes
    urlcolor     = blue, %Colour for external hyperlinks
    linkcolor    = blue, %Colour of internal links
    citecolor   = red %Colour of citations
}

% Definition of numbered environments.
% Usage: \begin{theorem} ... \end{theorem}
% Remark has no numbering.
\theoremstyle{plain}
\newtheorem{question}{Question}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

% Question with box around it
\newenvironment{qbox}{\begin{mdframed}\begin{question}}{\end{question}\end{mdframed}}

% A proof with "solution" instead of "proof" and no QED symbol
\newenvironment{solution}{\begin{proof}[Solution]\renewcommand\qedsymbol{}}{\end{proof}}

% Some renewed commands
\renewcommand{\vec}{\mathbf}
\renewcommand{\emptyset}{\varnothing}
\renewcommand{\epsilon}{\varepsilon}
\renewcommand{\theta}{\vartheta}
\renewcommand{\phi}{\varphi}

% Frequently used math alphabets
\newcommand{\Bb}{\mathbb}
\newcommand{\Cal}{\mathcal}
\newcommand{\Bf}{\mathbf}
\newcommand{\Rm}{\mathrm}

% Frequently used letters in the blackboard alphabet
\newcommand{\CC}{\Bb C}
\newcommand{\NN}{\Bb N}
\newcommand{\PP}{\Bb P}
\newcommand{\QQ}{\Bb Q}
\newcommand{\RR}{\Bb R}
\newcommand{\EE}{\Bb E}

% Usage: \ang{...} is equivalent to \langle ... \rangle, while \ang*{...} is equivalent to \left\langle ... \right\rangle
% For other delimiters: use \qty from the physics package (i.e., \qty(...))
\DeclarePairedDelimiter{\ang}{\langle}{\rangle}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

% Frequently used commands
\newcommand{\T}{^\top} % Matrix transpose A\T
\newcommand{\C}{^\complement} % Set complement A\C
\renewcommand{\P}{^\perp}
\newcommand{\D}{^\dagger}
\newcommand\ceq\coloneqq % Definitions :=
\newcommand\pow{\Cal P} % Power sets
\newcommand\eps\epsilon
\newcommand\ind{\mathbbm 1} % Blackboard 1 for indicator functions
\newcommand\restr{\mathord\restriction}
\newcommand\TODO{{\color{red} TODO: }}
\newcommand\clos\overline

% Functions that appear frequently
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\Int}{Int}
\DeclareMathOperator{\Span}{Span}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\newcommand\XX{\Cal X}
\newcommand\YY{\Cal Y}
\newcommand\UU{\Cal U}
\newcommand\VV{\Cal V}
\newcommand\LL{\Cal L}
\newcommand\KK{\Cal K}
\newcommand\BB{\Cal B}
% \newcommand\LLL{\Bb L}
\DeclareMathOperator\Dom{\Cal D}
\DeclareMathOperator\Nul{\Cal N}
\DeclareMathOperator\Ran{\Cal R}

\title{Inverse Problems --- Example Sheet 1} % subject
\author{Lucas Riedstra}
\date{2 November 2020} % date

\begin{document}
\maketitle

\begin{question}
	For $\Omega = [0, 1]^2$ and $\XX \in L^2(\Omega)$, we consider the integral operator $A \colon \XX \to \XX$ with
	\[
	(Au)(y) \ceq \int_\Omega k(x, y) u(x) \dd{x},
	\]
	for $k \in L^2(\Omega \times \Omega)$.  Show that
	\begin{enumerate}[(a)]
		\item $A$ is linear with respect to $u$,
		\item $A$ is a bounded linear operator, i.e.\ $\norm{Au}_\XX \leq \norm{A}_{\LL(\XX, \XX)} \norm{u}_\XX$. Give also an estimate for $\norm{A}_{\LL(\XX, \XX)}$,
		\item the adjoint $A^*$ is given via
		\[
		(A^* v)(y) = \int_\Omega k(y, x) v(x) \dd{x}.
		\]
		\item $A$ is a compact operator, i.e.\ $A \in \KK(\XX, \XX)$.

		\emph{Hint:} you may use the fact that if an operator $A$ can be written as a limit (in the operator norm) of finite-rank operators then $A$ is compact. An operator $B$ is called finite-rank if $\dim(B) < \infty$. 
	\end{enumerate}
\end{question}

\begin{solution}
	Note: when writing a norm of a vector $v \in V$, I will simply write $\norm{v}$ and not $\norm{v}_V$, unless it is unclear in which space $v$ lives. The same holds for inner products.

	\begin{enumerate}[(a)]
		\item Let $\alpha, \beta \in \RR$, $u, v \in L^2(\Omega)$ and $y \in \Omega$. Then we have
		\begin{align*}
			(A(\alpha u + \beta v))(y) &= \int_\Omega k(x, y) (\alpha u + \beta v)(x) \dd{x} \\
			&= \int_\Omega k(x, y) \qty(\alpha u(x) + \beta v(x)) \dd{x}\\
			&= \alpha \int_\Omega k(x, y) u(x) \dd{x} + \beta \int_\Omega k(x, y) v(x) \dd{x} \\
			&= (\alpha A u)(y) + (\beta Av)(y) = (\alpha Au + \beta Av)(y). 
		\end{align*}
	Since equality holds for all $y \in \Omega$ we find $A(\alpha u + \beta v) = \alpha A u + \beta A v$, which proves that $A$ is linear. 
	\item Let $u \in L^2(\Omega)$, then we have
	\begin{align*}
	\norm{Au}^2 &= \int_\Omega ((Au)(y))^2 \dd{y} = \int_\Omega \qty( \int_\Omega k(x, y) u(x) \dd{x})^2 \dd{y} = \int_\Omega \ang{k(\cdot, y), u(\cdot)}^2 \dd{y}. 
	\end{align*}
Now we apply Cauchy-Schwarz and find
\begin{align*}
	\int_\Omega \ang{k(\cdot, y), u(\cdot)}^2 \dd{y} &\leq \int_\Omega \norm{k(\cdot, y)}^2 \norm{u}^2 \dd{y} \overset\star= \norm{u}^2 \iint_{\Omega^2} k^2(x, y) \dd{x}\dd{y} = \norm{u}^2 \norm{k}^2, 
\end{align*}
where $\star$ follows from Fubini's theorem since the integrand is nonnegative. 
Taking square roots on both sides we find that $\norm{Au}\leq \norm{k}\norm{u}$, so $A$ is bounded with $\norm{A} \leq \norm{k}$.

\item We know that the adjoint is the unique operator that satisfies $\ang{Au, v} = \ang{u, A^*v}$ for all $u, v \in \XX$. Let $u, v \in \XX$, then we compute
\begin{align*}
	\ang{Au, v} &= \int_\Omega (Au)(y)\cdot  v(y) \dd{y} = \int_\Omega \qty(\int_\Omega k(x, y) u(x) \dd{x}) v(y) \dd{y} \\
	&= \int_\Omega\int_\Omega k(x, y) u(x) v(y) \dd{x} \dd{y} \overset\star= \int_\Omega\int_\Omega k(x, y) u(x) v(y) \dd{y} \dd{x} \\
	&= \int_\Omega u(x) \qty(\int_\Omega k(x, y) v(y) \dd{y}) \dd{x} = \ang{u, A^*v}
\end{align*}
where $(A^* v)(x) = \int_\Omega k(x, y) v(y) \dd{y}$ as required. Here $\star$ follows from Fubini's theorem (\TODO justify).

\item It is known that for any compact set $X \subseteq \RR^n$, polynomials lie dense in $L^2(X)$. Therefore, there exists a sequence of polynomials $p_n$ such that $p_n \to k$ in $L^2(\XX)$. 
%\[
%(A_n u)(y) \ceq \int_\Omega p_n()
%\]
 \TODO Finish
	\end{enumerate}
\end{solution}

\begin{question}
	We consider the problem of differentiation, formulated as the inverse problem of finding $u$ from $Au = f$ with the integral operator $A \colon L^2([0, 1]) \to L^2([0, 1])$ defined as
	\[
	(Au)(y) \ceq \int_0^y u(x) \dd{x}.
	\]

	\begin{enumerate}[(a)]
		\item Let $f$ be given by
		\[
		f(x) \ceq \begin{cases} 0 & x < \frac12, \\ 1 &x > \frac12. \end{cases}
		\]
		Show that $f \in \clos{\Ran(A)}$. 
		\item Let $f$ be given as in a). Show that $f \in \clos{\Ran(A)} \setminus \Ran(A)$. \emph{Hint: Consider the Picard criterion}.

		\item Prove or falsify: ``The Moore-Penrose inverse of $A$ is continuous.''
	\end{enumerate}
\end{question}

\begin{solution}
	\begin{enumerate}[(a)]
		\item We want to show that we can approximate $f$ by a sequence $(Au_n)$ for some $(u_n) \subseteq L^2[0,1]$. To this end, define for $n \geq 2$
		\[
		u_n(x) =
		\begin{cases}
			0 &\abs{x - \frac12} > \frac1n, \\
			\frac n2 &\abs{x - \frac12} \leq \frac1n. 
		\end{cases}
		\]
		Clearly $u \in L^2[0, 1]$, and we have
		\[
		f_n(y) \ceq (Au_n)(y) = \int_0^y u_n(x) \dd{x} = \begin{cases}
			0 &\text{if $y < \frac12 - \frac1n$,}\\
			\frac n2(y - \frac12 + \frac1n) &\text{if $\frac12 - \frac1n \leq y \leq \frac12 + \frac1n$} \\
			1 &\text{ if $y > \frac12 + \frac1n$}. 
		\end{cases}
		\]

		Therefore we find
		\begin{align*}
			\norm{f_n - f}^2 &= \int_0^1 (f_n - f)^2(x) \dd{x} \\
			&= 2 \int_{\frac12 - \frac1n}^{\frac12} \frac{n^2}{4} (x - \frac12 - \frac1n)^2 \dd{x} \\
			&= \frac{n^2}{2} \int_0^{1/n}  x^2 \dd{x} = \frac{1}{6n} \to 0, 
		\end{align*}
	so $f_n \to f$ in $L^2[0, 1]$. Since $f_n \in \Ran(A)$ this shows $f \in \clos{\Ran(A)}$.

	% \item From Fourier theory it is known that $\qty{s_n(x) = \sqrt 2\sin(nx) : n \in \NN}$ and $\qty{1} \cup \qty{c_n(x) = \sqrt 2\cos(nx) \mid n \in \NN}$ are orthonormal bases of $L^2[0, 1]$. 
	%Note that 
	%\[
	%(A s_n)(y) = \int_0^y s_n(x) \dd{x} = \sqrt 2 \int_0^y \sin(nx) \dd{x} = \frac{\sqrt 2}{n} \qty(1 - \cos(ny)) = \frac{\sqrt 2}{n} c_0(y) - \frac1n c_n(y) 
	%\]
	\item Clearly, any operator in $\Ran(A)$ is continuous by the fundamental theorem of calculus. Since $f$ is not continuous (also not up to a measure zero set), we conclude $f \notin \Ran(A)$ and therefore $f \in \clos{\Ran(A)} \setminus \Ran(A)$.

	\end{enumerate}
\end{solution}
\end{document}

### 2020-10-21 distribution_theory/C1.tex
updatesssss

@@
 \[
-\ang{T_f, \phi} = \int_X f \phi \dd{x}. 
+\ang{\tau_f, \phi} = \int_X f \phi \dd{x}. 
 \]
-Henceforth we will abuse notation: if $f \in \locint(X)$, then we will write $f$ instead of $T_f$, i.e., $\ang{f, \phi} = \int_X f\phi \dd{x}$.
+Henceforth we will abuse notation: if $f \in \locint(X)$, then we will write $f$ instead of $\tau_f$, i.e., $\ang{f, \phi} = \int_X f\phi \dd{x}$.
 
@@
 \[
-(T_h\phi)(x) \ceq \phi(x - h),
+(\tau_h\phi)(x) \ceq \phi(x - h),
 \]
@@
     \[
-    \ang{T_hu, \phi} \ceq \ang{u, T_{-h}\phi} \quad\text{and}\quad \ang{\refl u, \phi} \ceq \ang{u, \refl{\phi}}. 
+    \ang{\tau_hu, \phi} \ceq \ang{u, \tau_{-h}\phi} \quad\text{and}\quad \ang{\refl u, \phi} \ceq \ang{u, \refl{\phi}}. 
     \]
@@
     \[
-    V_h \ceq \frac{T_{-h}u - u}{\norm{h}}
+    V_h \ceq \frac{\tau_{-h}u - u}{\norm{h}}
     \]
@@
 \begin{proof}
-    By definition, we can write $\ang{V_h, \phi} = \frac{1}{\norm{h}} \ang{u, T_h\phi - \phi}$. Now Taylor's theorem tells us that
+    By definition, we can write $\ang{V_h, \phi} = \frac{1}{\norm{h}} \ang{u, \tau_h\phi - \phi}$. Now Taylor's theorem tells us that
     \[
-    (T_h\phi - \phi)(x) = \phi(x - h) - \phi(x) = - \sum_i h_i \pdv{\phi}{x_i}(x) + R(x, h),
+    (\tau_h\phi - \phi)(x) = \phi(x - h) - \phi(x) = - \sum_i h_i \pdv{\phi}{x_i}(x) + R(x, h),
     \]
@@
 \subsubsection{Convolution}
-For $\phi \in \DD(\RR^n)$, note that $(T_x \refl\phi)(y) = \refl\phi(y-x) = \phi(x-y)$. 
+For $\phi \in \DD(\RR^n)$, note that $(\tau_x \refl\phi)(y) = \refl\phi(y-x) = \phi(x-y)$. 
 \begin{definition}
@@
     \[
-    (u * \phi)(x) \ceq \int_{\RR^n} u(x-y) \phi(y) \dd{y} = \int_{\RR^n} u(y) \phi(x - y) \dd{y} = \ang{u, T_x \refl\phi}. 
+    (u * \phi)(x) \ceq \int_{\RR^n} u(x-y) \phi(y) \dd{y} = \int_{\RR^n} u(y) \phi(x - y) \dd{y} = \ang{u, \tau_x \refl\phi}. 
     \]
@@
     \[
-    (u * \phi)(x) \ceq \ang{u, T_x\refl\phi}. 
+    (u * \phi)(x) \ceq \ang{u, \tau_x\refl\phi}. 
     \]
@@
 \end{proof}
+
+Due to the previous corollary, we often call $u * \phi$ a \emph{regularisation} of $u$.
+
+\begin{convention}
+	If $u \in \DD'(X)$ and $\phi \in \DD(X)$, then instead of $\ang{u, \phi}$ we also write $\ang{u(t), \phi(t)}$ (or with any other dummy variable) when the variable used for $\phi$ is not directly clear.
+\end{convention}
+
+\subsection{Density of $\DD(\RR^n)$ in $\DD'(\RR^n)$}
+\begin{lemma}
+	If $u \in \DD'(\RR^n)$ and $\phi, \psi \in \DD(\RR^n)$, then
+	\[
+	(u * \phi) * \psi = u * (\phi * \psi). 
+	\]
+\end{lemma}
+
+\begin{proof}
+	Fix $x \in \RR^n$. Now we write
+	\begin{align*}
+		((u * \phi) * \psi)(x) &= \int_{\RR^n} \ang{u(z), (\tau_y \refl\phi)(z)} \cdot \psi(x-y) \dd{y} \\
+		&= \int_{\RR^n} \ang{u(z), (\tau_{x-y}\refl\phi)(z)} \cdot \psi(y) \dd{y} \\
+		&= \int_{\RR^n} \ang{u(z), \psi(y) (\tau_{x-y} \refl\phi)(z)} \dd{y}.
+	\end{align*}
+We would like to interchange integral and application of $u$, and we will have to justify this using Riemann sums: 
+\begin{align*}
+\int_{\RR^n} \ang{u(z), \psi(y) (\tau_{x-y} \refl\phi)(z)} \dd{y} &= \lim_{\eps \downarrow 0} \sum_{m \in \ZZ^n} \ang{u(z), \psi(\eps m) \phi(x - z - \eps m)} \eps^n \\
+&\overset*= \lim_{\eps\downarrow0} \ang{u(z), \sum_{m \in \ZZ^n} \psi(\eps m) \phi(x - z - \eps m)\eps^n } \\
+&\overset{**}= \ang*{u(z), \int_{\RR^n} \psi(y) \phi(x - z - y) \dd{y}} \\
+&= \ang{u(z), (\phi * \psi)(x - z)} = \ang{u(z), (\tau_x\widecheck{\phi * \psi})(z)} = (u * (\phi * \psi))(x). 
+\end{align*}
+Here, $*$ is by the fact that the sum is finite since $\psi$ has compact support, while $**$ is by sequential continuity of $u$ and the fact that the Riemann sum converges to the convolution integral \emph{in the space of test functions} (non-examinable fact).
+\end{proof}
+
+\begin{theorem}
+	If $u \in \DD'(\RR^n)$, there exists a sequence $(\phi_k) \subseteq \DD(\RR^n)$ such that $\phi_k \to u$ in $\DD'(\RR^n)$. 
+\end{theorem}
+
+\begin{proof}
+	Fix $\psi \in \DD(\RR^n)$ with $\int_{\RR^n} \psi \dd{x} = 1$, and set $\psi_k(x) \ceq k^n \psi(kx)$. Note that $\int_{\RR^n} \psi_k \dd{x} = 1$.
+
+	Now, fix any $\chi \in \DD(\RR^n)$ with $\chi \equiv 1$ on $\qty{\norm{x} < 1}$ and $\chi \equiv 0$ on $\qty{\norm{x} < 2}$.
+Define $\chi_k(x) \ceq \chi(x/k)$, so that $\lim_{k\to\infty} \chi_k(x) = 1$ for all $x$.
+We will set
+	\[
+	\phi_k(x) \ceq \chi_k(x) (u * \psi_k)(x). 
+	\]
+	Clearly we have $\phi_k \in \DD(\RR^n)$ since each $\chi_k$ has compact support.
+
+	Note that $\ang{u, \phi} = (u * \refl\phi)(0)$ by definition. Now, take any $\theta \in \DD(\RR^n)$, then we have
+
+	\begin{align*}
+		\ang{\phi_k, \theta} &= \ang{\chi_k (u * \psi_k), \theta} = \ang{u * \psi_k, \chi_k \theta} = \qty[ (u * \psi_k) * \refl{\chi_k\theta}](0) \\
+		&= \qty[u * \qty(\psi_k * \refl{\chi_k \theta})](0).
+	\end{align*}
+
+	Now we compute $\psi_k * \refl{\chi_k\theta}$: note that 
+	\begin{align*}
+		(\psi_k * \refl{\chi_k\theta})(x) &= \int_{\RR^n} k^n\psi(k(x-y)) \chi\qty(-\frac yk) \theta(-y) \dd{y} \\
+		&= \int_{\RR^n} \psi(y) \chi\qty(\frac{y}{k^2} - \frac xk) \theta(\frac yk - x) \dd{y} \\
+		&= \theta(-x) + R_k(-x) = (\refl{\theta + R_k})(x)
+	\end{align*}
+where $R_k(x) = \int_{\RR^n} \psi(y) \qty[ \chi\qty(\frac y{k^2} + \frac xk) \theta\qty(\frac yk + x )- \theta(x)] \dd{y}$.
+
+So \[
+\ang{\phi_k, \theta} = (u * (\refl{\theta + R_k}))(0) = (u *\refl\theta)(0) + (u * \refl{R_k})(0) = \ang{u, \theta} + \ang{u, R_k}. 
+\]
+
+We must now only prove that $R_k \to 0$ in $\DD(\RR^n)$, and then by sequential continuity it follows that $\phi_k \to u$  in $\DD'(\RR^n)$. 
+\end{proof}

### 2020-10-21 inverse_problems/inverse_ES1.tex
updatesssss

@@
 	%\]
-	\item Clearly, any operator in $\Ran(A)$ is continuous by the fundamental theorem of calculus.
-Since $f$ is not continuous (also not up to a measure zero set), we conclude $f \notin \Ran(A)$ and therefore $f \in \clos{\Ran(A)} \setminus \Ran(A)$.
+	\item To apply the Picard criterion we must find the singular values and right singular vectors of $A$, which are equal to the square roots of the eigenvalues of $AA^*$ and the eigenvectors of $AA^*$.
 
+	Note that 
+	\begin{align*}
+		\ang{Au, v} &= \int_0^1 (Au)(y)\cdot v(y) \dd{y} \\
+		&= \int_0^1 \int_0^y u(x) \dd{x} v(y) \dd{y} \\
+		&= \int_0^1 \int_0^y u(x) v(y) \dd{x} \dd{y} \\
+		&= \int_0^1 \int_x^1 u(x) v(y) \dd{y} \dd{x} \\ 
+		&= \int_0^1 u(x) \int_x^1 v(y) \dd{y} \dd{x} \\
+		&= \ang{u, A^* v}
+	\end{align*}
+where $v(x) = \int_x^1 v(y) \dd{y}$.  Therefore, we find
+\begin{align*}
+	(AA^* u)(x) = \int_0^y \int_x^1 u(x) \dd{x} \dd{y}
+\end{align*}
 	\end{enumerate}

### 2020-10-21 modern_statistical_methods/C1.tex
updatesssss

@@
 
-This gives our first inequality $MSPE \leq \frac{\sigma^2}{n} \sum_{i=1}^n \frac{d_i^2}{(d_i + \lambda)^2} + \frac\lambda{4n}$.
+This gives our first inequality $\MSPE \leq \frac{\sigma^2}{n} \sum_{i=1}^n \frac{d_i^2}{(d_i + \lambda)^2} + \frac\lambda{4n}$.
 For the second inequality, note that $\frac{d_i^2}{(d_i + \lambda)^2} \leq 1$ and also $\frac{d_i^2}{(d_i + \lambda)^2} \leq \frac{d_i^2}{4 d_i\lambda} = \frac{d_i}{4\lambda}$.
@@
 \end{proof}
+
+We can now ask ourselves if kernel ridge regression is optimal in any sense.
+To this end, we define $\hat \mu_i \ceq d_i / n$ and $\lambda_n \ceq \lambda/n$.
+Then we can rewrite the upper bound from our precious theorem as
+
+\[
+\MSPE \leq \frac{\sigma^2}{n \lambda_n} \sum_{i=1}^n \min(\frac{\hat \mu_i}{4}, \lambda_n) + \frac{\lambda_n}{4} \eqqcolon \delta_n(\lambda_n).  
+\]
+
+Now, instead of taking $x_1, \dotsc, x_n \in \XX$ to be fixed, assume that they are i.i.d.\ random variables (then $K$ is random, so the $\hat\mu_i$ are random as well), and taking an expectation on both sides yields
+\[
+\EE[\MSPE] \leq \EE[\delta_n(\lambda_n)]. 
+\]
+
+We want to bound $\delta_n(\lambda_n)$ by some function of $n$. For this we will use Mercer's theorem:
+\begin{theorem}[Mercer]
+	Let $\XX = [a, b]$. If $k$ is a continuous kernel on $\XX$, then there exists an orthonormal basis $(e_i)$ of $L^2[a, b]$ such that
+	\[
+	k(x, x') = \sum_{j=1}^\infty \mu_j e_j(x) e_j(x')
+	\]
+	as well as a sequence of nonnegative eigenvalues $(\mu_j)$ such that
+	\[
+	\int_a^b k(x, x') e_j(x) \dd{x} = \mu_j e_j(x). 
+	\]
+\end{theorem}
+
+Applying Mercer's theorem to $k$, it can be proved that for some $C> 0$ we have
+\[
+\EE\qty[\frac1n \sum_{i=1}^n \min(\frac{\hat \mu_i}{4}, \lambda_n)] \leq \frac Cn \sum_{i=1}^\infty \min(\frac{\mu_i}{4}, \lambda_n).
+\]
+
+\begin{example}
+	Let $k$ be the Sobolev kernel. Then the eigenvalues satisfy 
+	\[
+	\frac{\mu_j}{4} = \frac{1}{\pi^2(2j - 1)^2}, 
+	\]
+	and with some calculations it can be shown that $\sum_{i=1}^\infty \min(\frac{\mu_i}{4}, \lambda_n) = O(\sqrt{\lambda_n})$ as $\lambda_n \to 0$. Combining this with the rest of the bound shows
+	\[
+	\EE[\delta_n(\lambda_n)] = O\qty(\frac{\sigma^2}{n} \lambda_n^{-1/2} + \lambda_n). 
+	\]
+	The optimal scaling $\lambda_n \sim (\sigma^2/n)^{2/3}$ gives an error rate of order $(\sigma^2/n)^{2/3}$.
+
+	It has been shown that this is in fact the optimal rate for estimating a function $f^0 \in \HH$ (up to multiplicative constants). 
+\end{example}
+
+\subsection{Other kernel machines}
+
+\subsubsection{The support vector machine}
+Consider a \emph{classification problem} with data $x_1, \dotsc, x_n \in \RR^p$ and binary response $Y_i \in \qty{\pm 1}$.
+Furthermore assume that $x_1, \dotsc, x_n$ are separated by a hyperplane through the origin, that is, for some  $\beta \in \RR^p$ we have $Y_i x_i\T \beta > 0$ for all $i$.
+
+To choose between different planes that separate the classes, we pick the hyperplane with the highest margin: that is, we compute
+\[
+\max_{\beta \in \RR^p, M \geq 0} M \qquad\text{such that } \frac{Y_i x_i\T\beta}{\norm{\beta}} \geq M \quad (i = 1,\dotsc, n). 
+\]
+When the classes are not separable, we can penalise according to the distance of a point ``over the margin''.
+The penalty should be 0 if $x$ is on the correct side and should equal the distance over the boundary otherwise, measured in units of $M$.
+In this case, by considering minimisation of $1/M^2$ instead of maximisation of $M$ we find
+\[
+\argmin_{M \geq 0, \beta \in \RR^p} \frac1{M^2} + \lambda \sum_{i=1}^n \qty(1 - \frac{Y_i x_i\T \beta}{\norm{\beta} M})_+. 
+\]
+Since the above equation is independent of the norm of $\beta$, we may set $\norm{\beta} = 1/M$ and remove $M$ from the equation entirely, thus obtaining
+\[
+\argmin_{\beta \in \RR^p} \norm{\beta}^2 + \lambda \sum_{i=1}^n (1 - Y_i x_i\T \beta)_+.
+\]
+Replacing $\lambda$ by $\frac1\lambda$ and multiplying the equation with $\lambda$ we obtain
+\[
+\argmin_{\beta \in \RR^p} \sum_{i=1}^n (1 - Y_i x_i\T \beta)_+ + \lambda \norm{\beta}^2. 
+\]
+Finally, if we lift the restriction that the hyperplane must go through the origin, we find the objective function
+\begin{equation} \label{eq:support_vector_classifier}
+(\hat \mu, \hat\beta) = \argmin_{\mu \in \RR, \beta \in \RR^p} \sum_{i=1}^n (1 - Y_i (x_i\T\beta + \mu))_+ + \lambda  \norm{\beta}^2. 
+\end{equation}
+The solution to this problem is known as the \emph{support vector classifier}.
+
+\paragraph{Introducing kernels}
+Now we introduce kernels to the problem. Letting $k$ be the linear kernel and $\HH$ the corresponding RKHS, we can rewrite \cref{eq:support_vector_classifier} as
+\[
+(\hat \mu, \hat f) = \argmin_{\mu \in \RR, f \in \HH} \sum_{i=1}^n (1 - Y_i\T (f(x_i) + \mu))_+ + \lambda \norm{f}^2.
+\]
+Suppose we change $\HH$ to a different RKHS with a different kernel $k$. By a variant of the representer theorem (Example Sheet 1, question 10) the solution is equivalent to
+\[
+(\hat\mu, \hat\alpha) = \argmin_{\mu \in \RR, \alpha \in \RR^n} \sum_{i=1}^n (1 - Y_i (K_i\T \alpha + \mu))_+ + \lambda \alpha\T K\alpha, 
+\]
+and the solution to the above problem is called the \emph{support vector machine}.  Predictions at a new point $x^*$ are given by
+\[
+\sign(\hat\mu + \sum_{i=1}^n \hat\alpha_i k(x^*, x_i)) \in \pm 1. 
+\]
+\begin{remark}
+Note that the support vector machine can correspond to a nonlinear boundary: the boundary is a hyperplane \emph{in the corresponding RKHS}, but can be nonlinear in the space $\XX$ where the data points live.
+This also makes it easy to overfit.
+\end{remark}

### 2020-10-21 topics_statistical_theory/C2.tex
updatesssss

\section{Kernel density estimation}
Let $X_1, \dotsc, X_n \iid f$, and suppose we wish to estimate the density function $f$. The oldest way to do this is with a histogram: we divide $\RR$ into equally sized intervals or \emph{bins}, and let $I_x$ denote the bin containing $x \in \RR$.
\begin{definition}
	The \emph{histogram density estimator} $\hat f_n^H$ with bin width $b > 0$ is given by
	\[
	\hat f_n^H(x) \ceq \frac{1}{nb} \sum_{i=1}^n \ind_{X_i \in I_x}.  
	\]
\end{definition}

There are a few major drawbacks to using histograms: it is difficult to choose $b$ and the positioning of bin edges, the theoretical performance is suboptimal (mostly due to their discontinuity) and graphical display in the multivariate case is difficult.

\subsection{The univariate kernel density estimator}
\begin{definition}
	A Borel measurable function $k \colon \RR \to \RR$ is called a \emph{kernel} if it satisfies $\int_\RR K(x) \dd{x} = 1$.

	A \emph{univariate kernel density estimator} of $f$ with kernel $K$ and \emph{bandwidth} $h > 0$ is defined as
	\[
	\hat f_n(x) \ceq \frac1{nh} \sum_{i=1}^n K\qty(\frac{x - X_i}{h}).
	\]
	Defining $K_h(x) \ceq \frac1h K\qty(\frac xh)$, we can rewrite this as
	\[
	\hat f_n(x) \ceq \frac1n \sum_{i=1}^n K_h(x - X_i). 
	\]
\end{definition}

Usually $K$ is chosen to be non-negative (which ensures that $K$ itself and $\hat f_n$ are themselves density functions), and $K$ is often chosen to be symmetric about 0. Generally, the choice of kernel $K$ is much less important than the choice of bandwidth $h$.

If we consider $\hat f_n(x)$ as a point estimator of $f(x)$, we typically wish to minimise the \emph{mean squared error}
\[
\MSE(\hat f_n(x)) \ceq \EE\qty[(\hat f_n(x) - f(x))^2]. 
\]
Other possibilities include the mean absolute error which (unlike the MSE) is scale-invariant. However, the MSE has an appealing decomposition into variance and bias terms:
\[
\MSE(\hat f_n(x)) = \Var(\hat f_n(x)) + \Bias^2(\hat f_n(x)). 
\]

We can express the MSE in terms of convolutions: 
\begin{definition}
	Let $g_1, g_2 \colon \RR \to \RR$ be measurable. Then the \emph{convolution} of $g_1$ and $g_2$, denoted $g_1 * g_2$, is defined by
	\[
	(g_1 * g_2)(x) \ceq \int_\RR g_1(x - z) g(z) \dd{z}. 
	\]
\end{definition}
We can compute 
\begin{align*}
\Bias \hat f_n(x) &= \EE[\hat f_n(x)] - f(x) = \EE[K_h(x - X_1)] - f(x) = \int_\RR  K_h(x - z) f(z) \dd{z} \\
&= (K_h * f)(x) - f(x). 
\end{align*}

Analogously, letting $\xi_i \ceq K_h(x - X_i)$ (note that these are i.i.d.\ random variables), we have
\begin{equation} \label{eq:variance_computation}
	\Var \hat f_n(x) = \frac1n \Var(\xi_1) = \frac1n \qty(\EE[\xi_1^2] - \EE^2[\xi_1]) = \frac1n \qty[(K_h^2 * f)(x) - (K_h * f)^2(x)].
\end{equation}

To assess performance of $h$ and $K$, we want to assess the performance of $\hat f_n$ as an estimation of $f$ as a function. This gives the following definition:
\begin{definition}
	We define the \emph{mean integrated squared error} or MISE as\[
	\MISE(\hat f_n) \ceq \EE\qty(\int_\RR \qty(\hat f_n(x) - f(x))^2 \dd{x}) \overset\star= \int_\RR \MSE(\hat f_n(x)) \dd{x}, 
	\]
	where $\star$ follows from Fubini's theorem since the integrand is nonnegative. 
\end{definition}

We now aim to find bounds on the bias and the variance of $\hat f_n$ in order to choose $h$ and $K$ appropriately.

\subsection{Bounds on variance and bias}
\begin{definition}
	For a kernel $K$, define $R(K) \ceq \int_\RR K^2(u) \dd{u}$. 
\end{definition}

\begin{proposition}
	Let $\hat{f_n}$ be the kernel density estimator with kernel $K$ and bandwidth $h > 0$ constructed from $X_1, \dotsc, X_n \iid f$. Then for any $x \in \RR, h > 0, n \in \NN$ we have
	\[
	\Var \hat f_n(x) \leq \frac{1}{nh} \norm{f}_\infty R(K). 
	\]
\end{proposition}

\begin{proof}
	By \cref{eq:variance_computation} we have
	\begin{align*}
	\Var \hat f_n(x) &\leq \frac1n (K_h^2 * f)(x) = \frac{1}{nh^2} \int_\RR K^2\qty(\frac{x - z}{h}) f(z) \dd{z} = \frac{1}{nh} \int_\RR K^2(u) f(x - uh) \dd{u} \\
	&\leq \frac1{nh} \norm{f}_\infty \int_\RR K^2(u) \dd{u} = \frac{1}{nh} \norm{f}_\infty R(K). 
	\end{align*}
\end{proof}

### 2020-10-21 topics_statistical_theory/C1.tex
worked hard on topics :-)

@@
 	Let $X_1, \dotsc, X_n$ be independent random variables with $\EE[X_i] = 0$, $\sigma_i^2 \ceq \Var(X_i) < \infty$, and $X_i \leq b$ for some $b > 0$.
-Define  $S \ceq \sum_{i=1}^n X_i$, $\nu \ceq \nu$ and $\phi \colon \RR \to \RR$ by $\phi(u) \ceq e^u - 1 - u = \sum_{k=2}^\infty \frac{u^k}{k!}$, then for every $t > 0$ we have
+Define  $S \ceq \sum_{i=1}^n X_i$, $\nu \ceq \overline{\Var(X_i)}$ and $\phi \colon \RR \to \RR$ by $\phi(u) \ceq e^u - 1 - u = \sum_{k=2}^\infty \frac{u^k}{k!}$, then for every $t > 0$ we have
 	\[
@@
 	\[
-	\PP(\bar X \geq x) \leq \exp\qty(- \frac{nx^2}{2(\sigma^2 + cx)}). 
+	\PP(\bar X \geq x) \leq \exp\qty(- \frac{nx^2}{2(\sigma^2 + cx)}),
 	\]
+	for all $x \geq 0$. 
 \end{theorem}

### 2020-10-21 topics_statistical_theory/topics_ES1.tex
worked hard on topics :-)

@@
 \DeclareMathOperator\Gam{Gamma}
+\DeclareMathOperator\Cov{Cov}
 
@@
 
-Finally, define $T = S_{n+1} - S_j$, so that $S_j$ and $T$ are independent. It is known that $S_j \sim \Gam(j, 1)$, $T \sim \gamma(n - j + 1)$, and furthermore that
+Finally, define $T_j = S_{n+1} - S_j$, so that $S_j$ and $T_j$ are independent. It is known that $S_j \sim \Gam(j, 1)$, $T_j \sim \gamma(n - j + 1, 1)$, and furthermore that
 \[
@@
         1 &y \geq 1. 
-    \end{cases}, \quad F_Y(0) = P(A). 
+    \end{cases}
     \]
+
     The empirical distribution function of $Y_1, \dotsc, Y_n \iid Y$ is given by
@@
     \]
-    and thus for $y = 0$ we have
+    and thus for $y \in [0, 1)$ we have
     \[
-    \hat F_n(0) = \frac1n \sum_{i=1}^n \ind_{Y_i \leq 0} = \frac1n \sum_{i=1}^n \ind_{U_{i} \in A} = \hat P_n(A).
+    \hat F_n(y) = \frac1n \sum_{i=1}^n \ind_{Y_i \leq y} = \frac1n \sum_{i=1}^n \ind_{X_i \in A} = \hat P_n(A).
     \]
 
-    The result now follows from the DKW inequality. \TODO make more precise. 
+    By the DKW inequality we find
+    \begin{align*}
+    	\PP\qty(\abs{\hat P_n(A) - P(A)} > \eps) = \PP \qty(\sup_{y \in \RR} \abs{\hat F_n(y) - F(y)} > \eps) \leq 2e^{-2n\eps^2}.
+    \end{align*}
 \end{proof}
+
+\begin{question}
+	Let $X \sim \Rm{Bin}(n, p)$. Compare the Hoeffding, Bennett, and Bernstein upper bounds on $\PP(X/n \geq \frac12)$ as $p \to 0$. 
+\end{question}
+
+\begin{solution}
+	Note that $X/n$ is the average of $n$ i.i.d.\ random variables $Y_i \sim \Rm{Bern}(p)$, where $Y_i \in [0, 1]$ for all $i$.
+
+	We start with Hoeffding's inequality. In this case, we have
+	\[
+	\PP(X/n - p \geq x) \leq \exp(-\frac{2n^2(1/2)^2}{n}) = \exp(- \frac n2),
+	\]
+	and this bound also holds as $p \to 0$.
+
+	We continue with Bennett's inequality.
+We consider the random variables $Y_i - p$, which are bounded from above by $b = 1 - p$.
+Now Bennett's inequality tells us, with $\sigma_p^2 = \Var(Y_i - p) = p(1-p)$ that
+	\begin{align*}
+		\PP(X/n \geq x) &\leq \exp(- \frac{n}{(1-p)^2} h\qty(\frac{1-p}{2p(1-p)})) = \exp\qty(- \frac{n}{(1-p)^2} h\qty(\frac{1}{2p})).
+	\end{align*}
+We compute
+\[
+h\qty(\frac1{2p}) = (1 + \frac1{2p}) \log(1 + \frac1{2p}) - \frac1{2p} = \log(1 + \frac1{2p}) + \frac1{2p} \qty(\log(1 + \frac1{2p}) - 1) \overset{p \downarrow 0}\to +\infty. 
+\]
+Since $\frac{n}{(1-p)^2}$ is clearly bounded by $n$, we conclude that 
+\[
+\PP(X/n \geq x) \to e^{-\infty} = 0. 
+\]
+
+%   Bennett's inequality tells us that
+%	\[
+%	\PP(X/n \geq \frac12) \leq \exp\qty(- n\nu h(1/(2\nu))), 
+%	\]
+%	where $h(u) = (1 + u) \log(1+ u)$ and $\nu = p(1-p)$. Plugging this in yields
+%	\[
+%	\exp(-n\nu h(x/\nu)) = \exp(-n p(1-p) \qty(1 + \frac{1}{2p(1-p)})\log(1 + \frac{1}{2p(1-p)}))
+%	\]
+
+We finish with Bernstein's inequality.  We have for $q \geq 3$ and $p \leq\frac12$ that 
+\begin{align*}
+	\EE[(Y_i - p)_+^q] &= p(1-p)^q = \sigma_p^2 (1-p)^{q-1} = (q! \sigma_p^2 (1-p)^{q-2}/2)\cdot (2(1-p)/q!) \\
+	&\leq q!\sigma_p^2 ((1-p)/3)^{q-2}/2.
+\end{align*}
+
+Now Bernstein's inequality tells us that
+\[
+\PP(X/n \geq \frac12 \leq \exp\qty(- \frac{n(1/2)^2}{2(\sigma_p^2 + (1-p)/6)}) = \exp(- \frac{n}{8\sigma_p^2 + 4(1-p)/3}) \overset{p \to 0}{\to} \exp(-\frac{3n}{4}). \]
+
+Of course, the true limit is 0 for any $n$, which is only given by Bennett's inequality. We also see that Hoeffding's inequality gives the worst result. 
+\end{solution}
+
+\begin{question}
+	Derive the following alternative form of Bernstein's inequality: under the same conditions, 
+	\[
+	\PP\qty(\bar X \geq \sqrt{\frac{2\sigma^2 \log(1/\delta)}{n}} + \frac cn \log(1/\delta)) \leq \delta
+	\]
+	for every $\delta \in (0, 1]$. 
+\end{question}
+
+\begin{proof}
+	We have 
+	\[
+	\PP(\bar X \geq x) \leq \exp(- \frac{nx^2}{2(\sigma^2 + cx)}) \eqqcolon \delta. 
+	\]
+	Now we just need express $x$ in terms of $\delta$: taking logarithms on both sides we obtain
+	\[
+	- \frac{nx^2}{2(\sigma^2 + cx)} = \log(\delta) \implies nx^2 = 2(\sigma^2 + cx) \log(1/\delta) \implies nx^2 - 2c\log(1/\delta) x - 2\sigma^2\log(1/\delta) = 0.
+	\]
+
+	Using the abc-formula with the fact that $x \geq 0$ yields 
+	\begin{align*}
+	x &= \frac{2c \log(1/\delta) + \sqrt{4c^2\log^2(1/\delta) + 8n\sigma^2\log(1/\delta)}}{2n} \\
+	&= \frac cn \log(1/\delta) + \sqrt{\frac{c^2}{n^2}\log^2(1/\delta) + \frac{2\sigma^2}{n} \log(1/\delta)} \\
+	&= ???? 
+	\end{align*}
+\end{proof}
+
+\begin{question}
+	\begin{enumerate}[(a)]
+		\item Let $X_1, \dotsc, X_n \iid F$ and let $\hat F_n$ denote their empirical distribution function. For $t_1 < \dotsb < t_k$, write down the distribution of 
+		\[
+		n\qty(\hat F_n(t_1), \hat F_n(t_2) - \hat F_n(t_1), \dotsc, \hat F_n(t_k) - \hat F_n(t_{k-1}), 1 - \hat F_n(t_k)). 
+		\]
+		\item Find the asymptotic distribution of $n^{1/2} (\hat F_n(t_1) - F(t_1), \dotsc, \hat F_n(t_k) - F(t_k))$. 
+	\end{enumerate}
+\end{question}
+
+\begin{solution}
+	\begin{enumerate}[(a)] \item
+	Write $n \hat F_n(t) = \sum_{i=1}^n \ind_{X_i \leq t} = \# \qty{i \mid X_i \leq t}$, and analogously, for $t < u$, 
+	$n (\hat F_n(u) - \hat F_n(t)) = \#\qty{i \mid t < X_i \leq u}$.
+
+	Then, defining $t_0 = -\infty$ and $t_{k+1} = \infty$,  we find that
+	\begin{gather*}
+	\PP\qty[ n \qty(\hat F_n(t_1), \dotsc, 1 - \hat F_n(t_k)) (a_1, \dotsc, a_{n+1})] \\ = \PP[\text{exactly $a_i$ of the $X_i$ lie in $(t_{i-1}, t_i]$ for $i = 1, \dotsc, n$}].
+	\end{gather*}
+
+This probability is 0 unless all $a_i$ are nonnegative integers with $\sum_i a_i = n$.
+In this case, we can first choose $a_1$ of the $X_i$ which must lie in $(\infty, t_1]$, then we choose $a_2$ of the remaining $X_i$ which must lie in $(t_1, t_2]$ and so forth.
+We find that the above probability equals
+\[
+\binom{n}{a_1} (F(t_1))^{a_1} \binom{n - a_1}{a_2} (F(t_2) - F(t_1))^{a_2} \dotsb \binom{a_k + a_{k_1}}{a_k} (F(t_k) - F(t_{k-1}))^{a_k} \cdot (1 - F(t_k))^{a_{k+1}}. 
+\]
+
+\item By the central limit theorem, the asymptotic distribution is $N(0, \Sigma)$, where $\Sigma$ is the covariance matrix of $(\hat F_n(t_1), \dotsc, \hat F_n(t_k))$.
+We will compute the entries of $\Sigma$.
+
+Choose $t \in \RR$ arbitrarily. Then we have
+\begin{align*}
+	\Var(\hat F_n(t)) &= \EE[\hat F_n^2(t)] - \EE[\hat F_n(t)]^2 = \EE\qty[\qty(\frac1n\sum_i \ind_{X_i \leq t})^2] - F^2(t) \\
+	&= \frac1{n^2}\EE\qty[\sum_i \ind_{X_i \leq t} + 2\sum_{i < j} \ind_{X_i \leq t} \ind_{X_j \leq t}] - F^2(t) \\
+	&= \frac{F(t) + (n-1) F^2(t)}{n} - F^2(t) = \frac{F(t)(1 - F(t))}{n},
+\end{align*}
+so we have computed the diagonal entries $\Sigma_{ii} = \frac{F(t_i)(1 - F(t_i))}{n}$.
+
+Now we must compute the covariances: assume $s < t$, then  
+\begin{align*}
+	\Cov(\hat F_n(s), \hat F_n(t)) &= \EE[\hat F_n(s) \hat F_n(t)] - \EE[\hat F_n(s)] \EE[\hat F_n(t)] \\
+	&= \frac1{n^2} \sum_{i, j} \EE[\ind_{X_i \leq s} \ind_{X_j \leq t}] - F(s)F(t) \\
+	&= \frac1{n^2} \qty(n F(s) + n(n-1) F(s)F(t)) - F(s)F(t) \\
+	&= \frac{F(s) + (n-1) F(s) F(t)}{n} - F(s)F(t) = \frac{F(s) - F(s)F(t)}{n}. 
+\end{align*}
+This gives the diagonal entries $\Sigma_{ij} = \frac{F(t_i) - F(t_i)F(t_j)}{n}$ for $i < j$. 
+In the end, we find
+\[
+\Sigma_{ij} = \frac1n \cdot \begin{cases}
+	F(t_i)(1 - F(t_i)) &\text{if $i = j$}, \\
+	F(t_{\min(i, j)}) - F(t_i) F(t_j) &\text{if $i \neq j$}. 
+\end{cases}
+\]
+\end{enumerate} 
+\end{solution}
 \end{document}

### 2020-10-23 distribution_theory/distribution_ES1.tex
finished msm example sheet 1git add . #partyyyy

@@
 
-\begin{solution}
+\begin{proof}
 	Since $\phi \in \DD(X)$, we know that $\supp\phi \subseteq \clos{B_N}$ for some $N \in \NN$. Now, suppose $\norm{h} < 1$, then
@@
 and therefore that $\pt^\beta R_N(x, h) = o(\norm{h}^n)$ for all multi-indices $\beta$. 
+\end{proof}
+
+\begin{question}
+	Which elements of $\DD(X)$ can be represented as a power series on $X$? 
+\end{question}
+
+\begin{solution}
+	It is known that if two power series agree on an open set, they agree on the entire space.
+Since every $\phi \in \DD(X)$ is identically zero on some open set (outside its support), the only element of $\DD(X)$ with a power series representation is the zero function.
+\end{solution}
+
+\begin{question}
+	Prove the $C^\infty$ Urysohn lemma: if $K$ is a compact subset of $X \subseteq \RR^n$, show that one can find a $\phi \in \DD(X)$ such that $0 \leq \phi \leq 1$ and $\phi = 1$ on a neighborhood of $K$.
+\end{question}
+
+\begin{solution}
+	Let $K \subseteq U_1$.
+Define $U_2 \ceq U_1 + B(0, 1)$ and let $\chi = \ind_{U_2}$.
+Now let $\psi \in \DD(\RR^n)$ satisfy $\int_{\RR^n} \psi \dd{x} = 1$ and $\supp\psi \subseteq B(0, 1)$.
+The we compute
+	\begin{align*}
+		(\chi * \psi)(x) &= \int_{\RR^n} \chi(y) \psi(x - y) \dd{y} = \int_{U_2} \psi(x - y) \dd{y}.
+	\end{align*}
+	Clearly, $\chi * \psi \in \DD(X)$, and furthermore, we have for $x \in U_1$ that
+	\[
+	\int_{U_2} \psi(x - y) \dd{y} = \int_{U_2 - x} \psi(z) \dd{z} \overset\star= 1,
+	\]
+	since $B(0, 1) \subseteq U_2 - x$. This proves the claim.
+\end{solution}
+
+\begin{question}
+	Given $T \in \DD'(X)$, the derivative $\pt^\alpha T$ is defined by
+	\[
+	\ang{\pt^\alpha T, \phi} = (-1)^\abs{\alpha} \ang{T, \pt^\alpha\phi} \quad \forall \phi \in \DD(X). 
+	\]
+	Show that $\pt^\alpha T \in \DD'(X)$. If $\ord(T) = m$ what can you say about $\ord(\pt^\alpha T)$? 
+\end{question}
+
+\begin{proof}
+	Let $K \subseteq X$ be compact and $\phi \in \DD(X)$. Since $T$ is a distribution, we know that there exists constants $C, N$ such that
+	\[
+	\abs{\ang{T, \phi}} \leq C \sum_{\abs{\beta} \leq N} \sup \abs{\pt^\beta \phi}. 
+	\]
+
+	Letting $M \ceq \abs{\alpha}$, we find
+		\begin{align*}
+		\abs{\ang{\pt^\alpha T, \phi}} &= \abs{\ang{T, \pt^\alpha\phi}} \leq C \sum_{\abs{\beta} \leq N} \sup \abs{\pt^{\alpha + \beta}\phi} \leq C \sum_{\abs{\beta} \leq M + N} \sup \abs{\pt^\beta \phi}.
+	\end{align*}
+We conclude that $\pt^\alpha T$ is a distribution, and that if $\ord(T) = m$, $\ord(\pt^\alpha T) \leq m + \abs{\alpha}$. 
+\end{proof}
+
+\begin{question}
+	Given $T \in \DD'(X)$ and $f \in C^\infty(X)$, prove that for each multi-index $\alpha$
+	\[
+	\pt^\alpha(Tf) = \sum_{\beta \leq \alpha} \binom{\alpha}{\beta} \pt^\beta f \pt^{\alpha - \beta}T
+	\]
+	in $\DD'(X)$. 
+\end{question}
+
+\begin{proof}
+	Let $\phi \in \DD(X)$, then by definition we have $\ang{\pt^\alpha(Tf), \phi} = \ang{T, (-1)^\abs{\alpha} f \pt^\alpha\phi}$. Approximate $T$ by a sequence $(\psi_n) \subseteq \DD'(X)$, then we find
+	\begin{align*}
+	\ang{T, (-1)^{\abs{\alpha}} f \pt^\alpha\phi} &= \lim_{n\to\infty} \ang{\psi_n,  (-1)^{\abs{\alpha}} f \pt^\alpha\phi} = \lim_{n\to\infty}  (-1)^{\abs{\alpha}} \int_X \psi_n(x) f(x) \pt^\alpha\phi(x) \dd{x} \\
+	&= \lim_{n\to\infty} \int_X \pt^\alpha (\psi_n(x) f(x)) \phi(x) \dd{x} \\
+	&= \lim_{n\to\infty} \int_X \qty(\sum_{\beta \leq \alpha} \binom\alpha\beta\pt^\beta f(x) \cdot \pt^{\alpha - \beta} \psi_n(x)) \phi(x) \dd{x} \\
+	&= \lim_{n\to\infty} \sum_{\beta \leq \alpha} \binom\alpha\beta \ang{\pt^\beta f \pt^{\alpha - \beta} \psi_n, \phi} =  \lim_{n\to\infty} \sum_{\beta \leq \alpha} \binom\alpha\beta (-1)^{\abs{\alpha - \beta}} \ang{\psi_n, \pt^\beta f \pt^{\alpha - \beta} \phi} \\
+	&= \sum_{\beta \leq \alpha} \binom\alpha\beta (-1)^{\abs{\alpha - \beta}} \ang{T, \pt^\beta f \pt^{\alpha - \beta} \phi} = \sum_{\beta \leq \alpha} \binom\alpha\beta \ang{\pt^\beta f \pt^{\alpha - \beta} T, \phi} \\
+	&= \ang*{\sum_{\beta \leq \alpha} \binom{\alpha}{\beta} \pt^\beta f \pt^{\alpha - \beta}T, \phi}. 
+	%&= \lim_{n\to\infty} \ang*{\sum_{\abs{\beta} \leq \abs{\alpha}} \binom\alpha\beta \pt^\beta f\cdot \pt^{\alpha - \beta} \psi_n, \phi} = \lim_{n\to\infty} \sum_{\abs{\beta} \leq \abs{\alpha}
+	\end{align*}
+\end{proof}
+
+\begin{question}
+	Let $(x_k)$ be a sequence in $X$ with no limit point in $X$. Consider the family of linear maps
+	$u_\alpha \colon \DD(X) \to \CC$ defined by 
+	\[ \ang{u_\alpha, \phi} = \sum_{k=1}^\infty \pt^\alpha \phi(x_k)
+	 \]
+	 for each multi-index $\alpha$. For what $\alpha$ is $u_\alpha \in \DD'(X)$? What is $\ord(u_\alpha)$? 
+\end{question}
+
+\begin{solution}
+	Let $K \subseteq X$ be compact.
+Since $(x_k)$ does not have a limit point, only finitely many of the $x_k$ lie in $K$ (otherwise $(x_k)$ would have a subsequence contained in $K$ which would have a convergent subsequence).
+Without loss of generality, assume that $x_1, \dotsc, x_n \in K$, and $x_{n+1}, x_{n+2}, \dotsc, \notin K$.
+Now, for any $\phi \in \DD(X)$ with $\supp(\phi) \subseteq K$ we find
+	\begin{align*}
+		\abs{\ang{u, \phi}} &= \abs{\sum_{k=1}^\infty \pt^\alpha\phi(x_k)} = \abs{\sum_{k=1}^n \pt^\alpha\phi(x_k)} \leq \sum_{k=1}^n \abs{\pt^\alpha \phi(x_k)} \leq n \cdot \sup \abs{\pt^\alpha\phi} \leq n \cdot \sum_{\abs{\beta} \leq \abs{\alpha}} \sup \abs{\pt^\beta \phi}.
+	\end{align*}
+	This shows that $u_\alpha \in \DD'(X)$ for any $\alpha$, with $\ord(u_\alpha) \leq \abs{\alpha}$. We claim that this is an equality, i.e., $\ord(u_\alpha) = \abs{\alpha}$. \TODO How to show??
+\end{solution}
+
+\begin{question}
+	Find the most general solution to the equations
+	\begin{enumerate}[(a)]
+		\item $u'  = 1$,
+		\item $xu' = \delta_0$,
+		\item $(e^{2\pi i x} - 1)u' = 0$
+	\end{enumerate}
+	in $\DD'(\RR)$. 
+\end{question}
+
+\begin{solution}
+	Let $\phi \in \DD(X)$. 
+	\begin{enumerate}[(a)]
+		\item If $u' = 1$ then we find 
+		\[
+		\int_\RR \phi(x) \dd{x} = \ang{1, \phi} = \ang{u', \phi} = -\ang{u, \phi'},
+		\]
+		so for any $c \in \RR$ we find by partial integration $\ang{u, \phi'} = - \int_\RR\phi(x) \dd{x} = \int_\RR (x + c) \phi'(x) \dd{x}$. From this we deduce that $u = x + c$ for some $c$. 
+	\end{enumerate}
 \end{solution}

### 2020-10-23 modern_statistical_methods/msm_ES1.tex
finished msm example sheet 1git add . #partyyyy

@@
 \newcommand\XX{\Cal X}
+\newcommand\HH{\Cal H}
 
@@
 \newcommand\TODO{{\color{red} TODO: }}
+\renewcommand\P{^\perp}
 
@@
 	\]
-	Then $\phi$ is continuous, and we can compute its Taylor series around 0. It is computed that 
+	Then $\phi$ is smooth, and we can compute its Taylor series around 0. It is computed that 
 	\[
@@
 \begin{question}
-	...
+	Consider minimising
+	\[
+	c(Y, X, f(x_1) + \mu, \dotsc, f(x_n) + \mu) + J(\norm{f}^2_\HH),
+	\]
+	over $f \in \HH$ and $\mu \in \RR$ where $\HH$ is an RKHS.
+Here $c$ is an arbirary loss function and $J$ is strictly increasing.
+Let $k$ be the reproducing kernel of $\HH$.
+Show that any minimiser $\hat g(\cdot) = \hat f(\cdot) + \hat \mu$ may be written as
+	\[
+	\hat g(\cdot) = \hat \mu + \sum_{i=1}^n \hat\alpha_i k(\cdot, x_i)
+	\]
+	where $\hat\alpha_i \in \RR$ for $i = 1, \dotsc, n$. 
 \end{question}
@@
 \begin{proof}
-	content...
+	As in the proof of the representer theorem, we project $f$ onto $U = \Span\qty{k(\cdot, x_1), \dotsc, k(\cdot, x_n)}$, and find that the problem is equivalent to finding
+	\[
+	\argmin_{u \in U, \mu \in \RR} c(Y, X, u(x_1) + \mu, \dotsc, u(x_n) + \mu) + J(\norm{u}^2),
+	\]
+	and writing $u = \sum_{i=1}^n \alpha_i k(\cdot, x_i)$, we find that any minimiser $\hat g$ satisfies
+	\[
+	\hat g(\cdot) = \hat \mu +  \sum_{i=1}^n \hat\alpha_i k(\cdot, x_i)
+	\]
+	as required. 
 \end{proof}
@@
 \begin{question}
-	...
+	This question proves a result needed for Theorem 7 in the notes.
+Let $\HH$ be a RKHS of functions on $\XX$ with reproducing kernel $k$ and suppose $f^0 \in \HH$.
+Let $x_1, \dotsc, x_n \in \XX$ and let $K$ be the kernel matrix $K_{ij} = k(x_i, x_j)$.
+Show that
+	\[
+	\qty(f^0(x_1), \dotsc, f^0(x_n))\T = K\alpha,
+	\]
+	for some $\alpha \in \RR^n$ and moreover that $\norm{f^0}^2_\HH \geq \alpha\T K\alpha$. 
 \end{question}
@@
 \begin{proof}
-	...
+	Let $U = \Span\qty{k(\cdot, x_1), \dotsc, k(\cdot, x_n)}$. Then we have $f^0 = u + v$ for some $u \in U, v \in U\P$, and writing $u = \sum_{m=1}^n \alpha_m k(\cdot, x_m)$ we see that 
+		\[
+	f^0(x_i) = \ang{k(\cdot, x_i), u + v} = \ang{k(\cdot, x_i), u} = u(x_i) = \sum_{m=1}^n \alpha_m k(x_i, x_m) = (K\alpha)_i,
+	\]
+	and from this it follows immediately that 
+	\[
+	\norm{f^0}^2 = \norm{u}^2 + \norm{v}^2 \geq \norm{u}^2 = \sum_{i, j} \alpha_i \alpha_j k(x_i, x_j) = \alpha\T K\alpha. 
+	\]
 \end{proof}
@@
 \end{solution}
+
+\begin{question}
+	Let $\HH$ be a RKHS with reproducing kernel $k$. Show that if $h_x \in \HH$ has the property that $\ang{h_x, f} = f(x)$ for all $f \in \HH$, then $h_x(\cdot) = k(\cdot, x)$. 
+\end{question}
+
+\begin{proof}
+	For any $y \in \XX$ we have
+	\[
+	h_x(y) = \ang{k(\cdot, y), h_x} = \ang{h_x, k(\cdot, y)} = k(x, y) = k(y, x),
+	\]
+	from which it follows that $h_x(\cdot) = k(\cdot, x)$. 
+\end{proof}
+
+\begin{question}
+	Prove that if $k$ is a reproducing kernel for RKHS's $\HH_1$ and $\HH_2$, then $\HH_1 = \HH_2$, so the RKHS is uniquely determined by $k$.
+\emph{Hint: First argue that it is enough to show the result for $\HH_1 \subseteq \HH_2$.
+Next consider decomposing each $f \in \HH_2$ as $f = u+ v$ with $u \in \HH_1, v \in \HH_1\P$ and argue that $v = 0$.
+}
+\end{question}
+
+\begin{proof}
+	Without loss of generality we can assume $\HH_1 = \HH$, where $\HH$ is the RKHS constructed in the proof of Theorem 5, so $\HH$ is the completion of $\Span\qty{k(\cdot, x) \mid x \in \XX}$.
+Now we can embed $\HH$ into $\HH_2$ by mapping $k(\cdot, x)$ to $k_x$.
+Clearly this mapping respects the inner product by construction, so this is a linear embedding.
+Therefore, we may assume $\HH_1 \subseteq \HH_2$.
+
+	Now suppose $\HH_1 \subseteq \HH_2$ and decompose each $f \in \HH_2$ as $f = u + v$ with $u \in \HH_1$, $v \in \HH_1\P$.
+By the previous question we know that for every $x$ there exists a unique $k_x$ with $\ang{k_x, f} = f(x)$ for all $f$, and clearly we have $k_x \in \HH_1$.
+
+	Then for each $f \in \HH_2, x \in \XX$ we have (since $k_x \in \HH_1$) that
+	\[
+	f(x) = \ang{k_x, f} = \ang{k_x, u} + \ang{k_x, v} = \ang{k_x, u} = u(x),
+	\]
+	which implies $v = 0$. Since this is the case for every $f \in \HH_2$, we conclude that $\HH_1\P = 0$ and therefore $\HH_1 = \HH_2$ (since Hilbert spaces are closed). 
+\end{proof}
 \end{document}

### 2020-10-23 modern_statistical_methods/C1.tex
finished msm example sheet 1git add . #partyyyy
updated msm summary w/ lecture 7

@@
 \]
-Suppose we change $\HH$ to a different RKHS with a different kernel $k$. By a variant of the representer theorem (Example Sheet 1, question 10) the solution is equivalent to
+Suppose we change $\HH$ to a different RKHS with a different kernel $k$. By a variant of the representer theorem (Example Sheet 1, question 9) the solution is equivalent to
 \[
@@
 \end{remark}
+
+\subsubsection{Logistic regression}
+\begin{recap}
+	The \emph{standard logistic regression} model is motivated by assuming 
+	\[
+	\log\qty(\frac{\PP(Y_i = 1)}{\PP(Y_i = -1)}) = x_i\T \beta^0. 
+	\]
+	The maximum-likelihood estimator $\hat\beta$ is given by (Example sheet 2, question 4)
+	\[
+	\hat\beta \ceq \argmin_{\beta \in \RR^p} \sum_{i=1}^n \log\qty(1 + \exp(- Y_i x_i\T\beta)). 
+	\]
+\end{recap}
+
+Just like in ridge regression, we will try to lower the variance of $\hat\beta$ by penalising large values, i.e.,
+\begin{align*}
+\hat\beta_\lambda &\ceq \argmin_{\beta \in \RR^p} \sum_{i=1}^n\log(1 + \exp(-Y_i x_i\T\beta)) + \lambda \norm{\beta}^2 \\
+&= \argmin_{f \in \HH} \sum_{i=1}^n \log(1 + \exp(- Y_i f(x_i))) + \lambda \norm{f}^2,
+\end{align*}
+where $\HH$ is the RKHS corresponding to the linear kernel.
+
+Note that this is exactly the same as the support vector machine, but with a different loss function.
+The loss function in the support vector machine is called ``hinge loss'', while the loss function in logistic regression is called ``logistic loss''.
+The main differences are:
+\begin{enumerate}
+	\item The logistic loss function increases much quicker for incorrectly classified data. 
+	\item Even when a point is correctly classified, the logistic loss function decreases the further the datapoint is from the boundary.
+The hinge loss function will always be 0 if a point is correctly classified and lies outside of the margin.
+\end{enumerate}
+
+The second point also explains the name \emph{support vector machine}: the only points that contribute to the loss function are those that either violate the boundary of are within the margin (the \emph{support vectors}), and perturbing any of the other points does not change the boundary.
+Because of this property, SVM's tend to be more stable.
+
+\subsection{Large scale kernel machines}
+We have seen that kernels give very flexible regression and classification estimators, and that the parameter $\lambda$ helps us with the bias-variance tradeoff.
+Also, when $p \gg n$, using the kernel matrix saves computational effort.
+
+We now turn to the case $n \gg p$.
+In this case, both kernel ridge regression and the SVM are very expensive to compute: the former has cost $O(n^3)$, while the latter can be computed using an iterative algorithm where every iteration has cost $O(n^2)$.
+
+One approach to speed this process up is that of \emph{random feature maps}: we develop a random map $\hat\phi \colon \XX \to \RR^b$ (with $b$ small) such that $\EE[\hat\phi(x)\T\hat\phi(x')] = k(x, x')$ for all $x, x' \in \XX$.
+Then we consider $L$ i.i.d.\ copies of $\hat\phi$ (denotes $\hat\phi_1, \dotsc, \hat\phi_n$), and we consider the feature map
+\[
+\psi(x) = L^{-1/2} (\hat\phi_1(x), \dotsc, \hat \phi_L(x))\T \in \RR^{Lb}, \text{ so that } \psi(x)\T\psi(x') = \frac1L \sum_{i=1}^L \hat\phi_i(x)\T \hat\phi_i(x'). 
+\]
+Letting $\Phi$ be the matrix with rows $\psi(x_1)\T, \dotsc, \psi(x_n)\T$, we find that $\EE(\Phi\Phi\T)_{ij} = k(x_i, k_j)$, and that $\Var(\Phi\Phi\T)_{ij}$ decreases with speed $L^{-1}$.
+So $\Phi\Phi\T$ is a good approximation for $K$ when $L$ is moderately large.
+
+One specific example for shift-invariant kernels ($k(x, x') = h(x - x')$ for all $x, x' \in \XX$) is based on the work of Rahimi and Recht. 
+\begin{theorem}[Bochner]
+	Let $k \colon \RR^p \times \RR^p \to \RR$ be a continuous kernel. Then $k$ is shift-invariant if and only if, for some $c > 0$ and some distribution $F$ on $\RR^p$, it holds for all $W \sim F$ that
+	\[
+	k(x, x') = c \EE\qty[ \exp(i(x - x')\T W)] = c\EE\qty[ \cos((x - x')\T W)]. 
+	\]
+\end{theorem}
+
+To make use of this theorem, we let $u \sim U[-\pi, \pi]$, and it can then be computed (using $u \overset d= -u$ and $\cos(u) \overset d= \sin(u)$ that
+\[
+2\EE\qty[\cos(x + u) \cos(y + u)] = \cos(x - y).  
+\]
+Therefore, given a shift-invariant kernel $k$ with associated distribution $F$, we define for $W \sim F$ the feature map
+\[
+\hat\phi(x) = \sqrt{2c} \cos(W\T x + u) \in \RR, 
+\]
+so that
+\[
+\EE\hat\phi(x) \hat\phi(x') = 2c \EE\qty[ \EE[ \cos(W\T x + u)\cos(W\T x' + u) \mid W]] = c \EE \cos((x - x\T)W) = k(x, x'). 
+\]
+
+\begin{example}
+	Let $k$ be the Gaussian kernel $k(x, x') = e^{-\norm{x - x'}/(2\sigma^2)}$. Let $W \sim N(0, \sigma^{-2}I)$, then using the characteristic function of a normal distribution we find that
+	\[
+	\EE\qty(e^{i(x - x')\T W}) = \exp(- \norm{x - x'}^2/(2\sigma^2)) = k(x, x'), 
+	\]
+	so we can set $c = 1$, i.e., define the feature map $\hat\phi(x) = \sqrt 2 \cos(W\T x + u)$. 
+\end{example}

### 2020-10-25 distribution_theory/C1.tex
more updates on this lovely sunday

@@
 
-\begin{lemma}
+\begin{lemma}[Sequential continuity] \label{lem:seq_continuity}
     Let $u \colon \DD(X) \to \CC$ be a linear map.
-Then $u$ is a distribution if and only if, for every sequence $(\phi_j) \subseteq \DD(X)$ with $\phi_j \to 0$ as in \cref{def:test_convergence}, we have $\ang{u, \phi_m} \to 0$.
+Then $u$ is a distribution if and only if, for every sequence $(\phi_m) \subseteq \DD(X)$ with $\phi_m \to 0$ as in \cref{def:test_convergence}, we have $\ang{u, \phi_m} \to 0$.
 \end{lemma}

### 2020-10-25 distribution_theory/C2.tex
more updates on this lovely sunday

\section{Distributions with compact support}
\begin{definition}
	Let $Y \subseteq X$ be open and $u \in \DD'(X)$. We say that $u$ \emph{vanishes} on $Y$ if $\ang{u, \phi} = 0$ for all $\phi \in \DD(Y)$. 
\end{definition}

\begin{definition}
	For $u \in \DD'(X)$, we define the \emph{support} of $u$ as
	\[
	\supp u \ceq X \setminus \bigcup\qty{Y \subseteq X \mid Y \text{ open}, u \text{ vanishes on $Y$}}. 
	\]
\end{definition}
For example, the support of $\delta_x$ is simply $\qty{x}$.

\subsection{Test functions and distributions}
We will now consider a bigger space of test functions, which yields a smaller space of distributions. 
\begin{definition}
	We define $\EE(X)$ as the space of smooth functions $\phi \colon X \to \CC$. We say that a sequence $(\phi_m) \subseteq \EE(X)$ converges to 0 if $\pt^\alpha\phi \to 0$ uniformly on compact subsets of $X$ for every multi-index $\alpha$. 
\end{definition}

\begin{definition}
	We define 
	%the space of \emph{distributions with compact support}, denoted 
	$\EE'(X)$ as the space of linear maps $u \colon \EE(X) \to \CC$ for which there exists a compact $K \subseteq X$ and nonnegative constants $C, N$ such that
	\begin{equation} \label{eq:dist_compact_support}
	\abs{\ang{u, \phi}} \leq C \sum_{\alpha \leq N} \sup_K \abs{\pt^\alpha\phi}
	\end{equation}
	for all $\phi \in \EE(X)$. 
\end{definition}

\begin{lemma}[Sequential continuity] \label{lem:seq_continuity_compact_support}
	Let $u \colon \EE(X) \to \CC$ be a linear map. Then $u \in \EE'(X)$ if and only if, for every sequence $(\phi_m) \subseteq \EE(X)$ with $\phi_m \to 0$, we have $\ang{u, \phi_m} \to 0$. 
\end{lemma}

\begin{proof}
	\TODO
\end{proof}

\begin{lemma}
	If $u \in \EE'(X)$, then $u\restr_{\DD(X)}$ defines an element of $\DD'(X)$ with compact support and finite order.

	Conversely, for each $u \in \DD'(X)$ with compact support there exists a unique extension $\tilde u \in \EE'(X)$ with $\supp(\tilde u) = \supp(u)$ and $\tilde u\restr_{\DD(X)} = u$. 
\end{lemma}

\begin{proof}
	Let $u \in \EE'(X)$, so that there exists a compact $K \subseteq X$ with $\abs{\ang{u, \phi}} \leq C \sum_{\abs{\alpha} \leq N} \sup_K \abs{\pt^\alpha \phi}$. 
	Now, for any compact $K' \subseteq X$ and any $\phi$ with $\supp\phi \subseteq K'$, \cref{eq:seminorm_cond} is clearly satisfied, and we can use the same $N$ for all compact $K'$, so clearly $u\restr_{\DD(X)}$ is an element of $\DD'(X)$ with finite order.
	Finally, suppose $\phi$ is supported in $X \setminus K$, then it is clear that $\ang{u, \phi} = 0$, which proves that $\supp u \subseteq K$ and therefore that $u$ has compact support.

	Now suppose $u \in \DD'(X)$ has compact support, let $\rho \in \DD(X)$ be 1 in a neighbourhood of $\supp u$, and define $\tilde u \in \EE'(X)$ by
	\[
	\ang{\tilde u, \phi} \ceq \ang{u, \rho\phi} \quad\forall \phi \in \EE(X). 
	\]
	Clearly $\tilde u$ is an element of $\EE'(X)$ since $\supp(\rho\phi) \subseteq \supp\rho$ and 
	\[
	\abs{\ang{\tilde u, \phi}} = \abs{\ang{u, \rho\phi}} \leq C \sum_{\abs{\alpha} \leq N} \sup_{\supp(\rho)} \abs{\pt^\alpha(\rho \phi)} \overset\star\leq C' \sum_{\abs{\alpha} \leq N} \sup_{\supp\rho} \abs{\pt^\alpha\phi},
	\]
	where $\star$ follows from the Leibniz rule. It is also clear that $\supp \tilde u = \supp u$.

	Finally we will show uniqueness: suppose $\tilde v$ is an extension of $u$ with $\supp \tilde v = \supp u$, and write any $\phi \in \EE(X)$ as $\phi = \rho\phi + (1-\rho)\phi = \phi_0 + \phi_1$. Then since $\phi_0 \in \DD(X)$ and $\phi_1$ vanishes on a neighbourhood of $\supp u$, we find
	\[
	\ang{\tilde v, \phi} = \ang{\tilde v, \phi_0 + \phi_1} = \ang{\tilde v, \phi_0} = \ang{u, \phi_0},
	\]
	which is independent of choice of extension. 
\end{proof}

\subsection{Convolution between $\EE'(\RR^n)$ and $\DD'(\RR^n)$}
\begin{definition}
	Define for $u \in \EE'(\RR^n), \phi \in \EE(\RR^n)$ the \emph{convolution}
	\[
	(u * \phi)(x) = \ang{u, \tau_x \refl\phi}. 
	\]
\end{definition}
This convolution satisfies the same properties as the convolution between $\DD'(\RR^n)$ and $\DD(\RR^n)$. 
Also, if $\phi \in \DD(\RR^n)$, then $u * \phi \in \DD(\RR^n)$.

\begin{definition}
	Let $u, v \in \DD'(\RR^n)$, at least one of which has compact support, define $u * v \colon \DD(\RR^n) \to \EE(\RR^n)$ by
	\[
	(u * v) * \phi = u * (v * \phi) \quad \forall \phi \in \DD(\RR^n). 
	\]
\end{definition}
We will show (example sheet 2, question 1) that $u * v$ is uniquely defined and gives rise to an element of $\DD'(\RR^n)$ via $\ang{u * v, \phi} \ceq [(u * v) * \refl\phi](0)$.

\begin{lemma}
	Given $u, v \in \DD'(\RR^n)$, at least one of which has compact support, we have $u * v = v * u$. 
\end{lemma}

\begin{proof}
	First we note that $(u * \phi) * \psi = u * (\phi * \psi)$ holds if $u$ has compact support and at least one of $\phi, \psi$ has compact support.

	Fix $\phi, \psi \in \DD(\RR^n)$, we see from our earlier shown properties that 
	\begin{align*}
	(u * v) * (\phi * \psi) &= u * \qty(v * (\phi * \psi)) = u * ((v * \phi) * \psi) = u * (\psi * (v * \phi)) = (u * \psi) * (v * \phi).
	\end{align*}
	If we interchange $u$ and $v$ in the above, that is equivalent to interchanging $\phi$ and $\psi$, which we know must yield the same result. This shows $u * v$ and $v * u$ agree on $\phi * \psi$ for all $\phi, \psi \in \DD(\RR^n)$, and from this it follows that $u * v = v * u$ (\TODO why?).
\end{proof}

### 2020-10-25 inverse_problems/C2.tex
IP up to lec6
more updates on this lovely sunday

@@
 	\]
-	If all $R_\alpha$ are linear, then we speak of a \emph{linear regularisation} of $A\D$. 
+	If all $R_\alpha$ are linear (\TODO and bounded?), then we speak of a \emph{linear regularisation} of $A\D$. 
 \end{definition}
@@
 We need some process to choose a parameter. To this end, note that we have
-\[
+\begin{equation} \label{eq:error_terms}
 \norm{R_\alpha f_\delta - A\D f} \leq \norm{R_\alpha (f_\delta - f)} + \norm{(R_\alpha - A\D) f} \\
-\leq  \norm{R_\alpha} \norm{f_\delta - f} + \norm{f} \norm{R_\alpha - A\D}. 
-\]
+\leq  \delta \norm{R_\alpha} +\norm{(R_\alpha - A\D) f}  . 
+\end{equation}
 The first term is called the \emph{data error} and is unbounded for $\alpha \to 0$, and the second term is called the \emph{approximation error} which does vanish for $\alpha \to 0$.
 Therefore, we want to choose $\alpha$ small enough to have a low approximation error, while keeping the data error at bay.
-We define this formally:
 
+\subsection{Parameter choice rules}
 \begin{definition}
@@
 \end{definition}
+
+\begin{definition}
+	Let $(R_\alpha)_{\alpha > 0}$ be a regularisation of $A\D$ and $\alpha$ a parameter choice rule. We call $(R_\alpha, \alpha)$ a \emph{convergent regularisation} if
+	\[
+	\lim_{\delta \to 0} \sup_{f_\delta : \norm{f - f_\delta} \leq \delta} \norm{R_\alpha f_\delta - A\D f} = 0
+	\]
+	and
+	\begin{equation} \label{eq:alpha_to_0}
+	\lim_{\delta \to 0} \sup_{f_\delta : \norm{f - f_\delta} \leq \delta} \alpha(\delta, f_\delta) = 0. 
+	\end{equation}
+\end{definition}
+
+\subsubsection{A priori parameter choice rules}
+
+We will not prove the following theorem, which guarantees the existence of a priori PCRs: 
+\begin{theorem}
+	Let $(R_\alpha)_{\alpha > 0}$ be a regularisation of $A\D$. Then there exists an a priori PCR $\alpha = \alpha(\delta)$ such that $(R_\alpha, \alpha)$ is convergent. 
+\end{theorem}
+
+We can characterise PCRs in the following way:
+\begin{theorem}
+	Let $(R_\alpha)_{\alpha > 0}$ be a linear regularisation of $A\D$, and $\alpha = \alpha(\delta)$ an a priori PCR. Then $(R_\alpha, \alpha)$ is convergent if and only if
+	\[
+	\lim_{\delta \to 0} \delta \norm{R_{\alpha(\delta)}} = 0 \quad\text{and}\quad \lim_{\delta\to 0} \alpha(\delta) = 0. 
+	\]
+\end{theorem}
+
+\begin{proof}
+	``$\implies$'' Suppose $(R_\alpha, \alpha)$ is convergent.
+It is clear that $\lim_{\delta \to 0} \alpha(\delta) = 0$ by \cref{eq:alpha_to_0}.
+Suppose $\lim_{\delta\to 0} \delta \norm{R_{\alpha(\delta)}} \neq 0$.
+Then there exists a sequence $(\delta_k) \to 0$ and a constant $C > 0$ such that $\delta_k \norm{R_{\alpha(\delta_k)}} \geq C$ for all $k$.
+This implies we can find a sequence $(g_k) \subseteq \YY$ with $\norm{g_k} = 1$ and $\delta_k \norm{R_{\alpha(\delta_k)}g_k} \geq C$ for all $k$.
+
+	Now let $f \in \Dom(A\D)$ and define $f_k \ceq f + \delta_k g_k$, then clearly we have $f_k \to f$, but also
+	\[
+	C \leq \norm{R_{\alpha(\delta_k)}(\delta_k g_k)} = \norm{R_{\alpha(\delta_k)}(f_{\delta_k} - f)} \leq \norm{R_{\alpha(\delta_k)}f_{\delta_k} - A\D f} + \norm{(R_{\alpha(\delta_k)} - A\D) f}.
+	\]
+	In particular we find that $\norm{(R_{\alpha(\delta_k)} - A\D) f} \geq C$, so clearly $R_\alpha$ is not convergent.
+
+	``$\impliedby$'' This follows immediately from \cref{eq:error_terms}. 
+\end{proof}
+
+A problem with a priori PCRs is that they are scale-invariant: if $\alpha = \alpha(\delta)$ gives a convergent regularisation, then $\hat\alpha = \alpha(k\delta)$ also gives a convergent regularisation for any $k$.
+In practice, it is not always clear which scale should be chosen.
+
+\subsubsection{A posteriori parameter choice rules}
+Let $f \in \Dom(A\D)$ and $f_\delta$ s.t.\ $\norm{f- f_\delta} \leq \delta$.
+Letting $u\D$ denote the minimum-norm solution of the problem $Au = f$, and defining $\mu \ceq \norm{Au\D - f} = \inf_{u \in \XX} \norm{Au - f}$, we see that
+\[
+\norm{Au\D - f_\delta} \leq \norm{Au\D - f} + \norm{f - f_\delta} \leq \mu + \delta. 
+\]
+Therefore, it is not useful to choose $\alpha(\delta, f_\delta)$ with $\norm{Au_\alpha - f_\delta} < \mu + \delta$: if this is the case, we are most likely overfitting.
+
+This motivates \emph{Morozov's discrepancy principle}:
+
+\begin{definition}
+	Let $(R_\alpha)$ be a (\TODO linear?) regularisation of $A\D$ and assume $\Ran(A)$ is dense in $\YY$. Fix $\eta > 1$, and define
+	\[
+	\alpha(\delta, f_\delta) = \sup\qty{\alpha > 0 : \norm{A R_{\alpha} f_\delta - f_\delta} \leq \eta\delta}.
+	\]
+	Then $\alpha(\delta, f_\delta)$ is said to satisfy \emph{Morozov's discrepancy principle}. 
+\end{definition}
+
+It can be shown that the above $\alpha$ indeed gives a convergent regularisation.
+
+\subsubsection{Heuristic parameter choice rules}
+Heuristic parameter choice rules unfortunately only work if the original problem was well-posed:
+\begin{theorem}[Bakushinskii]
+	Let $(R_\alpha)$ be a regularisation of $A\D$ and suppose there exists a heuristic parameter choice rule $\alpha$ such that $(R_\alpha, \alpha)$ is convergent.
+Then $A\D$ is continuous from $\YY$ to $\XX$.
+\end{theorem}
+
+\subsection{Spectral regularisation}
+We will now start with specific examples of regularisations. Spectral regularisations are derived from the spectral decomposition
+\[
+A\D f = \sum_{j=1}^\infty \sigma_j^{-1} \ang{f, y_j} x_j. 
+\]
+We construct a regularisation by replacing $\sigma_j^{-1}$ by some function $g_\alpha(\sigma_j)$, i.e.,
+\begin{equation} \label{eq:spectral_reg}
+R_\alpha f = \sum_{j=1}^\infty g_\alpha(\sigma_j) \ang{f, y_j} x_j. 
+\end{equation}
+Let us explore which conditions $g_\alpha$ must satisfy:
+
+\begin{theorem} \label{thm:spectral_reg}
+	Let, for $\alpha > 0$, the function $g_\alpha \colon \RR_{>0} \to \RR_{>0}$ satisfy
+	\begin{enumerate} \itemsep=0em
+		\item $\lim_{\alpha \to 0} g_\alpha(\sigma) = \frac1\sigma$ for all $\sigma > 0$;
+		\item $g_\alpha(\sigma) \leq C_\alpha$ for some $C_\alpha > 0$; 
+		\item $\sup_{\alpha, \sigma}  \sigma g_\alpha(\sigma) \leq \gamma$ for some $\gamma > 0$. 
+	\end{enumerate}
+	Then collection $(R_\alpha)$ defined by \cref{eq:spectral_reg} is a linear regularisation of $A\D$, and in particular, we have $\norm{R_\alpha} \leq C_\alpha$. 
+\end{theorem}
+
+\begin{proof}
+	From condition 2 it follows that all $R_\alpha$ are bounded. Since 
+	\[\ang{f, y_j} = \ang{P_{\clos{\Ran(A)}} f, y_j} = \ang{AA\D f, y_j} = \ang{A\D f, A^* y_j} = \sigma_j \ang{u\D, x_j},\]
+	we compute
+	\[
+	(R_\alpha - A\D) f = \sum_j (g_\alpha(\sigma_j) - \sigma_j)^{-1} \ang{f, y_j} x_j = \sum_j (\sigma_j g_\alpha(\sigma_j) - 1) \ang{u\D, x_j} x_j, 
+	\]
+	and since $\sigma g_\alpha(\sigma_j) \leq \gamma$, we have $(\sigma_j g_\alpha(\sigma_j) - 1)^2 \leq 1 + \gamma^2$, so that 
+	\[
+	\norm{(R_\alpha - A\D) f}^2 \leq (1 + \gamma^2) \norm{u\D}^2 < \infty. 
+	\]
+
+	Since $\norm{(R_\alpha - A\D) f}$ is finite, we may apply the reverse Fatou lemma to the sum and obtain
+	\[
+	\limsup_{\alpha \to 0} \norm{(R_\alpha - A\D)f}^2 \leq \sum_j \qty(\sigma_j \limsup_{\alpha \to 0} g_\alpha(\sigma_j) - 1)^2 \ang{u\D, x_j}^2 = 0,
+	\]
+	and therefore $R_\alpha f \to A\D f$ as $\alpha \to 0$. 
+\end{proof}
+
+\begin{example}
+	The first, very simple example is the \emph{truncated SVD}: we simply define
+	\[
+	g_\alpha(\sigma) = \begin{cases} 1/\sigma &\sigma \geq \alpha, \\ 0 &\sigma < \alpha. \end{cases}
+	\]
+	It is easy to check that $g_\alpha$ satisfies the conditions of \cref{thm:spectral_reg}, and that all $R_\alpha$ are finite-rank operators with $\norm{R_\alpha} \leq \frac1\alpha$.
+Therefore, if we choose $\alpha = \alpha(\delta)$ such that $\delta/\alpha(\delta) \to 0$, then we obtain a convergent regularisation.
+
+	This also highlights the problem with this method: as $\delta$ gets smaller, we need more and more singular vectors which are generally expensive to compute. 
+\end{example}
+
+\begin{example}
+	The second example is \emph{Tikhonov regularisation}.
+Here, we define $g_\alpha(\sigma) = \frac{\sigma}{\sigma^2 + \alpha}$, and again it is easily checked that the conditions of \cref{thm:spectral_reg} are satisfied, noting that
+	\[
+	\frac{\sigma}{\sigma^2 + \alpha} \leq \frac{\sigma}{2 \sigma \sqrt{\alpha}} = \frac{1}{2\sqrt\alpha} \eqqcolon C_\alpha. 
+	\]
+	Therefore, if $\delta/\sqrt{\alpha(\delta)} \to 0$, the regularisation is convergent.
+
+	This method does not require computing the SVD of $A$: it is easily shown that $u_\alpha \ceq R_\alpha f$ is the unique solution to the \emph{regularised normal equation}
+	\[
+	(A^*A + \alpha I) u_\alpha = A^* f. 	
+	\]
+
+	While $A^* A + \alpha I$ is always invertible, computing the inverse is expensive, so we usually use some approximation of the inverse.
+
+	Finally, it can also be shown that 
+	\[
+	u_\alpha = \min_{u \in \XX} \norm{A u - f}^2 + \alpha \norm{u}^2,
+	\]
+	so we can also view $u_\alpha$ as the solution of an optimisation problem. 
+\end{example}

### 2020-10-27 distribution_theory/C2.tex
worked on some example sheets, updated msm and topics summaries (todo: dist and inv prob)

@@
 	If we interchange $u$ and $v$ in the above, that is equivalent to interchanging $\phi$ and $\psi$, which we know must yield the same result.
-This shows $u * v$ and $v * u$ agree on $\phi * \psi$ for all $\phi, \psi \in \DD(\RR^n)$, and from this it follows that $u * v = v * u$ (\TODO why?).
+This shows $u * v$ and $v * u$ agree on $\phi * \psi$ for all $\phi, \psi \in \DD(\RR^n)$.
+Defining $E = u * v - v * u$, we find that $0 = E * (\phi * \psi) = (E * \phi) * \psi$ for all $\phi, \psi \in \DD(\RR^n)$, so $E * \phi = 0$ for all $\phi \in \DD(\RR^n)$, so $E = 0$.
 \end{proof}

### 2020-10-27 distribution_theory/distribution_ES1.tex
worked on some example sheets, updated msm and topics summaries (todo: dist and inv prob)

@@
 \newcommand{\RR}{\Bb R}
-\newcommand{\EE}{\Bb E}
+\newcommand{\EE}{\Cal E}
 \newcommand{\DD}{\Cal D}
+\newcommand\ZZ{\Bb Z}
 
@@
 		so for any $c \in \RR$ we find by partial integration $\ang{u, \phi'} = - \int_\RR\phi(x) \dd{x} = \int_\RR (x + c) \phi'(x) \dd{x}$. From this we deduce that $u = x + c$ for some $c$.
+
+		\item If $xu' = \delta_0$ then we have
+		\[
+		\phi(0) = \ang{\delta_0, \phi} = \ang{xu', \phi} = \ang{u', x\phi} = \ang{u, -(x\phi)'} = \ang{u, -x\phi' - \phi} %=  \ang{u, -x\phi'} + \ang{u, -\phi}. 
+		\]
+		Note that the above is satisfied for $u = - \delta_0 + c$ for any constant $c$. \TODO is this the most general solution?
+
+		\item Since $e^{2\pi i x} = 1 \iff x \in \ZZ$, intuitively it must be the case that $u'$ is 0, except ``on $\ZZ$'', whatever that may mean.
+Therefore, we guess that, for any sequence $(\alpha_n)_{n \in \ZZ} \subseteq \CC$ and constant $c \in \CC$, the map \[u = c + \sum_{n \in \ZZ} \alpha_n \ind_{x \geq n}.\] We compute the derivative of $u$.
+It is easily seen that $u' = \sum_{n \in \ZZ} \alpha_n \delta_n$ (the infiniteness of the sum does not pose a problem since the test functions are compactly supported, so $\ang{u, \phi}$ will always be a finite sum).
+%		We have for arbirary $\phi$ that
+%		\begin{align*}
+%		\ang{u', \phi} &= - \ang{u, \phi'} = -\int_\RR \sum_{n \in \ZZ} \qty(\alpha_n \ind_{x \geq n}) \phi'(x) \dd{x}	\overset\star= - \sum_{n \in \ZZ} \alpha_n \int_n^\infty \phi'(x) \dd{x} = \sum_{n \in \ZZ} \alpha_n \phi(n),
+%		\end{align*}
+%	where $\star$ follows from the fact that the sum is finite since $\phi'$ is compactly supported. 
+%	Therefore, we can conclude $u' = \sum_{n \in \ZZ} \alpha_n \delta_n$, and indeed we find
+From this, we see that
+	\[
+	\ang{(e^{2\pi i x} - 1) u', \phi} = \ang{u', (e^{2\pi i x} -1 )\phi} = \sum_{n \in \ZZ} \alpha_n (e^{2\pi i n} - 1) \phi(n) = 0,
+	\]
+	so $u$ satisfies the equation. \TODO Why is this the most general solution? Intuitively clear, but how to make this rigorous?
 	\end{enumerate}
 \end{solution}
+
+\begin{question}
+	Define the distribution $u \in \DD'(\RR^2)$ by the locally integrable function $u(x, y) = \ind_{x \geq y}$.
+Show that $\pt_x^2   u - \pt_y^2 u = 0$ in $\DD'(\RR^2)$.
+Can you give a physical interpretation of this result?
+\end{question}
+
+\begin{proof}
+	Let $f \in \DD(\RR^2)$, then we have
+	\begin{align*}
+		\ang{\pt_x^2 u - \pt_y^2 u, f} &= \ang{\pt_x^2 u, f}- \ang{\pt_y^2 u, f} = \ang{u, \pt_x^2 f} - \ang{u, \pt_y^2 f} = \ang{u, \pt_x^2 f - \pt_y^2 f} \\
+		&\overset\star= \int_{-\infty}^\infty \int_{y}^\infty \pt_x^2 f(x, y) \dd{x}\dd{y} - \int_{-\infty}^\infty \int_{-\infty}^x \pt_y^2 f(x, y) \dd{y} \dd{x} \\
+		&= - \int_{-\infty}^\infty \pt_x f(y, y) \dd{y} - \int_{-\infty}^\infty \pt_y f(x, x) \dd{x}
+		\\
+		&= - \int_{-\infty}^\infty (\pt_xf + \pt_yf)(x, x) \dd{x}.
+	\end{align*}
+Here, $\star$ follows from Fubini's theorem. 
+Define $g(x) = f(x, x)$, then it is easily seen that $g'(x) = \pt_x f(x, x) + \pt_y f(x, x)$, so we find that
+\[
+\ang{\pt_x^2 u - \pt_y^2 u, f} = 
+- \int_{-\infty}^\infty g'(x) \dd{x} = \lim_{x\to -\infty} g(x) - \lim_{x\to\infty} g(x) = 0 - 0 = 0. 
+\]
+\end{proof}
+
+This shows that $\pt_x u - \pt_y u = 0$, or equivalently, that $u$ satisfies the wave equation. 
 \end{document}

### 2020-10-27 inverse_problems/inverse_ES1.tex
worked on some example sheets, updated msm and topics summaries (todo: dist and inv prob)

@@
 \newcommand{\EE}{\Bb E}
+\newcommand\ZZ{\Bb Z}
+\newcommand\FF{\Cal F}
 
@@
 \DeclareMathOperator\Ran{\Cal R}
+\DeclareMathOperator\diag{diag}
 
@@
 
+\begin{mdframed}
+		Note: when writing a norm of a vector $v \in V$, I will simply write $\norm{v}$ and not $\norm{v}_V$, unless it is unclear in which space $v$ lives. The same holds for inner products. 	
+\end{mdframed}
 \begin{question}
@@
 \begin{solution}
-	Note: when writing a norm of a vector $v \in V$, I will simply write $\norm{v}$ and not $\norm{v}_V$, unless it is unclear in which space $v$ lives. The same holds for inner products.
 
@@
 
-\item It is known that for any compact set $X \subseteq \RR^n$, polynomials lie dense in $L^2(X)$. Therefore, there exists a sequence of polynomials $p_n$ such that $p_n \to k$ in $L^2(\XX)$. 
+\item It is known that for any compact set $X \subseteq \RR^n$, polynomials lie dense in $L^2(X)$.
+Therefore, there exists a sequence of polynomials $p_n$ such that $p_n \to k$ in $L^2([0, 1]^4)$.
+It is easily seen that for any polynomial $p$, the operator
+\[
+(A_p u)(y) \ceq \int_\Omega p(x, y) u(x) \dd{x} 
+\] 
+has finite rank: let $p(z) = \sum_{\abs{\alpha} \leq n} c_\alpha z^\alpha$ (where $z \in [0, 1]^4$ and $\alpha$ is a multi-index), then we find
+\[
+(A_p u)(y) = \sum_{\abs{\alpha} \leq n} c_\alpha \int_\Omega x_1^{\alpha_1} x_2^{\alpha_2} y_1^{\alpha_3} y_2^{\alpha_4} u(x) \dd{x} = \sum_{\abs{\alpha} \leq n} c_\alpha \qty(\int_\Omega x_1^{\alpha_1} x_2^{\alpha_2} \dd{x}) y_1^{\alpha_3} y_2^{\alpha_4},
+\]
+so $A_p u$ lies in the $\Span\qty{y_1^{\alpha_1} y_2^{\alpha_2} \mid \alpha_1 + \alpha_2 \leq n}$, and therefore has finite rank. 
+By (b), we find that $\norm{A - A_n} \leq \norm{k - p_n} \to 0$, which shows that $A_n \to A$ in operator norm. We conclude that $A$ is compact. 
 %\[
@@
 %\]
- \TODO Finish
 	\end{enumerate}
@@
 			\frac n2(y - \frac12 + \frac1n) &\text{if $\frac12 - \frac1n \leq y \leq \frac12 + \frac1n$} \\
-			1 &\text{ if $y > \frac12 + \frac1n$}. 
+			1 &\text{if $y > \frac12 + \frac1n$}. 
 		\end{cases}
@@
 	%\]
-	\item To apply the Picard criterion we must find the singular values and right singular vectors of $A$, which are equal to the square roots of the eigenvalues of $AA^*$ and the eigenvectors of $AA^*$.
+	\item 	Note that $A$ is compact, since we can write $(Au)(y) = \int_{[0, 1]} k(x, y) u(x) \dd{x}$ with $k(x, y) = \ind_{x \leq y}$.
 
-	Note that 
+	To apply the Picard criterion we must find the singular values and right singular vectors of $A$, which are equal to the square roots of the eigenvalues of $AA^*$ and the eigenvectors of $AA^*$.
+
 	\begin{align*}
@@
 \begin{align*}
-	(AA^* u)(x) = \int_0^y \int_x^1 u(x) \dd{x} \dd{y}
+	(AA^* u)(x) = \int_0^x \int_y^1 u(z) \dd{z} \dd{y}.
 \end{align*}
+Now we solve the eigenequation. Firstly, note that $\Nul(A) = \qty{0}$, so let $\lambda^2 > 0$. Then we have
+\begin{align}
+	\int_0^x \int_y^1 u(z) \dd{z} \dd{y} &= \lambda^2 u(x) \\
+	\int_x^1 u(z) \dd{z} &= \lambda^2 u'(x) \\
+	- u(x) &= \lambda^2 u''(x). 
+\end{align}
+Furthermore, from (1) we infer $u(0) = 0$ while from (2) we infer $u'(1) = 0$. 
+The general solution to (3) is given by $u(x) = a \cos(\lambda^{-1} x) + b \sin(\lambda^{-1} x)$. 
+Plugging $u(0) = 0$ gives $a = 0$, and $u'(1) = 0$ gives
+\[
+b\lambda^{-1} \cos(\lambda^{-1}) = 0 \implies b = 0 \text{ or } \lambda^{-1} = \pi\qty(n - \frac12) \text{ for some $n \in \ZZ_{>0}$}.
+\]
+
+Our singular values are therefore $\frac{1}{(n - \frac12)\pi}$ ($n \in \ZZ_{>0}$) with right singular vectors $y_n(x) = \sqrt 2\sin((n - \frac12) \pi x)$. 
+We now compute
+\[
+\ang{f, y_n} = \sqrt 2\int_{1/2}^1 \sin(\qty(n - \frac12) \pi x) \dd{x} = ...
+\]
+
+\item The Moore-Penrose inverse of $A$ is discontinuous.
+This can be seen by theorem 2.1.11: in (a) we have constructed a sequence $(f_n) \subseteq \Ran(A)$ that converges, and in (b) we have shown that its limit lies outside of $\Ran(A)$.
+Therefore, $\Ran(A)$ is not closed, so $A\D$ is discontinuous.
+
 \end{enumerate}
 \end{solution}
+
+\begin{question}
+	\begin{enumerate}[(a)]
+		\item
+
+	Let $m, n \in \NN$ with $m \geq n \geq 2$. Compute the Moore-Penrose inverses of the following matrices:
+	\begin{enumerate}[(i)]
+		\item $A = (1, 1, \dotsc, 1) \in \RR^{1 \times n}$;
+		\item $A = \diag(a_1, \dotsc, a_n) \in \RR^{n \times n}$ with $a_j \in \RR$ for $j = 1, \dotsc, n$;
+		\item $A \in \RR^{m \times n}$ with $A\T A = I_n$. 
+	\end{enumerate}
+
+\item Let $a, b \in \RR$ with $a < b$. Compute the Moore-Penrose inverse of the operator $A \colon L^2([a, b]) \to \RR$ with $Au = \int_a^b u(x) \dd{x}$. 
+	\end{enumerate}
+\end{question}
+
+\begin{solution}
+	\begin{enumerate}[(a)]
+		\item \begin{enumerate}[(i)]
+			\item Clearly $\Ran(A) = \RR$ and $\Nul(A)\P = \Span\qty{\vec e}$ where $\vec e$ is the all-ones vector. So $A\D$ must map $x \in \RR$ to $\vec e/n$, and therefore we have
+			\[
+			A\D = \vec e/n \in \RR^{n \times 1}.
+			\]
+
+			\item Clearly we have $\Ran(A) = \Span\qty{\vec e_j \mid a_j \neq 0} = \Nul(A)\P$ while $\Ran(A)\P = \Span\qty{\vec e_j \mid a_j = 0} = \Nul(A)$.
+
+			It is easily seen that
+			\[
+			A\D = \diag(b_1, \dotsc, b_n) \in \RR^{n \times n}, \quad\text{where } b_i = \begin{cases}
+				a_i^{-1} &\text{if $a_i \neq 0$}; \\
+				0 &\text{if $a_i = 0$}. 
+			\end{cases}
+			\]
+
+			\item If $A\T A = I$, then the columns $\vec u_1, \dotsc, \vec u_n$ of $A$ are orthogonal.
+In particular, $A$ is injective (so $\Nul(A) = \qty{0}$) and $\Ran(A) = \Span\qty{\vec u_1, \dotsc, \vec u_n}$.
+Extend $\vec u_1, \dotsc, \vec u_n$ to an orthonormal basis $\vec u_1, \dotsc, \vec u_m$ of $\RR^m$, then we must have
+			\[
+			A\D\qty(\sum_{i=1}^m \alpha_i \vec u_i) = \mqty(\alpha_1 \\ \vdots \\ \alpha_n) \in \RR^n.
+			\]
+			From this expression it is easily seen that $A\D = A\T$. 
+		\end{enumerate}
+
+	\item Clearly we have $\Ran(A) = \RR$ while $\Nul(A) = \qty{u \mid \int_a^b u(x) \dd{x} = 0}$. It is also easily seen that 
+	\[
+	\Nul(A)\P = \qty{v \mid \int_a^b v(x) u(x) \dd{x} = 0 \text{ if $\int_a^b u(x) \dd{x} = 0$}} = \Span\qty{1}. 
+	\]
+
+	Therefore we simply have that $A\D$ maps a constant $c \in \RR$ to the constant function $\frac{c}{b-a}$. 
+	\end{enumerate}
+\end{solution}
+
+\begin{question}
+	Many forward problems are either modelled as convolutions or they are modelled as the composition of several components, one of which is a convolution.
+Therefore convolutions play an important role in inverse problems.
+As in Exercise 1, let $\Omega =[0, 1]^2$ be the unit square and let $\XX = L^2(\Omega)$.
+A convolution is the special case of an integral operator $A \colon \XX \to \XX$ where the kernel has a simple structure:
+	\[
+	(Au)(y) \ceq \int_\Omega k(y - x) u(x) \dd{x},
+	\]
+	for $k \in L^2(\Omega)$. It follows easily from Exercise 1 that $A$ is linear and bounded. 
+	\begin{enumerate}[(a)]
+		\item Although shown in general in Exercise 1, give an explicit form for the adjoint of the convolution. 
+		\item Let $f = Au$. It follows from the convolution theorem that a convolution can be inverted by means of the Fourier transform
+		\begin{equation} \label{eq:fourier}
+		u = (2\pi)^{-\frac n2} \FF^{-1} \qty(\frac{\FF(f)}{\FF(k)}),
+		\end{equation}
+		where $\FF$ is the Fourier transform and $\FF^{-1}$ its inverse.
+Implement this formula in \emph{MATLAB} to deblur the blurry tree image $f$ generated by the script \verb|ex4b_generate_data.m|.
+Note that the script also outputs $\FF(k)$.
+Add some noise to the data and show that the inversion formula is ill-conditioned.
+
+		Hint: Make use of the \emph{MATLAB} commands \verb|fft2| and \verb|ifft2|.
+
+		\item Reformulate \cref{eq:fourier} so that the denominator is non-negative and give a stable approximation of this formula. Implement this formula in MATLAB and empirically show that it is stable.
+
+		Hint: Make use of the \emph{MATLAB} command \verb|conj|. 
+	\end{enumerate}
+\end{question}
+
+\begin{proof}
+	\begin{enumerate}[(a)]
+		\item We have by question 1 that
+		\[
+		(A^* v)(y) = \int_\Omega k(x - y)v(x) \dd{x}.
+		\]
+		\item 
+	\end{enumerate}
+\end{proof}
 \end{document}

### 2020-10-27 modern_statistical_methods/C2.tex
worked on some example sheets, updated msm and topics summaries (todo: dist and inv prob)

\section{The Lasso}
\subsection{Model selection}
We go back to the linear model $Y = X\beta^0 + \eps$ with $\EE[\eps] = 0$ and $\Var(\eps) = \sigma^2 I$. Using the trace trick, one can easily compute that the MSPE of the OLS estimator is given by
\[
\frac1n \EE\norm{X\beta^0 - X\hat\beta^\Rm{OLS}}_2^2 = \frac{\sigma^2 p}{n}. 
\]
Defining $S = \qty{k \mid (\beta^0)_k \neq 0}$, there is often reason to assume that $S$ is small, i.e., $s \ceq \abs{S} \ll p$. If we could fit a model using only the variables in $S$, the MSPE would be much $\frac{\sigma^2 s}{n} \ll \frac{\sigma^2 p}{n}$.

\paragraph{Best subset selection} A natural way to find $S$ is to consider all possible subsets of $\qty{1, \dotsc, p}$, and pick the best regression procedure using, for example, cross-validation. However, this can become computationally infeasible for moderately large $p$ (say $p \approx 10$).

\paragraph{Forward selection} This is a greedy way of performing best subset regression. Given a target model size $m$, we first compute the intercept-only model $M_0$, and then one-by-one add the predictor variable that reduces the residual sum of squares the most, until we have a model with $m$ variables.

\subsection{Lasso estimator}
The \emph{Least absolute shrinkage and selection operator} or \emph{Lasso} is given by
\[
(\hat\mu_\lambda^\Rm{L}, \hat\beta_\lambda^\Rm{L}) \ceq \argmin_{(\mu, \beta) \in \RR \times \RR^p} \frac1{2n}\norm{Y- \mu \vec 1 - X\beta}_2^2 + \lambda \norm{\beta}_1. 
\]
As with ridge regression, we usually centre and scale the matrix $X$, as well as centre the responses $Y$, in which case we find
\[
\hat\beta_\lambda^\Rm{L} = \argmin_{\beta \in \RR^p} \frac1{2n}\norm{Y - X\beta}_2^2 + \lambda \norm{\beta}_1. 
\]

The main difference between the lasso estimator and the ridge regression estimator is that it is likely that the lasso estimator has some zero components. This means that the lasso estimator also estimates which variables are relevant.

We have the following:
\begin{theorem}[Slow rate]
	Assume $X$ has centred and scaled columns, and assume that $Y$ has been centred, so $Y = X\beta^0 + \eps - \bar\eps \vec 1$. Let $A > 0$ and suppose
	\[
	\lambda = A\sigma\sqrt{\frac{\log(p)}{n}}.
	\]
	Let $\hat\beta = \hat\beta_\lambda^\Rm{L}$, then with probability at least $1 - 2p^{-A^2/2 - 1}$ we have that the MSPE satisfies
	\[
	\frac1n \norm{X(\beta^0 - \hat\beta)}_2^2 \leq 4\lambda \norm{\beta^0}_1 = 4A\sigma\sqrt{\frac{\log(p)}{n}} \norm{\beta^0}_1. 
	\]
\end{theorem}

\begin{proof}
	By definition we have
	\[
	\frac1{2n}\norm{Y - X\hat\beta }_2^2 + \lambda \norm{\hat \beta}_1 \leq \frac1{2n}\norm{Y - X\beta^0}_2^2 + \lambda \norm{\beta^0}_1,
	\]
	and rearranging the terms gives
	\[
	\frac1{2n} \norm{X(\beta^0 - \hat\beta)}_2^2 \leq\frac1n \eps\T X (\hat\beta - \beta^0) + \lambda\norm{\beta^0}_1 - \lambda \norm{\hat\beta}_1. 
	\]
	By H\"older's inequality we have $\abs{\eps\T X(\beta^0 - \hat\beta)} \leq \norm{X\T\eps}_\infty \norm{\hat\beta - \beta^0}_1$. 
	Define the event \\ $\Omega = \qty{\norm{X\T\eps}_\infty/n \leq \lambda}$, then conditional on $\Omega$ we find
	\[
	\frac1{n} \norm{X(\beta^0 - \hat\beta)}_2^2 \leq 2\lambda \qty(\norm{\beta^0 - \hat\beta}_1 + \norm{\beta^0}_1 - \norm{\hat\beta}_1) \leq 4\lambda \norm{\beta^0}_1.
	\]
	In (\TODO refer), we will show that $\PP(\Omega) \geq 1 - 2p^{-(A^2/2 - 1)}$, which completes the proof. 
\end{proof}

### 2020-10-27 modern_statistical_methods/msm_ES1.tex
worked on some example sheets, updated msm and topics summaries (todo: dist and inv prob)

@@
 
+¨
+
 \begin{question}
@@
 \begin{proof}
-	Without loss of generality we can assume $\HH_1 = \HH$, where $\HH$ is the RKHS constructed in the proof of Theorem 5, so $\HH$ is the completion of $\Span\qty{k(\cdot, x) \mid x \in \XX}$.
+%	Without loss of generality we can assume $\HH_1 = \HH$, where $\HH$ is the RKHS constructed in the proof of Theorem 5, so $\HH$ is the completion of $\Span\qty{k(\cdot, x) \mid x \in \XX}$.
 Now we can embed $\HH$ into $\HH_2$ by mapping $k(\cdot, x)$ to $k_x$.
@@
 Therefore, we may assume $\HH_1 \subseteq \HH_2$.
-
-	Now suppose $\HH_1 \subseteq \HH_2$ and decompose each $f \in \HH_2$ as $f = u + v$ with $u \in \HH_1$, $v \in \HH_1\P$.
+%	
+%	Now suppose $\HH_1 \subseteq \HH_2$ and decompose each $f \in \HH_2$ as $f = u + v$ with $u \in \HH_1$, $v \in \HH_1\P$.
 By the previous question we know that for every $x$ there exists a unique $k_x$ with $\ang{k_x, f} = f(x)$ for all $f$, and clearly we have $k_x \in \HH_1$.
-
-	Then for each $f \in \HH_2, x \in \XX$ we have (since $k_x \in \HH_1$) that
+%	
+%	Then for each $f \in \HH_2, x \in \XX$ we have (since $k_x \in \HH_1$) that
+%	\[
+%	f(x) = \ang{k_x, f} = \ang{k_x, u} + \ang{k_x, v} = \ang{k_x, u} = u(x),
+%	\]
+%	which implies $v = 0$. Since this is the case for every $f \in \HH_2$, we conclude that $\HH_1\P = 0$ and therefore $\HH_1 = \HH_2$ (since Hilbert spaces are closed). 
+Let $U = \HH_1 \cap \HH_2$.
+By the previous question, we know that $V$ must contain all functions of the form $k(\cdot, x)$.
+Now, take any $f \in \HH_1$ and write $f = u + v$ with $u \in U$, $v \in U\P$.
+Then we have
 \[
-	f(x) = \ang{k_x, f} = \ang{k_x, u} + \ang{k_x, v} = \ang{k_x, u} = u(x),
+v(x) = \ang{v, k(\cdot, x)} = 0 \quad\text{for all $x$}, 
 \]
-	which implies $v = 0$. Since this is the case for every $f \in \HH_2$, we conclude that $\HH_1\P = 0$ and therefore $\HH_1 = \HH_2$ (since Hilbert spaces are closed). 
+so $v = 0$. Therefore we find $f \in V$, so in particular we have $\HH_1 \subseteq V$. Analogously we have $\HH_2 \subseteq V$, which shows $\HH_1 = \HH_2$.
 \end{proof}

### 2020-10-27 topics_statistical_theory/C2.tex
worked on some example sheets, updated msm and topics summaries (todo: dist and inv prob)

@@
 We can compute 
-\begin{align*}
-\Bias \hat f_n(x) &= \EE[\hat f_n(x)] - f(x) = \EE[K_h(x - X_1)] - f(x) = \int_\RR  K_h(x - z) f(z) \dd{z} \\
-&= (K_h * f)(x) - f(x). 
-\end{align*}
+\begin{align}
+\Bias \hat f_n(x) &= \EE[\hat f_n(x)] - f(x) = \EE[K_h(x - X_1)] - f(x) = \int_\RR  K_h(x - z) f(z) \dd{z} \nonumber \\
+&= (K_h * f)(x) - f(x) \label{eq:bias_computation}. 
+\end{align}
 
@@
 
-\begin{proposition}
+\begin{proposition} \label{prop:variance_bound}
 	Let $\hat{f_n}$ be the kernel density estimator with kernel $K$ and bandwidth $h > 0$ constructed from $X_1, \dotsc, X_n \iid f$. Then for any $x \in \RR, h > 0, n \in \NN$ we have
@@
 \end{proof}
+
+Obtaining a bound on the bias is not at all straightforward: we wil need to introduce conditions on both the density $f$ and the kernel $k$. 
+\begin{definition}
+	Let $I \subseteq \RR$ be an interval, fix $\beta, L > 0$, and let $m \ceq \ceil{\beta} - 1$.
+A function $f \colon \RR \to \RR$ is said to belong to the \emph{Hölder class} $\HH(\beta, L)$ if $f$ is $m$ times differentiable on $I$ and
+	\[
+	\abs{f^{(m)}(x) - f^{(m)}(y)} \leq L \abs{x - y}^{\beta - m} \quad\text{for all $x, y \in I$}. 
+	\]
+	If $I$ is unspecified, we let $I = \RR$.
+
+	The densities in $\HH(\beta, L)$ are denoted by
+	\[
+	\FF(\beta, L) \ceq \qty{f \in \HH(\beta, L) \mid f \geq 0 \text{ and } \int_\RR f \dd{x} = 1}. 
+	\]
+\end{definition}
+
+\begin{definition}
+	Fix $\ell \in \NN$. We say a kernel $k$ is of \emph{order} $\ell$ if $\int_\RR x^j k(x) \dd{x} =0 $ for $j = 1, \dotsc, \ell - 1$. 
+\end{definition}
+
+\begin{remark}
+	Most kernels used in practice are of order 2.
+Note that a kernel of order $\geq 3$ cannot be nonnegative, since we have $\int_\RR x^2k(x) \dd{x} = 0$.
+Therefore, the kernels are not themselves densities and the corresponding kernel density estimate is not guaranteed to be a density.
+\end{remark}
+
+\begin{proposition} \label{prop:bias_bound}
+	Assume that $f \in \FF(\beta, L)$ and that $k$ is a kernel of order $\ell \ceq \ceil{\beta}$, and furthermore assume that
+	\[
+	\mu_\beta(k) \ceq \int_\RR \abs{u}^\beta \abs{k(u)} \dd{u} < \infty. 
+	\]
+	Then the kernel density estimate with bandwidth $h$ and kernel $k$ based on $X_1, \dotsc, X_n \sim f$ satisfies
+	\[
+	\abs{\Bias \hat f_n(x)} \leq \frac{L}{(\ell - 1)!} \mu_\beta(k) h^\beta \quad\text{for all $x \in \RR, h > 0, n \in \NN$}. 
+	\]
+\end{proposition}
+
+\begin{proof}
+	By \cref{eq:bias_computation}, we have 
+	\[
+	\Bias \hat f_n(x) = \frac1h \int_\RR k\qty(\frac{x - z}{h}) f(z) \dd{z} - f(x) = \int_\RR k(u)\qty( f(x - uh) - f(x)) \dd{x}. 
+	\]
+	By applying Taylor's theorem with the Lagrange remainder we obtain, with $m = \ceil{\beta} - 1$, that
+	\[
+	f(x - uh) - f(x) = \sum_{j=1}^{m-1} \frac{(-uh)^j}{j!} f^{(j)}(x) + \frac{(-uh)^m}{m!} f^{(m)}(x - \tau u h) \quad\text{for some $\tau \in [0, 1]$}. 
+	\]
+	Since $\int_\RR u^j k(u) \dd{u} = 0$ for all $j \leq m$, plugging the sum into the integral will give 0. Therefore, we find
+	\[
+	\Bias \hat f_n(x) =  \frac{(-h)^m}{m!} \int_\RR u^m k(u) f^{(m)}(x - \tau uh) \dd{u} = \frac{(-h)^m}{m!} \int_\RR u^m k(u) \qty[f^{(m)}(x - \tau uh)
+- f^{(m)}(x)] \dd{u},	\]
+	where the last inequality follows again from the fact that $k$ is of order $m + 1$.
+
+	Now we use that $f \in \FF(\beta, L)$, and conclude
+	\[
+	\abs{\Bias \hat f_n(x)} \leq \frac{Lh^m}{m!} \int_\RR \abs{u}^m \abs{k(u)} \abs{\tau u h}^{\beta - m} \dd{u} \leq \frac{Lh^\beta}{m!} \int_\RR \abs{u}^\beta \abs{k(u)} \dd{u} = \frac{L}{(\ell - 1)!} \mu_\beta(k) h^\beta,
+	\]	
+	which concludes the proof. 
+\end{proof}
+
+Combining \cref{prop:variance_bound,prop:bias_bound}, we find that
+\[
+\MSE \hat f_n(x) \leq \frac{1}{nh} R(k) \norm{f}_\infty + \frac{L^2}{((\ell - 1)!)^2} \mu^2_\beta(k) h^{2\beta}. 
+\]
+By minimising this w.r.t.\ $h$, we find that the optimal $h$ is given by
+\[
+h_n^* = \qty(\frac{((\ell - 1)!)^2 \norm{f}_\infty R(k)}{2\beta L^2 \mu_\beta^2(k)})^{1/(2\beta + 1)} n^{-1/(2\beta + 1)}, 
+\]
+and the corresponding $\MSE$ is given by
+\[
+\MSE \hat f_n(x) \leq C n^{-2\beta/(2\beta + 1)},
+\]
+where $C$ is independent of $n$. This bound on the rate is slower than the $O(1/n)$ rate found in parametric problems, but such a rate is only obtained when the assumed model is correct.
+
+\subsection{Bounds on the integrated variance and bias}
+To bound the MISE, we will give bounds on the integrated variance and bias. 
+\begin{proposition}
+	Let $\hat{f_n}$ denote the kernel density estimate with bandwidth $h$ and kernel $k$ based on $X_1, \dotsc, X_n \iid P$ (where $P$ is a distribution on $\RR$). Then 
+\end{proposition}

### 2020-10-27 inverse_problems/ex_sheet_1_matlab/ex1_generate_data.m

@@ -1,0 +1,75 @@
+% Copyright 2018 Matthias J. Ehrhardt, Lukas F. Lang
+%
+% This file is part of IP18.
+%
+%    IP18 is free software: you can redistribute it and/or modify
+%    it under the terms of the GNU General Public License as published by
+%    the Free Software Foundation, either version 3 of the License, or
+%    (at your option) any later version.
+%
+%    IP18 is distributed in the hope that it will be useful,
+%    but WITHOUT ANY WARRANTY; without even the implied warranty of
+%    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+%    GNU General Public License for more details.
+%
+%    You should have received a copy of the GNU General Public License
+%    along with IP18. If not, see <http://www.gnu.org/licenses/>.
+%
+% This script takes the image 'trees.tif' and 'kernel.png' and creates a
+% blurry image by convolving the two.
+clear;
+close all;
+clc;
+
+% Load the image.
+u = imread('trees.tif');
+[m, n] = size(u);
+
+% Plot the image with colorbar.
+figure(1);
+colormap gray;
+subplot(2, 2, 1);
+imagesc(u);
+title('Original image u.');
+axis image;
+colorbar;
+
+% Load and create normalised convolution kernel.
+filename = fullfile('data', 'kernel.png');
+k = sum(255 - double(imread(filename)), 3);
+k = k ./ sum(abs(k(:)));
+
+% Plot kernel.
+subplot(2, 2, 2);
+imagesc(k);
+title('Kernel k.');
+axis image;
+colorbar;
+
+% Create zero-padded version of kernel.
+kk = zeros([m, n]);
+kk(1:size(k, 1), 1:size(k, 2)) = k;
+
+% Compute Fourier transforms of kernel and image.
+Fk = fft2(kk);
+Ff = fft2(u);
+
+% Compute convolved image.
+f = ifft2(Ff .* Fk);
+
+% Show data and kernel in Fourier domain.
+subplot(2, 2, 3);
+imagesc(f);
+title('Data f.');
+axis image;
+colorbar;
+subplot(2, 2, 4);
+imagesc(ifftshift(abs(Fk)));
+title('Kernel in Fourier domain (abs).');
+axis image;
+colorbar;
+
+% Save the data and the figure.
+mkdir('results');
+save(fullfile('results', 'ex5b_data.mat'), 'k', 'u', 'f', 'Fk');
+saveas(gcf, fullfile('results', 'ex5b_data.png')); 

### 2020-10-28 distribution_theory/C2.tex
I. Hate. Matlab.

@@
 \end{proof}
+
+\section{Tempered distributions and Fourier analysis}
+\subsection{Functions of rapid decay}
+\begin{definition}
+	For any $f \colon \RR^n \to \CC$ and multi-indices $\alpha, \beta$ we define $\norm{f}_{\alpha, \beta} \ceq \sup_{x \in \RR^n} \abs{x^\alpha \pt^\beta\phi}$.
+
+	We define the \emph{Schwarz space}
+	\[
+	\Cal S(\RR^n) \ceq \qty{f \in C^\infty(\RR^n) : \norm{f}_{\alpha, \beta} < \infty \text{ for all $\alpha, \beta$}}.
+	\]
+
+	We say that a sequence $(\phi_n) \subseteq \SS(\RR^n)$ converges to 0 if $\norm{\phi_m}_{\alpha, \beta} \to 0$ for every $\alpha, \beta$.  
+\end{definition}
+
+\begin{example}
+	The function $x \mapsto \exp(-\norm{x}^2)$ lies in $\SS(\RR^n)$. 
+\end{example}
+
+\begin{proposition}
+	For all $n$ we have that $\SS(\RR^n) \subseteq L^1(\RR^n)$. 
+\end{proposition}
+
+\begin{proof}
+	Let $\phi \in \SS(\RR^n)$, then for all $N \in \NN$ we have
+	\[
+	\int_{\RR^n} \abs{\phi(x)} \dd{x} = \int_{\RR^n} (1 + \norm{x})^{-N} (1 + \norm{x})^N \abs{\phi(x)} \dd{x} \overset?\leq C\sum_{\abs{\alpha} \leq N} \norm{\phi}_{\alpha, 0} \int_{\RR^n} (1 + \norm{x})^{-N} \dd{x}.
+	\]
+
+	Since $\int_{\RR^n} (1 + \norm{x})^{-N} \dd{x}$ is finite for $N$ large enough (??), this proves the claim. 
+\end{proof}
+
+\begin{definition}
+	A linear map $u \colon \SS(\RR^n) \to \CC$ is called a \emph{tempered distribtion} if there exists constants $C, N$ such that
+	\[
+	\abs{\ang{u, \phi}} \leq C \sum_{\abs{\alpha}, \abs{\beta} \leq N} \norm{\phi}_{\alpha, \beta} \quad\text{for all $\phi \in \SS(\RR^n)$}. 
+	\]
+\end{definition}
+This definition is equivalent to sequential continuity.
+
+\subsection{The Fourier transform on $\SS(\RR^n)$}
+
+\begin{convention}
+	We write $D \ceq -i\pt$ and $D^\alpha = (-i)^{\abs{\alpha}} \pt^\alpha$. 
+\end{convention}
+
+\begin{definition}
+	For $f \in L^1(\RR^n)$, define the \emph{Fourier transform} of $f$ by  
+	\[
+	[\FF(f)](\lambda) = \hat f(\lambda) \ceq \int_{\RR^n} e^{-i\lambda x} f(x) \dd{x} \quad\text{where $\lambda \in \RR^n$}. 
+	\]
+\end{definition}
+
+\begin{lemma}
+	If $f \in L^1(\RR^n)$, then $\hat f$ is continuous. 
+\end{lemma}
+
+\begin{proof}
+	If $\lambda_m \to \lambda \in \RR^n$, then by the dominated convergence theorem we have
+	\[
+	\hat f(\lambda_m) = \int_{\RR^n} e^{-i\lambda_m x} f(x) \dd{x} \overset{\text{DCT}}= \int_{\RR^n} e^{-i\lambda x} f(x) \dd{x} = \hat f(\lambda),
+	\]
+	where we were able to use the dominated convergence theorem since the integrand is absolutely bounded by $\abs{f}$ and $f \in L^1$. 
+\end{proof}
+
+It turns out that this idea generalises: the faster the function $f$ decays, the smoother the Fourier transform $\hat f$ is.
+
+\begin{lemma}
+	For $\phi \in \SS(\RR^n)$, we have $\FF[D_x^\alpha \phi](\lambda) = \lambda^\alpha \hat\phi(\lambda)$ and $\FF[x^\beta\phi](\lambda) = (-1)^{\abs{\beta}} D_\lambda^\beta \hat\phi(\lambda)$. 
+\end{lemma}
+
+\begin{proof}
+	Since $\abs{x^\alpha D^\beta\phi} \to 0$ as $\norm{x} \to \infty$, we have using integration by parts
+	\begin{align*}
+		\FF[D_\lambda^\alpha\phi](\lambda) &= \int_{\RR^n} e^{-i\lambda x} D_x^\alpha \phi(x) \dd{x} \\
+		&= (-1)^{\abs{\alpha}} \int_{\RR^n} D_x^\alpha(e^{-i\lambda x}) \phi(x) \dd{x} \\
+		&= \lambda^\alpha\hat\phi(\lambda). 
+	\end{align*}
+
+	For the second part of the lemma, by differentiation under the integral sign we get
+	\begin{align*}
+		 \FF[x^\beta \phi](\lambda) &=  \int_{\RR^n} e^{-i\lambda x} x^\beta \phi(x) \dd{x} \\
+		 &= \int_{\RR^n} ((-D_\lambda)^\beta e^{-i\lambda x}) \phi(x) \dd{x} \\
+		 &= (-1)^{\abs{\beta}} D_\lambda^\beta \hat\phi(\lambda).
+	\end{align*}
+\end{proof}
+
+We define the \emph{inverse Fourier transform} by 
+\[
+\FF^{-1}[\hat f](x)  \ceq \frac{1}{(2\pi)^n} \int_{\RR^n} e^{i\lambda x} \hat f(\lambda) \dd{\lambda}. 
+\]
+We will now show that on $\SS(\RR^n)$, the inverse Fourier transform is indeed an inverse:
+\begin{theorem}
+	The Fourier transform $\FF$ defines a continuous isomorphism from $\SS(\RR^n)$ to itself. 
+\end{theorem}
+
+\begin{proof}
+	Let $\phi \in \SS(\RR^n)$. First, we show that $\hat\phi \in \SS(\RR^n)$: by the previous lemma we have for multi-indices $\alpha, \beta$ that
+	\begin{align}
+		\abs{\lambda^\alpha (-D_\lambda)^\beta \hat\phi(\lambda)} &= \abs{\lambda^\alpha \FF[x^\beta\phi](\lambda)} = \abs{\FF[D^\alpha_x (x^\beta\phi)](\lambda)} = \abs{\int_{\RR^n} e^{-i\lambda x} D^\alpha(x^\beta\phi) \dd{x}} \nonumber \\
+		&\leq \int_{\RR^n} \abs{D^\alpha (x^\beta\phi)} \dd{x} \label{eq:fourier_schwarz_estimate},
+	\end{align}
+which is finite since $D^\alpha (x^\beta\phi)$ is also a Schwarz function and therefore integrable.
+
+From the previous lemma we also infer that $\hat\phi$ is smooth, so indeed we have $\hat\phi \in \SS(\RR^n)$. 
+From \cref{eq:fourier_schwarz_estimate} it is also easily seen that if $\phi_m \to 0$ in $\SS(\RR^n)$, then $\hat\phi_m \to 0$ in $\SS(\RR^n)$ also, which shows that $\FF$ is continuous.
+
+To prove surjectivity and injectivity, we will show that $\FF^{-1}[\FF[f]](x) = f(x)$ (???). Indeed we have
+\begin{align*}
+	\FF^{-1} [\FF[\phi]](x) &= \frac{1}{(2\pi)^n} \int_{\RR^n} \int_{\RR^n} e^{i\lambda (x-y)} \phi(y) \dd{y}\dd{\lambda} \\
+	&= \frac{1}{(2\pi)^n} \lim_{\eps \downarrow 0} \int_{\RR^n} \int_{\RR^n} e^{i \lambda ( x- y) - \eps\norm{\lambda}^2} \phi(y) \dd{y}\dd{\lambda} \\
+	&\overset\star= \frac{1}{(2\pi)^n}\lim_{\eps \downarrow 0} \int_{\RR^n} \phi(y)\int_{\RR^n} e^{i\lambda(x- y) - \eps\norm{\lambda}^2} \dd{\lambda} \dd{y},
+	%&= \frac{1}{(2\pi)^n} \int_{\RR^n}\int_{\RR^n} e^{i\lambda(x- y)} \dd{\lambda} \dd{y}
+\end{align*}
+where $\star$ follows from Fubini's theorem.
+
+Now we note that 
+\[
+\int_{\RR^n} e^{i\lambda(x-y) - \eps\norm{\lambda}^2} \dd{\lambda} = \prod_{i=1}^n \int_{\RR^n} e^{i\lambda_j(x_j - y_j) - \eps\lambda_j^2} \dd{\lambda} \overset{\star\star}= \prod_{i=1}^n \qty(\frac\pi e)^{1/2} e^{- \frac{(x_i - y_i)^2}{4\eps}} = \qty(\frac\pi\eps)^{n/2} e^{- \frac{\norm{x - y}^2}{4\eps}}.
+\] 
+To explain $\star\star$, \TODO.
+
+and plugging that into the above yields 
+\begin{align*}
+\FF^{-1}[\FF[\phi]](x) &= \lim_{\eps\downarrow 0} 2^{-n}(\pi \eps)^{-n/2} \int_{\RR^n} \phi(y) e^{-\norm{x-y}^2/(4\eps)} \dd{y} \\
+&\overset{\star\star\star}= \pi^{-n/2} \lim_{\eps\downarrow0}\int_{\RR^n} \phi(x - 2\sqrt\eps y') e^{-\norm{y'}^2} \dd{y} \\
+&\overset{\text{DCT}}=\pi^{-n/2} \phi(x) \int_{\RR^n} e^{-\norm{y'}^2} \dd{y} = \phi(x),
+\end{align*}
+where $\star\star\star$ follows from the substitution $x - y = 2\sqrt\eps y'$.
+
+Finally, continuity of $\FF^{-1}$ is easily shown with an argument analogous to that for continuity of $\FF$. 
+\end{proof}

### 2020-10-28 distribution_theory/distribution_ES1.tex
I. Hate. Matlab.

@@
 \]
+This shows that $\pt_x u - \pt_y u = 0$, or equivalently, that $u$ satisfies the wave equation. 
 \end{proof}
 
-This shows that $\pt_x u - \pt_y u = 0$, or equivalently, that $u$ satisfies the wave equation. 
+\begin{question}
+	Compute $\Delta(\norm{x}^{2-n})$ in $\DD'(\RR^n)$ for $n \geq 3$, i.e.\ compute
+	\[
+	\ang{\Delta (\norm{x}^{2-n}), \phi} = \ang{\norm{x}^{2-n}, \Delta\phi} = \int \frac{\Delta\phi}{\norm{x}^{n-2}} \dd{x}
+	\]
+	for arbitrary $\phi \in \DD(\RR^n)$. Note that $\norm{x} = (x_1^2 +\dotsb +x_n^2)^{1/2}$ and $\Delta = \sum_i (\pdv{x_i})^2$.
+
+	\emph{Hint:} use $\int\dd{x} = \int_{\norm{x}\leq\eps}\dd{x} + \int_{\norm{x}>\eps}\dd{x}$ and treat each integral separately. 
+\end{question}
+
+\begin{solution}
+	We follow the hint: let $\eps > 0$, then we first compute
+	\[
+	\int_{\norm{x} > \eps} \frac{\Delta\phi}{\norm{x}^{n-2}} \dd{x} = \sum_{i=1}^n \int_{\norm{x} > \eps} \frac{\pdv[2]{x_i} \phi}{\norm{x}^{n-2}}\dd{x} = \sum_{i=1}^n \int_{\norm{x} > \eps} \phi \cdot \pdv[2]{x_i} \norm{x}^{2-n} \dd{x}.
+	\]
+	Now, it is easily computed that $\pdv{x_i} \norm{x} = \frac{x_i}{\norm{x}}$, and therefore 
+	\begin{align*}
+	\pdv[2]{x_i} \norm{x}^{2-n} &= \pdv{x_i} (2-n) x_i \norm{x}^{-n} = (2-n) \qty(\norm{x}^{-n} - n x_i^2 \norm{x}^{-n-2}) \\
+	&= (n-2)\norm{x}^{-n} \qty(n \qty(\frac{x_i}{\norm{x}})^2 - 1). 
+	\end{align*}
+\end{solution}
+
 \end{document}

### 2020-10-28 inverse_problems/inverse_ES1.tex
I. Hate. Matlab.

@@
 \end{align*}
-where $(A^* v)(x) = \int_\Omega k(x, y) v(y) \dd{y}$ as required. Here $\star$ follows from Fubini's theorem (\TODO justify).
+where $(A^* v)(x) = \int_\Omega k(x, y) v(y) \dd{y}$ as required. Here $\star$ follows from Fubini's theorem.
 
@@
 \[
-(A_p u)(y) = \sum_{\abs{\alpha} \leq n} c_\alpha \int_\Omega x_1^{\alpha_1} x_2^{\alpha_2} y_1^{\alpha_3} y_2^{\alpha_4} u(x) \dd{x} = \sum_{\abs{\alpha} \leq n} c_\alpha \qty(\int_\Omega x_1^{\alpha_1} x_2^{\alpha_2} \dd{x}) y_1^{\alpha_3} y_2^{\alpha_4},
+(A_p u)(y) = \sum_{\abs{\alpha} \leq n} c_\alpha \int_\Omega x_1^{\alpha_1} x_2^{\alpha_2} y_1^{\alpha_3} y_2^{\alpha_4} u(x) \dd{x} = \sum_{\abs{\alpha} \leq n} c_\alpha \qty(\int_\Omega x_1^{\alpha_1} x_2^{\alpha_2} u(x) \dd{x}) y_1^{\alpha_3} y_2^{\alpha_4},
 \]
@@
 
-	% \item From Fourier theory it is known that $\qty{s_n(x) = \sqrt 2\sin(nx) : n \in \NN}$ and $\qty{1} \cup \qty{c_n(x) = \sqrt 2\cos(nx) \mid n \in \NN}$ are orthonormal bases of $L^2[0, 1]$. 
-	%Note that 
-	%\[
-	%(A s_n)(y) = \int_0^y s_n(x) \dd{x} = \sqrt 2 \int_0^y \sin(nx) \dd{x} = \frac{\sqrt 2}{n} \qty(1 - \cos(ny)) = \frac{\sqrt 2}{n} c_0(y) - \frac1n c_n(y) 
-	%\]
-	\item 	Note that $A$ is compact, since we can write $(Au)(y) = \int_{[0, 1]} k(x, y) u(x) \dd{x}$ with $k(x, y) = \ind_{x \leq y}$.
+	\item 	In example 2.2.12, it is shown that for this operator, the Picard criterion is
+	\begin{equation} \label{eq:picard}
+	2\sum_{j=1}^\infty  \sigma_j^{-2} \qty(\int_0^1 f(s) \sin(\sigma_j^{-1} s) \dd{s})^2,
+	\end{equation}
+	where $\sigma_j = \frac2{(2j - 1)\pi}$.
 
-	To apply the Picard criterion we must find the singular values and right singular vectors of $A$, which are equal to the square roots of the eigenvalues of $AA^*$ and the eigenvectors of $AA^*$.
-
-	\begin{align*}
-		\ang{Au, v} &= \int_0^1 (Au)(y)\cdot v(y) \dd{y} \\
-		&= \int_0^1 \int_0^y u(x) \dd{x} v(y) \dd{y} \\
-		&= \int_0^1 \int_0^y u(x) v(y) \dd{x} \dd{y} \\
-		&= \int_0^1 \int_x^1 u(x) v(y) \dd{y} \dd{x} \\ 
-		&= \int_0^1 u(x) \int_x^1 v(y) \dd{y} \dd{x} \\
-		&= \ang{u, A^* v}
-	\end{align*}
-where $v(x) = \int_x^1 v(y) \dd{y}$.  Therefore, we find
-\begin{align*}
-	(AA^* u)(x) = \int_0^x \int_y^1 u(z) \dd{z} \dd{y}.
-\end{align*}
-Now we solve the eigenequation. Firstly, note that $\Nul(A) = \qty{0}$, so let $\lambda^2 > 0$. Then we have
-\begin{align}
-	\int_0^x \int_y^1 u(z) \dd{z} \dd{y} &= \lambda^2 u(x) \\
-	\int_x^1 u(z) \dd{z} &= \lambda^2 u'(x) \\
-	- u(x) &= \lambda^2 u''(x). 
-\end{align}
-Furthermore, from (1) we infer $u(0) = 0$ while from (2) we infer $u'(1) = 0$. 
-The general solution to (3) is given by $u(x) = a \cos(\lambda^{-1} x) + b \sin(\lambda^{-1} x)$. 
-Plugging $u(0) = 0$ gives $a = 0$, and $u'(1) = 0$ gives
+	We compute
 	\[
-b\lambda^{-1} \cos(\lambda^{-1}) = 0 \implies b = 0 \text{ or } \lambda^{-1} = \pi\qty(n - \frac12) \text{ for some $n \in \ZZ_{>0}$}.
+	\int_0^1 f(s) \sin(\sigma_j^{-1} s) \dd{s} = \int_{1/2}^1 \sin(\sigma_j^{-1} s) \dd{s} = \sigma_j \qty[ \cos(\frac12 \sigma_j^{-1}) - \cos(\sigma_j^{-1})]. 
+	\]
+	We have
+	\[
+	\cos(\sigma_j^{-1}) = \cos(\frac{(2j - 1)\pi}{2}) = 0 \quad\text{and}\quad \cos(\frac12\sigma_j^{-1}) = \cos(\frac{(2j-1)\pi}{4}) = \pm\frac1{\sqrt2 }. 
 	\]
 
-Our singular values are therefore $\frac{1}{(n - \frac12)\pi}$ ($n \in \ZZ_{>0}$) with right singular vectors $y_n(x) = \sqrt 2\sin((n - \frac12) \pi x)$. 
-We now compute
+	Plugging this into \cref{eq:picard} gives that
 	\[
-\ang{f, y_n} = \sqrt 2\int_{1/2}^1 \sin(\qty(n - \frac12) \pi x) \dd{x} = ...
+	2\sum_{j=1}^\infty  \sigma_j^{-2} \qty(\int_0^1 f(s) \sin(\sigma_j^{-1} s) \dd{s})^2 = 2 \sum_{j=1}^\infty \sigma_j^{-2} (\sigma_j^2/2) = \sum_{j=1}^\infty 1 = \infty,
 	\]
+	so $f$ does not satisfy the Picard criterion and therefore $f \in \clos{\Ran(A)} \setminus \Ran(A)$. 
+%	Note that $A$ is compact, since we can write $(Au)(y) = \int_{[0, 1]} k(x, y) u(x) \dd{x}$ with $k(x, y) = \ind_{x \leq y}$. 
+%	
+%	To apply the Picard criterion we must find the singular values and right singular vectors of $A$, which are equal to the square roots of the eigenvalues of $AA^*$ and the eigenvectors of $AA^*$. 
+%	
+%
+%	\begin{align*}
+%		\ang{Au, v} &= \int_0^1 (Au)(y)\cdot v(y) \dd{y} \\
+%		&= \int_0^1 \int_0^y u(x) \dd{x} v(y) \dd{y} \\
+%		&= \int_0^1 \int_0^y u(x) v(y) \dd{x} \dd{y} \\
+%		&= \int_0^1 \int_x^1 u(x) v(y) \dd{y} \dd{x} \\ 
+%		&= \int_0^1 u(x) \int_x^1 v(y) \dd{y} \dd{x} \\
+%		&= \ang{u, A^* v}
+%	\end{align*}
+%where $v(x) = \int_x^1 v(y) \dd{y}$.  Therefore, we find
+%\begin{align*}
+%	(AA^* u)(x) = \int_0^x \int_y^1 u(z) \dd{z} \dd{y}.
+%\end{align*}
+%Now we solve the eigenequation. Firstly, note that $\Nul(A) = \qty{0}$, so let $\lambda^2 > 0$. Then we have
+%\begin{align}
+%	\int_0^x \int_y^1 u(z) \dd{z} \dd{y} &= \lambda^2 u(x) \\
+%	\int_x^1 u(z) \dd{z} &= \lambda^2 u'(x) \\
+%	- u(x) &= \lambda^2 u''(x). 
+%\end{align}
+%Furthermore, from (1) we infer $u(0) = 0$ while from (2) we infer $u'(1) = 0$. 
+%The general solution to (3) is given by $u(x) = a \cos(\lambda^{-1} x) + b \sin(\lambda^{-1} x)$. 
+%Plugging $u(0) = 0$ gives $a = 0$, and $u'(1) = 0$ gives
+%\[
+%b\lambda^{-1} \cos(\lambda^{-1}) = 0 \implies b = 0 \text{ or } \lambda^{-1} = \pi\qty(n - \frac12) \text{ for some $n \in \ZZ_{>0}$}.
+%\]
+%
+%Our singular values are therefore $\frac{1}{(n - \frac12)\pi}$ ($n \in \ZZ_{>0}$) with right singular vectors $y_n(x) = \sqrt 2\sin((n - \frac12) \pi x)$. 
+%We now compute
+%\[
+%\ang{f, y_n} = \sqrt 2\int_{1/2}^1 \sin(\qty(n - \frac12) \pi x) \dd{x} = ...
+%\]
 
 \item The Moore-Penrose inverse of $A$ is discontinuous.
-This can be seen by theorem 2.1.11: in (a) we have constructed a sequence $(f_n) \subseteq \Ran(A)$ that converges, and in (b) we have shown that its limit lies outside of $\Ran(A)$.
-Therefore, $\Ran(A)$ is not closed, so $A\D$ is discontinuous.
+This can be seen by theorem 2.1.11: we have in (b) an element $f \in \clos{\Ran(A)} \setminus \Ran(A)$, so   $\Ran(A)$ is not closed, so $A\D$ is discontinuous.
 
@@
 
-			\item If $A\T A = I$, then the columns $\vec u_1, \dotsc, \vec u_n$ of $A$ are orthogonal.
+			\item 
+%			If $A\T A = I$, then the columns $\vec u_1, \dotsc, \vec u_n$ of $A$ are orthogonal.
 In particular, $A$ is injective (so $\Nul(A) = \qty{0}$) and $\Ran(A) = \Span\qty{\vec u_1, \dotsc, \vec u_n}$.
 Extend $\vec u_1, \dotsc, \vec u_n$ to an orthonormal basis $\vec u_1, \dotsc, \vec u_m$ of $\RR^m$, then we must have
+%			\[
+%			A\D\qty(\sum_{i=1}^m \alpha_i \vec u_i) = \mqty(\alpha_1 \\ \vdots \\ \alpha_n) \in \RR^n.
+%			\]
+%			From this expression it is easily seen that $A\D = A\T$.
+			We claim $A\D = A\T$. For this, we check the first two Moore-Penrose equations: for the first one, we have $A\T A = I_n = P_{\RR^n} = P_{\Nul(A)\P}$ since $\Nul(A) = \qty{0}$.  
+			For the second one, let $\vec u_1, \dotsc, \vec u_n \in \RR^m$ be the columns of $A$ (which are orthonormal), then we have
 			\[
-			A\D\qty(\sum_{i=1}^m \alpha_i \vec u_i) = \mqty(\alpha_1 \\ \vdots \\ \alpha_n) \in \RR^n.
+			AA\T = \sum_{i=1}^n \vec u_i \vec u_i\T = P_{\Span\qty{\vec u_1, \dotsc, \vec u_n}} = P_{\clos{\Ran(A)}}. 
 			\]
-			From this expression it is easily seen that $A\D = A\T$. 
 		\end{enumerate}

### 2020-10-28 modern_statistical_methods/C2.tex
I. Hate. Matlab.

@@
 	\]
-	Let $\hat\beta = \hat\beta_\lambda^\Rm{L}$, then with probability at least $1 - 2p^{-A^2/2 - 1}$ we have that the MSPE satisfies
+	Let $\hat\beta = \hat\beta_\lambda^\Rm{L}$, then with probability at least $1 - 2p^{-(A^2/2 - 1)}$ we have that the MSPE satisfies
 	\[
@@
 	\]
-	In (\TODO refer), we will show that $\PP(\Omega) \geq 1 - 2p^{-(A^2/2 - 1)}$, which completes the proof. 
+	In \cref{lem:probability_omega_bound}, we will show that $\PP(\Omega) \geq 1 - 2p^{-(A^2/2 - 1)}$, which completes the proof. 
 \end{proof}
+
+\subsection{Concentration inequalities}
+Let $W$ be any random variable and $\phi \colon \RR\to[0,\infty)$ strictly increasing. Then by Markov's inequality we have
+\[
+\PP(W \geq t) = \PP(\phi(W) \geq \phi(t)) \leq \frac{\EE[\phi(W)]}{\phi(t)}. 
+\]
+Plugging in $\phi(x) = e^{\alpha x}$ (for some $\alpha > 0$), we get
+\[
+\PP(W \geq t) \leq e^{-\alpha t} \EE[e^{\alpha W}] = e^{-\alpha t} M_W(\alpha).
+\]
+Now we can take the infimum over all $\alpha$ on the right-hand side, and we get what is called the \emph{Chernoff bound}:
+\[
+\PP(W \geq t) \leq \inf_{\alpha > 0} e^{-\alpha t} M_W(\alpha). 
+\]
+
+\begin{definition}
+	A random variable  $W$ with mean $\mu$ is called \emph{sub-Gaussian with parameter $\sigma > 0$} or $\sigma$-sub-Gaussian if 
+	\[
+	M_{W - \mu} \leq M_{N(0, \sigma^2)} \quad\text{or equivalently}\quad \EE[e^{\alpha(W - \mu)}] \leq e^{\alpha^2\sigma^2/2} \text{ for all $\alpha \in \RR$}. 
+	\]
+\end{definition}
+
+We need the following lemma, which characterises an important class of sub-Gaussian random variables:
+\begin{lemma}[Hoeffding]
+	If $W$ is a mean-zero random variable which takes values in $[a, b]$, then $W$ is sub-Gaussian with parameter $(b-a)/2$. 
+\end{lemma}
+
+By Chernoff bounding, we obtain for a $\sigma$-sub-Gaussian random variable $W$ that \[\PP(W - \mu \geq t) \leq e^{-t^2/(2\sigma^2)}.\]
+
+\begin{proposition}
+	Let $W_1, \dotsc, W_n$ be independent, mean-zero random variables where $W_i$ is $\sigma_i$-sub-Gaussian.
+For any $\gamma \in \RR^n$, the random variable $\gamma\T W$ is sub-Gaussian with parameter $\qty(\sum_i \sigma_i^2\gamma_i^2)^{1/2}$.
+\end{proposition}
+\begin{proof}
+	Since the $W_i$ are independent we have for all $\alpha \in \RR$ that
+	\begin{align*}
+		\EE[e^{\alpha \sum_i \gamma_i W_i}] = \prod_{i=1}^n \EE[e^{\alpha \gamma_i W_i}]  \leq \prod_{i=1}^n e^{\alpha^2\gamma_i^2\sigma_i^2/2} = e^{\alpha^2 \sum_i \gamma_i^2\sigma_i^2/2}.
+	\end{align*}
+\end{proof}
+
+\begin{lemma} \label{lem:probability_omega_bound}
+	Suppose $\eps_1, \dotsc, \eps_n$ are independent mean-zero $\sigma$-sub-Gaussian random variables and let $\lambda \ceq A\sigma \sqrt{\log(p)/n}$. Then
+	\[
+	\PP\qty(\frac{\norm{X\T\eps}_\infty}{n} \leq \lambda) \geq 1 - 2p^{-(A^2/2 - 1)}. 
+	\]
+\end{lemma}
+
+\begin{proof}
+	We have
+	\[
+	\PP\qty(\frac{\norm{X\T\eps}_\infty}{n} > \lambda) = \PP\qty(\bigcup_i \frac{\abs{X_i\T\eps}}{n} > \lambda) \leq \sum_i \PP\qty(\frac{\abs{X_i\T\eps}}{n} > \lambda). 
+	\]
+	By the previous proposition, both $X_i\T\eps$ and $-X_i\T\eps$ are sub-Gaussian with parameter $\sigma/\sqrt n$, so we have
+	\[
+	\sum_i \PP\qty(\frac{\abs{X_i\T\eps}}{n} > \lambda) \leq 2p \exp(\frac{-\lambda^2}{2(\sigma/\sqrt n)^2}) = 2p\exp(-A^2\log(p)/2) = 2 p^{-(A^2/2 - 1)}.
+	\]
+\end{proof}

### 2020-10-28 inverse_problems/ex_sheet_1_matlab/ex4b_deblur.m

@@ -1,0 +1,6 @@
+data = load('results/ex5b_data.mat');
+Fk = data.Fk;
+f = data.f;
+Ff = fft2(f);
+u =  (2 * pi)^(-1) * ifft2(Ff ./ Fk);
+imshow(u)

### 2020-10-29 inverse_problems/C3.tex
moaaaaar updates

\section{Variational regularisation}
\subsection{Background}
\subsubsection{Banach spaces and weak convergence}
A \emph{Banach space} $\XX$ is a complete normed vector space. We define the \emph{dual space} $\XX^* \ceq \LL(X, \RR)$, and for $p \in \XX^*, u \in \XX$ we usually write $\ang{p, u}$ instead of $p(u)$.  For any $A \in \LL(\XX, \YY)$ we define the \emph{adjoint} $A^* \colon \YY^* \to \XX^*$ by $\ang{A^*p, u} \ceq \ang{p, Au}$ for all $p \in \XX^*, u \in \XX$. The dual space $\XX'$ is equipped with the norm
\[
\norm{p}_{\XX^*} \ceq \sup_{\norm{u}\leq 1} \ang{p, u}, 
\]
and with this norm $\XX^*$ is a Banach space.

The \emph{bi-dual space} is defined as $\XX^{**} \ceq (\XX^*)^*$. The mapping $E \colon \XX \to (\XX)^{**}$ defined by $\ang{E(u), p} \ceq \ang{p, u}$ is a continuous linear isometry, and we will regard $\XX$ as a subspace of $\XX^{**}$ using this isometry. If $\XX = \XX^{**}$ (i.e., $E$ is surjective), the space $\XX$ is called \emph{reflexive}.
A space $\XX$ is called \emph{separable} if $\XX$ has a countable dense subset.

A sequence $(u_k) \subseteq \XX$ is said to \emph{converge weakly} to $u \in \XX$, denoted $u_k \cw u$, if $\ang{p, u_k} \to \ang{p, u}$ for all $p \in \XX^*$.

A sequence $(p_k) \subseteq \XX^*$ is said to \emph{converge weakly-*} to $p \in \XX'$, denoted $p_k \cws p$, if $\ang{p_k, u} \to \ang{p, u}$ for all $u \in \XX$.

\begin{theorem}
	Let $\XX$ be Banach, then the unit ball is compact in $\XX^*$ w.r.t.\ the weak-$*$ topology. If $\XX$ is separable, then the weak-$*$ topology is metrisable and every bounded sequence in $\XX^*$ has a weakly-$*$ convergent subsequence. 
\end{theorem}

\begin{theorem}
	Let $\XX$ be reflexive, then every bounded sequence in $\XX$ has a weakly convergent subsequence. 
\end{theorem}

We define $\rext \ceq \RR\cup \qty{\pm\infty}$.

\begin{definition}
	Let $\XX$ be a Banach space with topology $\tau_X$. A functional $E \colon \XX \to\rext$ is said to be \emph{sequentially lower-semicontinuous with respect to $\tau_\XX$} or simply $\tau_\XX$-LSC if
	\[
	E(u) \leq \liminf_{n\to\infty} E(u_n)\quad\text{if $u_n \overset\tau\to u$}. 
	\]

	Specifically, if $\tau_\XX$ is the weak topology, then $E$ is called \emph{weakly} LSC. If $\tau_\XX$ is the topology induced by the norm on $\XX$, then $E$ is called \emph{strongly} LSC or simply LSC.
\end{definition}

\subsubsection{Convex analysis}
\begin{definition}
	Let $C \subseteq \XX$. Then the \emph{characteristic function} of $C$ is defined as 
	\[
	\chi_C(u) \ceq \begin{cases} 0, &u \in C, \\ \infty,& u \notin C. \end{cases}
	\]
\end{definition}
Using characteristic functions, we have $\min_{u \in C} E(u) = \min_{u \in \XX} E(u) + \chi_C(u)$.

\begin{definition}
	Let $E \colon \XX \to \rext$, then the \emph{effective domain} is $\dom(E) \ceq \qty{u \mid E(u) < \infty}$.

	The functional $E$ is called \emph{proper} if $\dom(E) \neq\emptyset$. 
\end{definition}

\begin{definition}
	A functional $E \colon \XX \to \rext$ is called:
	\begin{enumerate}
		\item \emph{convex} if for all $u \neq v \in \XX$ and $\lambda \in (0, 1)$ we have $E(\lambda u + (1-\lambda) v) \leq \lambda E(u) + (1-\lambda) E(v)$;
		\item \emph{strictly convex} if the above inequality is strict;
		\item \emph{strongly convex} with constant $\theta > 0$ if $u \mapsto E(u) - \theta \norm{u}^2$ is convex. 
	\end{enumerate}

Note that $C \subseteq \XX$ is a convex set if and only if $\chi_C$ is a convex function. 
\end{definition}

\begin{lemma}
	Nonnegative linear combinations of convex functionals are convex. If one of the components is strictly convex, then the nonnegative linear combination is also strictly convex. 
\end{lemma}

\begin{definition}
	Let $E \colon \XX \to \rext$ be a functional. We define the \emph{Fenchel conjugate}
	\[
	E^* \colon \XX^* \to \rext \colon p \mapsto \sup_{u \in \XX} \qty[ \ang{p, u} - E(u)]. 
	\]
\end{definition}

\begin{theorem}
	For any $E \colon \XX \to \RR$ we have $E^{**} \restr_{\XX} \leq E$. If $E$ is proper and LSC, then $E^{**} \restr_{\XX} = E$. 
\end{theorem}

\begin{definition}
	A functional $E \colon \XX \to \rext$ is called \emph{subdifferentiable} at $u \in \XX$ if there exists a $p \in \XX^*$ such that
	\[
	E(v) \geq E(u) + \ang{p, v - u} \quad\text{for all $v \in \XX$}. 
	\]
	In this case, we call $p$ a \emph{subgradient} of $E$ at position $u$. The collection of all subgradients of $E$ at $u$ is denoted by $\pt E(u)$ and is called the \emph{subdifferential} of $E$ at $u$. 
\end{definition}

\begin{lemma}
	Let $E \colon \XX \to \rext$ be convex, then $E$ is subdifferentiable at all points $u \in \dom(E)$. If $E$ is also proper, then $E$ is not subdifferentiable at any $u \notin \dom(E)$. 
\end{lemma}

\begin{theorem}
	Let $E \colon \XX \to \rext$ be proper and convex and $u \in \dom(E)$. Then $\pt E(u)$ is convex and weakly-$*$ compact in $\XX^*$. 
\end{theorem}

\begin{theorem}
	Let $E, F$ be proper LSC convex functionals and $u \in \dom(E) \cap \dom(F)$ such that at least one of $E$ and $F$ is continuous at $u$. Then $\pt(E + F)(u) = \pt E(u) + \pt F(u)$. 
\end{theorem}

\begin{theorem}
	Let $E$ be convex. Then $u$ is a global minimiser of $E$ if and only if $0 \in \pt E(u)$. 
\end{theorem}

\begin{definition}
	Let $E$ be convex, $u, v \in \XX$, $E(v) < \infty$ and $q \in \pt E(v)$. Then the \emph{Bregman distance} of $E$ between $u$ and $v$ is defined as
	\[
	D_E^q(u, v) \ceq E(u) - E(v) - \ang{q, u -v} \geq 0. 
	\]

	If we also have $E(u) < \infty, p \in \pt E(u)$, then we define the \emph{symmetric Bregman distance}
	\[
	D_E^{p, q}(u, v) \ceq D_E^p(v, u) + D_E^q(u, v) = \ang{p - q, u - v}. 
	\]
\end{definition}

\begin{definition}
	A functional $E$ is called \emph{absolutely one-homogeneous} if $E(\lambda u) = \abs{\lambda} E(u)$ for all $\lambda \in \RR, u \in \XX$. 
\end{definition}

\begin{proposition}
	Let $E$ be a convex, proper and absolutely one-homogeneous, and $p \in \pt E(u)$. Then:
	\begin{enumerate}
		\item $E(u) = \ang{p, u}$;
		\item $D^p(v, u) = E(v) - \ang{p, v}$ for all $v \in \XX$;
		\item $E^*(p) = \chi_{\pt E(0)}(p)$.
	\end{enumerate}  
\end{proposition}

Furthermore, we have the following:
\begin{proposition}
	Let $E$ be proper, convex, and absolutely one-homogeneous, and let $u \in \XX$. Then $p \in \pt E(u)$ if and only if $p \in \pt E(0)$ and $\ang{p, u} = E(u)$. 
\end{proposition}

\subsubsection{Minimisers}
\begin{definition}
	We say that $u^* \in \XX$ is a \emph{minimiser} of a  functional $E$ if $u$ minimises $E$ and $E(u) < \infty$. 
\end{definition}

\begin{definition}
	A functional $E$ is called \emph{coercive} if $\norm{u_j} \to \infty \implies \abs{E(u_j)} \to \infty$. 
\end{definition}

\begin{lemma}
	Let $E$ be proper, coercive and bounded from below. Then $\inf_{u \in \XX} E(u) > -\infty$ and there exists a (bounded) minimising sequence $(u_j)$ with $E(u_j) \to \inf_u E(u)$. \end{lemma}

\begin{theorem}[Direct method]
	Let $\XX$ be Banach and $\tau_\XX$ a topology on $\XX$ such that any bounded sequence in $\XX$ has a $\tau_\XX$ convergent subsequence. Then any proper, bounded from below, coercive, $\tau_X$-LSC functional has a minimiser.  
\end{theorem}

\begin{proof}
	Since $E$ is bounded from below, we have $\inf_u E(u) > -\infty$, so there exists a bounded minimising sequence $(u_j)$, which we can assume is $\tau_\XX$ convergent with limit $u^*$ after taking a subsequence if necessary. By lower-semicontinuity of $E$ we have
	\[
	E(u^*) \leq \liminf_{k\to\infty} E(u_j) = \lim_{j\to\infty} E(u_j) = \inf_u E(u),
	\]
	so $u^*$ is a minimiser. 
\end{proof}

\begin{theorem}
	If a strictly convex functional has a minimiser, it is unique. 
\end{theorem}

\begin{proof}
	Suppose $u \neq v$ are two minimisers, then by strict convexity, we have $E(\frac12 u + \frac12 v) < E(u)$, a contradiction. 
\end{proof}

\subsubsection{Duality in convex optimisation}
Consider the \emph{primal} optimisation problem 
\[
(P) \ceq \inf_{u \in \XX} E(Au) + F(u),
\]
where $E, F$ are proper, convex and LSC, and $A \in \BB(\XX, \YY)$. Since $E$ is convex and LSC, we have $E = E^{**}$ so we can rewrite the primal problem as the \emph{saddle point problem}
\[
\inf_{u \in \XX} \sup_{\eta \in \YY^*} \ang{\eta, Au} - E^*(\eta) + F(u).
\]
Since $\inf\sup \geq \sup\inf$ always holds we have
\[
(P) \geq \sup_{\eta \in \YY^*} \inf_{u \in \XX} \ang{\eta, y} - E^*(\eta) + F(u) = \sup_{\eta \in \YY^*} -E^*(\eta) - F^*(-A^*\eta) \eqqcolon (D). 
\]

The problem (D) is called the \emph{dual problem}, and the fact that $(D) \leq (P)$ is called \emph{weak duality}. The value $(P) - (D)$ is called the \emph{duality gap}, and if $(P) = (D)$, we speak of \emph{strong duality}.

We have the following:
\begin{theorem}
	Suppose the function $E(Au) + F(u)$ is proper, convex, LSC and coercive. Suppose also that there exists $u_0 \in \XX$ s.t.\ $F(u)< \infty$, $E(Au_0) < \infty$, and $E(y)$ is continuous at $y = A u_0$. Then:
	\begin{enumerate}
		\item The dual problem (D) has at least one solution $\hat\eta$;
		\item There is no duality gap;
		\item If $(P)$ has an optimal solution $\hat u$, then we have
		\[
		A^*\hat\eta \in \pt F(\hat u), \quad -\hat\eta \in \pt E(A\hat u). 
		\]
	\end{enumerate}
\end{theorem}

### 2020-10-29 topics_statistical_theory/C2.tex
moaaaaar updates

--- 2020-10-21 topics_statistical_theory/C2.tex
@@
 \begin{definition}
-	A Borel measurable function $k \colon \RR \to \RR$ is called a \emph{kernel} if it satisfies $\int_\RR K(x) \dd{x} = 1$.
+	A Borel measurable function $K \colon \RR \to \RR$ is called a \emph{kernel} if it satisfies $\int_\RR K(x) \dd{x} = 1$.
 
@@
 	By \cref{eq:variance_computation} we have
+	\begin{align}
+	\Var \hat f_n(x) &\leq \frac1n (K_h^2 * f)(x) = \frac{1}{nh^2} \int_\RR K^2\qty(\frac{x - z}{h}) f(z) \dd{z} = \frac{1}{nh} \int_\RR K^2(u) f(x - uh) \dd{u}  \label{eq:var_bound}\\
+	&\leq \frac1{nh} \norm{f}_\infty \int_\RR K^2(u) \dd{u} = \frac{1}{nh} \norm{f}_\infty R(K) .\nonumber 
+	\end{align}
+\end{proof}
+
+Obtaining a bound on the bias is not at all straightforward: we wil need to introduce conditions on both the density $f$ and the kernel $K$. 
+\begin{definition}
+	Let $I \subseteq \RR$ be an interval, fix $\beta, L > 0$, and let $m \ceq \ceil{\beta} - 1$.
+A function $f \colon \RR \to \RR$ is said to belong to the \emph{Hölder class} $\HH(\beta, L)$ if $f$ is $m$ times differentiable on $I$ and
+	\[
+	\abs{f^{(m)}(x) - f^{(m)}(y)} \leq L \abs{x - y}^{\beta - m} \quad\text{for all $x, y \in I$}. 
+	\]
+	If $I$ is unspecified, we let $I = \RR$.
+
+	The densities in $\HH(\beta, L)$ are denoted by
+	\[
+	\FF(\beta, L) \ceq \qty{f \in \HH(\beta, L) \mid f \geq 0 \text{ and } \int_\RR f \dd{x} = 1}. 
+	\]
+\end{definition}
+
+\begin{definition}
+	Fix $\ell \in \NN$. We say a kernel $K$ is of \emph{order} $\ell$ if $\int_\RR x^j k(x) \dd{x} =0 $ for $j = 1, \dotsc, \ell - 1$. 
+\end{definition}
+
+\begin{remark}
+	Most kernels used in practice are of order 2.
+Note that a kernel of order $\geq 3$ cannot be nonnegative, since we have $\int_\RR x^2K(x) \dd{x} = 0$.
+Therefore, the kernels are not themselves densities and the corresponding kernel density estimate is not guaranteed to be a density.
+\end{remark}
+
+\begin{proposition} \label{prop:bias_bound}
+	Assume that $f \in \FF(\beta, L)$ and that $K$ is a kernel of order $\ell \ceq \ceil{\beta}$, and furthermore assume that
+	\[
+	\mu_\beta(K) \ceq \int_\RR \abs{u}^\beta \abs{K(u)} \dd{u} < \infty. 
+	\]
+	Then the kernel density estimate with bandwidth $h$ and kernel $K$ based on $X_1, \dotsc, X_n \sim f$ satisfies
+	\[
+	\abs{\Bias \hat f_n(x)} \leq \frac{L}{(\ell - 1)!} \mu_\beta(K) h^\beta \quad\text{for all $x \in \RR, h > 0, n \in \NN$}. 
+	\]
+\end{proposition}
+
+\begin{proof}
+	By \cref{eq:bias_computation}, we have 
+	\[
+	\Bias \hat f_n(x) = \frac1h \int_\RR K\qty(\frac{x - z}{h}) f(z) \dd{z} - f(x) = \int_\RR K(u)\qty( f(x - uh) - f(x)) \dd{x}. 
+	\]
+	By applying Taylor's theorem with the Lagrange remainder we obtain, with $m = \ceil{\beta} - 1$, that
+	\[
+	f(x - uh) - f(x) = \sum_{j=1}^{m-1} \frac{(-uh)^j}{j!} f^{(j)}(x) + \frac{(-uh)^m}{m!} f^{(m)}(x - \tau u h) \quad\text{for some $\tau \in [0, 1]$}. 
+	\]
+	Since $\int_\RR u^j K(u) \dd{u} = 0$ for all $j \leq m$, plugging the sum into the integral will give 0. Therefore, we find
+	\[
+	\Bias \hat f_n(x) =  \frac{(-h)^m}{m!} \int_\RR u^m K(u) f^{(m)}(x - \tau uh) \dd{u} = \frac{(-h)^m}{m!} \int_\RR u^m K(u) \qty[f^{(m)}(x - \tau uh)
+- f^{(m)}(x)] \dd{u},	\]
+	where the last inequality follows again from the fact that $K$ is of order $m + 1$.
+
+	Now we use that $f \in \FF(\beta, L)$, and conclude
+	\[
+	\abs{\Bias \hat f_n(x)} \leq \frac{Lh^m}{m!} \int_\RR \abs{u}^m \abs{K(u)} \abs{\tau u h}^{\beta - m} \dd{u} \leq \frac{Lh^\beta}{m!} \int_\RR \abs{u}^\beta \abs{K(u)} \dd{u} = \frac{L}{(\ell - 1)!} \mu_\beta(K) h^\beta,
+	\]	
+	which concludes the proof. 
+\end{proof}
+
+Combining \cref{prop:variance_bound,prop:bias_bound}, we find that
+\[
+\MSE \hat f_n(x) \leq \frac{1}{nh} R(K) \norm{f}_\infty + \frac{L^2}{((\ell - 1)!)^2} \mu^2_\beta(K) h^{2\beta}. 
+\]
+By minimising this w.r.t.\ $h$, we find that the optimal $h$ is given by
+\[
+h_n^* = \qty(\frac{((\ell - 1)!)^2 \norm{f}_\infty R(K)}{2\beta L^2 \mu_\beta^2(K)})^{1/(2\beta + 1)} n^{-1/(2\beta + 1)}, 
+\]
+and the corresponding $\MSE$ is given by
+\[
+\MSE \hat f_n(x) \leq \qty(\frac{\norm{f}_\infty^{2\beta} R(K)^{2\beta} L^2 \mu_\beta^2(K) \qty[(2\beta)^{2\beta + 1} + 1]}{((\ell - 1)!)^2 (2\beta)^{2\beta}}) n^{-2\beta/(2\beta + 1)},
+\]
+This $O(n^{-2\beta/(2\beta + 1)})$ bound on the rate is slower than the $O(1/n)$ rate found in parametric problems, but such a rate is only obtained when the assumed model is correct.
+
+We can strengthen this as follows:
+\begin{theorem}
+	Assume that $K$ is a kernel of order $\ell \ceq \ceil{\beta}$ and that $\mu_\beta(K)$ and $R(K)$ are both finite.
+Fix $\alpha > 0$ and let $h = \alpha n^{-1/(2\beta + 1)}$.
+Then there exists $C > 0$, independent of $n$, such that
+	\[
+	\sup_{x \in \RR} \sup_{f \in \FF(\beta, L)} \MSE \hat f_n(x) \leq C n^{-2\beta/(2\beta + 1)}. 
+	\]
+\end{theorem}
+
+\begin{proof}
+	We will show that the class $\FF(\beta, L)$ is uniformly bounded in supremum norm.
+Let $K^*$ be a bounded kernel of order $\ell$ (see example sheet \TODO), then by the previous proposition with $h = 1$ we have by nonnegativity of $f$ that
 	\begin{align*}
-	\Var \hat f_n(x) &\leq \frac1n (K_h^2 * f)(x) = \frac{1}{nh^2} \int_\RR K^2\qty(\frac{x - z}{h}) f(z) \dd{z} = \frac{1}{nh} \int_\RR K^2(u) f(x - uh) \dd{u} \\
-	&\leq \frac1{nh} \norm{f}_\infty \int_\RR K^2(u) \dd{u} = \frac{1}{nh} \norm{f}_\infty R(K). 
+		f(x) &\leq \abs{f(x)- \int_{-\infty}^\infty K^*(x - z) f(z) \dd{z}} + \abs{\int_{-\infty}^\infty K^*(x-z) f(z) \dd{z}}.	\\
+		&\leq \abs{\Bias \hat f_{n, K^*}(x)} + \norm{K^*}_\infty \int_{-\infty}^\infty  f(z) \dd{z} \\
+		&\leq \frac{L}{(\ell - 1)!} \mu_\beta(K^*) + \norm{K^*}_\infty,
+	\end{align*}
+and this bound is independent of $f$ and $x$.
+
+Now we have
+\[
+\MSE \hat f_n(x) \leq \frac{R(K) \norm{f}_\infty}{nh} + \frac{L^2}{((\ell - 1)!)^2} \mu^2_\beta(K) h^{2\beta} \leq C n^{-2\beta/(2\beta + 1)}. 
+\]
+\end{proof}
+
+\subsection{Bounds on the integrated variance and bias}
+To bound the MISE, we will give bounds on the integrated variance and bias. 
+\begin{proposition}
+	Let $\hat{f_n}$ denote the kernel density estimate with bandwidth $h$ and kernel $K$ based on $X_1, \dotsc, X_n \iid P$ (where $P$ is a distribution on $\RR$). Then
+	\[
+	\int_{-\infty}^\infty \Var \hat f_n(x) \dd{x} = \frac1{nh} R(K). 
+	\]
+\end{proposition}
+
+\begin{proof}
+	We have by Fubini and \cref{eq:var_bound} that
+	\begin{align*}
+		\int_{-\infty}^\infty \Var\hat f_n(x) \dd{x} &\leq \frac1{nh^2} \int_{-\infty}^\infty \int_{-\infty}^\infty K^2\qty(\frac{x - z}{h}) f(z) \dd{z} \dd{x} = \frac1{nh^2} \int_{-\infty}^\infty f(z) \int_{-\infty}^\infty K^2\qty(\frac{x-z}{h}) \dd{x} \dd{z} \\
+		&= \frac1{nh} R(K) \int_{-\infty}^\infty f(z) \dd{z} = \frac1{nh} R(K). 
 	\end{align*}
 \end{proof}
+
+\begin{recap}
+	Let $[a, b] = I \subseteq \RR$ be an interval, then $f \colon I \to \RR$ is called \emph{absolutely continuous} if, for every $\eps > 0$, there exists $\delta > 0$ such that, whenever $(x_1, y_1), \dotsc, (x_m, y_m)$ are disjoint subintervals of $I$ with $\sum_{i=1}^m (y_i - x_i) < \delta$, we have $\sum_{i=1}^m \abs{f(y_i) - f(x_i)} < \eps$.
+
+	It is known that absolute continuity is equivalent to being differentiable Lebesgue almost everywhere with a so-called \emph{weak derivative} $f'$ that satisfies $f(x) = f(a) + \int_a^x f'(t) \dd{t}$ for all $x \in [a, b]$.
+\end{recap}
+
+\begin{recap}
+	The \emph{generalised Minkowski inequality} states that any Borel measurable function $g \colon \RR^2 \to \RR$ we have that
+	\[
+	\int_\RR \qty(\int_\RR g(u, x) \dd{u})^2 \dd{x} \leq \qty(\int_\RR \qty(\int_\RR g^2(u, x) \dd{x})^{1/2} \dd{u})^2. 
+	\] 
+\end{recap}
+To obtain bounds on the integrated squared bias, we will require smoothness conditions w.r.t.\ the $L^2(\RR)$ norm. 
+\begin{definition}
+	Fix $\beta, L > 0$ and let $m \ceq \ceil{\beta} - 1$.
+The \emph{Nikolski class} $\Nn(\beta, L)$ consists of functions $f \colon \RR \to \RR$ that are $(m-1)$ times differentiable and for  which $f^{(m-1)}$ is absolutely continuous with weak derivative $f^{(m)}$ satisfying
+	\[
+	\qty{\int_{-\infty}^\infty \qty[ f^{(m)}(x + t) - f^{(m)}(x)]^2 \dd{x}}^{1/2} \leq L \abs{t}^{\beta - m} \quad\text{for all $t \in \RR$}. 
+	\]
+
+	The densities in $\Nn(\beta, L)$ are denoted by $\FF_\Nn(\beta, L)$. 
+\end{definition}
+
+\begin{proposition}
+	Fix $\beta, L> 0$ and let $K$ be a kernel of order $\ell \ceq \ceil\beta$.
+Let $\hat f_n$ denote the KDE with kernel $K$ and bandwidth $h$ based on $X_1, \dotsc, X_n \iid f \in \FF_\Nn(\beta, L)$.
+Then we have
+	\[
+	\int_{-\infty}^\infty \Bias^2 \hat f_n(x) \dd{x} \leq \frac{L^2}{((\ell - 1)!)^2} \mu_\beta^2(K) h^{2\beta}. 
+	\]
+\end{proposition}
+
+\begin{proof}
+	\TODO write this out (integration + taylor expansion + 2x minkowski). 
+\end{proof}
+
+Putting everything together, we obtain the following:
+\begin{theorem}
+	Fix $\beta, L > 0$, and let $K$ be a kernel of order $\ell = \ceil\beta$ with $R(k)$ and $\mu_\beta(K)$ finite.
+Let $\hat f_n$ be the KDE with kernel $K$ and bandwidth $h$ based on $X_1, \dotsc, X_n \iid f \in \FF_\Nn(\beta, L)$.
+Then we have
+	\[
+	\MISE\hat f_n \leq \frac{R(K)}{nh} + \frac{L^2}{((\ell - 1)!)^2} \mu_\beta^2(K) h^{2\beta}. 
+	\]
+	In particular, fixing $\alpha > 0$ and taking $h = \alpha n^{-1/(2\beta + 1)}$, there exists $C > 0$ independent of $n$ such that
+	\[
+	\sup_{f \in \FF_\Nn(\beta, L)} \MISE \hat f_{n, h, K} \leq C n^{-2\beta/(2\beta + 1)}. 
+	\]
+\end{theorem}
+
+\subsection{Bandwidth selection}
+The choice of bandwidth in the previous theorem is not practical since we have not specified $\alpha$ and $\beta$ is typically unknown.
+One possible approach is \emph{least squares cross validation}.
+For this, note that minimising the $\MISE$ is equivalent to minimising
+\[
+\MISE(\hat f_n) - \int_\RR f^2(x) \dd{x} = \EE\qty[ \int_\RR \hat f_n^2(x) \dd{x}] - 2\EE\qty[ \int_\RR \hat f_n(x) f(x) \dd{x}],
+\]
+and it can be checked that an unbiased estimator for the above is given by 
+\[
+\Rm{LSCV}(h) \ceq \int_\RR \hat f_n^2(x) \dd{x} - \frac2n \sum_{i=1}^n \hat f_{n, -i}(X_i),
+\]
+where $\hat f_{n, -i}$ is the KDE based on all observations except $X_i$. We now choose $h$ such that LSCV is minimised.

### 2020-10-29 inverse_problems/ex_sheet_1_matlab/ex4b_deblur.m

@@ -1,6 +1,27 @@
 data = load('results/ex5b_data.mat');
+colormap gray;
+u_org = data.u;
+% u_org_noise = imnoise(u_org, 'gaussian');
+
+% Restore data from array
 Fk = data.Fk;
 f = data.f;
+[m, n] = size(f);
+r = normrnd(0, 0.1, m, n);
+
+% Add noise to f (REMOVE THIS IF TRUE RECONSTRUCTION IS DESIRED)
+f = f + r;
+
+% Compute inverse Fourier transform
 Ff = fft2(f);
-u =  (2 * pi)^(-1) * ifft2(Ff ./ Fk);
-imshow(u)
+% prod = Ff ./ Fk;
+prod = Ff .* conj(Fk) ./ (abs(Fk) .^ 2);
+u_inv =  ifft2(prod);
+
+% Show image
+figure(1)
+subplot(2, 1, 1)
+imagesc(u_org)
+
+subplot(2, 1, 2)
+imagesc(u_inv)

### 2020-10-29 distribution_theory/distribution_ES1.tex
why is distribution theory so *****ing DIFFICULT

@@
 \DeclareMathOperator{\ord}{ord}
+\DeclareMathOperator{\prv}{p.\!v.}
 
@@
 	\end{align*}
+\TODO finish
 \end{solution}
 
+\begin{question}
+	Let $(f_k)$ be the sequence of smooth functions defined by $f_k(x) = \frac1\pi \frac k{(kx)^2 + 1}$. Prove that $f_k \to \delta_0$ in $\DD'(\RR)$. Compute the limits
+	\begin{enumerate}[(a)]
+		\item $\lim_{k\to\infty} k^2 x e^{-k^2 x^2}$,
+		\item $\lim_{k\to\infty} k^3 e^{ikx}$,
+		\item $\lim_{k\to\infty} \frac{\sin(kx)}{\pi x}$,
+	\end{enumerate}
+in $\DD'(\RR)$. 
+\end{question}
+
+\begin{solution}
+	It is easily seen that every $f_k$ is locally integrable. Furthermore, noting that $\arctan(kx)' = \frac{k}{(kx)^2 + 1}$, we see by the dominated convergence theorem that
+	\begin{align*}
+		\lim_{k\to\infty}\ang{f_k, \phi} &= \frac1\pi \lim_{k\to\infty} \int_\RR \frac{k}{(kx)^2 + 1} \phi(x) \dd{x} = -\frac1\pi \lim_{k\to\infty} \int_\RR \arctan(kx) \phi'(x) \dd{x} \\
+		&= -\frac1\pi \int_\RR \lim_{k\to\infty} \arctan(kx) \phi'(x) \dd{x} = -\frac1\pi \qty(2 \frac\pi2 \int_0^\infty \phi'(x) \dd{x} ) = \phi(0), 
+	\end{align*}
+which proves $f_k \to \delta_0$.
+
+\begin{enumerate}[(a)]
+	\item We have $- (\frac12 e^{-k^2x^2})' = k^2 x e^{-k^2 x^2}$ and therefore
+	\begin{align*}
+		\lim_{k\to\infty} \int_\RR k^2 x e^{-k^2x^2} \phi(x) \dd{x} &= \frac12 \lim_{k\to\infty}  \int_\RR  e^{-k^2 x^2} \phi'(x) \dd{x} = \frac12 \int_\RR \lim_{k\to\infty} e^{-k^2 x^2} \phi'(x) \dd{x} = 0,
+	\end{align*}
+so the sequence converges to 0 in $\DD'(\RR)$.
+
+\item We have 
+\begin{align*}
+	\lim_{k\to\infty}  k^3 \int_\RR e^{ikx} \phi(x) \dd{x} = \lim_{k\to\infty} 2\pi k^3 \Cal F^{-1}[\phi](k). 
+\end{align*}
+Since $\phi$ is a Schwarz function, its inverse Fourier transform is also a Schwarz function, and in particular $k^3 \Cal F^{-1}[\phi](k) \to 0$ as $k \to \infty$.
+
+\item This sequence converges to $\delta_0$, although I have no idea how to prove it.
+
+\end{enumerate}
+\end{solution}
+
+\begin{question}
+	Compute the limit
+	\[
+	\lim_{k\to\infty} \frac12 + \sum_{m=1}^k \cos(\pi m x)
+	\]
+	in $\DD'(-1, 1)$. 
+\end{question}
+
+\begin{solution}
+	Note that \[
+	f_k(x) \ceq 
+	\frac12 + \sum_{m=1}^k \cos(\pi m x) = \frac12\qty(1 + e^{i \pi x} + e^{-i\pi x} + \dotsb + e^{i \pi m x} + e^{-i\pi m x}) = \frac12 \sum_{m=-k}^k (e^{i\pi x})^m.
+	\]
+	By viewing the last term as a geometric series $\frac12 e^{-ik\pi x} \sum_{m=0}^{2k+1} (e^{ix})^m$ we can compute that 
+	\[
+	f_k(x) = \frac{\sin((n + \frac12) \pi x)}{2\sin(\frac12\pi x)}. 
+	\]
+	It can be shown (???) that $f_k \to \delta_0$ in $\DD'(\RR)$. 
+\end{solution}
+
+\begin{question}
+	We define the \emph{principal value} of $1/x$, written $\prv(1/x)$, by
+	\[
+	\ang*{\prv(1/x), \phi} = \lim_{\eps \to 0} \int_{\abs{x} > \eps} \frac{\phi(x)}{x} \dd{x} \quad\text{for all $\phi \in \DD(\RR)$}. 
+	\]
+	Prove that $\prv(1/x) \in \DD'(\RR)$ and that $\ord(\prv(1/x)) = 1$. Show that
+	\[
+	\lim_{\eps \to 0} \frac{1}{x - i\eps} = \prv\qty(\frac1x) + i\pi \delta_0 \quad\text{in $\DD'(\RR)$}. 
+	\]
+\end{question}
+
+\begin{proof}
+	First we must show that $\prv(1/x)$ is well-defined.  ??
+\end{proof}
+
+\begin{question}
+	??
+\end{question}
+\begin{proof}
+	??
+\end{proof}
+
+\begin{question}
+	??
+\end{question}
+
+\begin{proof}
+	??
+\end{proof}
+
+\begin{question}
+	Define the distribution $u \in \EE'(\RR^3)$ by the locally integrable function $u(x) = \ind_{\abs{x}\leq 1}$.
+Prove that $-\sum_i x_i (\pdv{u}{x_i}) = \dd{\sigma_2}$ in $\EE'(\RR^3)$, where $\dd{\sigma_2}$ is the surface element on the sphere $S^2 \subseteq \RR^3$.
+\end{question}
+
+\begin{proof}
+	We have
+	\begin{align*}
+		\ang*{- \sum_i x_i \pdv{u}{x_i}, \phi} &= - \sum_i \ang*{\pdv{u}{x_i}, x_i\phi} = \sum_i \ang*{u, \pdv{x_i}(x_i \phi)} = \ang*{u, \sum_i \pdv{x_i}(x_i\phi)} \\
+		&= \int_{\norm{x} \leq 1} \sum_i \pdv{x_i}(x_i\phi) \dd{x}
+	\end{align*}
+\end{proof}
 \end{document}

### 2020-10-29 modern_statistical_methods/msm_ES2.tex
why is distribution theory so *****ing DIFFICULT

--- 2020-10-20 modern_statistical_methods/msm_ES1.tex
@@
 
-\title{Modern Statistical Methods --- Example Sheet 1} % subject
+\newcommand\Nul{\Cal N}
+\newcommand\Ran{\Cal R}
+
+\title{Modern Statistical Methods --- Example Sheet 2} % subject
 \author{Lucas Riedstra}
@@
 \maketitle
+\begin{question}
+	Let $Y \in \RR^n$ be a vector of responses, $\Phi \in \RR^{n \times p}$ a design matrix, $J \colon [0, \infty) \to [0, \infty)$ a strictly increasing function and $c \colon \RR^n \to \RR^n$ some cost function.
+Set $K = \Phi\Phi\T$.
+Show, without using the representer theorem, that $\hat\theta$ minimises
+	\[
+	Q_1(\theta) \ceq c(Y, \Phi\theta) + J(\norm{\theta}_2^2)
+	\]
+	over $\theta \in \RR^p$ if and only if $\Phi\hat\theta = K\hat\alpha$ and $\hat\alpha$ minimises
+	\[
+	Q_2(\alpha) \ceq c(Y, K\alpha) + J(\alpha\T K\alpha)
+	\]
+	over $\alpha \in \RR^n$. 
+\end{question}
 
-\begin{qbox}
-    Consider minimising the following objective involving response $Y \in \RR^n$ and design matrix $X \in \RR^{n \times p}$ over $(\mu, \beta) \in \RR \times \RR^p$: 
+\begin{proof}
+	Let $\hat\theta$ be a minimiser of $Q_1$, and write $\hat\theta = \Phi\T\hat\alpha + \hat\beta$ with $\Phi\T \hat\alpha \in \Nul(\Phi)\P = \Ran(\Phi\T)$, $\hat\beta \in \Nul(\Phi)$.
+
+	Noting that $K\hat\alpha = \Phi\Phi\T \hat\alpha = \Phi\hat\theta$ and  $\norm{\Phi\T\hat\alpha} = \alpha\T K\alpha$ we see
 	\[
-    \norm{Y - \mu \Bf{1} - X\beta}_2^2 + J(\beta).
+	Q_1(\theta) = c(Y, K\hat\alpha) + J(\alpha\T K\alpha + \|\hat\beta\|^2),
 	\]
-    Here $J \colon \RR^p \to \RR$ is an arbitrary penalty function.
-Suppose $\bar{X_k} = 0$ for $k = 1, \dotsc, p$.
-Assuming that a minimiser $(\hat\mu, \hat\beta)$ exists, show that $\hat\mu = \bar Y$.
-Now take $J(\beta) = \lambda \norm{\beta}_2^2$ so we have the ridge regression objective.
-Show that
-    \[
-    \hat\beta = (X\T X + \lambda I)^{-1} X\T Y.
-    \] 
-    From here onwards, whenever we refer to ridge regression, we will assume $X$ has had its columns mean-centred. 
-\end{qbox}
-
-\begin{solution}
-    If a minimiser $\hat\mu$ exists, then the derivative of the objective must be 0. We expand the objective function as
-    \begin{align*}
-    \norm{Y - \mu\Bf{1} - X\beta}_2^2 + J(\beta) &= \sum_{i=1}^n \qty(Y_i - \mu - (X\beta)_i)^2 + J(\beta) \\
-    &= \sum_{i=1}^n \qty(Y_i - \mu - \sum_{j=1}^p X_{ij}\beta_j)^2. 
-    \end{align*}
-Now, differentiating w.r.t.\ $\mu$ yields
-\begin{align*}
-\pdv{\mu} \norm{Y - \mu\Bf{1} - X\beta}_2^2 + J(\beta) &= -2 \sum_{i=1}^n \qty( Y_i - \mu - \sum_{j=1}^p X_{ij}\beta_j) \\
-&= -2 \qty(\sum_{i=1}^n Y_i  - n\mu - \sum_{i=1}^n \sum_{j=1}^p X_{ij}\beta_j) \\
-&= -2 \qty(\sum_{i=1}^n Y_i  - n\mu - \sum_{j=1}^p \beta_j (\sum_{i=1}^p X_{ij})) \\
-&= -2 \qty( \sum_{i=1}^n Y_i - n\mu), 
-\end{align*}
-since $\sum_{i=1}^p X_{ij} = 0$ for all $i$ by assumption.
-Now, setting the derivative to 0 shows that $\hat\mu = \frac1n \sum_{i=1}^n Y_i = \bar Y$ is the minimiser.
-Since the second derivative of the objective function is constantly $+2$, this $\hat\mu$ is indeed a minimum.
-
-The computation of $\hat\beta$ is given in the lecture notes. 
-\end{solution}
+	and therefore it is necessary that $\hat\beta = 0$. The claim follows. 
+\end{proof}
 
 \begin{question}
-    Consider performing ridge regression when $Y = X \beta^0 + \eps$, where $X \in \RR^{n \times p}$ has full column rank, and $\Var(\eps) = \sigma^2 I$.
-Let the SVD of $X$ be $UDV\T$ and write $U\T X \beta^0 = \gamma$.
-Show that
-    \[
-    \frac1n \EE\norm{X \beta^0 - X \hat\beta_\lambda^R}_2^2 = \frac1n \sum_{j=1}^p \qty(\frac\lambda{\lambda + D_{jj}^2})^2 \gamma_j^2 + \frac{\sigma^2}{n} \sum_{j=1}^p \frac{D_{jj}^4}{(\lambda + D_{jj}^2)^2}.
-    \]
+	Let $x, x' \in \RR^p$ and let $\psi \in \qty{-1, 1}^p$ be a random vector with independent components taking values $-1, 1$ each with probability $1/2$.
+Show that $\EE(\psi\T x \psi\T x') = x\T x'$.
+Construct a random feature map $\hat\phi \colon \RR^p \to \RR$ such that $\EE\qty{\hat\phi(x)\hat\phi(x')} = (x\T x)^2$.
 \end{question}
@@
 \begin{solution}
-    We write $\hat\beta \ceq \hat\beta_\lambda^R$. We write out
+	We have 
+	\begin{align*}
+		\psi\T x \psi\T x' = \qty(\sum_i \psi_i x_i) \qty(\sum_j \psi_j x'_j) = \sum_i x_i x_i' + 2 \sum_{i < j} \psi_i \psi_j x_i x_j'. 
+	\end{align*}
+	Noting that for $i \neq j$ we have $\EE[\psi_i \psi_j] = \EE[\psi_i] \EE[\psi_j] = 0$ it follows that $\EE[\psi\T x \psi\T x'] = \sum_i x_i x_i' = x\T x'$.
+
+	Let $\psi_*$  be an identical independent copy of $\psi$ and define $\hat\phi(x) = \psi\T x \psi_*\T x$. Then we find
 	\[
-    X\hat\beta = UDV\T (V\T D^2 V + \lambda I)^{-1} V\T D U\T Y =  UD^2 (D^2 + \lambda I)^{-1} U\T Y = \sum_j \frac{D_{jj}^2}{\lambda + D_{jj}^2} U_j U_j\T Y. 
-    \]
-    Since $X\beta^0 = UDV\T \beta^0$ lies in the range of $U$, we have $X\beta^0 = \sum_j U_j U_j\T X\beta^0$. 
-    \begin{align*}
-        X\beta^0 - X\hat\beta &= \sum_j U_j U_j\T X \beta^0 -  \sum_j \frac{D_{jj}^2}{\lambda + D_{jj}^2} U_j U_j\T Y \\
-        &= \sum_j \qty(U_j\T X \beta^0 - \frac{D_{jj}^2 U_j\T Y}{\lambda + D_{jj}^2}) U_j \\
-        &= \sum_j \qty(\frac{D_{jj}^2 U_j\T (X\beta^0 - Y) + \lambda U_j\T X \beta^0}{\lambda + D_{jj}^2})
-    \end{align*}
-and therefore
-\begin{align*}
-    \frac1n \EE\norm{X\beta^0 - X\hat\beta}_2^2 &= \frac1n \sum_j \EE \qty(\frac{D_{jj}^2 U_j\T (X\beta^0 - Y) + \lambda U_j\T X \beta^0}{\lambda + D_{jj}^2})^2 \\
-    &\overset{*}= \frac1n \sum_j \frac{D_{jj}^4}{(\lambda + D_{jj}^2)^2} \EE[(U_j\T (X\beta^0 - Y))^2] + \frac1n \sum_j \qty(\frac{\lambda}{\lambda + D_{jj}^2})^2  (U_j\T X \beta^0)^2 \\
-&\overset{**}= \frac{\sigma^2}{n} \sum_j \frac{D_{jj}^4}{(\lambda + D_{jj}^2)^2}  + \frac1n \sum_j \qty(\frac{\lambda}{\lambda + D_{jj}^2})^2  \gamma_j^2.
-\end{align*}
-Note that the cross-terms in $(*)$ disappear since they are linear combinations of $\EE(X\beta^0 - Y) = 0$, while the equality $**$ holds since 
-\[
-\EE[(U_j\T(X \beta^0 - Y))^2] = \Var(U_j\T (X\beta^0 - Y)) = U_j\T \Var(X\beta^0 - Y) U_j = \sigma^2 U_j\T U_j = \sigma^2. 
+	\EE[\hat\phi(x)\hat\phi(x')] = \EE[\psi\T x \psi\T x']\EE[\psi_*\T x \psi_*\T x'] = (x\T x')^2. 
 	\]
@@
 \begin{question}
-    Show that the ridge regression estimates can be otained by ordinary least squares regression on an augmented data set with $\sqrt\lambda I$ added to the botten of $X$, and $p$ zeroes added to the end of the response $Y$.
-\end{question}
-
-\begin{solution}
-    Let $\tilde X, \tilde Y$ be the new data set, then our ordinary least squares objective function is
-    \begin{align*}
-    L(\beta) & \ceq \norm{\tilde Y - \tilde X \beta}^2 \\
-    &= \norm{Y - X\beta}^2 + \norm{\sqrt\lambda\beta}^2 \\
-    &= \norm{Y - X\beta}^2 + \lambda \norm{\beta}^2,
-    \end{align*}
-which is exactly our ridge regression objective function. 
-\end{solution}
-
-\begin{question}
-    In the following, assume that forming $AB$ where $A \in \RR^{a \times b}, B \in \RR^{b \times c}$ requires $O(abc)$ computational operations, and that if $M \in \RR^{d \times d}$ is invertible, then forming $M^{-1}$ requires $O(d^3)$ operations.
-    \begin{enumerate}[(a)]
-        \item Suppose we wish to apply ridge regression to data $(Y, X) \in \RR^n \times \RR^{n \times p}$ with $n \gg p$.
-A complication is that the data is split into $m$ separate datasets of size $n/m \in \NN$,
+	Let $\XX = \pow(\qty{1, \dotsc, p})$ and $z, z' \in \XX$.
+Let $k$ be the Jaccard similarity kernel.
+Let $\pi$ be a random permutation of $\qty{1, \dotsc, p}$.
+Let $M = \min\qty{\pi(j) \mid j \in z}$, $M' = \min\qty{\pi(j) \mid j \in z'}$.
+Show that
 	\[
-        Y = \mqty( Y^{(1)} \\ \vdots \\ Y^{(m)}), \quad X = \mqty( X^{(1)} \\\vdots\\X^{(m)}),
+	\PP(M = M') = k(z, z'),
 	\]
-        with each dataset located on a different server.
-Moving large amounts of data between servers is expensive.
-Explain how one can produce ridge estimates $\hat\beta_\lambda$ by communicating only $O(p^2)$ numbers from each server to some central server.
-What is the total order of the computation time required at each server, and at the central server for your approach?
-
-        \item Now suppose instead that $p \gg n$ and it is instead the variable sthat are split across $m$ servers, so each server has only a subset of $p/m \in \NN$ variables for each observation, and some central server stores $Y$.
-Explain how one can obtain the fitted values $X\hat\beta_\lambda$ communicating only $O(n^2)$ numbers from each server to the central server.
-What is the total order of the computation time required at each server, and at the central server for your approach?
-    \end{enumerate}
-\end{question}
-
-\begin{solution}
-    \begin{enumerate}[(a)]
-        \item Note that
-        \[
-        X\T Y = \mqty( X^{(1)\top} \dotsc X^{(m)\top}) \mqty( Y^{(1)}\\\vdots\\Y^{(m)}) = \sum_{i=1}^m X^{(i)\top} Y^{(i)}. 
-        \]
-        Each of the $X^{(i)\top} Y^{(i)}$ can be computed at the data server (requiring $O(np)$ computation) and produces $p$ numbers per data server.
-
-        Analogously, we have
-        \[
-        X\T X = \sum_{i=1}^m X^{(i)\top} X^{(i)},
-        \]
-        which can also be computed at the data server (requiring $O(np^2)$ computation) and produces $p^2$ numbers per data server.
-In the end, each server will therefore have $O(p^2)$ numbers to send, having done $O(np^2)$ computations.
-
-        When each server sends their numbers to the central data server, the central data server can then compute $(X\T X + \lambda I)^{-1}$ using $O(p^3)$ computations and then multiply that with $X\T Y$ which will require $O(np)$ computations.
-
-        \item In this case, we have 
-        \[
-        X = \mqty( X^{(1)} \dotsb X^{(m)} ). 
-        \]
-        We use the alternative form of  $\hat\beta_\lambda$, namely
-        \[
-        X \hat\beta_\lambda^R = XX\T (XX\T + \lambda I)^{-1} Y.
-        \]
-
-        Now, we have $XX\T = \sum_{i=1}^m X^{(i)} X^{(i)\top}$, and each $X^{(i)} X^{(i)\top}$ can be computed at the data server, requiring $O(n^2 p)$ computations and producing $O(n^2)$ numbers to send.
-
-        After sending these numbers over to the central server, the inverse $(XX\T + \lambda I)^{-1}$ can be computed in $O(n^3)$ time, then multiplied from the left with $XX\T$ in $O(n^3)$ time, and multiplied with $Y$ in $O(n^2)$ time.
-The total amount of computation at the central server will therefore be $O(n^3)$ as well.
-    \end{enumerate}
-\end{solution}
-
-\begin{question}
-	Prove Proposition 4 in the notes.
-\emph{Hint: for part (ii) it may help to consider the eigendecompositions of positive semi-definite matrices $K^{(1)}$ and $K^{(2)}$ derived from kernels $k_1$ and $k_2$ in the form $K^{(1)} = PDP\T = \sum_{i=1}^n P_iP_i\T D_{ii}$ for example.
-}
+	when $z, z' \neq \emptyset$.
+Now let $\psi \in \qty{-1, 1}^p$ be a random vector with i.i.d.\ components taking the values -1 or 1, each with probability $1/2$.
+By considering $\EE[\psi_M\psi_{M'}]$ show that the Jaccard similarity kernel is indeed a kernel.
+Explain how we can use the ideas above to approximate kernel ridge regression with Jaccard similarity, when $n$ is very large (you may assume none of the data points are the empty set).
 \end{question}
@@
 \begin{proof}
-	Proposition 4 is the following:
-	\begin{mdframed}
-		Suppose $k_1, k_2, \dotsc$ are kernels. 
-		\begin{enumerate}[(i)]
-			\item If $\alpha_1, \alpha_2 \geq 0$ then $\alpha_1 k_1 + \alpha_2 k_2$ is a kernel. If $\lim_{k\to\infty} k(x, x') \eqqcolon k(x, x')$ exists for all $x, x' \in \Cal X$, then $k$ is a kernel. 
-			\item The pointwise product $k = k_1k_2$ is a kernel.
-		\end{enumerate}
-	\end{mdframed}
+	We have \[
+	\PP(M = M') = \PP\qty(\argmin_{j \in z \cup z'} \pi(j) \in z \cap z') = \frac{\abs{z \cap z'}}{\abs{z \cup z'}} = k(z, z') \quad\text{since $\pi$ is random.}
+	\]
 
-Let $x_1, \dotsc, x_n \in \Cal X$ and define the positive definite matrix $K^{(m)}$ by $K^{(m)}_{i, j} = k_m(x_i, x_j)$. 
-\begin{enumerate}[(i)]
-	\item Since $(\alpha_1 k_1 + \alpha_2 k_2)(x_i, x_j) = \alpha_1 k_1(x_i, x_j) + \alpha_2 k_2(x_i, x_j)$, the kernel matrix $K_{ij} \ceq (\alpha_1 k_1 + \alpha_2 k_2)(x_i, x_j)$ equals $\alpha_1 K_1 + \alpha_2 K_2$, and is therefore clearly symmetric positive semi-definite.
-This shows that $\alpha_1 k_1 + \alpha_2 k_2$ is a kernel.
+	Furthermore, we have
+	\[
+	\EE[\psi_M \psi_{M'}] = \PP(M = M') \EE[\psi_M^2] + \PP(M \neq M') \EE[\psi_M \psi_{M'}] = k(z, z'), 
+	\]
+	since for $M \neq M'$ we have $\EE[\psi_M\psi_M'] = \EE[\psi_M]\EE[\psi_{M'}] = 0$. 
+	Let $z_1, \dotsc, z_n \in \XX$ with corresponding $M_1, \dotsc, M_n$, and write $\hat \psi = (\psi_{M_1}, \dotsc, \psi_{M_n})\T$, then the kernel matrix $K$ is given by $\EE[\hat\psi \hat\psi\T]$ which is positive semidefinite.
 
-	For the second part (the pointwise limit of kernels is a kernel), it suffices to show that the entrywise limit of positive semi-definite matrices is again positive semidefinite.
-But this is clear: if $K^{(m)} \to K$ entrywise, then for any $\vec x \in \RR^n$ we have
-	\[
-	\vec x\T K \vec x = \vec x\T (\lim_{m\to\infty} K^{(m)}) \vec x = \lim_{m\to\infty} \vec x\T K^{(m)}\vec x \geq 0,
-	\]
-	so $K$ is positive semi-definite.
-
-	\item Let $K^{(1)}$ and $K^{(2)}$ be positive semi-definite, then we must show that the pointwise product $K^{(1)} \circ K^{(2)}$ is positive semi-definite as well.
-Write $K^{(1)} = \sum_{\ell=1}^n \lambda_\ell P_\ell P_\ell\T$ and $K^{(2)} = \sum_{\ell=1}^n \mu_\ell Q_\ell Q_\ell\T$, where $\lambda_i, \mu_i \geq 0$ and $\qty{P_\ell}$, $\qty{Q_\ell}$ are orthonormal bases of $\RR^n$.
-
-	Then we find
-	\begin{align*}
-	(K^{(1)} \circ K^{(2)})_{ij} &= \qty(\sum_m \lambda_m P_{im} P_{jm}) \qty(\sum_\ell \mu_\ell Q_{i\ell} Q_{j\ell}) = \sum_{m, \ell} \lambda_m \mu_\ell (P_{im} P_{jm} Q_{i\ell} Q_{j\ell}) \\
-	&= \sum_{m, \ell} \lambda_m \mu_\ell \qty[(P_m \circ Q_\ell)(P_m \circ Q_\ell)\T]_{ij},
-	\end{align*}
-
-	and therefore $K^{(1)} \circ K^{(2)} = \sum_{m, \ell} \lambda_m \mu_\ell \qty[(P_m \circ Q_\ell)(P_m \circ Q_\ell)\T]$, which is a nonnegative combination of positive semi-definite matrices and therefore positive semi-definite.
-\end{enumerate}
+	Using the random feature map $\hat\phi(z) = \psi_{M_z}$ we can approximate kernel ridge regression using the random feature map method. 
 \end{proof}
-
-\begin{question}
-	Let $\XX \ceq \qty{x \in \RR^d : \norm{x} < 1}$. Show that $k(x, x') \ceq (1 - x\T x')^{-\alpha}$ defined on $\XX \times \XX$, where $\alpha > 0$, is a kernel.
-\end{question}
-
-\begin{proof}
-	Note by Cauchy-Schwarz we have for $x, x' \in \XX$ that $\abs{x\T x'} \leq \norm{x} \norm{x'} < 1$, so $k$ is well-defined.
-
-	Symmetry of $k$ is clear. Define the function 
-	\[
-	\phi \colon (-\infty, 1) \to \RR \colon x \mapsto (1-x)^{-\alpha}. 
-	\]
-	Then $\phi$ is continuous, and we can compute its Taylor series around 0. It is computed that 
-	\[
-	\phi^{(k)}(x) = (1-x)^{-\alpha - k} \prod_{j=0}^{k-1}(\alpha + j) \implies \phi^{(k)}(0) = \prod_{j=0}^{k-1} (\alpha + j). 	\]
-	Therefore, the Taylor series of $\phi$ is given by
-	\[
-	\sum_{k=0}^\infty \qty(\prod_{j=0}^{k-1} (\alpha + j)) \frac{x^k}{k!}, 
-	\]
-	and its radius of convergence is given by
-	\[
-	\lim_{k\to\infty} \frac{\frac{1}{k!}\prod_{j=0}^{k-1} (\alpha + j)}{\frac{1}{(k+1)!} \prod_{j=0}^k (\alpha + j)} = \lim_{k\to\infty} \frac{k}{\alpha + k} = 1,
-	\]
-	so $\phi$ agrees with its Taylor series on $(-1, 1)$. By Cauchy-Schwarz, for $x, x' \in \XX$ we have $x_i\T x_j \leq \norm{x_i} \norm{x_j} < 1$, so we can write for $x, x' \in \XX$
-	\[
-	k(x, x') = \phi(x\T x') = \sum_{k=0}^\infty \qty(\prod_{j=0}^{k-1} (\alpha + j)) \frac{(x\T x')^k}{k!}.
-	\]
-	By proposition 4, every term in the sum is a kernel, and therefore the limit is a kernel as well. 
-\end{proof}
-
-\begin{question}
-	Suppose we have a matrix of predictors $X \in \RR^{n \times p}$ where $p \gg n$. Explain how to obtain the fitted values of the following ridge regression using the kernel trick:
-	\begin{gather*}
-		\text{Minimise over $\beta \in \RR^p$, $\theta \in \RR^{p(p-1)/2}$, $\gamma \in \RR^p$,} \\
-		\sum_{i=1}^n \qty(Y_i - \sum_{k=1}^p X_{ik}\beta_k - \sum_{k=1}^p \sum_{j=1}^{k-1} X_{ik} X_{ij} \theta_{jk} - \sum_{k=1}^p X_{ik}^2\gamma_k)^2 + \lambda_1 \norm{\beta}^2 + \lambda_2 \norm{\theta}^2 + \lambda_3\norm{\gamma}^2.
-	\end{gather*}
-Note that we have indexed $\theta$ with two numbers for convenience. 
-\end{question}
-
-\begin{proof}
-	Our linear model has predictors
-	\[
-	\qty{X_{ik} \mid 1 \leq k \leq p} \cup \qty{X_{ij} X_{ik} \mid 1 \leq j < k \leq p} \cup \qty{X_{ik}^2 \mid 1 \leq k \leq p}
-	\]
-	for $i = 1, \dotsc, k$.
-We want to assume that $\lambda_1 = \lambda_2 = \lambda_3$: to this end, define $\xi = \frac{\lambda_1}{\lambda_2}$ and $\eta = \frac{\lambda_3}{\lambda_1}$, and replace $\theta$ by $\xi\theta$ and $\gamma$ by $\eta\gamma$.
-This means we will have to replace our predictors by
-	\[
-	Z_i \ceq 
-	(X_{ik} \mid 1 \leq k \leq p)\T \cup (\xi X_{ij} X_{ik} \mid 1 \leq j < k \leq p)\T \cup (\gamma X_{ik}^2 \mid 1 \leq k \leq p)\T
-	\]
-	We use the kernel trick: note that
-	\begin{align*}
-	\ang{Z_i, Z_j} &= \sum_k X_{ik}X_{jk} + \xi^2 \sum_{k < \ell} X_{ik} X_{i\ell} X_{jk} X_{j\ell} + \gamma^2 \sum_k X_{ik}^2 X_{jk}^2 \\
-	&= \qty(\frac{1}{\sqrt 2\xi} + \frac{\xi}{\sqrt 2}\sum_k X_{ik}X_{jk})^2 + (\gamma^2 - \frac{\xi^2}{2})\sum_k X_{ik}^2 X_{jk}^2 - \frac{1}{2\xi^2}.
-	\end{align*}
-	We have therefore rewritten $\ang{Z_i, Z_j}$ into something that can be computed in $O(p)$, which is exactly the aim of the kernel trick.
-\end{proof}
-
-\begin{question}
-	Let $\hat\alpha$ be a minimiser of $\norm{Y - K\alpha}^2 + \lambda \alpha\T K\alpha$, with $K$ being a kernel matrix as usual. Show that $K\hat\alpha = K(K + \lambda I)^{-1} Y$. 
-\end{question}
-
-\begin{solution}
-	Suppose $\hat\alpha$ minimises $Q(\alpha) \ceq \norm{Y - K\alpha}^2 + \lambda \alpha\T K\alpha$. Then we have
-	\begin{align*}
-		0 = \nabla_\alpha Q(\hat\alpha) &= -2 K (Y - K\hat\alpha) + 2\lambda K\hat\alpha \\
-		\lambda K\hat\alpha &= KY - K^2\hat\alpha \\
-		(K + \lambda) K\hat\alpha &= KY \\
-		K\hat\alpha &= (K + \lambda)^{-1} KY.
-	\end{align*}
-Since $K$ and $(K + \lambda)^{-1}$ are simultaneously diagonalisable, they commute, and therefore $K\hat\alpha = K(K + \lambda)^{-1} Y$. 
-%\begin{align*}
-%(K + \lambda)^{-1} KY &= (U(D + \lambda) U\T)^{-1} UDU\T Y \\
-%&= U(D + \lambda)^{-1} D U\T Y \\
-%&= UD (D+ \lambda)^{-1} U\T Y\\
-%&= UDU\T (U (D+ \lambda) U\T) Y\\
-%&= K(K + \lambda)^{-1} Y,
-%\end{align*}
-%which proves the claim.
-\end{solution}
-
-\begin{question}
-	...
-\end{question}
-
-\begin{proof}
-	content...
-\end{proof}
-
-\begin{question}
-	...
-\end{question}
-
-\begin{proof}
-	...
-\end{proof}
-
-\begin{question}
-	Show from first principles that the Sobolev kernel is indeed a (positive definite) kernel.
-\end{question}
-
-\begin{solution}
-	Let $\XX = [0, 1]$ and $k(x, x') = \min(x, x')$ the Sobolev kernel.
-We must then show that for any $x_1, \dotsc, x_n \in \XX$, the matrix $K_{ij} = \min(x_i, x_j)$ is positive semi-definite (symmetry of $k$ is clear).
-
-	To this end, assume $x_1 \leq \dotsb \leq x_n$ (this may require relabeling of the $x_i$ which does not affect positive semi-definiteness of $K$).
-Now let $\vec f_i$ denote the vector $f_{i1}, \dotsc, f_{i(i-1)} = 0$, $f_{ii}, \dotsc, f_{in} = 1$.
-	\[
-	K = \mqty[ x_1 & x_1 & x_1 & \cdots & x_1 \\
-	x_1 & x_2 & x_2 & \dotsb & x_2 \\
-	x_1 & x_2 & x_3 & \dotsb & x_3 \\
-	\vdots & \vdots & \vdots & \ddots & \vdots \\
-	x_1 & x_2 & x_3 & \cdots & x_n] = x_1 \vec f_1 \vec f_1\T + (x_2 - x_1) \vec f_2 \vec f_2\T + \dotsb + (x_n - x_{n-1}) \vec f_n \vec f_n\T. 
-	\]
-	Since this is a nonnegative combination of positive semidefinite matrices, we conclude that $K$ is positive semidefinite. 
-\end{solution}
 \end{document}

### 2020-10-31 modern_statistical_methods/msm_ES2.tex
progress?

@@
 \end{proof}
+
+\begin{question}
+	Consider the logistic regression model where we assume $Y_1, \dotsc, Y_n \in \qty{-1, 1}$ are independent and
+	\[
+	\log(\frac{\PP(Y_i = 1)}{\PP(Y_i = -1)}) = x_i\T\beta^0. 
+	\]
+	Show that the maximum likelihood estimate $\beta$ minimises
+	\[
+	\sum_{i=1}^n \log(1 + \exp(-Y_i x_i\T \beta))
+	\]
+	over $\beta \in \RR^p$. 
+\end{question}
+
+\begin{proof}
+	Let $(y_1, \dotsc, y_n)$ be the responses, then the likelihood function becomes
+	\[
+	L(\beta) = \PP(Y_1, \dotsc, Y_n \mid \beta)
+	\]
+
+	Note that
+	\begin{align*}
+		\frac{\PP(Y_i = 1)}{1 - \PP(Y_i = 1)} = \exp(x_i\T\beta) \implies \PP(Y_i = 1) = \frac{1}{1 + \exp(-x_i\T \beta)},
+	\end{align*}
+and analogously
+\begin{align*}
+	\PP(Y_i = -1) =  \frac{1}{1 + \exp(x_i\T\beta)}. 
+\end{align*}
+We can combine the above formulas as
+\[
+\PP(Y_i = y_i) = \frac{1}{1 + \exp(-y_i x_i\T\beta)}. 
+\]
+
+Therefore our the MLE $\hat\beta$ maximises
+\[
+L(\beta) = \prod_{i=1}^n \frac{1}{1 + \exp(-Y_i x_i\T\beta)}, 
+\]
+and it also maximises the log-likelihood function
+\[
+\log(L(\beta)) = - \sum_{i=1}^n \log(1 + \exp(-Y_i x_i\T\beta))
+\]
+which is of course equivalent to minimising
+\[
+-\log(L(\beta)) = \sum_{i=1}^n \log(1 + \exp(-Y_i x_i\T\beta)). 
+\]
+\end{proof}
+
+\begin{question}
+	Consider the following algorithm for model selection when we have a response $Y \in \RR^n$ and a matrix of predictors $X \in \RR^{n \times p}$.
+	\begin{enumerate}[(a)]
+		\item First centre $Y$ and all the columns of $X$. Initiale the current model $M \subseteq \qty{1, \dotsc, p}$ to be $\emptyset$ and set the current residual $R$ to be $Y$. 
+		\item Find the variable $k^*$ in $M^c$ most correlated with the current residual $R$.
+Set $M$ to be $M \cup \qty{k^*}$.
+Replace $R$ with the residual from regressing $R$ onto $X_{k^*}$.
+Further replac eeach variable in $M^c$ with the residual from regressing itself onto $X_{k^*}$.
+		\item Continue the previous step unto $R = 0$. 
+	\end{enumerate}
+Show that this algorithm is equivalent to forward selection.
+
+\emph{Hint:} Use induction on the iteration $m$ of the algorithm.
+Consider strengthening the natural induction hypothesis that the model at iteration $m$ is the same as that selected after $m$ steps of forward selection.
+\end{question}
+
+\begin{solution}
+	content...
+\end{solution}
+
+\begin{question}
+	Show that if $W$ is mean-zero and sub-Gaussian with parameter $\sigma$, then $\Var(W) \leq \sigma^2$. 
+\end{question}
+
+\begin{proof}
+	If $W$ is mean-zero, then $\Var(W) = \EE(W^2)$. By the proof of lemma 15 we have $\EE(W^2) \leq 2 \sigma^2$, but this bound is too loose.
+
+	Since $W$ is sub-Gaussian, we have for all $\alpha \in \RR$ that
+	\begin{align*}
+		\EE[e{\alpha W}] &\leq e^{\alpha^2\sigma^2/2} \\
+		\sum_{k=0}^\infty \frac{\EE[W^k] }{k!} \alpha^k &\leq \sum_{k=0}^\infty \frac{\sigma^{2k}}{2^k k!} \alpha^{2k} \\
+		\frac{\EE[W^2]}{2} \alpha^2 + \sum_{k=3}^\infty \frac{\EE[W^k]}{k!} \alpha^k &\leq \frac{\sigma^2}{2} \alpha^2 + \sum_{k=2}^\infty \frac{\sigma^{2k}}{2^k k!} \alpha^{2k} \\
+		\frac12\alpha^2 \qty(\sigma^2 - \EE[W^2]) &\geq \alpha^3 P(\alpha),
+		\end{align*}
+	where $P$ is a power series in $\alpha$. Rescaling we find \[
+	\frac12(\sigma^2 - \EE[W^2]) \geq \alpha P(\alpha),
+	\]
+	and letting $\alpha \to 0$ also lets $\alpha P(\alpha) \to 0$, and therefore $\sigma^2 - \EE[W^2] \geq 0$ so $\Var(W) \leq \sigma^2$. 
+\end{proof}
+
+\begin{question}
+	Verify Hoeffding's lemma for the special case where $W$ is a Rademacher random variable, so $W$ takes the values $-1, 1$ each with probability $1/2$. 
+\end{question}
+
+\begin{proof}
+		If $W$ is a Rademacher random variable then we have, using $(2k)! \geq 2^k k!$ for each $k \in \NN$, that
+	\begin{align*}
+		\EE[e^{\alpha W}] = \frac12 e^{-\alpha} + \frac12 e^{\alpha} = \sum_{k=0}^\infty \frac{\alpha^{2k}}{(2k)!} \leq \sum_{k=0}^\infty \frac{\alpha^{2k}}{2^k k!} = e^{\alpha^2/2}
+	\end{align*}
+which shows that $W$ is sub-Gaussian with parameter $1 = (1 -(-1))/2$, so Hoeffding's lemma holds indeed. 
+\end{proof}
+
+\begin{question}
+	\begin{enumerate}[(a)]
+		\item Let $W \sim \chi_d^2$. Show that
+		\[
+		\PP(\abs{W/d - 1} \geq t) \leq 2e^{-dt^2/8}
+		\]
+		for $t \in (0, 1)$.
+You may use the facts that the mgf of a $\chi_1^2$ random variable is $(1 - 2\alpha)^{-1/2}$ for $\alpha < 1/2$, and $e^{-\alpha} (1 - 2\alpha)^{-1/2} \leq e^{2\alpha^2}$ when $\abs{\alpha} < 1/4$.
+
+		\item Let $A \in \RR^{d\times p}$ have i.i.d.\ standard normal entries. Fix $u \in \RR^p$. Use the result above to conclude that
+		\[
+		\PP\qty(\abs{\frac{\norm{Au}_2^2}{d \norm{u}_2^2} - 1} \geq t) \leq 2e^{-dt^2/8}. 
+		\]
+
+		\item Suppose we have data $u_1, \dotsc, u_n \in \RR^p$, with $p$ large and $n \geq 2$.
+Show that for a given $\eps \in (0, 1)$ and $d > 16\log(n/\sqrt\eps)/t^2$, each data point may be compressed down to $u_i \mapsto Au_i/\sqrt d = w_i$, whilst approximately preserving the distance between the points:
+		\[
+		\PP\qty(1 - t \leq \frac{\norm{w_i - w_j}_2^2}{\norm{u_i - u_j}_2^2} \leq 1 + t \text{ for all } i \neq j \in \qty{1, \dotsc, n}) \geq 1 - \eps. 
+		\]
+		This is the famous Johnson-Lindenstrauss Lemma. 
+	\end{enumerate}
+\end{question}
+
+\begin{proof}
+	\begin{enumerate}[(a)]
+		\item 
+	\end{enumerate}
+\end{proof}
 \end{document}

### 2020-10-31 topics_statistical_theory/topics_ES1.tex
progress?

@@
 \DeclareMathOperator\Cov{Cov}
+\DeclareMathOperator\diag{diag}
 
@@
 	\begin{gather*}
-	\PP\qty[ n \qty(\hat F_n(t_1), \dotsc, 1 - \hat F_n(t_k)) (a_1, \dotsc, a_{n+1})] \\ = \PP[\text{exactly $a_i$ of the $X_i$ lie in $(t_{i-1}, t_i]$ for $i = 1, \dotsc, n$}].
+	\PP\qty[ n \qty(\hat F_n(t_1), \dotsc, 1 - \hat F_n(t_k)) = (a_1, \dotsc, a_{k+1})] \\ = \PP[\text{exactly $a_i$ of the $X_i$ lie in $(t_{i-1}, t_i]$ for $i = 1, \dotsc, n$}].
 	\end{gather*}
 
-This probability is 0 unless all $a_i$ are nonnegative integers with $\sum_i a_i = n$.
-In this case, we can first choose $a_1$ of the $X_i$ which must lie in $(\infty, t_1]$, then we choose $a_2$ of the remaining $X_i$ which must lie in $(t_1, t_2]$ and so forth.
-We find that the above probability equals
+In this case, we have a multinomial distribution with $n$ trials and probabilities $F(t_1), F(t_2) - F(t_1), \dotsc, F(t_k) - F(t_{k-1}), 1 - F(t_k)$.
+Therefore, the probability is 0 if $\sum_i a_i \neq n$ and else it is
 \[
-\binom{n}{a_1} (F(t_1))^{a_1} \binom{n - a_1}{a_2} (F(t_2) - F(t_1))^{a_2} \dotsb \binom{a_k + a_{k_1}}{a_k} (F(t_k) - F(t_{k-1}))^{a_k} \cdot (1 - F(t_k))^{a_{k+1}}. 
+\frac{n!}{a_1! \dotsb a_{k+1}!} F(t_1)^{a_1} \dotsb (1 - F(t_k))^{a_{k+1}}. 
 \]
@@
 \end{solution}
+
+\begin{question}
+	We say that a continuous process $(B_t)_{t \in [0, 1]}$ is a \emph{standard Brownian motion} on $[0, 1]$ if $B_0 = 0$ and if, for $0 \leq s_1 \leq t_1 \leq \dotsb \leq s_k \leq t_k \leq 1$, we have $(B_{t_1} - B_{s_1}, \dotsb, B_{t_k} - B_{s_k}) \sim N_k(0, \Sigma)$, where $\Sigma \ceq \diag(t_ 1- s_1, \dotsb, t_k - s_k)$.
+The process $(W_t)_{t \in [0, 1]}$ defined by $W_t \ceq B_t - tB_1$ is called a \emph{Brownian bridge}, or \emph{tied-down Brownian motion}, because $W_0 = W_1 = 0$.
+Compute the distribution of $(W_{t_1}, \dotsc, W_{t_k})$.
+\end{question}
+
+\begin{solution}
+	Note that $W_t = B_t - t B_1 = (1-t) (B_t - B_0) - t (B_1 - B_t)$.
+Now, since $(B_t - B_0)$ and $(B_1 - B_t)$ are independent with distributions $N(0, t)$ and $N(0, 1-t)$ distributions respectively, we find that
+	\[
+	W_t \sim (1-t) N(0, t) + t N(0, 1 - t) = N\qty(0, \frac{t}{\sqrt{1 - t}}) + N\qty(0, \frac{1-t}{\sqrt t}) = N\qty(0, \frac{t}{\sqrt{1 - t}} + \frac{1-t}{\sqrt t}). 
+	\]
+
+	??
+\end{solution}
+
+\begin{question}
+	Let $\phi$ denote the standard normal density function, which is a bounded, second-order kernel.
+For $r \in \NN_0$, define the $r$-th Hermite polynomial $H_r$ by $H_r(x) \ceq (-1)^r \phi^{(r)}(x) / \phi(x)$.
+Prove that $H_r$ is a monic polynomial of degree $r$ that is even if $r$ is even and odd if $r$ is odd.
+Show further that
+	\[
+	\int_{-\infty}^\infty H_r(u) H_s(u) \phi(u) \dd{u} = \begin{cases}
+		(2\pi)^{1/2} r!, & r=s, \\ 0 , & r \neq s.
+	\end{cases}
+	\]
+	Now fix an integer $\ell \geq 2$ and define 
+	\[
+	K_\ell(u) \ceq \sum_{r=0}^{\ell -1 } \frac{H_r(0) H_r(u)}{(2\pi)^{1/2} r!} e^{-u^2/2}. 
+	\]
+	Prove that $K_\ell$ is a bounded kernel of order $\ell$. 
+\end{question}
+
+\begin{proof}
+ 	We prove this by induction on $r$. For $r = 0$, we have $H_0(x) = 1$, which is indeed an even monic polynomial of degree 0. 
+ 	Now, suppose the claim holds for a given $r$, that is, $H_r(x) = (-1)^r \phi^{(r)}(x)/\phi(x) = p(x)$ for some monic polynomial $p$ of degree $r$, which is even if $r$ is even and odd if $r$ is odd.
+Then we have
+ 	\begin{align*}
+ 	\phi^{(r)}(x) &= (-1)^r p(x) \phi(x) = (-1)^r (2\pi)^{-1/2} p(x) \exp(-x^2/2) \\
+ 	\phi^{(r+1)}(x) &= (-1)^r 2\pi^{-1/2} \qty(p'(x) - x p(x)) \exp(-x^2/2) \\
+ 	H_{r+1 }(x) &= (-1)^r (-1)^{r+1} \qty(p'(x) - x p(x)) = xp(x) - p'(x). 
+ 	\end{align*}
+ 	Now it is clear that $H$ is a monic polynomial of degree $r$ since $p$ was assumed monic.
+Furthermore, since derivatives of even functions are odd and vice versa, it is clear that $H$ is odd if $p$ is even and vice versa.
+
+	Now, suppose $r < s$, then 
+	\begin{align*}
+		\int_{-\infty}^\infty H_r(u) H_s(u) \phi(u) \dd{u} &= (-1)^s \int_{-\infty}^\infty H_r(u) \phi^{(s)}(x) \dd{u} \overset{\text{IBP}}= \int_{-\infty}^\infty H_r^{(s)}(u) \phi(u) \dd{u} = 0,
+	\end{align*}
+since $H_r^{(s)} = 0$ if $r < s$.
+
+However, if $r = s$, then following the same line of reasoning as above and using the fact that $H_r^{(r)} = r!$, we find
+\[
+\int_{-\infty}^\infty H_r^2(u) \phi(u) \dd{u} = r! \int_{-\infty}^\infty \phi(u) \dd{u} = r!. 
+\]
+
+Now we consider $K_\ell$: we have
+\[
+\int_{-\infty}^\infty K_\ell(u) \dd{u} = \sum_{r=0}^{\ell - 1} \frac{H_r(0)}{r!} \int_{-\infty}^\infty H_r(u) \phi(u) \dd{u} =  \sum_{r=0}^{\ell - 1} \frac{(-1)^r H_r(0)}{r!} \int_{-\infty}^\infty \phi^{(r)}(u) \dd{u}.
+\]
+Note that every term in the above sum vanishes except for the $r = 0$ term due to the integral, and the $r = 0$ term is 1, so $K_\ell$ is indeed a kernel.
+
+We verify that $K_\ell$ has order $\ell$: let $j \in \qty{1, \dotsc, \ell - 1}$, then we have
+\begin{align*}
+	\int_{-\infty}^\infty u^j K_\ell(u) \dd{u} = \sum_{r=0}^{\ell - 1} \frac{(-1)^r H_r(0)}{r!} \int_{-\infty}^\infty u^j \phi^{(r)}(u) \dd{u}.
+\end{align*}
+Clearly, if $j < r$, integral in the sum vanishes. If $j = r$, the integral yields $r!$. So we have
+\[
+\int_{-\infty}^\infty u^j K_\ell(u) \dd{u} = (-1)^j H_j(0) + \sum_{r=0}^{j-1} \int_{-\infty}^\infty u^j \phi^{(r)}(u) \dd{u}. 
+\]
+\end{proof}
 \end{document}

### 2020-11-01 inverse_problems/inverse_ES1_handin.tex
yolo

--- 2020-10-28 inverse_problems/inverse_ES1.tex
@@
 	\begin{equation} \label{eq:picard}
-	2\sum_{j=1}^\infty  \sigma_j^{-2} \qty(\int_0^1 f(s) \sin(\sigma_j^{-1} s) \dd{s})^2,
+	2\sum_{j=1}^\infty  \sigma_j^{-2} \qty(\int_0^1 f(s) \sin(\sigma_j^{-1} s) \dd{s})^2 < \infty,
 	\end{equation}
@@
 
-\begin{question}
-	\begin{enumerate}[(a)]
-		\item
-
-	Let $m, n \in \NN$ with $m \geq n \geq 2$. Compute the Moore-Penrose inverses of the following matrices:
-	\begin{enumerate}[(i)]
-		\item $A = (1, 1, \dotsc, 1) \in \RR^{1 \times n}$;
-		\item $A = \diag(a_1, \dotsc, a_n) \in \RR^{n \times n}$ with $a_j \in \RR$ for $j = 1, \dotsc, n$;
-		\item $A \in \RR^{m \times n}$ with $A\T A = I_n$. 
-	\end{enumerate}
-
-\item Let $a, b \in \RR$ with $a < b$. Compute the Moore-Penrose inverse of the operator $A \colon L^2([a, b]) \to \RR$ with $Au = \int_a^b u(x) \dd{x}$. 
-	\end{enumerate}
-\end{question}
-
-\begin{solution}
-	\begin{enumerate}[(a)]
-		\item \begin{enumerate}[(i)]
-			\item Clearly $\Ran(A) = \RR$ and $\Nul(A)\P = \Span\qty{\vec e}$ where $\vec e$ is the all-ones vector. So $A\D$ must map $x \in \RR$ to $\vec e/n$, and therefore we have
-			\[
-			A\D = \vec e/n \in \RR^{n \times 1}.
-			\]
-
-			\item Clearly we have $\Ran(A) = \Span\qty{\vec e_j \mid a_j \neq 0} = \Nul(A)\P$ while $\Ran(A)\P = \Span\qty{\vec e_j \mid a_j = 0} = \Nul(A)$.
-
-			It is easily seen that
-			\[
-			A\D = \diag(b_1, \dotsc, b_n) \in \RR^{n \times n}, \quad\text{where } b_i = \begin{cases}
-				a_i^{-1} &\text{if $a_i \neq 0$}; \\
-				0 &\text{if $a_i = 0$}. 
-			\end{cases}
-			\]
-
-			\item 
-%			If $A\T A = I$, then the columns $\vec u_1, \dotsc, \vec u_n$ of $A$ are orthogonal.
-In particular, $A$ is injective (so $\Nul(A) = \qty{0}$) and $\Ran(A) = \Span\qty{\vec u_1, \dotsc, \vec u_n}$.
-Extend $\vec u_1, \dotsc, \vec u_n$ to an orthonormal basis $\vec u_1, \dotsc, \vec u_m$ of $\RR^m$, then we must have
-%			\[
-%			A\D\qty(\sum_{i=1}^m \alpha_i \vec u_i) = \mqty(\alpha_1 \\ \vdots \\ \alpha_n) \in \RR^n.
-%			\]
-%			From this expression it is easily seen that $A\D = A\T$.
-			We claim $A\D = A\T$. For this, we check the first two Moore-Penrose equations: for the first one, we have $A\T A = I_n = P_{\RR^n} = P_{\Nul(A)\P}$ since $\Nul(A) = \qty{0}$.  
-			For the second one, let $\vec u_1, \dotsc, \vec u_n \in \RR^m$ be the columns of $A$ (which are orthonormal), then we have
-			\[
-			AA\T = \sum_{i=1}^n \vec u_i \vec u_i\T = P_{\Span\qty{\vec u_1, \dotsc, \vec u_n}} = P_{\clos{\Ran(A)}}. 
-			\]
-		\end{enumerate}
-
-	\item Clearly we have $\Ran(A) = \RR$ while $\Nul(A) = \qty{u \mid \int_a^b u(x) \dd{x} = 0}$. It is also easily seen that 
-	\[
-	\Nul(A)\P = \qty{v \mid \int_a^b v(x) u(x) \dd{x} = 0 \text{ if $\int_a^b u(x) \dd{x} = 0$}} = \Span\qty{1}. 
-	\]
-
-	Therefore we simply have that $A\D$ maps a constant $c \in \RR$ to the constant function $\frac{c}{b-a}$. 
-	\end{enumerate}
-\end{solution}
-
-\begin{question}
-	Many forward problems are either modelled as convolutions or they are modelled as the composition of several components, one of which is a convolution.
-Therefore convolutions play an important role in inverse problems.
-As in Exercise 1, let $\Omega =[0, 1]^2$ be the unit square and let $\XX = L^2(\Omega)$.
-A convolution is the special case of an integral operator $A \colon \XX \to \XX$ where the kernel has a simple structure:
-	\[
-	(Au)(y) \ceq \int_\Omega k(y - x) u(x) \dd{x},
-	\]
-	for $k \in L^2(\Omega)$. It follows easily from Exercise 1 that $A$ is linear and bounded. 
-	\begin{enumerate}[(a)]
-		\item Although shown in general in Exercise 1, give an explicit form for the adjoint of the convolution. 
-		\item Let $f = Au$. It follows from the convolution theorem that a convolution can be inverted by means of the Fourier transform
-		\begin{equation} \label{eq:fourier}
-		u = (2\pi)^{-\frac n2} \FF^{-1} \qty(\frac{\FF(f)}{\FF(k)}),
-		\end{equation}
-		where $\FF$ is the Fourier transform and $\FF^{-1}$ its inverse.
-Implement this formula in \emph{MATLAB} to deblur the blurry tree image $f$ generated by the script \verb|ex4b_generate_data.m|.
-Note that the script also outputs $\FF(k)$.
-Add some noise to the data and show that the inversion formula is ill-conditioned.
-
-		Hint: Make use of the \emph{MATLAB} commands \verb|fft2| and \verb|ifft2|.
-
-		\item Reformulate \cref{eq:fourier} so that the denominator is non-negative and give a stable approximation of this formula. Implement this formula in MATLAB and empirically show that it is stable.
-
-		Hint: Make use of the \emph{MATLAB} command \verb|conj|. 
-	\end{enumerate}
-\end{question}
-
-\begin{proof}
-	\begin{enumerate}[(a)]
-		\item We have by question 1 that
-		\[
-		(A^* v)(y) = \int_\Omega k(x - y)v(x) \dd{x}.
-		\]
-		\item 
-	\end{enumerate}
-\end{proof}
 \end{document}

### 2020-11-01 modern_statistical_methods/C2.tex
updated msm notes

@@
 \end{proof}
+
+We now move to finding tail bounds for products of sub-Gaussian random variables. To this end, we consider Bernstein's in equality. 
+\begin{definition}
+	A random variable $W$ with mean $\mu$ is said to satisfy \emph{Bernstein's condition} with parameters $\sigma > 0, b > 0$ if 
+	\[
+	\EE\qty(\abs{W - \mu}^k) \leq \frac12 k! \sigma^2 b^{k-2} \qquad k = 2, 3, \dotsc
+	\]
+\end{definition}
+
+\begin{proposition}[Bernstein's inequality]
+	Let $W_1, W_2, \dotsc$ be independent random variables with mean $\mu$, where each $W_i$ satisfies Bernstein's condition with parameter $(\sigma, b)$. Then 
+	\begin{enumerate}
+		\item $\EE[\exp(\alpha(W_i - \mu))] \leq \exp\qty(\frac{\alpha^2\sigma^2}{2(1 - b \abs{\alpha})})$ for all $\abs{\alpha} < b^{-1}$;
+		\item $\PP\qty(\overline{W} - \mu \geq t) \leq \exp\qty(\frac{-nt^2}{2(\sigma^2 + bt)})$ for all $t \geq 0$.  
+	\end{enumerate}
+\end{proposition}
+
+\begin{proof}
+	See notes from topics in statistical theory.
+\end{proof}
+
+\begin{lemma}
+	Let $W, Z$ be mean-zero, sub-Gaussian random variables with parameters $\sigma_W, \sigma_Z$. Then $WZ$ satisfies the Bernstein condition with parameters $(8\sigma_W\sigma_Z, 4\sigma_W\sigma_Z)$. 
+\end{lemma}
+
+\begin{proof}
+	Using the tail probability formula for expectations one can prove that any mean-zero sub-Gaussian random variable $X$ satisfies $\EE[X^{2k}] \leq 2^{k+1} \sigma_X^{2k} k!$.
+
+	Furthermore for any random variable $Y$ we have
+	\begin{align*}
+		\EE\abs{Y - \EE Y}^k &\leq \EE\qty(\abs{Y} + \abs{\EE Y})^k \\
+		&= \EE \sum_{t=0}^k \binom kt \abs{Y}^t \abs{\EE Y}^{k-t} \\
+		&\overset*\leq \sum_{t=0}^k \binom kt \EE(\abs{Y}^t) \EE(\abs{Y}^{k-t}) \\
+		&\overset*\leq \sum_{t=0}^k \binom kt \EE(\abs{Y}^k)^{t/k} \EE(\abs{Y}^k)^{k - t/k} \\
+		&= \EE(\abs{Y}^k) \sum_{t=0}^k \binom kt = 2^k \EE\abs{Y}^k,
+	\end{align*}
+where both equations $*$ are applications of Jensen's inequality.
+
+From the above bound, we obtian 
+\begin{align*}
+\EE\abs{WZ - \EE WZ}^k &\leq 2^k \EE(\abs{W}^k \abs{Z}^k) \overset{\Rm{CS}}\leq 2^k (\EE W^{2k})^{1/2} (\EE Z^{2k})^{1/2} \leq 2^{2k+1} \sigma_W^k \sigma_Z^k k! \\
+&= \frac{k!}{2} (8\sigma_W\sigma_Z)^2 (4\sigma_W\sigma_Z)^{k-2},
+\end{align*}
+which proves the claim. 
+\end{proof}
+
+\subsubsection{Optimisation theory and convex analysis}
+\begin{proposition}
+	Let $f \colon \RR^d \to \RR$ be twice continuously differentiable.
+Then $f$ is convex (resp.\ strictly convex) if and only if its Hessian is positive semi-definite (resp.\ positive definite) for all $x \in \RR^d$.
+\end{proposition}

### 2020-11-02 distribution_theory/distribution_ES1.tex
distribution theory is still difficult :((

@@
 \usepackage{hyperref}
-\usepackage{mathtools, amsmath, amssymb, amsthm, mathabx, enumerate, mdframed, bbm, graphicx, float, physics, xcolor, cleveref}
+\usepackage[shortlabels]{enumitem}
+\usepackage{mathtools, amsmath, amssymb, amsthm, mathabx, mdframed, bbm, graphicx, float, physics, xcolor, cleveref}
 
@@
 \author{Lucas Riedstra}
-\date{...} % date
+\date{4 November 2020} % date
 
@@
 
+\begin{mdframed}
+For both questions 2 and 4, we need the following lemma:
+\begin{lemma} \label{lem:distance_compact_closed}
+	Let $K, V \subseteq \RR^n$ where $K$ is compact, $V$ is closed, and $K \cap V = \emptyset$. Then there is a nonzero distance between $K$ and $V$, i.e., 
+	\[
+	\inf_{x \in K, v \in V} \norm{x - v} > 0. 
+	\]
+\end{lemma}
+
+\begin{proof}
+	We know that $K \subseteq V\C$ and that $V\C$ is open, so for every $x \in K$ there exists an open ball $B(x, \eps_x)$ around $x$ such that $B(x, 2\eps_x) \subseteq V\C$.
+Since $\qty{B(x, \eps_x)}$ is an open covering of $K$, there exist finitely many balls $B(x_1, \eps_1), \dotsc, B(x_n, \eps_n)$ that cover $K$.
+Let $\eps \ceq \min\qty{\eps_1, \dotsc, \eps_n}$ and $x \in K$, then there is an $x_i$ such that $\norm{x - x_i} < \eps$, and since $B(x_i, 2\eps) \subseteq B(x_i, 2\eps_i) \subseteq V\C$ it is clear that $B(x, \eps) \subseteq V\C$ as well.
+
+	We conclude that $B(x, \eps) \subseteq V\C$ for any $x \in K$, and therefore that $\inf\limits_{x \in K, v \in V} \norm{x - v} \geq \eps > 0$. 
+\end{proof}
+\end{mdframed}
+
 \begin{question}
@@
 \begin{proof}
-	Since $\phi \in \DD(X)$, we know that $\supp\phi \subseteq \clos{B_N}$ for some $N \in \NN$. Now, suppose $\norm{h} < 1$, then
+	Let $\phi \in \DD(X)$ with $K = \supp\phi$, then by \cref{lem:distance_compact_closed} we know there exists a nonzero distance $d > 0$ between $K$ and $\RR^n \setminus X$. 
+	We claim that if $\norm{h} \leq \frac d2$, then 
 	\[
-	\phi(x+h) \neq 0 \implies \norm{x + h} \leq N \implies \norm{x} \leq \norm{x + h} + \norm{h} \leq N + 1, 
+	\supp(R_N) \subseteq \qty{x \in X \mid d(x, K) \leq \frac d2} \eqqcolon \hat K,
 	\]
-	so if we define $\psi_h(x) = \phi(x+h)$ then we know that $\supp\phi_h \subseteq \clos{B_{N+1}}$.
+	which is clearly a compact set contained in $X$. Indeed, if $\norm{h} \leq \frac d2$ we have
+	\[
+	\phi(x+h) \neq 0 \implies x + h \in K \implies d(x, K) \leq \norm{h} \leq \frac d2. 
+	\]
 
+%	Since $\phi \in \DD(X)$, we know that $\supp\phi \subseteq \clos{B_N}$ for some $N \in \NN$. Now, suppose $\norm{h} < 1$, then
+%	\[
+%	\phi(x+h) \neq 0 \implies \norm{x + h} \leq N \implies \norm{x} \leq \norm{x + h} + \norm{h} \leq N + 1, 
+%	\]
+%	so if we define $\psi_h(x) = \phi(x+h)$ then we know that $\supp\phi_h \subseteq \clos{B_{N+1}}$. 
+%	
 	By Taylor's theorem, we have
@@
 	\]
-	and since $\sum_{\abs{\alpha} \leq N} \frac{h^\alpha}{\alpha!} \pt^\alpha \phi(x)$ vanishes for $x \notin \clos{B_N}$, it is clear that $\supp(R_N(\cdot, h))$ must also be contained in $\clos{B_{N+1}}$ (again, for $\norm{h} \leq 1$).
-This shows that $\supp(R_N)$ is contained in $\clos{B_{N+1}}$ for $\abs{h}$ sufficiently small.
+	and since $\sum_{\abs{\alpha} \leq N} \frac{h^\alpha}{\alpha!} \pt^\alpha \phi(x)$ vanishes for $x \notin K$, it is clear that $\supp(R_N(\cdot, h))$ must be contained in $\hat K$ (for $\norm{h} \leq \frac d2$).
 
@@
 	&\leq \qty[\max_{\abs{\alpha} = N+1, x \in \RR^n} \abs{(\pt^{\alpha + \beta})\phi(x)}] \sum_{\abs{\alpha} = N+1} \frac{\abs{h^\alpha}}{\alpha!} (N+1) \\
-	&\leq C (N+1)! \sum_{\abs{\alpha}= N+1}\frac{\abs{h^\alpha}}{\alpha!} = C(\abs{h_1} + \dotsb + \abs{h_n})^{N+1}.
+	&= C (N+1)! \sum_{\abs{\alpha}= N+1}\frac{\abs{h^\alpha}}{\alpha!} = C(\abs{h_1} + \dotsb + \abs{h_n})^{N+1}.
 \end{align*}
+Here, $\star$ follows from differentiation under the integral sign since the integrand is bounded.
+
 Since this upper bound does not depend on $x$, we also have
@@
 \begin{solution}
-	Let $K \subseteq U_1$.
-Define $U_2 \ceq U_1 + B(0, 1)$ and let $\chi = \ind_{U_2}$.
-Now let $\psi \in \DD(\RR^n)$ satisfy $\int_{\RR^n} \psi \dd{x} = 1$ and $\supp\psi \subseteq B(0, 1)$.
-The we compute
+	Let $K \subseteq X$, then by \cref{lem:distance_compact_closed}, there is a nonzero distance $d$ from $K$ to $\RR^n \setminus X$, so we have neighbourhoods
+	\[
+	U_1 \ceq \qty{u \in X \mid d(u, K) < \frac d4}, \quad U_2 \ceq \qty{u \in X \mid d(u, K) < \frac{3d}{4}}
+	\]
+	which is an open neighbourhood of $K$ with $\clos U \subseteq X$.
+
+	 Let $\chi = \ind_{U_2}$. Now let $\psi \in \DD(\RR^n)$ satisfy $\int_{\RR^n} \psi \dd{x} = 1$ and $\supp\psi \subseteq B(0, \frac d4)$. The we compute
 	\begin{align*}
@@
 	\]
-	since $B(0, 1) \subseteq U_2 - x$. This proves the claim.
+	since $\supp\psi \subseteq B(0, \frac d4) \subseteq U_2 - x$. This proves the claim.
 \end{solution}
@@
 
-		\item Since $e^{2\pi i x} = 1 \iff x \in \ZZ$, intuitively it must be the case that $u'$ is 0, except ``on $\ZZ$'', whatever that may mean.
-Therefore, we guess that, for any sequence $(\alpha_n)_{n \in \ZZ} \subseteq \CC$ and constant $c \in \CC$, the map \[u = c + \sum_{n \in \ZZ} \alpha_n \ind_{x \geq n}.\] We compute the derivative of $u$.
+		\item Since $e^{2\pi i x} = 1 \iff x \in \ZZ$, it is clear that $\supp(u') \subseteq \ZZ$.
+We will show that this is also sufficient, i.e., that any distribution $u$ with $\supp(u') \subseteq \ZZ$ yields a solution.
+
+		It is easily seen that 
+		\begin{align*}
+		\supp(u') \subseteq  \ZZ &\iff u' = \sum_{n \in \ZZ} \alpha_n \delta_n \quad\text{for some $(\alpha_n)_{n \in \ZZ} \subseteq \CC$} \\
+		&\iff u = c + \sum_{n\in \ZZ} \alpha_n \ind_{x \geq n} \quad\text{for some $c \in \CC$, $(\alpha_n)_{n \in \ZZ} \subseteq \CC$}. 
+		\end{align*}
+
+		Indeed, if $u = c + \sum_{n \in \ZZ} \alpha_n \ind_{x \geq n}$ then 
+		\[
+		\ang{u', \phi} = -\ang{u, \phi'} \overset\star= - \alpha_n \sum_{n \in \ZZ \cap \supp\phi} \int_n^\infty \phi'(x) \dd{x} = \sum_{n \in \ZZ \cap \supp\phi} \alpha_n \phi(n) = \ang{\sum_{n \in \ZZ} \alpha_n\delta_n, \phi},
+		\]
+		where $\star$ follows from the fact that there are only finitely many $n$ in $\ZZ \cap \supp\phi$ (since $\phi$ has compact support).
+
+%		Therefore, we guess that, for any sequence $(\alpha_n)_{n \in \ZZ} \subseteq \CC$ and constant $c \in \CC$, the map \[u = c + \sum_{n \in \ZZ} \alpha_n \ind_{x \geq n}.\] We compute the derivative of $u$.
 It is easily seen that $u' = \sum_{n \in \ZZ} \alpha_n \delta_n$ (the infiniteness of the sum does not pose a problem since the test functions are compactly supported, so $\ang{u, \phi}$ will always be a finite sum).
@@
 %	Therefore, we can conclude $u' = \sum_{n \in \ZZ} \alpha_n \delta_n$, and indeed we find
-From this, we see that
+Finally, we compute that 
 	\[
@@
 	\]
-	so $u$ satisfies the equation. \TODO Why is this the most general solution? Intuitively clear, but how to make this rigorous?
+	so $u$ satisfies the equation. 
 	\end{enumerate}

### 2020-11-02 inverse_problems/C3.tex
distribution theory is still difficult :((

@@
 \end{definition}
+
+\begin{lemma} \label{lem:lsc_closed_sets}
+	Let $E \colon \XX \to \rext$ be $\tau_\XX$-LSC and $a \in \RR$. Then $\qty{u \in \XX \mid E(u) \leq a}$ is $\tau_\XX$-closed. 
+\end{lemma}
+
+\begin{proof}
+	Let $(u_n) \overset\tau\to u$ and suppose $E(u_n) \leq a$ for all $n$. Then clearly $E(u) \leq \liminf E(u_n) \leq a$. 
+\end{proof}
 
@@
 \begin{definition}
-	A functional $E$ is called \emph{absolutely one-homogeneous} if $E(\lambda u) = \abs{\lambda} E(u)$ for all $\lambda \in \RR, u \in \XX$. 
+	For $p > 0$, a functional $E$ is called \emph{absolutely $p$-homogeneous} if $E(\lambda u) = \abs{\lambda}^p E(u)$ for all $\lambda \in \RR, u \in \XX$. 
 \end{definition}
@@
 
-\begin{theorem}[Direct method]
+\begin{theorem}[Direct method] \label{thm:direct_method}
 	Let $\XX$ be Banach and $\tau_\XX$ a topology on $\XX$ such that any bounded sequence in $\XX$ has a $\tau_\XX$ convergent subsequence.
@@
 \end{theorem}
+
+\subsection{Regularisation properties}
+We consider the problem $Au = f$ where $A \in \BB(\XX, \YY)$ and where $\YY$ is a Banach space and $\XX$ is the dual of a separable Banach space.  Recall that in Tychonoff regularisation, we have
+\[
+R_\alpha(f_\delta) = \argmin_{u \in \XX} \norm{Au - f_\delta}^2 + \alpha \norm{u}^2. 
+\]
+Here, the term $\norm{u}^2$ is known as the \emph{regularisation term}, and penalising large values of $\norm{u}$ ensures regularity of the solution.
+Inspired by this, suppose we have any function $\JJ(u)$ as regulariser, then the \emph{variational regularisation problem} is given by
+\[
+R_\alpha f_\delta \in \argmin_{u \in \XX} \frac12 \norm{Au - f_\delta}^2 + \alpha \JJ(u), 
+\]
+\begin{definition}
+	Let $u_\JJ\D$ be a least-squares solution that minimises $\JJ$ over all least-squares solutions. Then we call $u_\JJ\D$ a \emph{$\JJ$-minimising solution} of the problem $Au = f$. 
+\end{definition}
+
+\begin{convention}
+	We will assume there exists t least one least-squares solution with a finite value of $\JJ$. 
+\end{convention}
+
+\begin{lemma} \label{lem:null_space_subspace}
+	Let $\JJ(u) \ceq \sum_{i=1}^n \JJ_i(u)$, where each $\JJ_i(u)$ is convex and absolutely $p_i$-homogeneous ($p_i > 0$). Then $\Nul(J)$ is a linear subspace of $\XX$. 
+\end{lemma}
+
+\begin{proof}
+	We note that $\JJ_i$ is nonnegative, since
+	\[
+	0 = \JJ_i(0) = \JJ_i\qty(\frac12 u - \frac12 u) \leq \frac12\JJ_i(u) + \frac12\JJ_i(-u) = \JJ_i(u).
+	\]
+	Therefore we have $\Nul(\JJ) = \cap_i \Nul(\JJ_i)$, and we will show that every $\Nul(\JJ_i)$ is a subspace. Let $\lambda \in \RR, u, v \in \Nul(\JJ_i)$, then 
+	\[
+	\JJ_i(\lambda u + v) = 2^{p_i} \JJ_i\qty(\frac{\lambda u}{2} + \frac v2) \leq 2^{p_i - 1} \qty(\abs{\lambda}^{p_i} \JJ_i(u) + \JJ_i(v)) = 0. 
+	\]
+	so $\lambda u + v \in \Nul(\JJ_i)$. This completes the proof. 
+\end{proof}
+
+\begin{lemma}
+	Let the assumptions of \cref{lem:null_space_subspace} be satisfied, and suppose $u \in \XX, v \in \Nul(\JJ)$. Then $\JJ(u + v) = \JJ(u)$. 
+\end{lemma}
+
+\begin{proof}
+	Clearly it suffices to prove this lemma for each $\JJ_i$. It is easily seen that for any $t \in (0, 1)$ we have
+	\[
+	\JJ_i(u + v) = \JJ_i \qty(t \frac ut + (1-t) \frac{v}{1-t}) \leq t \JJ_i \qty(\frac ut) = t^{1 - p_i} \JJ(u). 
+	\]
+	Letting $t \to 1$, we find $\JJ_i(u + v) \leq \JJ_i(u)$.
+
+	Similarly, we have
+	\[
+	\JJ_i(u) = \JJ_i(u + v - v) = \JJ_i \qty(t \frac{ u+ v}{t} + (1-t) \frac{-v}{1-t}) \leq t \JJ_i\qty(\frac{u+ v}{t}) = t^{1 - p_i} \JJ_i(u + v), 
+	\]
+	and letting $t \to 1$ we obtain $\JJ_i(u) \leq \JJ_i(u+v)$, so $\JJ_i(u) = \JJ_i(u+v)$, so $\JJ(u) = \JJ(u + v)$. 
+\end{proof}
+
+\begin{recap}
+	Let $\XX$ be a Banach space and let $U \subseteq \XX$ be a closed subspace of $\XX$.
+Then $U$ is called \emph{complemented} in $\XX$ if there exists a closed subspace $V \subseteq \XX$ such that $\XX = U \oplus V$.
+
+	In general, it is difficult to determine which closed subspaces of $\XX$ are complemented. However, it is known that all \textbf{finite-dimensional} subspaces are complemented. 
+\end{recap}
+\begin{lemma} \label{lem:regulariser_coercive}
+	Suppose $\JJ$ is proper, convex, and satisfies the conditions of \cref{lem:null_space_subspace}. Suppose also that:
+	\begin{enumerate}[(i)]
+		\item $\dim \Nul(\JJ) < \infty$ (so that $\XX = \Nul(\JJ) + \XX_0$ for some closed subspace $\XX_0$) and $\JJ$ is coercive on $\XX_0$;
+		\item $\Nul(A) \cap \Nul(\JJ) = \qty{0}$. 
+	\end{enumerate}
+
+Then the function $\Phi_\alpha(u) \ceq \frac12\norm{Au - f}^2 + \alpha \JJ(u)$ is coercive on $\XX$ for any $\alpha > 0$. 
+\end{lemma}
+
+\begin{proof}
+	Let $(u_j) \subseteq \XX$ be a sequence such that $(\Phi_\alpha(u_j))$ is bounded. We will prove that $(u_j)$ is bounded.
+
+	Decompose every $u_j$ as $u_j = u_j^0 + u_j^{\Cal N}$ with $u_j^0 \in \XX_0$ and $u_j^{\Cal N} \in \Nul(\JJ)$.
+Now, since $(\Phi_\alpha(u_j))$ is bounded, the sequence $(\JJ(u_j)) = (\JJ(u_j^0))$ is bounded as well, and by coercivity of $\JJ$ on $\XX^0$ we conclude that $(u_j^0)$ is bounded.
+
+	Now define $\tilde A \ceq A\restr_{\Nul(\JJ)}$.
+Since $\Nul(A) \cap \Nul(\JJ) = \qty{0}$, $\tilde A$ has an inverse on $A \Nul(\JJ)$, and since $\dim A\Nul(\JJ) < \infty$, that inverse is bounded.
+Therefore we find
+	\begin{align*}
+		\norm{u_j^{\Cal N}} &= \norm{\tilde A^{-1} ( A u_j^{\Cal N})} \leq C \norm{\tilde A u_j^{\Cal N}} = C \norm{\tilde A u_j - f - (A u_j^0 - f)} \leq C \qty(\norm{Au_j - f} + \norm{A} \norm{u_j^0} + \norm{f}).
+	\end{align*}
+	Since $\Phi_\alpha(u_j)$ is bounded, $\norm{Au_j - f}$ is also bounded, and the other terms are also all bounded, so we conclude that $(u_j^{\Cal N})$ is bounded, and therefore that $(u_j)$ is bounded.
+\end{proof}
+
+\begin{theorem}
+	Let $\XX, \YY$ be Banach spaces with topologies $\tau_\XX, \tau_\YY$. Assume that:
+	\begin{enumerate}[(a)]
+		\item bounded sequences in $\XX$ have $\tau_\XX$-convergent subsequences;
+		\item $\JJ\colon \XX \to [0, \infty]$ is proper, convex, $\tau_\XX$-LSC and satisfies the assumptions of \cref{lem:regulariser_coercive};
+		\item $A$ is continuous w.r.t.\ $\tau_\XX$ and $\tau_\YY$;
+		\item $\norm{\cdot}_\YY$ is $\tau_\YY$-LSC. 
+	\end{enumerate}
+
+Then:
+\begin{enumerate}
+	\item There exists a $\JJ$-minimising solution $u_\JJ\D$ of the equation $Au = f$;
+	\item For any $\alpha > 0$, $f \in \YY$ there exists a minimiser
+	\[
+ 	u_\alpha = R_\alpha f \in \argmin \frac12 \norm{Au - f}^2 + \alpha \JJ(u). 
+	\]
+\end{enumerate}
+\end{theorem}
+
+\begin{proof}
+	\begin{enumerate}
+		\item Let $\LLL$ be the set of least-squares solutions and $\mu \ceq \inf\qty{\norm{Au - f} \mid u \in \XX}$, then we can write
+		\[
+		\LLL = \qty{u \in \XX : \norm{Au - f} \leq \mu}. 
+		\]
+		Since $\norm{\cdot}_\YY$ is $\tau_\YY$-LSC and $A \colon \tau_\XX \to \tau_\YY$ is continuous, we have by \cref{lem:lsc_closed_sets} that $\LLL$ is $\tau_\XX$-closed.
+
+		Now consider the problem 
+		\[
+		\inf_{u \in \LLL} \JJ(u) = \inf_{u \in \XX} \JJ(u) + \chi_\LLL(u). 
+		\] 
+		By assumption there exists $u \in \LL$ with $\JJ(u) < \infty$, and the objective function $\JJ + \chi_\LL$ is bounded from below (by 0).
+Using similar arguments as in the previous lemma (???), it can be shown that it is coercive.
+Finally, characteristic functions of closed sets are LSC, so $\chi_\LLL$ and $\JJ$ are both $\tau_\XX$-LSC, and therefore their sum is $\tau_\XX$-LSC as well.
+By the direct method (\cref{thm:direct_method}), we conclude that a $\JJ$-minimising solution exists.
+
+		\item The objective function $\Phi_\alpha$ is coercive by the previous lemma, and also bounded from below.
+It is easily seen that $\Phi_\alpha$ is $\tau_\XX$-LSC, and using the direct method we conclude that $\Phi_\alpha$ has a minimiser.
+	\end{enumerate}
+\end{proof}

### 2020-11-02 modern_statistical_methods/msm_ES1_handin.tex
handing in one msm question

--- 2020-10-20 modern_statistical_methods/msm_ES1.tex
@@
 \author{Lucas Riedstra}
-\date{...} % date
-
 \begin{document}
@@
 
-\begin{qbox}
-    Consider minimising the following objective involving response $Y \in \RR^n$ and design matrix $X \in \RR^{n \times p}$ over $(\mu, \beta) \in \RR \times \RR^p$: 
-    \[
-    \norm{Y - \mu \Bf{1} - X\beta}_2^2 + J(\beta).
-    \]
-    Here $J \colon \RR^p \to \RR$ is an arbitrary penalty function.
-Suppose $\bar{X_k} = 0$ for $k = 1, \dotsc, p$.
-Assuming that a minimiser $(\hat\mu, \hat\beta)$ exists, show that $\hat\mu = \bar Y$.
-Now take $J(\beta) = \lambda \norm{\beta}_2^2$ so we have the ridge regression objective.
-Show that
-    \[
-    \hat\beta = (X\T X + \lambda I)^{-1} X\T Y.
-    \] 
-    From here onwards, whenever we refer to ridge regression, we will assume $X$ has had its columns mean-centred. 
-\end{qbox}
-
-\begin{solution}
-    If a minimiser $\hat\mu$ exists, then the derivative of the objective must be 0. We expand the objective function as
-    \begin{align*}
-    \norm{Y - \mu\Bf{1} - X\beta}_2^2 + J(\beta) &= \sum_{i=1}^n \qty(Y_i - \mu - (X\beta)_i)^2 + J(\beta) \\
-    &= \sum_{i=1}^n \qty(Y_i - \mu - \sum_{j=1}^p X_{ij}\beta_j)^2. 
-    \end{align*}
-Now, differentiating w.r.t.\ $\mu$ yields
-\begin{align*}
-\pdv{\mu} \norm{Y - \mu\Bf{1} - X\beta}_2^2 + J(\beta) &= -2 \sum_{i=1}^n \qty( Y_i - \mu - \sum_{j=1}^p X_{ij}\beta_j) \\
-&= -2 \qty(\sum_{i=1}^n Y_i  - n\mu - \sum_{i=1}^n \sum_{j=1}^p X_{ij}\beta_j) \\
-&= -2 \qty(\sum_{i=1}^n Y_i  - n\mu - \sum_{j=1}^p \beta_j (\sum_{i=1}^p X_{ij})) \\
-&= -2 \qty( \sum_{i=1}^n Y_i - n\mu), 
-\end{align*}
-since $\sum_{i=1}^p X_{ij} = 0$ for all $i$ by assumption.
-Now, setting the derivative to 0 shows that $\hat\mu = \frac1n \sum_{i=1}^n Y_i = \bar Y$ is the minimiser.
-Since the second derivative of the objective function is constantly $+2$, this $\hat\mu$ is indeed a minimum.
-
-The computation of $\hat\beta$ is given in the lecture notes. 
-\end{solution}
-
-\begin{question}
-    Consider performing ridge regression when $Y = X \beta^0 + \eps$, where $X \in \RR^{n \times p}$ has full column rank, and $\Var(\eps) = \sigma^2 I$.
-Let the SVD of $X$ be $UDV\T$ and write $U\T X \beta^0 = \gamma$.
-Show that
-    \[
-    \frac1n \EE\norm{X \beta^0 - X \hat\beta_\lambda^R}_2^2 = \frac1n \sum_{j=1}^p \qty(\frac\lambda{\lambda + D_{jj}^2})^2 \gamma_j^2 + \frac{\sigma^2}{n} \sum_{j=1}^p \frac{D_{jj}^4}{(\lambda + D_{jj}^2)^2}.
-    \]
-\end{question}
-
-\begin{solution}
-    We write $\hat\beta \ceq \hat\beta_\lambda^R$. We write out
-    \[
-    X\hat\beta = UDV\T (V\T D^2 V + \lambda I)^{-1} V\T D U\T Y =  UD^2 (D^2 + \lambda I)^{-1} U\T Y = \sum_j \frac{D_{jj}^2}{\lambda + D_{jj}^2} U_j U_j\T Y. 
-    \]
-    Since $X\beta^0 = UDV\T \beta^0$ lies in the range of $U$, we have $X\beta^0 = \sum_j U_j U_j\T X\beta^0$. 
-    \begin{align*}
-        X\beta^0 - X\hat\beta &= \sum_j U_j U_j\T X \beta^0 -  \sum_j \frac{D_{jj}^2}{\lambda + D_{jj}^2} U_j U_j\T Y \\
-        &= \sum_j \qty(U_j\T X \beta^0 - \frac{D_{jj}^2 U_j\T Y}{\lambda + D_{jj}^2}) U_j \\
-        &= \sum_j \qty(\frac{D_{jj}^2 U_j\T (X\beta^0 - Y) + \lambda U_j\T X \beta^0}{\lambda + D_{jj}^2})
-    \end{align*}
-and therefore
-\begin{align*}
-    \frac1n \EE\norm{X\beta^0 - X\hat\beta}_2^2 &= \frac1n \sum_j \EE \qty(\frac{D_{jj}^2 U_j\T (X\beta^0 - Y) + \lambda U_j\T X \beta^0}{\lambda + D_{jj}^2})^2 \\
-    &\overset{*}= \frac1n \sum_j \frac{D_{jj}^4}{(\lambda + D_{jj}^2)^2} \EE[(U_j\T (X\beta^0 - Y))^2] + \frac1n \sum_j \qty(\frac{\lambda}{\lambda + D_{jj}^2})^2  (U_j\T X \beta^0)^2 \\
-&\overset{**}= \frac{\sigma^2}{n} \sum_j \frac{D_{jj}^4}{(\lambda + D_{jj}^2)^2}  + \frac1n \sum_j \qty(\frac{\lambda}{\lambda + D_{jj}^2})^2  \gamma_j^2.
-\end{align*}
-Note that the cross-terms in $(*)$ disappear since they are linear combinations of $\EE(X\beta^0 - Y) = 0$, while the equality $**$ holds since 
-\[
-\EE[(U_j\T(X \beta^0 - Y))^2] = \Var(U_j\T (X\beta^0 - Y)) = U_j\T \Var(X\beta^0 - Y) U_j = \sigma^2 U_j\T U_j = \sigma^2. 
-\]
-\end{solution}
-
-\begin{question}
-    Show that the ridge regression estimates can be otained by ordinary least squares regression on an augmented data set with $\sqrt\lambda I$ added to the botten of $X$, and $p$ zeroes added to the end of the response $Y$.
-\end{question}
-
-\begin{solution}
-    Let $\tilde X, \tilde Y$ be the new data set, then our ordinary least squares objective function is
-    \begin{align*}
-    L(\beta) & \ceq \norm{\tilde Y - \tilde X \beta}^2 \\
-    &= \norm{Y - X\beta}^2 + \norm{\sqrt\lambda\beta}^2 \\
-    &= \norm{Y - X\beta}^2 + \lambda \norm{\beta}^2,
-    \end{align*}
-which is exactly our ridge regression objective function. 
-\end{solution}
-
-\begin{question}
-    In the following, assume that forming $AB$ where $A \in \RR^{a \times b}, B \in \RR^{b \times c}$ requires $O(abc)$ computational operations, and that if $M \in \RR^{d \times d}$ is invertible, then forming $M^{-1}$ requires $O(d^3)$ operations.
-    \begin{enumerate}[(a)]
-        \item Suppose we wish to apply ridge regression to data $(Y, X) \in \RR^n \times \RR^{n \times p}$ with $n \gg p$.
-A complication is that the data is split into $m$ separate datasets of size $n/m \in \NN$,
-        \[
-        Y = \mqty( Y^{(1)} \\ \vdots \\ Y^{(m)}), \quad X = \mqty( X^{(1)} \\\vdots\\X^{(m)}),
-        \]
-        with each dataset located on a different server.
-Moving large amounts of data between servers is expensive.
-Explain how one can produce ridge estimates $\hat\beta_\lambda$ by communicating only $O(p^2)$ numbers from each server to some central server.
-What is the total order of the computation time required at each server, and at the central server for your approach?
-
-        \item Now suppose instead that $p \gg n$ and it is instead the variable sthat are split across $m$ servers, so each server has only a subset of $p/m \in \NN$ variables for each observation, and some central server stores $Y$.
-Explain how one can obtain the fitted values $X\hat\beta_\lambda$ communicating only $O(n^2)$ numbers from each server to the central server.
-What is the total order of the computation time required at each server, and at the central server for your approach?
-    \end{enumerate}
-\end{question}
-
-\begin{solution}
-    \begin{enumerate}[(a)]
-        \item Note that
-        \[
-        X\T Y = \mqty( X^{(1)\top} \dotsc X^{(m)\top}) \mqty( Y^{(1)}\\\vdots\\Y^{(m)}) = \sum_{i=1}^m X^{(i)\top} Y^{(i)}. 
-        \]
-        Each of the $X^{(i)\top} Y^{(i)}$ can be computed at the data server (requiring $O(np)$ computation) and produces $p$ numbers per data server.
-
-        Analogously, we have
-        \[
-        X\T X = \sum_{i=1}^m X^{(i)\top} X^{(i)},
-        \]
-        which can also be computed at the data server (requiring $O(np^2)$ computation) and produces $p^2$ numbers per data server.
-In the end, each server will therefore have $O(p^2)$ numbers to send, having done $O(np^2)$ computations.
-
-        When each server sends their numbers to the central data server, the central data server can then compute $(X\T X + \lambda I)^{-1}$ using $O(p^3)$ computations and then multiply that with $X\T Y$ which will require $O(np)$ computations.
-
-        \item In this case, we have 
-        \[
-        X = \mqty( X^{(1)} \dotsb X^{(m)} ). 
-        \]
-        We use the alternative form of  $\hat\beta_\lambda$, namely
-        \[
-        X \hat\beta_\lambda^R = XX\T (XX\T + \lambda I)^{-1} Y.
-        \]
-
-        Now, we have $XX\T = \sum_{i=1}^m X^{(i)} X^{(i)\top}$, and each $X^{(i)} X^{(i)\top}$ can be computed at the data server, requiring $O(n^2 p)$ computations and producing $O(n^2)$ numbers to send.
-
-        After sending these numbers over to the central server, the inverse $(XX\T + \lambda I)^{-1}$ can be computed in $O(n^3)$ time, then multiplied from the left with $XX\T$ in $O(n^3)$ time, and multiplied with $Y$ in $O(n^2)$ time.
-The total amount of computation at the central server will therefore be $O(n^3)$ as well.
-    \end{enumerate}
-\end{solution}
-
-\begin{question}
-	Prove Proposition 4 in the notes.
-\emph{Hint: for part (ii) it may help to consider the eigendecompositions of positive semi-definite matrices $K^{(1)}$ and $K^{(2)}$ derived from kernels $k_1$ and $k_2$ in the form $K^{(1)} = PDP\T = \sum_{i=1}^n P_iP_i\T D_{ii}$ for example.
-}
-\end{question}
-
-\begin{proof}
-	Proposition 4 is the following:
-	\begin{mdframed}
-		Suppose $k_1, k_2, \dotsc$ are kernels. 
-		\begin{enumerate}[(i)]
-			\item If $\alpha_1, \alpha_2 \geq 0$ then $\alpha_1 k_1 + \alpha_2 k_2$ is a kernel. If $\lim_{k\to\infty} k(x, x') \eqqcolon k(x, x')$ exists for all $x, x' \in \Cal X$, then $k$ is a kernel. 
-			\item The pointwise product $k = k_1k_2$ is a kernel.
-		\end{enumerate}
-	\end{mdframed}
-
-Let $x_1, \dotsc, x_n \in \Cal X$ and define the positive definite matrix $K^{(m)}$ by $K^{(m)}_{i, j} = k_m(x_i, x_j)$. 
-\begin{enumerate}[(i)]
-	\item Since $(\alpha_1 k_1 + \alpha_2 k_2)(x_i, x_j) = \alpha_1 k_1(x_i, x_j) + \alpha_2 k_2(x_i, x_j)$, the kernel matrix $K_{ij} \ceq (\alpha_1 k_1 + \alpha_2 k_2)(x_i, x_j)$ equals $\alpha_1 K_1 + \alpha_2 K_2$, and is therefore clearly symmetric positive semi-definite.
-This shows that $\alpha_1 k_1 + \alpha_2 k_2$ is a kernel.
-
-	For the second part (the pointwise limit of kernels is a kernel), it suffices to show that the entrywise limit of positive semi-definite matrices is again positive semidefinite.
-But this is clear: if $K^{(m)} \to K$ entrywise, then for any $\vec x \in \RR^n$ we have
-	\[
-	\vec x\T K \vec x = \vec x\T (\lim_{m\to\infty} K^{(m)}) \vec x = \lim_{m\to\infty} \vec x\T K^{(m)}\vec x \geq 0,
-	\]
-	so $K$ is positive semi-definite.
-
-	\item Let $K^{(1)}$ and $K^{(2)}$ be positive semi-definite, then we must show that the pointwise product $K^{(1)} \circ K^{(2)}$ is positive semi-definite as well.
-Write $K^{(1)} = \sum_{\ell=1}^n \lambda_\ell P_\ell P_\ell\T$ and $K^{(2)} = \sum_{\ell=1}^n \mu_\ell Q_\ell Q_\ell\T$, where $\lambda_i, \mu_i \geq 0$ and $\qty{P_\ell}$, $\qty{Q_\ell}$ are orthonormal bases of $\RR^n$.
-
-	Then we find
-	\begin{align*}
-	(K^{(1)} \circ K^{(2)})_{ij} &= \qty(\sum_m \lambda_m P_{im} P_{jm}) \qty(\sum_\ell \mu_\ell Q_{i\ell} Q_{j\ell}) = \sum_{m, \ell} \lambda_m \mu_\ell (P_{im} P_{jm} Q_{i\ell} Q_{j\ell}) \\
-	&= \sum_{m, \ell} \lambda_m \mu_\ell \qty[(P_m \circ Q_\ell)(P_m \circ Q_\ell)\T]_{ij},
-	\end{align*}
-
-	and therefore $K^{(1)} \circ K^{(2)} = \sum_{m, \ell} \lambda_m \mu_\ell \qty[(P_m \circ Q_\ell)(P_m \circ Q_\ell)\T]$, which is a nonnegative combination of positive semi-definite matrices and therefore positive semi-definite.
-\end{enumerate}
-\end{proof}
-
-\begin{question}
-	Let $\XX \ceq \qty{x \in \RR^d : \norm{x} < 1}$. Show that $k(x, x') \ceq (1 - x\T x')^{-\alpha}$ defined on $\XX \times \XX$, where $\alpha > 0$, is a kernel.
-\end{question}
-
-\begin{proof}
-	Note by Cauchy-Schwarz we have for $x, x' \in \XX$ that $\abs{x\T x'} \leq \norm{x} \norm{x'} < 1$, so $k$ is well-defined.
-
-	Symmetry of $k$ is clear. Define the function 
-	\[
-	\phi \colon (-\infty, 1) \to \RR \colon x \mapsto (1-x)^{-\alpha}. 
-	\]
-	Then $\phi$ is continuous, and we can compute its Taylor series around 0. It is computed that 
-	\[
-	\phi^{(k)}(x) = (1-x)^{-\alpha - k} \prod_{j=0}^{k-1}(\alpha + j) \implies \phi^{(k)}(0) = \prod_{j=0}^{k-1} (\alpha + j). 	\]
-	Therefore, the Taylor series of $\phi$ is given by
-	\[
-	\sum_{k=0}^\infty \qty(\prod_{j=0}^{k-1} (\alpha + j)) \frac{x^k}{k!}, 
-	\]
-	and its radius of convergence is given by
-	\[
-	\lim_{k\to\infty} \frac{\frac{1}{k!}\prod_{j=0}^{k-1} (\alpha + j)}{\frac{1}{(k+1)!} \prod_{j=0}^k (\alpha + j)} = \lim_{k\to\infty} \frac{k}{\alpha + k} = 1,
-	\]
-	so $\phi$ agrees with its Taylor series on $(-1, 1)$. By Cauchy-Schwarz, for $x, x' \in \XX$ we have $x_i\T x_j \leq \norm{x_i} \norm{x_j} < 1$, so we can write for $x, x' \in \XX$
-	\[
-	k(x, x') = \phi(x\T x') = \sum_{k=0}^\infty \qty(\prod_{j=0}^{k-1} (\alpha + j)) \frac{(x\T x')^k}{k!}.
-	\]
-	By proposition 4, every term in the sum is a kernel, and therefore the limit is a kernel as well. 
-\end{proof}
+\setcounter{question}{6}
 
@@
 	\end{align*}
-	We have therefore rewritten $\ang{Z_i, Z_j}$ into something that can be computed in $O(p)$, which is exactly the aim of the kernel trick.
+	We have therefore rewritten $\ang{Z_i, Z_j}$ into something that can be computed in $O(p)$, which is exactly the aim of the kernel trick, since this allows computation of the kernel matrix $K$ in $O(n^2 p)$.
 \end{proof}
 
-\begin{question}
-	Let $\hat\alpha$ be a minimiser of $\norm{Y - K\alpha}^2 + \lambda \alpha\T K\alpha$, with $K$ being a kernel matrix as usual. Show that $K\hat\alpha = K(K + \lambda I)^{-1} Y$. 
-\end{question}
-
-\begin{solution}
-	Suppose $\hat\alpha$ minimises $Q(\alpha) \ceq \norm{Y - K\alpha}^2 + \lambda \alpha\T K\alpha$. Then we have
-	\begin{align*}
-		0 = \nabla_\alpha Q(\hat\alpha) &= -2 K (Y - K\hat\alpha) + 2\lambda K\hat\alpha \\
-		\lambda K\hat\alpha &= KY - K^2\hat\alpha \\
-		(K + \lambda) K\hat\alpha &= KY \\
-		K\hat\alpha &= (K + \lambda)^{-1} KY.
-	\end{align*}
-Since $K$ and $(K + \lambda)^{-1}$ are simultaneously diagonalisable, they commute, and therefore $K\hat\alpha = K(K + \lambda)^{-1} Y$. 
-%\begin{align*}
-%(K + \lambda)^{-1} KY &= (U(D + \lambda) U\T)^{-1} UDU\T Y \\
-%&= U(D + \lambda)^{-1} D U\T Y \\
-%&= UD (D+ \lambda)^{-1} U\T Y\\
-%&= UDU\T (U (D+ \lambda) U\T) Y\\
-%&= K(K + \lambda)^{-1} Y,
-%\end{align*}
-%which proves the claim.
-\end{solution}
-
-\begin{question}
-	...
-\end{question}
-
-\begin{proof}
-	content...
-\end{proof}
-
-\begin{question}
-	...
-\end{question}
-
-\begin{proof}
-	...
-\end{proof}
-
-\begin{question}
-	Show from first principles that the Sobolev kernel is indeed a (positive definite) kernel.
-\end{question}
-
-\begin{solution}
-	Let $\XX = [0, 1]$ and $k(x, x') = \min(x, x')$ the Sobolev kernel.
-We must then show that for any $x_1, \dotsc, x_n \in \XX$, the matrix $K_{ij} = \min(x_i, x_j)$ is positive semi-definite (symmetry of $k$ is clear).
-
-	To this end, assume $x_1 \leq \dotsb \leq x_n$ (this may require relabeling of the $x_i$ which does not affect positive semi-definiteness of $K$).
-Now let $\vec f_i$ denote the vector $f_{i1}, \dotsc, f_{i(i-1)} = 0$, $f_{ii}, \dotsc, f_{in} = 1$.
-	\[
-	K = \mqty[ x_1 & x_1 & x_1 & \cdots & x_1 \\
-	x_1 & x_2 & x_2 & \dotsb & x_2 \\
-	x_1 & x_2 & x_3 & \dotsb & x_3 \\
-	\vdots & \vdots & \vdots & \ddots & \vdots \\
-	x_1 & x_2 & x_3 & \cdots & x_n] = x_1 \vec f_1 \vec f_1\T + (x_2 - x_1) \vec f_2 \vec f_2\T + \dotsb + (x_n - x_{n-1}) \vec f_n \vec f_n\T. 
-	\]
-	Since this is a nonnegative combination of positive semidefinite matrices, we conclude that $K$ is positive semidefinite. 
-\end{solution}
 \end{document}

### 2020-11-03 distribution_theory/C2.tex
moaaaaar updates
updated distribution theory, still struggling with topics ugh

@@
 
-\subsection{Convolution between $\EE'(\RR^n)$ and $\DD'(\RR^n)$}
+\subsection{Convolution between distributions}
 \begin{definition}
@@
 
-	We define the \emph{Schwarz space}
+	We define the \emph{Schwartz space}
 	\[
@@
 	\]
+	The space of tempered distributions is denoted by $\SS'(\RR^n)$. 
 \end{definition}
+
 This definition is equivalent to sequential continuity.
 
-\subsection{The Fourier transform on $\SS(\RR^n)$}
+\subsection{The Fourier transform on Schwartz functions}
 
@@
 	\[
-	[\FF(f)](\lambda) = \hat f(\lambda) \ceq \int_{\RR^n} e^{-i\lambda x} f(x) \dd{x} \quad\text{where $\lambda \in \RR^n$}. 
+	[\FF(f)](\lambda) = \hat f(\lambda) \ceq \int_{\RR^n} e^{-i\lambda \vdot x} f(x) \dd{x} \quad\text{where $\lambda \in \RR^n$}. 
 	\]
@@
 	\[
-	\hat f(\lambda_m) = \int_{\RR^n} e^{-i\lambda_m x} f(x) \dd{x} \overset{\text{DCT}}= \int_{\RR^n} e^{-i\lambda x} f(x) \dd{x} = \hat f(\lambda),
+	\hat f(\lambda_m) = \int_{\RR^n} e^{-i\lambda_m \vdot x} f(x) \dd{x} \overset{\text{DCT}}= \int_{\RR^n} e^{-i\lambda \vdot x} f(x) \dd{x} = \hat f(\lambda),
 	\]
@@
 
-\begin{lemma}
+\begin{lemma} \label{lem:differentiate_fourier}
 	For $\phi \in \SS(\RR^n)$, we have $\FF[D_x^\alpha \phi](\lambda) = \lambda^\alpha \hat\phi(\lambda)$ and $\FF[x^\beta\phi](\lambda) = (-1)^{\abs{\beta}} D_\lambda^\beta \hat\phi(\lambda)$. 
@@
 	\begin{align*}
-		\FF[D_\lambda^\alpha\phi](\lambda) &= \int_{\RR^n} e^{-i\lambda x} D_x^\alpha \phi(x) \dd{x} \\
-		&= (-1)^{\abs{\alpha}} \int_{\RR^n} D_x^\alpha(e^{-i\lambda x}) \phi(x) \dd{x} \\
+		\FF[D_\lambda^\alpha\phi](\lambda) &= \int_{\RR^n} e^{-i\lambda \vdot x} D_x^\alpha \phi(x) \dd{x} \\
+		&= (-1)^{\abs{\alpha}} \int_{\RR^n} D_x^\alpha(e^{-i\lambda \vdot x}) \phi(x) \dd{x} \\
 		&= \lambda^\alpha\hat\phi(\lambda). 
@@
 	\begin{align*}
-		 \FF[x^\beta \phi](\lambda) &=  \int_{\RR^n} e^{-i\lambda x} x^\beta \phi(x) \dd{x} \\
-		 &= \int_{\RR^n} ((-D_\lambda)^\beta e^{-i\lambda x}) \phi(x) \dd{x} \\
+		 \FF[x^\beta \phi](\lambda) &=  \int_{\RR^n} e^{-i\lambda \vdot x} x^\beta \phi(x) \dd{x} \\
+		 &= \int_{\RR^n} ((-D_\lambda)^\beta e^{-i\lambda \vdot x}) \phi(x) \dd{x} \\
 		 &= (-1)^{\abs{\beta}} D_\lambda^\beta \hat\phi(\lambda).
@@
 \[
-\FF^{-1}[\hat f](x)  \ceq \frac{1}{(2\pi)^n} \int_{\RR^n} e^{i\lambda x} \hat f(\lambda) \dd{\lambda}. 
+\FF^{-1}[\hat f](x)  \ceq \frac{1}{(2\pi)^n} \int_{\RR^n} e^{i\lambda \vdot x} \hat f(\lambda) \dd{\lambda}. 
 \]
 We will now show that on $\SS(\RR^n)$, the inverse Fourier transform is indeed an inverse:
-\begin{theorem}
+\begin{theorem} \label{thm:fourier_transform_isomorphism}
 	The Fourier transform $\FF$ defines a continuous isomorphism from $\SS(\RR^n)$ to itself. 
@@
 	\begin{align}
-		\abs{\lambda^\alpha (-D_\lambda)^\beta \hat\phi(\lambda)} &= \abs{\lambda^\alpha \FF[x^\beta\phi](\lambda)} = \abs{\FF[D^\alpha_x (x^\beta\phi)](\lambda)} = \abs{\int_{\RR^n} e^{-i\lambda x} D^\alpha(x^\beta\phi) \dd{x}} \nonumber \\
-		&\leq \int_{\RR^n} \abs{D^\alpha (x^\beta\phi)} \dd{x} \label{eq:fourier_schwarz_estimate},
+		\abs{\lambda^\alpha (-D_\lambda)^\beta \hat\phi(\lambda)} &= \abs{\lambda^\alpha \FF[x^\beta\phi](\lambda)} = \abs{\FF[D^\alpha_x (x^\beta\phi)](\lambda)} = \abs{\int_{\RR^n} e^{-i\lambda \vdot x} D^\alpha(x^\beta\phi) \dd{x}} \nonumber \\
+		&\leq \int_{\RR^n} \abs{D^\alpha (x^\beta\phi)} \dd{x} \label{eq:fourier_Schwartz_estimate},
 	\end{align}
-which is finite since $D^\alpha (x^\beta\phi)$ is also a Schwarz function and therefore integrable.
+which is finite since $D^\alpha (x^\beta\phi)$ is also a Schwartz function and therefore integrable.
 
 From the previous lemma we also infer that $\hat\phi$ is smooth, so indeed we have $\hat\phi \in \SS(\RR^n)$. 
-From \cref{eq:fourier_schwarz_estimate} it is also easily seen that if $\phi_m \to 0$ in $\SS(\RR^n)$, then $\hat\phi_m \to 0$ in $\SS(\RR^n)$ also, which shows that $\FF$ is continuous.
+From \cref{eq:fourier_Schwartz_estimate} it is also easily seen that if $\phi_m \to 0$ in $\SS(\RR^n)$, then $\hat\phi_m \to 0$ in $\SS(\RR^n)$ also, which shows that $\FF$ is continuous.
 
@@
 \begin{align*}
-	\FF^{-1} [\FF[\phi]](x) &= \frac{1}{(2\pi)^n} \int_{\RR^n} \int_{\RR^n} e^{i\lambda (x-y)} \phi(y) \dd{y}\dd{\lambda} \\
-	&= \frac{1}{(2\pi)^n} \lim_{\eps \downarrow 0} \int_{\RR^n} \int_{\RR^n} e^{i \lambda ( x- y) - \eps\norm{\lambda}^2} \phi(y) \dd{y}\dd{\lambda} \\
-	&\overset\star= \frac{1}{(2\pi)^n}\lim_{\eps \downarrow 0} \int_{\RR^n} \phi(y)\int_{\RR^n} e^{i\lambda(x- y) - \eps\norm{\lambda}^2} \dd{\lambda} \dd{y},
+	\FF^{-1} [\FF[\phi]](x) &= \frac{1}{(2\pi)^n} \int_{\RR^n} \int_{\RR^n} e^{i\lambda \vdot (x-y)} \phi(y) \dd{y}\dd{\lambda} \\
+	&= \frac{1}{(2\pi)^n} \lim_{\eps \downarrow 0} \int_{\RR^n} \int_{\RR^n} e^{i \lambda \vdot ( x- y) - \eps\norm{\lambda}^2} \phi(y) \dd{y}\dd{\lambda} \\
+	&\overset\star= \frac{1}{(2\pi)^n}\lim_{\eps \downarrow 0} \int_{\RR^n} \phi(y)\int_{\RR^n} e^{i\lambda \vdot (x- y) - \eps\norm{\lambda}^2} \dd{\lambda} \dd{y},
 	%&= \frac{1}{(2\pi)^n} \int_{\RR^n}\int_{\RR^n} e^{i\lambda(x- y)} \dd{\lambda} \dd{y}
@@
 \[
-\int_{\RR^n} e^{i\lambda(x-y) - \eps\norm{\lambda}^2} \dd{\lambda} = \prod_{i=1}^n \int_{\RR^n} e^{i\lambda_j(x_j - y_j) - \eps\lambda_j^2} \dd{\lambda} \overset{\star\star}= \prod_{i=1}^n \qty(\frac\pi e)^{1/2} e^{- \frac{(x_i - y_i)^2}{4\eps}} = \qty(\frac\pi\eps)^{n/2} e^{- \frac{\norm{x - y}^2}{4\eps}}.
+\int_{\RR^n} e^{i\lambda\vdot (x-y) - \eps\norm{\lambda}^2} \dd{\lambda} = \prod_{i=1}^n \int_{\RR} e^{i\lambda_j(x_j - y_j) - \eps\lambda_j^2} \dd{\lambda} \overset{\star\star}= \prod_{i=1}^n \qty(\frac\pi e)^{1/2} e^{- \frac{(x_i - y_i)^2}{4\eps}} = \qty(\frac\pi\eps)^{n/2} e^{- \frac{\norm{x - y}^2}{4\eps}}.
 \] 
@@
 
-Finally, continuity of $\FF^{-1}$ is easily shown with an argument analogous to that for continuity of $\FF$. 
+Finally, continuity of $\FF^{-1}$ is easily shown with an argument analogous to that for continuity of $\FF$ (????). 
 \end{proof}
+
+\subsection{The Fourier transform on tempered distributions}
+\begin{lemma}
+	For $\phi, \psi \in \SS(\RR^n)$, we have $\int_{\RR^n} \phi(x) \hat\psi(x) \dd{x} = \int_{\RR^n} \hat\phi(x) \psi(x) \dd{x}$. 
+\end{lemma}
+
+\begin{proof}
+	This follows from Fubini's theorem:
+	\[
+	\int_{\RR^n}\phi(x) \hat \psi(x) \dd{x} = \int_{\RR^n}\int_{\RR^n} \phi(x) \psi(\lambda) e^{-i\lambda \vdot x} \dd{\lambda} \dd{x} = \int_{\RR^n}\int_{\RR^n} \psi(\lambda)  \phi(x) e^{-i\lambda \vd x} \dd{x} \dd{\lambda} = \psi(\lambda) \psi(\lambda) \hat\phi(\lambda) \dd{\lambda}.
+	\]
+\end{proof}
+
+The above result can be rewritten as $\ang{\hat\phi, \psi} = \ang{\phi, \hat\psi}$, which motivates the definition of the Fourier transform for tempered distributions:
+\begin{definition}
+	For $u \in \SS'(\RR^n)$, we define its Fourier transform by
+	\[
+	\ang{\hat u, \phi} \ceq \ang{u, \hat\phi} \quad\text{for all $\phi \in \SS(\RR^n)$}. 
+	\]
+\end{definition}
+Using sequential continuity and \cref{thm:fourier_transform_isomorphism}, it is easily seen that $\hat u$ is indeed an element of $\SS'(\RR^n)$.
+
+\begin{example}
+	It is easily checked that $\delta_0 \in \SS'(\RR^n)$, and we can compute
+	\[
+	\ang{\hat{\delta_0}, \phi} = \ang{\delta_0, \hat\phi} = \hat\phi(0) = \int_{\RR^n} \phi(x) \dd{x} = \ang{1, \phi},
+	\]
+	so we can write $\hat{\delta_0} = 1$. Analogously, by the Fourier inversion theorem we have
+	\[
+	\ang{\hat 1, \phi} = \ang{1, \hat\phi} = \int_{\RR^n} \hat\phi(\lambda) \dd{\lambda} = (2\pi)^n \phi(0) = \ang{(2\pi)^n \delta_0, \phi}, 
+	\]
+	so we write $\hat 1 = (2\pi)^n \delta_0$. 
+\end{example}
+
+We can easily generalise \cref{lem:differentiate_fourier} to the Fourier transform on $\SS'(\RR^n)$, so
+\[
+\FF[D^\alpha u] = \lambda^\alpha \hat u, \quad \FF[x^\beta u] = (-D^\beta) \hat u. 
+\]
+
+\begin{theorem}
+	The Fourier transform $\FF$ extends to a continuous isomorphism on $\SS'(\RR^n)$. 
+\end{theorem}
+
+\begin{proof}
+	We claim that $\refl u = (2\pi)^{-n} \FF[\FF[u]]$. To check this, note that by the Fourier inversion theorem we have for $\phi \in \SS(\RR^n)$ that
+	\[
+	\refl\phi(x) = \phi(-x) = (2\pi)^{-n} \int_{\RR^n} e^{-i\lambda\vd x} \hat\phi(\lambda) \dd{\lambda} = (2\pi)^{-n} \FF[\FF[\phi]](x),
+	\]
+	and therefore 
+	\[
+	\ang{\refl u, \phi} = \ang{u, \refl\phi} = \ang{u, (2\pi)^{-n} \FF[\FF[\phi]]} = \ang{(2\pi)^{-n} \FF[\FF[u]], \phi}. 
+	\]
+
+	This shows that $\FF$ is bijective (since $\FF\circ \FF$ is bijective). For continuity of $\FF$ and its inverse: using \cref{thm:fourier_transform_isomorphism}, we find
+	\begin{align*}
+	&\phantom{\iff} u_m \to 0 \text{ in $\SS'(\RR^n)$} \\
+	&\iff \ang{u_m, \phi} \to 0 \ \forall \phi \in \SS(\RR^n) \\
+	&\iff \ang{u_m, \hat\phi} \to 0 \ \forall \phi \in \SS(\RR^n) \\
+	&\iff \ang{\hat u_m, \phi} \to 0 \ \forall \phi \in \SS(\RR^n) \\
+	&\iff \hat u_m \to 0 \text{ in $\SS'(\RR^n)$}.
+	\end{align*}
+\end{proof}
+
+\subsection{Sobolev spaces}
+\begin{convention}
+	We write $\ang{\lambda} \ceq (1 + \norm{\lambda})^2$ for $\lambda \in \RR^n$.
+Note that $\ang{\lambda} \sim 1$ as $\norm{\lambda} \to 0$ and $\ang{\lambda} \to \norm{\lambda}$ as $\norm{\lambda} \to \infty$.
+\end{convention}
+
+\begin{definition}
+	For $s \in \RR$, define the \emph{Sobolev space} $H^s(\RR^n)$ to be the set of tempered distributions $u \in \SS'(\RR^n)$ for which $\hat u$ can be identified with a measurable function $\hat u \colon \RR^n \to \CC$ such that $\ang{\lambda}^s \hat u \in L^2(\RR^n)$.
+\end{definition}
+
+% Suppose we have an open subset $X \subseteq \RR^n$, then we can define the \emph{localised} Sobolev space $H^s_\Rm{loc}(X)$ by sayin
+
+\begin{lemma}[Sobolev lemma]
+	If $u \in H^s(\RR^n)$ with $s > \frac n2$, then  $u$ is continuous. 
+\end{lemma}
+
+\begin{proof}
+	We will show that $\hat u$ is integrable. By Cauchy-Schwarz, we have
+	\begin{align*}
+		\int_{\RR^n} \abs{\hat u(\lambda)} \dd{\lambda} &= \int_{\RR^n}  \ang{\lambda}^{-s} \ang{\lambda}^s \abs{\hat u(\lambda)} \dd{\lambda} \\
+		&\leq \qty(\int_{\RR^n}  \ang{\lambda}^{-2s}\dd{\lambda})^{1/2} \norm{u}_{H^s} \\
+		&= C \norm{u}_{H^s} \qty(\int_0^\infty r^{n-1} (1 + r^2)^{-s} \dd{r})^{1/2}, 
+	\end{align*}
+where the last line follows from using polar coordinates and $C$ is the area of the $(n-1)$-sphere.
+
+Writing $s = \frac n2 + \eps$, we find that the integrand $r^{n-1}(1 + r^2)^{-s}$ is of order $O(r^{-1 - 2\eps})$ as $r \to\infty$, and therefore the integral is finite, so indeed we have $\hat u \in L^1(\RR^n)$.
+
+By applying \cref{thm:fourier_transform_isomorphism} to a test function, we can show that $u = (2\pi)^{-n} \int_{\RR^n} e^{i\lambda \vd x} \hat u(\lambda) \dd{\lambda}$, which is continuous by the dominated convergence theorem.
+\end{proof}

### 2020-11-03 topics_statistical_theory/topics_ES1.tex
distribution theory is still difficult :((
updated distribution theory, still struggling with topics ugh

@@
 \usepackage{hyperref}
-\usepackage{mathtools, amsmath, amssymb, amsthm, enumerate, mdframed, bbm, graphicx, float, physics, xcolor, cleveref}
+\usepackage[shortlabels]{enumitem}
+\usepackage{mathtools, amsmath, amssymb, amsthm, mdframed, bbm, graphicx, float, physics, xcolor, cleveref}
 
@@
 \newcommand{\EE}{\Bb E}
+\newcommand\HH{\Cal H}
+\newcommand\FF{\Cal F}
+\newcommand\Nn{\Cal N}
+\renewcommand\SS{\Cal S}
 
@@
 \DeclareMathOperator\Exp{Exp}
+\DeclareMathOperator\MSE{MSE}
+\DeclareMathOperator\MISE{MISE}
+\DeclareMathOperator\Bias{Bias}
 \DeclareMathOperator\Beta{Beta}
@@
 \author{Lucas Riedstra}
-\date{...} % date
+\date{8 November 2020} % date
 
@@
 \]
+
+??
+\end{proof}
+
+\begin{question}
+	For $\beta \in \NN$ and $L > 0$, define the \emph{Sobolev class} $\SS(\beta, L)$ to be the set of $(\beta - 1)$ times differentiable functions $f \colon \RR \to \RR$ for which $f^{(\beta - 1)}$ is absolutely continuous with $L^1$ derivative satisfying $\norm{f^{(\beta)}}_{L^2} \leq L$.
+Recalling the Nikolski class $\Nn(\beta, L)$ from lectures, prove that $\SS(\beta, L) \subseteq \Nn(\beta, L)$.
+Writing $\FF_\SS(\beta, L)$ for the densities in $\SS(\beta, L)$, deduce that a kernel density estimator $\hat f_n$ constructed from $X_1, \dotsc, X_n \iid f \in \FF_\SS(\beta, L)$ with a kernel $K$ of order $\ell \ceq \beta$ and bandwidth $h > 0$ satisfies
+	\[
+	\MISE(\hat f_n) \leq \frac1{nh} R(K) + \frac1{((\ell - 1)!)^2} R(f^{(\beta)}) \mu_\beta^2(K) h^{2\beta}. 
+	\]
+\end{question}
+
+\begin{proof}
+	Let $f \in \SS(\beta, L)$ and $t \in \RR$, then we have 
+	\begin{align*}
+		\int_\RR \qty[f^{(\beta - 1)}(x + t) - f^{(\beta - 1)}(x)]^2 \dd{x} &= \int_\RR \qty[\int_x^{x+t} f^{(\beta)}(y) \dd{y}]^2 \dd{x} \\
+		&= \int_\RR \qty[ \int_\RR \ind_{(x, x+t)}(y) f^{(\beta)}(y) \dd{y} ]^2 \dd{x} \\
+		&\overset{GM}\leq  \qty{ \int_\RR \qty( \int_\RR \ind_{(y-t, y)}(x) f^{(\beta)}(y)^2 \dd{x})^{1/2} \dd{y}}^2 
+	\end{align*}
+??
+\end{proof}
+
+\begin{question}
+	\begin{enumerate}
+		\item 	Verify the algebraic identity
+		\[
+		\phi_\sigma(x-\mu)\phi_{\sigma'}(x-\mu') = \phi_{\sigma\sigma'/(\sigma^2 + \sigma'^2)^{1/2}}(x - \mu^*) \phi_{(\sigma + \sigma')^{1/2}}(\mu - \mu')
+		\]
+		where $\mu^* \ceq (\sigma'^2 \mu + \sigma^2\mu') / (\sigma^2 + \sigma'^2)$, and $\phi_\sigma$ is the $N(0, \sigma^2)$ density.
+
+		\item Let $X_1, \dotsc, X_n \iid N(0, \sigma^2)$.
+Taking $K$ to be the $N(0, 1)$ density, show that the MISE of the kernel density estimate $\hat f_n$ with kernel $K$ and bandwidth $h$ can be expressed exactly as
+		\[
+		\MISE(\hat f_n) = (2\pi)^{-1/2} \qty{\frac1{nh} + (1 - \frac1n) \frac{1}{(h^2 + \sigma^2)^{1/2}} - \frac{2^{3/2}}{(h^2 + 2\sigma^2)^{1/2}} + \frac1\sigma}. 
+		\]
+	\end{enumerate}
+
+\end{question}
+
+\begin{proof}
+	\begin{enumerate}
+		\item \TODO
+		\item no.
+%		We have 
+%		\begin{gather*}
+%			\phi_{\sigma\sigma'/(\sigma^2 + \sigma'^2)^{1/2}}(x - \mu^*) \phi_{(\sigma + \sigma')^{1/2}}(\mu - \mu') \\
+%			= \frac{1}{2\pi \sigma \sigma'} \exp(-\frac12\qty(\frac{(x - \mu^*)^2(\sigma^2 + \sigma'^2)}{\sigma^2 \sigma'^2} + \frac{(\mu - \mu^*)^2}{\sigma^2 + \sigma'^2})) \\
+%			= \frac{1}{2\pi \sigma \sigma'} \exp\qty(-\frac12\qty(\frac{(x - \mu^*)^2(\sigma^2 + \sigma'^2)^2 + \sigma^2 \sigma'^2 (\mu - \mu^*)^2}{\sigma^2 \sigma'^2(\sigma^2 + \sigma'^2)})) \\
+%			=  \frac{1}{2\pi \sigma \sigma'}\exp\qty(-\frac12\qty(\frac{
+%			x^2(\sigma^2 + \sigma'^2)^2 + (\sigma'^2\mu + \sigma^2\mu')^2 - 2x (\sigma'^2\mu + \sigma^2\mu') (\sigma^2 + \sigma'^2)	
+%		}{\sigma^2 \sigma'^2(\sigma^2 + \sigma'^2)}))
+%		\end{gather*}
+	\end{enumerate}
 \end{proof}

### 2020-11-04 distribution_theory/distribution_ES1_handin.tex
progress is being made

--- 2020-11-02 distribution_theory/distribution_ES1.tex
@@
 
-\begin{question}
-	Construct a non-zero element of $\DD(\RR)$ that vanishes outside $(0, 1)$. Construct a non-zero of $\DD(\RR^n)$ that vanishes outside the ball $B_\eps = \qty{x \in \RR^n : \norm{x} < \eps}$. 
-\end{question}
-
-\begin{proof}
-	It is well-known that the function 
-	\[
-	\phi \colon \RR \to \RR \colon x \mapsto \begin{cases}
-		0 &\text{if $x \leq 0$}; \\
-		e^{-1/x} &\text{if $x > 0$},
-	\end{cases}
-	\]
-	is smooth and vanishes outside $(0, \infty)$. The function $\psi(x) \ceq \phi(x)\phi(1-x)$ is therefore also smooth and vanishes outside $(0, 1)$.
-
-	Since $\psi$ vanishes outside $(0, 1)$, the function $\psi(x/\eps)$ vanishes outside $(0, \eps)$, and therefore the function $\vec x \mapsto \psi(\norm{\vec x}/\eps)$ vanishes outside $B_\eps$. 
-\end{proof}
-
 \begin{mdframed}
-For both questions 2 and 4, we need the following lemma:
+For both question 2, we need the following lemma:
 \begin{lemma} \label{lem:distance_compact_closed}
@@
 
+\setcounter{question}{1}
 \begin{question}
@@
 
-	\emph{Hint: you may find it convenient to use the following form of the remainder}
+	\emph{Hint:} you may find it convenient to use the following form of the remainder
 	\[
@@
 	\]
-	\emph{and note that} $(N+1)! \sum_{\abs{\alpha} = N+1} \frac{h^\alpha}{\alpha!} = (h_1 + \dotsb + h_n)^{N+1}$. 
+	and note that $(N+1)! \sum_{\abs{\alpha} = N+1} \frac{h^\alpha}{\alpha!} = (h_1 + \dotsb + h_n)^{N+1}$. 
 \end{question}
@@
 
-\begin{question}
-	Which elements of $\DD(X)$ can be represented as a power series on $X$? 
-\end{question}
-
-\begin{solution}
-	It is known that if two power series agree on an open set, they agree on the entire space.
-Since every $\phi \in \DD(X)$ is identically zero on some open set (outside its support), the only element of $\DD(X)$ with a power series representation is the zero function.
-\end{solution}
-
-\begin{question}
-	Prove the $C^\infty$ Urysohn lemma: if $K$ is a compact subset of $X \subseteq \RR^n$, show that one can find a $\phi \in \DD(X)$ such that $0 \leq \phi \leq 1$ and $\phi = 1$ on a neighborhood of $K$.
-\end{question}
-
-\begin{solution}
-	Let $K \subseteq X$, then by \cref{lem:distance_compact_closed}, there is a nonzero distance $d$ from $K$ to $\RR^n \setminus X$, so we have neighbourhoods
-	\[
-	U_1 \ceq \qty{u \in X \mid d(u, K) < \frac d4}, \quad U_2 \ceq \qty{u \in X \mid d(u, K) < \frac{3d}{4}}
-	\]
-	which is an open neighbourhood of $K$ with $\clos U \subseteq X$.
-
-	 Let $\chi = \ind_{U_2}$. Now let $\psi \in \DD(\RR^n)$ satisfy $\int_{\RR^n} \psi \dd{x} = 1$ and $\supp\psi \subseteq B(0, \frac d4)$. The we compute
-	\begin{align*}
-		(\chi * \psi)(x) &= \int_{\RR^n} \chi(y) \psi(x - y) \dd{y} = \int_{U_2} \psi(x - y) \dd{y}.
-	\end{align*}
-	Clearly, $\chi * \psi \in \DD(X)$, and furthermore, we have for $x \in U_1$ that
-	\[
-	\int_{U_2} \psi(x - y) \dd{y} = \int_{U_2 - x} \psi(z) \dd{z} \overset\star= 1,
-	\]
-	since $\supp\psi \subseteq B(0, \frac d4) \subseteq U_2 - x$. This proves the claim.
-\end{solution}
-
-\begin{question}
-	Given $T \in \DD'(X)$, the derivative $\pt^\alpha T$ is defined by
-	\[
-	\ang{\pt^\alpha T, \phi} = (-1)^\abs{\alpha} \ang{T, \pt^\alpha\phi} \quad \forall \phi \in \DD(X). 
-	\]
-	Show that $\pt^\alpha T \in \DD'(X)$. If $\ord(T) = m$ what can you say about $\ord(\pt^\alpha T)$? 
-\end{question}
-
-\begin{proof}
-	Let $K \subseteq X$ be compact and $\phi \in \DD(X)$. Since $T$ is a distribution, we know that there exists constants $C, N$ such that
-	\[
-	\abs{\ang{T, \phi}} \leq C \sum_{\abs{\beta} \leq N} \sup \abs{\pt^\beta \phi}. 
-	\]
-
-	Letting $M \ceq \abs{\alpha}$, we find
-		\begin{align*}
-		\abs{\ang{\pt^\alpha T, \phi}} &= \abs{\ang{T, \pt^\alpha\phi}} \leq C \sum_{\abs{\beta} \leq N} \sup \abs{\pt^{\alpha + \beta}\phi} \leq C \sum_{\abs{\beta} \leq M + N} \sup \abs{\pt^\beta \phi}.
-	\end{align*}
-We conclude that $\pt^\alpha T$ is a distribution, and that if $\ord(T) = m$, $\ord(\pt^\alpha T) \leq m + \abs{\alpha}$. 
-\end{proof}
-
-\begin{question}
-	Given $T \in \DD'(X)$ and $f \in C^\infty(X)$, prove that for each multi-index $\alpha$
-	\[
-	\pt^\alpha(Tf) = \sum_{\beta \leq \alpha} \binom{\alpha}{\beta} \pt^\beta f \pt^{\alpha - \beta}T
-	\]
-	in $\DD'(X)$. 
-\end{question}
-
-\begin{proof}
-	Let $\phi \in \DD(X)$, then by definition we have $\ang{\pt^\alpha(Tf), \phi} = \ang{T, (-1)^\abs{\alpha} f \pt^\alpha\phi}$. Approximate $T$ by a sequence $(\psi_n) \subseteq \DD'(X)$, then we find
-	\begin{align*}
-	\ang{T, (-1)^{\abs{\alpha}} f \pt^\alpha\phi} &= \lim_{n\to\infty} \ang{\psi_n,  (-1)^{\abs{\alpha}} f \pt^\alpha\phi} = \lim_{n\to\infty}  (-1)^{\abs{\alpha}} \int_X \psi_n(x) f(x) \pt^\alpha\phi(x) \dd{x} \\
-	&= \lim_{n\to\infty} \int_X \pt^\alpha (\psi_n(x) f(x)) \phi(x) \dd{x} \\
-	&= \lim_{n\to\infty} \int_X \qty(\sum_{\beta \leq \alpha} \binom\alpha\beta\pt^\beta f(x) \cdot \pt^{\alpha - \beta} \psi_n(x)) \phi(x) \dd{x} \\
-	&= \lim_{n\to\infty} \sum_{\beta \leq \alpha} \binom\alpha\beta \ang{\pt^\beta f \pt^{\alpha - \beta} \psi_n, \phi} =  \lim_{n\to\infty} \sum_{\beta \leq \alpha} \binom\alpha\beta (-1)^{\abs{\alpha - \beta}} \ang{\psi_n, \pt^\beta f \pt^{\alpha - \beta} \phi} \\
-	&= \sum_{\beta \leq \alpha} \binom\alpha\beta (-1)^{\abs{\alpha - \beta}} \ang{T, \pt^\beta f \pt^{\alpha - \beta} \phi} = \sum_{\beta \leq \alpha} \binom\alpha\beta \ang{\pt^\beta f \pt^{\alpha - \beta} T, \phi} \\
-	&= \ang*{\sum_{\beta \leq \alpha} \binom{\alpha}{\beta} \pt^\beta f \pt^{\alpha - \beta}T, \phi}. 
-	%&= \lim_{n\to\infty} \ang*{\sum_{\abs{\beta} \leq \abs{\alpha}} \binom\alpha\beta \pt^\beta f\cdot \pt^{\alpha - \beta} \psi_n, \phi} = \lim_{n\to\infty} \sum_{\abs{\beta} \leq \abs{\alpha}
-	\end{align*}
-\end{proof}
-
-\begin{question}
-	Let $(x_k)$ be a sequence in $X$ with no limit point in $X$. Consider the family of linear maps
-	$u_\alpha \colon \DD(X) \to \CC$ defined by 
-	\[ \ang{u_\alpha, \phi} = \sum_{k=1}^\infty \pt^\alpha \phi(x_k)
-	 \]
-	 for each multi-index $\alpha$. For what $\alpha$ is $u_\alpha \in \DD'(X)$? What is $\ord(u_\alpha)$? 
-\end{question}
-
-\begin{solution}
-	Let $K \subseteq X$ be compact.
-Since $(x_k)$ does not have a limit point, only finitely many of the $x_k$ lie in $K$ (otherwise $(x_k)$ would have a subsequence contained in $K$ which would have a convergent subsequence).
-Without loss of generality, assume that $x_1, \dotsc, x_n \in K$, and $x_{n+1}, x_{n+2}, \dotsc, \notin K$.
-Now, for any $\phi \in \DD(X)$ with $\supp(\phi) \subseteq K$ we find
-	\begin{align*}
-		\abs{\ang{u, \phi}} &= \abs{\sum_{k=1}^\infty \pt^\alpha\phi(x_k)} = \abs{\sum_{k=1}^n \pt^\alpha\phi(x_k)} \leq \sum_{k=1}^n \abs{\pt^\alpha \phi(x_k)} \leq n \cdot \sup \abs{\pt^\alpha\phi} \leq n \cdot \sum_{\abs{\beta} \leq \abs{\alpha}} \sup \abs{\pt^\beta \phi}.
-	\end{align*}
-	This shows that $u_\alpha \in \DD'(X)$ for any $\alpha$, with $\ord(u_\alpha) \leq \abs{\alpha}$. We claim that this is an equality, i.e., $\ord(u_\alpha) = \abs{\alpha}$. \TODO How to show??
-\end{solution}
-
+\setcounter{question}{3}
 \begin{question}
@@
 	\begin{enumerate}[(a)]
-		\item If $u' = 1$ then we find 
+		\item It is clear that $x' = 1$ in the distributional sense, since for any test function $\phi$ we have
 		\[
-		\int_\RR \phi(x) \dd{x} = \ang{1, \phi} = \ang{u', \phi} = -\ang{u, \phi'},
+		\ang{x', \phi} = -\ang{x, \phi'} = - \int_\RR x \phi'(x) \dd{x} = \int_\RR \phi(x)\dd{x} = \ang{1, \phi}. 
 		\]
-		so for any $c \in \RR$ we find by partial integration $\ang{u, \phi'} = - \int_\RR\phi(x) \dd{x} = \int_\RR (x + c) \phi'(x) \dd{x}$. From this we deduce that $u = x + c$ for some $c$.
+		Therefore, the equation $u' = 1$ is equivalent to the equation $(u - x)' = 0$. We know that this implies that $u - x = c$ for some constant $c \in \CC$, so the most general solution is $u = x + c$.
 
@@
 		\]
-		Note that the above is satisfied for $u = - \delta_0 + c$ for any constant $c$. \TODO is this the most general solution?
+		Note that the above is satisfied for $u = - \delta_0 + c$ for any constant $c\in \CC$.
+
+		I do not know if this is the most general solution and/or how one would show this.
 
@@
 
-\begin{question}
-	Define the distribution $u \in \DD'(\RR^2)$ by the locally integrable function $u(x, y) = \ind_{x \geq y}$.
-Show that $\pt_x^2   u - \pt_y^2 u = 0$ in $\DD'(\RR^2)$.
-Can you give a physical interpretation of this result?
-\end{question}
-
-\begin{proof}
-	Let $f \in \DD(\RR^2)$, then we have
-	\begin{align*}
-		\ang{\pt_x^2 u - \pt_y^2 u, f} &= \ang{\pt_x^2 u, f}- \ang{\pt_y^2 u, f} = \ang{u, \pt_x^2 f} - \ang{u, \pt_y^2 f} = \ang{u, \pt_x^2 f - \pt_y^2 f} \\
-		&\overset\star= \int_{-\infty}^\infty \int_{y}^\infty \pt_x^2 f(x, y) \dd{x}\dd{y} - \int_{-\infty}^\infty \int_{-\infty}^x \pt_y^2 f(x, y) \dd{y} \dd{x} \\
-		&= - \int_{-\infty}^\infty \pt_x f(y, y) \dd{y} - \int_{-\infty}^\infty \pt_y f(x, x) \dd{x}
-		\\
-		&= - \int_{-\infty}^\infty (\pt_xf + \pt_yf)(x, x) \dd{x}.
-	\end{align*}
-Here, $\star$ follows from Fubini's theorem. 
-Define $g(x) = f(x, x)$, then it is easily seen that $g'(x) = \pt_x f(x, x) + \pt_y f(x, x)$, so we find that
-\[
-\ang{\pt_x^2 u - \pt_y^2 u, f} = 
-- \int_{-\infty}^\infty g'(x) \dd{x} = \lim_{x\to -\infty} g(x) - \lim_{x\to\infty} g(x) = 0 - 0 = 0. 
-\]
-This shows that $\pt_x u - \pt_y u = 0$, or equivalently, that $u$ satisfies the wave equation. 
-\end{proof}
-
-\begin{question}
-	Compute $\Delta(\norm{x}^{2-n})$ in $\DD'(\RR^n)$ for $n \geq 3$, i.e.\ compute
-	\[
-	\ang{\Delta (\norm{x}^{2-n}), \phi} = \ang{\norm{x}^{2-n}, \Delta\phi} = \int \frac{\Delta\phi}{\norm{x}^{n-2}} \dd{x}
-	\]
-	for arbitrary $\phi \in \DD(\RR^n)$. Note that $\norm{x} = (x_1^2 +\dotsb +x_n^2)^{1/2}$ and $\Delta = \sum_i (\pdv{x_i})^2$.
-
-	\emph{Hint:} use $\int\dd{x} = \int_{\norm{x}\leq\eps}\dd{x} + \int_{\norm{x}>\eps}\dd{x}$ and treat each integral separately. 
-\end{question}
-
-\begin{solution}
-	We follow the hint: let $\eps > 0$, then we first compute
-	\[
-	\int_{\norm{x} > \eps} \frac{\Delta\phi}{\norm{x}^{n-2}} \dd{x} = \sum_{i=1}^n \int_{\norm{x} > \eps} \frac{\pdv[2]{x_i} \phi}{\norm{x}^{n-2}}\dd{x} = \sum_{i=1}^n \int_{\norm{x} > \eps} \phi \cdot \pdv[2]{x_i} \norm{x}^{2-n} \dd{x}.
-	\]
-	Now, it is easily computed that $\pdv{x_i} \norm{x} = \frac{x_i}{\norm{x}}$, and therefore 
-	\begin{align*}
-	\pdv[2]{x_i} \norm{x}^{2-n} &= \pdv{x_i} (2-n) x_i \norm{x}^{-n} = (2-n) \qty(\norm{x}^{-n} - n x_i^2 \norm{x}^{-n-2}) \\
-	&= (n-2)\norm{x}^{-n} \qty(n \qty(\frac{x_i}{\norm{x}})^2 - 1). 
-	\end{align*}
-\TODO finish
-\end{solution}
-
-\begin{question}
-	Let $(f_k)$ be the sequence of smooth functions defined by $f_k(x) = \frac1\pi \frac k{(kx)^2 + 1}$. Prove that $f_k \to \delta_0$ in $\DD'(\RR)$. Compute the limits
-	\begin{enumerate}[(a)]
-		\item $\lim_{k\to\infty} k^2 x e^{-k^2 x^2}$,
-		\item $\lim_{k\to\infty} k^3 e^{ikx}$,
-		\item $\lim_{k\to\infty} \frac{\sin(kx)}{\pi x}$,
-	\end{enumerate}
-in $\DD'(\RR)$. 
-\end{question}
-
-\begin{solution}
-	It is easily seen that every $f_k$ is locally integrable. Furthermore, noting that $\arctan(kx)' = \frac{k}{(kx)^2 + 1}$, we see by the dominated convergence theorem that
-	\begin{align*}
-		\lim_{k\to\infty}\ang{f_k, \phi} &= \frac1\pi \lim_{k\to\infty} \int_\RR \frac{k}{(kx)^2 + 1} \phi(x) \dd{x} = -\frac1\pi \lim_{k\to\infty} \int_\RR \arctan(kx) \phi'(x) \dd{x} \\
-		&= -\frac1\pi \int_\RR \lim_{k\to\infty} \arctan(kx) \phi'(x) \dd{x} = -\frac1\pi \qty(2 \frac\pi2 \int_0^\infty \phi'(x) \dd{x} ) = \phi(0), 
-	\end{align*}
-which proves $f_k \to \delta_0$.
-
-\begin{enumerate}[(a)]
-	\item We have $- (\frac12 e^{-k^2x^2})' = k^2 x e^{-k^2 x^2}$ and therefore
-	\begin{align*}
-		\lim_{k\to\infty} \int_\RR k^2 x e^{-k^2x^2} \phi(x) \dd{x} &= \frac12 \lim_{k\to\infty}  \int_\RR  e^{-k^2 x^2} \phi'(x) \dd{x} = \frac12 \int_\RR \lim_{k\to\infty} e^{-k^2 x^2} \phi'(x) \dd{x} = 0,
-	\end{align*}
-so the sequence converges to 0 in $\DD'(\RR)$.
-
-\item We have 
-\begin{align*}
-	\lim_{k\to\infty}  k^3 \int_\RR e^{ikx} \phi(x) \dd{x} = \lim_{k\to\infty} 2\pi k^3 \Cal F^{-1}[\phi](k). 
-\end{align*}
-Since $\phi$ is a Schwarz function, its inverse Fourier transform is also a Schwarz function, and in particular $k^3 \Cal F^{-1}[\phi](k) \to 0$ as $k \to \infty$.
-
-\item This sequence converges to $\delta_0$, although I have no idea how to prove it.
-
-\end{enumerate}
-\end{solution}
-
-\begin{question}
-	Compute the limit
-	\[
-	\lim_{k\to\infty} \frac12 + \sum_{m=1}^k \cos(\pi m x)
-	\]
-	in $\DD'(-1, 1)$. 
-\end{question}
-
-\begin{solution}
-	Note that \[
-	f_k(x) \ceq 
-	\frac12 + \sum_{m=1}^k \cos(\pi m x) = \frac12\qty(1 + e^{i \pi x} + e^{-i\pi x} + \dotsb + e^{i \pi m x} + e^{-i\pi m x}) = \frac12 \sum_{m=-k}^k (e^{i\pi x})^m.
-	\]
-	By viewing the last term as a geometric series $\frac12 e^{-ik\pi x} \sum_{m=0}^{2k+1} (e^{ix})^m$ we can compute that 
-	\[
-	f_k(x) = \frac{\sin((n + \frac12) \pi x)}{2\sin(\frac12\pi x)}. 
-	\]
-	It can be shown (???) that $f_k \to \delta_0$ in $\DD'(\RR)$. 
-\end{solution}
-
-\begin{question}
-	We define the \emph{principal value} of $1/x$, written $\prv(1/x)$, by
-	\[
-	\ang*{\prv(1/x), \phi} = \lim_{\eps \to 0} \int_{\abs{x} > \eps} \frac{\phi(x)}{x} \dd{x} \quad\text{for all $\phi \in \DD(\RR)$}. 
-	\]
-	Prove that $\prv(1/x) \in \DD'(\RR)$ and that $\ord(\prv(1/x)) = 1$. Show that
-	\[
-	\lim_{\eps \to 0} \frac{1}{x - i\eps} = \prv\qty(\frac1x) + i\pi \delta_0 \quad\text{in $\DD'(\RR)$}. 
-	\]
-\end{question}
-
-\begin{proof}
-	First we must show that $\prv(1/x)$ is well-defined.  ??
-\end{proof}
-
-\begin{question}
-	??
-\end{question}
-\begin{proof}
-	??
-\end{proof}
-
-\begin{question}
-	??
-\end{question}
-
-\begin{proof}
-	??
-\end{proof}
-
-\begin{question}
-	Define the distribution $u \in \EE'(\RR^3)$ by the locally integrable function $u(x) = \ind_{\abs{x}\leq 1}$.
-Prove that $-\sum_i x_i (\pdv{u}{x_i}) = \dd{\sigma_2}$ in $\EE'(\RR^3)$, where $\dd{\sigma_2}$ is the surface element on the sphere $S^2 \subseteq \RR^3$.
-\end{question}
-
-\begin{proof}
-	We have
-	\begin{align*}
-		\ang*{- \sum_i x_i \pdv{u}{x_i}, \phi} &= - \sum_i \ang*{\pdv{u}{x_i}, x_i\phi} = \sum_i \ang*{u, \pdv{x_i}(x_i \phi)} = \ang*{u, \sum_i \pdv{x_i}(x_i\phi)} \\
-		&= \int_{\norm{x} \leq 1} \sum_i \pdv{x_i}(x_i\phi) \dd{x}
-	\end{align*}
-\end{proof}
 \end{document}

### 2020-11-04 topics_statistical_theory/topics_ES1.tex
progress is being made

@@
 \begin{proof}
-	We have 
+	Define $x^* \ceq \frac{2^{1/2}\sigma}{n^{1/2}} \log^{1/2}(\frac1\delta) + \frac cn \log(\frac 1\delta)$. Then we have
 	\[
-	\PP(\bar X \geq x) \leq \exp(- \frac{nx^2}{2(\sigma^2 + cx)}) \eqqcolon \delta. 
+	(x^*)^2 = \frac{2\sigma^2}{n}\log(\frac1\delta) + \frac{2^{3/2}\sigma c}{n^{3/2}} \log^{3/2}\qty(\frac1\delta) + \frac{c^2}{n^2} \log^2\qty(\frac1\delta),
 	\]
+	and therefore
+	\begin{align*}
+	-\frac{n(x^*)^2}{2(\sigma^2 + cx)} &= - \frac{2\sigma^2 \log(\frac1\delta) + 2^{3/2}\sigma c \log^{3/2}(\frac1\delta)/n^{1/2} + c^2 \log^2(\frac1\delta)/n}{2\sigma^2 + 2^{3/2} \sigma c \log^{1/2}(\frac1\delta)/n^{1/2} + 2c^2\log(\frac1\delta)/n} \\
+	&= -\log(\frac1\delta) \frac{2\sigma^2 + 2^{3/2} \sigma c/n^{1/2} + c^2\log(1/\delta)/n}{2\sigma^2 + 2^{3/2} \sigma c/n^{1/2} + 2c^2\log(1/\delta)/n} \\
+	&\geq -\log(\frac1\delta) = \log(\delta),
+	\end{align*}
+
+	so by Bernstein's inequality we have
+	\[
+	\PP(\bar X \geq x^*) \leq \exp(\log(\delta)) = \delta, 
+	\]
+	which is what we wanted to prove.
+
+%	We have 
+%	\[
+%	\PP(\bar X \geq x) \leq \exp(- \frac{nx^2}{2(\sigma^2 + cx)}) \eqqcolon \delta. 
+%	\]
+
 	Now we just need express $x$ in terms of $\delta$: taking logarithms on both sides we obtain
@@
 	&= \frac cn \log(1/\delta) + \sqrt{\frac{c^2}{n^2}\log^2(1/\delta) + \frac{2\sigma^2}{n} \log(1/\delta)} \\
-	&= ???? 
+	&\geq \frac cn \log(1/\delta) + \sqrt{\frac{2\sigma^2}{n} \log(1/\delta)}. 
 	\end{align*}
+
+So we have ?????
 \end{proof}
@@
 \begin{align*}
-	\int_{-\infty}^\infty u^j K_\ell(u) \dd{u} = \sum_{r=0}^{\ell - 1} \frac{(-1)^r H_r(0)}{r!} \int_{-\infty}^\infty u^j \phi^{(r)}(u) \dd{u}.
+	\int_{-\infty}^\infty u^j K_\ell(u) \dd{u} = \sum_{r=0}^{\ell - 1} \frac{(-1)^r H_r(0)}{r!} \int_{-\infty}^\infty u^j H_r(u) \phi(u) \dd{u}.
 \end{align*}
-Clearly, if $j < r$, integral in the sum vanishes. If $j = r$, the integral yields $r!$. So we have
+Write $u^j = \sum_{k=0}^j c_k H_k(u)$, then the integral will vanish unless $k = r$, so we get
 \[
-\int_{-\infty}^\infty u^j K_\ell(u) \dd{u} = (-1)^j H_j(0) + \sum_{r=0}^{j-1} \int_{-\infty}^\infty u^j \phi^{(r)}(u) \dd{u}. 
+\int_{-\infty}^\infty u^j K_\ell(u) = \sum_{r=0}^{j} (-1)^r c_r H_r(0) = \sum_{r=0}^j c_r H_r(0) = 0^j = 0,
 \]
-
-??
+since $H_r(0) = 0$ for $r$ odd. 
+%Clearly, if $j < r$, integral in the sum vanishes. If $j = r$, the integral yields $r!$. So we have
+%\[
+%\int_{-\infty}^\infty u^j K_\ell(u) \dd{u} = (-1)^j H_j(0) + \sum_{r=0}^{j-1} \int_{-\infty}^\infty u^j \phi^{(r)}(u) \dd{u}. 
+%\]
+%
+%??
 \end{proof}
@@
 \begin{question}
-	\begin{enumerate}
+	\begin{enumerate}[(a)]
 		\item 	Verify the algebraic identity
 		\[
-		\phi_\sigma(x-\mu)\phi_{\sigma'}(x-\mu') = \phi_{\sigma\sigma'/(\sigma^2 + \sigma'^2)^{1/2}}(x - \mu^*) \phi_{(\sigma + \sigma')^{1/2}}(\mu - \mu')
+		\phi_\sigma(x-\mu)\phi_{\sigma'}(x-\mu') = \phi_{\sigma\sigma'/(\sigma^2 + \sigma'^2)^{1/2}}(x - \mu^*) \phi_{(\sigma^2 + \sigma'^2)^{1/2}}(\mu - \mu')
 		\]
@@
 \begin{proof}
-	\begin{enumerate}
-		\item \TODO
-		\item no.
-%		We have 
-%		\begin{gather*}
-%			\phi_{\sigma\sigma'/(\sigma^2 + \sigma'^2)^{1/2}}(x - \mu^*) \phi_{(\sigma + \sigma')^{1/2}}(\mu - \mu') \\
-%			= \frac{1}{2\pi \sigma \sigma'} \exp(-\frac12\qty(\frac{(x - \mu^*)^2(\sigma^2 + \sigma'^2)}{\sigma^2 \sigma'^2} + \frac{(\mu - \mu^*)^2}{\sigma^2 + \sigma'^2})) \\
-%			= \frac{1}{2\pi \sigma \sigma'} \exp\qty(-\frac12\qty(\frac{(x - \mu^*)^2(\sigma^2 + \sigma'^2)^2 + \sigma^2 \sigma'^2 (\mu - \mu^*)^2}{\sigma^2 \sigma'^2(\sigma^2 + \sigma'^2)})) \\
-%			=  \frac{1}{2\pi \sigma \sigma'}\exp\qty(-\frac12\qty(\frac{
-%			x^2(\sigma^2 + \sigma'^2)^2 + (\sigma'^2\mu + \sigma^2\mu')^2 - 2x (\sigma'^2\mu + \sigma^2\mu') (\sigma^2 + \sigma'^2)	
-%		}{\sigma^2 \sigma'^2(\sigma^2 + \sigma'^2)}))
-%		\end{gather*}
+	\begin{enumerate}[(a)]
+		\item We have
+		\begin{align*}
+			&\phantom{=\ }\frac{(x-\mu)^2}{\sigma^2} + \frac{(x - \mu')^2}{\sigma'^2} \\
+			&= \frac{\sigma'^2(x - \mu)^2 + \sigma^2(x - \mu')^2}{\sigma^2\sigma'^2} \\
+			&= \frac{(\sigma^2 + \sigma'^2)x^2 - 2 (\sigma'^2 \mu + \sigma^2 \mu')x + \sigma'^2 \mu^2 + \sigma^2\mu'^2}{\sigma^2 \sigma'^2} \\
+			&=  \frac{(\sigma^2 + \sigma'^2)(x^2 - 2\mu^*x) + \sigma'^2 \mu^2 + \sigma^2\mu'^2}{\sigma^2 \sigma'^2} \\
+			&= \frac{(\sigma^2 + \sigma'^2)(x^2 - 2\mu^* + \mu^{*2}) - (\sigma^2 + \sigma'^2)\mu^{*2} + \sigma'^2 \mu^2 + \sigma^2\mu'^2}{\sigma^2 \sigma'^2} \\
+			&= \frac{(x - \mu^*)^2}{(\sigma\sigma'/(\sigma^2 + \sigma'^2)^{1/2})^2} + \frac{\sigma'^2 \mu + \sigma^2 \mu'^2 - (\sigma'^2\mu + \sigma^2\mu')^2/(\sigma^2 + \sigma'^2)}{\sigma^2\sigma'^2} \\
+			&= \frac{(x - \mu^*)^2}{(\sigma\sigma'/(\sigma^2 + \sigma'^2)^{1/2})^2} + \frac{(\sigma^2 + \sigma'^2)(\sigma'^2\mu + \sigma^2\mu'^2) - (\sigma'^2\mu + \sigma^2\mu')^2}{\sigma^2 \sigma'^2(\sigma^2 + \sigma'^2)} \\
+			&= \frac{(x - \mu^*)^2}{(\sigma\sigma'/(\sigma^2 + \sigma'^2)^{1/2})^2} + \frac{(\mu - \mu')^2}{\sigma^2 + \sigma'^2} \\
+			&= \frac{(x - \mu^*)^2}{(\sigma\sigma'/(\sigma^2 + \sigma'^2)^{1/2})^2} + \frac{(\mu - \mu')^2}{((\sigma^2 + \sigma'^2)^{1/2})^2}, 
+		\end{align*}
+	which proves the claim. 
+	\item Let $K = \phi_1$ and define $K_h(x) \ceq h^{-1} K(x/h)$ so $K_h = \phi_h$. Then recall from the lectures that
+	\[
+	\MISE(\hat f_n) = \frac1n \int_\RR \qty[(\phi_h^2 * \phi_\sigma)(x) - (\phi_h * \phi_\sigma)^2(x)] \dd{x} + \int_{-\infty}^\infty \qty[ (\phi_h * \phi_\sigma)(x) - \phi_\sigma]^2 \dd{x}
+	\]
 	\end{enumerate}

### 2020-11-08 distribution_theory/C3.tex
lots of work done, lots of confusion going on

\section{Tempered distributions and Fourier analysis}
\subsection{Functions of rapid decay}
\begin{definition}
	For any $f \colon \RR^n \to \CC$ and multi-indices $\alpha, \beta$ we define $\norm{f}_{\alpha, \beta} \ceq \sup_{x \in \RR^n} \abs{x^\alpha \pt^\beta\phi}$.

	We define the \emph{Schwartz space}
	\[
	\Cal S(\RR^n) \ceq \qty{f \in C^\infty(\RR^n) : \norm{f}_{\alpha, \beta} < \infty \text{ for all $\alpha, \beta$}}.
	\]

	We say that a sequence $(\phi_n) \subseteq \SS(\RR^n)$ converges to 0 if $\norm{\phi_m}_{\alpha, \beta} \to 0$ for every $\alpha, \beta$.  
\end{definition}

\begin{example}
	The function $x \mapsto \exp(-\norm{x}^2)$ lies in $\SS(\RR^n)$. 
\end{example}

\begin{proposition}
	For all $n$ we have that $\SS(\RR^n) \subseteq L^1(\RR^n)$. 
\end{proposition}

\begin{proof}
	Let $\phi \in \SS(\RR^n)$, then for all $N \in \NN$ we have
	\[
	\int_{\RR^n} \abs{\phi(x)} \dd{x} = \int_{\RR^n} (1 + \norm{x})^{-N} (1 + \norm{x})^N \abs{\phi(x)} \dd{x} \overset?\leq C\sum_{\abs{\alpha} \leq N} \norm{\phi}_{\alpha, 0} \int_{\RR^n} (1 + \norm{x})^{-N} \dd{x}.
	\]

	Since $\int_{\RR^n} (1 + \norm{x})^{-N} \dd{x}$ is finite for $N$ large enough (??), this proves the claim. 
\end{proof}

\begin{definition}
	A linear map $u \colon \SS(\RR^n) \to \CC$ is called a \emph{tempered distribtion} if there exists constants $C, N$ such that
	\[
	\abs{\ang{u, \phi}} \leq C \sum_{\abs{\alpha}, \abs{\beta} \leq N} \norm{\phi}_{\alpha, \beta} \quad\text{for all $\phi \in \SS(\RR^n)$}. 
	\]
	The space of tempered distributions is denoted by $\SS'(\RR^n)$. 
\end{definition}

This definition is equivalent to sequential continuity.

\subsection{The Fourier transform on Schwartz functions}

\begin{convention}
	We write $D \ceq -i\pt$ and $D^\alpha = (-i)^{\abs{\alpha}} \pt^\alpha$. 
\end{convention}

\begin{definition}
	For $f \in L^1(\RR^n)$, define the \emph{Fourier transform} of $f$ by  
	\[
	[\FF(f)](\lambda) = \hat f(\lambda) \ceq \int_{\RR^n} e^{-i\lambda \vdot x} f(x) \dd{x} \quad\text{where $\lambda \in \RR^n$}. 
	\]
\end{definition}

\begin{lemma}
	If $f \in L^1(\RR^n)$, then $\hat f$ is continuous. 
\end{lemma}

\begin{proof}
	If $\lambda_m \to \lambda \in \RR^n$, then by the dominated convergence theorem we have
	\[
	\hat f(\lambda_m) = \int_{\RR^n} e^{-i\lambda_m \vdot x} f(x) \dd{x} \overset{\text{DCT}}= \int_{\RR^n} e^{-i\lambda \vdot x} f(x) \dd{x} = \hat f(\lambda),
	\]
	where we were able to use the dominated convergence theorem since the integrand is absolutely bounded by $\abs{f}$ and $f \in L^1$. 
\end{proof}

It turns out that this idea generalises: the faster the function $f$ decays, the smoother the Fourier transform $\hat f$ is.

\begin{lemma} \label{lem:differentiate_fourier}
	For $\phi \in \SS(\RR^n)$, we have $\FF[D_x^\alpha \phi](\lambda) = \lambda^\alpha \hat\phi(\lambda)$ and $\FF[x^\beta\phi](\lambda) = (-1)^{\abs{\beta}} D_\lambda^\beta \hat\phi(\lambda)$. 
\end{lemma}

\begin{proof}
	Since $\abs{x^\alpha D^\beta\phi} \to 0$ as $\norm{x} \to \infty$, we have using integration by parts
	\begin{align*}
		\FF[D_\lambda^\alpha\phi](\lambda) &= \int_{\RR^n} e^{-i\lambda \vdot x} D_x^\alpha \phi(x) \dd{x} \\
		&= (-1)^{\abs{\alpha}} \int_{\RR^n} D_x^\alpha(e^{-i\lambda \vdot x}) \phi(x) \dd{x} \\
		&= \lambda^\alpha\hat\phi(\lambda). 
	\end{align*}

	For the second part of the lemma, by differentiation under the integral sign we get
	\begin{align*}
		 \FF[x^\beta \phi](\lambda) &=  \int_{\RR^n} e^{-i\lambda \vdot x} x^\beta \phi(x) \dd{x} \\
		 &= \int_{\RR^n} ((-D_\lambda)^\beta e^{-i\lambda \vdot x}) \phi(x) \dd{x} \\
		 &= (-1)^{\abs{\beta}} D_\lambda^\beta \hat\phi(\lambda).
	\end{align*}
\end{proof}

We define the \emph{inverse Fourier transform} by 
\[
\FF^{-1}[\hat f](x)  \ceq \frac{1}{(2\pi)^n} \int_{\RR^n} e^{i\lambda \vdot x} \hat f(\lambda) \dd{\lambda}. 
\]
We will now show that on $\SS(\RR^n)$, the inverse Fourier transform is indeed an inverse:
\begin{theorem} \label{thm:fourier_transform_isomorphism}
	The Fourier transform $\FF$ defines a continuous isomorphism from $\SS(\RR^n)$ to itself. 
\end{theorem}

\begin{proof}
	Let $\phi \in \SS(\RR^n)$. First, we show that $\hat\phi \in \SS(\RR^n)$: by the previous lemma we have for multi-indices $\alpha, \beta$ that
	\begin{align}
		\abs{\lambda^\alpha (-D_\lambda)^\beta \hat\phi(\lambda)} &= \abs{\lambda^\alpha \FF[x^\beta\phi](\lambda)} = \abs{\FF[D^\alpha_x (x^\beta\phi)](\lambda)} = \abs{\int_{\RR^n} e^{-i\lambda \vdot x} D^\alpha(x^\beta\phi) \dd{x}} \nonumber \\
		&\leq \int_{\RR^n} \abs{D^\alpha (x^\beta\phi)} \dd{x} \label{eq:fourier_Schwartz_estimate},
	\end{align}
which is finite since $D^\alpha (x^\beta\phi)$ is also a Schwartz function and therefore integrable.

From the previous lemma we also infer that $\hat\phi$ is smooth, so indeed we have $\hat\phi \in \SS(\RR^n)$. 
From \cref{eq:fourier_Schwartz_estimate} it is also easily seen that if $\phi_m \to 0$ in $\SS(\RR^n)$, then $\hat\phi_m \to 0$ in $\SS(\RR^n)$ also, which shows that $\FF$ is continuous.

To prove surjectivity and injectivity, we will show that $\FF^{-1}[\FF[f]](x) = f(x)$ (???). Indeed we have
\begin{align*}
	\FF^{-1} [\FF[\phi]](x) &= \frac{1}{(2\pi)^n} \int_{\RR^n} \int_{\RR^n} e^{i\lambda \vdot (x-y)} \phi(y) \dd{y}\dd{\lambda} \\
	&= \frac{1}{(2\pi)^n} \lim_{\eps \downarrow 0} \int_{\RR^n} \int_{\RR^n} e^{i \lambda \vdot ( x- y) - \eps\norm{\lambda}^2} \phi(y) \dd{y}\dd{\lambda} \\
	&\overset\star= \frac{1}{(2\pi)^n}\lim_{\eps \downarrow 0} \int_{\RR^n} \phi(y)\int_{\RR^n} e^{i\lambda \vdot (x- y) - \eps\norm{\lambda}^2} \dd{\lambda} \dd{y},
	%&= \frac{1}{(2\pi)^n} \int_{\RR^n}\int_{\RR^n} e^{i\lambda(x- y)} \dd{\lambda} \dd{y}
\end{align*}
where $\star$ follows from Fubini's theorem.

Now we note that 
\[
\int_{\RR^n} e^{i\lambda\vdot (x-y) - \eps\norm{\lambda}^2} \dd{\lambda} = \prod_{i=1}^n \int_{\RR} e^{i\lambda_j(x_j - y_j) - \eps\lambda_j^2} \dd{\lambda} \overset{\star\star}= \prod_{i=1}^n \qty(\frac\pi e)^{1/2} e^{- \frac{(x_i - y_i)^2}{4\eps}} = \qty(\frac\pi\eps)^{n/2} e^{- \frac{\norm{x - y}^2}{4\eps}}.
\] 
To explain $\star\star$, \TODO.

and plugging that into the above yields 
\begin{align*}
\FF^{-1}[\FF[\phi]](x) &= \lim_{\eps\downarrow 0} 2^{-n}(\pi \eps)^{-n/2} \int_{\RR^n} \phi(y) e^{-\norm{x-y}^2/(4\eps)} \dd{y} \\
&\overset{\star\star\star}= \pi^{-n/2} \lim_{\eps\downarrow0}\int_{\RR^n} \phi(x - 2\sqrt\eps y') e^{-\norm{y'}^2} \dd{y} \\
&\overset{\text{DCT}}=\pi^{-n/2} \phi(x) \int_{\RR^n} e^{-\norm{y'}^2} \dd{y} = \phi(x),
\end{align*}
where $\star\star\star$ follows from the substitution $x - y = 2\sqrt\eps y'$.

Finally, continuity of $\FF^{-1}$ is easily shown with an argument analogous to that for continuity of $\FF$ (????). 
\end{proof}

\subsection{The Fourier transform on tempered distributions}
\begin{lemma}
	For $\phi, \psi \in \SS(\RR^n)$, we have $\int_{\RR^n} \phi(x) \hat\psi(x) \dd{x} = \int_{\RR^n} \hat\phi(x) \psi(x) \dd{x}$. 
\end{lemma}

\begin{proof}
	This follows from Fubini's theorem:
	\[
	\int_{\RR^n}\phi(x) \hat \psi(x) \dd{x} = \int_{\RR^n}\int_{\RR^n} \phi(x) \psi(\lambda) e^{-i\lambda \vdot x} \dd{\lambda} \dd{x} = \int_{\RR^n}\int_{\RR^n} \psi(\lambda)  \phi(x) e^{-i\lambda \vd x} \dd{x} \dd{\lambda} = \psi(\lambda) \psi(\lambda) \hat\phi(\lambda) \dd{\lambda}. 
	\]
\end{proof}

The above result can be rewritten as $\ang{\hat\phi, \psi} = \ang{\phi, \hat\psi}$, which motivates the definition of the Fourier transform for tempered distributions:
\begin{definition}
	For $u \in \SS'(\RR^n)$, we define its Fourier transform by
	\[
	\ang{\hat u, \phi} \ceq \ang{u, \hat\phi} \quad\text{for all $\phi \in \SS(\RR^n)$}. 
	\]
\end{definition}
Using sequential continuity and \cref{thm:fourier_transform_isomorphism}, it is easily seen that $\hat u$ is indeed an element of $\SS'(\RR^n)$.

\begin{example}
	It is easily checked that $\delta_0 \in \SS'(\RR^n)$, and we can compute
	\[
	\ang{\hat{\delta_0}, \phi} = \ang{\delta_0, \hat\phi} = \hat\phi(0) = \int_{\RR^n} \phi(x) \dd{x} = \ang{1, \phi},
	\]
	so we can write $\hat{\delta_0} = 1$. Analogously, by the Fourier inversion theorem we have
	\[
	\ang{\hat 1, \phi} = \ang{1, \hat\phi} = \int_{\RR^n} \hat\phi(\lambda) \dd{\lambda} = (2\pi)^n \phi(0) = \ang{(2\pi)^n \delta_0, \phi}, 
	\]
	so we write $\hat 1 = (2\pi)^n \delta_0$. 
\end{example}

We can easily generalise \cref{lem:differentiate_fourier} to the Fourier transform on $\SS'(\RR^n)$, so
\[
\FF[D^\alpha u] = \lambda^\alpha \hat u, \quad \FF[x^\beta u] = (-D^\beta) \hat u. 
\]

\begin{theorem}
	The Fourier transform $\FF$ extends to a continuous isomorphism on $\SS'(\RR^n)$. 
\end{theorem}

\begin{proof}
	We claim that $\refl u = (2\pi)^{-n} \FF[\FF[u]]$. To check this, note that by the Fourier inversion theorem we have for $\phi \in \SS(\RR^n)$ that
	\[
	\refl\phi(x) = \phi(-x) = (2\pi)^{-n} \int_{\RR^n} e^{-i\lambda\vd x} \hat\phi(\lambda) \dd{\lambda} = (2\pi)^{-n} \FF[\FF[\phi]](x),
	\]
	and therefore 
	\[
	\ang{\refl u, \phi} = \ang{u, \refl\phi} = \ang{u, (2\pi)^{-n} \FF[\FF[\phi]]} = \ang{(2\pi)^{-n} \FF[\FF[u]], \phi}. 
	\]

	This shows that $\FF$ is bijective (since $\FF\circ \FF$ is bijective). For continuity of $\FF$ and its inverse: using \cref{thm:fourier_transform_isomorphism}, we find
	\begin{align*}
	&\phantom{\iff} u_m \to 0 \text{ in $\SS'(\RR^n)$} \\
	&\iff \ang{u_m, \phi} \to 0 \ \forall \phi \in \SS(\RR^n) \\
	&\iff \ang{u_m, \hat\phi} \to 0 \ \forall \phi \in \SS(\RR^n) \\
	&\iff \ang{\hat u_m, \phi} \to 0 \ \forall \phi \in \SS(\RR^n) \\
	&\iff \hat u_m \to 0 \text{ in $\SS'(\RR^n)$}.
	\end{align*}
\end{proof}

\subsection{Sobolev spaces}
\begin{convention}
	We write $\ang{\lambda} \ceq (1 + \norm{\lambda}^2)^{1/2}$ for $\lambda \in \RR^n$. Note that $\ang{\lambda} \sim 1$ as $\norm{\lambda} \to 0$ and $\ang{\lambda} \to \norm{\lambda}$ as $\norm{\lambda} \to \infty$. 
\end{convention}

\begin{definition}
	For $s \in \RR$, define the \emph{Sobolev space} $H^s(\RR^n)$ to be the set of tempered distributions $u \in \SS'(\RR^n)$ for which $\hat u$ can be identified with a measurable function $\hat u \colon \RR^n \to \CC$ such that $\ang{\lambda}^s \hat u \in L^2(\RR^n)$.

	For $X \subseteq \RR^n$ open, we define the \emph{localised Sobolev space} $H^s\loc(X)$ by setting
	\[
	u \in H^s\loc(X) \iff \phi u \in H^s(\RR^n) \text{ for all $\phi \in \DD(X)$}. 
	\]
\end{definition}

% Suppose we have an open subset $X \subseteq \RR^n$, then we can define the \emph{localised} Sobolev space $H^s_\Rm{loc}(X)$ by sayin

\begin{lemma}[Sobolev lemma]
	If $u \in H^s(\RR^n)$ with $s > \frac n2$, then  $u$ is continuous. 
\end{lemma}

\begin{proof}
	We will show that $\hat u$ is integrable. By Cauchy-Schwarz, we have
	\begin{align*}
		\int_{\RR^n} \abs{\hat u(\lambda)} \dd{\lambda} &= \int_{\RR^n}  \ang{\lambda}^{-s} \ang{\lambda}^s \abs{\hat u(\lambda)} \dd{\lambda} \\
		&\leq \qty(\int_{\RR^n}  \ang{\lambda}^{-2s}\dd{\lambda})^{1/2} \norm{u}_{H^s} \\
		&= C \norm{u}_{H^s} \qty(\int_0^\infty r^{n-1} (1 + r^2)^{-s} \dd{r})^{1/2}, 
	\end{align*}
where the last line follows from using polar coordinates and $C$ is the area of the $(n-1)$-sphere.

Writing $s = \frac n2 + \eps$, we find that the integrand $r^{n-1}(1 + r^2)^{-s}$ is of order $O(r^{-1 - 2\eps})$ as $r \to\infty$, and therefore the integral is finite, so indeed we have $\hat u \in L^1(\RR^n)$.

By applying \cref{thm:fourier_transform_isomorphism} to a test function, we can show that $u = (2\pi)^{-n} \int_{\RR^n} e^{i\lambda \vd x} \hat u(\lambda) \dd{\lambda}$, which is continuous by the dominated convergence theorem. 
\end{proof}

\begin{corollary}
	If $u \in H^s(\RR^n)$ for every $s > n/2$, then $u \in C^\infty(\RR^n)$. 
\end{corollary}

%\begin{proof}
%	We will show for any multi-index $\alpha$ of order $N$ that $\lambda^\alpha \hat u$ is integrable: this shows (??) that $D^\alpha u$ exists. We have for every $s > n/2$ that
%	\begin{align*}
%		\int_{\RR^n} \abs{\lambda^\alpha \hat u(\lambda)} \dd{\lambda} \les \int_{\RR^n} \ang{\lambda}^N  \abs{\hat u(\lambda)}\dd{\lambda} = \int_{\RR^n} \ang{\lambda}^{N - s} \ang{\lambda}^s \abs{\hat u(\lambda)}\dd\lambda \leq \qty(\int_{\RR^n} \ang{\lambda}^{2(N -s)})^{1/2} \norm{u}_{H^s}.
%	\end{align*}
%Analogously as in the proof of the previous theorem, we get an integrand $r^{n-1}(1 + r^2)^{N-s} = O(n-1-2s + 2N)$
%\end{proof}

### 2020-11-08 distribution_theory/C4.tex
lots of work done, lots of confusion going on

\section{Applications of Fourier transform}
\subsection{Elliptic regularity}
Recall that $D = -i\pt$. If $p$ is an $N$-th order polynomial, then $p(D)$ is called an $N$-th order differential operator.

\begin{definition}
	For an $N$-th order differential operator $p(D) = \sum_{\abs{\alpha} \leq N} c_\alpha D^\alpha$, define its \emph{principal symbol} $\sigma_p(\lambda)$ by \[
	\sigma_p(\lambda) \ceq \sum_{\abs{\alpha} = N} c_\alpha \lambda^\alpha \qquad(\lambda \in \RR^n). 
	\]
	The operator $p(D)$ is called \emph{elliptic} if $\sigma_p(\lambda) \neq 0$ for $\lambda \neq 0$. 
\end{definition}

\begin{lemma}
	If $p(D)$ is an $N$-th order elliptic partial differential operator, then there exist $R > 0$ such that, $C > 0$ such that 
	\[
	\abs{p(\lambda)} \geq C \ang{\lambda}^N \qquad\text{if } \norm{\lambda} > R. 
	\]
\end{lemma}

\begin{proof}
Let $C_0 > 0$ be the minimum of $\abs{\sigma_p}$ on $S^{n-1}$, then for $\lambda \neq 0$ we have
\[
\abs{\sigma_p(\lambda)} = \abs{\sum_{\abs{\alpha} = N} c_\alpha \lambda^\alpha} = \norm{\lambda}^N \abs{\sigma_p(\lambda/\norm{\lambda})}  \geq \norm{\lambda}^N C_0. 
\]
By the triangle inequality we find
\[
\abs{p(\lambda)} \geq \abs{\sigma_p(\lambda)} - \abs{\sigma_p(\lambda) - p(\lambda)} \geq \qty[ C_0 - \abs{\frac{p(\lambda) - \sigma_p(\lambda)}{\norm{\lambda}^N}}] \norm{\lambda}^N
\]
Since $p - \sigma_p$ is a polynomial of order $N - 1$, we can choose $R$ sufficiently large s.t.\ $\abs{\frac{p(\lambda) - \sigma_p(\lambda)}{\norm{\lambda}^N}} < C_0/2$. Since $\ang{\lambda} \sim \norm{\lambda}$ for $\lambda$ large enough, we find that there exists $C$ such that
\[
\abs{p(\lambda)} \geq \frac{C_0}{2} \norm{\lambda}^N \geq C \ang{\lambda}^N
\]
for $\norm{\lambda} > R$. 
\end{proof}

We will try to prove the \emph{elliptic regularity theorem}:
\begin{theorem}[Elliptic regularity] \label{thm:elliptic_regularity}
	Suppose $p(D)$ is an $N$-th order elliptic partial differential operator and $u \in \DD'(X)$ satisfies $p(D) u \in H^s\loc(X)$, then $u \in H^{s+N}\loc(X)$. 
\end{theorem}

\begin{corollary}
	If $p(D)$ is $N$-th order elliptic and $p(D)u \in C^\infty(X)$, then $u \in C^\infty(X)$. 
\end{corollary}

We will first prove an ``easy version'' of \cref{thm:elliptic_regularity} using a \emph{parametrix}: 
\begin{definition}
	If $p(D)$ is an $N$-th order differential operator, then $E \in D'(\RR^n)$ is called a \emph{parametrix} for $p(D)$ if
	\[
	p(D)E = \delta_0 + \omega \qquad\text{for some $\omega \in \EE(\RR^n)$}. 
	\]
\end{definition}

\begin{lemma}
	Every elliptic partial differential operator $p(D)$ has a parametrix which is smooth on $\RR^n \setminus \qty{0}$. 
\end{lemma}

\begin{proof}
	Since $p(D)$ is elliptic, we can choose $R > 0$, $C > 0$ such that $\abs{p(\lambda)} \geq C \ang{\lambda}^N$  for $\norm{\lambda} > R$.

	Fix some $\chi_R \in \DD(\RR^n)$ such that $\chi_R = 1$ on $\norm{\lambda} \leq R$ and $\chi_R = 0$ on $\norm{\lambda} > R + 1$, and define
	\[
	\hat E(\lambda) \ceq \frac{1 - \chi_R(\lambda)}{p(\lambda)}. 
	\]
	Then $\tilde E$ is smooth and for $\lambda$ sufficiently large we have $\abs*{\hat E(\lambda)} \les \ang{\lambda}^{-N}$ since $\chi_R$ vanishes for large $\lambda$, which implies $\hat E \in \SS'(\RR^n)$. Therefore, $p(\lambda) \hat E = 1 - \chi_R(\lambda)$ is also a tempered distribution and we can take its inverse Fourier transform $p(D) E = \delta_0 + \omega$ for some $\omega \in \SS(\RR^n)$, which shows that $E$ is a parametrix.

	To prove that $E$ is smooth on $\RR^n \setminus \qty{0}$, consider for $\norm{\lambda} > R + 1$
	\[
	\abs{\FF[D^\beta (x^\alpha E)]} = \abs{\lambda^\beta D^\alpha \hat E} = \abs{\lambda^\beta D^\alpha\qty(\frac1{p(\lambda)})} \overset\star\les \norm{\lambda}^{\abs{\beta} - \abs{\alpha} - N},
	\]
	where $\star$ can be shown with an induction argument. For each $\beta$, we can simply choose $\abs{\alpha}$ large enough such that $\FF[D^\beta (x^\alpha E)] \in L^1(\RR^n)$, and therefore $D^\beta(x^\alpha E)$ is continuous for $\abs{\alpha}$ large enough. Since $\beta$ was randomly chosen, $E$ will be smooth outside the origin. 
\end{proof}

We will now consider the proof of \cref{thm:elliptic_regularity} in the special case that $u$ and $f \ceq p(D)u$ have compact support. 
\begin{proof}
	Let $E$ be a parametrix for $P$, then we have
	\[
	u= \delta_0 * u = [p(D) E - \omega] * u = p(D)E * u - \omega * u = E * f - \omega * u. 
	\]
	Since $u$ has compact support, $\omega * u$ will be a Schwartz function, and it can be shown that
	\[
	\abs{\FF[E * f](\lambda)} = \abs{\hat E(\lambda) \hat f(\lambda)} \les \ang{\lambda}^{-N} \abs{\hat f(\lambda)}, 
	\]
	which shows that $f \in H^s(\RR^n) \implies u \in H^{s + N}(\RR^n)$. 
\end{proof}

To prove \cref{thm:elliptic_regularity} in general, we will need some facts which are proved on the second example sheet:
\begin{enumerate}
	\item If $s > t$ then $H^s(\RR^n) \subseteq H^t(\RR^n)$;
	\item If $\phi \in \SS(\RR^n)$, $u \in H^s(\RR^n)$, then $\phi u \in H^s(\RR^n)$;
	\item If $u \in \EE'(\RR^n)$, then $u \in H^t(\RR^n)$ for some $t \in \RR$;
	\item If $u \in H^s(\RR^n)$, then $D^\alpha u \in H^{s - \abs{\alpha}}(\RR^n)$. 
\end{enumerate}
Now we prove the theorem:
\begin{proof}
	Fix $\phi \in \DD(X)$, we wish to prove that $\phi u \in H^{s + N}(\RR^n)$ given that $p(D) u \in H^s\loc(X)$.  Choosing $M \in \NN$, we introduce a collection $\qty{\psi_0, \dotsc, \psi_M} \subseteq \DD(X)$ such that
	\[
	\supp(\phi) \subseteq \supp(\psi_M) \subseteq \dotsb \subseteq \supp(\psi_0), \quad \psi_{i-1} = 1 \text{ on $\supp \psi_i$}, \quad \psi_M = 1 \text{ on $\supp\phi$}. 
	\]

	Consider $\psi_0 u \in \EE'(\RR^n)$. Then there exists $t \in \RR$ for which $\phi_0 u \in H^t(\RR^n)$. We compute
	\begin{align*}
		p(D)(\psi_1 u) = \psi_1 p(D) u + [p(D), \psi_1](u) = \psi_1 f + [p(D), \psi_1](\psi_0 u),
	\end{align*}
where the last equality follows from the fact that $\psi_0 u \equiv u$ on $\supp \psi_1$. Now note that $[p(D), \psi_1]$ is an order $N-1$ differential operator. So we have $\psi_1 f \in H^s(\RR^n)$ and $[p(D), \psi_1](\psi_0 u) \in H^{t - N + 1}(\RR^n)$.
Setting $\tilde A_1 \ceq \min(s, t - N + 1)$ we find that $p(D)(\psi_1 u) \in H^{\tilde A_1}(\RR^n)$.

Since $\abs{p(\lambda)} \ges \ang{\lambda}^N$, we find that
\begin{align*}
p(D)(\psi_1 u) \in H^{\tilde A_1}(\RR^n) &\implies \int_{\RR^n} \ang{\lambda}^{2 \tilde A_1} \abs{p(\lambda) \FF[\psi_1u](\lambda)}^2 \dd{\lambda} \\
&\implies \int_{\RR^n} \ang{\lambda}^{2\tilde A_1 + 2N} \abs{\FF[\psi_1 u](\lambda)}^2 \dd{\lambda} \\
&\implies \psi_1 u \in H^{\tilde A_1 + N}(\RR^n).
\end{align*}
Define $A_1 \ceq \tilde A_1 + N = \min\qty{s + N, t + 1}$, then we have shown that $\psi_1 u \in H^{A_1}(\RR^n)$. By carrying on inductively, we can show that $\psi_M u \in H^{A_M}(\RR^n)$ where $A_M = \min\qty{s + N, t + M}$. By choosing $M$ large enough we conclude $\psi_M u \in H^{s + N}(\RR^n)$, and since $\psi_M = 1$ on $\supp\phi$, this also shows that $\phi u \in H^{s + N}(\RR^n)$. Since $\phi$ was randomly chosen, it follows that $u \in H^{s+N}\loc(X)$.  
\end{proof}

\subsection{Fundamental solutions}
\begin{definition}
	Let $p(D)$ be a partial differential operator, then $E \in \DD'(\RR^n)$ is called a \emph{fundamental solution} for $p(D)$ if $p(D)E = \delta_0$. 
\end{definition}

\begin{example}
	Let $z = x_1 + i x_2 \in \CC$ and define the Cauchy-Riemann operator as $\pdv{\bar z} \ceq \frac12 \qty(\pdv{x_1} + i \pdv{x_2})$. 
	It can be shown that $E \ceq \frac1{\pi z}$ is a fundamental solution of this equation. 
\end{example}

\begin{example}
	Let $p(D) = \pdv{t} - \Delta x$ be the heat operator (where $\Delta = \pdv[2]x_1 + \dotsb + \pdv[2]{x_n}$) with coordinates $(t, x) \in \RR \times \RR^n$. Then it can be shown that
	\[
	E \ceq \begin{cases}
		(4\pi t)^{-n/2} \exp(-\frac{\norm{x}}{4t}), &t > 0, \\ 0, & t \leq 0,
	\end{cases}
	\]
	is a fundamental solution.

	Furthermore, if $f$ has compact support, then $u = E * f$ solves $p(D) u = f$, since in this case
	\[
	p(D)(E * f) = (p(D) E * f) = \delta_0 * f = f. 
	\]
\end{example}

### 2020-11-08 inverse_problems/C3.tex
lots of work done, lots of confusion going on

@@
 
-\begin{theorem}
+\begin{theorem} \label{thm:minimiser_existence}
 	Let $\XX, \YY$ be Banach spaces with topologies $\tau_\XX, \tau_\YY$. Assume that:
@@
 \end{proof}
+
+\begin{theorem}
+	Let the assumptions of \cref{thm:minimiser_existence} be satisfied and suppose that $\inf\limits_{v \in \XX} \norm{Av - f} = 0$. 
+	Let $\alpha = \alpha(\delta)$ satisfy \[
+	\lim_{\delta\to0} \alpha(\delta) = 0  = \limsup_{\delta \to 0} \frac{\delta^2}{\alpha(\delta)}. 
+	\]
+	Then any sequence $(\delta_n) \to 0$ has a subsequence $(\delta_n) \to 0$ such that $u_{\delta_n} \ceq u_{\alpha(\delta_n)} \to u_\JJ \D$ and $\JJ(u_\delta) \to \JJ(u_\JJ\D)$, where $u_\JJ\D$ is a $\JJ$-minimising solution.
+\end{theorem}
+
+\begin{proof} 
+	By \cref{lem:regulariser_coercive}, we know that $\Phi_\alpha$ is coercive, and by \cref{thm:minimiser_existence}, we know there exists a $\JJ$-minimising solution $u_\JJ\D$.
+For any $(u_{\delta_n})$ with $\delta_n \to 0$ we find
+	\begin{equation} \label{eq:phi_alpha}
+	\Phi_\alpha(u_{\delta_n}) \leq \Phi_\alpha(u_\JJ\D) = \frac12\norm{A u_\JJ\D - f_\delta}^2 + \alpha(\delta) \JJ(u_\JJ\D) \leq \frac{\delta^2}{2} + \alpha(\delta)\JJ(u_\JJ\D),
+	\end{equation}
+	which is bounded, so by coercivity of $\Phi_\alpha$ we conclude that $(u_{\delta_n})$ is bounded.
+
+	After taking a subsequence if necessary, we conclude that $u_{\delta_n} \overset{\tau_X}\to u_0$.
+We show that $u_0$ is a $\JJ$-minimising solution.
+Since $A \colon \tau_X \to \tau_Y$ is continuous and $\norm{\cdot}_\YY$ is $\tau_\YY$-LSC, we get
+	\[
+	\norm{Au_0 - f} \leq \liminf_{n\to\infty} \norm{Au_{\delta_n} - f} \leq \liminf_{n\to\infty} \norm{Au_{\delta_n} - f_{\delta_n}} + \norm{f - f_{\delta_n}} = 0,
+	\]
+	since $\norm{Au_{\delta_n} - f\delta_n} \leq 2 \Phi_\alpha(u_{\delta_n}) \to 0$ by \cref{eq:phi_alpha}. This shows that $u_0$ is a least-squares solution.
+
+	Furthermore, we have 
+	\[
+	\JJ(u_0) \leq \liminf_{n\to\infty} \JJ(u_{\delta_n}) \leq \limsup_{n\to\infty} \JJ(u_{\delta_n}) \leq \limsup_{n\to\infty} \frac{\Phi_\alpha(u_{\delta_n})}{\alpha(\delta)} \leq \limsup_{n\to\infty} \frac12 \frac{\delta_n^2}{\alpha(\delta_n)} + \JJ(u_\JJ\D) = \JJ(u_\JJ\D)
+	\]
+	which proves that $u_0$ minimises $\JJ$.
+
+	To prove that $u_{\delta_n}$ is actually a minimising sequence, note that all above inequalities are in fact equalities, so $\liminf_{n\to\infty} \JJ(u_{\delta_n}) = \limsup_{n\to\infty} \JJ(u_{\delta_n}) = \lim_{n\to\infty} \JJ(u_{\delta_n}) = \JJ(u_\JJ\D)$.
+
+%	By the previous theorem, we know there exists a $\JJ$-minimising solution $u_0$, and for every $\alpha > 0$ we know there exists
+%	\[
+%	u_\delta \ceq u_{\alpha(\delta)} \in \argmin_u \frac12 \norm{Au - f}^2 + \alpha \JJ(u). 
+%	\]
+%	
+%	Therefore, we have 
+%	\begin{equation} \label{eq:estimate}
+%	\frac12 \norm{Au_\delta - f_\delta}^2 + \alpha(\delta) \JJ(u_\delta) \leq \frac12 \norm{A u_0 - f_\delta}^2 + \alpha(\delta) \JJ(u_0) \leq \frac{\delta^2}{2} + \alpha(\delta) \JJ(u_0). 
+%	\end{equation}
+%	\Cref{eq:estimate} gives us two estimates: the first one is
+%	\[
+%	\alpha(\delta) \JJ(u_\delta) \leq \frac{\delta^2}{2} + \alpha(\delta) \JJ(u_0) \implies \JJ(u_\delta) \leq \frac12 \frac{\delta^2}{\alpha(\delta)} + \JJ(u_0), 
+%	\]
+%	so $\JJ(u_\delta)$ is bounded uniformly in $\delta$. From \cref{eq:estimate} we also infer
+%	\[
+%	\frac12 \norm{Au_\delta - f_\delta}^2 \leq \frac{\delta^2}{2} + \alpha(\delta) \JJ(u_0),
+%	\]
+%	from which it follows that $\norm{Au_\delta - f_\delta}$ is also uniformly bounded in $\delta$. 
+%	
+%	Writing $u_\delta = u_\delta^0 + u_\delta^{\Cal N}$
+\end{proof}
+
+\subsection{Total variation regularisation}
+\begin{definition}
+	Let $\Omega \subseteq \RR^n$ be a bounded comain and $u \in L^1(\Omega)$. Let
+	\[
+	\DD(\Omega, \RR^n) \ceq \qty{\phi \in C_0^\infty(\Omega, \RR^n) \mid \sup_{x \in \Omega} \norm{\phi(x)}_2 \leq 1}. 
+	\]
+	We define the \emph{total variation} of $u \in L^1(\Omega)$ as
+	\[
+	\TV(u) \ceq \sup_{\phi \in \DD(\Omega, \RR^n)} \int_\Omega u(x) (\div \phi)(x) \dd{x}. 
+	\]
+\end{definition}
+
+\begin{example}
+	If $u \in L^1(\Omega)$ has a weak derivative $\grad u \in L^1$ (i.e., $u \in W^{1, 1}$), then since any $\phi \in \DD(\Omega, \RR^n)$ is 0 on the boundary of $\Omega$, we have
+	\[
+	\TV(u) = \sup_{\phi \in \DD(\Omega, \RR^n)} \int_\Omega u(x) (\div\phi)(x) \dd{x} =  \sup_{\phi \in \DD(\Omega, \RR^n)} \int_\Omega -\grad u \vdot \phi \dd{x}, 
+	\]
+	By Cauchy-Schwarz we have $\abs{\nabla u(x) \vdot \phi(x)} \leq \norm{\nabla u(x)}_2 \norm{\phi(x)}_2$ which gives an upper bound for $\TV(u)$, and it can be shown that this upper bound is indeed an equality, so
+	\[
+	\TV(u) = \int_\Omega \norm{\nabla u}_2 \dd{x} = \norm{\nabla u}_{L^1}.
+	\]
+\end{example}
+
+\begin{proposition}
+	The functional $\TV$ is proper, convex, and absolutely one-homogeneous, and for any constant $c$ and any $u \in L^1(\Omega)$ we have
+	\[
+	\TV(c) = 0, \quad \TV(u + c) = \TV(u). 
+	\]
+\end{proposition}
+
+\begin{proof}
+	For any test function $\phi$ we have 
+	\[
+	\int_\Omega \div\phi \dd{x} = \int_{\pt\Omega} \phi \vdot \vec n \dd{x} = 0
+	\]
+	since $\phi = 0$ on the boundary. This shows immediately that $\TV(c) = 0$ and that $\TV(u + c) = \TV(u)$. 
+\end{proof}
+
+\begin{remark}
+	In fact, the converse also holds: $\TV(u) = 0$ if and only if $u$ is constant. 
+\end{remark}
+
+\begin{example}
+	Let $C \subseteq \Omega$ be a bounded domain with smooth boundary. Then we have
+	\[
+	\TV(\ind_C) = \sup_{\phi \in \DD(\Omega, \RR^n)} \int_C\div\phi \dd{x} = \sup_{\phi \in \DD(\Omega, \RR^n)} \int_{\pt C} \phi \vdot \vec n \dd{s} \leq \sup_{\phi \in \DD(\Omega, \RR^n)} \int_{\pt C} \norm{\phi}_2 \norm{\vec n}_2  \dd{s} = \int_{\pt C} 1 \dd{s},
+	\]
+	and we can show that the inequality here is in fact an equality, i.e., $\TV(\ind_C) = \int_{\pt C} 1 \dd{s}$, which is the perimeter of $C$. 
+\end{example}
+
+\begin{definition}
+	We define the space of functions of \emph{bounded variation} as
+	\[
+	\BV(\Omega) \ceq \qty{u \in L^1(\Omega) : \norm{u}_{\BV} \ceq \norm{u}_{L^1} + \TV(u) < \infty}.
+	\]
+\end{definition}
+\begin{remark}
+	It can be shown that BV is the dual of a separable Banach space, and that weak-$*$ convergence in BV is equivalent to strong convergence in $L^1$ and convergence of the values $\TV(u_n) \to \TV(u)$. 
+\end{remark}
+
+\begin{definition}
+	We define a \emph{domain} in $\RR^n$ as an open connected subset, and it is called a \emph{Lipschitz domain} if its boundary is locally the graph of a Lipschitz continuous function (see Wikipedia for the details).
+\end{definition}
+
+\begin{theorem}
+	Let $\Omega \subseteq \RR^n$ be a bounded Lipschitz domain and $p, m \in \NN$. Define 
+	\[
+	p^* \ceq \begin{cases} \frac{np}{n - mp}, &n > mp, \\ \infty,& n \leq mp. \end{cases}
+	\]
+	Then the embedding $W^{m, p} \to L^q(\Omega)$ is continuous for all $q \in [1, p^*]$ and compact for all $q \in [1, p^*)$. 
+\end{theorem}
+
+A consequence ($p = m = 1$) is the following:
+\begin{corollary}
+	For any bounded Lipschitz domain $\Omega \subseteq \RR^n$, the $\BV(\Omega)$ is compactly embedded in $L^1(\Omega)$ for $n \geq 2$.
+For $n = 2$, the embedding $\BV(\Omega) \to L^2(\Omega)$ is continuous.
+\end{corollary}
+
+\begin{theorem}
+	Let $\Omega \subseteq \RR^n$ be open and bounded. Then $\TV$ is LSC in $L^1(\Omega)$. 
+\end{theorem}
+
+\begin{proof}
+	Let $(u_j) \subseteq \BV(\Omega)$ converge to $u \in L^1(\Omega)$. Then for any test function we have
+	\begin{align*}
+		\TV(u) &= \sup_{\phi \in \DD(\Omega, \RR^n)}\int_\Omega u_j(x) (\div \phi)(x) \dd{x} \\
+		&= \sup_{\phi \in \DD(\Omega, \RR^n)} \lim_{j\to\infty} \int_\Omega u_j(x) (\div \phi)(x) \dd{x} \\
+		&\leq \liminf_{j\to\infty} \sup_{\phi \in \DD(\Omega, \RR^n)} \int_\Omega u_j(x) (\div\phi)(x) \dd{x} \\
+		&= \liminf_{j\to\infty} \TV(u_j). 
+	\end{align*}
+\end{proof}

### 2020-11-08 modern_statistical_methods/C2.tex
lots of work done, lots of confusion going on

@@
 
-We have the following:
-\begin{theorem}[Slow rate]
+We have the following bound on the prediction error:
+\begin{theorem}[Slow rate] \label{thm:lasso_slow_rate}
 	Assume $X$ has centred and scaled columns, and assume that $Y$ has been centred, so $Y = X\beta^0 + \eps - \bar\eps \vec 1$. Let $A > 0$ and suppose
@@
 	\[
-	\PP\qty(\frac{\norm{X\T\eps}_\infty}{n} \leq \lambda) \geq 1 - 2p^{-(A^2/2 - 1)}. 
+	\PP\qty(\frac{\norm{X\T\eps}_\infty}{n} \leq \lambda) \geq 1 - 2p\exp(\frac{-\lambda^2}{2(\sigma/\sqrt n)^2}) =  1 - 2p^{-(A^2/2 - 1)}. 
 	\]
@@
 \subsubsection{Optimisation theory and convex analysis}
+\subsubsection*{Lagrangian optimisation}
+\begin{definition}
+	For the optimisation problem
+	\begin{equation} \label{eq:opt_primal}
+	\inf \qty{f(x) \mid g(x) = 0, h(x) \leq 0},
+	\end{equation}
+ 	its \emph{Lagrangian} is given by
+ 	\[
+ 	L(x, \lambda,\nu) \ceq f(x) + \lambda h(x) + \nu g(x),
+ 	\]
+ 	and its \emph{Lagrange dual function} is
+ 	\begin{equation} \label{eq:opt_dual}
+ 	\tilde f(\lambda, \nu) = \inf_x L(x, \lambda, \nu). 
+ 	\end{equation}
+\end{definition}
+
+Note that if $p^*$ is the minimum value of \eqref{eq:opt_primal}, then for any feasible $x$ (i.e., $x$ with $g(x) = 0$, $h(x) \leq 0$)
+we have for $\lambda \geq 0$ that 
+\[
+f(x) \geq L(x, \lambda, \nu) \geq \tilde f(\lambda, \nu).
+\]
+By taking the infimum over $x$ in the left-hand side and the supremum over $\lambda, \nu$ in the right-hand side we find that
+\[
+p^*  \geq \sup_{\lambda \geq 0, \nu} \tilde f(\lambda, \nu),
+\]
+Sometimes we actually have equality here, and in this case we say that the \emph{duality gap} is zero. Note that $\tilde f$ is a concave function, and is therefore easy to maximise.
+
+A sufficient condition for a zero duality gap is that $L(\cdot, \lambda, \nu)$ is convex and that the problem is feasible.
+
+\subsubsection*{Convex analysis}
+
+Let $\rext = \RR \cup \qty{\infty}$, and define for any $f \colon \RR^d \to \rext$ its (\emph{effective}) \emph{domain} $\dom(f) \ceq \qty{x \mid f(x) < \infty}$.
 \begin{proposition}
@@
 \end{proposition}
+
+\begin{definition}
+	Let $f \colon \RR^d \to \rext$  be convex and $x \in \RR^d$. Then $v \in \RR^d$ is called a \emph{subgradient} of $f$ at $x$ if
+	\[
+	f(y) \geq f(x) + v\T(y-x) \qquad\forall y \in \RR^d. 
+	\]
+	The set of all subgradients of $f$ at $x$ is called the \emph{subdifferential} of $f$ at $x$, denoted by $\pt f(x)$. 
+\end{definition}
+
+\begin{proposition}
+	If $f$ is convex and differentiable at $x \in \Int(\dom f)$, then $\pt f(x) = \qty{\nabla f(x)}$. 
+\end{proposition}
+
+\begin{proposition}
+	Let $f, g \colon \RR^d \to \rext$ be convex with $\Int(\dom f) \cap \Int(\dom g) \neq\emptyset$, $x \in \RR^d$ and $\alpha > 0$.
+Then $\pt(\alpha f)(x) = \alpha \pt f(x)$ and $\pt(f+g)(x) = \pt f(x) + \pt g(x)$.
+\end{proposition}
+
+\begin{proposition}[KKT conditions]
+	Let $f \colon \RR^d \to \rext$ be convex. Then $x^*$ is a global minimiser of $f$ if and only if $0 \in \pt f(x^*)$. 
+\end{proposition}
+
+\begin{proof}
+	Clearly $x^*$ is a global minimiser of $f$ if and only if $f(y) \geq f(x^*) + 0\T (y - x)$ for all $y \in \RR^d$. 
+\end{proof}
+
+We define for $x \in \RR$
+\[
+\sign(x) \ceq\begin{cases}
+	-1, &x<0,\\ 0, &x=0, \\ 1, &x>0,
+\end{cases}
+\]
+and for $x \in \RR^d$ we define $\sign(x) \ceq (\sign(x_1), \dotsc, \sign(x_n))$.
+
+\begin{proposition}
+	Let $x \in \RR^d$ and $A \ceq \qty{j \mid x_j \neq 0}$. Then
+	\[
+	\pt \norm{x}_1 = \qty{v \in \RR^d : \norm{v}_\infty \leq 1 \text{ and } v_A = \sign(x_A)}. 
+	\]
+\end{proposition}
+
+\begin{proof}
+	Define $g_j(x) = \abs{x_j}$ so that $\norm{\cdot}_1 = \sum_j g_j$ and $\pt \norm{x}_1 = \sum_j \pt g_j(x)$.
+Now, if $x_j \neq 0$, then $g_j$ is differentiable at $x$ so $\pt g_j(x) = \qty{\sign(x_j) e_j}$.
+If $x_j = 0$, then it can be shown that
+	\[
+	\pt g_j(x) = \qty{c e_j \mid c \in [-1, 1]}.
+	\]
+	Combining the above facts proves the claim. 
+\end{proof}
+
+With the previous proposition, we can write characterise solutions to the Lasso: namely we have
+\[
+\hat\beta_\lambda^L = \argmin_\beta \frac1{2n} \norm{Y- X\beta}_2^2 + \lambda \norm{\beta}_1 = \argmin_\beta Q(\lambda)
+\]
+if and only if $0 \in \pt Q(\hat \beta_\lambda^L)$. A simple computation gives, with $A = \qty{k \mid \beta_k \neq 0}$, 
+\begin{align*}
+\pt Q(\beta) &= \pt(\frac1{2n} \norm{Y- X\beta}_2^2) + \pt(\lambda \norm{\beta}_1) \\
+&= \qty{-\frac1n X\T(Y - X\beta)} + \qty{\lambda \nu : \norm{\nu}_\infty \leq 1, \nu_A = \sign(\beta_A)}. 
+\end{align*}
+
+We conclude that $\hat\beta$ is a Lasso solution if and only if there exists a vector $\hat\nu$ with $\norm{\hat\nu}_{\infty} \leq 1$ and $\hat\nu_A = \sign(\hat\beta_A)$ such that
+\[
+\frac1n X\T(Y - X\hat\beta) = \lambda \hat\nu. 
+\]
+
+Note that $\hat\beta^L_\lambda$ need not be unique, however, the fitted values are unique:
+\begin{proposition}
+	$X\hat\beta_\lambda^L$ is unique. 
+\end{proposition}
+
+\begin{proof}
+	Write $\frac1{2n}\norm{Y- X\beta}_2^2 = Q_1(X \beta)$ and $\norm{\beta}_1 = Q_2(\beta)$.
+Note that $Q_1$ is strictly convex.
+Now, for any $t\in (0, 1)$ and any two lasso solutions $\hat\beta, \hat\gamma$ with $Q(\hat\beta) = Q(\hat\gamma) = c^*$ we have
+
+	\begin{align*}
+		c^* &\leq Q(t\hat\beta + (1-t)\hat\gamma) = Q_1(tX\hat\beta+(1-t)X\hat\gamma) + Q_2(t\hat\beta + (1-t)\hat\gamma) \\
+		&\overset\star\leq t Q_1(X\hat\beta) + (1-t)Q_2(X\hat\gamma) + Q_2(t\hat\beta + (1-t)\hat\gamma) \\
+		&\leq t Q(\hat\beta) + (1-t)Q(\hat\gamma) = c^*,
+	\end{align*}
+so all inequalities must be equalities.
+In particular, by $\star$ we find $Q_1(tX\hat\beta + (1-t)X\hat\gamma) = t Q_1(X\hat\beta) + (1-t)Q_1(X\hat\gamma)$, and since $Q_1$ is strictly convex this implies $X\hat\beta = X\hat\gamma$.
+\end{proof}
+
+\subsection{Variable selection}
+We consider the noiseless model $Y = X\beta^0$, where no randomness is involved.
+Let $S = \qty{k \mid \beta_k \neq 0}$, $N = \qty{1, \dotsc, p} \setminus S$, and assume that $S = \qty{1, \dotsc, s}$ (so $\beta_1, \dotsc, \beta_s$ are nonzero and $\beta_{s+1}, \dotsc, \beta_p$ are zero).
+
+Furthermore, write $X = \mqty[ X_S & X_N]$ and assume that $\rank(X_S) = s$, so that $X_S\T X_S$ is invertible (NB in particular this implies $n \geq s$). 
+\begin{theorem}
+Let $\lambda > 0$ and define $\Delta = X_N\T X_S(X_S\T X_S)^{-1} \sign(\beta^0_S)$. 
+\begin{enumerate}
+	\item If $\norm{\Delta}_{\infty} \leq 1$ and for $k = 1, \dotsc, s$ we have
+	\[
+	\abs{\beta^0_k} > \lambda \abs{\sign(\beta^0_S)\T \qty[\frac1n X_S\T X_S]^{-1}_k},
+	\]
+	then there exists a Lasso solution $\hat\beta_\lambda^L$ with $\sign(\hat\beta_\lambda^L) = \sign(\beta^0)$.
+
+	\item If there exists a Lasso solution $\hat\beta_\lambda^L$ with $\sign(\hat\beta_\lambda^L) = \sign(\beta^0)$, then $\norm{\Delta}_{\infty} = 1$. 
+\end{enumerate}
+\end{theorem}
+
+\begin{remark}
+	The condition $\norm{\Delta}_{\infty} \leq 1$ is called the \emph{irrepresentable condition}, and it roughly states that none of the insignificant predictors align ``too well'' with the response.
+\TODO Elaborate on this (lecture 12).
+\end{remark}
+\begin{proof}
+	Let $\hat\beta$ be any lasso solution and write $\hat S = \qty{k \mid \beta_k \neq 0}$. Using the fact that $\beta^0_N = 0$, the KKT conditions for the Lasso can be expanded as
+	\begin{equation} \label{eq:expanded_lasso_conditions}
+	\frac1n \mqty(X_S\T X_S & X_S\T X_N \\ X_N\T X_S & X_N\T X_N) \mqty(\beta^0_S - \hat\beta_S\\ -\hat\beta_N) = \lambda \mqty(\hat\nu_S\\ \hat\nu_N)
+	\end{equation}
+	for some $\hat\nu$ with $\norm{\hat\nu}_\infty \leq 1$ and $\hat\nu_{\hat S} = \sign(\hat\beta_{\hat S})$. 
+	\begin{enumerate}
+		\item It is easily checked that taking
+		\begin{align*}
+		(\hat\beta_S, \hat\beta_N) &= \qty(\beta^0_S - \lambda \qty(\frac1n X_S\T X_S)^{-1} \sign(\beta^0_S), 0) \\
+		(\hat\nu_S, \hat\nu_N) &= (\sign(\beta^0_S), \Delta)
+		\end{align*}
+		is a Lasso solution with $\sign(\hat\beta^0_S) = \sign(\hat\beta_S)$. 
+		\item If $\sign(\hat\beta) = \sign(\beta^0)$, then $\hat S = S$, $\hat \nu_S= \sign(\beta^0_S)$ and $\hat\beta_N = 0$. The top block of \cref{eq:expanded_lasso_conditions} can be written as
+		\[
+		\frac1n X_S\T X_S (\beta^0_S - \hat\beta_S) = \lambda \sign(\beta^0_S) \implies \beta^0_S - \hat\beta_S = \lambda \qty[\frac1n X_S\T X_S]^{-1} \sign(\beta^0_S). 
+		\]
+		Plugging that into the bottom block of the same equation, we get
+		\[
+		\lambda \frac1n X_N\T X_S \qty[\frac1nX_S\T X_S]^{-1} \sign(\beta^0_S) = \lambda \hat\nu_N \implies \norm{\Delta}_\infty \leq 1. 
+		\]
+	\end{enumerate}
+\end{proof}
+
+\subsubsection{Prediction and estimation}
+\begin{definition}
+	Let $X \in \RR^{n \times p}$ be a design matrix and $S \subseteq \qty{1, \dotsc, p}$ where $s \ceq \abs{S} > 0$, then the \emph{compatibility constant} $\phi^2 = \phi^2(X, S)$ is defined as
+	\[
+	\phi^2 \ceq \inf\qty{\frac{\frac1n \norm{X\beta}_2^2}{\frac1s \norm{\beta_S}_1^2} : \beta \in \RR^p, \beta_S \neq 0, \norm{\beta_N}_1 \leq 3 \norm{\beta_S}_1}. 
+	\]
+	The \emph{compatibility condition} is that $\phi^2 > 0$. 
+\end{definition}
+
+Note that if $X_S$ is rank-deficient, then there exists  $\beta_S \neq 0$ with $X_S\beta_S = 0$, and by setting $\beta_N = 0$ we find that $\phi^2 = 0$.
+Therefore, $X_S$ must have full column rank for the compatibility condition to be satisfied, and in particular we must have $n \geq s$.
+
+\begin{proposition}
+	If $X\T X/n$ has minimal eigenvalue $c_\Rm{min}$ then $\phi^2 \geq c_\Rm{min}$. In particular, if $X$ has full column rank (i.e., $c_\Rm{min} > 0$), the compatibility condition is always satisfied. 
+\end{proposition}
+
+\begin{proof}
+	We have
+	\[
+	\norm{\beta_S}_1 = \sign(\beta_S)\T \beta_S\leq \norm{\sign(\beta_S)}_2 \norm{\beta_S}_2 = \sqrt{s} \norm{\beta}_2.
+	\]
+	Therefore, we find
+	\[
+	\phi^2 \geq \inf_{\beta \neq 0}  \frac{\frac1n \norm{X\beta}_2^2}{\frac1s \norm{\beta}_1^2} \geq \frac1n \inf_{\beta \neq 0} \frac{\norm{X\beta}_2^2}{\norm{\beta}_2^2} = \inf_{\norm{\beta} =1 } \beta\T \qty(\frac{X\T X}{n}) \beta = c_\Rm{min}.
+	\]
+\end{proof}
+
+Note that the converse of this proposition does not necessarily hold: $X$ does not need to have full column rank to satisfy the compatibility condition.
+In particular, it does not need to be the case that $n < p$.
+
+Now we present the ``fast'' convergence rate of the lasso (compare with \cref{thm:lasso_slow_rate}):
+\begin{theorem}[Fast rate] \label{thm:lasso_fast_rate}
+	Let the conditions of \cref{thm:lasso_slow_rate} be satisfied and assume furthermore that the compatibility condition holds. 
+\end{theorem}

### 2020-11-08 modern_statistical_methods/msm_ES2.tex
distribution theory is still difficult :((
lots of work done, lots of confusion going on

@@
 Replace $R$ with the residual from regressing $R$ onto $X_{k^*}$.
-Further replac eeach variable in $M^c$ with the residual from regressing itself onto $X_{k^*}$.
+Further replace each variable in $M^c$ with the residual from regressing itself onto $X_{k^*}$.
 		\item Continue the previous step unto $R = 0$. 
@@
 \begin{solution}
-	content...
+We follow the hint and use induction on the iteration $m$ of the algorithm. For the base case, note that centering $Y$  is equivalent to fitting an intercept-only model. ???
 \end{solution}
@@
 	\begin{enumerate}[(a)]
-		\item 
+		\item We have
+		\begin{align*}
+		\PP(\abs{W/d - 1} \geq t) &= \PP(W/d - 1 \geq t)+ \PP(W/d - 1 \leq -t) \\
+		&= \PP(W\geq d(1+t)) + \PP(W \leq d(1 -t)). 
+		\end{align*}
+
+		Now we use Chernoff bounds: we have 
+		\begin{align*}
+		\PP(W \geq d(1+t))  &\leq \inf_{\alpha > 0} e^{-\alpha d(1 + t)} \EE[e^{\alpha W}] \leq \inf_{0 < \alpha < \frac12} e^{-\alpha dt} \qty(\frac{e^{-\alpha}}{\sqrt{1 - 2\alpha}})^d  \\
+		&\leq \inf_{0 < \alpha < \frac14} e^{d(2\alpha^2 - \alpha t)} \overset\star=  e^{dt^2/8}, 
+		\end{align*}
+		where $\star$ is a consequence of plugging in the minimum $\alpha = t/4 \in (0, 1/4)$.
+
+		Analogously, we have for $\alpha > 0$ that
+		\begin{align*}
+			\PP(W \leq d(1-t)) &\leq \inf_{\alpha > 0} e^{\alpha d(1-t)} \EE[e^{-\alpha W}] \leq \inf_{-\frac12 < \alpha < 0} e^{\alpha d t} \qty(\frac{e^{-\alpha}}{\sqrt{1 - 2\alpha}})^d \\
+			&\leq \inf_{-\frac14 < \alpha < 0} e^{d(2\alpha^2 + \alpha t)} \overset\star= e^{dt^2/8},
+		\end{align*}
+	where $\star$ is now a consequence of plugging in the minimum $\alpha = -t/4 \in (-1/4, 0)$.
+
+	Plugging what we found back into our original equation we obtain for $t \in (0, 1)$ indeed
+	\[
+	\PP(\abs{W/d - 1} \geq t) \leq 2e^{dt^2/8}. 
+	\]
+
+	\item We wish to show that $\norm{Au}^2/\norm{u}^2 \sim \chi_d^2$. For this, let $Z_1, \dotsc, Z_p$ be i.i.d.\ $N(0, 1)$ random variables, then 
+	\[
+	(Au)_i = \sum_{j=1}^p A_{ij} u_j \sim \sum_{j=1}^p u_j Z_j \sim N(0, \sum_i u_i^2) = N(0, \norm{u}_2^2). 
+	\]
+	Therfore, we have 
+	\[
+	\frac{\norm{Au}^2}{\norm{u^2}} = \frac{\sum_i (Au)_i^2}{\norm{u}^2} = \sum_i \qty(\frac{(Au)_i}{\norm{u}})^2 \sim \sum_i Z_i^2 \sim \chi_d^2. 
+	\]
+	Combining this with (a) proves the claim.
+
+	\item We have
+	\begin{align*}
+			&\PP\qty(1 - t \leq \frac{\norm{w_i - w_j}_2^2}{\norm{u_i - u_j}_2^2} \leq 1 + t \text{ for all } i \neq j \in \qty{1, \dotsc, n})  \\
+			&= \PP\qty(\abs{\frac{\norm{A (u_i - u_j)}^2}{d \norm{u_i - u_j}^2} - 1} \leq t \text{ for all } i \neq j \in \qty{1, \dotsc, n}) \\
+			&= 1 - \PP\qty(\abs{\frac{\norm{A (u_i - u_j)}^2}{d \norm{u_i - u_j}^2} - 1} \geq t \text{ for some } i \neq j \in \qty{1, \dotsc, n}) \\
+			&\overset\star\geq - \frac{n^2}{2} \cdot 2e^{-dt^2/8} \geq 1 - \eps, 
+	\end{align*}
+where $\star$ follows from the fact that there are $\frac{n(n-1)}{2} \leq \frac{n^2}{2}$ pairs $i \neq j$, and the last inequality follows from 
+\[
+e^{-dt^2/8} \leq e^{-2 \log(n/\sqrt\eps)} = \qty(\frac{n}{\sqrt\eps})^{-2} = \frac\eps{n^2}. 
+\]
 	\end{enumerate}
 \end{proof}
+
+\begin{mdframed}
+	In the following questions, assume that $X$ has had its columns centred and scaled to have $\ell_2$-norm $\sqrt n$, and that $Y$ is also centred.
+\end{mdframed}
+
+\begin{question}
+	Show that any two Lasso solutions when $\lambda > 0$ must have the same $\ell_1$-norm. 
+\end{question}
+
+\begin{proof}
+	By proposition 21, $X\hat\beta_\lambda^L$ is unique, so if $\hat\beta$, $\hat\gamma$ are Lasso solutions then $X\hat\beta = X\hat\gamma$ so 
+	\[
+	Q_\lambda(\hat\beta) = \frac1{2n} \norm{Y - X\hat\beta}_2^2 + \lambda \norm{\hat\beta}_1 = \frac1{2n} \norm{Y - X\hat\beta}_2^2 + \lambda \norm{\hat\gamma}_1 = Q_\lambda(\hat\gamma), 
+	\]
+	which implies $\lambda \norm*{\hat\beta}_1 = \lambda \norm{\hat\gamma}_1$. For $\lambda > 0$ this shows $\norm{\hat\beta}_1 = \norm{\hat\gamma_1}$. 
+\end{proof}
+
+\begin{question}
+	Carath\'eodory's Lemma states that if $S \subseteq \RR^{d'}$ is in a subspace of dimension $d$, any $v$ that is a convex combination of points in $S$ can be expressed as a convex combination of $d+1$ points from $S$.
+
+	With this knowledge, show that for any value of $\lambda$, there is always a Lasso solution with no more than $n$ non-zero coefficients. 
+\end{question}
+
+\begin{proof}
+	By the previous question, all Lasso solutions have the same norm $x$, so the set of Lasso solutions is given by
+	\[
+	\argmin_{\norm{\beta}_1 = x} \norm{Y - X\beta}_2^2.
+	\]
+	Let $\hat\beta$ be any fixed Lasso solution, then we find that the entire solution set can also be described as
+	\[
+	\qty{\hat\gamma \mid \norm{\gamma} = x, \hat\beta - \hat\gamma \in \Nul(X)}. 
+	\]
+	??
+\end{proof}
+
+\begin{question}
+	Show that if $\lambda \geq \lambda_\Rm{max} \ceq \norm{X\T Y}_\infty / n$ then $\hat\beta_\lambda^L = 0$. 
+\end{question}
+
+\begin{proof}
+	By the KKT conditions any lasso solution satisfies $\frac1n X\T (Y - X\hat\beta) = \frac1n X\T Y - \frac1n X\T X\hat\beta= \lambda \hat\nu$ for some $\nu$ with $\norm{\hat\nu}_\infty \leq 1$ and $\hat\nu_{\hat S_{\lambda}} = \sign(\hat\beta^L_{\lambda, \hat S_\lambda})$.
+
+	Suppose $\lambda \geq \lambda_\Rm{max}$, and suppose also that $\hat\beta \neq 0$, so $\norm{\hat\nu}_\infty = 1$. Then we find
+	\[
+	\frac1n \norm{X\T (Y - X\hat\beta)}_\infty = \norm{\lambda\hat\nu} = \lambda  \geq \frac1n \norm{X\T Y}_\infty, 
+	\]
+	so 
+	\[
+	\norm{X\T Y - X\T X\hat\beta}_\infty \geq \norm{X\T Y}_\infty. 
+	\]
+	??
+\end{proof}
 \end{document}

### 2020-11-08 topics_statistical_theory/topics_ES1.tex
lots of work done, lots of confusion going on

@@
 	Note that $X/n$ is the average of $n$ i.i.d.\ random variables $Y_i \sim \Rm{Bern}(p)$, where $Y_i \in [0, 1]$ for all $i$.  
+	\begin{enumerate}
+		\item
 
@@
 	\[
-	\PP(X/n - p \geq x) \leq \exp(-\frac{2n^2(1/2)^2}{n}) = \exp(- \frac n2),
+	\PP\qty(X/n - p \geq \frac12) \leq \exp(-\frac{2n^2(1/2)^2}{n}) = \exp(- \frac n2),
 	\]
@@
 
-	We continue with Bennett's inequality.
-We consider the random variables $Y_i - p$, which are bounded from above by $b = 1 - p$.
-Now Bennett's inequality tells us, with $\sigma_p^2 = \Var(Y_i - p) = p(1-p)$ that
+\item 	We continue with Bennett's inequality. We consider the mean-zero random variables $Y_i - p$, which are bounded from above by $b = 1 - p$. Now Bennett's inequality tells us that
 	\begin{align*}
-		\PP(X/n \geq x) &\leq \exp(- \frac{n}{(1-p)^2} h\qty(\frac{1-p}{2p(1-p)})) = \exp\qty(- \frac{n}{(1-p)^2} h\qty(\frac{1}{2p})).
+		\PP\qty(X/n \geq \frac12) &\leq \exp(- \frac{np(1-p)}{(1-p)^2} h\qty(\frac{1-p}{2p(1-p)})) = \exp\qty( -\frac{np}{1-p} \cdot  h\qty(\frac{1}{2p})).
 	\end{align*}
@@
 \[
-h\qty(\frac1{2p}) = (1 + \frac1{2p}) \log(1 + \frac1{2p}) - \frac1{2p} = \log(1 + \frac1{2p}) + \frac1{2p} \qty(\log(1 + \frac1{2p}) - 1) \overset{p \downarrow 0}\to +\infty. 
+h\qty(\frac1{2p}) = \qty(1 + \frac1{2p}) \log(1 + \frac1{2p}) - \frac1{2p} = \log(1 + \frac1{2p}) + \frac1{2p} \qty(\log(1 + \frac1{2p}) - 1),
 \]
-Since $\frac{n}{(1-p)^2}$ is clearly bounded by $n$, we conclude that 
+and therefore
 \[
-\PP(X/n \geq x) \to e^{-\infty} = 0. 
+\frac{np}{1-p} h\qty(\frac1{2p}) \geq  \frac{n}{2(1-p)} \qty(\log(1 + \frac1{2p}) - 1) \overset{p \to 0}\to \infty. 
 \]
+It follows that the Bennett upper bound converges to 0 as $p \to 0$.
 
@@
 
-We finish with Bernstein's inequality.  We have for $q \geq 3$ and $p \leq\frac12$ that 
+\item We finish with Bernstein's inequality.  We have for $q \geq 3$ that
+\[
+\frac{2(1-p)}{q!} \leq \frac2{q!} \leq 3^{2-q}, 
+\]
+and therefore we have
 \begin{align*}
 	\EE[(Y_i - p)_+^q] &= p(1-p)^q = \sigma_p^2 (1-p)^{q-1} = (q! \sigma_p^2 (1-p)^{q-2}/2)\cdot (2(1-p)/q!) \\
-	&\leq q!\sigma_p^2 ((1-p)/3)^{q-2}/2.
+	&\leq q!\sigma_p^2 ((1-p)/3)^{q-2}/2,
 \end{align*}
-
+so $Y_i - p$ satisfies Bernstein's condition with $c = \frac{1-p}{3}$.
 Now Bernstein's inequality tells us that
@@
 \PP(X/n \geq \frac12 \leq \exp\qty(- \frac{n(1/2)^2}{2(\sigma_p^2 + (1-p)/6)}) = \exp(- \frac{n}{8\sigma_p^2 + 4(1-p)/3}) \overset{p \to 0}{\to} \exp(-\frac{3n}{4}). \]
+	\end{enumerate}
+Of course, the true limit is 0 for any $n$, which is only given by Bennett's inequality. We also see that Hoeffding's inequality gives the most loose bound.
 
-Of course, the true limit is 0 for any $n$, which is only given by Bennett's inequality. We also see that Hoeffding's inequality gives the worst result. 
 \end{solution}
@@
 	\[
-	W_t \sim (1-t) N(0, t) + t N(0, 1 - t) = N\qty(0, \frac{t}{\sqrt{1 - t}}) + N\qty(0, \frac{1-t}{\sqrt t}) = N\qty(0, \frac{t}{\sqrt{1 - t}} + \frac{1-t}{\sqrt t}). 
+	W_t \sim (1-t) N(0, t) + t N(0, 1 - t) = N\qty(0, t(1-t)^2) + N\qty(0, t^2(1-t)) = N\qty(0, t(1-t)). 
 	\]
@@
 		\[
-		\MISE(\hat f_n) = (2\pi)^{-1/2} \qty{\frac1{nh} + (1 - \frac1n) \frac{1}{(h^2 + \sigma^2)^{1/2}} - \frac{2^{3/2}}{(h^2 + 2\sigma^2)^{1/2}} + \frac1\sigma}. 
+		\MISE(\hat f_n) = \frac1{2\pi^{1/2}} \qty{\frac1{nh} + (1 - \frac1n) \frac{1}{(h^2 + \sigma^2)^{1/2}} - \frac{2^{3/2}}{(h^2 + 2\sigma^2)^{1/2}} + \frac1\sigma}. 
 		\]
@@
 	\[
-	\MISE(\hat f_n) = \frac1n \int_\RR \qty[(\phi_h^2 * \phi_\sigma)(x) - (\phi_h * \phi_\sigma)^2(x)] \dd{x} + \int_{-\infty}^\infty \qty[ (\phi_h * \phi_\sigma)(x) - \phi_\sigma]^2 \dd{x}
+	\MISE(\hat f_n) = \frac1n \int_\RR \qty[(\phi_h^2 * \phi_\sigma)(x) - (\phi_h * \phi_\sigma)^2(x)] \dd{x} + \int_{-\infty}^\infty \qty[ (\phi_h * \phi_\sigma)(x) - \phi_\sigma(x)]^2 \dd{x}
 	\]
+
+	We will use the previous exercise to compute all these terms. 
+	We have in general 
+	\begin{align}
+		(\phi_h * \phi_\sigma)(x) &= \int_\RR \phi_h(x-z) \phi_\sigma(z) \dd{z}\nonumber \\
+		&= \int_\RR \phi_\sigma(z) \phi_h(z-x) \dd{z} \nonumber\\
+		&= \phi_{(\sigma^2 + h^2)^{1/2}}(x) \int_\RR \phi_\xi(z - \mu^*) \dd{z} \nonumber\\
+		&= \phi_{(\sigma^2 + h^2)^{1/2}}(x) \label{eq:normal_convolve}. 
+	\end{align}
+
+	We also have
+	\begin{equation} \label{eq:normal_squared}
+	\phi_\sigma^2(x-\mu) = \phi_{\sigma/\sqrt 2}(x - \mu) \phi_{\sqrt 2 \sigma}(0) = \frac{1}{2 \sigma \sqrt\pi} \phi_{\sigma/\sqrt 2}(x - \mu). 
+	\end{equation}
+
+	Combining \cref{eq:normal_convolve,eq:normal_squared} we get 
+	\begin{align*}
+		(\phi_h^2 * \phi_\sigma)(x) &= \int_\RR \phi_h^2(x-z) \phi_\sigma(z) \dd{z} \\
+		%&= \phi_{\sqrt 2 h}(0) \int_\RR \phi_{h/\sqrt 2}(x - z)  \phi_\sigma(z) \dd{z} \\
+		&= \frac{1}{2h \sqrt \pi } \int_\RR \phi_{h/\sqrt 2}(x - z) \phi_{\sigma}(z)  \dd{z} \\
+		&= \frac{1}{2h\sqrt \pi} (\phi_{h/\sqrt 2} * \phi_\sigma)(x) \\
+		&= \frac{1}{2h \sqrt\pi} \phi_{(\sigma^2 + h^2/2)^{1/2}}(x)
+%		&= \frac{1}{2h \sqrt \pi} \phi_{\xi'}(x) \int_\RR \phi_\xi(z - \mu^*) \dd{z} \\
+%		&= \frac{1}{2h\sqrt\pi} \phi_{\xi'}(x),
+	%\\ &= \frac{1}{2\sqrt{\pi h}} \int_\RR \phi_{\xi}(z - \mu^*) \phi_{(\sigma^2 + h^2/2)^{1/2}}(x-z)
+	\end{align*}
 	\end{enumerate}
+
+	We also get
+	\[
+	(\phi_h * \phi_\sigma)^2(x) = \phi_{(\sigma^2 + h^2)^{1/2}}^2(x) = \frac{1}{2 (\sigma^2 + h^2)^{1/2} \sqrt\pi} \phi_{(\sigma^2 + h^2)^{1/2}/\sqrt 2}(x).
+	\]
+
+	Finally, we have
+	\begin{align*}
+		((\phi_h * \phi_\sigma)(x) - \phi_\sigma(x))^2 &= (\phi_h * \phi_\sigma)^2(x) - 2 (\phi_h * \phi_\sigma)(x) \phi_\sigma(x) + \phi_\sigma^2(x).
+	\end{align*}
+	The first term we already computed, the third term is $\frac{1}{2\sigma\sqrt\pi} \phi_{\sigma/\sqrt 2}(x)$, so we only need to compute
+	\begin{align*}
+		(\phi_h * \phi_\sigma)(x) \phi_\sigma(x) &= \phi_{(\sigma^2 + h^2)^{1/2}}(x) \phi_\sigma(x) = \phi_{\xi}(x) \phi_{(2\sigma^2 + h^2)^{1/2}}(0) = \frac{1}{\sqrt{2\pi} (2\sigma^2 + h^2)^{1/2}} \phi_\xi(x),
+	\end{align*}
+where $\xi$ is an irrelevant constant.
+
+Combining all these terms and using that $\phi_\sigma(x-\mu)$ integrates to 1 for any $\mu, \sigma$, we get
+\begin{align*}
+	\MISE \hat f_n &= \frac1n \qty(\frac1{2h\sqrt\pi} - \frac{1}{2(\sigma^2 + h^2)^{1/2}\sqrt\pi}) + \frac{1}{2(\sigma^2 + h^2)^{1/2}\sqrt\pi} - \frac{\sqrt 2}{\sqrt\pi (2\sigma^2 + h^2)^{1/2}} + \frac{1}{2\sigma\sqrt\pi} \\
+	&= \frac{1}{2\sqrt\pi} \qty(\frac{1}{nh} + (1 - \frac1n) \frac{1}{(\sigma^2 + h^2)^{1/2}} - \frac{2^{3/2}}{(2\sigma^2 + h^2)^{1/2}} + \frac1\sigma). 
+\end{align*}
+\end{proof}
+
+\begin{question}
+	Use the expression from 10(b) to derive an upper bound of the form $\MISE \hat f_n \leq C_1/(nh) + C_2^2 h^4$ (where you should specify $C_1, C_2$).
+Show that $\phi_\sigma \in \FF_\Nn(2, L)$ with $L^2 = 3/(8 \pi^{1/2} \sigma^5)$, and hence compare the bound from the first part of this question with that obtained from the general theory from lectures.
+\end{question}
+
+\begin{solution}
+	We have 
+	\[
+	\qty(1 - \frac1n) \frac{1}{(\sigma^2 + h^2)^{1/2}} - \frac{2^{3/2}}{(2\sigma^2 + h^2)^{1/2}} \leq \frac{1}{(\sigma^2 + h^2)^{1/2}}  - \frac{2}{(\sigma^2 + h^2/2)^{1/2}} < 0,
+	\]
+%	and therefore
+%	\[
+%	\MISE \hat f_n < \frac{1}{2\sqrt\pi} \qty(\frac1{nh} + \frac1\sigma). 
+%	\]
+	I do not know how to obtain an upper bound of the form $C_1/(nh) + C_2^2 h_4$ from this expression.
+
+	To show that $\phi_\sigma \in \FF_\Nn(2, L)$,  we must show that $\phi_\sigma \in \FF_\Nn(2, L)$. By question 9, it suffices to show that $\norm{\phi''_\sigma}_{L^2}^2 \leq L^2$. 
+	A simple computation gives
+	\[
+	\phi_\sigma''(x) = \frac{1}{\sqrt{2\pi} \sigma^5} \qty(x^2 - \sigma^2) \exp(-\frac{x^2}{2\sigma^2}) \leq \frac1{\sqrt{2\pi} \sigma^5} x^2 \exp(-\frac{x^2}{2\sigma^2})
+	\]
+	so
+	\begin{align*}
+	\norm{\phi_\sigma''}_{L^2}^2 &\leq \frac{1}{2\pi\sigma^{10}} \int_\RR x^4 \exp(-\frac{x^2}{\sigma^2}) \dd{x} \overset\star= \frac{1}{2\pi \sigma^{10}} \cdot \frac34 \sqrt{\pi} \sigma^5 = \frac{3}{8 \sqrt\pi \sigma^5} = L^2,
+	\end{align*}
+where $\star$ can be computed using the fact that the integral is, up to scaling, the fourth moment of $N(0, \sqrt2\sigma)$ distribution.
+
+Note that for $K = \phi_1$, we have 
+\[
+R(K) = \int_{-\infty}^\infty \phi_1^2(x) \dd{x} = \frac1{2\sqrt\pi}, 
+\]
+while
+\[
+\mu_2^2(K)=  \int_{-\infty}^\infty x^2 \phi_1(x) \dd{x} = 1.
+\]
+
+Plugging all the above into theorem 27 shows that 
+\[
+\MISE(\hat f_n) \leq \frac{1}{2\sqrt\pi} \frac1{nh} + \frac{3}{8\sqrt\pi \sigma^5} h^4.
+\]
+\end{solution}
+
+\begin{question}
+	Let $f$ be a bounded density that is twice differentiable at $x \in \RR$ and satisfies $R(f'') < \infty$.
+Let $h = h_n$ be deterministic, with $h \to 0$ and $nh \to \infty$ as $n\to\infty$, and let $K$ be a second-order kernel with $\max\qty{R(K), \mu_2(K)} < \infty$.
+Show that the KDE $\hat f_n \equiv \hat f_{n, h, K}$ satisfies
+	\[
+	\MSE (\hat f_n(x)) = \frac1{nh} R(K) f(x) + \frac14 \mu_2^2(K) f''(x)^2 h^4 + o\qty(\frac1{nh} + h^4)
+	\]
+	as $n\to\infty$. 
+\end{question}
+
+\begin{proof}
+	As in the proof of proposition 19, we have
+	\[
+	\Var \hat f_n(x) %= \frac1{nh} \int_{-\infty}^\infty K^2(u) f(x - uh) \dd{u}
+	\]
+%	Now we use a Taylor expansion 
+%	\[
+%	f(x - uh) = f(x) - f'(x) uh + \frac{(uh)^2}{2} \int_0^1 f''(x - \tau uh) (1 - \tau) \dd{\tau},
+%	\]
+%	and we obtain
+%	\[
+%	\Var \hat f_n(x) = \frac1{nh} \qty(f(x) \int_{-\infty}^\infty K^2(u) \dd{u})
+%	\]
 \end{proof}

### 2020-11-10 inverse_problems/C3.tex
I'm still alive

@@
 
-\begin{theorem}
+\begin{theorem} \label{thm:minimiser_convergence}
 	Let the assumptions of \cref{thm:minimiser_existence} be satisfied and suppose that $\inf\limits_{v \in \XX} \norm{Av - f} = 0$. 
@@
 \end{proof}
+
+\TODO what is the point/use of proposition 4.3.12 and corollary 4.3.13? 
+%\begin{proposition}
+%	Let $\Omega \subseteq \RR^n$ be a bounded Lipschitz domain, then there exists $C > 0$ such that for all $u \in \BV(\Omega)$, we have
+%	\[
+%	\norm{u - u_\Omega}_{L^1} \leq C \TV(u),
+%	\]
+%	where $u_\Omega$ is the mean value of $u$ over $\Omega$. 
+%\end{proposition}
+
+\begin{theorem}
+	Let $\XX = L^1(\Omega)$, where $\Omega \subseteq \RR^n$ is bounded Lipschitz, and $\YY$ a Banach space.
+Let $A \in \BB(L^1, \YY)$ such that $A(1) \neq 0$.
+Then $u_\delta$ converges strongly in $L^1$ (along a subsequence) to a TV-minimising solution as $\delta \to 0$ if $\alpha(\delta)$ is chosen as required by \cref{thm:minimiser_convergence}.
+\end{theorem}
+
+\begin{proof}
+	We need to check that the conditions of  \cref{thm:minimiser_convergence} are satisfied.
+We have assumed that $\Ran(A)$ is dense, so $\inf_{v\in \XX} \norm{Av - f} = 0$.
+Since $\BV(\Omega)$ is compactly embedded in $L^1(\Omega)$, we know that sequences with a finite TV-value have (strongly) convergent subsequences.
+We have checked that $\TV$ is convex and LSC, that $\dim(\Nul(\TV)) < \infty$, and by assumption we have $\Nul(A) \cap \Nul(\TV) = \qty{0}$.
+Finally, $\TV$ is absolutely one-homogeneous, so all assumptions are valid.
+\end{proof}

### 2020-11-10 inverse_problems/C4.tex
I'm still alive

\section{Convex Duality}
We are looking for a convergence rate
\[
D_\JJ^\Rm{symm}(u_\delta, u_\JJ\D) \leq \psi(\delta) \qquad \psi \colon \RR_+ \to \RR_+, \quad \lim_{\delta\to0}\psi(\delta) = 0. 
\]

We will assume for simplicity that $\JJ$ is absolutely one-homogeneous.  We define
\[
E(y) \ceq \frac12 \norm{y - f}_\YY^2, \quad F(u) \ceq \alpha \JJ(u), 
\]
so that $u_\delta \in \argmin_{u \in \XX} E(u) + F(Au)$.

\begin{lemma}
	Let $X$ be a Banach space and $\phi(x) = \frac12\norm{x}_X^2$. Then the convex conjugate of $\phi$ is
	\[
	\phi^*(\xi) = \frac12 \norm{\xi}_{X^*}^2. 
	\]
\end{lemma}

\begin{proof}
	We have
	\[
	\phi^*(\xi) = \sup_{x \in X} \ang{\xi, x} - \frac12 \norm{x}^2 \leq \sup_{x \in X} \norm{x} \norm{\xi} - \frac12 \norm{x}^2.
	\]
	The right-hand is a quadratic function in $\norm{x}$ with maximum $\frac12\norm{\xi}^2$. The inequality is in fact equality (\TODO show) which proves the result. 
\end{proof}

Now we compute
\[
E^*(\eta) = \ang{\eta, f} + \frac12 \norm{\eta}_{\YY^*}, \quad F^*(p) = \chi_{\pt \JJ(0)} (p/\alpha), 
\]
which gives the dual problem
\[
\sup_{\eta \in \YY^*} - \ang{\eta, f} - \frac12\norm{\eta}^2 - \chi_{\pt \JJ(0)} \qty(-\frac{A^*\eta}{\alpha}),
\]
which is equivalent to
\[
\sup_{\mu \in \YY^*, A^*\mu \in \pt \JJ(0)} \alpha \qty(\ang{\mu, f} - \frac\alpha2 \norm{\mu}^2)
\]

We need to

### 2020-11-10 inverse_problems/inverse_ES2.tex
I'm still alive

--- 2020-11-01 inverse_problems/inverse_ES1_handin.tex
@@
 \newcommand{\EE}{\Bb E}
-\newcommand\ZZ{\Bb Z}
-\newcommand\FF{\Cal F}
+\newcommand\XX{\Cal X}
+\newcommand\YY{\Cal Y}
+\newcommand\UU{\Cal U}
+\newcommand\VV{\Cal V}
+\newcommand\LL{\Cal L}
+\newcommand\KK{\Cal K}
+\newcommand\BB{\Cal B}
+\newcommand\JJ{\Cal J}
+\newcommand\LLL{\Bb L}
+\newcommand\DD{\Cal D}
 
@@
 \DeclareMathOperator*{\argmax}{arg\,max}
-
-\newcommand\XX{\Cal X}
-\newcommand\YY{\Cal Y}
-\newcommand\UU{\Cal U}
-\newcommand\VV{\Cal V}
-\newcommand\LL{\Cal L}
-\newcommand\KK{\Cal K}
-\newcommand\BB{\Cal B}
-% \newcommand\LLL{\Bb L}
 \DeclareMathOperator\Dom{\Cal D}
@@
 \DeclareMathOperator\Ran{\Cal R}
-\DeclareMathOperator\diag{diag}
+\DeclareMathOperator\TV{TV}
+\DeclareMathOperator\BV{BV}
 
-\title{Inverse Problems --- Example Sheet 1} % subject
+\newcommand\cw\rightharpoonup
+\newcommand\cws{\overset{*}{\rightharpoonup}}
+\newcommand\rext{\overline\RR}
+\newcommand\pt\partial
+\DeclareMathOperator\dom{dom}
+
+\title{Inverse Problems --- Example Sheet 2} % subject
 \author{Lucas Riedstra}
-\date{2 November 2020} % date
+\date{18 November 2020} % date
 
@@
 
-\begin{mdframed}
-		Note: when writing a norm of a vector $v \in V$, I will simply write $\norm{v}$ and not $\norm{v}_V$, unless it is unclear in which space $v$ lives. The same holds for inner products. 	
-\end{mdframed}
 \begin{question}
-	For $\Omega = [0, 1]^2$ and $\XX \in L^2(\Omega)$, we consider the integral operator $A \colon \XX \to \XX$ with
+	Let $\UU$ be a Banach space and $J \colon \UU\to \rext$ a functional. We define the \emph{subdifferential} of $J$ at any $v \in \UU$ as
+	\begin{equation} \nonumber %\label{def:subdifferentiable}
+	\pt\JJ(v) \ceq \qty{p \in \UU^* \mid J(u) \geq J(v) + \ang{p, u - v} \text{ for all $u \in \UU$}}. 
+	\end{equation}
+	Characterise the subdifferential for the
+	\begin{enumerate}[(a)]
+		\item absolute value function: $\UU= \RR$, $J(v) = \abs{v}$,
+		\item $\ell^1$-norm: $\UU = \ell^2$,
 		\[
-	(Au)(y) \ceq \int_\Omega k(x, y) u(x) \dd{x},
+		J(u) = \norm{u}_{\ell^1} \ceq \begin{cases}
+			\sum_{j=1}^\infty \abs{u_j} &\text{if $u \in \ell^1$}; \\ \infty &\text{else.}
+		\end{cases}
 		\]
-	for $k \in L^2(\Omega \times \Omega)$.  Show that
-	\begin{enumerate}[(a)]
-		\item $A$ is linear with respect to $u$,
-		\item $A$ is a bounded linear operator, i.e.\ $\norm{Au}_\XX \leq \norm{A}_{\LL(\XX, \XX)} \norm{u}_\XX$. Give also an estimate for $\norm{A}_{\LL(\XX, \XX)}$,
-		\item the adjoint $A^*$ is given via
+		\item characteristic function of the unit ball in $\RR$: $\UU = \RR$, $J(u) = \chi_C(u)$, $C \ceq \qty{ u\in \RR : \abs{u} \leq 1}$. 
+		\item Total Variation $\TV \colon L^1(\Omega) \to \rext$, where $\Omega \subset \RR^n$ is bounded Lipschitz
 		\[
-		(A^* v)(y) = \int_\Omega k(y, x) v(x) \dd{x}.
+		\TV(u) = \sup_{\phi \in \DD} \ang{u, \div\phi}, \quad \DD = \qty{\phi \in \Cal C_0^\infty(\Omega; \RR^n) : \norm{\phi(x)}_2 \leq 1 \ \forall x \in \Omega}. 
 		\]
-		\item $A$ is a compact operator, i.e.\ $A \in \KK(\XX, \XX)$.
-
-		\emph{Hint:} you may use the fact that if an operator $A$ can be written as a limit (in the operator norm) of finite-rank operators then $A$ is compact.
-An operator $B$ is called finite-rank if $\dim(B) < \infty$.
 	\end{enumerate}
@@
 \begin{solution}
+	Note: the spaces $\UU$ in parts (a) to (c) are Hilbert spaces, which means we can identify $\UU^*$ with $\UU$ (since any functional in $\UU^*$ is of the form $\ang{u, \cdot}$ for some $u \in \UU$). 
+\begin{enumerate}[(a)]
+\item Let $v \in \RR$. We know that $\abs{\cdot}$ is differentiable at $v \neq 0$, so 
+\[
+v > 0 \implies \pt J(v) = \qty{1} \quad\text{and}\quad v < 0 \implies \pt J(v) = \qty{-1}. 
+\]
+For $v = 0$ we have
+\begin{align*}
+	p \in \pt J(v) &\iff \abs{u} \geq p \cdot u  \text{ for all $u \in \RR$} \\
+	&\iff p \in [-1, 1],
+\end{align*}
+so $\pt J(0) = [-1, 1]$.
 
-	\begin{enumerate}[(a)]
-		\item Let $\alpha, \beta \in \RR$, $u, v \in L^2(\Omega)$ and $y \in \Omega$. Then we have
-		\begin{align*}
-			(A(\alpha u + \beta v))(y) &= \int_\Omega k(x, y) (\alpha u + \beta v)(x) \dd{x} \\
-			&= \int_\Omega k(x, y) \qty(\alpha u(x) + \beta v(x)) \dd{x}\\
-			&= \alpha \int_\Omega k(x, y) u(x) \dd{x} + \beta \int_\Omega k(x, y) v(x) \dd{x} \\
-			&= (\alpha A u)(y) + (\beta Av)(y) = (\alpha Au + \beta Av)(y). 
-		\end{align*}
-	Since equality holds for all $y \in \Omega$ we find $A(\alpha u + \beta v) = \alpha A u + \beta A v$, which proves that $A$ is linear. 
-	\item Let $u \in L^2(\Omega)$, then we have
-	\begin{align*}
-	\norm{Au}^2 &= \int_\Omega ((Au)(y))^2 \dd{y} = \int_\Omega \qty( \int_\Omega k(x, y) u(x) \dd{x})^2 \dd{y} = \int_\Omega \ang{k(\cdot, y), u(\cdot)}^2 \dd{y}. 
-	\end{align*}
-Now we apply Cauchy-Schwarz and find
-\begin{align*}
-	\int_\Omega \ang{k(\cdot, y), u(\cdot)}^2 \dd{y} &\leq \int_\Omega \norm{k(\cdot, y)}^2 \norm{u}^2 \dd{y} \overset\star= \norm{u}^2 \iint_{\Omega^2} k^2(x, y) \dd{x}\dd{y} = \norm{u}^2 \norm{k}^2, 
-\end{align*}
-where $\star$ follows from Fubini's theorem since the integrand is nonnegative. 
-Taking square roots on both sides we find that $\norm{Au}\leq \norm{k}\norm{u}$, so $A$ is bounded with $\norm{A} \leq \norm{k}$.
+\item Let $v \in \ell^2$. Firstly, if $v \notin \ell^1 = \dom(J)$, then we have $\pt J(v) = \emptyset$. Assume now that $v \in \ell_1 \cap \ell_2$. Then we have, for $p \in \ell^2$, that
+\begin{align}
+	p \in \pt J(v) &\iff \norm{u}_{\ell^1} \geq \norm{v}_{\ell^1} + \ang{p, u - v} &\text{for all $u \in \ell^2$}  \nonumber \\
+	&\iff \norm{u}_{\ell^1} - \norm{v}_{\ell^1} - \ang{p, u- v} \geq 0 &\text{for all $u \in \ell^2$} \nonumber \\
+	&\iff \sum_{j=1}^\infty \abs{u_i} - \abs{v_i} - p_i(u_i - v_i) \geq 0 &\text{for all $u \in \ell^2$} \label{subdiff1} \\
+	&\overset\star\iff \abs{x} - \abs{v_i} - p_i(x - v_i) \geq 0 &\text{for all $i \in \NN$ and $x \in \RR$}. \label{subdiff2}  
+\end{align}
+We first prove the bi-implication $\star$.
+If \eqref{subdiff2} holds, it is clear that \eqref{subdiff1} holds.
+If \eqref{subdiff2} does not hold, then we can find $x, i$ such that $\abs{x} - \abs{v_i} - p_i(x - v_i) < 0$.
+By now letting $u = x e_i$ in \eqref{subdiff1} we find  that \eqref{subdiff1} does not hold.
 
-\item We know that the adjoint is the unique operator that satisfies $\ang{Au, v} = \ang{u, A^*v}$ for all $u, v \in \XX$. Let $u, v \in \XX$, then we compute
-\begin{align*}
-	\ang{Au, v} &= \int_\Omega (Au)(y)\cdot  v(y) \dd{y} = \int_\Omega \qty(\int_\Omega k(x, y) u(x) \dd{x}) v(y) \dd{y} \\
-	&= \int_\Omega\int_\Omega k(x, y) u(x) v(y) \dd{x} \dd{y} \overset\star= \int_\Omega\int_\Omega k(x, y) u(x) v(y) \dd{y} \dd{x} \\
-	&= \int_\Omega u(x) \qty(\int_\Omega k(x, y) v(y) \dd{y}) \dd{x} = \ang{u, A^*v}
-\end{align*}
-where $(A^* v)(x) = \int_\Omega k(x, y) v(y) \dd{y}$ as required. Here $\star$ follows from Fubini's theorem.
-
-\item It is known that for any compact set $X \subseteq \RR^n$, polynomials lie dense in $L^2(X)$.
-Therefore, there exists a sequence of polynomials $p_n$ such that $p_n \to k$ in $L^2([0, 1]^4)$.
-It is easily seen that for any polynomial $p$, the operator
+However, if we define $H(x) \ceq \abs{x}$,  we see that \cref{subdiff2} is equivalent to $p_i \in \pt H(v_i)$ for all $i$. Therefore, by (a) we have
 \[
-(A_p u)(y) \ceq \int_\Omega p(x, y) u(x) \dd{x} 
-\] 
-has finite rank: let $p(z) = \sum_{\abs{\alpha} \leq n} c_\alpha z^\alpha$ (where $z \in [0, 1]^4$ and $\alpha$ is a multi-index), then we find
-\[
-(A_p u)(y) = \sum_{\abs{\alpha} \leq n} c_\alpha \int_\Omega x_1^{\alpha_1} x_2^{\alpha_2} y_1^{\alpha_3} y_2^{\alpha_4} u(x) \dd{x} = \sum_{\abs{\alpha} \leq n} c_\alpha \qty(\int_\Omega x_1^{\alpha_1} x_2^{\alpha_2} u(x) \dd{x}) y_1^{\alpha_3} y_2^{\alpha_4},
-\]
-so $A_p u$ lies in the $\Span\qty{y_1^{\alpha_1} y_2^{\alpha_2} \mid \alpha_1 + \alpha_2 \leq n}$, and therefore has finite rank. 
-By (b), we find that $\norm{A - A_n} \leq \norm{k - p_n} \to 0$, which shows that $A_n \to A$ in operator norm. We conclude that $A$ is compact. 
-%\[
-%(A_n u)(y) \ceq \int_\Omega p_n()
-%\]
-	\end{enumerate}
-\end{solution}
-
-\begin{question}
-	We consider the problem of differentiation, formulated as the inverse problem of finding $u$ from $Au = f$ with the integral operator $A \colon L^2([0, 1]) \to L^2([0, 1])$ defined as
-	\[
-	(Au)(y) \ceq \int_0^y u(x) \dd{x}.
+\pt J(v) = \qty{p \in \ell^2 \mid p_i = \sign(v_i) \text{ if $v_i \neq 0$ and } p_i \in [-1, 1] \text{ for all $i$}}. 
 \]
 
-	\begin{enumerate}[(a)]
-		\item Let $f$ be given by
+\item Clearly, if $\abs{v} < 1$, then $\chi_C$ is differentiable with derivative 0 so $\pt J(v) = \qty{0}$. If $\abs{v} > 1$, then $v \notin \dom(J)$, and therefore $\pt J(v) = \emptyset$.
+
+Consider the point $v = 1$, then we have
 \[
-		f(x) \ceq \begin{cases} 0 & x < \frac12, \\ 1 &x > \frac12. \end{cases}
+p \in \pt J(\chi_C) \iff \chi_C(u) \geq p \cdot (u - 1) \ \forall u.
 \]
-		Show that $f \in \clos{\Ran(A)}$. 
-		\item Let $f$ be given as in a). Show that $f \in \clos{\Ran(A)} \setminus \Ran(A)$. \emph{Hint: Consider the Picard criterion}.
-
-		\item Prove or falsify: ``The Moore-Penrose inverse of $A$ is continuous.''
-	\end{enumerate}
-\end{question}
-
-\begin{solution}
-	\begin{enumerate}[(a)]
-		\item We want to show that we can approximate $f$ by a sequence $(Au_n)$ for some $(u_n) \subseteq L^2[0,1]$. To this end, define for $n \geq 2$
+For $u > 1$, this equation is satisfied regardless of $p$. Therefore, the above equation is equivalent to
 \[
-		u_n(x) =
-		\begin{cases}
-			0 &\abs{x - \frac12} > \frac1n, \\
-			\frac n2 &\abs{x - \frac12} \leq \frac1n. 
-		\end{cases}
+p \cdot(u-1) \leq 0 \ \forall u \leq 1, 
 \]
-		Clearly $u \in L^2[0, 1]$, and we have
+which is satisfied for all $p \geq 0$, so we conclude $\pt J(1) = [0, \infty)$. 
+Analogously, we find $\pt J(-1) = (-\infty, 0]$. We conclude that
 \[
-		f_n(y) \ceq (Au_n)(y) = \int_0^y u_n(x) \dd{x} = \begin{cases}
-			0 &\text{if $y < \frac12 - \frac1n$,}\\
-			\frac n2(y - \frac12 + \frac1n) &\text{if $\frac12 - \frac1n \leq y \leq \frac12 + \frac1n$} \\
-			1 &\text{if $y > \frac12 + \frac1n$}. 
+\pt J(v) = \begin{cases}
+	\emptyset &\text{if $\abs{v} > 1$}; \\
+	(-\infty, 0] &\text{if $v = -1$}; \\
+	\qty{0} &\text{if $v \in (-1, 1)$}; \\
+	[0, \infty) &\text{if $v = 1$}. 
 \end{cases}
@@
 
-		Therefore we find
-		\begin{align*}
-			\norm{f_n - f}^2 &= \int_0^1 (f_n - f)^2(x) \dd{x} \\
-			&= 2 \int_{\frac12 - \frac1n}^{\frac12} \frac{n^2}{4} (x - \frac12 - \frac1n)^2 \dd{x} \\
-			&= \frac{n^2}{2} \int_0^{1/n}  x^2 \dd{x} = \frac{1}{6n} \to 0, 
-		\end{align*}
-	so $f_n \to f$ in $L^2[0, 1]$. Since $f_n \in \Ran(A)$ this shows $f \in \clos{\Ran(A)}$.
-
-	\item 	In example 2.2.12, it is shown that for this operator, the Picard criterion is
-	\begin{equation} \label{eq:picard}
-	2\sum_{j=1}^\infty  \sigma_j^{-2} \qty(\int_0^1 f(s) \sin(\sigma_j^{-1} s) \dd{s})^2 < \infty,
-	\end{equation}
-	where $\sigma_j = \frac2{(2j - 1)\pi}$.
-
-	We compute
-	\[
-	\int_0^1 f(s) \sin(\sigma_j^{-1} s) \dd{s} = \int_{1/2}^1 \sin(\sigma_j^{-1} s) \dd{s} = \sigma_j \qty[ \cos(\frac12 \sigma_j^{-1}) - \cos(\sigma_j^{-1})]. 
-	\]
-	We have
-	\[
-	\cos(\sigma_j^{-1}) = \cos(\frac{(2j - 1)\pi}{2}) = 0 \quad\text{and}\quad \cos(\frac12\sigma_j^{-1}) = \cos(\frac{(2j-1)\pi}{4}) = \pm\frac1{\sqrt2 }. 
-	\]
-
-	Plugging this into \cref{eq:picard} gives that
-	\[
-	2\sum_{j=1}^\infty  \sigma_j^{-2} \qty(\int_0^1 f(s) \sin(\sigma_j^{-1} s) \dd{s})^2 = 2 \sum_{j=1}^\infty \sigma_j^{-2} (\sigma_j^2/2) = \sum_{j=1}^\infty 1 = \infty,
-	\]
-	so $f$ does not satisfy the Picard criterion and therefore $f \in \clos{\Ran(A)} \setminus \Ran(A)$. 
-%	Note that $A$ is compact, since we can write $(Au)(y) = \int_{[0, 1]} k(x, y) u(x) \dd{x}$ with $k(x, y) = \ind_{x \leq y}$. 
-%	
-%	To apply the Picard criterion we must find the singular values and right singular vectors of $A$, which are equal to the square roots of the eigenvalues of $AA^*$ and the eigenvectors of $AA^*$. 
-%	
-%
-%	\begin{align*}
-%		\ang{Au, v} &= \int_0^1 (Au)(y)\cdot v(y) \dd{y} \\
-%		&= \int_0^1 \int_0^y u(x) \dd{x} v(y) \dd{y} \\
-%		&= \int_0^1 \int_0^y u(x) v(y) \dd{x} \dd{y} \\
-%		&= \int_0^1 \int_x^1 u(x) v(y) \dd{y} \dd{x} \\ 
-%		&= \int_0^1 u(x) \int_x^1 v(y) \dd{y} \dd{x} \\
-%		&= \ang{u, A^* v}
-%	\end{align*}
-%where $v(x) = \int_x^1 v(y) \dd{y}$.  Therefore, we find
-%\begin{align*}
-%	(AA^* u)(x) = \int_0^x \int_y^1 u(z) \dd{z} \dd{y}.
-%\end{align*}
-%Now we solve the eigenequation. Firstly, note that $\Nul(A) = \qty{0}$, so let $\lambda^2 > 0$. Then we have
-%\begin{align}
-%	\int_0^x \int_y^1 u(z) \dd{z} \dd{y} &= \lambda^2 u(x) \\
-%	\int_x^1 u(z) \dd{z} &= \lambda^2 u'(x) \\
-%	- u(x) &= \lambda^2 u''(x). 
-%\end{align}
-%Furthermore, from (1) we infer $u(0) = 0$ while from (2) we infer $u'(1) = 0$. 
-%The general solution to (3) is given by $u(x) = a \cos(\lambda^{-1} x) + b \sin(\lambda^{-1} x)$. 
-%Plugging $u(0) = 0$ gives $a = 0$, and $u'(1) = 0$ gives
-%\[
-%b\lambda^{-1} \cos(\lambda^{-1}) = 0 \implies b = 0 \text{ or } \lambda^{-1} = \pi\qty(n - \frac12) \text{ for some $n \in \ZZ_{>0}$}.
-%\]
-%
-%Our singular values are therefore $\frac{1}{(n - \frac12)\pi}$ ($n \in \ZZ_{>0}$) with right singular vectors $y_n(x) = \sqrt 2\sin((n - \frac12) \pi x)$. 
-%We now compute
-%\[
-%\ang{f, y_n} = \sqrt 2\int_{1/2}^1 \sin(\qty(n - \frac12) \pi x) \dd{x} = ...
-%\]
-
-\item The Moore-Penrose inverse of $A$ is discontinuous.
-This can be seen by theorem 2.1.11: we have in (b) an element $f \in \clos{\Ran(A)} \setminus \Ran(A)$, so   $\Ran(A)$ is not closed, so $A\D$ is discontinuous.
-
+\item 
 \end{enumerate}
 \end{solution}
-
 \end{document}

### 2020-11-10 modern_statistical_methods/C2.tex
I'm still alive

@@
 \begin{theorem}[Fast rate] \label{thm:lasso_fast_rate}
-	Let the conditions of \cref{thm:lasso_slow_rate} be satisfied and assume furthermore that the compatibility condition holds. 
+	Let the conditions of \cref{thm:lasso_slow_rate} be satisfied and assume furthermore that the compatibility condition holds. Then with probability at least $1 - 2p^{-(A^2/8 - 1)}$ we have
+	\[
+	\frac1n \norm{X(\beta^0 - \hat\beta)}_2^2 + \lambda \norm{\hat\beta - \beta^0}_1 \leq \frac{16\lambda^2s}{\phi^2} = \frac{16A^2 \log(p) \sigma^2 s}{\phi^2n }. 
+	\]
 \end{theorem}
+
+\begin{proof}
+	As in the proof of \cref{thm:lasso_slow_rate}, we have
+	\[
+	\frac1{2n} \norm{X(\beta^0 - \hat\beta)}_2^2 + \lambda \norm{\hat\beta}_1 \leq \frac1n \eps\T X(\hat\beta - \beta^0) + \lambda \norm{\beta^0}_1. 
+	\]
+	Define the event $\Omega = \qty{2 \norm{X\T \eps}_\infty/n \leq \lambda}$, then by H\"older's inequality we get, conditional on $\Omega$,
+	\[
+	\frac1n \norm{X(\hat\beta - \beta^0)}_2^2 + 2\lambda \norm{\hat\beta}_1  \leq \lambda \norm{\hat\beta - \beta^0}_1 + 2\lambda \norm{\beta^0}_1. 
+	\]
+	By \cref{lem:probability_omega_bound}, we find that $\PP(\Omega) \geq 1 - 2p^{-(A^2/8 - 1)}$.
+
+	Dividing by $\lambda$, we have
+	\begin{align*}
+		\frac1{\lambda n} \norm{X(\hat\beta - \beta^0)}_2^2 + 2 \norm{\hat\beta}_1 &\leq \norm{\hat\beta - \beta^0}_1 + 2 \norm{\beta^0}_1 \\
+		\frac1{\lambda n} \norm{X(\hat\beta - \beta^0)}_2^2 + 2 (\norm{\hat\beta_S}_1 + \norm{\hat\beta_N}_1) &\leq \norm{\hat\beta_S - \beta^0_S}_1 + \norm{\hat\beta_N}_1 + 2 \norm{\beta^0_S}_1\\
+		\frac1{\lambda n} \norm{X(\hat\beta - \beta^0)}_2^2 +  \norm{\hat\beta_N}_1 &\leq \norm{\hat\beta_S - \beta^0_S}_1 - 2 \norm{\hat\beta_S}_1 + 2 \norm{\beta^0_S}_1 \\
+		&\leq 3 \norm{\hat\beta_S - \beta^0_S}_1. 
+	\end{align*}
+	Now we note that $\norm{\hat\beta_N}_1 = \norm{\hat\beta_N + \beta^0_N}_1$, and we add $\norm{\hat\beta_S - \beta^0_S}_1$ to both sides of the inequality to get
+	\begin{equation} \label{eq:error_inequality}
+	\frac1{\lambda n} \norm{X(\hat\beta - \beta^0)}_2^2 + \norm{\hat\beta - \beta^0}_1  \leq 4 \norm{\beta_S - \beta^0_S}_1. 
+	\end{equation}
+	Assume that $\norm{\hat\beta - \beta^0} \neq 0$ (otherwise the theorem is definitely true), and note that $\norm{(\hat\beta - \beta^0)_N}_1 \leq 3 \norm{\hat\beta}_1 \leq 3 \norm{(\hat \beta - \beta^0)_S}$ by our previous inequality, we find that the compatibility constant satisfies
+	\[
+	\phi^2 \leq \frac{\frac1n \norm{X(\hat\beta - \beta^0)}_2^2}{\frac1s \norm{\hat\beta - \beta^0}_1^2} \implies \frac1{\phi} \geq \sqrt{\frac ns} \frac{\norm{\hat\beta - \beta^0}_1}{\norm{X(\hat\beta - \beta^0)}_2},
+	\]
+	and plugging this into \cref{eq:error_inequality} we get
+	\begin{equation} \label{eq:error_inequality_2}
+	\frac1{\lambda n} \norm{X(\hat\beta - \beta^0)}_2^2 + \norm{\hat\beta - \beta^0}_1  \leq 4 \norm{\beta_S - \beta^0_S}_1 \leq \frac4\phi \sqrt{\frac sn} \norm{X(\hat\beta - \beta^0)}_2,
+	\end{equation}
+	or equivalently
+	\[
+	\frac1{\lambda n} \norm{X(\hat\beta - \beta^0)}_2^2 - \frac4\phi \sqrt{\frac sn} \norm{X(\hat\beta - \beta^0)}_2 + \norm{\hat\beta - \beta^0}_1 \leq 0. 
+	\]
+	This is a quadratic inequality in $\norm{X(\hat\beta - \beta^0)}_2$, and solving it yields
+	\[
+	\frac1{\sqrt n} \norm{X(\hat\beta - \beta^0)}_2 \leq \frac{4\lambda \sqrt s}{\phi}. 
+	\] 
+	Plugging that into the right hand side of \cref{eq:error_inequality_2} yields the required result. 
+\end{proof}

### 2020-11-10 modern_statistical_methods/msm_ES2.tex
I'm still alive

@@
 % Functions that appear frequently
-\DeclareMathOperator{\sign}{sign}
+\DeclareMathOperator{\sign}{sgn}
 \DeclareMathOperator{\Int}{Int}
@@
 \begin{proof}
-	By the previous question, all Lasso solutions have the same norm $x$, so the set of Lasso solutions is given by
+	Let $v_1, \dotsc, v_p \subseteq \RR^n$ be the columns of $X$.
+Because $X$ has centred columns, we know that $v_1, \dotsc, v_p$ lie in the $(n-1)$-dimensional subspace of vectors in $\RR^n$ with mean zero.
+
+	Now, let $\hat\beta$ be any lasso solution. If $\hat\beta = 0$, we are done, so suppose $\hat\beta \neq 0$, and write
 	\[
-	\argmin_{\norm{\beta}_1 = x} \norm{Y - X\beta}_2^2.
+	X\hat\beta = \mqty[ v_1 & \dotsb & v_p] \mqty[ \hat\beta_1 \\ \vdots \\ \hat\beta_p] = \norm{\hat\beta}_1 \mqty[\sign(\hat\beta_1) v_1 & \dotsb & \sign(\hat\beta_p) v_p] \mqty[ \abs{\hat\beta_1} / \norm{\hat\beta}_1\\ \vdots \\ \abs{\hat\beta_p}/\norm{\hat\beta}_1].
 	\]
-	Let $\hat\beta$ be any fixed Lasso solution, then we find that the entire solution set can also be described as
+	If we define $S = \qty{\norm{\hat\beta}_1 \sign(\hat\beta_1)v_1, \dotsc, \norm{\hat\beta}_1 \sign(\hat\beta_p)v_p}$, then we see that $X\hat\beta$ can be written as a convex combination of points in $S$.
+Since the points in $S$ also all have mean zero, by Caratheodory's lemma there exists a vector $\gamma \in \RR^p$ with at most $n$ nonzero coefficients, nonnegative entries, and $\norm{\gamma}_1 \geq 0$ such that
 	\[
-	\qty{\hat\gamma \mid \norm{\gamma} = x, \hat\beta - \hat\gamma \in \Nul(X)}. 
+	X\hat\beta = \norm{\hat\beta}_1 \mqty[\sign(\hat\beta_1) v_1 & \dotsb & \sign(\hat\beta_p) v_p] \gamma = X \qty(\norm{\hat\beta}_1 \mqty[\sign(\hat\beta_1) \gamma_1 \\ \vdots \\ \sign(\hat\beta_p) \gamma_p]) \eqqcolon X\hat\beta'.
 	\]
-	??
+	Now, we have $X\hat\beta = X\hat\beta'$ and $\norm{\hat\beta}'_1 = \norm{\hat\beta}_1$, which shows that $\hat\beta'$ is a Lasso solution with no more than $n$ nonzero coefficients. 
 \end{proof}
@@
 \begin{proof}
-	By the KKT conditions any lasso solution satisfies $\frac1n X\T (Y - X\hat\beta) = \frac1n X\T Y - \frac1n X\T X\hat\beta= \lambda \hat\nu$ for some $\nu$ with $\norm{\hat\nu}_\infty \leq 1$ and $\hat\nu_{\hat S_{\lambda}} = \sign(\hat\beta^L_{\lambda, \hat S_\lambda})$.
+	We first show that 0 is a lasso solution: we have
+	\[
+	\frac1n X\T (Y - X0) = \frac1n X\T Y = \lambda \cdot \frac{\frac1n X\T Y}{\lambda} \eqqcolon \lambda\hat\nu, 
+	\]
+	and since 
+	\[
+	\norm{\hat\nu}_\infty = \frac{\norm{X\T Y}_\infty / n}{\lambda} \leq \frac{\norm{X\T Y}_\infty / n}{\norm{X\T Y}_\infty /n} \leq 1
+	\]
+	it follows that 0 is a Lasso solution. By question 9, it follows that any other Lasso solution must also have norm 0, and therefore 0 is the only Lasso solution. 
+%	By the KKT conditions any lasso solution satisfies $\frac1n X\T (Y - X\hat\beta) = \frac1n X\T Y - \frac1n X\T X\hat\beta= \lambda \hat\nu$ for some $\nu$ with $\norm{\hat\nu}_\infty \leq 1$ and $\hat\nu_{\hat S_{\lambda}} = \sign(\hat\beta^L_{\lambda, \hat S_\lambda})$.
+%	
+%	Suppose $\lambda \geq \lambda_\Rm{max}$, and suppose also that $\hat\beta \neq 0$, so $\norm{\hat\nu}_\infty = 1$. Then we find
+%	\[
+%	\frac1n \norm{X\T (Y - X\hat\beta)}_\infty = \norm{\lambda\hat\nu} = \lambda  \geq \frac1n \norm{X\T Y}_\infty, 
+%	\]
+%	so 
+%	\[
+%	\norm{X\T Y - X\T X\hat\beta}_\infty \geq \norm{X\T Y}_\infty. 
+%	\]
+%	??
+\end{proof}
 
-	Suppose $\lambda \geq \lambda_\Rm{max}$, and suppose also that $\hat\beta \neq 0$, so $\norm{\hat\nu}_\infty = 1$. Then we find
+\begin{question}
+	Show that when the columns of $X$ are orthogonal (so necessarily $p \leq n$) and scaled to have $\ell_2$-norm $\sqrt n$, the $k$-th component of the Lasso estimator is given by
 	\[
-	\frac1n \norm{X\T (Y - X\hat\beta)}_\infty = \norm{\lambda\hat\nu} = \lambda  \geq \frac1n \norm{X\T Y}_\infty, 
+	\hat\beta^L_{\lambda, k} = (\abs{\hat\beta_k^\Rm{OLS}} - \lambda)_+ \sign(\hat\beta_k^\Rm{OLS}). 
 	\]
-	so 
+	What is the corresponding estimator if the $\ell_1$ penalty $\norm{\beta}_1$ in the Lasso objective is replaced by the $\ell_0$ penalty $\norm{\beta}_0 \ceq \#\qty{k \mid \beta_k \neq 0}$? 
+\end{question}
+
+\begin{proof}
+	For simplicity, let $\hat\beta$ denote the proposed Lasso estimator and $\tilde\beta$ the OLS estimator.
+Since the columns of $X$ are orthogonal, $X$ is injective, so the Lasso estimator is unique (since $X\hat\beta_\lambda^L$ is unique).
+
+	Now, define $\gamma \in \RR^p$ by 
 	\[
-	\norm{X\T Y - X\T X\hat\beta}_\infty \geq \norm{X\T Y}_\infty. 
+	\gamma_i = \begin{cases}
+		\tilde\beta_i, &\abs{\tilde\beta_i}\leq \lambda, \\ \lambda, &\abs{\tilde\beta_i} > \lambda, \\ -\lambda, &\abs{\tilde\beta_i} < -\lambda
+	\end{cases},
 	\]
-	??
+	so that $\hat\beta= \tilde\beta - \gamma$. Then we find
+	\[
+	\frac1n X\T (Y- X\hat\beta) = \frac1n X\T(Y - X(\tilde\beta - \gamma)) = \frac1n X\T (Y - X\tilde\beta)+ \frac1n X\T X\gamma = \frac1n \gamma = \lambda \frac1{n\lambda}\gamma, 
+	\]
+	since $Y - X\tilde\beta$ is orthogonal to the columns of $X$ so $X\T(Y - X\tilde\beta) = 0$.
+
+	Now, we have $\norm{\frac1{n\lambda} \gamma}_\infty = \frac1{n\lambda} \norm{\gamma}_\infty \leq \frac1n \leq 1$, and furthermore, if $\hat\beta_i \neq 0$, then it is clear that $\sign(\hat\beta_i) = \sign(\gamma_i)$.
+This shows that $\hat\beta$ is indeed a Lasso solution.
+
+	Suppose we replace the Lasso objective by the $\ell_0$ penalty.
+Let $\qty{X_1, \dotsc, X_p}$ be the columns of $X$ and expand them to an orthonormal basis $\qty{X_1, \dotsc, X_n}$ of $\RR^n$.
+Then we can write
+	\[
+	\norm{Y - X\beta}_2^2 = \norm{\sum_{i=1}^n P_{X_i} Y - \sum_{i=1}^p \beta_i X_i}_2^2 = \sum_{i=1}^p \qty(\ang{X_i, Y} - \beta_i)^2  + \sum_{j = p+1}^n \ang{X_{j}, Y}^2. 
+	\]
+	Since the latter sum does not depend on $\beta$, our objective function is equivalent to minimising
+	\[
+	\sum_{i=1}^p \frac1{2n}\qty(\ang{X_i, Y} - \beta_i)^2 + \lambda \ind_{\beta_i \neq 0}. 
+	\]
+	Clearly, this is minimised by choosing $\beta_i = \tilde\beta_i$ if $\lambda \leq \frac1{2n} \ang{X_i, Y}^2$ and $\beta_i = 0$ otherwise (note, if $\lambda = \frac1{2n} \ang{X_i, Y}^2$, then choosing $\beta_i =0 $ or $\beta_i = \tilde\beta_i$ gives the same outcome).
+%	If we replace the Lasso objective by the $\ell_0$ penalty, note that with $P_{X_i}$ the projection onto $X_i$, we have
+%	\[
+%	\norm{Y - X\tilde\beta}_2^2 = \norm{Y}_2^2 - \sum_{i=1}^p \norm{P_{X_i}Y}_2^2,
+%	\]
+%	so we obtain the $\norm{\cdot}_0$ estimator by replacing $\tilde{\beta_i}$ by 0 if $\norm{P_{X_i} Y} > \lambda$. 
 	\end{proof}

### 2020-11-10 topics_statistical_theory/C2.tex
I'm still alive

@@
 Putting everything together, we obtain the following:
-\begin{theorem}
+\begin{theorem} \label{thm:mise_bound}
 	Fix $\beta, L > 0$, and let $K$ be a kernel of order $\ell = \ceil\beta$ with $R(k)$ and $\mu_\beta(K)$ finite.
@@
 The choice of bandwidth in the previous theorem is not practical since we have not specified $\alpha$ and $\beta$ is typically unknown.
-One possible approach is \emph{least squares cross validation}.
-For this, note that minimising the $\MISE$ is equivalent to minimising
+
+\subsubsection{Least squares cross validation}
+One possible approach is \emph{least squares cross validation}. For this, note that minimising the $\MISE$ is equivalent to minimising
 \[
@@
 where $\hat f_{n, -i}$ is the KDE based on all observations except $X_i$. We now choose $h$ such that LSCV is minimised.
+
+\subsubsection{Lepski}
+\TODO write this subsection (\TODO understand this first)
+
+\subsection{Choice of kernel}
+To choose a kernel, we first fix the scale of the kernel by setting $\mu_2(K) = 1$.
+Now, by our bound on the MISE (\cref{thm:mise_bound}) it is reasonable to minimise $R(K)$, where for simplicity we assume that $K$ is a nonnegative second-order kernel.
+The solution is the Epanechnikov kernel
+\[
+K_E(x) \ceq \frac{3}{4\sqrt 5} \qty(1 - \frac{x^2}{5}) \ind_{\abs{x} \leq \sqrt 5}, 
+\]
+and the ratio $R(K_E)/R(K)$ is called the \emph{efficiency} of a kernel $K$.
+We find that for different kernels, the efficiency is greater than 0.9, which shows that kernel shape does not affect efficiency greatly.
+
+\subsection{Multivariate density estimation}
+The general $d$-dimensional KDE is 
+\[
+\hat f_n(x) \ceq \frac{1}{n \sqrt{\det H}} \sum_{i=1}^n K(H^{-1/2}(x - X_i)),
+\]
+where $H$ is a symmetric positive-definite bandwidth bandwidth matrix (often chosen to be diagonal or a multiple of $I$).
+If $H = h^2 I$, it can be shown that, under an appropriate definition of a $\beta$ smoothness class, we have an optimal bandwidth of order $n^{-1/(d + 2\beta)}$, and with this choise, a MISE of order $n^{-2\beta/(d + 2\beta)}$.
+This is called the ``curse of dimensionality'': the higher the dimension becomes, the harder nonparametric estimation gets.

### 2020-11-10 topics_statistical_theory/C3.tex
I'm still alive

\section{Nonparametric regression}
\subsection{Fixed and random design}
In \emph{fixed design}, we assume we have data $x_1 \leq \dotsb \leq x_n$ and the response variables satisfy 
\[
Y_i \ceq m(x_i) + \sqrt{v(x_i)} \eps_i,
\]
where the $\eps_i$ are independent, mean-zero random variables with $\Var(\eps_i) = 1$. The function $m$ is called the \emph{regression function}, and the function $v$ is the \emph{variance function}. If $v$ is constant, the model is called \emph{homoscedastic}, else it is called \emph{heteroscedastic}.

In \emph{random design}, we assume we have i.i.d.\ data pairs $(X_i, Y_i)$ with 
\[
Y_i = m(X_i) + \sqrt{v(X_i)} \eps_i, 
\]
where the $\eps_i$ are again independent with $\EE[\eps_1 | X_1] = 0$ and $\Var(\eps_1 | X_1) = 1$. The regression function is given by $m(x) = \EE(Y_1 | X_1 = x)$ and the variance function by $v(x) = \Var(Y_1 | X_1 = x)$.

### 2020-11-11 distribution_theory/distribution_ES2.tex
I actually UNDERSTOOD topics

--- 2020-10-28 distribution_theory/distribution_ES1.tex
@@
 \newcommand{\NN}{\Bb N}
+\newcommand\ZZ{\Bb Z}
 \newcommand{\PP}{\Bb P}
@@
 \newcommand{\RR}{\Bb R}
-\newcommand{\EE}{\Cal E}
+\newcommand{\KK}{\Bb K}
 \newcommand{\DD}{\Cal D}
-\newcommand\ZZ{\Bb Z}
+\newcommand\EE{\Cal E}
+\renewcommand\SS{\Cal S}
+\newcommand\FF{\Cal F}
+\newcommand\vd{\vdot}
+%\newcommand\ZZ{\Bb Z}
 
@@
 \newcommand\refl\widecheck
+\newcommand\les\lesssim
+\newcommand\ges\gtrsim
+\newcommand\loc{_\Rm{loc}}
 
@@
 
-\newcommand\clos\overline
-
-\title{Distribution Theory and Applications --- Example Sheet 1} % subject
+\title{Distribution Theory --- Example Sheet 2} % subject
 \author{Lucas Riedstra}
-\date{...} % date
+\date{18 November 2020} % date
 
@@
 \maketitle
-
 \begin{question}
-	Construct a non-zero element of $\DD(\RR)$ that vanishes outside $(0, 1)$. Construct a non-zero of $\DD(\RR^n)$ that vanishes outside the ball $B_\eps = \qty{x \in \RR^n : \norm{x} < \eps}$. 
+Let $u, v \in \DD'(\RR^n)$, one of which has compact support. Show that the convolution $u * v$, defined as in your notes, is uniquely defined and gives rise to an element of $\DD'(\RR^n)$. 
 \end{question}
@@
 \begin{proof}
-	It is well-known that the function 
+	The convolution between $u, v \in \DD'(\RR^n)$ is defined by the formula
+	\begin{equation} \label{def:convolution}
+	(u * v) * \phi = u * (v * \phi) \quad\text{for all $\phi \in \DD(\RR^n)$}. 
+	\end{equation}
+	To show that this is uniquely defined, recall that for all $u \in \DD'(\RR^n) \phi\in \DD(\RR^n)$, we have $\ang{u, \phi} = (u * \refl\phi)(0)$. Therefore, we have
 	\[
-	\phi \colon \RR \to \RR \colon x \mapsto \begin{cases}
-		0 &\text{if $x \leq 0$}; \\
-		e^{-1/x} &\text{if $x > 0$},
-	\end{cases}
+	\ang{u * v, \phi} = ((u * v) * \refl\phi)(0) = (u * (v * \refl\phi))(0),
 	\]
-	is smooth and vanishes outside $(0, \infty)$. The function $\psi(x) \ceq \phi(x)\phi(1-x)$ is therefore also smooth and vanishes outside $(0, 1)$.
+	which shows that the formula \cref{def:convolution} uniquely defines $\ang{u * v, \phi}$ for any $\phi \in \DD(\RR^n)$, and therefore that $u * v$ is well-defined.
 
-	Since $\psi$ vanishes outside $(0, 1)$, the function $\psi(x/\eps)$ vanishes outside $(0, \eps)$, and therefore the function $\vec x \mapsto \psi(\norm{\vec x}/\eps)$ vanishes outside $B_\eps$. 
+	Now we prove that $u * v \in \DD'(\RR^n)$: by the previous equation we have
+	\[
+	\ang{u * v, \phi} = (u * (v * \refl\phi))(0) = \ang{u, \refl{v * \refl\phi}}.
+	\]
+
+	Suppose $u$ is compactly supported. Since $\refl{v * \refl\phi} \in \EE(\RR^n)$, there exists a compact $K \subseteq X$ and nonnegative $C, N$ such that
+	\begin{align*}
+		\abs{\ang{u, \refl{v * \refl\phi}}} &\leq C \sum_{\alpha \leq N} \sup_{x \in K} \abs{\pt^\alpha (\refl{v * \refl\phi})} = C \sum_{\alpha \leq N} \sup_{x \in -K} \abs{\pt^\alpha (v * \refl\phi)} = C \sum_{\alpha \leq N} \sup_{x \in -K} \abs{v * \pt^\alpha\refl\phi}\\
+		&= C \sum_{\abs{\alpha} \leq N} \sup_{x \in -K} \abs{\ang{v, \tau_x \refl{\pt^\alpha \refl\phi}}}.
+	\end{align*}
+	Note that if $\supp\phi \subseteq K'$, then $\supp\refl\phi \subseteq -K'$, and for $x \in -K$ we find $\supp \tau_x \refl{\pt^\alpha \refl\phi} \subseteq -K' -K$.
+Then by the previous equation we find that there exists $C', M$ with
+	\[
+	\abs{\ang{u * v, \phi}} \leq C' \sum_{\abs{\alpha} \leq N} \sum_{\abs{\beta} \leq M} \sup_{x \in -K' - K} \pt^\beta (\tau_x \refl{\pt^\alpha \refl\phi}) = C' \sum_{\abs{\alpha} \leq N} \sum_{\abs{\beta} \leq M} \sup_x \abs{\pt^{\alpha + \beta} \phi} \leq C'' \sum_{\abs{\alpha} \leq M + N} \sup_x \abs{\pt^\alpha \phi},
+	\]
+	which shows that $u* v \in \DD'(\RR^n)$. An analogous argument holds if $v$ is compactly supported. 
 \end{proof}
@@
 \begin{question}
-	Given $\phi \in \DD(X)$, Taylor's theorem gives
-	\[
-	\phi(x+h) = \sum_{\abs{\alpha} \leq N} \frac{h^\alpha}{\alpha!} \pt^\alpha \phi(x) + R_N(x, h).
-	\]
-	Prove that $\supp(R_N)$ is contained in some fixed compact $K \subseteq X$ for $\abs{h}$ sufficiently small.
-Show also that $\pt^\alpha R_N = o(\abs{h}^N)$ uniformly in $x$ for each multi-index $\alpha$, i.e.\ prove
-	\[
-	\lim_{\abs{h}\to 0} \frac{\sup_x \abs{\pt^\alpha R_N(x, h)}}{\abs{h}^N} = 0
-	\]
-	for each multi-index $\alpha$.
-
-	\emph{Hint: you may find it convenient to use the following form of the remainder}
-	\[
-	R_N(x, h) = \sum_{\abs{\alpha} = N+1} \frac{h^\alpha}{\alpha!}(N+1)\int_0^1 (1-t)^N (\pt^\alpha\phi)(x + th)\dd{t},
-	\]
-	\emph{and note that} $(N+1)! \sum_{\abs{\alpha} = N+1} \frac{h^\alpha}{\alpha!} = (h_1 + \dotsb + h_n)^{N+1}$. 
+	Show that if $u, v, w \in \DD'(\RR^n)$ and at least two of them have compact support, then the convolution is associative (i.e., $(u * v) * w) = u* (v * w)$). 
 \end{question}
@@
 \begin{proof}
-	Since $\phi \in \DD(X)$, we know that $\supp\phi \subseteq \clos{B_N}$ for some $N \in \NN$. Now, suppose $\norm{h} < 1$, then
-	\[
-	\phi(x+h) \neq 0 \implies \norm{x + h} \leq N \implies \norm{x} \leq \norm{x + h} + \norm{h} \leq N + 1, 
-	\]
-	so if we define $\psi_h(x) = \phi(x+h)$ then we know that $\supp\phi_h \subseteq \clos{B_{N+1}}$.
-
-	By Taylor's theorem, we have
-	\[
-	\phi(x+h) = \sum_{\abs{\alpha} \leq N} \frac{h^\alpha}{\alpha!} \pt^\alpha \phi(x) + R_N(x, h),  
-	\]
-	and since $\sum_{\abs{\alpha} \leq N} \frac{h^\alpha}{\alpha!} \pt^\alpha \phi(x)$ vanishes for $x \notin \clos{B_N}$, it is clear that $\supp(R_N(\cdot, h))$ must also be contained in $\clos{B_{N+1}}$ (again, for $\norm{h} \leq 1$).
-This shows that $\supp(R_N)$ is contained in $\clos{B_{N+1}}$ for $\abs{h}$ sufficiently small.
-
-	Now let $\beta$ be a multi-index and define $C \ceq \frac{1}{N!} \max_{\abs{\alpha} = N+1, x \in \RR^n} \abs{(\pt^{\alpha + \beta})\phi(x)}$ (note that $C$ exists and is finite since all partial derivatives of $\phi$ have compact support), then we have
+	Note that the convolution between two compactly supported distributions is again compactly supported, which ensures that both expressions `make sense'. Now, let $\phi \in \DD(\RR^n)$, then we have
 	\begin{align*}
-	\abs{\pt^\beta R_N(x, h)} &= \abs{\pt^\beta \sum_{\abs{\alpha} = N+1} \frac{h^\alpha}{\alpha!}(N+1)\int_0^1 (1-t)^N (\pt^\alpha\phi)(x + th)\dd{t}} \\
-	%&= \abs{\sum_{\abs{\alpha} = N+1} \frac{h^\alpha}{\alpha!}(N+1) \cdot \pt^\beta\int_0^1 (1-t)^N (\pt^\alpha\phi)(x + th)\dd{t}} \\
-	&\overset\star= \abs{\sum_{\abs{\alpha} = N+1} \frac{h^\alpha}{\alpha!}(N+1) \int_0^1 (1-t)^N (\pt^{\alpha + \beta}\phi)(x + th)\dd{t}} \\
-	&\leq \sum_{\abs{\alpha} = N+1} \frac{\abs{h^\alpha}}{\alpha!}(N+1) \int_0^1 (1-t)^N \abs{\qty(\pt^{\alpha + \beta}\phi)(x+th)} \dd{t} \\
-	&\leq \qty[\max_{\abs{\alpha} = N+1, x \in \RR^n} \abs{(\pt^{\alpha + \beta})\phi(x)}] \sum_{\abs{\alpha} = N+1} \frac{\abs{h^\alpha}}{\alpha!} (N+1) \\
-	&\leq C (N+1)! \sum_{\abs{\alpha}= N+1}\frac{\abs{h^\alpha}}{\alpha!} = C(\abs{h_1} + \dotsb + \abs{h_n})^{N+1}.
+		((u * v) * w) * \phi &= (u* v) * (w * \phi) = u * (v * ( w* \phi)) = u * ((v * w) * \phi) = (u * (v * w)) * \phi,
 	\end{align*}
-Since this upper bound does not depend on $x$, we also have
-\[
-\sup_x \abs{\pt^\beta R_N(x, h)} \leq C(\abs{h_1} + \dotsb + \abs{h_n})^{N+1}, 
-\]
-and we conclude that
-\[ 
-\frac{\sup_x \abs{\pt^\beta R_N(x, h)}}{\norm{h}^N} \leq \frac{C(\abs{h_1} + \dotsb + \abs{h_n})^{N+1}}{\norm{h}^N} \leq \frac{C N^{N+1} \norm{h}^{N+1}}{\norm{h}^N} = CN^{N+1} \norm{h} \to 0, 
-\]
-and therefore that $\pt^\beta R_N(x, h) = o(\norm{h}^n)$ for all multi-indices $\beta$. 
+	which proves the theorem. 
 \end{proof}
@@
 \begin{question}
-	Which elements of $\DD(X)$ can be represented as a power series on $X$? 
-\end{question}
-
-\begin{solution}
-	It is known that if two power series agree on an open set, they agree on the entire space.
-Since every $\phi \in \DD(X)$ is identically zero on some open set (outside its support), the only element of $\DD(X)$ with a power series representation is the zero function.
-\end{solution}
-
-\begin{question}
-	Prove the $C^\infty$ Urysohn lemma: if $K$ is a compact subset of $X \subseteq \RR^n$, show that one can find a $\phi \in \DD(X)$ such that $0 \leq \phi \leq 1$ and $\phi = 1$ on a neighborhood of $K$.
-\end{question}
-
-\begin{solution}
-	Let $K \subseteq U_1$.
-Define $U_2 \ceq U_1 + B(0, 1)$ and let $\chi = \ind_{U_2}$.
-Now let $\psi \in \DD(\RR^n)$ satisfy $\int_{\RR^n} \psi \dd{x} = 1$ and $\supp\psi \subseteq B(0, 1)$.
-The we compute
-	\begin{align*}
-		(\chi * \psi)(x) &= \int_{\RR^n} \chi(y) \psi(x - y) \dd{y} = \int_{U_2} \psi(x - y) \dd{y}.
-	\end{align*}
-	Clearly, $\chi * \psi \in \DD(X)$, and furthermore, we have for $x \in U_1$ that
+	Let $\phi \in \DD(\RR)$ and choose $\eps > 0$ sufficiently small so that $\supp(\phi) \subset I_\eps = (-1/\eps, 1/\eps)$.
+Given that $\phi$ has a uniformly convergent Fourier series on $I_\eps$ in the form
 	\[
-	\int_{U_2} \psi(x - y) \dd{y} = \int_{U_2 - x} \psi(z) \dd{z} \overset\star= 1,
+	\phi(x) = \sum_{n \in \ZZ} c_n e^{i \eps \pi n x}, \quad c_n = \frac\eps2 \int_\RR \phi(x) e^{-i\eps \pi n x} \dd{x},
 	\]
-	since $B(0, 1) \subseteq U_2 - x$. This proves the claim.
-\end{solution}
-
-\begin{question}
-	Given $T \in \DD'(X)$, the derivative $\pt^\alpha T$ is defined by
-	\[
-	\ang{\pt^\alpha T, \phi} = (-1)^\abs{\alpha} \ang{T, \pt^\alpha\phi} \quad \forall \phi \in \DD(X). 
-	\]
-	Show that $\pt^\alpha T \in \DD'(X)$. If $\ord(T) = m$ what can you say about $\ord(\pt^\alpha T)$? 
+	prove the Fourier inversion theorem on $\DD(\RR)$ by taking a suitable limit. 
 \end{question}
@@
 \begin{proof}
-	Let $K \subseteq X$ be compact and $\phi \in \DD(X)$. Since $T$ is a distribution, we know that there exists constants $C, N$ such that
-	\[
-	\abs{\ang{T, \phi}} \leq C \sum_{\abs{\beta} \leq N} \sup \abs{\pt^\beta \phi}. 
+	Let $\psi \in \DD(\RR)$, then we want to prove that \[
+	\psi(x) = \frac1{(2\pi)^n} \iint e^{i \lambda (x - y)} \psi(y) \dd{y} \dd{\lambda}. 
 	\]
-
-	Letting $M \ceq \abs{\alpha}$, we find
-		\begin{align*}
-		\abs{\ang{\pt^\alpha T, \phi}} &= \abs{\ang{T, \pt^\alpha\phi}} \leq C \sum_{\abs{\beta} \leq N} \sup \abs{\pt^{\alpha + \beta}\phi} \leq C \sum_{\abs{\beta} \leq M + N} \sup \abs{\pt^\beta \phi}.
-	\end{align*}
-We conclude that $\pt^\alpha T$ is a distribution, and that if $\ord(T) = m$, $\ord(\pt^\alpha T) \leq m + \abs{\alpha}$. 
+	??
 \end{proof}
@@
 \begin{question}
-	Given $T \in \DD'(X)$ and $f \in C^\infty(X)$, prove that for each multi-index $\alpha$
-	\[
-	\pt^\alpha(Tf) = \sum_{\beta \leq \alpha} \binom{\alpha}{\beta} \pt^\beta f \pt^{\alpha - \beta}T
-	\]
-	in $\DD'(X)$. 
+	For $\phi \in \SS(\RR^n)$ prove that $\sum_m \phi(m) = \sum_n \hat\phi(2\pi n)$. This is the famous Poisson summation formula. 
 \end{question}
@@
 \begin{proof}
-	Let $\phi \in \DD(X)$, then by definition we have $\ang{\pt^\alpha(Tf), \phi} = \ang{T, (-1)^\abs{\alpha} f \pt^\alpha\phi}$. Approximate $T$ by a sequence $(\psi_n) \subseteq \DD'(X)$, then we find
+	We have
 	\begin{align*}
-	\ang{T, (-1)^{\abs{\alpha}} f \pt^\alpha\phi} &= \lim_{n\to\infty} \ang{\psi_n,  (-1)^{\abs{\alpha}} f \pt^\alpha\phi} = \lim_{n\to\infty}  (-1)^{\abs{\alpha}} \int_X \psi_n(x) f(x) \pt^\alpha\phi(x) \dd{x} \\
-	&= \lim_{n\to\infty} \int_X \pt^\alpha (\psi_n(x) f(x)) \phi(x) \dd{x} \\
-	&= \lim_{n\to\infty} \int_X \qty(\sum_{\beta \leq \alpha} \binom\alpha\beta\pt^\beta f(x) \cdot \pt^{\alpha - \beta} \psi_n(x)) \phi(x) \dd{x} \\
-	&= \lim_{n\to\infty} \sum_{\beta \leq \alpha} \binom\alpha\beta \ang{\pt^\beta f \pt^{\alpha - \beta} \psi_n, \phi} =  \lim_{n\to\infty} \sum_{\beta \leq \alpha} \binom\alpha\beta (-1)^{\abs{\alpha - \beta}} \ang{\psi_n, \pt^\beta f \pt^{\alpha - \beta} \phi} \\
-	&= \sum_{\beta \leq \alpha} \binom\alpha\beta (-1)^{\abs{\alpha - \beta}} \ang{T, \pt^\beta f \pt^{\alpha - \beta} \phi} = \sum_{\beta \leq \alpha} \binom\alpha\beta \ang{\pt^\beta f \pt^{\alpha - \beta} T, \phi} \\
-	&= \ang*{\sum_{\beta \leq \alpha} \binom{\alpha}{\beta} \pt^\beta f \pt^{\alpha - \beta}T, \phi}. 
-	%&= \lim_{n\to\infty} \ang*{\sum_{\abs{\beta} \leq \abs{\alpha}} \binom\alpha\beta \pt^\beta f\cdot \pt^{\alpha - \beta} \psi_n, \phi} = \lim_{n\to\infty} \sum_{\abs{\beta} \leq \abs{\alpha}
+		\sum_m \phi(m) = \frac1{(2\pi)^n} \sum_m \int e^{i\lambda m} \hat \phi(\lambda) \dd{\lambda} = \sum_m \int e^{2\pi i \lambda m} \hat \phi(2\pi \lambda) \dd{\lambda}
 	\end{align*}
 \end{proof}
-
-\begin{question}
-	Let $(x_k)$ be a sequence in $X$ with no limit point in $X$. Consider the family of linear maps
-	$u_\alpha \colon \DD(X) \to \CC$ defined by 
-	\[ \ang{u_\alpha, \phi} = \sum_{k=1}^\infty \pt^\alpha \phi(x_k)
-	 \]
-	 for each multi-index $\alpha$. For what $\alpha$ is $u_\alpha \in \DD'(X)$? What is $\ord(u_\alpha)$? 
-\end{question}
-
-\begin{solution}
-	Let $K \subseteq X$ be compact.
-Since $(x_k)$ does not have a limit point, only finitely many of the $x_k$ lie in $K$ (otherwise $(x_k)$ would have a subsequence contained in $K$ which would have a convergent subsequence).
-Without loss of generality, assume that $x_1, \dotsc, x_n \in K$, and $x_{n+1}, x_{n+2}, \dotsc, \notin K$.
-Now, for any $\phi \in \DD(X)$ with $\supp(\phi) \subseteq K$ we find
-	\begin{align*}
-		\abs{\ang{u, \phi}} &= \abs{\sum_{k=1}^\infty \pt^\alpha\phi(x_k)} = \abs{\sum_{k=1}^n \pt^\alpha\phi(x_k)} \leq \sum_{k=1}^n \abs{\pt^\alpha \phi(x_k)} \leq n \cdot \sup \abs{\pt^\alpha\phi} \leq n \cdot \sum_{\abs{\beta} \leq \abs{\alpha}} \sup \abs{\pt^\beta \phi}.
-	\end{align*}
-	This shows that $u_\alpha \in \DD'(X)$ for any $\alpha$, with $\ord(u_\alpha) \leq \abs{\alpha}$. We claim that this is an equality, i.e., $\ord(u_\alpha) = \abs{\alpha}$. \TODO How to show??
-\end{solution}
-
-\begin{question}
-	Find the most general solution to the equations
-	\begin{enumerate}[(a)]
-		\item $u'  = 1$,
-		\item $xu' = \delta_0$,
-		\item $(e^{2\pi i x} - 1)u' = 0$
-	\end{enumerate}
-	in $\DD'(\RR)$. 
-\end{question}
-
-\begin{solution}
-	Let $\phi \in \DD(X)$. 
-	\begin{enumerate}[(a)]
-		\item If $u' = 1$ then we find 
-		\[
-		\int_\RR \phi(x) \dd{x} = \ang{1, \phi} = \ang{u', \phi} = -\ang{u, \phi'},
-		\]
-		so for any $c \in \RR$ we find by partial integration $\ang{u, \phi'} = - \int_\RR\phi(x) \dd{x} = \int_\RR (x + c) \phi'(x) \dd{x}$. From this we deduce that $u = x + c$ for some $c$.
-
-		\item If $xu' = \delta_0$ then we have
-		\[
-		\phi(0) = \ang{\delta_0, \phi} = \ang{xu', \phi} = \ang{u', x\phi} = \ang{u, -(x\phi)'} = \ang{u, -x\phi' - \phi} %=  \ang{u, -x\phi'} + \ang{u, -\phi}. 
-		\]
-		Note that the above is satisfied for $u = - \delta_0 + c$ for any constant $c$. \TODO is this the most general solution?
-
-		\item Since $e^{2\pi i x} = 1 \iff x \in \ZZ$, intuitively it must be the case that $u'$ is 0, except ``on $\ZZ$'', whatever that may mean.
-Therefore, we guess that, for any sequence $(\alpha_n)_{n \in \ZZ} \subseteq \CC$ and constant $c \in \CC$, the map \[u = c + \sum_{n \in \ZZ} \alpha_n \ind_{x \geq n}.\] We compute the derivative of $u$.
-It is easily seen that $u' = \sum_{n \in \ZZ} \alpha_n \delta_n$ (the infiniteness of the sum does not pose a problem since the test functions are compactly supported, so $\ang{u, \phi}$ will always be a finite sum).
-%		We have for arbirary $\phi$ that
-%		\begin{align*}
-%		\ang{u', \phi} &= - \ang{u, \phi'} = -\int_\RR \sum_{n \in \ZZ} \qty(\alpha_n \ind_{x \geq n}) \phi'(x) \dd{x}	\overset\star= - \sum_{n \in \ZZ} \alpha_n \int_n^\infty \phi'(x) \dd{x} = \sum_{n \in \ZZ} \alpha_n \phi(n),
-%		\end{align*}
-%	where $\star$ follows from the fact that the sum is finite since $\phi'$ is compactly supported. 
-%	Therefore, we can conclude $u' = \sum_{n \in \ZZ} \alpha_n \delta_n$, and indeed we find
-From this, we see that
-	\[
-	\ang{(e^{2\pi i x} - 1) u', \phi} = \ang{u', (e^{2\pi i x} -1 )\phi} = \sum_{n \in \ZZ} \alpha_n (e^{2\pi i n} - 1) \phi(n) = 0,
-	\]
-	so $u$ satisfies the equation. \TODO Why is this the most general solution? Intuitively clear, but how to make this rigorous?
-	\end{enumerate}
-\end{solution}
-
-\begin{question}
-	Define the distribution $u \in \DD'(\RR^2)$ by the locally integrable function $u(x, y) = \ind_{x \geq y}$.
-Show that $\pt_x^2   u - \pt_y^2 u = 0$ in $\DD'(\RR^2)$.
-Can you give a physical interpretation of this result?
-\end{question}
-
-\begin{proof}
-	Let $f \in \DD(\RR^2)$, then we have
-	\begin{align*}
-		\ang{\pt_x^2 u - \pt_y^2 u, f} &= \ang{\pt_x^2 u, f}- \ang{\pt_y^2 u, f} = \ang{u, \pt_x^2 f} - \ang{u, \pt_y^2 f} = \ang{u, \pt_x^2 f - \pt_y^2 f} \\
-		&\overset\star= \int_{-\infty}^\infty \int_{y}^\infty \pt_x^2 f(x, y) \dd{x}\dd{y} - \int_{-\infty}^\infty \int_{-\infty}^x \pt_y^2 f(x, y) \dd{y} \dd{x} \\
-		&= - \int_{-\infty}^\infty \pt_x f(y, y) \dd{y} - \int_{-\infty}^\infty \pt_y f(x, x) \dd{x}
-		\\
-		&= - \int_{-\infty}^\infty (\pt_xf + \pt_yf)(x, x) \dd{x}.
-	\end{align*}
-Here, $\star$ follows from Fubini's theorem. 
-Define $g(x) = f(x, x)$, then it is easily seen that $g'(x) = \pt_x f(x, x) + \pt_y f(x, x)$, so we find that
-\[
-\ang{\pt_x^2 u - \pt_y^2 u, f} = 
-- \int_{-\infty}^\infty g'(x) \dd{x} = \lim_{x\to -\infty} g(x) - \lim_{x\to\infty} g(x) = 0 - 0 = 0. 
-\]
-This shows that $\pt_x u - \pt_y u = 0$, or equivalently, that $u$ satisfies the wave equation. 
-\end{proof}
-
-\begin{question}
-	Compute $\Delta(\norm{x}^{2-n})$ in $\DD'(\RR^n)$ for $n \geq 3$, i.e.\ compute
-	\[
-	\ang{\Delta (\norm{x}^{2-n}), \phi} = \ang{\norm{x}^{2-n}, \Delta\phi} = \int \frac{\Delta\phi}{\norm{x}^{n-2}} \dd{x}
-	\]
-	for arbitrary $\phi \in \DD(\RR^n)$. Note that $\norm{x} = (x_1^2 +\dotsb +x_n^2)^{1/2}$ and $\Delta = \sum_i (\pdv{x_i})^2$.
-
-	\emph{Hint:} use $\int\dd{x} = \int_{\norm{x}\leq\eps}\dd{x} + \int_{\norm{x}>\eps}\dd{x}$ and treat each integral separately. 
-\end{question}
-
-\begin{solution}
-	We follow the hint: let $\eps > 0$, then we first compute
-	\[
-	\int_{\norm{x} > \eps} \frac{\Delta\phi}{\norm{x}^{n-2}} \dd{x} = \sum_{i=1}^n \int_{\norm{x} > \eps} \frac{\pdv[2]{x_i} \phi}{\norm{x}^{n-2}}\dd{x} = \sum_{i=1}^n \int_{\norm{x} > \eps} \phi \cdot \pdv[2]{x_i} \norm{x}^{2-n} \dd{x}.
-	\]
-	Now, it is easily computed that $\pdv{x_i} \norm{x} = \frac{x_i}{\norm{x}}$, and therefore 
-	\begin{align*}
-	\pdv[2]{x_i} \norm{x}^{2-n} &= \pdv{x_i} (2-n) x_i \norm{x}^{-n} = (2-n) \qty(\norm{x}^{-n} - n x_i^2 \norm{x}^{-n-2}) \\
-	&= (n-2)\norm{x}^{-n} \qty(n \qty(\frac{x_i}{\norm{x}})^2 - 1). 
-	\end{align*}
-\end{solution}
-
 \end{document}

### 2020-11-11 topics_statistical_theory/C3.tex
I actually UNDERSTOOD topics

@@
 The regression function is given by $m(x) = \EE(Y_1 | X_1 = x)$ and the variance function by $v(x) = \Var(Y_1 | X_1 = x)$.
+
+\subsection{Local polynomial estimators}
+We will assume the fixed design setting. 
+\begin{definition}
+	Let $K$ be a kernel, $h > 0$ a bandwidth and $p \in \NN$.
+Then the \emph{local polynomial estimator of degree $p$ with bandwidth $h$ and kernel $K$}, denoted $\hat m_n (\cdot; p) \equiv \hat m_n(\cdot; p, h, K)$, is constructed at $x \in \RR$ by fitting a polynomial $p$ to the data using weighted least squares, where the pair $(x_i, Y_i)$ is assigned weight $K_h(x_i - x)$.
+\end{definition}
+
+To write this in formulas, define $Q(u) \ceq (1, u, \frac{u^2}{2}, \dotsc, \frac{u^p}{p!}) \in \RR^{p + 1}$ and $Q_h(\cdot) = Q(\cdot/h)$, we then have
+\[
+\hat m_n(x; p) = \hat\beta_0, \quad\text{where } \hat\beta \ceq \argmin_{\beta \in \RR^{p+1}} \sum_{i=1}^n \qty(Y_i - \beta\T Q_h(x_i - x))^2 K_h(x_i - x). 
+\]
+
+In matrix-vector notation, writing 
+\begin{align*}
+X &\equiv X(x;p, h) \ceq \mqty(Q_h(x_1 - x)\T \\ \vdots \\ Q_h(x_n - x)\T) \in \RR^{n \times (p + 1)}, \quad Y \ceq \mqty(Y_1 \\ \vdots \\ Y_n) \in \RR^n, \\
+W&\equiv W(x; h, K) \ceq \diag(K_h(x_1 - x), \dotsc, K_h(x_n - x)) \in \RR^{n \times n}, 
+\end{align*}
+we have 
+\[
+\hat\beta = \argmin_{\beta \in \RR^{p+1}} (Y - X\beta)\T W (Y - X\beta).
+\]
+
+By standard weighted least squares theory, we know that $\hat\beta$ must satisfy $X\T W X\hat\beta = X\T W Y$. 
+\begin{proposition}
+	Suppose $X\T W X$ is positive definite. Then we have
+	\[
+	\hat\beta = (X\T W X)^{-1} X\T W Y. 
+	\]
+
+\end{proposition}
+
+We will assume from here on that $X\T W X$ is indeed positive definite. 
+In this case, since the entries of $W$ and $X$ are functions of $x_i - x$, we can write the local polynomial estimator in the form
+\[
+\hat m_n(x) = n^{-1} \sum_{i=1}^n w(x, x_i) Y_i.
+\]
+(??)
+The set of weights $\qty{w(x, x_i)}$ is called the \emph{effective kernel} at $x$.
+
+For $p = 0$ and $p = 1$, there exist explicit formulas for the local polynomial estimator of degree $p$.
+
+\begin{proposition}[Reproducing property] \label{prop_reproducing}
+	Let $\qty{w_p(x, x_i)}$ denote the effective kernel of a local polynomial estimator of degree $p$ based on data $(x_1, Y_1), \dotsc, (x_n, Y_n)$, and let $R$ denote a polynomial of degree at most $p$.
+If $X\T W X$ is positive definite, then
+	\[
+	\frac1n \sum_{i=1}^n w_p(x, x_i) R(x_i) = R(x). 
+	\]
+\end{proposition}
+
+\begin{proof}
+	We have 
+	\begin{align*}
+		\frac1n \sum_{i=1}^n w_p(x, x_i) R(x_i) &= \hat\beta_0, 
+	\end{align*}
+where
+\begin{align*}
+	\hat\beta &= \argmin_{\beta \in \RR^{p + 1}} \sum_{i=1}^n \qty(R(x_i) - \beta\T Q_h(x_i - x))^2 K_h(x_i - x).
+\end{align*}
+By choosing $\hat\beta$ such that $\hat\beta\T Q_h(\cdot - x) = R$, we obtain 0, and therefore  this is indeed the minimiser.
+Since $Q_h(0) = e_1$, it follows that $R(x) = \hat\beta\T Q_h(x - x) = \hat\beta_0$, which concludes the proof.
+\end{proof}
+
+Before we can study the bias and variance of local polynomial estimators, we require the following lemma:
+\begin{lemma}
+	Let $K$ be a kernel that vanishes outside $[-1, 1]$, and suppose that $n^{-1} X\T W X$ is positive definite with minimal eigenvalue $\lambda_0 \equiv \lambda_{0, n, x} > 0$. Then 
+	\begin{enumerate}[(i)]
+		\item $\sup_{x\in[0, 1]} \max_{i = 1, \dotsc, n} \frac1n \abs{w(x, x_i)} \leq \frac{2\norm{K}_\infty}{\lambda_0 n h}$;
+		\item $n^{-1} \sum_{i=1}^n \abs{w(x, x_i)} \leq \frac{2\norm{K}_\infty}{\lambda_0 n h}\cdot  \sum_{i=1}^n \ind_{\abs{x_i - x} \leq h}$;
+		\item $w(x, x_i) = 0$ when $\abs{x_i - x} > h$. 
+	\end{enumerate}
+\end{lemma}
+
+\begin{proof}
+	\begin{enumerate}[(i)]
+		\item
+	Note that $n^{-1} w(x, x_i)$ is the $(0, i)$ entry of the matrix $(X\T W X)^{-1} X\T W$, and it is therefore less than the norm of the $i$-th column of $(X\T W X)^{-1} X\T W$.
+For $x \in [0, 1]$ and $i = 1, \dotsc, n$, we find
+	\begin{align*}
+		\frac1n \abs{w(x, x_i)} &\leq \norm{(X\T W X)^{-1} Q_h(x_i - x) K_h(x_i - x)} \overset\star\leq  \norm{K_h}_\infty \norm{(X\T W X)}^{-1} \norm{Q_h(x_i - x)} \ind_{\abs{x_i - x} \leq h} \\
+		&= \frac{\norm{K}_\infty}{h} \frac{1}{\lambda_0 n}\norm{Q_h(x_i - x)} \ind_{\abs{x_i - x} \leq h}  \leq \frac{\norm{K}_\infty}{\lambda_0 n h} \norm{Q(1)} = \frac{\norm{K}_\infty}{\lambda_0 n h} \qty(\sum_{j=0}^p \frac1{(j!)^2})^{1/2} \\
+		&\leq \frac{\norm{K}_\infty}{\lambda_0 n h} e^{1/2} \leq \frac{2 \norm{K}_\infty}{\lambda_0 nh}. 
+	\end{align*}
+	Here, the indicator function in $\star$ appears because $K$ vanishes outside $[-1, 1]$. 
+	Since the upper bound is independent of both $x$ and $i$, the claim is proven.
+
+	\item Similarly as before, we find 
+	\[
+	\frac1n \sum_{i=1}^n \abs{w(x, x_i)} \leq \frac{\norm{K}_\infty}{\lambda_0 n h} \sum_{i=1}^n \norm{Q_h(x_i - x)} \ind_{\abs{x_i - x} \leq h} \leq \frac{2 \norm{K}_\infty}{\lambda_0 n h} \sum_{i=1}^n \ind_{\abs{x_i - x} \leq h}.
+	\]
+
+	\item This follows immediately from inequality $\star$ in the proof of (i).
+
+	\end{enumerate}
+\end{proof}
+
+Now, we can compute bounds on the variance and bias of our local polynomial estimator. For simplicity, we will assume $x_i = i/n$. 
+\begin{proposition}
+Assume the model $Y_i = m(x_i) + v^{1/2}\eps_i$ with $m \in \HH(\beta, L)$ on $[0, 1]$ and $\max_i v(x_i) \leq \sigma_\Rm{max}^2$.
+Let $K$ be a kernel that vanishes outside $[-1, 1]$, and suppose that $\lambda_0$, the minimal eigenvalue of $n^{-1} X\T W X$, is strictly positive.
+Then, for $p \geq \ceil{\beta} - 1 \eqqcolon \beta_0$ and for each $x_0 \in [0, 1]$, $n \in \NN$ and $h \geq 1/(2n)$, we have
+\[
+\Var \hat m_n(x_0;p) \leq \frac{16 \norm{K}_\infty^2\sigma_\Rm{max}^2}{\lambda_0^2 n h}, \quad \abs{\Bias \hat m_n(x_0;p)} \leq \frac{8L \norm{K}_\infty}{\lambda_0 \beta_0!} h^\beta. 
+\]
+\end{proposition}
+
+\begin{proof}
+	Using the previous lemma, we obtain
+	\begin{align*}
+		\Var \hat m_n(x_0;p) &= \Var\qty(\frac1n \sum_{i=1}^n w(x_0, x_i) Y_i) = \frac1{n^2} \sum_{i=1}^n w(x_0, x_i)^2 \Var(Y_i) \\
+		&\leq \frac{\sigma_\Rm{max}^2}{n^2} \sum_{i=1}^n w(x_0, x_i)^2  \leq \sigma_\Rm{max}^2 \qty(\sup_{x \in [0, 1]} \max_i \frac1n \abs{w(x, x_i)}) \frac1n \sum_{i=1}^n \abs{w(x_0, x_i)} \\
+		&\leq \frac{4 \norm{K}_\infty^2 \sigma_\Rm{max}^2}{\lambda_0^2 n^2 h^2} \sum_{i=1}^n \ind_{\abs{i/n - x_0} \leq h}.
+	\end{align*}
+Now, we have $\abs{i/n - x_0} \leq h \iff i \in [n x_0 - nh, n x_0 + nh]$, and there are at most $2nh + 1$ integers in this interval. Recalling that $1 \leq 2nh$ we obtain 
+\[
+\Var \hat m_n(x_0; p) \leq \frac{4(2nh + 1) \norm{K}_\infty \sigma_\Rm{max}^2}{\lambda_0^2 n^2 h^2} \leq \frac{16 \norm{K}_\infty \sigma_\Rm{max}^2}{\lambda_0^2 nh} . 
+\]
+
+For the bias, we will first use a Taylor expansion combined with the reproducing property \cref{prop_reproducing}. 
+Firstly, since $\frac1n\sum_{i=1}^n w(x_0, x_i) = 1$ by this property, we have
+\[
+\Bias \hat m_n(x_0;p) = \qty(\frac1n \sum_{i=1}^n w(x_0, x_i) m(x_i)) - m(x_0) = \frac1n \sum_{i=1}^n w(x_0, x_i) \qty{m(x_i) - m(x_0)}.
+\] 
+Now, we apply a Taylor expansion to write \[
+m(x_i) - m(x_0) = P(x_i - x_0) + \frac{1}{\beta_0!} m^{(\beta_0)}(x_0 + \tau_i(x_i - x_0)) (x_i - x_0)^{\beta_0}, 
+\]
+where $P$ has degree at most $\beta_0 - 1 < p$ and a constant coefficient equal to 0,  and $\tau_i \in [0, 1]$.
+By the reproducing property we have $n^{-1} \sum_{i=1}^n w(x_0, x_i) P(x_i - x_0) = P(0) = 0$, so we obtain
+\begin{align*}
+\Bias \hat m_n(x;p) &= \frac1n w(x_0, x_i) \sum_{i=1}^n w(x_0, x_i) \frac{m^{(\beta_0)}(x_0 + \tau_i(x_i - x_0))}{\beta_0!} (x_i - x_0)^{\beta_0} \\
+&= \frac1n w(x_0, x_i) \sum_{i=1}^n w(x_0, x_i) \frac{m^{(\beta_0)}(x_0 + \tau_i(x_i - x_0)) - m^{(\beta_0)}(x_0)}{\beta_0!} (x_i - x_0)^{\beta_0},
+\end{align*}
+where the last line follows again from the reproducing property. Now we apply the fact that
+$m \in \HH(\beta, L)$ with the previous lemma and find
+\begin{align*}
+	\abs{\Bias \hat m_n(x;p)} &\leq \frac Ln \sum_{i=1}^n \abs{w(x_0, x_i)} \frac{\abs{x_i - x_0}^\beta}{\beta_0!} = \frac Ln \sum_{i=1}^n \abs{w(x_0, x_i)} \frac{\abs{x_i - x_0}^\beta}{\beta_0!} \ind_{\abs{x_i - x} \leq h} \\
+	&\leq \frac{Lh^\beta}{n\beta_0!} \sum_{i=1}^n \abs{w(x_0, x_i)} \leq \frac{8 L \norm{K}_\infty}{\lambda_0 \beta_0!}h^\beta. 
+\end{align*}
+\end{proof}
+
+So again, we have a variance term of order $1/nh$ and a bias term of order $h^\beta$. However, both our bounds depend on $\lambda_0$, which depends on both $n$ and $x$. 
+\begin{proposition}
+Suppose $x_i = i/n$ and that $K(u) \geq K_0 \ind_{\abs{u} \leq \Delta}$ for some $K_0, \Delta > 0$. Then, for $n \geq 2$ and $h \leq \frac1{4\Delta}$, 
+\[
+\inf_{x \in [0, 1]} \lambda_{0, n, x} \geq K_0 \inf_{v \in S^p} \min \qty{\int_0^\Delta (v\T Q(u))^2 \dd{u}, \int_{-\Delta}^0 (v\T Q(u))^2 \dd{u}} - \frac{(4\Delta + 2) K_0 e^{\Delta^2}}{nh}. 
+\]
+\end{proposition}
+
+\begin{proof}
+Since $\lambda_{0, n, x} = \inf_{v \in S^p} v\T \qty(\frac1n X\T W X)v$, we will try to bound this quantity from below.
+
+Let $u_i \ceq \frac{x_i - x}{h}$ for $i = 1, \dotsc, n$, so $u_1 \leq \frac{x_1}{h} = \frac{1}{nh}$, and let $u_0 \ceq 0$.
+
+First, we assume $x < 1 - h\Delta$, so that $u_n > \frac{1 - (1 - h\Delta)}{h} = \Delta$. Then, for any $v \in S^p$, we have
+\begin{align*}
+v\T \qty(\frac1n X\T W X) &= \frac1n (Xv)\T W (Xv) = \frac1n \sum_{i=1}^n (Xv)_i^2  W_{ii}\\
+& = \frac1{nh} \sum_{i=1}^n (v\T Q(u_i))^2 K(u_i) \geq \frac{K_0}{nh} \sum_{i=1}^n (v\T Q(u_i))^2 \ind_{u_i \in [0, \Delta]} .
+\end{align*}
+This is a Riemann sum (since $u_{i+1} - u_i = 1/(nh)$) and we will try to approximate it by the corresponding integral $K_0 \int_0^\Delta (v\T Q(u))^2 \dd{u}$.
+
+To do this, note that we have for $u \in [0, \Delta]$ that
+\[
+\norm{Q(u)}  \leq \qty{\sum_{\ell = 0}^p \frac{u^{2\ell}}{(\ell!)}}^{1/2} \leq \qty(\sum_{\ell = 0}^\infty \frac{u^{2\ell}}{\ell!})^{1/2} \leq e^{\Delta^2/2}. 
+\] 
+Furthermore, for $a, b \geq 0$, it can be shown that
+\[
+\abs{b^\ell - a^\ell} \leq \max(1, \ell / 2) (b^{\ell - 1} + a^{\ell - 1})\abs{b - a},
+\]
+and using this we obtain for $u, v \in [0, \Delta]$ that 
+\begin{align*}
+	\norm{Q(u)- Q(v)} &\leq \qty{\sum_{\ell =1 }^p \frac{\max(1, \ell^2/4) (u^{\ell - 1} + v^{\ell - 1})^2 (u - v)^2}{(\ell !)^2}}^{1/2} \\
+	&\leq \abs{u - v} \qty{\sum_{\ell = 1}^p \frac{\max(1, \ell^2/4)(2\Delta^{\ell - 1})^2}{(\ell!)^2}}^{1/2} \leq \abs{u - v} \qty{\sum_{\ell = 1}^p \frac{\max(4, \ell^2) \Delta^{2\ell - 2}}{(\ell!)^2}}^{1/2}
+	 \\
+	 &\overset\star\leq  2\abs{u - v} \qty{\sum_{\ell = 0}^{p - 1} \frac{\Delta^{2\ell}}{(\ell!)^2}}^{1/2} \leq 	2 \abs{u - v} \qty{\sum_{\ell = 0}^\infty \frac{\Delta^{2\ell}}{\ell!}}^{1/2} = 2 e^{\Delta^2/2}\abs{u-v},
+\end{align*}
+where in $\star$ we used that $\max(4, \ell^2) \leq 4\ell^2$.
+
+Now we estimate the difference between the Riemann sum and the corresponding integral. Write $i_- \ceq \min\qty{i : u_i > 0}$ and $i_+ \ceq \min\qty{i \mid u_i > \Delta}$.  Then 
+\begin{align*}
+\frac1{nh} \sum_{i=1}^n (v\T Q(u_i))^2 \ind_{u_i \in [0, \Delta]} &= \sum_{i=i_-}^{i_+ -1 } \int_{u_{i-1}}^{u_i} (v\T Q(u_i))^2 \dd{u} \\
+&\leq \sum_{i = i_-}^{i_+} \int_{\max(u_{i-1}, 0)}^{\min(u_i, \Delta)} (v\T Q(u_i))^2 \dd{u} + \frac1{nh} \sup_{u \in [0, \Delta]} (v\T Q(u))^2.
+\end{align*}
+(Note that the last integral, from $u_{i-1}$ to $\Delta$, is 0 since )
+Therefore we obtain
+\begin{align}
+	&\abs{\frac{1}{nh} \sum_{i=1}^n (v\T Q(u_i))^2 \ind_{u_i \in [0, \Delta]} -  \int_0^\Delta (v\T Q(u))^2 \dd{u}} \nonumber \\
+	&\leq \abs{\sum_{i = i_-}^{i_+} \int_{\max(u_{i-1}, 0)}^{\min(u_i, \Delta)}(v\T Q(u_i))^2 - (v\T Q(u))^2 \dd{u}} + \frac1{nh} \sup_{u \in [0, \Delta]} (v\T Q(u))^2. \label{eq:integral_expr}
+\end{align}
+Now note that by Cauchy-Schwarz since $\norm{v} = 1$ we have 
+\begin{align*}
+\abs{(v\T Q(u_i))^2 - (v\T Q(u))^2} &= \abs{(v\T Q(u_i) + v\T Q(u))}\abs{(v\T Q(u_i) - v\T Q(u))}\\
+&\leq (\norm{Q(u_i)} + \norm{Q(u)}) \norm{Q(u_i) - Q(u)} \\
+&\leq 2 e^{\Delta^2/2} \cdot 2e^{\Delta^2/2} \abs{u_i - u} = 4 e^{\Delta^2} \abs{u_i - u}, 
+\end{align*}
+and  therefore
+\[
+\eqref{eq:integral_expr} \leq 4e^{\Delta^2} \sum_{i=i_-}^{i_+} \int_{\max(u_{i-1}, 0)}^{\min(u_i, \Delta)}(u_i - u) \dd{u} + \frac{e^{\Delta^2}}{nh} \leq \frac{(4\Delta + 1) e^{\Delta^2}}{nh},
+\]
+which concludes the case $x < 1- h\Delta$.
+
+Suppose $x \geq 1 - h\Delta$, then we have $u_1 \leq -\Delta$ and $u_n \geq 0$, and we can apply very similar arguments to reach the desired conclusion. 
+\end{proof}
+%Suppose that $X\T W X$ is positive definite (this is a very weak condition), then we have
+%\begin{align*}
+%	\hat\beta = \argmin_{\beta \in \RR^{p+1}} (\beta - (X\T W X)^{-1} X\T W Y)\T (X\T W X) (\beta - (X\T W X)^{-1} X\T W Y),
+%\end{align*}
+%which is minimised at $\hat\beta = (X\T W X)^{-1} X\T WY$.

### 2020-11-14 distribution_theory/C1.tex
worked on some example sheets, updated msm and topics summaries (todo: dist and inv prob)
updated distribution theory, still struggling with topics ugh
it's weekend, party time

@@
 
-\subsection{Limits in $\DD'(X)$}
+\subsection{Limits in the distribution space}
 \begin{definition}
@@
 
-\subsection{Density of $\DD(\RR^n)$ in $\DD'(\RR^n)$}
-\begin{lemma}
+\subsection{Density of test functions in distribution space}
+\begin{lemma} \label{lem:convolution_associative}
 	If $u \in \DD'(\RR^n)$ and $\phi, \psi \in \DD(\RR^n)$, then
@@
 
+We will use the following trick many times:
+\begin{proposition}
+	For any $u \in \DD'(\RR^n), \phi \in \DD(\RR^n)$ we have $\ang{u, \phi} = (u * \refl\phi)(0)$. 
+\end{proposition}
+
+\begin{proof}
+	We have $(u * \refl\phi)(0) = \ang{u, \tau_0 \phi} = \ang{u, \phi}$. 
+\end{proof}
+For example, from this trick it follows that if $u * \phi = 0$ for all $\phi$, then $u = 0$.
+
 \begin{theorem}
@@
 
-	Note that $\ang{u, \phi} = (u * \refl\phi)(0)$ by definition. Now, take any $\theta \in \DD(\RR^n)$, then we have
+	Now, take any $\theta \in \DD(\RR^n)$, then we have
 

### 2020-11-14 distribution_theory/C4.tex
it's weekend, party time

@@
 \end{example}
+
+As a guess to construct fundamental solutions, we can use the Fourier transform: we have
+\begin{align*}
+	p(D) E = \delta_0 &\implies p(\lambda) \hat E = 1 \implies \hat E = \frac{1}{p(\lambda)}  \\
+	&\implies \ang{E, \phi} = \ang{E,\frac{1}{(2\pi)^n} \FF[\refl{\FF[\phi]}]} = \frac{1}{(2\pi)^n}\ang{\hat E, \refl{\FF[\phi]}} =\frac{1}{(2\pi)^n} \int \frac{\hat\phi(-\lambda)}{p(\lambda)} \dd{\lambda}.
+\end{align*}
+
+Indeed, one can check that this $E$ ``works'', but the problem is that we have no guarantee that $E \in \DD'(\RR^n)$, since $p(\lambda)$ may cause problems at its roots.
+To circumvent this, we have to use a construction called \emph{H\"ormander's staircase}.
+For this, we will first need a lemma. For $x \in \RR^n$, we will write $x = (x', x_n)$ with $x' \in \RR^{n-1}$. 
+\begin{lemma}
+	For each $\phi \in \DD(\RR^n)$ and $\lambda' \in \RR^{n-1}$, the function $z \mapsto \hat\phi(\lambda', z)$ is analytic in $z\in\CC$.
+Furthermore, for each $m \in \NN_0$ there exists constants $c_m, \delta > 0$ (independent of $\lambda'$) such that
+	\[
+	\abs{\hat\phi(\lambda', z)} \leq c_m(1 + \abs{z})^{-m} e^{\delta \abs{\Im z}}. 
+	\] 
+\end{lemma}
+
+\begin{proof}
+	By definition of the Fourier transform and Fubini's theorem, we have
+	\[
+	\hat\phi(\lambda', z) = \int e^{-i\lambda' \cdot x'} \int e^{-i z x_n} \phi(x', x) \dd{x_n} \dd{x'}. 
+	\]
+	It is easily seen that this function is smooth in $z$ and satisfies the Cauchy-Riemann equations, which means it is analytic.
+
+	Integrating by parts we find
+	\begin{align*}
+		\abs{z^m \hat\phi(\lambda', z)} &= \abs{\int e^{-i\lambda' \vdot x'} \int \qty(i \pdv{x_n})^m e^{-i z x_n} \phi(x', x_n) \dd{x_n} \dd{x'} } \\
+		&= \abs{\int e^{-i\lambda' \vdot x'} \int e^{-i z x_n} \qty(\pdv[m]{x_n} \phi(x', x_m)) \dd{x_m} \dd{x'}} \\
+		&\leq \iint \abs{e^{-iz x_n}} \cdot \abs{\pdv[m]{x_n} \phi(x', x_n)} \dd{x_n} \dd{x'} \\
+	&\leq c_m e^{\delta \abs{\Im z}}, 
+	\end{align*}
+	where $\delta$ is chosen such that $\phi(x', x_n) = 0$ if $\abs{x_n}> \delta$. 
+\end{proof}
+
+Now, we can prove the main theorem of this section, which \emph{almost} gives an explicit construction for a fundamental solution:
+\begin{theorem}
+	Every nonzero constant-coefficient partial differential operator has a fundamental solution. 
+\end{theorem}
+
+\begin{proof}
+	By rotating our coordinate axes, we can assume $p$ takes the form
+	\[
+	p(\lambda', \lambda_n) = \lambda_n^M + \sum_{m=1}^{M-1} a_m(\lambda') \lambda_n^m, 
+	\]
+	(??) (i.e., we simply write $p$ as a polynomial in $\lambda_n$). Fix $\mu' \in \RR^{n-1}$, then we can write
+	\[
+	p(\mu', \lambda_n) = \prod_{i=1}^M (\lambda_n - \tau_i(\mu')), 
+	\]
+	where the $\tau_i$ are the roots of the polynomial $\lambda_n \mapsto p(\mu', \lambda_n)$.
+Now, by the pigeonhole principle, there exists a horizontal line $\Im\lambda_n = c(\mu')$ in the region $\abs{\Im\lambda_n} \leq M + 1$ such that
+	\[
+	\abs{\lambda_n - \tau_i(\mu')} > 1 \quad\text{on $\Im\lambda_n = c(\mu')$} \quad (i = 1, \dotsc, m)
+	\]
+	Therefore, on $\Im(\lambda_n) = c(\mu')$ we have $\abs{p(\lambda', \lambda_n)} \ges 1$.
+
+	Since roots of a polynomial vary continuously with its coefficients, we can use the same horizontal line $\Im\lambda_n = c(\mu')$ for all $\lambda'$ in a (small) neighbourhood $N(\mu')$ of $\mu'$.
+We can cover all of $\RR^{n-1}$ with such neighbourhoods, and by the Heine-Borel theorem, we can extract a locally finite subcover $N_1 = N(\mu_1'), N_2 = N(\mu_2'), \dotsc$.
+Furthermore, we can modify these neighbourhoods so that they are disjoint by defining
+	\[
+	\Delta_i = N_i \setminus \qty(\bigcup_{j=1}^{i-1} \overline{N_j}). 
+	\]
+	The $\Delta_i$ are all open, disjoint, and satisfy $\RR^{n-1} = \cup_i \overline{\Delta_i}$.
+	Now we define
+	\[
+	\ang{E, \phi} \ceq \frac1{(2\pi)^n} \sum_{i=1}^\infty \int_{\Delta_i} \int_{\Im\lambda_n = c_i} \frac{\hat\phi(-\lambda' - \lambda_n)}{p(\lambda', \lambda_n)} \dd{\lambda_n} \dd{\lambda'}. 
+	\]
+	In ES3, it is shown that $E \in \DD'(\RR^n)$. Furthermore, we have
+	\begin{align*}
+		\ang{p(D)E, \phi} &= \ang{E, p(-D)\phi} = \frac1{(2\pi)^n} \sum_{i=1}^\infty \int_{\Delta_i} \int_{\Im\lambda_n = c_i} \frac{p(\lambda', \lambda_n)}{p(\lambda', \lambda_n)} \hat\phi(-\lambda' - \lambda_n) \dd{\lambda_n} \dd{\lambda'} \\
+		&= \frac1{(2\pi)^n} \sum_{i=1}^\infty \int_{\Delta_i} \int_{\Im\lambda_n = c_i}  \hat\phi(-\lambda' - \lambda_n) \dd{\lambda_n} \dd{\lambda'}  \\
+		&\overset\star= \frac1{(2\pi)^n} \sum_{i=1}^\infty \int_{\Delta_i} \int_{\Im\lambda_n = 0}  \hat\phi(-\lambda' - \lambda_n) \dd{\lambda_n} \dd{\lambda'} = \frac1{(2\pi)^n} \int_{\RR^n} \hat\phi(-\lambda) \dd{\lambda} = \phi(0).
+	\end{align*} 
+by the Fourier inversion theorem.
+Here, $\star$ follows from the Cauchy's theorem and the previous lemma ($\hat\phi$ decays rapidly in the horizontal direction, so taking a contour integral over a rectangle and letting the vertical side go to infinity shows that the integral over $\Im \lambda_n = c_i$ equals the integral over $\Im\lambda_n = 0$).
+We conclude that $p(D)E = \delta_0$.
+\end{proof}
+
+Note that the only nonconstructive part of the theorem is the extraction of a locally finite subcover of the neighbourhoods $N(\mu')$.
+
+\subsection{Structure theorem for distributions of compact support}
+In this section, we will prove that every $u \in \EE'(X)$ can be written as a finite sum $u = \sum_\alpha \pt^\alpha f_\alpha$ where the $f_\alpha$ are continuous.
+The theorem generalises to $u \in \DD'(X)$ (the sum can then be infinite, but locally finite), but we will not prove this, since it requires the use of partitions of unity.
+
+We start with a lemma:
+\begin{lemma}
+	For $u \in \EE'(\RR^n)$, the Fourier transform $\hat u \in \SS'(\RR^n)$ can be identified with the smooth (analytic) function $\lambda \mapsto \ang{u(x), e^{-i\lambda\vdot x}}$, which we will denote $\hat u(\lambda)$.
+\end{lemma}
+
+\begin{proof}
+	We will first prove the density of $\DD(\RR^n)$ in $\SS(\RR^n)$. 
+	Fix $\phi \in \SS(\RR^n)$ and $\chi \in \DD(\RR^n)$ with $\chi = 1$ on $\norm{x}\leq1$ and $\chi = 0$ on $\norm{x}>2$.
+Define $\phi_m \in \DD(\RR^n)$ by $\phi_m(x) \ceq \phi(x) \chi(x/m)$.
+We will show that $\phi_m \to \phi \in \SS(\RR^n)$.
+
+	For each pair of multi-indices $\alpha, \beta$, we have
+	\begin{align*}
+		\norm{\phi - \phi_m}_{\alpha, \beta} &= \norm{x^\alpha D^\beta( \phi - \phi_m)}_\infty = \norm{x^\alpha D^\beta(\phi \cdot \qty{1 - \chi(x/m)})} \\
+		&= \norm{x^\alpha \sum_{\gamma \leq \beta} \binom\beta\gamma (D^\gamma\phi)(x) \cdot D^{\beta - \gamma} (1 - \chi(x/m))}. 
+	\end{align*}
+	For $\gamma \neq \beta$, the derivative $D^\gamma\phi$ is bounded uniformly while the derivative $D^{\beta - \gamma} (1 - \chi(x/m))$ will converge uniformly to 0 since it will have at least one factor $1/m$.
+For $\gamma = \beta$, we have
+	\[
+	\norm{x^\alpha (1 - \chi(x/m)) D^\beta\phi}_\infty \leq \sup_{\norm{x} > M} \norm{x^\alpha D^\beta\phi} \to 0,
+	\]
+	since $D^\beta\phi$ decays rapidly. We conclude that $\norm{\phi - \phi_m}_{\alpha, \beta} \to 0$, and therefore that $\phi_m \to \phi$ in $\SS(\RR^n)$.
+
+	Now, by a Riemann sum argument (like the one we have used in \cref{lem:convolution_associative}) we have
+	\begin{align*}
+		\ang{\hat u, \phi_m} &= \ang{u, \hat \phi_m} = \ang*{u(x), \int e^{-i\lambda\vdot x} \phi_m(\lambda) \dd{\lambda}} 
+		\overset\star= \int \ang{u(x), e^{-i\lambda\cdot x}} \phi_m(\lambda) \dd{\lambda}, 
+		\end{align*}
+	where $\star$ is the Riemann sum argument (here, we need that $\phi_m$ has compact support).  Now, since $u \in \EE'(\RR^n)$, there exists a compact $K$ and constants $C', N > 0$ such that
+	\[
+	\abs{\ang{u(x), e^{-i\lambda\cdot x}}} \leq C' \sum_{\abs{\alpha} \leq N} \sup_K \abs{D_x e^{-i\lambda\cdot x}} \leq C \ang{\lambda}^N, 
+	\]
+	so $\lambda \mapsto \ang{u(x), e^{-i\lambda\cdot x}}$ is polynomially bounded, and therefore we can use the dominated convergence theorem to conclude
+	\[
+	\ang{\hat u, \phi} = \lim_{n\to\infty} \ang{\hat u, \phi_m} = \lim_{n\to\infty} \int \ang{u(x), e^{-i\lambda\cdot x}} \phi_m(\lambda) \dd{\lambda} = \int \ang{u(x), e^{-i\lambda\cdot x}} \phi(\lambda) \dd{\lambda},
+	\]
+	which proves that $\hat u$ can be identified with the function $\hat u(\lambda) = \ang{u(x), e^{-i\lambda\vdot x}}$. 
+\end{proof}
+
+Furthermore, it is clear that for $u \in \EE'(\RR^n)$, we have 
+\begin{equation} \label{eq:fourier_estimate}
+	\abs{\hat u(\lambda)} \leq C \sum_{\abs{\alpha} \leq N} \sup_K \abs{\pt^\alpha_x e^{-i\lambda x}} \les \ang{\lambda}^N.
+\end{equation}
+
+\begin{theorem}[Structure theorem]
+	For $u \in \EE'(X)$, there exists a finite collection $\qty{f_\alpha} \subseteq C(X)$ with $\supp(f_\alpha) \subseteq X$ such that $u = \sum_\alpha \pt^\alpha f_\alpha$ in $\EE'(X)$. 
+\end{theorem}
+
+\begin{proof}
+	Fix $\rho \in \DD(X)$ with $\rho = 1$ on a neighbourhood of $u$, then we can extend $u$ to $\EE'(\RR^n)$ by setting $\ang{u, \phi} \ceq \ang{u, \rho\phi}$ (note that $\rho\phi \in \DD(X)$ for all $\phi \in \EE(\RR^n)$).
+Since $\rho\phi \in \SS(\RR^n)$, we know there exist $\psi \in \SS(\RR^n)$ such that
+	\[
+	\rho\phi = \FF[\FF[\psi]] = (2\pi)^n \refl\psi, 
+	\]
+	and therefore
+	\[
+	\ang{u, \rho\phi} = \ang{u, \FF[\FF[\psi]]} = \ang{\hat u, \hat\psi}. 
+	\]
+	Using the Laplacian $\Delta = \sum_i \pt^i \pt^i$, we can write for any $m \in \NN$
+	\[
+	\hat\psi = \ang{\lambda}^{-2M} \FF\qty[(1 - \Delta)^M \psi],
+	\]
+	since $\FF\qty[(1 - \Delta)^m \psi] = (1 + \norm{\lambda}^2)^m \hat\psi = \ang{\lambda}^{2M}\hat\psi$.
+
+	Plugging this back into our previous equations, we have
+	\begin{equation} \label{eq:fourier_laplace}
+		\ang{\hat u, \hat\psi} = \ang{\hat u, \ang{\lambda}^{-2M} \FF[(1 - \Delta)^M \psi]} = \ang*{\FF[\hat u \ang{\lambda}^{-2M}], (1 - \Delta)^M \psi}.
+	\end{equation}
+Now, by \cref{eq:fourier_estimate}, we have $\abs{\hat u(\lambda)} \les \ang{\lambda}^N$, so we can choose $M$ large enough such that $\hat u(\lambda) \cdot \ang{\lambda}^{-2M} \in L^1(\RR^n)$, and by the dominated convergence theorem, the function
+\[
+f(x) = \frac1{(2\pi)^n} \int e^{-i\lambda \vdot x} \ang{\lambda}^{-2M} \hat u(\lambda) \dd{\lambda} 
+\]
+is continuous, and it is easily checked that $(2\pi)^n \refl f = \FF[\ang{\lambda}^{2M} \hat u(\lambda)]$.
+
+Using the fact that $(2\pi)^n \refl\psi = \rho\phi$, and going back to \cref{eq:fourier_laplace} we see
+\[
+\ang{u, \rho\phi} = \ang{(2\pi)^n \refl f, (1 - \Delta)^M \psi} = \ang{\refl f, (1 -\Delta)^M \refl{(\rho\phi)}} = \ang{f, (1 - \Delta)^M (\rho\phi)}, 
+\]
+where the last step follows from the fact that the Laplacian is reflection invariant.
+
+Expanding the derivatives using the Leibniz rule yields
+\[
+(1 -\Delta)^M (\rho\phi) = \sum_\alpha (-1)^{\abs{\alpha}} \rho_\alpha \pt^\alpha\phi
+\]
+for suitable $\rho_\alpha \in \DD(X)$, and therefore we have
+\[
+\ang{u, \phi} = \sum_\alpha \ang{f, (-1)^{\abs{\alpha}} \rho_\alpha \pt^\alpha\phi} = \ang*{\sum_\alpha \pt^\alpha(\rho_\alpha f), \phi},
+\]
+so $u = \sum_\alpha \pt^\alpha(\rho_\alpha f) = \sum_\alpha \pt^\alpha f_\alpha \in \EE'(\RR^n)$, where $f_\alpha$ is continuous and $\supp(f_\alpha) = \supp(\rho_\alpha f) \subseteq X$. 
+\end{proof}
+
+There also exist nonconstructive proofs for the previous theorem using Hahn-Banach.

### 2020-11-14 inverse_problems/inverse_ES2.tex
I actually UNDERSTOOD topics
it's weekend, party time

@@
 \author{Lucas Riedstra}
-\date{18 November 2020} % date
+\date{16 November 2020} % date
 
@@
 
-\item 
+\item Let $f \in L^1(\Omega) \setminus \BV(\Omega)$, then clearly $\pt \TV(f) = \emptyset$.
+Now suppose $f \in \BV(\Omega)$.
+It is known that the dual of $L^1(\Omega)$ is $L^\infty(\Omega)$.
+Therefore, we have for $p \in L^\infty(\Omega)$ that
+\begin{align*}
+	p \in \pt \TV(f) &\iff \TV(g) \geq \TV(f) + \int_{\Omega} p(x) (g - f)(x) \dd{x} \ \forall g \in L^1(\Omega)
+\end{align*}
+(??)
 \end{enumerate}
 \end{solution}
+
+\begin{question}
+	Let $\UU$ be a Banach space and let $E \colon \UU\to \rext$ be proper, lower semi-continuous and convex.
+Then the \emph{Fenchel conjugate} or \emph{convex conjugate} of $E$ is defined to be the mapping $E^* \colon \UU^* \to \RR$ with
+	\[
+	E^*(v) \ceq \sup_{u \in \UU} \qty{\ang{v, u} - E(u)}. 
+	\]
+	\begin{enumerate}[(a)]
+		\item Compute the convex conjugates of the following functionals.
+		\begin{enumerate}[(i)]
+			\item $E(u) = \norm{u}_\UU$ for a Banach space $\UU$,
+			\item $E(u \mid f) = \sum_{i=1}^n u_i \log(\frac{u_i}{f_i})$, where $f \in \RR^n_{>0}$ is a positive vector and $u \in \RR^n$. What is the effective domain of $E$? 
+		\end{enumerate}
+	\item Let $E \colon \UU\to\rext$ be a proper, lower semi-continuous and convex functional. Show that
+	\[
+	p \in \pt E(u) \iff u \in \pt E^*(p)
+	\]
+	for all $u, p \in \UU$.
+
+	\emph{Hint:} You may exploit the fact that under the stated assumptions $E = E^{**}$ holds true. 
+	\end{enumerate}
+\end{question}
+
+\begin{solution}
+	\begin{enumerate}[(a)]
+		\item \begin{enumerate}[(i)]
+			\item We have
+			\[
+			E^*(v) = \sup_{u \in \UU} \qty(\ang{v, u} - \norm{u}). 
+			\]
+			Suppose $\ang{v, u^*} - \norm{u^*} = \xi > 0$ for some $u^*$. Then we have for $\alpha > 0$ that
+			\[
+			\ang{v, \alpha u^*} - \norm{\alpha u^*} = \alpha \qty(\ang{v, u^*} - \norm{u^*}) = \alpha\xi, 
+			\]
+			and therefore clearly $E^*(v) = \infty$.
+
+			On the other hand, if $\ang{v, u} - \norm{u}\leq0$ for all $u$, then the supremum is attained in $u = 0$ with value $0$, and therefore $E^*(v) = 0$.
+
+			We see that
+			\begin{align*}
+				&\ang{v, u^*} - \norm{u^*} > 0 &\text{for some $u^* \in \UU$}; \\
+				&\iff \ang{v, u^*} > \norm{u^*} &\text{for some $u^* \in \UU$}; \\
+				&\iff \norm{v}_{\UU^*} > 1. 
+			\end{align*}
+		We conclude that
+		\[
+		E^*(v) = \chi_\qty{\norm{v} \leq 1} = \begin{cases}
+			0 &\text{if $\norm{v} \leq 1$}, \\ \infty &\text{else.}
+		\end{cases}
+		\]
+
+		\item Suppose first that $p \in \pt E(u)$. Then we have
+		\begin{align*}
+&\phantom{\implies}	p \in \pt E(u) \\
+&\implies E(v) \geq E(u) + \ang{p, v - u} &\text{for all $v$} \\
+&\implies \ang{p, u} - E(u) \geq \ang{p, v} - E(v) &\text{for all $v$} \\
+&\implies \ang{p, u} - E(u) \geq \sup_v \qty(\ang{p, v} - E(v)) = E^*(p) \\
+&\implies \ang{p, u} - E(u) + \ang{q - p, u} \geq E^*(p) + \ang{q - p, u} &\text{for all $q$} \\
+&\implies \ang{q, u} - E(u) \geq E^*(p) + \ang{q - p, u} &\text{for all $q$} \\
+&\implies \sup_v \qty(\ang{q, v} - E(v)) \geq E^*(p) + \ang{q - p, u} &\text{for all $q$} \\
+&\implies E^*(q) \geq E^*(p) + \ang{q -p, u} &\text{for all $q$} \\
+&\implies u \in \pt E^*(p). 
+		\end{align*}
+	Now, for the reverse implication, note that by what we just proved we have
+	\[
+	u \in \pt E^*(p) \implies p \in \pt E^{**}(u) \iff p \in \pt E(u), 
+	\]
+	which proves the claim.
+		\end{enumerate}
+	\end{enumerate}
+\end{solution}
+
+\begin{question}
+	Let $u, v \in \UU$ and $p \in \pt J(v)$. Recall that the Bregman distance of $J$ at $u, v$ is defined as
+	\[
+	D_J^p(u, v) \ceq J(u) - J(v) - \ang{p, u - v}.
+	\]
+	\begin{enumerate}[(a)]
+		\item Show that Bregman distances are non-negative. 
+		\item Show that Bregman distances may not be symmetric, i.e., there exists a $J$ and $u, v \in \UU$ with $p \in \pt J(v), q \in \pt J(u)$ such that $D_J^p(u, v) \neq D_J^q(v, u)$.
+		\item Show that a vanishing Bregman distance may not imply that the two arguments are the same. What if $J$ is strictly convex?
+	\end{enumerate}
+
+\end{question}
+
+\begin{proof}
+	\begin{enumerate}[(a)]
+		\item Since $p \in \pt J(v)$, we have $J(u) \geq J(v) + \ang{p, u - v}$, or equivalently $J(u) - J(v) - \ang{p, u- v} \geq 0$.
+
+		\item Let $\UU = \RR$, $J(x) = \abs{x}$, and choose $u = 0, v = 1$ and $p = 1, q = 0$. Then 
+		\[
+		D^p_J(u, v) = -1 - (-1) =  0 \quad\text{and}\quad D^q_J(v, u) = 1. 
+		\]
+
+		\item In the previous part we had an example $J(x)= \abs{x}, u = 0, v = 1, p = 1$, where $D_J^p(u, v) = 0$ while $u \neq v$.
+
+		Suppose that $J$ is strictly convex and that $u \neq v$ but $D_J^p(u, v) = 0$, so $J(u) = J(v) + \ang{p, u-v}$. Then we have for all $t \in (0, 1)$ that 
+		\begin{align*}
+			J(v) + \ang{p, (1-t)(u-v)} &= J(v) + \ang{p,  \qty(tv + (1-t)u) -v} \\
+			&\overset\star\leq J(tv + (1-t) u) \\
+			&< t J(v) + (1-t) J(u) \\
+			&= t J(v) + (1-t) \qty(J(v) + (1-t)\ang{p, u-v}) \\
+			&= J(v) + \ang{p, (1-t)(u-v)}, 
+		\end{align*}
+	a contradiction (here $\star$ follows from $p \in \pt J(v)$). We conclude that, if $J$ is strictly convex, the Bregman distance does satisfy $u \neq v \implies D_J^p(u, v) > 0$.
+
+	\end{enumerate}
+\end{proof}
+
+\begin{question}
+	Recall that a function $J \colon \UU\to \rext$ is called absolutely one-homogeneous if $J(\lambda u) = \abs{\lambda} J(u)$ for all $\lambda \in \RR, u \in \UU$.
+Let $J$ be convex, proper, l.s.c.\ and absolutely one-homogeneous.
+	\begin{enumerate}[(a)]
+		\item Show that $p \in \pt J(v)$ if and only if $p \in \pt J(0)$ and $J(v) = \ang{p, v}$.
+		Therefore, 
+		\[
+		D_J^p(u, v) = J(u) - \ang{p, u}. 
+		\]
+		Show that
+		\[
+		\pt J(0) = \bigcup_{u \in \UU} \pt J(u). 
+		\]
+
+		\item Show that the Bregman distances associated with absolutely one-homogeneous functionals fulfill the triangle inequality in the first argument, i.e., for all $u, v, w \in \UU$ and $p \in \pt J(w)$ there is
+		\[
+		D_J^p(u + v, w) \leq D_J^p(u, w) + D_J^p(v, w). 
+		\]
+
+		\item Show that the convex conjugate $J^*(\cdot)$ is the characteristic function of the convex set $\pt J(0)$. Compare this to the results of Exercise 2(a)(i). 
+	\end{enumerate}
+\end{question}
+
+\begin{proof}
+	It is clear that $J(0) = 0$. 
+	\begin{enumerate}[(a)]
+		\item Suppose $p \in \pt J(v)$.
+Then we have $J(u) \geq J(v) + \ang{p, u - v}$ for all $u$, which we can rewrite as $J(u) - \ang{p, u} \geq J(v) - \ang{p, v}$.
+Plugging in $u = 0$ we obtain $J(v) - \ang{p, v} \leq 0$, but plugging in $u = 2v$ we obtain
+		\[
+		2 \qty(J(v) - \ang{p, v}) = J(2v) - \ang{p, 2v} \geq J(v) - \ang{p, v} \implies J(v) - \ang{p, v} \geq 0, 
+		\]
+		so we conclude $J(v) - \ang{p, v} = 0$ or $J(v) = \ang{p, v}$. 	
+		This also implies that
+		\[
+		J(u) \geq \ang{p, u} \text{ for all $u$} \implies p \in \pt J(0). 		
+		\]
+
+		Conversely, if $p \in \pt J(0)$ and $J(v) = \ang{p, v}$, then for all $u$ we have
+		\[
+		J(u) \geq \ang{p, u} + (J(v) - \ang{p, v}) \implies p \in \pt J(v).
+		\]
+		This concludes the first claim.
+
+		From this claim, it follows that $\pt J(u) \subseteq \pt J(0)$ for all $u \in \UU$, and therefore trivially $\pt J(0) = \cup_u \pt J(u)$.
+
+		\item Note that we have
+		\[
+		J(u+v) = 2 J\qty(\frac12u + \frac12 v) \leq 2\qty(\frac12 J(u) + \frac12J(v)) = J(u) + J(v), 
+		\]
+		and therefore
+		\[
+		D_J^p(u+v, w) = J(u+v) - \ang{p, u +v} \leq J(u) + J(v) - \ang{p, u} - \ang{p, v} = D_J^p(u, w) + D_J^p(v, w). 
+		\]
+
+		\item We can reason analogously to 2(a)(i): we have
+		\[
+		J^*(v) = \sup_{u \in U} (\ang{v, u} - J(u)).
+		\]
+		Suppose that $v \notin \pt J(0)$, i.e., $\ang{v, u^*} - J(u) = \xi > 0$ for some $u^*$. Then we have for all $\lambda > 0$ that 
+		\[
+		\ang{v, \lambda u^*} - J(\lambda u^*) = \lambda \xi, 
+		\]
+		and letting $\lambda \to\infty$ shows $J^*(v) = \infty$.
+
+		On the other hand, suppose that $v \in \pt J(0)$, i.e., $\ang{v, u} - J(u) \leq 0$ for all $u$. Then the supremum is attained in $u = 0$ and therefore we have $J^*(v) = 0$.
+
+		It follows that $J^*(v) = \pt J(0)$, which is indeed also what we saw in 2(a)(i), since the subdifferential of the norm at 0 is exactly $\qty{v \in \UU^* : \norm{v}_{\UU^*} \leq 1}$. 
+	\end{enumerate}
+\end{proof}
 \end{document}

### 2020-11-14 modern_statistical_methods/C2.tex
it's weekend, party time

@@
 
+\subsubsection{Slow prediction error rate}
 We have the following bound on the prediction error:
@@
 
-\subsection{Concentration inequalities}
+\subsubsection{Concentration inequalities}
 Let $W$ be any random variable and $\phi \colon \RR\to[0,\infty)$ strictly increasing. Then by Markov's inequality we have
@@
 
+\subsubsection{Lasso solutions}
+
 With the previous proposition, we can write characterise solutions to the Lasso: namely we have
@@
 
-\subsection{Variable selection}
+\subsubsection{Variable selection}
 We consider the noiseless model $Y = X\beta^0$, where no randomness is involved.
@@
 \end{proof}
+
+\subsubsection*{The compatibility condition and random design}
+We will show that if $X$ is a random matrix, then the compatibility condition holds with high probability.
+
+Define, for $\Sigma \in \RR^{p \times p}$ and $S \subseteq \qty{1, \dotsc, p}$, the quantity
+\[
+\phi_\Sigma^2(S) = \inf_{\beta_S \neq 0, \norm{\beta_N}_1 \leq 3 \norm{\beta_S}_1} \frac{\beta\T \Sigma\beta}{\frac1{\abs{S}}\norm{\beta_S}_1^2},
+\]
+so that the compatiblity constant satisfies $\phi^2 = \phi_{\hat\Sigma}^2(S)$ with $\hat\Sigma \ceq X\T X /n $ and $S$ the support of $\beta^0$.
+
+To prove our main result, we first need a lemma:
+\begin{lemma}
+	Suppose $\phi_{\check\Sigma}^2(S) > 0$ and $\max_{jk} \abs{\hat\Sigma_{jk} - \check\Sigma_{jk}} \leq \phi_{\check\Sigma}^2(S) / (32\abs{S})$.
+Then $\phi_{\hat\Sigma}^2(S) \geq \phi_{\check\Sigma}^2(S)$.
+\end{lemma}
+
+\begin{proof}
+	Let $s = \abs{S}$ and $t \ceq \phi_{\check\Sigma}^2/(32s)$. Then by applying H\"older's inequality we obtain
+	\[
+	\abs{\beta\T (\check\Sigma - \hat\Sigma)\beta} \leq \norm{\beta}_1 \norm{(\check\Sigma - \hat\Sigma)\beta}_\infty \leq \norm{\beta}_1^2 \max_{j,k} \abs{\hat\Sigma_{jk} - \check\Sigma_{jk}}  \leq \norm{\beta}_1^2 t.
+	\]
+	Furthermore, if $\norm{\beta_N}_1 \leq 3 \norm{\beta_S}_1$, then
+	\[
+	\norm{\beta}_1 = \norm{\beta_N}_1 + \norm{\beta_S}_1 \leq 4 \norm{\beta_S}_1 \leq \frac{4\sqrt{\beta\T \check\Sigma\beta}}{\phi_{\check\Sigma} / s} \implies \norm{\beta_1}^2 \leq \frac{16 \beta\T \check\Sigma\beta}{\phi_{\check\Sigma}^2/s^2},
+	\]
+	and therefore
+	\[
+	\frac12 \beta\T \check\Sigma\beta = \beta\T \check\Sigma\beta - \frac{\phi_{\check\Sigma}^2}{32s} \cdot \frac{16 \beta\T \check\Sigma\beta}{\phi_{\check\Sigma}^2/s^2} \leq \beta\T \check\Sigma \beta -  t \norm{\beta}_1^2 \leq \beta\T\check\Sigma\beta - \beta\T(\check\Sigma - \hat\Sigma)\beta = \beta\T \hat\Sigma\beta.
+	\]
+	Taking the infimum over all $\beta$ with $\norm{\beta_N}_1 \leq 3 \norm{\beta_S}_1$ and $\beta_S \neq 0$, we obtain the required result. 
+\end{proof}
+
+Now we can state the main theorem. We consider a sequence of models with sample size $n$ (where $n\to\infty$), and with $p$ and $s$ increasing to $\infty$ as functions of $n$.
+
+\begin{theorem} \label{thm:lasso_compatibility_condition_asymptotics}
+	Suppose the rows of $X$ are i.i.d.\ and each entry is a mean-zero sub-Gaussian random variable with parameter $v$.
+Let $\Sigma^0 \ceq\EE[X\T X/n]$, and suppose that $s \sqrt{\log(p)/n} \to 0$ as $n\to\infty$.
+Define
+	\[
+	\phi^2_{\hat\Sigma, s} \ceq \min_{\abs{S} = s} \phi_{\hat\Sigma}^2(S), \quad \phi_{\Sigma^0, s}^2 \ceq \min_{\abs{S} = s} \phi_{\Sigma^0}^2(S). 
+	\]
+	If $\phi_{\Sigma^0, s}^2 > c > 0$ (for all $n$), then $\PP(\phi_{\hat\Sigma, s}^2 \geq \phi_{\Sigma^0, s}^2/2) \to 1$ as $n\to\infty$. 
+\end{theorem}
+
+\begin{proof}
+	By the previous lemma, it suffices to show that 
+	\[
+	(*) = \PP(\max_{jk}\abs{\hat\Sigma_{jk} - \Sigma^0_{jk}} \geq \phi_{\Sigma^0, s}^2 / (32s)) \to 0 \quad\text{ as } n \to\infty. 
+	\]
+	Letting $t \ceq \phi_{\Sigma^0, s}^2/(32s)$, then 
+	by a union bound and Bernstein's theorem, we have that 
+	\begin{align*}
+	(*) &\leq p^2 \max_{j, k} \PP\qty(\abs{\sum_{i=1}^n X_{ij} X_{ik}/n - \Sigma^0_{jk}} \geq t) \\
+	&\overset\star\leq p^2 2 \exp\qty(-\frac{nt^2}{2(64 v^4 + 4v^2 t)}) \\
+	&\leq c_1 \exp(-c_2 \frac{n}{s^2} + c_3 \log(p)) \to 0 \quad\text{if $s \sqrt{\log(p)/n} \to 0$ as $n\to\infty$}. 
+	\end{align*}
+where $\star$ follows from the fact that $X_{ij}X_{ik}$ satisfies Bernstein's condition with parameters $(8v^2, 4v^2)$. 
+\end{proof}
+
+What does the previous theorem tell us? We have a weak growth condition: for example, the theorem holds if $p = n^d$ and $s = (\sqrt n)^{1 -\alpha}$ for any $d, \alpha > 0$.
+
+\begin{example}
+	Suppose the rows of $X$ are taken form a $N_p(0, \Sigma^0)$ distribution, where $\Sigma^0$ has bounded diagonal entries and the smallest eigenvalue of $\Sigma^0$ is $c_\Rm{min} > c > 0$ (for all $n$).
+Then $\PP(\phi_{\hat\Sigma, s}^2 \geq \frac{c_\Rm{min}}{2}) \to 1$ as $n\to\infty$ if the growth condition is satisfied.
+This holds for example if $\Sigma^0 = I_p$.
+\end{example}
+
+\begin{example}
+	Suppose $X$ has i.i.d.\ entries $\PP(X_{ij} = 1) = \frac12 = \PP(X_{ij} = -1)$. Then by Hoeffding's lemma, $X_{ij}$ is sub-Gaussian, and therefore the theorem holds under appropriate conditions. 
+\end{example}
+
+\begin{example}
+	Let $W$ be a matrix whose rows are vectors in a basis, and let $X$ have rows which are random subsets of $W$.
+This does not satisfy the conditions of \cref{thm:lasso_compatibility_condition_asymptotics}, yet the lasso still works very well in this case.
+\end{example}
+
+Generally, the Lasso demonstrates a very low estimation if $n/p$ is much larger than $s/p$, but a very high estimation error if $s/p$ is much larger than $n/p$.
+Usually, there is a sharp ``phase transition'' here.
+
+\subsubsection{Computation}
+The \emph{coordinate descent} algorithm is an algorithm for minimising functions of the form $f \colon \RR^d \to \RR$, and works particularly well for functions of the form
+\[
+f(x) = g(x) + \sum_{j=1}^d h_j(x), 
+\]
+where $g$ is convex and differentiable and the $h_j \colon \RR \to \RR$ are convex and continuous. Note that the Lasso objective function is clearly of this form. 
+We start with an initial guess $x^{(0)}$ for a minimiser, and then for $m = 1, 2, \dotsc$ we choose, for $i = 1, \dotsc, d$, 
+\[
+x^{(m)}_i = \argmin_{x_i \in \RR} f\qty(x_1^{(m)}, \dotsc, x_{i-1}^{(m)}, x_i, x_{i+1}^{(m-1)}, \dotsc, x_d^{(m-1)}). 
+\]
+It can be shown that for the Lasso objective function, each step of the coordinate descent algorithm admits a closed-form solution.
+
+We have the following:
+\begin{theorem}[Tseng] \label{thm:tseng} If $A_0 \ceq \qty{x \mid f(x) \leq f(x^0)}$ is compact, then every converging subsequence of $(x^{(m)})$ converges to a minimiser of $f$. 
+\end{theorem}
+
+Using this theorem, we can see the following:
+\begin{corollary}
+	Suppose $A^0$ is compact. Then 
+	\begin{enumerate}
+		\item There exists a mimiser $x^*$ of $f$, and $f(x^{(m)}) \to f(x^*)$;
+		\item If $x^*$ is the unique minimiser of $f$, then $x^{(m)} \to x^*$. 
+	\end{enumerate}
+\end{corollary}
+
+\begin{proof}
+	Since $f$ is continuous, it attains a minimum $f(x^*)$ on $A_0$.
+Suppose $f(x^{(m)}) \not\to f(x^*)$.
+Note that $f(x^{(m)})$ is a monotone decreasing sequence, so it must converge.
+Furthermore, $(x^{(m)}) \subseteq A_0$, so there exists a subsequence of $(f(x^{(m)}))$ converging to $f(x^*)$ by \cref{thm:tseng}.
+This proves that $f(x^{(m)}) \to f(x^*)$.
+
+	Suppose that $x^*$ is the unique minimiser of $f$, then by \cref{thm:tseng} we know that every convergent subsequence of $x^{(m)}$ converges to $x^*$.
+Since $x^{(m)}$ is bounded, this proves that $x^{(m)} \to x^*$ as well.
+\end{proof}
+
+Often, we want to solve the Lasso on a grid of values $\lambda_0 > \dotsb > \lambda_L$.
+To speed up the coordinate descent procedure, we can use the minimiser for $\lambda_{\ell -1}$ as a starting point for coordinate descent on $\lambda_\ell$.
+An ``active set'' strategy can speed up computation even further:
+\begin{enumerate}
+	\item Initialise $A_\ell \ceq \qty{k \mid \beta^L_{\ell - 1, k} \neq 0}$;
+	\item Perform coordinate descent with $\lambda_\ell$ only on the coordinates in $A_\ell$, call this result $\hat\beta$. 
+	\item Let $V$ be the set of coordinates which violate the KKT conditions;
+	\item If $V = \emptyset$, set $\hat\beta_{\lambda_\ell}^L = \hat\beta$. Else, update $A_\ell = A_\ell \cup V$ and return to step 2. 
+\end{enumerate}
+
+\begin{remark}
+	There are multiple possible algorithms for computing Lasso solutions, coordinate descent is simply one option. 
+\end{remark}
+
+\subsection{Extensions of the Lasso}
+One of the ways to extend the Lasso is to change the loss function (for example,
+the logistic loss function can be used if $Y_i \in \qty{-1, 1}$). 
+There are other extensions.
+
+\subsubsection{Structural penalties}
+\begin{example}[Group lasso]
+	Let $G_1, \dotsc, G_q$ be a partition of $\qty{1, \dotsc, p}$, then instead of the Lasso penalty we can use
+	\[
+	\lambda \sum_{j=1}^q m_j \norm{\beta_{G_j}}_2,
+	\]
+	where $m_j$ is typically chosen as $\sqrt{\abs{G_j}}$. This estimator will generally ensure that either the entire group $G_j$ is shrunk to 0 or not at all. 
+\end{example}
+
+\begin{example}[Fused lasso]
+	Suppose the $(\beta_j)$ are ordered, for example if $j$ fixes time.
+If we expect $(\beta_j)$ to be piecewise constant (or at least that $\beta^0_j$ is close to $\beta^0_{j+1}$), we can use the penalty
+	\[
+	\lambda_1 \sum_{j=1}^{p-1} \abs{\beta_j - \beta_{j+1}} + \lambda_2 \norm{\beta}_1, 
+	\]
+	where the term $\lambda_2 \norm{\beta}_1$ can be omitted if shrinkage is not desired. 
+\end{example}
+
+\subsubsection{Debiasing the Lasso}
+A useful property of the Lasso is that non-significant coefficients can be shrunk towards 0, but this same effect also causes many significant coefficients to be shrunk  towards 0.
+A possible solution is to only use the Lasso to estimate which coefficients are 0 and then use a different estimator (i.e., OLS) for the other coefficients.
+
+Another option is to use a non-convex penalty function which does not shrink large values of $\hat\beta$ as much.

### 2020-11-14 modern_statistical_methods/C3.tex
it's weekend, party time

\section{High-dimensional covariance estimation and PCA}
Until now, we've focused on regression. In this chapter, we will assume the samples $x_1, \dotsc, x_n$ are i.i.d.\ from some distribution in $\RR^d$. We are interested in estimating the covariance matrix of $x_i$. There are two reasons for this:
\begin{enumerate}
	\item Covariance can reveal underlying structure relating the variables.
	\item When $d$ is large, interpreting the covariance matrix directly is hard. However, its largest eigenvalues and corresponding eigenvectors can reveal the principal modes of variation in the data. 
\end{enumerate}

\subsection{Covariance estimation}
\subsubsection{Maximum likelihood in multivariate normal model}
Let $x_1, \dotsc, x_n$ be i.i.d.\ samples from a $N_d(\mu, \Sigma)$ distribution. We will assume that $\Sigma$ is invertible with inverse $\Omega$ (this inverse is sometimes called the \emph{precision matrix}). Furthermore, we define the \emph{sample covariance} \[
\hat\Sigma \ceq \frac1n \sum_{i=1}^n (x_i - \bar X)(x_i - \bar X)\T. 
\]

It can be shown that the maximum likelihood estimate for $\mu$ is $\bar X \ceq \frac1n \sum_{i=1}^n x_i$, and that the maximum likelihood estimate $\Omega$ is
\[
\Omega = \min_{\Omega \succ 0} \qty(-\log(\det\Omega) + \tr(\hat\Sigma\Omega)). 
\]
Since this is the minimisation of a convex function over a convex set, we know that $\Omega$ is a minimiser if and only if the derivative w.r.t.\ each element of $\Omega$ is 0. From this, we can compute that $(\hat\Omega^\Rm{ML})^{-1} = \hat\Sigma$. Therefore, if $X$ has full column rank so that $\hat\Sigma$ is invertible, we find $\hat\Omega^\Rm{ML} = \hat\Sigma^{-1}$, and since matrix inversion is a bijective map on the set of positive definite matrices, we conclude $\hat\Sigma^\Rm{ML} = \hat\Sigma$.

\subsubsection{Non-asymptotic error bounds}
In the classical setting, we know that the maximum likelihood estimator is asymptotically optimal if we let $n \to \infty$ and keep all other parameters fixed. However, we are interested in \emph{non-asymptotic error bounds}, which do not rely on $n$ going to $\infty$. From these bounds, we can infer much more about the $n\to\infty$ case, namely, how much the other parameters are allowed to grow with $n$.

We will assume for simplicity that $\mu = 0$ and that $\hat\Sigma = \frac1n \sum_{i=1}^n x_i x_i\T$.  Recall that the largest (resp.\ smallest) eigenvalue of a symmetric matrix $A \in \RR^{n\times n}$ is given by the maximum (resp.\ minimum) of $v\T A v$ over all $v \in S^{n-1}$. It follows that the operator norm of $A$ is given by $\norm{A}_\Rm{op} = \max_{v \in S^{n-1}} \abs{v\T A v}$.

\begin{theorem}
	Let $X \in \RR^{n\times d}$ have i.i.d.\ rows $x_1, \dotsc, x_n$ with mean 0 and covariance $\Sigma$, so that 
	\[
	\EE\qty[ e^{\lambda v\T x_1}] \leq e^{\lambda^2\sigma^2/2} \quad\text{for all $v \in S^{d-1}$}. 
	\]
	Then there are universal constants $c, C$ such that $\hat\Sigma$ satisfies 
	\[
	\EE\qty[e^{\lambda\norm{\hat\Sigma - \Sigma}_\Rm{op}}] \leq \exp\qty(c \qty[ \frac{\lambda^2\sigma^4}{n} + d]),
	\]
	and for all $\delta > 0$, we have
	\[
	\PP\qty(\frac{\norm{\hat\Sigma - \Sigma}_\Rm{op}}{\sigma^2} \leq C \qty(\frac{d + \delta}{n} \lor \sqrt{\frac{d + \delta}{n}})) \leq 1 - e^{-\delta}. 
	\]
\end{theorem}

This is a non-asymptotic error bound. Note that, as long as $d/n \to 0$, the estimator $\hat\Sigma$ will be consistent.

### 2020-11-16 distribution_theory/distribution_ES2.tex
pffft

@@
 \DeclareMathOperator{\ord}{ord}
+%\DeclareMathOperator{\PV}{PV}
 
@@
 	\[
-	\abs{\ang{u * v, \phi}} \leq C' \sum_{\abs{\alpha} \leq N} \sum_{\abs{\beta} \leq M} \sup_{x \in -K' - K} \pt^\beta (\tau_x \refl{\pt^\alpha \refl\phi}) = C' \sum_{\abs{\alpha} \leq N} \sum_{\abs{\beta} \leq M} \sup_x \abs{\pt^{\alpha + \beta} \phi} \leq C'' \sum_{\abs{\alpha} \leq M + N} \sup_x \abs{\pt^\alpha \phi},
+	\abs{\ang{u * v, \phi}} \leq C' \sum_{\abs{\alpha} \leq N} \sum_{\abs{\beta} \leq M} \sup_{x \in -K' - K} \pt^\beta (\tau_x \refl{\pt^\alpha \refl\phi}) \leq C' \sum_{\abs{\alpha} \leq N} \sum_{\abs{\beta} \leq M} \sup_x \abs{\pt^{\alpha + \beta} \phi} \leq C'' \sum_{\abs{\alpha} \leq M + N} \sup_x \abs{\pt^\alpha \phi},
 	\]
@@
 \begin{proof}
-	Let $\psi \in \DD(\RR)$, then we want to prove that \[
-	\psi(x) = \frac1{(2\pi)^n} \iint e^{i \lambda (x - y)} \psi(y) \dd{y} \dd{\lambda}. 
-	\]
-	??
+Since $\DD(\RR) \subseteq \SS(\RR)$, we know that the Fourier inversion formula holds. We only need to show that the Fourier tranform of $\phi$ is again an element of $\DD(\RR)$. (??)
 \end{proof}
@@
 	\end{align*}
+(??)
+\end{proof}
+
+\begin{question}
+	If $u \in H^s(\RR^n)$ show that $D^\alpha u\in H^{s - \abs{\alpha}} (\RR^n)$. If $s > t$ show that $H^s(\RR^n) \subseteq H^t(\RR^n)$. 
+\end{question}
+
+\begin{proof}
+	Assuming $u \in H^s(\RR^n)$, we have
+	\begin{align*}
+		\int_{\RR^n} \ang{\lambda}^{2(s - \abs{\alpha})} \abs{\widehat{D^\alpha u}(\lambda)}^2 \dd{\lambda}  &= \int_{\RR^n} \ang{\lambda}^{2(s-\abs{\alpha})} \norm{\lambda}^{2\abs{\alpha}} \abs{\hat u(\lambda)}^2 \dd{\lambda} \\
+		&\les \int_{\RR^n} \ang{\lambda}^{2s} \abs{\hat u(\lambda)}^2 \dd{\lambda} < \infty, 
+	\end{align*}
+which proves $D^\alpha u \in H^{s- \abs{\alpha}}(\RR^n)$.
+
+The second claim follows immediately from the fact that $\ang{\lambda}^{t} \leq \ang{\lambda}^s$ for $s \geq t$ and $\lambda$ sufficiently large. 
+\end{proof}
+
+\begin{question}
+	Show that $S(\RR^n)$ is dense in $L^2(\RR^n) = H^0(\RR^n)$ and deduce that $\SS(\RR^n)$ is dense in $H^s(\RR^n)$.
+
+	\emph{Hint:} Use Parseval's theorem. 
+\end{question}
+
+\begin{proof}
+	We will show that $\DD(\RR^n)$ is dense in $L^2(\RR^n)$: since $\DD(\RR^n) \subseteq \SS(\RR^n)$, this shows that $\SS(\RR^n)$ is dense in $L^2(\RR^n)$ as well.
+
+	\TODO Give proof.
+
+	Now, for $u \in H^s(\RR^n)$, let $(\phi_m) \to u$ in $L^2$. Then we have
+	\begin{align*}
+		\norm{\phi_m - u}_{H^s}^2 = \int \ang{\lambda}^{2s} \abs{(\phi_m - u)(\lambda)}^2 \dd{\lambda}
+	\end{align*}
+\end{proof}
+%-----------------------------------------------------------------------------
+\setcounter{question}{18}
+
+\begin{question}
+	Compute the Fourier transforms of the functions
+	\begin{enumerate}[(a)]
+		\item $\sign(x)$;
+		\item $\arctan(x)$;
+		\item $x\log \abs{x} - x$;
+		\item $\exp(i\omega x^2)$
+	\end{enumerate}
+	in $\SS'(\RR)$, where $\omega \in \RR$. 
+\end{question}
+
+\begin{proof}
+	\begin{enumerate}[(a)]
+		\item We have for $\phi \in \SS(\RR)$ that 
+		\begin{align*}
+			\ang{\widehat{\sign}, \phi} &= \ang{\sign, \hat\phi} = \int_\RR \sign(\lambda) \hat\phi(\lambda)\dd{\lambda} = \int_\RR \sign(\lambda) \int_\RR e^{-i\lambda x} \phi(x) \dd{x} \dd{\lambda}
+			\\
+			&\overset\star= \int_\RR \phi(x) \int_\RR \sign(\lambda) e^{-i\lambda x} \dd{\lambda} \dd{x} 
+			=  \int_\RR \phi(x) \lim_{R\to\infty} \qty(\int_0^R e^{-i\lambda x} \dd{\lambda}  -  \int_{-R}^0 e^{-i\lambda x}  \dd{\lambda})\dd{x} \\
+			&= \lim_{\eps\to0} \lim_{R\to\infty} \int_{\abs{x}>\eps}  \phi(x) \qty(\frac{e^{ix R} + e^{-ix R}}{ix} - \frac2{ix}) \dd{x}  \\
+			&= \lim_{\eps \to 0} \lim_{R\to\infty} \int_\RR \frac{\phi(x) \ind_{\abs{x}> \eps}}{ix} \cdot e^{ixR} \dd{x} + \lim_{\eps \to 0} \lim_{R\to\infty} \int_\RR \frac{\phi(x) \ind_{\abs{x}> \eps}}{ix} \cdot e^{-ixR} \dd{x} + 2i \PV\qty(\frac1x).
+%			&= \lim_{R\to\infty}\qty( \int_\RR \phi(x) \int_0^R e^{-i\lambda x}\dd{x} - \int_\RR \phi(x) \int_{-R}^0 e^{-i\lambda x} \dd{x})
+		\end{align*}
+	We claim the first two terms go to 0: this is because the term in the integral is the Fourier transform of $\frac{\phi(x)\ind_{\abs{x} > \eps}}{ix}$ evaluated at $\pm R$, and since the function is in $L^1$, its Fourier transform decays to 0 as $\abs{R} \to\infty$.
+
+	We conclude $\widehat\sign = 2i \PV(\frac1x)$.
+
+	\item We know that $\arctan'(x) = \frac{1}{1 + x^2} \eqqcolon f(x)$, then we have $\widehat\arctan(\lambda) = \frac1{i\lambda} \hat f(\lambda)$ (in the distributional sense).
+
+	We have, using Fubini and the fact that $\ang{\hat f, \phi}$ is finite, that 
+	\begin{align*}
+		\ang{\hat f, \phi} &= \int_\RR \frac{\hat\phi(\lambda)}{1 + \lambda^2} \dd{\lambda} = \int_\RR \phi(x) \lim_{R\to\infty} \int_{-R}^R \frac{e^{-i\lambda x}}{1 + \lambda^2} \dd{\lambda} \dd{x} \overset\star= \int_\RR \phi(x) \qty(\pi e^{-\abs{x}}) \dd{x},
+	\end{align*}
+	from which it follows that the Fourier transform of $\frac{1}{1 + x^2}$ is given by $\pi e^{-\abs{\lambda}}$, and therefore the Fourier transform of $\arctan$ is given by $\frac{\pi}{i\lambda} e^{-\abs{\lambda}}$.
+
+	\item The derivative of this function, outside of 0, is $\log\abs{x}$.
+
+	\item  Clearly, if $\omega = 0$, the function is 1 and its Fourier transform is $2\pi\delta_0$, so assume $\omega \neq 0$. 
+	We have analogously to (b), with $f(x) = \exp(i\omega x^2)$, that 
+	\[
+	\ang{\widehat{f}, \phi} = \int_\RR \hat\phi(\lambda) e^{i\omega \lambda^2} \dd{\lambda} = \int_\RR \phi(x) \lim_{R\to\infty} \int_{-R}^R e^{i\omega\lambda^2 - i\lambda x} \dd{\lambda} \dd{x}.
+	\]
+	Now, by completing the square we have
+	\[
+	i(\omega\lambda^2 - x\lambda) = i\qty(\sqrt\omega\lambda - \frac{x}{2\sqrt\omega})^2 - \frac{ix^2}{4\omega},
+	\]
+	and therefore
+	\begin{align*}
+	\lim_{R\to\infty} \int_{-R}^R e^{i\omega\lambda^2 - i\lambda x} \dd{\lambda} &= e^{-ix^2/(4\omega)} \lim_{R\to\infty} \int_{-R}^R e^{i(\sqrt\omega\lambda - \frac{x}{2\sqrt\omega})^2}\dd{\lambda} = \frac{e^{-ix^2/(4\omega)}}{\sqrt\omega} \lim_{R\to\infty} \int_{-R}^R e^{i\lambda^2} \dd{\lambda} \\
+	&= \frac{e^{-ix^2/(4\omega)}}{\sqrt\omega} (1+i)\sqrt{\frac\pi2}, 
+	\end{align*}
+	where we use that the \emph{Fresnel integral} $\int_{-\infty}^\infty e^{ix^2} \dd{x}$ is known.
+
+	Plugging this back into our original equation yields
+	\[
+	\ang{\hat f, \phi} = \int_\RR \phi(x) \frac{e^{-ix^2/(4\omega)}}{\sqrt\omega} (1+i)\sqrt{\frac\pi2} \dd{x}, 
+	\]
+	which shows that
+	\[
+	\hat f(\lambda) =  (1+i) e^{-i\lambda^2/(4\omega)} \sqrt{\frac{\pi}{2\omega}}. 
+	\]
+	in the distributional sense. 
+	\end{enumerate}
 \end{proof}

### 2020-11-16 inverse_problems/inverse_ES2_handin.tex
pffft

--- 2020-11-10 inverse_problems/inverse_ES2.tex
@@
 
-\item 
+\item Let $f \in L^1(\Omega) \setminus \BV(\Omega)$, then clearly $\pt \TV(f) = \emptyset$.
+Now suppose $f \in \BV(\Omega)$.
+It is known that the dual of $L^1(\Omega)$ is $L^\infty(\Omega)$.
+Therefore, we have for $p \in L^\infty(\Omega)$ that
+\begin{align*}
+	p \in \pt \TV(f) &\iff \TV(g) \geq \TV(f) + \int_{\Omega} p(x) (g - f)(x) \dd{x} \ \forall g \in L^1(\Omega)
+\end{align*}
+I do not know how to continue from here. 
 \end{enumerate}
 \end{solution}
+
+\begin{question}
+	Recall that a function $J \colon \UU\to \rext$ is called absolutely one-homogeneous if $J(\lambda u) = \abs{\lambda} J(u)$ for all $\lambda \in \RR, u \in \UU$.
+Let $J$ be convex, proper, l.s.c.\ and absolutely one-homogeneous.
+	\begin{enumerate}[(a)]
+		\item Show that $p \in \pt J(v)$ if and only if $p \in \pt J(0)$ and $J(v) = \ang{p, v}$.
+		Therefore, 
+		\[
+		D_J^p(u, v) = J(u) - \ang{p, u}. 
+		\]
+		Show that
+		\[
+		\pt J(0) = \bigcup_{u \in \UU} \pt J(u). 
+		\]
+
+		\item Show that the Bregman distances associated with absolutely one-homogeneous functionals fulfill the triangle inequality in the first argument, i.e., for all $u, v, w \in \UU$ and $p \in \pt J(w)$ there is
+		\[
+		D_J^p(u + v, w) \leq D_J^p(u, w) + D_J^p(v, w). 
+		\]
+
+		\item Show that the convex conjugate $J^*(\cdot)$ is the characteristic function of the convex set $\pt J(0)$. Compare this to the results of Exercise 2(a)(i). 
+	\end{enumerate}
+\end{question}
+
+\begin{proof}
+	It is clear that $J(0) = 0$. 
+	\begin{enumerate}[(a)]
+		\item Suppose $p \in \pt J(v)$.
+Then we have $J(u) \geq J(v) + \ang{p, u - v}$ for all $u$, which we can rewrite as $J(u) - \ang{p, u} \geq J(v) - \ang{p, v}$.
+Plugging in $u = 0$ we obtain $J(v) - \ang{p, v} \leq 0$, but plugging in $u = 2v$ we obtain
+		\[
+		2 \qty(J(v) - \ang{p, v}) = J(2v) - \ang{p, 2v} \geq J(v) - \ang{p, v} \implies J(v) - \ang{p, v} \geq 0, 
+		\]
+		so we conclude $J(v) - \ang{p, v} = 0$ or $J(v) = \ang{p, v}$. 	
+		This also implies that
+		\[
+		J(u) \geq \ang{p, u} \text{ for all $u$} \implies p \in \pt J(0). 		
+		\]
+
+		Conversely, if $p \in \pt J(0)$ and $J(v) = \ang{p, v}$, then for all $u$ we have
+		\[
+		J(u) \geq \ang{p, u} + (J(v) - \ang{p, v}) \implies p \in \pt J(v).
+		\]
+		This concludes the first claim.
+
+		From this claim, it follows that $\pt J(u) \subseteq \pt J(0)$ for all $u \in \UU$, and therefore trivially $\pt J(0) = \cup_u \pt J(u)$.
+
+		\item Note that we have
+		\[
+		J(u+v) = 2 J\qty(\frac12u + \frac12 v) \leq 2\qty(\frac12 J(u) + \frac12J(v)) = J(u) + J(v), 
+		\]
+		and therefore
+		\[
+		D_J^p(u+v, w) = J(u+v) - \ang{p, u +v} \leq J(u) + J(v) - \ang{p, u} - \ang{p, v} = D_J^p(u, w) + D_J^p(v, w). 
+		\]
+
+		\item We can reason analogously to 2(a)(i): we have
+		\[
+		J^*(v) = \sup_{u \in U} (\ang{v, u} - J(u)).
+		\]
+		Suppose that $v \notin \pt J(0)$, i.e., $\ang{v, u^*} - J(u) = \xi > 0$ for some $u^*$. Then we have for all $\lambda > 0$ that 
+		\[
+		\ang{v, \lambda u^*} - J(\lambda u^*) = \lambda \xi, 
+		\]
+		and letting $\lambda \to\infty$ shows $J^*(v) = \infty$.
+
+		On the other hand, suppose that $v \in \pt J(0)$, i.e., $\ang{v, u} - J(u) \leq 0$ for all $u$. Then the supremum is attained in $u = 0$ and therefore we have $J^*(v) = 0$.
+
+		It follows that $J^*(v) = \pt J(0)$, which is indeed also what we saw in 2(a)(i), since the subdifferential of the norm at 0 is exactly $\qty{v \in \UU^* : \norm{v}_{\UU^*} \leq 1}$. 
+	\end{enumerate}
+\end{proof}
 \end{document}

### 2020-11-16 modern_statistical_methods/msm_ES3.tex
pffft

--- 2020-10-31 modern_statistical_methods/msm_ES2.tex
@@
 
-\title{Modern Statistical Methods --- Example Sheet 2} % subject
+\title{Modern Statistical Methods --- Example Sheet 3} % subject
 \author{Lucas Riedstra}
-\date{...} % date
+%\date{} % date
 
@@
 \maketitle
+\begin{mdframed}
+	In all of the below, assume that any design matrices $X$ are $n \times p$ and have their columns centred and then scaled to have $\ell^2$-norm $\sqrt n$, and that any responses $Y \in \RR^n$ are centred.
+\end{mdframed}
 \begin{question}
-	Let $Y \in \RR^n$ be a vector of responses, $\Phi \in \RR^{n \times p}$ a design matrix, $J \colon [0, \infty) \to [0, \infty)$ a strictly increasing function and $c \colon \RR^n \to \RR^n$ some cost function.
-Set $K = \Phi\Phi\T$.
-Show, without using the representer theorem, that $\hat\theta$ minimises
+	When proving the theorems on the prediction error of the Lasso, we started with the so-called basic inequality that
 	\[
-	Q_1(\theta) \ceq c(Y, \Phi\theta) + J(\norm{\theta}_2^2)
+	\frac1{2n} \norm{X(\beta^0 - \hat\beta)}_2^2 \leq \frac1n \eps\T X(\hat\beta - \beta^0) + \lambda \norm{\beta^0}_1 - \lambda \norm{\hat\beta}_1. 
 	\]
-	over $\theta \in \RR^p$ if and only if $\Phi\hat\theta = K\hat\alpha$ and $\hat\alpha$ minimises
+	Show that in fact we can improve this to 
 		\[
-	Q_2(\alpha) \ceq c(Y, K\alpha) + J(\alpha\T K\alpha)
+	\frac1{n} \norm{X(\beta^0 - \hat\beta)}_2^2 \leq \frac1n \eps\T X(\hat\beta - \beta^0) + \lambda \norm{\beta^0}_1 - \lambda \norm{\hat\beta}_1. 
 	\]
-	over $\alpha \in \RR^n$. 
 \end{question}
@@
 \begin{proof}
-	Let $\hat\theta$ be a minimiser of $Q_1$, and write $\hat\theta = \Phi\T\hat\alpha + \hat\beta$ with $\Phi\T \hat\alpha \in \Nul(\Phi)\P = \Ran(\Phi\T)$, $\hat\beta \in \Nul(\Phi)$.
-
-	Noting that $K\hat\alpha = \Phi\Phi\T \hat\alpha = \Phi\hat\theta$ and  $\norm{\Phi\T\hat\alpha} = \alpha\T K\alpha$ we see
-	\[
-	Q_1(\theta) = c(Y, K\hat\alpha) + J(\alpha\T K\alpha + \|\hat\beta\|^2),
-	\]
-	and therefore it is necessary that $\hat\beta = 0$. The claim follows. 
+	By the KKT conditions, we there exists $\hat\nu \in \RR^p$ with $\norm{\hat\nu}_\infty \leq 1$ and $\ang{\hat\nu, \hat\beta} = \norm{\beta}_1$ such that 
+	\begin{align*}
+		\frac1n X\T (Y - X\hat\beta) &= \lambda\hat\nu \\
+		\frac1n X\T (X(\beta^0 - \hat\beta) + \eps) &= \lambda\hat\nu \\
+		\frac1n X\T X(\beta^0 - \hat\beta) &= - \frac1n X\T\eps + \lambda\hat\nu \\
+		\frac1n (\beta^0 - \hat\beta)\T X\T X(\beta^0 - \hat\beta) &= -\frac1n (\beta^0 - \hat\beta)\T X\T\eps + \lambda (\beta^0 - \hat\beta)\T \hat\nu \\
+		\frac1n \norm{X(\beta^0 -  \hat\beta)}_2^2 &= \frac1n \eps\T X(\hat\beta - \beta^0) + \lambda \ang{\hat\nu, \beta^0} - \lambda \norm{\hat\beta}_1,
+		\end{align*}
+	and now plugging in $\ang{\beta^0, \hat\nu} \leq \norm{\beta^0}_1 \norm{\hat\nu}_\infty \leq \norm{\beta^0}_1$ yields the result. 
 \end{proof}
@@
 \begin{question}
-	Let $x, x' \in \RR^p$ and let $\psi \in \qty{-1, 1}^p$ be a random vector with independent components taking values $-1, 1$ each with probability $1/2$.
-Show that $\EE(\psi\T x \psi\T x') = x\T x'$.
-Construct a random feature map $\hat\phi \colon \RR^p \to \RR$ such that $\EE\qty{\hat\phi(x)\hat\phi(x')} = (x\T x)^2$.
+	Under the assumptions of Theorem 23 on the prediction and estimation properties of the Lasso under a compatibility condition, show that, with probability $1 - 2p^{-(A^2/8 - 1)}$, we have
+	\[
+	\frac1n \norm{X(\beta^0 - \hat\beta)}_2^2 \leq \frac{9A^2 \log(p)}{4\phi^2} \frac{\sigma^2 s}{n}. 
+	\]
 \end{question}
-
-\begin{solution}
-	We have 
-	\begin{align*}
-		\psi\T x \psi\T x' = \qty(\sum_i \psi_i x_i) \qty(\sum_j \psi_j x'_j) = \sum_i x_i x_i' + 2 \sum_{i < j} \psi_i \psi_j x_i x_j'. 
-	\end{align*}
-	Noting that for $i \neq j$ we have $\EE[\psi_i \psi_j] = \EE[\psi_i] \EE[\psi_j] = 0$ it follows that $\EE[\psi\T x \psi\T x'] = \sum_i x_i x_i' = x\T x'$.
-
-	Let $\psi_*$  be an identical independent copy of $\psi$ and define $\hat\phi(x) = \psi\T x \psi_*\T x$. Then we find
-	\[
-	\EE[\hat\phi(x)\hat\phi(x')] = \EE[\psi\T x \psi\T x']\EE[\psi_*\T x \psi_*\T x'] = (x\T x')^2. 
-	\]
-\end{solution}
-
-\begin{question}
-	Let $\XX = \pow(\qty{1, \dotsc, p})$ and $z, z' \in \XX$.
-Let $k$ be the Jaccard similarity kernel.
-Let $\pi$ be a random permutation of $\qty{1, \dotsc, p}$.
-Let $M = \min\qty{\pi(j) \mid j \in z}$, $M' = \min\qty{\pi(j) \mid j \in z'}$.
-Show that
-	\[
-	\PP(M = M') = k(z, z'),
-	\]
-	when $z, z' \neq \emptyset$.
-Now let $\psi \in \qty{-1, 1}^p$ be a random vector with i.i.d.\ components taking the values -1 or 1, each with probability $1/2$.
-By considering $\EE[\psi_M\psi_{M'}]$ show that the Jaccard similarity kernel is indeed a kernel.
-Explain how we can use the ideas above to approximate kernel ridge regression with Jaccard similarity, when $n$ is very large (you may assume none of the data points are the empty set).
-\end{question}
-
-\begin{proof}
-	We have \[
-	\PP(M = M') = \PP\qty(\argmin_{j \in z \cup z'} \pi(j) \in z \cap z') = \frac{\abs{z \cap z'}}{\abs{z \cup z'}} = k(z, z') \quad\text{since $\pi$ is random.}
-	\]
-
-	Furthermore, we have
-	\[
-	\EE[\psi_M \psi_{M'}] = \PP(M = M') \EE[\psi_M^2] + \PP(M \neq M') \EE[\psi_M \psi_{M'}] = k(z, z'), 
-	\]
-	since for $M \neq M'$ we have $\EE[\psi_M\psi_M'] = \EE[\psi_M]\EE[\psi_{M'}] = 0$. 
-	Let $z_1, \dotsc, z_n \in \XX$ with corresponding $M_1, \dotsc, M_n$, and write $\hat \psi = (\psi_{M_1}, \dotsc, \psi_{M_n})\T$, then the kernel matrix $K$ is given by $\EE[\hat\psi \hat\psi\T]$ which is positive semidefinite.
-
-	Using the random feature map $\hat\phi(z) = \psi_{M_z}$ we can approximate kernel ridge regression using the random feature map method. 
-\end{proof}
-
-\begin{question}
-	Consider the logistic regression model where we assume $Y_1, \dotsc, Y_n \in \qty{-1, 1}$ are independent and
-	\[
-	\log(\frac{\PP(Y_i = 1)}{\PP(Y_i = -1)}) = x_i\T\beta^0. 
-	\]
-	Show that the maximum likelihood estimate $\beta$ minimises
-	\[
-	\sum_{i=1}^n \log(1 + \exp(-Y_i x_i\T \beta))
-	\]
-	over $\beta \in \RR^p$. 
-\end{question}
-
-\begin{proof}
-	Let $(y_1, \dotsc, y_n)$ be the responses, then the likelihood function becomes
-	\[
-	L(\beta) = \PP(Y_1, \dotsc, Y_n \mid \beta)
-	\]
-
-	Note that
-	\begin{align*}
-		\frac{\PP(Y_i = 1)}{1 - \PP(Y_i = 1)} = \exp(x_i\T\beta) \implies \PP(Y_i = 1) = \frac{1}{1 + \exp(-x_i\T \beta)},
-	\end{align*}
-and analogously
-\begin{align*}
-	\PP(Y_i = -1) =  \frac{1}{1 + \exp(x_i\T\beta)}. 
-\end{align*}
-We can combine the above formulas as
-\[
-\PP(Y_i = y_i) = \frac{1}{1 + \exp(-y_i x_i\T\beta)}. 
-\]
-
-Therefore our the MLE $\hat\beta$ maximises
-\[
-L(\beta) = \prod_{i=1}^n \frac{1}{1 + \exp(-Y_i x_i\T\beta)}, 
-\]
-and it also maximises the log-likelihood function
-\[
-\log(L(\beta)) = - \sum_{i=1}^n \log(1 + \exp(-Y_i x_i\T\beta))
-\]
-which is of course equivalent to minimising
-\[
--\log(L(\beta)) = \sum_{i=1}^n \log(1 + \exp(-Y_i x_i\T\beta)). 
-\]
-\end{proof}
-
-\begin{question}
-	Consider the following algorithm for model selection when we have a response $Y \in \RR^n$ and a matrix of predictors $X \in \RR^{n \times p}$.
-	\begin{enumerate}[(a)]
-		\item First centre $Y$ and all the columns of $X$. Initiale the current model $M \subseteq \qty{1, \dotsc, p}$ to be $\emptyset$ and set the current residual $R$ to be $Y$. 
-		\item Find the variable $k^*$ in $M^c$ most correlated with the current residual $R$.
-Set $M$ to be $M \cup \qty{k^*}$.
-Replace $R$ with the residual from regressing $R$ onto $X_{k^*}$.
-Further replac eeach variable in $M^c$ with the residual from regressing itself onto $X_{k^*}$.
-		\item Continue the previous step unto $R = 0$. 
-	\end{enumerate}
-Show that this algorithm is equivalent to forward selection.
-
-\emph{Hint:} Use induction on the iteration $m$ of the algorithm.
-Consider strengthening the natural induction hypothesis that the model at iteration $m$ is the same as that selected after $m$ steps of forward selection.
-\end{question}
-
-\begin{solution}
-	content...
-\end{solution}
-
-\begin{question}
-	Show that if $W$ is mean-zero and sub-Gaussian with parameter $\sigma$, then $\Var(W) \leq \sigma^2$. 
-\end{question}
-
-\begin{proof}
-	If $W$ is mean-zero, then $\Var(W) = \EE(W^2)$. By the proof of lemma 15 we have $\EE(W^2) \leq 2 \sigma^2$, but this bound is too loose.
-
-	Since $W$ is sub-Gaussian, we have for all $\alpha \in \RR$ that
-	\begin{align*}
-		\EE[e{\alpha W}] &\leq e^{\alpha^2\sigma^2/2} \\
-		\sum_{k=0}^\infty \frac{\EE[W^k] }{k!} \alpha^k &\leq \sum_{k=0}^\infty \frac{\sigma^{2k}}{2^k k!} \alpha^{2k} \\
-		\frac{\EE[W^2]}{2} \alpha^2 + \sum_{k=3}^\infty \frac{\EE[W^k]}{k!} \alpha^k &\leq \frac{\sigma^2}{2} \alpha^2 + \sum_{k=2}^\infty \frac{\sigma^{2k}}{2^k k!} \alpha^{2k} \\
-		\frac12\alpha^2 \qty(\sigma^2 - \EE[W^2]) &\geq \alpha^3 P(\alpha),
-		\end{align*}
-	where $P$ is a power series in $\alpha$. Rescaling we find \[
-	\frac12(\sigma^2 - \EE[W^2]) \geq \alpha P(\alpha),
-	\]
-	and letting $\alpha \to 0$ also lets $\alpha P(\alpha) \to 0$, and therefore $\sigma^2 - \EE[W^2] \geq 0$ so $\Var(W) \leq \sigma^2$. 
-\end{proof}
-
-\begin{question}
-	Verify Hoeffding's lemma for the special case where $W$ is a Rademacher random variable, so $W$ takes the values $-1, 1$ each with probability $1/2$. 
-\end{question}
-
-\begin{proof}
-		If $W$ is a Rademacher random variable then we have, using $(2k)! \geq 2^k k!$ for each $k \in \NN$, that
-	\begin{align*}
-		\EE[e^{\alpha W}] = \frac12 e^{-\alpha} + \frac12 e^{\alpha} = \sum_{k=0}^\infty \frac{\alpha^{2k}}{(2k)!} \leq \sum_{k=0}^\infty \frac{\alpha^{2k}}{2^k k!} = e^{\alpha^2/2}
-	\end{align*}
-which shows that $W$ is sub-Gaussian with parameter $1 = (1 -(-1))/2$, so Hoeffding's lemma holds indeed. 
-\end{proof}
-
-\begin{question}
-	\begin{enumerate}[(a)]
-		\item Let $W \sim \chi_d^2$. Show that
-		\[
-		\PP(\abs{W/d - 1} \geq t) \leq 2e^{-dt^2/8}
-		\]
-		for $t \in (0, 1)$.
-You may use the facts that the mgf of a $\chi_1^2$ random variable is $(1 - 2\alpha)^{-1/2}$ for $\alpha < 1/2$, and $e^{-\alpha} (1 - 2\alpha)^{-1/2} \leq e^{2\alpha^2}$ when $\abs{\alpha} < 1/4$.
-
-		\item Let $A \in \RR^{d\times p}$ have i.i.d.\ standard normal entries. Fix $u \in \RR^p$. Use the result above to conclude that
-		\[
-		\PP\qty(\abs{\frac{\norm{Au}_2^2}{d \norm{u}_2^2} - 1} \geq t) \leq 2e^{-dt^2/8}. 
-		\]
-
-		\item Suppose we have data $u_1, \dotsc, u_n \in \RR^p$, with $p$ large and $n \geq 2$.
-Show that for a given $\eps \in (0, 1)$ and $d > 16\log(n/\sqrt\eps)/t^2$, each data point may be compressed down to $u_i \mapsto Au_i/\sqrt d = w_i$, whilst approximately preserving the distance between the points:
-		\[
-		\PP\qty(1 - t \leq \frac{\norm{w_i - w_j}_2^2}{\norm{u_i - u_j}_2^2} \leq 1 + t \text{ for all } i \neq j \in \qty{1, \dotsc, n}) \geq 1 - \eps. 
-		\]
-		This is the famous Johnson-Lindenstrauss Lemma. 
-	\end{enumerate}
-\end{question}
-
-\begin{proof}
-	\begin{enumerate}[(a)]
-		\item 
-	\end{enumerate}
-\end{proof}
 \end{document}

### 2020-11-16 modern_statistical_methods/C3.tex
Up to speed w/ lecture 17

@@
 	\[
-	\EE\qty[e^{\lambda\norm{\hat\Sigma - \Sigma}_\Rm{op}}] \leq \exp\qty(c \qty[ \frac{\lambda^2\sigma^4}{n} + d]),
+	\EE\qty[e^{\lambda\norm{\hat\Sigma - \Sigma}_\Rm{op}}] \leq \exp\qty(c \qty[ \frac{\lambda^2\sigma^4}{n} + d]) \quad\text{for all $\abs{\lambda} \leq \frac n{16\sigma^2}$},
 	\]
@@
 	\[
-	\PP\qty(\frac{\norm{\hat\Sigma - \Sigma}_\Rm{op}}{\sigma^2} \leq C \qty(\frac{d + \delta}{n} \lor \sqrt{\frac{d + \delta}{n}})) \leq 1 - e^{-\delta}. 
+	\PP\qty(\frac{\norm{\hat\Sigma - \Sigma}_\Rm{op}}{\sigma^2} \leq C \qty(\frac{d + \delta}{n} \lor \sqrt{\frac{d + \delta}{n}})) \geq 1 - e^{-\delta}. 
 	\]
@@
 This is a non-asymptotic error bound. Note that, as long as $d/n \to 0$, the estimator $\hat\Sigma$ will be consistent.
+
+\begin{proof}
+	The tail probability bound follows from Chernoff bounding, see lecture 17 for the details.
+
+	For the proof of (i), cover $S^{d-1}$ with balls of radius $1/8$ centred at points $v_1, \dotsc, v_N \in S^{d-1}$.
+It is a known geometric fact that this is possible with $N \leq e^{4d}/2$.
+We will prove that for any symmetric $Q$ we have $ \norm{Q}_\Rm{op} \leq 2 \max_i \abs{v_i\T Q v_i}$.
+For this, let $v \in S^{d-1}$ and choose $i$ such that $\norm{v - v_i} \leq \frac18$, then
+	\begin{align*}
+	\abs{v\T Qv} &= \abs{(v - v_i)\T Q (v - v_i) + 2 v_i\T Q(v - v_i) - v_i\T Q v_i} \\
+	&\leq \abs{(v - v_i)\T Q(v - v_i)} + 2 \abs{v_i\T Q(v - v_i)} + \abs{v_i\T Q v_i} \\
+	&\leq \norm{v - v_i}^2 \norm{Q}_\Rm{op} + 2 \norm{v_i} \norm{Q}_\Rm{op} \norm{v - v_i} + \abs{v_i\T Q v_i} \\
+	&\leq \qty(\frac28 + \frac1{64}) \norm{Q}_\Rm{op} + \abs{v_i\T Q v_i} \leq \frac12 \norm{Q}_\Rm{op} + \abs{v_i\T Q v_i} \\
+	&\leq \frac12\norm{Q}_\Rm{op} + \max_i \abs{v_i\T Q v_i}. 
+	\end{align*}
+	Since the right-hand side is independent of $v$ we can maximise the left-hand side over $v$ and obtain 
+	\[
+	\norm{Q}_\Rm{op} \leq \frac12\norm{Q}_\Rm{op} + \max_i \abs{v_i\T Q v_i} \implies \norm{Q}_\Rm{op} \leq 2 \max_i \abs{v_i\T Q v_i}. 
+	\]
+
+	Now, we apply this to $Q \ceq \hat\Sigma - \Sigma$. Then we have
+	\begin{equation} \label{eq:crude_moment_bound}
+		\EE[e^{\lambda \norm{Q}_\Rm{op}}] \leq \EE[e^{2\lambda \max_i \abs{v_i\T Qv_i}}] \leq \sum_{i=1}^N \EE[e^{2\lambda v_i\T Q v_i}] + \EE[ e^{-2\lambda v_i\T Q v_i}], 
+		\end{equation}
+	where for the final step we use that $\max_i \abs{v_i\T Q v_i} \in \qty{\pm v_i\T Q v_i \mid i = 1, \dotsc, n}$.
+
+	Next, we claim that for fixed $u \in S^{d-1}$, we have $\EE[e^{tu\T Q u}] \leq \exp(64\frac{t^2\sigma^4}{n})$ for $\abs{t} \leq \frac n{8\sigma^2}$.  For this, observe that 
+	\[
+	u\T Q u = \frac1n \sum_{i=1}^n (u\T x_i)^2 - u\T \Sigma u, 
+	\]
+	and note that $u\T \Sigma u = \Var(u\T x_i) = \EE[(u\T x_i)^2]$ since the $x_i$ have mean zero. Using the fact that the $x_i$ are independent we have
+	\[
+	\EE[\exp(t u\T Q u)] = \EE\qty[\exp(\frac tn \qty[(u\T x_1)^2 - u\T \Sigma u])]^n = \EE[\exp(Z^2 - \EE[Z^2])]^n, 
+	\]
+	where $Z \ceq \sqrt{\frac tn} u\T x_1$ is sub-Gaussian with parameter $\sqrt{\abs{t}\sigma^2/n}$.
+Therefore, $Z^2$ satisfies the Bernstein condition with parmeters $(8\abs{t} \sigma^2/n, 4 \abs{t}\sigma^2/n)$, and recalling that $\abs{t}\leq \frac n{8\sigma^2}$ we obtain
+	\begin{align*}
+		\EE[\exp(Z^2 - \EE[Z^2])]  &\leq 1 + \sum_{k=2}^\infty \frac{\EE[(Z^2 - \EE Z^2)^k]}{k!} \\
+		&\overset\star\leq 1 + \sum_{k=2}^\infty 2^{2k + 1} \sigma^{2k} \qty(\frac{\abs{t}}{n})^k \\
+		&\overset{\star\star}\leq 1 + \frac{2^5 \sigma^4 t^2}{n} \frac{1}{1 - 4\sigma^2 \abs{t}/n} \\
+		&\leq 1 + \frac{64 \sigma^2 t^2}{n} \leq e^{64 \sigma^4 t^2/n},
+	\end{align*}
+where $\star$ follows from Bernstein's inequality and $\star\star$ from a geometric series, which proves the claim.
+
+Substituting this back into \cref{eq:crude_moment_bound} using $t =  \pm 2\lambda$ shows that for $\abs{\lambda} < \frac{n}{16\sigma^2}$ we have
+\[
+	\EE[e^{\lambda \norm{Q}_\Rm{op}}] \leq 2N e^{256 \lambda^2 \sigma^4/n} \leq  e^{256\lambda^2\sigma^4/n + 4d} \leq  e^{256(\lambda^2 \sigma^4/n + d)}. 
+\]
+\end{proof}
+
+The upper bound in this theorem is tight when $\Sigma = I$.
+However, note that our bound in \cref{eq:crude_moment_bound} is quite crude, and we can obtain much better rates if we assume that the $x_i$ are normally distributed (in this case, more is known about the supremum).
+
+Furthermore, when the spectrum of $\Sigma$ is concentrated on a few directions, the rate can be improved significantly. For this, we need a definition.
+\begin{definition}
+	The \emph{effective rank} $r(\Sigma)$ of a nonzero symmetric positive semi-definite matrix $\Sigma$ with eigenvalues $0 \leq \gamma_d(\Sigma) \leq \dotsb \leq \gamma_1(\Sigma)$ is given by
+	\[
+	r(\Sigma) = \frac{\Tr(\Sigma)}{\norm{\Sigma}_\Rm{op}} = \frac{\sum_{i=1}^d \gamma_i(\Sigma)}{\gamma_1(\Sigma)}. 
+	\]
+\end{definition}
+It is easily seen that $r(\Sigma) \leq \Rm{rank}(\Sigma)$, with equality if and only if all nonzero eigenvalues of $\Sigma$ are equal.
+The following theorem (which we will not prove) gives a tighter bound if we can assume that the rows of $X$ are normally distributed and that $r(\Sigma)$ is constrained:
+\begin{theorem}
+	Suppose $X$ has i.i.d.\ $N(0, \Sigma)$ rows with $r(\Sigma) \leq cn$ (??). Then there exists $C$ such that for all $\delta > 0$ we have
+	\[
+	\PP\qty(\frac{\norm{\hat\Sigma - \Sigma}_\Rm{op}}{\norm{\Sigma}_\Rm{op}} \leq C \qty(\frac{r(\Sigma) + \delta}{n} \lor \sqrt{\frac{r(\Sigma) + \delta}{n}})) \geq 1 - e^{-\delta}. 
+	\]
+\end{theorem}
+
+In comparison with the previous bound, the dimension of the samples $d$ is replaced by the effective rank $r(\Sigma)$ (in fact, $d$ does not occur at all in this theorem!).
+This means that $d$ can grow rapidly with $n$, as long as $r(\Sigma)/n \to 0$.
+This can even be generalised to an infinite-dimensional setting with appropriate definitions.

### 2020-11-16 distribution_theory/defs.tex
aaarhhh

\setcounter{section}{-1}
\section{Most important definitions}
\paragraph{Spaces of test functions} a function $f \in X \to \CC$ is in: 
\begin{enumerate}
	\itemsep=0em
	\item $\DD(X)$ if $f$ is smooth and $\supp f \subseteq X$ is compact;
	\item $\SS(\RR^n)$ if $f$ is smooth and $\norm{f}_{\alpha, \beta} \ceq \sup_{x \in \RR^n} \abs{x^\alpha (D^\beta f)(x)} < \infty$ for all $\alpha, \beta$;
	\item $\EE(X)$ if $f$ is smooth. 
\end{enumerate}
Note $\DD(\RR^n) \subseteq \SS(\RR^n) \subseteq \EE(\RR^n)$. 
We have the following modes of convergence in these spaces:
\begin{enumerate}
	\itemsep=0em
	\item $\phi_m \to0$ in $\DD(X)$ if there exists a compact $K \subseteq X$ with $\supp\phi_m \subseteq K$ for all $m$, and $\pt^\alpha \phi_m \to 0$ uniformly for each $\alpha$;
	\item $\phi_m \to0$ in $\SS(\RR^n)$ if $\norm{\phi_m}_{\alpha, \beta} \to 0$ for all $\alpha, \beta$;
	\item $\phi_m \to0$ in $\EE(X)$ if $\pt^\alpha \phi_m \to 0$ uniformly on compact subsets of $X$ for all $\alpha$. 
\end{enumerate}

\paragraph{Spaces of distributions}
The continuous linear maps from $\DD(X)$, $\SS(X)$, and $\EE(X)$ to $\CC$  are called \emph{distributions}, \emph{tempered distributions}, and \emph{compactly supported distributions} respectively, and these spaces are denoted $\DD'(X)$, $\SS'(\RR^n)$, $\EE'(X)$, employed with weak-$*$ convergence. Note that $\EE'(\RR^n) \subseteq \SS'(\RR^n) \subseteq \DD'(\RR^n)$. We have the following characterisations:
\begin{enumerate}
	\item $u \in \DD'(X)$ iff for every compact $K \subseteq X$ there exist non-negative $C, N$ such that for all $\phi \in \DD(X)$ with $\supp(\phi) \subseteq K$ we have 
	\[
	\abs{\ang{u, \phi}} \leq C \sum_{\abs{\alpha} \leq N} \sup_x \abs{\pt^\alpha\phi(x)}. 
	\]

	\item $u \in \SS'(\RR^n)$ iff there exist constants $C, N$ such that for all $\phi \in \SS(\RR^n)$ we have
	\[
	\abs{\ang{u, \phi}} \leq C \sum_{\abs{\alpha}, \abs{\beta} \leq N} \norm{\phi}_{\alpha, \beta}. 
	\]

	\item $u \in \EE'(X)$ iff there exists a compact $K\subseteq X$ and non-negative $C, N$ such that 
	\[
	\abs{\ang{u, \phi}} \leq C \sum_{\abs{\alpha} \leq N} \sup_{x \in K} \abs{\pt^\alpha \phi(x)}.
	\]
\end{enumerate}

\paragraph{Basic operations}
We define the following basic operations: 
\begin{enumerate}
	\itemsep=0em
	\item if $f$ is smooth, then $\ang{fu, \phi} \ceq \ang{u, f\phi}$ (for Schwarz functions, we must have that $f\phi \in \SS(\RR^n)$ for all $\phi \in \SS(\RR^n)$);
	\item For a distribution $u$: $\ang{\pt^\alpha u, \phi} \ceq (-1)^{\abs{\alpha}} \ang{u, \pt^\alpha\phi}$;
	\item For a test function $\phi$ we define $(\tau_h\phi)(x) = \phi(x - h)$, and for a distribution $u$ we then define $\ang{\tau_h u, \phi} \ceq \ang{u, \tau_{-h}\phi}$. 
	\item For a test function $\phi$ we define $\Rr[\phi](x) = \refl\phi(x) \ceq \phi(-x)$, and for a distribution $u$ we then define $\ang{\Rr[u], \phi} = \ang{\refl u, \phi} \ceq \ang{u, \refl\phi}$.
\end{enumerate}

\paragraph{Convolution} 
\begin{enumerate}
	\item For $u \in C^\infty(X)$, $\phi \in \DD(X)$, we define 
	\[
	(u * \phi)(x) \ceq \int_{\RR^n} u(x-y) \phi(y) \dd{y} = \int_{\RR^n} u(y) \phi(x-y) \dd{y} = \ang{u, \tau_x \refl\phi}.
	\]
	\item For $u \in \DD'(\RR^n)$, $\phi \in \DD(\RR^n)$, (or $u \in \EE'(\RR^n)$, $\phi \in \EE(\RR^n)$, or $u \in \SS'(\RR^n)$, $\phi \in \SS(\RR^n)$), we define
	\[
	(u* \phi)(x) \ceq \ang{u, \tau_x \refl\phi}. 
	\]
	It can be shown that $u * \phi$ is smooth, and that $\ang{u, \phi} = (u * \refl\phi)(0)$. 
	\item For $u \in \DD'(\RR^n), v \in \EE'(\RR^n)$ or $u \in \EE'(\RR^n), v \in \DD'(\RR^n)$, define $u * v \in \DD'(\RR^n)$  by the property
	\[
	(u * v) * \phi = u * (v * \phi) \qquad\forall \phi\in \DD(\RR^n). 
	\]
\end{enumerate}

\paragraph{Fourier transform}
\begin{enumerate}
	\item For $f \in L^1(\RR^n)$, define the Fourier transform by
	\[
	\FF[f](\lambda) = \hat f(\lambda) \ceq \int_{\RR^n} e^{-i\lambda\vdot x} f(x) \dd{x}. 
	\]
	It is known that $\FF$ is a continuous bijection from $\SS(\RR^n)$ to itself with inverse
	\[
	\FF^{-1}[\hat f](x) \ceq \frac1{(2\pi)^n} \int_{\RR^n} e^{i\lambda \vdot x} \hat f(\lambda) \dd{\lambda}. 
	\]
	Note that we can write $\FF^{-1} = \frac1{(2\pi)^n} \Rr \FF = \frac1{(2\pi)^n} \FF \Rr$.

	\item For $u \in \SS'(\RR^n)$, we define the Fourier transform of $u$ by 
	$\ang{\FF[u], \phi} = \ang{\hat u, \phi} \ceq \ang{u, \hat\phi}$. It is known that $\FF$ extends to a continuous bijection from $\SS'(\RR^n)$ to itself, with inverse $\FF^{-1} = (2\pi)^{-n} \Rr \FF = (2\pi)^{-n} \FF \Rr$. 
\end{enumerate}

\paragraph{Sobolev space}
We define $\ang{\lambda} \ceq \sqrt{1 + \norm{\lambda}^2}$ for $\lambda \in \RR^n$, and note that $\ang{\lambda} \sim \norm{\lambda}$ for large $\lambda$.

For $s \in \RR$, we define the \emph{Sobolev space} $H^s(\RR^n)$ as the set of tempered distributions $u \in \SS'(\RR^n)$ for which $\hat u$ can be identified with a measurable function $\hat u(\lambda)$ such that 
\[
\int_{\RR^n} \ang{\lambda}^s \hat u(\lambda) \dd{\lambda} < \infty. 
\]

### 2020-11-18 distribution_theory/distribution_ES2.tex
aaarhhh
hellow

@@
 \newcommand\vd{\vdot}
+\newcommand\Rr{\Cal R}
 %\newcommand\ZZ{\Bb Z}
@@
 \maketitle
+
+\begin{mdframed}
+	We will write $\Rr$ and $\FF$ for the reflection and Fourier transform operators. 
+\end{mdframed}
 \begin{question}
@@
 	\end{equation}
-	To show that this is uniquely defined, recall that for all $u \in \DD'(\RR^n) \phi\in \DD(\RR^n)$, we have $\ang{u, \phi} = (u * \refl\phi)(0)$. Therefore, we have
+Recall that for all $u \in \DD'(\RR^n) \phi\in \DD(\RR^n)$, we have $\ang{u, \phi} = (u * \refl\phi)(0)$, and therefore $u * v$ should satisfy
 \[
-	\ang{u * v, \phi} = ((u * v) * \refl\phi)(0) = (u * (v * \refl\phi))(0),
+\ang{u * v, \phi} = ((u * v) * \refl\phi)(0) = (u * (v * \refl\phi))(0) = \ang{u, \Rr(v * \refl\phi)}. 
+\] Therefore 
+\begin{enumerate}
+	\item Define $w \colon \DD(\RR) \to \CC$ by 
+	\[
+	\ang{w, \phi} = \ang{u, \Rr(v * \refl\phi)}, 
 	\]
-	which shows that the formula \cref{def:convolution} uniquely defines $\ang{u * v, \phi}$ for any $\phi \in \DD(\RR^n)$, and therefore that $u * v$ is well-defined.
+	then we will show that $w$ satisfies $w * \phi = u * (v * \phi)$ for all $\phi \in \DD(\RR)$. 
+	\begin{align*}
+		(w * \phi)(x) &= \ang{w, \tau_x\refl\phi} = \ang{u, \Rr(v * \Rr(\tau_x \refl\phi))} = \ang{u, \Rr(v * \tau_{-x}\phi)} \\
+		&= \ang{u, \Rr \tau_{-x} (v * \phi)} = \ang{u, \tau_x \Rr(v * \phi)} = (u * (v * \phi))(x). 
+	\end{align*}
+\item Uniqueness: we have shown in the lectures that if $w * \phi = w' * \phi$ for all $\phi$, then $w = w'$. This shows that \cref{def:convolution} uniquely defines $u * v$. 
+\end{enumerate}
+%	\[
+%	\ang{u * v, \phi} = ((u * v) * \refl\phi)(0) = (u * (v * \refl\phi))(0),
+%	\]
+%	which shows that the formula \cref{def:convolution} uniquely defines $\ang{u * v, \phi}$ for any $\phi \in \DD(\RR^n)$, and therefore that $u * v$ is well-defined. For uniqueness,
 
@@
 
-	\TODO Give proof.
+	(??)
 
@@
 	\end{align*}
+
+(??)
+\end{proof}
+
+\begin{question}
+	Prove that multiplication by a Schwarz function gives rise to a continuous map from $H^s(\RR^n)$ to itself, i.e., $\norm{\phi u}_{H^s} \les \norm{u}_{H^s}$ for $\phi \in \SS(\RR^n)$.
+You may assume Peetre's inequality: for $\lambda, \nu \in \RR^n$ and $s \in \RR$
+	\[
+	\qty(\frac{1 + \norm{\lambda}^2}{1 + \norm{\mu}^2})^s \leq 2^{\abs{s}} \qty(1 + \norm{\lambda - \mu}^2)^{\abs{s}}. 
+	\]
+\end{question}
+
+\begin{proof}
+	We have
+	\begin{align*}
+		\norm{\phi u}_{H^s}^2 &= \int \abs{\FF[\phi u](\lambda)}^2 (1 + \norm{\lambda})^2 \dd{\lambda}
+	\end{align*}
+(??) (how to bound fourier transform of product??)
 \end{proof}

### 2020-11-18 modern_statistical_methods/C3.tex
aaarhhh
hellow

@@
 \begin{theorem}
-	Suppose $X$ has i.i.d.\ $N(0, \Sigma)$ rows with $r(\Sigma) \leq cn$ (??). Then there exists $C$ such that for all $\delta > 0$ we have
+	Suppose $X$ has i.i.d.\ $N(0, \Sigma)$ rows with $r(\Sigma) \leq cn$. Then there exists $C$ (depending on $c$) such that for all $\delta > 0$ we have
 	\[
@@
 This can even be generalised to an infinite-dimensional setting with appropriate definitions.
+
+\subsubsection{The graphical Lasso}
+We saw that if we make the structural assumption that the effective rank of $\Sigma$ is small, the sample covariance matrix $\hat\Sigma$ is a good estimator.
+However, there are other possible structural assumptions.
+One is based on the following:
+\begin{proposition}
+	Let $Z \sim N(\mu, \Omega^{-1})$. Then $z_k \perp z_j \mid z_{-jk}$ ($z_k$ is independent of $z_j$ conditional on the other $z_i$) if and only if $\Omega_{jk} =0 $.  
+\end{proposition}
+
+\begin{proof}
+	We have
+	\begin{align*}
+		f(z_j, z_k \mid z_{-jk}) &= \frac{f(z_1, \dotsc, z_n)}{f(z_{-jk})} \propto f(z_1, \dotsc, z_n) \propto \exp(-\frac12 (z-\mu)\T \Omega(z-\mu)) \\
+		&= g(z_j) h(z_k) \exp(- z_j z_k \Omega_{jk}), 
+	\end{align*}
+and this factorises if and only if $\Omega_{jk} = 0$. 
+\end{proof}
+
+Assuming that many variables are independent of each other conditional on the other variables (or in other words, that $z_k$ and $z_j$ depend on each other only ``through the other $z_i$''), is a fairly natural assumption, which means it is sensible to assume that $\Omega$ is sparse.
+If this is the case, $\hat\Sigma$ may not be the best estimator.
+
+The \emph{graphical lasso} explicitly penalises estimators of $\Omega$ with many non-zero entries, and is defined as
+\[
+\hat\Omega^\Rm{GL} \ceq \argmin_{\Omega \succ 0} \qty(-\log(\det\Omega) + \Tr(\hat\Sigma\Omega) + \lambda \norm{\Omega}_1), 
+\]
+where the 1-norm of a matrix is defined as the sum of the absolute values of its entries.
+This is exactly minus the log-likelihood plus a penalty term, and finding $\hat\Omega^\Rm{GL}$ is a convex optimisation problem which generally leads to sparse estimators.
+
+Note that when we penalise $\norm{\Omega}_1$, we are also penalising the diagonal entries, which have nothing to do with conditional independence.
+Therefore, sometimes we penalise only $\sum_{j \neq k} \abs{\Omega_{jk}}$.
+
+\subsection{Principal Component Analysis (PCA)}
+Let $X \in \RR^{n\times d}$ have i.i.d.\ rows with mean 0 and covariance $\Sigma$.
+Let $\lambda_1 \geq \dotsb \geq \lambda_d \geq 0$ be the eigenvalues of $\Sigma$ with eigenvectors $u_1, \dotsc, u_ d\in S^{d-1}$.
+For any $s \leq d$ we have
+\[
+\Span(u_1, \dotsc, u_s) = \argmax_{\dim(S) = s} \EE[ \norm{P_S x_1}^2] = \argmin_{\dim(S) = s} \EE[\norm{x_1 - P_S x_1}^2], 
+\]
+i.e., $\Span(u_1, \dotsc, u_s)$ is the $s$-dimensional subspace of $\RR^d$ with maximum variance and minimal residual variance.
+
+The quantity $q_s = \frac{\sum_{i=1}^s \lambda_i}{\Tr(\Sigma)}$ is the proportion of variance captured by the largest $s$ eigenvalues, and this may be close to 1 for $s \ll d$.
+The idea of PCA is to approximate the eigenvalues and eigenvectors of $\Sigma$ by those of $\hat\Sigma = \frac1n (X - \bar X \mathbf{1})\T (X - \bar X \mathbf{1}) \ceq \frac1n \tilde X\T \tilde X$.
+
+The main question is: how robust are eigenvectors and eigenvalues under perturbation of the data?
+
+\subsubsection{Perturbation of eigenvalues and eigenvectors}
+Let $E = \hat\Sigma - \Sigma$, and suppose $E$ is a deterministic perturbation. We have proven bounds on $\norm{E}_\Rm{op}$. It is easily seen that
+\[
+\hat{\lambda_1} = \max_{v \in S^{d-1}} v\T (\Sigma + E)v \leq\max_{v \in S^{d-1}} v\T \Sigma v + \max_{v \in S^{d-1}} v\T E v  \leq \lambda_1 + \norm{E}_\Rm{op}, 
+\]
+and analogously it can be seen that $\lambda_1 \leq \hat\lambda_1 + \norm{E}_\Rm{op}$, so we obtain $\abs{\lambda_1 - \hat\lambda_1} \leq \norm{E}_\Rm{op}$. This can be generalised:
+\begin{theorem}[Weyl]
+	We have $\max_i \abs*{\hat\lambda_i - \lambda_i} \leq \norm{E}_\Rm{op}$. 
+\end{theorem}
+
+We conclude that eigenvalues are generally robust to small perturbations.
+However, eigenvectors are genearlly not robust to small perturbations.
+This is because when $\lambda_i$ and $\lambda_{i+1}$ are close to each other, then $\Span(\hat u_i, \hat u_{i+1})$ may be a good approximation of $\Span(u_i, u_{i+1})$, but it is not possible to identify the individual eigenvectors.
+Therefore we must place further assumptions on the eigenvectors.

### 2020-11-18 topics_statistical_theory/C3.tex
hellow

@@
 \end{proof}
+
+Now that we have bounds on the variance and bias, we can prove our uniform bound: 
+\begin{theorem}
+	Under the conditions of the previous two propositions, if we choose $h = \alpha n^{-1/(2\beta + 1)}$ for some $\alpha > 0$, there exists $n_0 \in \NN, C > 0$, depending only on $\beta, L, \alpha, K, \sigma_\Rm{max}^2$, such that
+	\[
+	\sup_{m \in \HH(\beta, L)} \sup_{x_0 \in [0, 1]} \EE\qty[ \qty{\hat m_n(x_0;p) - m(x_0)}^2] \leq Cn^{-2\beta/(2\beta + 1)}.
+	\]
+\end{theorem}
+
 %Suppose that $X\T W X$ is positive definite (this is a very weak condition), then we have
@@
 %which is minimised at $\hat\beta = (X\T W X)^{-1} X\T WY$.
+
+\subsection{Splines}
+\subsubsection{Cubic splines}
+Let $n \geq 3$ and $a \leq x_1 < \dotsb < x_n \leq b$.  
+\begin{definition}
+	A function $g \colon [a, b] \to \RR$ is called a \emph{cubic spline} with \emph{knots} $x_1, \dotsc, x_n$ if 
+	\begin{enumerate}
+		\item $g$ is a cubic polynomial on each interval $(a, x_1), (x_1, x_2), \dotsc, (x_n, b)$;
+		\item $g \in C^2[a, b]$. 
+	\end{enumerate}
+	Furthermore, $g$ is called \emph{natural} if it is linear on $[a, x_1]$ and $[x_n, b]$, (i.e., $g''(a) = g'''(a) = g''(b) = g'''(b) = 0$). 
+\end{definition}
+
+We often represent a natural cubic spline by the vectors $\vec g \in \RR^n$ with $g_i = g(x_i)$, and $\vec \gamma \in \RR^{n-2}$ with $\gamma_i = g''(x_i)$ (excluding $\gamma_1$ and $\gamma_n$).
+Writing $h_i \ceq x_{i+1} - x_i$ we have for $x \in [x_i, x_{i+1}]$ that
+\[
+g(x) = \frac{(x - x_i) g_{i+1} - (x_{i+1} - x) g_i}{h_i} - \frac16(x - x_i)(x_{i+1} - x) \qty{\qty(1 + \frac{x - x_i}{h_i})\gamma_{i+1} + \qty(1 + \frac{x_{i+1} - x}{h_i}\gamma_i)}.
+\]
+\begin{proposition}
+	Given $\vec g \in \RR^n$, there exists a unique natural cubic spline $g$ with knots at $x_1, \dotsc, x_n$ satisfying $g(x_i) = g_i$ for all $i$, and there exists $K \succeq 0$ (depending on $x_1, \dotsc, x_n$) such that
+	\[
+	\int_a^b g''(x)^2 \dd{x} = \vec g\T K \vec g. 
+	\]
+\end{proposition}
+We call $g$ the \emph{natural cubic spline interpolant to $\vec g$ at $x_1, \dotsc, x_n$}.
+
+\begin{definition}
+	We define $\SS_2[a, b]$ as the set of real-valued functions on $[a, b]$ with an absolutely continuous first derivative.
+For $f \in \SS_2[a, b]$, we define the \emph{roughness} of $f$ by $R(f) \ceq \int_a^b f''(x)^2 \dd{x}$.
+\end{definition}
+
+\begin{proposition}
+	For any $\vec g \in \RR^n$, the natural cubic spline interpolant to $\vec g$ at $x_1, \dotsc, x_n$ is the unique minimiser of $R$ over all $g \in \SS_2[a, b]$ that satisfy $g(x_i) = g_i$ for all $i$.
+\end{proposition}
+
+\subsubsection{Natural cubic smoothing splines}
+Consider the nonparametric regression model $Y_i = g(x_i) + \sigma \eps_i$, where $\EE[\eps_i] = 0$ and $\Var[\eps_i] = 1$.
+A way to estimate a nonparametric regression function is to balance data fidelity against roughness of the curve, which can be done by minimising
+\[
+S_\lambda(g) \ceq \sum_{i=1}^n (Y_i - g(x_i))^2 + \lambda R(g),
+\]
+where $\lambda > 0$ is a regularisation parameter.
+For small $\lambda$, this is almost an exact fit to the data.
+For large $\lambda$, we are basically minimising $\abs{g''}$, which means we will approximate the linear regression fit.
+
+\begin{theorem}
+For each $\lambda \in (0, \infty)$, there exists a unique minimiser $\hat g_\lambda$ of $S_\lambda$ over $\SS_2[a, b]$.
+It is the natural cubic spline with knots at $x_1, \dotsc, x_n$ and $\vec g = (I + \lambda K)^{-1} \vec Y$.
+\end{theorem}
+
+\begin{proof}
+	If $g$ is not a natural cubic spline, we know that the natural cubic spline $g^*$ which interpolates $g(x_1), \dotsc, g(x_n)$ at $x_1, \dotsc, x_n$ has a strictly lower value of $S_\lambda$, so we know the minimiser must be a natural cubic spline.
+
+	If $g$ is a natural cubic spline, then there exists $K \succeq 0$ such that
+	\begin{align*}
+		S_\lambda(g) &= (\vec Y - \vec g)\T (\vec Y - \vec g) + \lambda \vec g\T K \vec g \\
+		&= \vec g\T (I + \lambda K)\vec g - 2 \vec Y\T \vec g + \vec Y\T\vec Y,
+	\end{align*}
+and by ``completing the square'' we write, for some $Z$ independent of $\vec g$, 
+\[
+S_\lambda(g) = \qty(\vec g - (I + \lambda K)^{-1} \vec Y)\T (I + \lambda K) \qty(\vec g - (I + \lambda K)^{-1} \vec Y) + Z
+\]
+Since $I + \lambda K$ is positive definite it follows that $\vec g = (I + \lambda K)^{-1} \vec Y$ gives the minimiser. 
+\end{proof}

### 2020-11-18 distribution_theory/distribution_ES2_handin.tex
opirseajf

--- 2020-11-18 distribution_theory/distribution_ES2.tex
@@
 \ang{u * v, \phi} = ((u * v) * \refl\phi)(0) = (u * (v * \refl\phi))(0) = \ang{u, \Rr(v * \refl\phi)}. 
-\] Therefore 
+\] 
+We can now show existence and uniqueness:
 \begin{enumerate}
-	\item Define $w \colon \DD(\RR) \to \CC$ by 
+	\item Existence: define $w \colon \DD(\RR) \to \CC$ by 
 	\[
@@
 
-\begin{question}
-	Show that if $u, v, w \in \DD'(\RR^n)$ and at least two of them have compact support, then the convolution is associative (i.e., $(u * v) * w) = u* (v * w)$). 
-\end{question}
-
-\begin{proof}
-	Note that the convolution between two compactly supported distributions is again compactly supported, which ensures that both expressions `make sense'. Now, let $\phi \in \DD(\RR^n)$, then we have
-	\begin{align*}
-		((u * v) * w) * \phi &= (u* v) * (w * \phi) = u * (v * ( w* \phi)) = u * ((v * w) * \phi) = (u * (v * w)) * \phi,
-	\end{align*}
-	which proves the theorem. 
-\end{proof}
-
-\begin{question}
-	Let $\phi \in \DD(\RR)$ and choose $\eps > 0$ sufficiently small so that $\supp(\phi) \subset I_\eps = (-1/\eps, 1/\eps)$.
-Given that $\phi$ has a uniformly convergent Fourier series on $I_\eps$ in the form
-	\[
-	\phi(x) = \sum_{n \in \ZZ} c_n e^{i \eps \pi n x}, \quad c_n = \frac\eps2 \int_\RR \phi(x) e^{-i\eps \pi n x} \dd{x},
-	\]
-	prove the Fourier inversion theorem on $\DD(\RR)$ by taking a suitable limit. 
-\end{question}
-
-\begin{proof}
-Since $\DD(\RR) \subseteq \SS(\RR)$, we know that the Fourier inversion formula holds. We only need to show that the Fourier tranform of $\phi$ is again an element of $\DD(\RR)$. (??)
-\end{proof}
-
-\begin{question}
-	For $\phi \in \SS(\RR^n)$ prove that $\sum_m \phi(m) = \sum_n \hat\phi(2\pi n)$. This is the famous Poisson summation formula. 
-\end{question}
-
-\begin{proof}
-	We have
-	\begin{align*}
-		\sum_m \phi(m) = \frac1{(2\pi)^n} \sum_m \int e^{i\lambda m} \hat \phi(\lambda) \dd{\lambda} = \sum_m \int e^{2\pi i \lambda m} \hat \phi(2\pi \lambda) \dd{\lambda}
-	\end{align*}
-(??)
-\end{proof}
-
-\begin{question}
-	If $u \in H^s(\RR^n)$ show that $D^\alpha u\in H^{s - \abs{\alpha}} (\RR^n)$. If $s > t$ show that $H^s(\RR^n) \subseteq H^t(\RR^n)$. 
-\end{question}
-
-\begin{proof}
-	Assuming $u \in H^s(\RR^n)$, we have
-	\begin{align*}
-		\int_{\RR^n} \ang{\lambda}^{2(s - \abs{\alpha})} \abs{\widehat{D^\alpha u}(\lambda)}^2 \dd{\lambda}  &= \int_{\RR^n} \ang{\lambda}^{2(s-\abs{\alpha})} \norm{\lambda}^{2\abs{\alpha}} \abs{\hat u(\lambda)}^2 \dd{\lambda} \\
-		&\les \int_{\RR^n} \ang{\lambda}^{2s} \abs{\hat u(\lambda)}^2 \dd{\lambda} < \infty, 
-	\end{align*}
-which proves $D^\alpha u \in H^{s- \abs{\alpha}}(\RR^n)$.
-
-The second claim follows immediately from the fact that $\ang{\lambda}^{t} \leq \ang{\lambda}^s$ for $s \geq t$ and $\lambda$ sufficiently large. 
-\end{proof}
-
-\begin{question}
-	Show that $S(\RR^n)$ is dense in $L^2(\RR^n) = H^0(\RR^n)$ and deduce that $\SS(\RR^n)$ is dense in $H^s(\RR^n)$.
-
-	\emph{Hint:} Use Parseval's theorem. 
-\end{question}
-
-\begin{proof}
-	We will show that $\DD(\RR^n)$ is dense in $L^2(\RR^n)$: since $\DD(\RR^n) \subseteq \SS(\RR^n)$, this shows that $\SS(\RR^n)$ is dense in $L^2(\RR^n)$ as well.
-
-	(??)
-
-	Now, for $u \in H^s(\RR^n)$, let $(\phi_m) \to u$ in $L^2$. Then we have
-	\begin{align*}
-		\norm{\phi_m - u}_{H^s}^2 = \int \ang{\lambda}^{2s} \abs{(\phi_m - u)(\lambda)}^2 \dd{\lambda}
-	\end{align*}
-
-(??)
-\end{proof}
-
-\begin{question}
-	Prove that multiplication by a Schwarz function gives rise to a continuous map from $H^s(\RR^n)$ to itself, i.e., $\norm{\phi u}_{H^s} \les \norm{u}_{H^s}$ for $\phi \in \SS(\RR^n)$.
-You may assume Peetre's inequality: for $\lambda, \nu \in \RR^n$ and $s \in \RR$
-	\[
-	\qty(\frac{1 + \norm{\lambda}^2}{1 + \norm{\mu}^2})^s \leq 2^{\abs{s}} \qty(1 + \norm{\lambda - \mu}^2)^{\abs{s}}. 
-	\]
-\end{question}
-
-\begin{proof}
-	We have
-	\begin{align*}
-		\norm{\phi u}_{H^s}^2 &= \int \abs{\FF[\phi u](\lambda)}^2 (1 + \norm{\lambda})^2 \dd{\lambda}
-	\end{align*}
-(??) (how to bound fourier transform of product??)
-\end{proof}
-%-----------------------------------------------------------------------------
 \setcounter{question}{18}
@@
 
-	\item The derivative of this function, outside of 0, is $\log\abs{x}$.
+	\item The derivative of this function is $\log(\abs{x})$, and the derivative of $\log(\abs{x})$ is $\PV(1/x)$ which we saw in the previous example sheet.
+By (a) we know that the Fourier transform of $\sign$ is $2i \PV(1/x)$, so we have
+	\[
+	\FF[2i \PV(1/x)] = \FF[\FF[\sign]] = (2\pi)^n \refl\sign \implies \FF[\PV(1/x)] = \frac{(2\pi)^n}{2i} \refl\sign.
+	\]
+	Note that $\refl\sign = -\sign$ we conclude
+	\[
+	\FF[x\log\abs{x}- x](\lambda) = \frac{(2\pi)^n}{2i\lambda^2} \sign(\lambda). 
+	\]
 

### 2020-11-22 inverse_problems/C3.tex
afwejinuo

@@
 \begin{lemma}
-	Let $E \colon \XX \to \rext$ be convex, then $E$ is subdifferentiable at all points $u \in \dom(E)$. If $E$ is also proper, then $E$ is not subdifferentiable at any $u \notin \dom(E)$. 
+	Let $E \colon \XX \to \rext$ be convex, then $E$ is subdifferentiable at all points $u \in \dom(E)$ where $E$ is continuous.
+If $E$ is also proper, then $E$ is not subdifferentiable at any $u \notin \dom(E)$.
 \end{lemma}
@@
 We have the following:
-\begin{theorem}
+\begin{theorem} \label{thm:primal_dual_gap}
 	Suppose the function $E(Au) + F(u)$ is proper, convex, LSC and coercive.

### 2020-11-22 inverse_problems/C5.tex
afwejinuo

\section{Bayesian probability and statistics}
\subsection{Motivation}
Let $\XX, \YY$ be separable Banach spaces and $A \colon \XX\to\YY$ a measurable operator (possibly nonlinear). We consider the inverse problem
\[
\text{Find $u \in \XX$ such that } A(u) + n = f_n,
\]
where $f_n \in \YY$ is data and $n \in \YY$ is noise.

In Bayesian probability theory, we attempt to give an estimate for $u$ while also commenting on the uncertainty left in $u$ after measurement. The main idea is to \textbf{model uncertain parameters as random variables}. In this case, we assume the noise $n$ and the parameter $u$ are random variables $N, U$. The distribution of $N$ is usually known, while the distribution of $U$ (called the \emph{prior distribution $\mu_0 = \PP(U \in \cdot)$}) describes our prior knowledge of $u$. We then observe data $f_n$, which is described as an occurrence of the event $\qty{f_n = A(U) + N}$. After our observation, we replace the distribution $\mu_0$ by the distribution
\[
\mu_\Rm{post} \ceq \PP(U \in \cdot \mid f_n = A(U) + N). 
\]

\subsection{Measure theory}
\begin{definition}
	Let $\Omega$ be a topological space, and let $\Cal O$ be the topology on $\Omega$. Then the $\sigma$-algebra generated by $\Cal O$ is denoted $\BB\Omega$ and called the \emph{Borel $\sigma$-algebra} on $\Omega$. 
\end{definition}
Note that we do not explicitly state the topology used when writing $\BB\Omega$, but this is always clear from context.

\begin{definition}
	Let $(\Omega, \FF, \mu)$ be a measure space, $(\Omega', \FF')$ a measurable space, and $g \colon (\Omega, \FF) \to (\Omega', \FF')$ measurable. Then, the map $\mu \circ g^{-1} = \mu(g \in \cdot)$ defines a measure on $(\Omega', \FF')$ called the \emph{pushforward measure}.

	Note that if $\mu$ is a probability measure, then so is $\mu(g \in \cdot)$. 
\end{definition}

\begin{definition}
	Let $(\Omega, \FF)$ be a measurable space and $\mu, \nu$ two measures on $\Omega$. Then $\nu$ is called \emph{absolutely continuous} with respect to $\mu$, denoted $\nu \ll \mu$, if \[\mu(F) = 0 \implies \nu(F) =0 \qquad\text{for all $F \in \FF$}. 
\]
\end{definition}

\begin{theorem}[Radon-Nikodym]
	Let $(\Omega, \FF)$ be a measurable space and let $\mu, \nu$ be $\sigma$-finite measures on $(\Omega, \FF)$. Then, the following are equivalent:
	\begin{enumerate}
		\item $\nu \ll \mu$;
		\item There is a measurable $g \colon (\Omega, \FF) \to (\RR, \BB\RR)$, with
		\[
		\nu(F) = \int_F g \dd{\mu} \qquad\text{for all $F \in \FF$}. 
		\]
	\end{enumerate}
The function $g$ is $\mu$-a.e.\ unique, and is called the \uline{Radon-Nikodym derivative}, denoted $\frac{\dd{\nu}}{\dd{\mu}}$. 
\end{theorem}

\subsection{Conditional probability}
We will consider $(\Omega, \FF, \PP)$ as underlying probability space for any random variable, and assume $\Omega$ is separable and completely metrisable (this is also called \emph{Polish}) and $\FF = \BB\Omega$. We also write
\[
\EE[\phi] \ceq \int_\Omega \phi \dd{\PP}
\]
for any $\phi \in L^1(\Omega)$.

\begin{theorem}
	Let $U \colon (\Omega, \FF) \to (\RR, \BB\RR)$ and $Y \colon (\Omega, \FF) \to (\YY, \BB\YY)$ be random variables and let $U$ be integrable. Then there exists a measurable function $h \colon (\YY, \BB\YY) \to(\RR, \BB\RR)$ such that
	\[
	\int_F h(y) \PP(Y \in \dd{y}) = \int_{\qty{Y\in F}} U \dd{\PP} \qquad\text{for all $F \in \FF$}. 
	\]
	The function $h$ is $\PP(Y \in \cdot)$-a.s.\ unique.
\end{theorem}

\begin{definition}
	For $y \in \YY$, we call $h(y)$ (from the previous theorem) the \uline{conditional expectation of $U$ given $Y = y$}, denoted $h(y) \eqqcolon \EE[U \mid Y = y]$ (which is well-defined up to $\PP(Y \in \cdot)$ null sets).

	For an event $F \in \FF$, we define the \emph{probability of $F$ given $Y = y$} as $\EE[\ind_F \mid Y = y]$.
\end{definition}

\begin{definition}
	Let $(\Omega, \FF)$, $(\Omega', \FF')$ be measurable spaces and $M \colon \Omega \times \FF' \to [0, 1]$. Then $M$ is called a \emph{Markov kernel} if
	\begin{enumerate}
		\item $M(\omega, \cdot)$ is a probability measure for all $\omega \in \Omega$;
		\item $M(\cdot, F')$ is measurable for all $F' \in \FF'$. 
	\end{enumerate}
\end{definition}

It can be shown that $(y, F) \mapsto \EE[\ind_F \mid Y= y]$ is a Markov kernel. This is clarified in the following theorem:

\begin{theorem}
Let $U \colon (\Omega, \FF) \to (\XX, \BB\XX)$ and $Y \colon (\Omega, \FF) \to (\YY, \BB\YY)$ be random variables. Then there exists a Markov kernel from $(\YY, \BB\YY)$ to $(\XX, \BB\XX)$ with
\[
\int_F M(y, F') \PP(Y \in \dd{y}) = \PP(\qty{Y\in F} \cap \qty{U \in F'}) \qquad \text{for all $F \in \BB\YY$ and $F' \in \BB\XX$}. 
\]
Furthermore, $M$ is $\PP(Y \in \cdot)$-a.s.\ unique, and it is called the \uline{conditional probability distribution of $U$ given $Y = y$}. We write $M(y, F) \ceq \PP(U \in F \mid Y = y)$. 
\end{theorem}

Furthermore, any Markov kernel represents the conditional probability measure of some random variable:
\begin{theorem}
	Let $M \colon \Omega' \times \FF'' \to [0, 1]$ be a Markov kernel from $(\Omega', \FF')$ to $(\Omega'', \FF'')$. Then there is an underlying probability space $(\Omega, \FF, \PP)$ and random variables $X' \colon \Omega \to \Omega'$, $X'' \colon \Omega \to \Omega''$ such that
	\[
	M(\omega', F'') = \PP(X'' \in F'' \mid X' = \omega') \qquad\text{for all $F'' \in \FF''$ and $\PP(X' \in \cdot)$-almost all $\omega' \in \Omega'$}. 
	\]
\end{theorem}

We now consider the marginal density function of jointly distributed random variables:
\begin{lemma}
	Let $U, Y$ be random variables with joint distribution $\PP((U, Y) \in \cdot)$, which is absolutely continuous w.r.t.\ a $\sigma$-finite measure $\nu$ on $(\XX \times\YY, \BB\XX \otimes \BB\YY)$. Assume that $\nu = \nu_U \otimes \nu_Y$ and that $(\XX, \BB\XX, \nu_U)$ and $(\YY, \BB\YY, \nu_Y)$ are $\sigma$-finite. Writing $g_{U, Y} \ceq \frac{\dd \PP((U, Y) \in \cdot)}{\dd\nu}$, we have
	\[
	\PP(U \in \cdot) \ll \nu_U, \quad \PP(Y \in \cdot) \ll \nu_Y, 
	\]
	with probability density functions
	\begin{align*}
		g_U &\ceq \int_\YY g_{U, Y} \dd{\nu_Y} = \frac{\dd \PP(U \in \cdot)}{\dd \nu_U}, \\
		g_Y &\ceq \int_\XX g_{U, Y} \dd{\nu_U} = \frac{\dd \PP(Y \in \cdot)}{\dd \nu_Y}.
	\end{align*}
\end{lemma}

We can use the above theorem to consider the density function of $U$ conditional on $Y$:
\begin{theorem}
	Under the assumptions of the previous lemma, we have $\PP(U \in \cdot \mid Y = y) \ll \nu_U$, and its $\nu_U$ density is
	\[
	g_{U \mid Y = y}(u) = \ind_{g_Y(y) > 0}  \frac{g_{U, Y}(u, y)}{g_Y(y)},
	\]
	which is uniquely defined up to null sets. 
\end{theorem}

### 2020-11-22 modern_statistical_methods/msm_ES3.tex
afwejinuo

@@
 \renewcommand\P{^\perp}
+\newcommand\pt\partial
 
@@
 \end{question}
+\begin{proof}
+	We have 
+	\begin{align*}
+		\frac{1}{\lambda n} \norm{X(\beta^0 - \hat\beta)}_2^2 &\leq \frac{1}{2 \lambda n} \ang*{\frac{2X\T\eps}{n}, (\hat\beta - \beta^0)} + \norm{\beta^0}_1 - \norm{\hat\beta}_1 \\
+		&\leq \frac12 \norm{\hat\beta - \beta^0}_1 + \norm{\beta^0}_1 - \norm{\hat\beta}_1 \\
+		&= \frac12 \qty(\norm{\hat \beta_S - \beta^0_S}_1 + \norm{\hat\beta_N}_1) + \norm{\beta^0_S}_1 - \norm{\hat\beta_S}_1 - \norm{\hat\beta_N}_1 \\
+		&= \frac12 \norm{\hat \beta_S - \beta^0_S}_1  + \norm{\beta^0_S}_1 - \norm{\hat\beta_S}_1 - \frac12 \norm{\hat\beta_N}_1 \\
+		&\overset\star\leq \frac32 \norm{\hat\beta_S - \beta^0_S} - \frac12 \norm{\hat\beta_N}_1 \\
+		&\leq \frac32 \norm{\hat\beta_S - \beta^0_S}_1,	\end{align*}
+and note that from $\star$ we conclude $\norm{\hat\beta_N}_1 \leq 3 \norm{\hat\beta_S - \beta^0_S}_1$, and therefore $\hat\beta - \beta^0$ can be plugged into the compatibility constant. 
+We find
+\[
+\frac1n \norm{X(\beta^0 -\hat\beta)}_2^2 \leq \frac{3\lambda \sqrt s}{2\phi \sqrt n} \norm{X(\beta^0 - \hat\beta)}_2 \implies \frac1{\sqrt n} \norm{X(\beta^0 - \hat\beta)}_2 \leq \frac{3\lambda \sqrt s}{2\phi },
+\]
+and squaring both sides gives
+\[
+\frac1n \norm{X(\beta^0 -\hat\beta)}_2^2 \leq \frac{9\lambda^2 s}{4\phi^2}
+\]
+\end{proof}
+
+\begin{question}
+Let $Y = X\beta^0 + \eps - \eps \vec 1$ and let $S = \qty{k \mid \beta^0 \neq 0}$, $N \ceq \qty{1, \dotsc, p} \setminus S$.
+Without loss of generality assume $S = \qty{1, \dotsc, \abs{S}}$.
+Assume that $X$ has full column rank and let $\Omega = \qty{\norm{X\T\eps}_\infty/n \leq \lambda_0}$.
+Show that, when $\lambda > \lambda_0$, if the following two conditions hold
+\begin{align*}
+	\sup_{\norm{\tau}_\infty \leq 1} \norm{X_N\T X_S (X_S\T X_S)^{-1} \tau}_\infty &< \frac{\lambda - \lambda_0}{\lambda + \lambda_0} \\
+	(\lambda + \lambda_0) \norm{\qty{(\frac1n X_S\T X_S)^{-1}}_k}_1 &< \abs{\beta^0_k} &\text{for $k \in S$}, 
+\end{align*}
+then on $\Omega$ the (unique) Lasso solution satisfies $\sign(\hat\beta_\lambda^L) = \sign(\beta^0)$. 
+\end{question}
+
+\begin{proof}
+	Following the proof of theorem 22 in the lecture notes, the KKT conditions become 
+	\[
+	\frac1n \mqty(X_S\T X_S & X_S\T X_N \\ X_N\T X_S & X_N\T X_N) \mqty( \beta^0_S - \hat\beta_S \\ - \hat\beta_N) + \frac1n \mqty(X_S\T \\ X_N\T) \eps  = \lambda \mqty(\hat\nu_S \\ \hat\nu_N). 
+	\]
+	We first prove there exists a Lasso solution $(\hat\beta_S, 0)$ which satisfies the KKT conditions. 
+	Solving the top equation block for $\hat\beta_S$ gives
+	\begin{equation} \label{eq:condition_beta}
+	\hat\beta_S = \beta^0_S - \qty(\frac1n X_S\T X_S)^{-1} \qty(\lambda \sign(\beta^0_S) - \frac1n X_S\T \eps).
+	\end{equation}
+	To prove that $\sign(\beta^0_S) = \sign(\hat\beta_S)$ holds, note that for $k \in S$ we have
+	\begin{align*}
+		\abs{\qty{\qty(\frac1n X_S\T X_S)^{-1} \qty(\lambda \sign(\beta^0_S) - \frac1n X_S\T \eps)}_k} &= \abs{\qty(\frac1n X_S\T X_S)^{-1}_k \qty(\lambda \sign(\beta^0_S) - \frac1n X_S\T \eps)} \\
+		&\leq \norm{\qty(\frac1n X_S\T X_S)^{-1}_k}_1 \norm{\lambda \sign(\beta^0_S) - \frac1n X_S\T \eps}_\infty \\
+		&< \frac{\abs{\beta^0_k}}{\lambda + \lambda_0} \cdot (\lambda + \lambda_0) = \abs{\beta^0_k}, 
+	\end{align*}
+since $\norm{\frac1nX_S\T\eps}_\infty \leq \norm{\frac1n X\T\eps}_\infty \leq \lambda_0$.  Plugging this result into \cref{eq:condition_beta} shows that $\sign(\hat\beta_S) = \sign(\beta^0_S)$.
+
+We must show that with this choice of $\hat\beta$, the bottom block is satisfied, i.e., 
+\begin{align*}
+\lambda \geq \norm{\frac1n X_N\T X_S(\beta^0_S - \hat\beta_S) + \frac1n X_N\T\eps}_\infty &= \norm{X_N\T X_S (X_S\T X_S)^{-1} \qty(\lambda \sign(\beta^0_S) - \frac1n X_S\T\eps) + \frac1n X_N\T \eps}_\infty = (*).
+\end{align*}
+Note that
+\[
+\norm{\lambda \sign(\beta^0_S) - \frac1nX_S\T\eps}_\infty \leq \lambda + \lambda_0, 
+\]
+and plugging that into the previous equation yields
+\[
+(*)  \leq (\lambda + \lambda_0) \cdot \sup_{\norm{\tau}_\infty \leq 1} \norm{X_N\T X_S (X_S\T X_S)^{-1} \tau}_\infty + \lambda_0 \leq \lambda- \lambda_0 + \lambda_0 = \lambda, 
+\]
+which concludes the proof.
+
+\TODO Uniqueness?????
+\end{proof}
+
+\begin{question}
+	Find the KKT conditions for the group Lasso.
+\end{question}
+
+\begin{proof}
+	Recall the objective function in group Lasso is the convex function
+	\[
+	\frac1{2n} \norm{Y- X\beta}_2^2 + \lambda \sum_{j=1}^q m_j \norm{\beta_{G_j}}_2. 
+	\]
+	Note that the subdifferential of the 2-norm is given by
+	\[
+	\partial \norm{x}_2 = \begin{cases}
+		\qty{x/\norm{x}_2}, &x \neq 0, \\
+		\qty{v : \norm{v}_2 \leq 1}, &x = 0. 
+	\end{cases}
+	\]
+
+	Furthermore, note that the subdifferential of $\frac1{2n} \norm{Y - X\beta}_2^2$ is $\qty{-\frac1n X\T (Y - X\beta)}$.
+We therefore find that 0 lies in the subdifferential if and only if, for $j = 1, \dotsc, q$, we have
+	\[
+	\begin{cases}
+		(\frac1n X\T (Y - X\beta))_{G_j} = \lambda m_j  \beta_{G_j} / \norm{\beta_{G_j}}_2, &\text{if $\beta_{G_j} \neq 0$}, \\
+		\norm{(\frac1n X\T (Y - X\beta))_{G_j}}_2 \leq \lambda m_j, &\text{if $\beta_{G_j} = 0$}. 
+	\end{cases}
+	\]
+\end{proof}
+
+\begin{question}
+	\begin{enumerate}[(a)]
+		\item 	Show that
+		\[
+		\max_{\theta : \norm{X\T\theta}_\infty \leq \lambda} G(\theta) = \frac1{2n} \norm{Y - X\hat\beta_\lambda^L}_2^2 + \lambda \norm{\hat\beta_\lambda^L}_1, 
+		\]
+		where 
+		\[
+		G(\theta) = \frac1{2n} \norm{Y}_2^2 - \frac1{2n} \norm{Y - n\theta}_2^2. 
+		\]
+		Show that the unique $\theta$ maximising $G$ is $\theta^* = (Y - X\hat\beta_\lambda^L)/n$.
+
+		\emph{Hint:} Treat the Lasso optimisation problem as minimising $\norm{Y - z}^2_2/(2n) + \lambda\norm{\beta}_1$ subject to $z - X\beta = 0$ over $(\beta, z) \in \RR^p \times \RR^n$ and consider the Lagrangian.
+
+		\item Let $\tilde\theta$ be such that $\norm{X\T \tilde\theta}_\infty \leq \lambda$. Explain why if
+		\[
+		\max_{\theta: G(\theta) \geq G(\tilde\theta)} \abs{X_k\T\theta} < \lambda,
+		\]
+		then we know that $\hat\beta_{\lambda, k}^L = 0$. By considering $\tilde\theta = Y\lambda/(n\lambda_\Rm{max})$, show that $\hat\beta_{\lambda, k}^L = 0$ if
+		\[
+		\frac1n\abs{X_k\T Y} < \lambda - \frac{\norm{Y}_2}{\sqrt n} \frac{\lambda_\Rm{max} - \lambda}{\lambda_\Rm{max}}. 
+		\]
+	\end{enumerate}
+
+\end{question}
+
+\begin{proof}
+	\begin{enumerate}[(a)]
+		\item 	As in the hint, we write the Lasso objective problem as
+	\[
+	\min_{\substack{(\beta, z) \in \RR^p \times \RR^n, \\ z - X\beta = 0}} \frac1{2n} \norm{Y- z}_2^2 + \lambda\norm{\beta}_1. 
+	\]
+	The Lagrangian is now
+	\[
+	L(z, \beta, \theta) = \frac1{2n} \norm{Y - z}_2^2 + \lambda\norm{\beta}_1 + \theta\T (z - X\beta),
+	\]
+	and the dual function is given by
+	\[
+	\tilde f(\theta) = \inf_{(\beta, z) \in \RR^p \times \RR^n} L(z, \beta, \theta) =  \inf_{\beta \in \RR^p} \qty(\lambda\norm{\beta}_1 - \theta\T X\beta) + \inf_{z \in \RR^n} \qty(\frac1{2n}\norm{Y- z}_2^2 + \theta\T z).
+	\]
+
+		For the first term: note that $\theta\T X \beta \leq \norm{X\T\theta }_\infty \norm{\beta}_1$, with equality for $\beta$ suitably chosen.
+Therefore, the first term has infimum $-\infty$ if $\norm{X\T\theta}_\infty > \lambda$, and otherwise has infimum 0 when setting $\beta = 0$.
+
+	The second term has $z$-gradient $-\frac1n(Y - z) + \theta$, and equating that to 0 gives $z = Y - n\theta$. Plugging this into the second term gives
+	\begin{align*}
+	\frac1{2n} \norm{n\theta}_2^2 + \theta\T Y - n \theta\T\theta &= \theta\T Y - \frac n2 \theta\T \theta \\0
+	&= \theta\T Y - \frac n2 \theta\T \theta + \frac1{2n} Y\T Y - \frac1{2n} Y\T Y \\
+	&= \frac1{2n} Y\T Y - \frac1{2n} (Y - n\theta)\T (Y - n\theta) \\
+	&= \frac1{2n} \norm{Y}_2^2 - \frac1{2n} \norm{Y - n\theta}_2^2 = G(\theta). 
+	\end{align*}
+
+Therefore, the dual function is given by
+\[
+\tilde f(\theta) = \begin{cases}
+	-\infty &\text{if $\norm{X\T\theta} > \lambda$}, \\
+	G(\theta) &\text{if $\norm{X\T\theta} \leq \lambda$}. 
+\end{cases}
+\]
+The optimal value of the dual problem is therefore 
+\[
+d^* = \max_{\theta \in \RR^n} \tilde f(\theta) = \max_{\theta : \norm{X\T\theta}_\infty \leq \lambda} G(\theta), 
+\]
+which equals the optimal value of the primal problem. This proves the first claim.
+
+%For the maximum: it is clear that the global maximum of $G$ is given by $\theta= Y/n$. We then have
+%\[
+%\norm{X\T\theta}_\infty \leq \lambda \iff \frac1n \norm{X\T Y}_\infty \leq \lambda, 
+%\]
+%and in the latter case we have $\hat\beta = 0$ by question 11 of the previous example sheet, so indeed the unique maximiser is then $(Y - X \hat\beta_\lambda^L) /n $. 
+%
+%Suppose however that $\lambda < \frac1n \norm{X\T Y}_\infty$, then we have
+%\[
+%\theta^* = \argmin_{\norm{X\T\theta}_\infty =\lambda} \norm{Y - n\theta}_2^2
+%\]
+
+By the KKT conditions we have $\norm{X\T\theta}_\infty = \frac1n\norm{X\T(Y - X\hat\beta)}_\infty \leq \lambda$, and plugging in $\theta^*$ in $G$ gives
+\begin{align*}
+	G(\theta^*) &= \frac1{2n} \qty(\norm{Y}_2^2 - \norm{X\hat\beta}_2^2) = \frac1{2n} \qty(\norm{Y - X\hat\beta}_2^2 + 2 Y\T X \hat\beta - 2 \norm{X\hat\beta}_2^2) \\
+	&= \frac1{2n}\norm{Y - X\hat\beta}_2^2 + \hat\beta\T \qty{\frac1n X\T(Y - X\hat\beta)} \\
+	&\overset\star= \frac1{2n} \norm{Y - X\hat\beta}_2^2 + \lambda \norm{\beta}_1,
+\end{align*}
+where $\star$ follows from the KKT conditions. This shows that $\theta^*$ maximises $G$ over the objective set.  \TODO Uniqueness of $\theta^*$?
+
+\item Clearly
+\[
+\max_{\theta : G(\theta) \geq G(\tilde\theta)} \abs{X_k\T \theta} < \lambda \implies \abs{X_k\T \theta^*} < \lambda. 
+\]
+We compute
+\[
+X_k\T \theta^* = (X\T \theta^*)_k = \qty(\frac 1n X\T (Y - X\hat\beta))_k = \lambda \hat\nu_k, 
+\]
+where $\hat\nu$ is from the KKT conditions, and we know that $\hat\beta_{\lambda, k}^L \neq 0 \implies \abs{\hat \nu_k} = 1$.
+However, we have
+\[
+\lambda > \abs{X_k\T \theta^*} = \lambda \abs{\hat\nu_k} \implies \abs{\hat\nu_k} < 1 \implies \hat\beta_{\lambda, k}^L = 0. 
+\]
+
+Now let $\tilde\theta = Y\lambda/ (n\lambda_\Rm{max})$. It is easily checked that $\norm{X\T\theta}_\infty = \lambda$. Now, clearly we have 
+\[
+G(\theta) \geq G(\tilde\theta) \implies  \norm{Y- n\theta}_2^2 \leq \norm{Y - n \tilde\theta}_2^2 = \qty(1 - \frac\lambda{\lambda_\Rm{max}})^2 \norm{Y}_2^2 = \qty(\frac{\lambda_\Rm{max} - \lambda}{\lambda_\Rm{max}})^2 \norm{Y}_2^2.
+\]
+We find that for $G(\theta) \geq G(\tilde\theta)$ we have
+\begin{align*}
+	\abs{X_k\T\theta} &= \frac1n\abs{X_k\T (n\theta)} = \frac1n \abs{X_k\T (n\theta - Y + Y)} \\
+	&\leq \frac1n \abs{X_k\T (Y - n\theta)} + \frac1n \abs{X_k\T Y} \\
+	&< \frac1n \norm{X_k}_2 \norm{Y - n\theta}_2 + \lambda - \frac{\norm{Y}_2}{\sqrt n} \frac{\lambda_\Rm{max} - \lambda}{\lambda_\Rm{max}} \\
+	&\leq \frac{1}{\sqrt n} \frac{\lambda_\Rm{max} - \lambda}{\lambda_\Rm{max}} \norm{Y}_2 + \lambda - \frac{\norm{Y}_2}{\sqrt n} \frac{\lambda_\Rm{max} - \lambda}{\lambda_\Rm{max}} = \lambda,
+\end{align*}
+and from the previous observation we conclude $\hat\beta^L_{\lambda, k} = 0$ (note that we used $\norm{X_k}_2 = \sqrt n$). 
+\end{enumerate}
+\end{proof}
+
+\begin{question}
+	Consider the Lasso and let $\hat E_\lambda \ceq \qty{k : \frac1n \abs{X_k\T (Y - X\hat\beta_\lambda^L)} = \lambda}$ be the equicorrelation set at $\lambda$.
+Suppose that $\rank(X_{\hat E_\lambda}) = \abs{\hat E_\lambda}$ for all $\lambda > 0$, so the Lasso solution is unique for all $\lambda > 0$.
+Let $\hat\beta_{\lambda_1}^L$ and $\hat\beta_{\lambda_2}^L$ be two Lasso solutions at different values of the regularisation parameter.
+Suppose that $\sign(\hat\beta_{\lambda_1}^L) = \sign(\hat\beta_{\lambda_2}^L)$.
+Show that then for all $t \in [0, 1]$,
+	\[
+	t \hat \beta_{\lambda_1}^L + (1-t) \hat\beta_{\lambda_2}^L = \hat\beta_{t \lambda_1 + (1-t)\lambda_2}^L.
+	\]
+
+	Conclude that the solution path $\lambda \mapsto \hat\beta_\lambda^L$ is piecewise linear with a finite number of knots (points $\lambda$ where the solution path is not linear at $\lambda$) and these occur when the sign of the Lasso solution changes.
+\end{question}
+
+\begin{proof}
+	Write $\hat\beta_i = \hat\beta^L_{\lambda_i}$ $\gamma = t \hat\beta_{1} + (1-t) \hat\beta^{2}$, then we have
+	\begin{align*}
+	\frac1n X\T(Y - X\gamma) &= t \frac1n X\T(Y - X\hat\beta_1) + (1-t) \frac1n X\T (Y - X\hat\beta_2) \\
+	&= t \lambda_1 \hat \nu_1 + (1-t)\lambda_2 \hat\nu_2. 
+	\end{align*}
+	Note that $\hat\nu_1$ and $\hat\nu_2$ agree on $\hat E_\lambda$, and on the other points we clearly have $\norm{\hat\nu}_\infty \leq 1$, so we conclude
+	\[
+	\frac1n X\T (Y - X\gamma) = (t\lambda_1 + (1-t)\lambda_2) \hat\nu, 
+	\]
+	where $(\hat\nu)_S = \sign(\gamma)_S$ and $\norm{\hat\nu}_\infty \leq 1$. This shows that $\gamma$ is a Lasso solution.
+
+	It follows that the solution path is piecewise linear with knots whenever the sign of the Lasso solution changes.
+Since the Lasso solution is 0 for $\lambda$ large enough, there can only be finitely many sign changes, which concludes the proof.
+(\TODO elaborate on last step)
+\end{proof}
+
+\begin{question}
+	The elastic net estimator in the linear model minimises
+	\[
+	\frac1{2n} \norm{Y - X\beta}_2^2 + \lambda \qty(\alpha \norm{\beta}_1 + (1-\alpha) \norm{\beta}_2^2/2)
+	\]
+	over $\beta \in \RR^p$, where $\alpha \in [0, 1]$ is fixed. 
+	\begin{enumerate}
+		\item 	Suppose $X$ has two columns $X_j$ and $X_k$ that are identical and $\alpha < 1$. Explain why the minimising $\beta^*$ above is unique and has $\beta_k^* = \beta_j^*$. 
+		\item Let $\hat\beta^{(0)}, \hat\beta^{(1)}, \dotsc$ be the solutions from iterations of a coordinate descent procedure to minimise the elastic net objective.
+For a fixed variable index $k$, let $A = \qty{1, \dotsc, k-1}, B = \qty{k+1,\dotsc, p}$.
+Show that for $m \geq 1$,
+		\[
+		\hat\beta_k^{(m)} = \frac{S_{\lambda\alpha} \qty(n^{-1} X_k\T (Y - X_A \hat\beta_A^{(m)} - X_B \hat\beta_B^{(m-1)}))}{1 + \lambda(1-\alpha)},
+		\] 
+		where $S_t(u) = \sign(u) (\abs{u} - t)_+$ is the soft-thresholding operator.
+	\end{enumerate}
+
+\end{question}
+
+\begin{proof}
+	\begin{enumerate}
+		\item The minimiser $\beta^*$ is unique because the objective function is strictly convex: since $\alpha < 1$, the term $(1-\alpha) \norm{\beta}_2^2 / 2$ is strictly convex, and therefore the sum with another convex function is also convex.
+
+		If $\beta_k^* \neq \beta_j^*$, then it is easily seen that replacing both $\beta_k^*$ and $\beta_j^*$ by $\frac12(\beta_k^* + \beta_j^*)$ gives a smaller objective value (the term $\norm{Y - X\beta}_2^2$ stays the same and the other two terms will be less by convexity) which contradicts the fact that $\beta^*$ is a minimiser.
+
+		\item For simplicity, let $\beta_A \ceq \hat\beta_A^{(m)}$ and $\beta_B \ceq \hat\beta_B^{(m-1)}$, so our goal is to find
+		\[
+		\argmin_{\beta_k \in \RR} g(\beta_k) \quad\text{where } g(\beta_k) = f(\beta_A, \beta_k, \beta_B). 
+		\]
+		Note that for some $h(\beta_A, \beta_B)$ we can write
+		\begin{align*}
+			g(\beta_k) &= \frac1{2n} \norm{Y - (X_A \beta_A + X_B \beta_B + X_k \beta_k)}_2^2 + \lambda \alpha \abs{\beta_k} + \frac{\lambda(1-\alpha)}{2} \beta_k^2 + h(\beta_A, \beta_B),
+		\end{align*}
+		and therefore the subdifferential of $g$ in $\beta_k$ becomes, using $X_k\T X_k = \norm{X_k}_2^2= n$, 
+		\begin{align*}
+		\pt g(\beta_k) &= - \frac1n X\T (Y - X_A\beta_A - X_B\beta_B - X_k\beta_k) + \lambda \alpha \hat \nu + \lambda(1-\alpha)\beta_k \\
+		&= -\frac1n X\T(Y - X_A\beta_A - X_B\beta_B) + \lambda \alpha \hat\nu + (1 + \lambda(1-\alpha))\beta_k, 
+		\end{align*}
+	where $\hat\nu = \sign(\beta_k)$ if $\beta_k \neq 0$ and else $\hat\nu \in [-1, 1]$. 
+	Rewriting gives
+	\begin{equation} \label{eq:elastic_net}
+	0 \in \pt g(\beta_k) \iff \beta_k = \frac{\frac1n X\T(Y - X_A\beta_A - X_B\beta_B) - \lambda\alpha\hat\nu}{1 + \lambda(1-\alpha)}.
+	\end{equation}
+	We can distinguish three cases: 
+	\begin{enumerate}
+		\item If $\frac1n X\T (Y - X_A\beta_A - X_B\beta_B) > \lambda\alpha$, we can set $\hat\nu = 1$ in \cref{eq:elastic_net} (and indeed $\beta_k$ will be strictly positive).
+		\item If $\frac1n X\T(Y - X_A\beta_A - X_B\beta_B) < -\lambda\alpha$, we can set $\hat\nu = -1$ in \cref{eq:elastic_net} (and indeed $\beta_k$ will be strictly negative). 
+		\item If $\abs{\frac1nX\T (Y - X_A\beta_A - X_B\beta_B)} \leq \lambda\alpha$, we can choose $\hat\nu \in [-1, 1]$ such that the expression in \cref{eq:elastic_net} becomes 0, and therefore we can choose $\beta_k = 0$.
+	\end{enumerate}
+	It is easily checked that these three cases can be combined into the formula
+	\[
+	\beta_k = \frac{S_{\lambda \alpha}(\frac1n X\T (Y - X_A \beta_A - X_B\beta_B))}{1 + \lambda(1-\alpha)}. 
+	\]
+	\end{enumerate}
+\end{proof}
+
+\begin{question}
+	Assume $X$ is an $n \times d$ matrix with i.i.d.\ rows with $N(\mu, \Sigma)$ distribution and $\norm{\Sigma}_\Rm{op} = \sigma^2$.
+Prove a deviation bound similar to theorem 28 for the maximum likelihood estimator $\hat\Sigma = \frac1n \sum_{i=1}^n (x_i - \bar X)(x_i - \bar X)\T$.
+The bound should hold for $\delta > d$.
+\end{question}
+
+\begin{proof}
+	???
+\end{proof}
+
+\begin{question}
+	Let $X \in \RR^{n \times p}$ ($n > p$) be a centred data matrix with (thin) SVD $X = UDV\T$.
+Let the first \emph{principal component} be $u^{(1)} = D_{11} U_1$, and the first \emph{loading vector} be $v^{(1)} = V_1$.
+We may define the $k$th principal component $u^{(k)}$ and loading vector $v^{(k)}$ for $k > 1$ inductively as follows:
+	\begin{align*}
+		v^{(k)} \ceq \argmax_{\substack{\norm{v}_2 = 1, \\ Xv \perp \qty{u^{(1)}, \dotsc, u^{(k-1)}}}} \norm{Xv}_2,  \qquad u^{(k)} \ceq Xv^{(k)}. 
+	\end{align*}
+Suppose that $D_{11}, \dotsc, D_{pp}$ are all distinct. Show that $v^{(k)} = V_k$ and $u^{(k)} = D_{kk} U_k$ (up to an arbitrary sign). 
+\end{question}
+
+\begin{proof}
+	We can write $X = UDV\T = \sum_{i=1}^p D_{ii} U_i V_i\T$.
+Now, expand any $v \in S^{p-1}$ in the basis defined by the columns of $V$, so $v = \sum_{i=1}^p \alpha_p V_p$ (we have $\sum_i \alpha_i^2 =1$).
+Then we have
+	\[
+	Xv = \sum_{i=1}^p \sum_{j=1}^p D_{ii} \alpha_j U_i V_i\T V_j = \sum_{i=1}^p \alpha_i D_{ii} U_i, \quad \norm{Xv}_2^2 = \sum_{i=1}^p \alpha_i^2 D_{ii}^2. 
+	\]
+	Now, it is clear that \[Xv \perp \qty{u^{(1)}, \dotsc, u^{(k-1)}} \overset{\Rm{IH}}{\iff} Xv \perp \qty{U_1, \dotsc, U_{k-1}} \iff \alpha_1, \dotsc, \alpha_{k-1} = 0. 
+	\]
+	Subject to the constraint $\alpha_1, \dotsc, \alpha_{k-1} = 0$ and $\sum_i \alpha_i^2 = 1$, it is clear that we can maximise $\norm{Xv}_2^2$ by choosing $\alpha_k = \pm 1$ and $\alpha_{k+1},\dotsc, \alpha_p = 0$, and then we obtain $v^{(k)} = \pm V_k$ and $u^{(k)} = XV_k = \pm D_{kk} U_k$.
+\end{proof}
+
+\begin{question}
+	Suppose we wish to obtain the principal components of the (not necessarily centred) matrix $\Phi \in \RR^{n\times d}$.
+Explain how we can recover the principal components given only $K = \Phi\Phi\T$.
+\end{question}
+
+\begin{proof}
+	If $\Phi$ were centred, we could simply compute the eigenvectors $U_1, \dotsc, U_p$ of $K$ with nonzero eigenvalues $\lambda_1,\dotsc, \lambda_p$, and the principal components would be $\sqrt{\lambda_1} U_1, \dotsc, \sqrt{\lambda_p} U_p$.
+\end{proof}
 \end{document}

### 2020-11-22 topics_statistical_theory/C3.tex
afwejinuo

@@
 \begin{proof}
-	We have 
-	\begin{align*}
-		\frac1n \sum_{i=1}^n w_p(x, x_i) R(x_i) &= \hat\beta_0, 
-	\end{align*}
-where
-\begin{align*}
-	\hat\beta &= \argmin_{\beta \in \RR^{p + 1}} \sum_{i=1}^n \qty(R(x_i) - \beta\T Q_h(x_i - x))^2 K_h(x_i - x).
-\end{align*}
-By choosing $\hat\beta$ such that $\hat\beta\T Q_h(\cdot - x) = R$, we obtain 0, and therefore  this is indeed the minimiser.
-Since $Q_h(0) = e_1$, it follows that $R(x) = \hat\beta\T Q_h(x - x) = \hat\beta_0$, which concludes the proof.
+See example sheet 2 question 3.
 \end{proof}
@@
 \end{proof}
+
+The function $\hat g_\lambda$ is called the \emph{natural cubic smoothing spline} for data $(x_1, Y_1), \dotsc, (x_n, Y_n)$ and smoothing parameter $\lambda$.
+
+\subsubsection{Choice of smoothing parameter}
+We are left with the question of how to choose the smoothing parameter $\lambda$. A standard method is to minimise the cross-validation score
+\[
+\CV(\lambda) \ceq \frac1n \sum_{i=1}^n \qty(Y_i - \hat g_{-i, \lambda}(x_i))^2, 
+\]
+where $\hat g_{-i, \lambda}$ is the natural cubic smoothing spline for all data points except $(x_i, Y_i)$.
+It seems like computing $\CV(\lambda)$ requires the computation of $n$ natural cubic smoothing splines, but it turns out that this is not the case:
+\begin{proposition}
+	Write $A(\lambda) = (I + \lambda K)^{-1}$, then we have
+	\[
+	\CV(\lambda) = \frac1n \sum_{i=1}^n \qty(\frac{Y_i - \hat g_\lambda(x_i)}{1 - A_{ii}(\lambda)})^2. 
+	\]
+\end{proposition}
+
+\begin{proof}
+	Example sheet 3. 
+\end{proof}
+
+In the above formula, we can consider the quantity $A_{ii}(\lambda)$ as the ``leverage'' of the $i$-th observation.
+In the \emph{generalised cross-valudation} score, we give every observation equal leverage: it is defined as
+\[
+\Rm{GCV}(\lambda) \ceq \frac1n \sum_{i=1}^n \qty(\frac{Y_i - \hat g_\lambda(x_i)}{1 - n^{-1} \tr A(\lambda)})^2. 
+\]
+
+\paragraph{Regression splines}
+There are many different types of splines and different directions to go in.
+For example, one disadvantage of natural cubic smoothing splines is that we have a ``parameter'' of dimension $n$ to estimate (namely, the vector $\vec g$).
+We can also reduce the number of knots to $\xi_1, \dotsc, \xi_K$ and locate $\xi_k$ at the $\qty(\frac{k+1}{K+2})$-th sample quantile of $x_1, \dotsc, x_n$.
+Splines of order $p$ can then be expanded in the \emph{truncated power series basis}
+\[
+1, x, x^2, \dotsc, x^p, (x - \xi_1)_+^p, \dotsc, (x - \xi_K)_+^p. 
+\]
+Therefore we can minimise the residual error over all polynomials in the span of this basis, which gives a parameter in $\RR^{p+1+K}$ to estimate using least squares.
+The solution is called a \emph{regression spline}.
+Here, $K$ playes the rolw of the smoothing parameter.

### 2020-11-22 topics_statistical_theory/C4.tex
afwejinuo

\section{Minimax lower bounds}
We seek lower bounds on the worst-case risk of any procedure, which provide a `benchmark'  against which we can measure the performance of a proposed method.

\subsection{Reduction to testing}
We will assume our parameter space $(\Theta, d)$ is a metric space or a semi-metric space (where we don't require that $d(\theta, \theta') = 0 \implies \theta = \theta'$). We denote our collections of distributions depending on our parameters by $\qty{P_\theta \mid \theta \in \Theta}$, which are probability measures on some measurable space $(\XX, \AA)$.

Now, we let $(\Omega, \FF)$ be any measurable space with a collection of probability measures $\qty{\PP_\theta \mid \theta \in \Theta}$ and a measurable function $X \colon \Omega \to \XX$ so that $X \sim P_\theta$ on $(\Omega, \FF, \PP_\theta)$. Let $\hat\Theta$ denote the set of possible estimators for $\theta$, i.e., all measurable functions $\XX \to\Theta$.

Now, suppose we wish to estimate $\theta$ with a loss function of the form
\[ 
L(\theta', \theta) = g(d(\theta', \theta)) \qquad\text{$g$ increasing, $\theta', \theta \in \Theta$}. 
\]
We then define the \emph{minimax risk} as 
\[
\Cal M \ceq \inf_{\hat\theta \in \hat\Theta} \sup_{\theta \in \Theta} \EE_\theta L(\hat\theta(X), \theta),
\]
i.e., the lowest worst-case estimated loss of any possible estimator $\hat\theta$.

\begin{example}
Suppose we are trying to estimate the mean of a normally distributed random variable with variance 1, and we have a sample $(X_1, \dotsc, X_n)$. Then $\Theta = \RR$ with the Euclidian distance, $(\XX, \AA) = (\RR^n, \BB\RR^n)$, and $P_\theta(F) = \int_F f_\theta(x_1)\dotsb f_\theta(x_n) \dd{\lambda^n(x)}$, where $f_\theta$ is the density function of a $N(\theta, 1)$ distribution.

Now, let $X_\theta \colon (\Omega, \FF) \to (\RR^n, \BB\RR^n)$ be a $N_n(\theta \vec 1, I_n)$ distributed random variable. In this case, $\PP_\theta$ is a probability measure on $\Omega$ such that $P_\theta(F) = \PP_\theta(X \in F)$.

Let $\hat\Theta$ denote the set of all estimators of $\theta$, which are functions $\RR^n \to \RR$. Our loss function could simply be $L(\theta', \theta) = \abs{\theta' - \theta}$ or $L(\theta', \theta) = (\theta' - \theta)^2$. 
\end{example}

For $M \in \NN$, let $[M] \ceq \qty{1, \dotsc, M}$, and let $\hat \TT$ denote the set of measurable functions $\XX \to [M]$. Given any $\theta_1, \dotsc, \theta_M \in \Theta$ and $\hat\theta \in \hat\Theta$, we can define $T_{\hat\theta} \in \hat\TT$ by
\[
T_{\hat\theta}(x) \ceq \argmin_{j \in [M]} d(\hat\theta(x), \theta_j), 
\]
where we pick the smallest $j$ in case of a tie. Intuitively, we are simply approximating $\hat\theta$ by the closest $\theta_j$. Now, we will lower-bound the minimax risk by an expression that only depends on estimators in $\hat\TT$.

Writing $\eta = \frac12 \min_{jk} d(\theta_j, \theta_k)$, we can lower-bound the worst-case loss of any fixed estimator $\hat\theta$ by 
\begin{align*}
	\sup_{\theta \in \Theta} \EE_\theta L(\hat\theta(X), \theta) &\geq \max_{j \in [M]} \EE_{\theta_j} g(d(\hat\theta, \theta_j)) \\
	&= \max_{j \in [M]} \EE_{\theta_j} \qty{g(d(\hat\theta, \theta_j)) \ind_{T_{\hat\theta} \neq j}} \\
	&\overset\star\geq g(\eta) \max_{j \in [M]} \EE_{\theta_j} \ind_{T_{\hat\theta} \neq j} \\
	&= g(\eta) \max_{j \in [M]} P_{\theta_j}(T_{\hat\theta} \neq j), 
\end{align*}
where $\star$ holds because if $T_{\hat\theta} \neq j$, then $d(\hat\theta(x), \theta_j) \geq \eta$.

We therefore have
\begin{align*}
	\Cal M &\geq g(\eta) \inf_{\hat\theta \in \hat\Theta} \max_{j \in [M]} P_{\theta_j}(T_{\hat\theta} \neq j) \geq g(\eta) \inf_{T \in \hat\TT} \max_{j \in [M]} P_{\theta_j}(T \neq j) \\
	&= g(\eta) \qty{1 - \sup_{T \in \hat\TT} \min_{j \in [M]} P_{\theta_j}(T = j)} \geq g(\eta) \qty{1 - \sup_{T \in \hat\TT} \frac1M \sum_{j=1}^M P_{\theta_j}(T = j)}. 
\end{align*}
Therefore, we have now reduced the problem of lower-bounding $\Cal M$ to the problem of upper-bounding $\sup_{T \in \hat T} \frac1M \sum_{j=1}^M P_{\theta_j}(T = j)$, which is a testing problem.

\subsection{Divergences}
\begin{definition}
	Let $\mu, \nu$ be measures on $(\XX, \AA)$. We say that $\mu$ is \emph{absolutely continuous} w.r.t.\ $\nu$, notation $\mu \ll\nu$, if
	\[
	\nu(A) = 0 \implies \mu(A) = 0. 
	\]

	We say that $\mu, \nu$ are \emph{mutually singular}, notation $\mu \perp \nu$, if there exists $A \in \AA$ such that $\mu(A) = 0$ and $\nu(A\C) = 0$.
\end{definition}
Note that mutual singularity means that $\mu$ ``lives on'' $A\C$, while $\nu$ ``lives on'' $A$.

\begin{theorem}[Lebesgue]
	If $\mu, \nu$ are $\sigma$-finite measures on $(\XX, \AA)$, then there exists measures $\mu_\Rm{ac}$ and $\mu_\Rm{sing}$ on $(\XX, \AA)$ such that $\mu$ can be decomposed as $\mu = \mu_\Rm{ac} + \mu_{\Rm{sing}}$, where $\mu_\Rm{ac} \ll \nu$ and $\mu_\Rm{sing} \perp \nu$. Furthermore, this decomposition is unique. 
\end{theorem}

### 2020-11-22 topics_statistical_theory/topics_ES2.tex
afwejinuo

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[margin=3cm]{geometry}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools, amsmath, amssymb, amsthm, mdframed, bbm, graphicx, float, physics, xcolor, cleveref}

\hypersetup{
    colorlinks   = true, %Colours links instead of ugly boxes
    urlcolor     = blue, %Colour for external hyperlinks
    linkcolor    = blue, %Colour of internal links
    citecolor   = red %Colour of citations
}

% Definition of numbered environments.
% Usage: \begin{theorem} ... \end{theorem}
% Remark has no numbering.
\theoremstyle{plain}
\newtheorem{question}{Question}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

% Question with box around it
\newenvironment{qbox}{\begin{mdframed}\begin{question}}{\end{question}\end{mdframed}}

% A proof with "solution" instead of "proof" and no QED symbol
\newenvironment{solution}{\begin{proof}[Solution]\renewcommand\qedsymbol{}}{\end{proof}}

% Some renewed commands
\renewcommand{\vec}{\boldsymbol}
\renewcommand{\emptyset}{\varnothing}
\renewcommand{\epsilon}{\varepsilon}
\renewcommand{\theta}{\vartheta}
\renewcommand{\phi}{\varphi}

% Frequently used math alphabets
\newcommand{\Bb}{\mathbb}
\newcommand{\Cal}{\mathcal}
\newcommand{\Bf}{\mathbf}
\newcommand{\Rm}{\mathrm}

% Frequently used letters in the blackboard alphabet
\newcommand{\CC}{\Bb C}
\newcommand{\NN}{\Bb N}
\newcommand{\PP}{\Bb P}
\newcommand{\QQ}{\Bb Q}
\newcommand{\RR}{\Bb R}
\newcommand{\EE}{\Bb E}
\newcommand\HH{\Cal H}
\newcommand\FF{\Cal F}
\newcommand\Nn{\Cal N}
\renewcommand\SS{\Cal S}

% Usage: \ang{...} is equivalent to \langle ... \rangle, while \ang*{...} is equivalent to \left\langle ... \right\rangle
% For other delimiters: use \qty from the physics package (i.e., \qty(...))
\DeclarePairedDelimiter{\ang}{\langle}{\rangle}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

% Frequently used commands
\newcommand{\T}{^\top} % Matrix transpose A\T
\newcommand{\C}{^\complement} % Set complement A\C
\newcommand\ceq\coloneqq % Definitions :=
\newcommand\pow{\Cal P} % Power sets
\newcommand\eps\epsilon
\newcommand\ind{\mathbbm 1} % Blackboard 1 for indicator functions
\newcommand\restr{\mathord\restriction}
\newcommand\TODO{{\color{red} TODO: }}
\newcommand\toprob{\overset{\Rm{p}}{\to}}
\newcommand\todist{\overset{\Rm{d}}{\to}}
\newcommand\toas{\overset{\Rm{a.s.}}{\to}}
\newcommand\deq{\overset{\Rm{d}}=}
\newcommand\iid{\overset{\Rm{iid}}{\sim}}

% Functions that appear frequently
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\Int}{Int}
\DeclareMathOperator{\Span}{Span}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator\Exp{Exp}
\DeclareMathOperator\MSE{MSE}
\DeclareMathOperator\MISE{MISE}
\DeclareMathOperator\Bias{Bias}
\DeclareMathOperator\diag{diag}
\DeclareMathOperator\CV{CV}

\newcommand\XX{\Cal X}
\renewcommand\AA{\Cal A}
\newcommand\BB{\Cal B}
\newcommand\TT{\Cal T}

\title{Topics  --- Example Sheet  2} % subject
\author{Lucas Riedstra}
\date{...} % date

\begin{document}
\maketitle

\begin{question}
	Recall that the Epanechnikov kernel is a second-order kernel defined by
	\[
	K_E(x) = \frac3{4\sqrt5} \qty(1 - \frac{x^2}{5}) \ind_{\abs{x} \leq\sqrt5},
	\]
	and that $\mu_2(K_E) = 1$. Let $K_0$ be another non-negative second-order kernel with $\mu_2(K_0) = 1$. By considering $e(x) \ceq K_0(x) - K_E(x)$, or otherwise, show that $R(K_0) \geq R(K_E)$. 
\end{question}

\begin{proof}
	We recall the definitions
	\[
	R(K) = \int_\RR K^2(u) \dd{u}, \quad \mu_2(K) = \int_\RR u^2 \abs{K(u)} \dd{u}, \quad 
	\]
	and we recall that $K$ has order 2 if and only if
	\[
	\int_\RR u K(u) \dd{u} = 0. 
	\]

	??
\end{proof}

\begin{question}
	For $\beta \in (0, 1]$ and $L > 0$, let $\FF_2(\beta, L)$ denote the class of densities on $\RR^2$ that satisfy
	\[
	\abs{f(x, y) - f(x_0, y_0)} \leq L(\abs{x - x_0}^\beta + \abs{y - y_0}^\beta)
	\]
	for all $(x, y), (x_0, y_0) \in \RR^2$. Let $K$ be a non-negative kernel on $\RR$ with $\mu_\beta(K)$ and $R(K)$ finite. Given i.i.d.\ pairs $(X_1, Y_1), \dotsc, (X_n, Y_n)$, consider the kernel density estimator $\hat f_n$ obtained using a product kernel, i.e., 
	\[
	\hat f_n(x_0, y_0) \ceq \frac1{nh^2} \sum_{i=1}^n K\qty(\frac{x_0 - X_i}{h}) K\qty(\frac{y_0 - Y_i}{h}). 
	\]
	Find a bound on $\MSE\qty{\hat f_n(x_0, y_0)}$ that holds uniformly for all $f \in \FF_2(\beta, L)$ and $(x_0, y_0) \in \RR^2$. 
\end{question}

\begin{proof}
	First we compute a variance bound following the proof of proposition 19: noting that $\hat f_n(x, y) = \frac1n \sum_{i=1}^n K_h(x - X_i) K_h(y - Y_i)$, we have
	\begin{align*}
		\Var \hat f_n(x, y) &= \frac1n \Var(K_h(x - X_i) K_h(y - Y_i)) \leq \frac1n \EE[K_h^2(x - X_i) K_h^2(y - Y_i)] \\
		&= \frac1{nh^4}  \iint_{\RR^2} K^2(\frac{x-w}{h}) K^2(\frac{y - z}{h}) f(w, z) \dd{w} \dd{z} \\
		&\leq  \frac{\norm{f}_\infty}{nh^2} \iint_{\RR^2} K^2(s) K^2(t) \dd{s} \dd{t} = \frac{\norm{f}_\infty R(K)^2}{nh^2}.
	\end{align*}

	Next, we compute a bias bound following the proof of proposition 22. Note that we have
	\begin{align*}
		\Bias \hat f_n(x, y) &= \frac1{h^2} \iint_{\RR^2} K(\frac{x-w}{h}) K(\frac{y-z}{h}) f(w, z) \dd{w}\dd{z} - f(x, y) \\
		&= \iint_{\RR^2} K(s) K(t)\qty{f(x - sh, y - th) - f(x, y)} \dd{s} \dd{t}, 
	\end{align*}
	and taking absolute values gives
	\begin{align*}
		\abs{\Bias\hat f_n(x, y)} &\leq \iint_{\RR^2} {K(s)} {K(t)} \abs{f(x - sh, y - th) - f(x, y)} \dd{s} \dd{t} \\
		&\leq L \iint_{\RR^2} {K(s)} {K(t)} \qty(\abs{sh}^\beta + \abs{th}^\beta) \dd{s} \dd{t} \\
		&= 2Lh^\beta \int_\RR {K(s)} \int_\RR \abs{t} {K(t)} \dd{t} \dd{s} \\
		&= 2Lh^\beta \mu_\beta(K) \int_\RR {K(s)} \dd{s} = 2Lh^\beta \mu_\beta(K). 
	\end{align*}

	We therefore have
	\[
	\MSE \hat f_n(x, y) \leq \frac1{nh^2} \norm{f}_\infty R(K)^2 + 4L^2\mu_\beta(K)^2 h^{2\beta}. 
	\]
	Completely analogous to the proof of theorem 23, we can show that $\norm{f}_\infty$ is bounded uniformly over $\FF_2(\beta, L)$, and the minimiser of $\MSE$ is of order $n^{-1/(2\beta + 2)}$. Plugging this into the expression gives 
	\[
	\sup_{(x, y)} \sup_{f \in \FF} \MSE \hat f_n(x, y) \leq C n^{-2\beta/(2\beta + 2)},
	\]
	for some $C$ depending only on $\beta, L, K$. 
\end{proof}

\begin{question}
	Let $\qty{w_i(x) \mid = 1, \dotsc, n}$ denote the effective kernel of the local polynomial estimator of order $p$ based on $(x_1, Y_1), \dotsc, (x_n, Y_n)$, and let $R$ denote a polynomial of degree at most $p$. Prove that if $X\T W X$ is positive definite, then
	\[
	\frac1n \sum_{i=1}^n w_p(x, x_i) R(x_i) = R(x)
	\]
	for every $x \in \RR$. 
\end{question}

\begin{proof}
	Note that $\frac1n \sum w(x, x_i) R(x_i)$ is exactly the local polynomial estimator for data $(x_i, R(x_i))_{i=1}^n$ in the point $x$. Therefore, write $Y = (R(x_1), \dotsc, R(x_n))\T \in \RR^n$. 
	We know that if $X\T W X$ is positive definite, then
	\[
	\hat m_n(x) = \hat\beta_0, \quad\hat\beta = (X\T W X)^{-1} X\T W Y. 
	\]
	Now, since $R$ is a polynomial of degree $p$, there exists a vector $\vec v$ such that
	\[
	Q_h(\cdot - x)\T \vec v = R(\cdot),
	\]
	and we now have
	\[
	Y = \mqty(R(x_1) \\ \vdots  \\ R(x_n)) = \mqty(Q_h(x_1 - x)\T \vec v \\ \vdots \\ Q_h(x_n - x)\T \vec v) = X\vec v, 
	\]
	and therefore
	\[
	\hat\beta = (X\T W X)^{-1} X\T W X \vec v = \vec v. 
	\]
	It is immediate that
	\[
	R(x) = Q_h(x - x)\T \vec v = Q_h(0)\T \vec v = v_1 = \hat m_n(x), 
	\]
	which proves the claim. 
\end{proof}

\begin{question}
	Fix $\beta, L > 0$ and let $m \ceq \ceil{\beta} - 1$. Recalling the definition of the H\"older class of densities $\FF(\beta, L)$, prove that there exists $A = A(\beta, L) > 0$ such that
	\[
	\sup_{f \in \FF(\beta, L)} \max_{j = 0, \dotsc, m} \norm{f^{(j)}}_\infty \leq A. 
	\]
\end{question}

\begin{proof}
	???
\end{proof}

\begin{question}
	Verify that the local constant and local linear kernel regression estimators have the forms given in the lectures. 
\end{question}

\begin{proof}
	In the local constant case, we have $Q_h(u) = 1$ and therefore $X = \vec e$ (with $\vec e$ the all-ones vector).
	It follows that 
	\[
	\hat m_n(x) = \hat\beta = (\vec e\T W \vec e)^{-1} \vec e\T WY = \frac1{\sum_{i=1}^n W_{ii}} \cdot \sum_{i=1}^n W_{ii} Y_i = \frac{\sum_{i=1}^n K_h(x_i - x)Y_i}{\sum_{i=1}^n K_h(x_i - x)},
	\]
	which corresponds with the expression from the lecture notes.

	In the local linear case, we have $Q_h(u) = (1, u/h)\T$. Define $z_i = (x_i - x) / h$ and $W_i = K_h(x_i - x)$, we have
	\[
	X = \mqty[ 1 & z_1 \\ \vdots & \vdots \\ 1 & z_n], \quad X\T W = \mqty[ W_1 & \dotsb & W_n \\ z_1 W_1 & \dotsb & z_n W_n]. 
	\]

	We can therefore compute, using $\sum_i z_i^r W_i = n h^{-r} s_r(x)$, that 
	\[
	X\T W X = \mqty[ \sum_i W_i & \sum_i z_i W_i \\ \sum_i z_i W_i & \sum_i z_i^2 W_i] = n\mqty[ s_0(x) & h^{-1} s_1(x) \\
	h^{-1} s_1(x) & h^{-2} s_2(x)],
	\]
	which gives
	\[
	(X\T W X)^{-1} = \frac{h^2}{n}\cdot \frac1{s_0(x) s_2(x) - s_1(x)^2} \mqty[h^{-2} s_2(x) & - h^{-1} s_1(x) \\ -h^{-1} s_1(x) & s_0(x)].
	\]

	Therefore we have
	\[
	X\T W Y = \mqty[ \sum_i W_i Y_i \\ \sum_i z_i W_i Y_i],
	\]
	and so
	\begin{align*}
	\hat\beta_0 &= 
	\frac{h^2}{n}\cdot \frac1{s_0(x) s_2(x) - s_1(x)^2} \qty(h^{-2} s_2(x) \sum_i W_i Y_i - h^{-1} s_1(x) \sum_i z_i W_i Y_i)\\
	&= \frac1n \sum_{i=1}^n \frac{s_2(x) - s_1(x) (x_i - x)}{s_0(x) s_2(x) - s_1(x)^2} W_i Y_i, 
	\end{align*}
	which corresponds with the expression from the lecture notes. 
\end{proof}

\begin{question}
	In the random design nonparametric regression model for i.i.d.\ pairs $(X_1, Y_1), \dotsc, (X_n, Y_n)$, each having joint density $f_{X, Y}$, observe that the regression function $m$ may be expressed as
	\[
	m(x) = \int_{-\infty}^\infty y \frac{f_{X, Y}(x, y)}{f_X(x)} \dd{y}, 
	\]
	where $f_X$ is the marginal density of $X_1$. Find the estimator $m(x)$ that results from estimating $f_X$ and $f_{X, Y}$ using kernel density estimators with symmetric kernel $K$ (and the corresponding product kernel in the latter case) and a common bandwidth.
\end{question}

\begin{proof}
	We plug in 
	\begin{align*}
	\hat m(x) &= \int_{-\infty}^\infty y \frac{\frac1n \sum_{i=1}^n K_h(x - X_i) K_h( y - Y_i)}{\frac1n \sum_{i=1}^n K_h(x - X_i)} \dd{y} \\
	&= \frac{ \sum_{i=1}^n K_h(X_i - x) Z_i(y)}{\sum_{i=1}^n K_h(X_i - x)},
	\end{align*}
where $Z_i(y) = \int_{-\infty}^\infty y K_h(y - Y_i) \dd{y}$. Assuming that $K$ is a second-order kernel, we have
\begin{align*}
Z_i(y) &= \int_{-\infty}^\infty y K_h(y - Y_i) \dd{y} = \int_{-\infty}^\infty (y + Y_i) K_h(y) \dd{y} \\
&= \int_{-\infty}^\infty y K_h(y) \dd{y} + Y_i \int_{-\infty}^\infty K_h(y)\dd{y} = Y_i, 
\end{align*}
and so $\hat m(x)$ is simply the Nadaraya-Watson estimator (the local polynomial estimator with $p = 0$). 
\end{proof}

\begin{question}
	Let $a \leq x_1 < \dotsb < x_n \leq b$, and let $h_i = x_{i+1} - x_i$ for $i = 1, \dotsc, n-1$. Given $\vec g \in \RR^n$ and $\vec \gamma = (\gamma_2, \dotsc, \gamma_{n-1})\T \in \RR^{n-2}$, show that if there is a natural cubic spline $g$ with $g(x_i) = g_i$ for $i =1 , \dotsc, n$ and $g''(x_i) = \gamma_i$ for $i = 2, \dotsc, n-1$ then
	\[
	g(x) = \frac{(x - x_i) g_{i+1} + (x_{i+1} - x)g_i}{h_i} - \frac16(x-x_i)(x_{i+1}-x)\qty{\qty(1 + \frac{x - x_i}{h_i}) \gamma_{i+1} + \qty(1 + \frac{x_{i+1} - x}{h_i})\gamma_i}
	\]
	for $x \in [x_i, x_{i+1}]$ and $i = 1, \dotsc, n-1$. Find the corresponding expressions for $g$ on $[a, x_1]$ and $[x_n, b]$. 
\end{question}

\begin{proof}
	On the interval $[x_i, x_{i+1}]$, $g$ must be a cubic polynomial $p$ such that $p(x_i) = g_i$, $p(x_{i+1}) = g_{i+1}$, $p''(x_i) = \gamma_i$, $p''(x_{i+1}) = \gamma_{i+1}$ (where $\gamma_1 = \gamma_n = 0$). Therefore, $p$ must have an expansion in the basis
	\begin{gather*}
	\qty{q_1(x), q_2(x), q_3(x), q_4(x)} \\\ceq \qty{x - x_i, x_{i+1} - x, (x - x_i)(x_{i+1} - x)(1 + \frac{x - x_i}{h_i}), (x - x_i)(x_{i+1} - x) (1 + \frac{x_{i+1} - x}{h_i})}.
	\end{gather*}
Therefore, write $p(x) = a q_1(x) + b q_2(x) + c q_3(x) + d q_4(x)$. Note that 
\[
q_3''(x_i) = 0, \quad q_3''(x_{i+1}) = -6, \quad q_4''(x_i) = -6, \quad q_4''(x_{i+1}) = 0. 
\]
We compute $a, b, c, d$: 
\begin{enumerate}
	\item Since $q_2, q_3, q_4$ all give 0 when evaluated in $x_{i+1}$, we must have $g_{i+1} = p(x_{i+1}) = a q_1(x_{i+1}) = ah_i$ or $a = g_{i+1}/h_i$;
	\item Analogously, we must have $b = g_i / h_i$;
	\item Since $q_1'', q_2'', q_4''$ all give 0 when evaluated in $x_{i+1}$, we must have $\gamma_{i+1} = p''(x_{i+1}) = c q_3''(x_{i+1}) = -6c$ or $c = - \gamma_{i+1}/6$;
	\item Analogously, we must have $d = -\gamma_i / 6$. 
\end{enumerate}
Combining the above proves the claim.

On $[a, x_1]$, $g$ simply needs to be a linear function with $g(x_1) = g_1$, which means it has the form $g(x) = c(x - x_1) + g_1$ for some $c$. We must then choose $c$ such that the first derivative is continuous. An analogous argument holds for $g$ on $[x_n, b]$. 
\end{proof}

\begin{question}
	Define tri-diagonal matrices $Q = (q_{ij})_{i=1, j=2}^{n,n-1} \in \RR^{n \times (n-2)}$ and $R = (r_{ij})_{i=2, j=2}^{n-1,n-1} \in \RR^{(n-2)\times(n-2)}$ by
	\[
	q_{ij} \ceq \begin{cases}
		1/h_i, & j = i +1, \\ -1/h_{i-1} - 1/h_i, & j = i, \\ 1/h_{i-1}, & j = i- 1, 
	\end{cases}
\qquad r_{ij} = \begin{cases}
	h_i/6, &j = i+1, \\
	(h_{i-1} + h_i)/3, &j=i, \\
	h_{i-1}/6, &j=i-1. 
\end{cases}
	\]
	Prove that $\vec g, \vec \gamma$ represent a natural cubic spline if and only if $Q\T \vec g = R\vec\gamma$. 
\end{question}

\begin{proof}
	In the previous question, we saw that there was at most one candidate $g$ for the natural cubic spline which satisfies $g(x_i) = g_i$ ($i =1, \dotsc, n$) and $g''(x_i) = \gamma_i$ ($i = 2, \dotsc, n-1$). For this $g$, we know that $g$ and $g''$ are continuous, however, we do not know if $g'$ is continuous. We find that $g$ represents a natural cubic spline if and only if $g'$ is continuous at the points $x_2, \dotsc, x_{n-1}$.

	Fix $j \in \qty{2, \dotsc, n-1}$.  On the interval $[x_{j-1}, x_j]$, we have 
	\[
	g'(x_j) = \frac{g_j - g_{j-1}}{h_{j-1}} + \frac16 h_{j-1}  \qty{2 \gamma_j + \gamma_{j-1}},
	\]
	while on the interval $[x_j, x_{j+1}]$, we have
	\[
	g'(x_j) = \frac{g_{j+1} - g_j}{h_j} - \frac16 h_j \qty{\gamma_{j+1} + 2 \gamma_j}. 
	\]
	Equating the two and rearranging terms, we obtain
	\[
	\frac{1}{h_j} g_{j+1} - \qty(\frac{1}{h_j} + \frac1{h_{j-1}}) g_j + \frac1{h_{j-1}} g_{j-1} = \frac{h_j}{6} \gamma_{j+1} + \frac{h_{i-1} + h_i}{3} \gamma_j + \frac{h_{j-1}}{6} \gamma_{j-1}.
	\]
	It is easily seen that these equations can be rewritten as $Q\T \vec g = R\vec \gamma$, which proves the claim. 
\end{proof}

\begin{question}[Continuation]
	Prove that $R$ is positive definite. Deduce that, given $\vec g \in \RR^n$, there exists a unique natural cubic spline $g$ with knots at $x_1, \dotsc, x_n$ satisfying $g(x_i) = g_i$ for $i = 1, \dotsc, n$. Show further that there exists a non-negative definite matrix $K \in \RR^{n \times n}$ such that 
	\[
	\int_a^b g''(x)^2 \dd{x} = \vec g\T K\vec g. 
	\]
\end{question}

\begin{proof}
	It is immediate that the columns of $R$ are linearly independent, so $R$ is invertible and therefore positve definite. By question 8, there is exactly one $\vec \gamma$ such that there exists a natural cubic spline represented by $\vec g, \vec \gamma$, namely $\vec \gamma = R^{-1} Q\T \vec g$. 
	Note that, on $[x_i, x_{i+1}]$, the polynomial $g$ has leading coefficient $c_i = \frac16 \qty(\frac{\gamma_{i+1} - \gamma_i}{h_i})$. Therefore, by partial integration we find (analogous to the next question)
	\begin{align*}
		\int_a^b g''(x)^2 \dd{x} &= \sum_{i=1}^{n-1}\int_{x_i}^{x_{i+1}} g''(x)^2 \dd{x} = -\sum_{i=1}^{n-1} \int_{x_i}^{x_{i+1}} g'''(x) g'(x) \dd{x} \\
		&= - \sum_{i=1}^{n-1} c_i \qty(g_{i+1} - g_i) = - \sum_{i=1}^n \frac16 \qty(\frac{\gamma_{i+1} - \gamma_i}{h_i}) (g_{i+1} - g_i). 
	\end{align*}
Since every $\gamma_i$ can be written as a linear combination of the $g_i$ via $\vec \gamma = R^{-1}Q\T \vec g$, it follows that there exists a matrix $K$ such that $0 \leq \int_a^b g''(x)^2 \dd{x} = \sum_{i, j} K_{ij} g_i g_j$ for all $\vec g$, and therefore $\int_a^b g''(x)^2 \dd{x} = \vec g\T K\vec g$ and $K$ is positive semi-definite. 
\end{proof}
\begin{question}
Let $n \geq 3$, let $a \leq x_1 < \dotsb < x_n \leq b$, and let $\vec g \in \RR^n$. Prove that the natural cubic spline interpolant to $\vec g$ at $x_1, \dotsc, x_n$ is the unique minimiser of $R(\tilde g) = \int_a^b \tilde g''(x)^2 \dd{x}$ over all $\tilde g \in \SS_2[a, b]$ that interpolate $\vec g$ at $x_1, \dotsc, x_n$.

\emph{Hint:} Let $g$ be the natural cubic spline interpolant, $h \ceq \tilde g - g$, and consider $\int_a^b g''(x) h''(x) \dd{x}$. 
\end{question}

\begin{solution}
As in the hint we let $g$ be the natural cubic spline interpolant, $\tilde g\in \SS_2[a, b]$, and $h \ceq \tilde g - g$. We know that $g$ is a cubic polynomial $p_i$ on each interval $[x_i, x_{i+1}]$ (denote its leading coefficient by $c_i$), that $g''$ is continuous, and that $g'' = 0$ on $[a, x_1]$ and on $[x_n, b]$. Furthermore, we know that $h(x_i) = 0$ for all $i$. 
Using this, we can write
\begin{align*}
	\int_a^b g''(x) h''(x) \dd{x} &= \sum_{i=1}^{n-1} \int_{x_i}^{x_{i+1}} p_i''(x) h''(x) \dd{x} \\
	&= \sum_{i=1}^{n-1} \qty(\qty[p_i''(x) h'(x)]_{x = x_i}^{ x_{i+1}} - \int_{x_i}^{x_{i+1}} p_i'''(x) h'(x) \dd{x}) \\
	&= - \sum_{i=1}^{n-1} c_i \int_{x_i}^{x_{i+1}} h'(x) \dd{x} \\
	&= - \sum_{i=1}^{n-1} c_i \qty(h(x_{i+1}) - h(x_i)) = 0. 
\end{align*} 
Now we find
\begin{align*}
	R(g) &= \int_a^b g''(x)^2 \dd{x} = \int_a^b g''(x) \qty(\tilde g''(x) - h''(x)) \dd{x} = \int_a^b g''(x) \tilde g''(x) \overset{\Rm{CS}}{\leq} \sqrt{R(g)} \sqrt{R(\tilde g)},
\end{align*}
with equality if and only if $g'' = \tilde g''$ (by Cauchy-Schwarz).
Rearranging gives $\sqrt{R(g)} \leq \sqrt{R(\tilde g)} \implies R(g) \leq R(\tilde g)$. Since $\tilde g \in \SS_2[a, b]$ was arbitrary, we deduce that $g$ is a minimiser of $R$ over all function in $\SS_2$.

For uniqueness: we already know that any other minimiser $h \in \SS_2$ must satisfy $g'' = h''$ a.e., so $g - h$ is a polynomial of degree 1 a.e., and by continuity we know that $g - h$ is a polynomial of degree 1. However, since $g$ and $h$ must agree on $n \geq 3$ points it follows that $g = h$. 
\end{solution}
\end{document}

### 2020-11-22 topics_statistical_theory/topics_ES2_handin.tex
afwejinuo

--- 2020-11-22 topics_statistical_theory/topics_ES2.tex
@@
 \author{Lucas Riedstra}
-\date{...} % date
+\date{22 November 2020} % date
 
@@
 
-\begin{question}
-	Recall that the Epanechnikov kernel is a second-order kernel defined by
-	\[
-	K_E(x) = \frac3{4\sqrt5} \qty(1 - \frac{x^2}{5}) \ind_{\abs{x} \leq\sqrt5},
-	\]
-	and that $\mu_2(K_E) = 1$. Let $K_0$ be another non-negative second-order kernel with $\mu_2(K_0) = 1$. By considering $e(x) \ceq K_0(x) - K_E(x)$, or otherwise, show that $R(K_0) \geq R(K_E)$. 
-\end{question}
-
-\begin{proof}
-	We recall the definitions
-	\[
-	R(K) = \int_\RR K^2(u) \dd{u}, \quad \mu_2(K) = \int_\RR u^2 \abs{K(u)} \dd{u}, \quad 
-	\]
-	and we recall that $K$ has order 2 if and only if
-	\[
-	\int_\RR u K(u) \dd{u} = 0. 
-	\]
-
-	??
-\end{proof}
+\setcounter{question}{1}
 
@@
 
-\begin{question}
-	Let $\qty{w_i(x) \mid = 1, \dotsc, n}$ denote the effective kernel of the local polynomial estimator of order $p$ based on $(x_1, Y_1), \dotsc, (x_n, Y_n)$, and let $R$ denote a polynomial of degree at most $p$.
-Prove that if $X\T W X$ is positive definite, then
-	\[
-	\frac1n \sum_{i=1}^n w_p(x, x_i) R(x_i) = R(x)
-	\]
-	for every $x \in \RR$. 
-\end{question}
-
-\begin{proof}
-	Note that $\frac1n \sum w(x, x_i) R(x_i)$ is exactly the local polynomial estimator for data $(x_i, R(x_i))_{i=1}^n$ in the point $x$. Therefore, write $Y = (R(x_1), \dotsc, R(x_n))\T \in \RR^n$. 
-	We know that if $X\T W X$ is positive definite, then
-	\[
-	\hat m_n(x) = \hat\beta_0, \quad\hat\beta = (X\T W X)^{-1} X\T W Y. 
-	\]
-	Now, since $R$ is a polynomial of degree $p$, there exists a vector $\vec v$ such that
-	\[
-	Q_h(\cdot - x)\T \vec v = R(\cdot),
-	\]
-	and we now have
-	\[
-	Y = \mqty(R(x_1) \\ \vdots  \\ R(x_n)) = \mqty(Q_h(x_1 - x)\T \vec v \\ \vdots \\ Q_h(x_n - x)\T \vec v) = X\vec v, 
-	\]
-	and therefore
-	\[
-	\hat\beta = (X\T W X)^{-1} X\T W X \vec v = \vec v. 
-	\]
-	It is immediate that
-	\[
-	R(x) = Q_h(x - x)\T \vec v = Q_h(0)\T \vec v = v_1 = \hat m_n(x), 
-	\]
-	which proves the claim. 
-\end{proof}
-
-\begin{question}
-	Fix $\beta, L > 0$ and let $m \ceq \ceil{\beta} - 1$. Recalling the definition of the H\"older class of densities $\FF(\beta, L)$, prove that there exists $A = A(\beta, L) > 0$ such that
-	\[
-	\sup_{f \in \FF(\beta, L)} \max_{j = 0, \dotsc, m} \norm{f^{(j)}}_\infty \leq A. 
-	\]
-\end{question}
-
-\begin{proof}
-	???
-\end{proof}
-
-\begin{question}
-	Verify that the local constant and local linear kernel regression estimators have the forms given in the lectures. 
-\end{question}
-
-\begin{proof}
-	In the local constant case, we have $Q_h(u) = 1$ and therefore $X = \vec e$ (with $\vec e$ the all-ones vector).
-	It follows that 
-	\[
-	\hat m_n(x) = \hat\beta = (\vec e\T W \vec e)^{-1} \vec e\T WY = \frac1{\sum_{i=1}^n W_{ii}} \cdot \sum_{i=1}^n W_{ii} Y_i = \frac{\sum_{i=1}^n K_h(x_i - x)Y_i}{\sum_{i=1}^n K_h(x_i - x)},
-	\]
-	which corresponds with the expression from the lecture notes.
-
-	In the local linear case, we have $Q_h(u) = (1, u/h)\T$. Define $z_i = (x_i - x) / h$ and $W_i = K_h(x_i - x)$, we have
-	\[
-	X = \mqty[ 1 & z_1 \\ \vdots & \vdots \\ 1 & z_n], \quad X\T W = \mqty[ W_1 & \dotsb & W_n \\ z_1 W_1 & \dotsb & z_n W_n]. 
-	\]
-
-	We can therefore compute, using $\sum_i z_i^r W_i = n h^{-r} s_r(x)$, that 
-	\[
-	X\T W X = \mqty[ \sum_i W_i & \sum_i z_i W_i \\ \sum_i z_i W_i & \sum_i z_i^2 W_i] = n\mqty[ s_0(x) & h^{-1} s_1(x) \\
-	h^{-1} s_1(x) & h^{-2} s_2(x)],
-	\]
-	which gives
-	\[
-	(X\T W X)^{-1} = \frac{h^2}{n}\cdot \frac1{s_0(x) s_2(x) - s_1(x)^2} \mqty[h^{-2} s_2(x) & - h^{-1} s_1(x) \\ -h^{-1} s_1(x) & s_0(x)].
-	\]
-
-	Therefore we have
-	\[
-	X\T W Y = \mqty[ \sum_i W_i Y_i \\ \sum_i z_i W_i Y_i],
-	\]
-	and so
-	\begin{align*}
-	\hat\beta_0 &= 
-	\frac{h^2}{n}\cdot \frac1{s_0(x) s_2(x) - s_1(x)^2} \qty(h^{-2} s_2(x) \sum_i W_i Y_i - h^{-1} s_1(x) \sum_i z_i W_i Y_i)\\
-	&= \frac1n \sum_{i=1}^n \frac{s_2(x) - s_1(x) (x_i - x)}{s_0(x) s_2(x) - s_1(x)^2} W_i Y_i, 
-	\end{align*}
-	which corresponds with the expression from the lecture notes. 
-\end{proof}
-
-\begin{question}
-	In the random design nonparametric regression model for i.i.d.\ pairs $(X_1, Y_1), \dotsc, (X_n, Y_n)$, each having joint density $f_{X, Y}$, observe that the regression function $m$ may be expressed as
-	\[
-	m(x) = \int_{-\infty}^\infty y \frac{f_{X, Y}(x, y)}{f_X(x)} \dd{y}, 
-	\]
-	where $f_X$ is the marginal density of $X_1$.
-Find the estimator $m(x)$ that results from estimating $f_X$ and $f_{X, Y}$ using kernel density estimators with symmetric kernel $K$ (and the corresponding product kernel in the latter case) and a common bandwidth.
-\end{question}
-
-\begin{proof}
-	We plug in 
-	\begin{align*}
-	\hat m(x) &= \int_{-\infty}^\infty y \frac{\frac1n \sum_{i=1}^n K_h(x - X_i) K_h( y - Y_i)}{\frac1n \sum_{i=1}^n K_h(x - X_i)} \dd{y} \\
-	&= \frac{ \sum_{i=1}^n K_h(X_i - x) Z_i(y)}{\sum_{i=1}^n K_h(X_i - x)},
-	\end{align*}
-where $Z_i(y) = \int_{-\infty}^\infty y K_h(y - Y_i) \dd{y}$. Assuming that $K$ is a second-order kernel, we have
-\begin{align*}
-Z_i(y) &= \int_{-\infty}^\infty y K_h(y - Y_i) \dd{y} = \int_{-\infty}^\infty (y + Y_i) K_h(y) \dd{y} \\
-&= \int_{-\infty}^\infty y K_h(y) \dd{y} + Y_i \int_{-\infty}^\infty K_h(y)\dd{y} = Y_i, 
-\end{align*}
-and so $\hat m(x)$ is simply the Nadaraya-Watson estimator (the local polynomial estimator with $p = 0$). 
-\end{proof}
-
-\begin{question}
-	Let $a \leq x_1 < \dotsb < x_n \leq b$, and let $h_i = x_{i+1} - x_i$ for $i = 1, \dotsc, n-1$.
-Given $\vec g \in \RR^n$ and $\vec \gamma = (\gamma_2, \dotsc, \gamma_{n-1})\T \in \RR^{n-2}$, show that if there is a natural cubic spline $g$ with $g(x_i) = g_i$ for $i =1 , \dotsc, n$ and $g''(x_i) = \gamma_i$ for $i = 2, \dotsc, n-1$ then
-	\[
-	g(x) = \frac{(x - x_i) g_{i+1} + (x_{i+1} - x)g_i}{h_i} - \frac16(x-x_i)(x_{i+1}-x)\qty{\qty(1 + \frac{x - x_i}{h_i}) \gamma_{i+1} + \qty(1 + \frac{x_{i+1} - x}{h_i})\gamma_i}
-	\]
-	for $x \in [x_i, x_{i+1}]$ and $i = 1, \dotsc, n-1$. Find the corresponding expressions for $g$ on $[a, x_1]$ and $[x_n, b]$. 
-\end{question}
-
-\begin{proof}
-	On the interval $[x_i, x_{i+1}]$, $g$ must be a cubic polynomial $p$ such that $p(x_i) = g_i$, $p(x_{i+1}) = g_{i+1}$, $p''(x_i) = \gamma_i$, $p''(x_{i+1}) = \gamma_{i+1}$ (where $\gamma_1 = \gamma_n = 0$).
-Therefore, $p$ must have an expansion in the basis
-	\begin{gather*}
-	\qty{q_1(x), q_2(x), q_3(x), q_4(x)} \\\ceq \qty{x - x_i, x_{i+1} - x, (x - x_i)(x_{i+1} - x)(1 + \frac{x - x_i}{h_i}), (x - x_i)(x_{i+1} - x) (1 + \frac{x_{i+1} - x}{h_i})}.
-	\end{gather*}
-Therefore, write $p(x) = a q_1(x) + b q_2(x) + c q_3(x) + d q_4(x)$. Note that 
-\[
-q_3''(x_i) = 0, \quad q_3''(x_{i+1}) = -6, \quad q_4''(x_i) = -6, \quad q_4''(x_{i+1}) = 0. 
-\]
-We compute $a, b, c, d$: 
-\begin{enumerate}
-	\item Since $q_2, q_3, q_4$ all give 0 when evaluated in $x_{i+1}$, we must have $g_{i+1} = p(x_{i+1}) = a q_1(x_{i+1}) = ah_i$ or $a = g_{i+1}/h_i$;
-	\item Analogously, we must have $b = g_i / h_i$;
-	\item Since $q_1'', q_2'', q_4''$ all give 0 when evaluated in $x_{i+1}$, we must have $\gamma_{i+1} = p''(x_{i+1}) = c q_3''(x_{i+1}) = -6c$ or $c = - \gamma_{i+1}/6$;
-	\item Analogously, we must have $d = -\gamma_i / 6$. 
-\end{enumerate}
-Combining the above proves the claim.
-
-On $[a, x_1]$, $g$ simply needs to be a linear function with $g(x_1) = g_1$, which means it has the form $g(x) = c(x - x_1) + g_1$ for some $c$.
-We must then choose $c$ such that the first derivative is continuous.
-An analogous argument holds for $g$ on $[x_n, b]$.
-\end{proof}
-
-\begin{question}
-	Define tri-diagonal matrices $Q = (q_{ij})_{i=1, j=2}^{n,n-1} \in \RR^{n \times (n-2)}$ and $R = (r_{ij})_{i=2, j=2}^{n-1,n-1} \in \RR^{(n-2)\times(n-2)}$ by
-	\[
-	q_{ij} \ceq \begin{cases}
-		1/h_i, & j = i +1, \\ -1/h_{i-1} - 1/h_i, & j = i, \\ 1/h_{i-1}, & j = i- 1, 
-	\end{cases}
-\qquad r_{ij} = \begin{cases}
-	h_i/6, &j = i+1, \\
-	(h_{i-1} + h_i)/3, &j=i, \\
-	h_{i-1}/6, &j=i-1. 
-\end{cases}
-	\]
-	Prove that $\vec g, \vec \gamma$ represent a natural cubic spline if and only if $Q\T \vec g = R\vec\gamma$. 
-\end{question}
-
-\begin{proof}
-	In the previous question, we saw that there was at most one candidate $g$ for the natural cubic spline which satisfies $g(x_i) = g_i$ ($i =1, \dotsc, n$) and $g''(x_i) = \gamma_i$ ($i = 2, \dotsc, n-1$).
-For this $g$, we know that $g$ and $g''$ are continuous, however, we do not know if $g'$ is continuous.
-We find that $g$ represents a natural cubic spline if and only if $g'$ is continuous at the points $x_2, \dotsc, x_{n-1}$.
-
-	Fix $j \in \qty{2, \dotsc, n-1}$.  On the interval $[x_{j-1}, x_j]$, we have 
-	\[
-	g'(x_j) = \frac{g_j - g_{j-1}}{h_{j-1}} + \frac16 h_{j-1}  \qty{2 \gamma_j + \gamma_{j-1}},
-	\]
-	while on the interval $[x_j, x_{j+1}]$, we have
-	\[
-	g'(x_j) = \frac{g_{j+1} - g_j}{h_j} - \frac16 h_j \qty{\gamma_{j+1} + 2 \gamma_j}. 
-	\]
-	Equating the two and rearranging terms, we obtain
-	\[
-	\frac{1}{h_j} g_{j+1} - \qty(\frac{1}{h_j} + \frac1{h_{j-1}}) g_j + \frac1{h_{j-1}} g_{j-1} = \frac{h_j}{6} \gamma_{j+1} + \frac{h_{i-1} + h_i}{3} \gamma_j + \frac{h_{j-1}}{6} \gamma_{j-1}.
-	\]
-	It is easily seen that these equations can be rewritten as $Q\T \vec g = R\vec \gamma$, which proves the claim. 
-\end{proof}
-
-\begin{question}[Continuation]
-	Prove that $R$ is positive definite.
-Deduce that, given $\vec g \in \RR^n$, there exists a unique natural cubic spline $g$ with knots at $x_1, \dotsc, x_n$ satisfying $g(x_i) = g_i$ for $i = 1, \dotsc, n$.
-Show further that there exists a non-negative definite matrix $K \in \RR^{n \times n}$ such that
-	\[
-	\int_a^b g''(x)^2 \dd{x} = \vec g\T K\vec g. 
-	\]
-\end{question}
-
-\begin{proof}
-	It is immediate that the columns of $R$ are linearly independent, so $R$ is invertible and therefore positve definite.
-By question 8, there is exactly one $\vec \gamma$ such that there exists a natural cubic spline represented by $\vec g, \vec \gamma$, namely $\vec \gamma = R^{-1} Q\T \vec g$.
-	Note that, on $[x_i, x_{i+1}]$, the polynomial $g$ has leading coefficient $c_i = \frac16 \qty(\frac{\gamma_{i+1} - \gamma_i}{h_i})$.
-Therefore, by partial integration we find (analogous to the next question)
-	\begin{align*}
-		\int_a^b g''(x)^2 \dd{x} &= \sum_{i=1}^{n-1}\int_{x_i}^{x_{i+1}} g''(x)^2 \dd{x} = -\sum_{i=1}^{n-1} \int_{x_i}^{x_{i+1}} g'''(x) g'(x) \dd{x} \\
-		&= - \sum_{i=1}^{n-1} c_i \qty(g_{i+1} - g_i) = - \sum_{i=1}^n \frac16 \qty(\frac{\gamma_{i+1} - \gamma_i}{h_i}) (g_{i+1} - g_i). 
-	\end{align*}
-Since every $\gamma_i$ can be written as a linear combination of the $g_i$ via $\vec \gamma = R^{-1}Q\T \vec g$, it follows that there exists a matrix $K$ such that $0 \leq \int_a^b g''(x)^2 \dd{x} = \sum_{i, j} K_{ij} g_i g_j$ for all $\vec g$, and therefore $\int_a^b g''(x)^2 \dd{x} = \vec g\T K\vec g$ and $K$ is positive semi-definite.
-\end{proof}
+\setcounter{question}{9}
 \begin{question}

### 2020-11-23 distribution_theory/C4.tex
wtfff

@@
 \begin{lemma}
-	For $u \in \EE'(\RR^n)$, the Fourier transform $\hat u \in \SS'(\RR^n)$ can be identified with the smooth (analytic) function $\lambda \mapsto \ang{u(x), e^{-i\lambda\vdot x}}$, which we will denote $\hat u(\lambda)$.
+	For $u \in \EE'(\RR^n)$, the Fourier transform $\hat u \in \SS'(\RR^n)$ can be identified with the smooth (real-analytic) function $\lambda \mapsto \ang{u(x), e^{-i\lambda\vdot x}}$, which we will denote $\hat u(\lambda)$.
 \end{lemma}
@@
 There also exist nonconstructive proofs for the previous theorem using Hahn-Banach.
+
+\subsection{Paley-Wiener-Schwartz theorem}
+We have seen that if $u \in \EE'(\RR^n)$, then $\hat u$ is equivalent to the real-analytic function $\lambda \mapsto \ang{u(x), e^{-i\lambda \vdot x}}$ and $\abs{\hat u(\lambda)} \les \ang{\lambda}^N$ for some $N \geq 0$.
+We consider the complex extension of $\hat u$, and we call $\hat u(z)$ the \emph{Fourier-Laplace transform} of $u$. 
+Note that
+\[
+\pdv{\hat u}{\bar z_i} = \ang{u, \pdv{\bar z_i} e^{-iz\vdot x}} = 0 \quad(i = 1, \dotsc, n), 
+\]
+so $\hat u(z)$ is complex-analytic in each component $z_i$.
+
+We can also estimate the size of $\hat u(z)$:
+\begin{lemma}
+	If $u \in \EE'(\RR^n)$ and $\supp(u) \subseteq \clos{B_\delta}$, then there exist nonnegative constant $C, N$ such that
+	\[
+	\norm{\hat u(z)} \leq C (1 + \norm{z})^N e^{\delta \abs{\Im z}} \qquad \forall z \in \CC^n. 
+	\]
+\end{lemma}
+
+\begin{proof}
+	Let $\psi \in C^\infty(\RR)$ be such that $\psi(\tau) = 1$ for $\tau \geq -\frac12$ and $\psi(\tau) = 0$ for $\tau \leq -1$.  Introduce for $\eps > 0$
+	\[
+	\phi_\eps(x) \ceq \psi(\eps(\delta - \norm{x})),
+	\]
+	then we have $\phi_\eps \in \DD(\RR^n)$ with $\phi_\eps(x) = 0$ for $\norm{x}\geq \delta + \eps^{-1}$, and $\phi_\eps(x) = 1$ for $\norm{x} \leq \delta + \frac12\eps^{-1}$.
+In particular, every $\phi_\eps$ is 1 on a neighbourhood of $\supp(u) \subseteq \clos{B_\delta}$.
+
+	Therefore, we have
+	\[
+	\hat u(z) = \ang{u(x), e^{-iz\vdot x}} = \ang{u(x), \phi_\eps(x) e^{-iz\vdot x}},
+	\]
+	and since $u \in \EE'(\RR^n)$, by the seminorm condition we have nonnegative $C, N$ such that
+	\[
+	\hat u(z) \leq C \sum_{\abs{\alpha} \leq N} \sup \abs{\pt^\alpha_x \qty(\phi_\eps(x) e^{-iz \vdot x})}.
+	\]
+	By definition we have $\abs{\pt_x^\beta \phi_\eps} \les \eps^{\abs{\beta}}$ while $\abs{\pt_x^\gamma e^{-iz \vdot x}} \les \norm{z}^{\abs{\gamma}} e^{(\delta + \eps^{-1}) \abs{\Im z}}$ on $\supp\phi_\eps$.
+
+	Applying Leibniz, we obtain
+	\[
+	\abs{\hat u(z)} \leq C \sum_{\abs{\beta} + \abs{\gamma} \leq N} \norm{z}^{\abs{\beta}} e^{\abs{\gamma}} e^{(\delta + \eps^{-1}) \Im z}, 
+	\]
+	and this holds for all $\eps > 0$, so we can plug in $\eps = \norm{z}$ and obtain the result (since $\Im z / \norm{z}$ is bounded). 
+\end{proof}
+
+So if $u \in \EE'(\RR^n)$ with $\supp(u) \subseteq \clos{B_\delta}$, then $\hat u(z)$ is complex analytic and obeys $\abs{\hat u(z)} \leq (1 + \norm{z})^N e^{\delta \abs{\Im z}}$.
+The PWS theorem addresses the converse: if a complex analytic function obeys the estimate we just saw, is it the fourier transform of some distribution?
+
+\begin{theorem}
+	\begin{enumerate}[(a)]
+		\item If $u \in \EE'(\RR^n)$ with $\ord(u) = N$ and $\supp u \subseteq \clos{B_\delta}$, then $\hat u(z)$ is entire and
+		\begin{equation} \label{eq:fourier_transform_estimate}
+			\abs{\hat u(z)} \les (1 + \norm{z})^N e^{\delta \abs{\Im z}}. 
+		\end{equation}
+
+	Conversely, if $U(z)$ is entire and satisfies \cref{eq:fourier_transform_estimate} for some $N$, then $U = \hat u$ for some $u \in \EE'(\RR^n)$ with $\supp(u) \subseteq \clos{B_\delta}$.
+
+	\item If $u \in \DD(\RR^n)$ and $\supp(u) \subseteq \clos{B_\delta}$, then for $M = 0, 1, 2, \dotsc$ we have 
+	\begin{equation}\label{eq:fourier_transform_estimate_2}
+		\abs{\hat u(z)} \les_M (1 + \norm{z})^{-M} e^{\delta \abs{\Im z}}.
+	\end{equation}
+	Conversely, if $U(z)$ is entire and satisfies \cref{eq:fourier_transform_estimate_2}, then $U = \hat u$ for some $u \in \DD(\RR^n)$ with $\supp(u) \subseteq \clos{B_\delta}$. 
+	\end{enumerate}
+\end{theorem}
+
+\begin{proof}
+	\TODO write this out (lecture 12)
+\end{proof}
+
+\subsection{Oscillatory integrals}
+We will study integrals of the form $\int_{\RR^n} e^{i\Phi(x, \theta)} a(x, \theta) \dd{\theta}$, where $\Phi$ is called the \emph{phase function} and $a$ the \emph{amplitude} of the signal.
+We wil use oscillations from the $e^{i\Phi}$-term to control the growth, while $a(\cdot, \theta)$ is allowed to grow modestly with $\theta$.
+
+\begin{lemma}[Riemann-Lebesgue]
+	If $f \in L^1(\RR)$, then $\abs{\hat f(\lambda)} \to 0$ when $\abs{\lambda} \to\infty$. 
+\end{lemma}
+\begin{proof}
+	Assume $f \in L^1$ is continuous, then setting $x' = x + \pi/\lambda$, we have
+	\begin{align*}
+		\hat f(\lambda) &= \int e^{-i\lambda \cdot x} f(x) \dd{x} = \frac12\int \qty( e^{-i\lambda \vd x}f(x) \dd{x} + e^{-i\lambda \vd (x + \pi/\lambda)} f(x + \pi/\lambda))\dd{x} \\
+		&= \frac12 \int e^{-i\lambda \vd x} \qty[ f(x) - f(x + \pi/\lambda)] \dd{x}
+	\end{align*}
+	Now let $\eps > 0$ and choose $R = R(\eps)$ such that $\int_{\norm{x}> R} \abs{f(x) - f(x + \pi/\lambda)} \dd{x} < \frac\eps2$.
+	By the dominated convergence theorem, we can also choose $\lambda = \lambda(\eps, R)$ large enough such that
+	\[
+	\int_{\norm{x}< R} \abs{e^{-i\lambda x} \qty[f(x) - f(x + \pi/\lambda)]} \dd{x} < \frac\eps2. 
+	\]  
+	So for $f$ we have $\abs{\hat f(\lambda)} < \frac\eps2$ for $\abs{\lambda}$ large enough, which proves the result for continuous functions.
+
+	Now we use that continuous functions are dense in $L^1$, so for any $f \in L^1$, pick $g \in C(\RR) \cap L^1$ such that $\norm{f - g}_{L^1} < \frac\eps2$, then it is easily checked that $\abs{\hat f(\lambda)} \leq \norm{f - g}_{L^1} + \abs{\hat g(\lambda)} < \eps $ for $\lambda$ sufficiently large, which proves the claim.
+
+\end{proof}
+
+Now suppose $\phi \in \DD(\RR)$ and $\Phi \in C^\infty(\RR)$ with $\Phi'$ nowhere 0. Consider
+\[
+I(\lambda) \ceq \int e^{i\lambda \Phi(\theta)} \phi(\theta) \dd{\theta}. 
+\]
+Note that $\abs{\Phi'(\theta)} \ges 1$ for $\theta \in \supp\phi$ (since $\supp\phi$ is compact), and we can write
+\[
+I(\lambda) = \int \frac1{i\lambda} \frac{\phi(\theta)}{\Phi'(\theta)} \dv{\theta}\qty(e^{i\lambda \Phi(\theta)}) \dd{\theta},
+\]
+and with repeated integration by parts we find that
+\[
+\abs{I(\lambda)} \leq \ang{\lambda}^{-N} \quad\text{for any $N \geq 0$}. 
+\]
+A natural question is what happens if $\Phi'(\theta) = 0$ somewhere.
+
+\begin{lemma}[Stationary phase]
+	Suppose $\Phi \in C^\infty(\RR)$ with $\Phi' \neq 0$ on $\RR \setminus \qty{0}$ and $\Phi(0) = \Phi'(0) = 0$, $\Phi''(0) \neq 0$. Then, for $\chi \in \DD(\RR)$, we have
+	\[
+	\int e^{i\lambda \Phi(\theta)} \chi(\theta) \dd{\theta} \les \ang{\lambda}^{-1/2}. 
+	\]
+\end{lemma}
+
+\begin{proof}
+	\TODO write this (lecture 13, a terrible proof with lots of differentiation and garbage). 
+\end{proof}

### 2020-11-23 inverse_problems/C5.tex
wtfff

@@
 	Note that if $\mu$ is a probability measure, then so is $\mu(g \in \cdot)$. 
+\end{definition}
+
+\begin{definition}
+	Let $(\Omega, \FF, \mu)$ be a measure space and let $g \colon (\Omega, \FF) \to (\RR, \BB \RR)$ be measurable and non-negative. Then the map 
+	\[
+	\nu \colon \FF \to [0, \infty] \colon F \mapsto \int_F g \dd{\mu}
+	\]
+	defines a measure on $(\Omega, \FF)$, called the \emph{measure with ($\mu$-)density $g$}. If $\nu$ is a probability measure, $g$ is called a \emph{($\mu$-)probability density}. 
 \end{definition}
@@
 
-	For an event $F \in \FF$, we define the \emph{probability of $F$ given $Y = y$} as $\EE[\ind_F \mid Y = y]$.
+	%For an event $F \in \FF$, we define the \emph{probability of $F$ given $Y = y$} as $\EE[\ind_F \mid Y = y]$.
 \end{definition}
@@
 
-It can be shown that $(y, F) \mapsto \EE[\ind_F \mid Y= y]$ is a Markov kernel. This is clarified in the following theorem:
+It can be shown that $(y, F) \mapsto \EE[\ind_F \mid Y= y]$ (which represents the probability of $F$ given $Y = y$) is a Markov kernel. This is clarified in the following theorem:
 
@@
 We can use the above theorem to consider the density function of $U$ conditional on $Y$:
-\begin{theorem}
+\begin{theorem}\label{thm:conditional_density}
+	% TODO niet helemaal hetzelfde als thm 6.3.10 maar whatever
 	Under the assumptions of the previous lemma, we have $\PP(U \in \cdot \mid Y = y) \ll \nu_U$, and its $\nu_U$ density is
@@
 \end{theorem}
+
+\begin{definition}
+	Let $g_U, g_Y, g_{U, Y}, g_{U\mid Y= y}, g_{Y \mid U = u}$ be the probability densities from the previous theorem.
+Then $g_U$ is called the \emph{marginal probability density} of $U$, $g_{U, Y}$ is called the \emph{joint probability density} of $U$ and $Y$, and $g_{U \mid Y = y}$ is called the \emph{conditional density} of $U$ given $Y = y$.
+\end{definition}
+
+\subsection{Bayesian statistics}
+\subsubsection{Statistical models}
+\begin{definition}
+	Let $\XX, \YY$ be separable Banach spaces.
+We call $\XX$ the \emph{parameter space} and $\YY$ the \emph{data space}.
+Let $\Pp \ceq \qty{M(\cdot \mid u) : u \in \XX}$, where $M$ is a Markov kernel from $(\XX, \BB\XX)$ to $(\YY, \BB\YY)$.
+The tuple $(\YY, \Pp)$ is called the \emph{statistical model}.
+\end{definition}
+
+Usually, we distinguish between \emph{parametric models} (where $\XX$ is finite-dimensional) and \emph{nonparametric models}. However, in Bayesian statistics, this distinction rarely matters.
+
+Given a ``true'' parameter $u^* \in \XX$, a distribution $Y \sim M(\cdot \mid u^*)$ and a realisation $y$ of $Y$, we aim to find $u^*$ based on $y$.
+The probability measure $M(\cdot \mid u^*)$ is called the \emph{data-generating distribution}.
+
+\begin{definition}
+	Let $(\YY, \Pp)$ be a statistical model and $L \colon (\XX \times \YY, \BB\XX \otimes \BB\YY) \to (\RR, \BB\RR)$ such that
+	\[
+	\Pp = \qty{F \mapsto \int_F L(y \mid u) \dd\mu(y) : u \in \XX}
+	\]
+	for some measure $\mu$ on $(\YY, \BB\YY)$. Then $L$ is called the \emph{(data) likelihood}. 
+\end{definition}
+The likelihood is a conditional density $g_{Y\mid U = u}$ for some random variable $U$.
+
+\subsubsection{Bayes' formula}
+Assume we are given a likelihood $L = f_{Y \mid U = u}$ for some $U$. We are interested in finding $f_{U \mid Y = y}$, the posterior density function of $U$. For this we use Bayes's formula:
+\begin{theorem}[Bayes]
+	Let $U, Y$ be random variables as in \cref{thm:conditional_density}. Then we have
+	\[
+	g_{U \mid Y =y}(u) = \frac{g_{Y \mid U = u}(y) g_{U}(u)}{g_Y(y)}
+	\]
+	for $u \in \XX$, $\nu_U$-a.e., and $y \in \YY$, $\PP(Y \in \cdot)$-a.e.\ with $g_Y(y) > 0$. 
+\end{theorem}

### 2020-11-23 modern_statistical_methods/C3.tex
wtfff

@@
 Therefore we must place further assumptions on the eigenvectors.
+
+\begin{definition}
+	Let $S, U \in \RR^{n \times d}$ be matrices with orthogonal columns. We define the \emph{$\sint$ distance} between the column space of $S$ and $U$ as
+	\[
+	\sint(S,U) \ceq \norm{\Pi_S(I - \Pi_U)}_F = \norm{\Pi_U(I - \Pi_S)}_F
+	\]
+\end{definition}
+To see that the last equality holds, note that
+\begin{align*}
+	\norm{\Pi_S(I - \Pi_U)}_F^2 &= \Tr\qty((I - \Pi_U) \Pi_S^2 (I - \Pi_U)) = \Tr\qty(\Pi_S(I - \Pi_U)) = \Tr(\Pi_S) - \Tr(\Pi_S\Pi_U) \\
+	&= \rank(S) - \Tr(\Pi_S\Pi_U) = \rank(U) - \Tr(\Pi_U\Pi_S) = \norm{\Pi_U(I - \Pi_S)}_F^2. 
+\end{align*}
+
+In the fourth example sheet we prove the following:
+\begin{proposition} \label{prop:sint_distance}
+	Let $S, U$ be matrices with orthogonal columns and equal rank. Then
+	\[
+	\min_{O \in \RR^{d \times d} \text{ }\Rm{orthogonal}} \norm{S - UO}_F \leq \sqrt2 \sint(S, U). 
+	\]
+\end{proposition}
+
+\begin{theorem}[Davis-Kahan]
+	Let $\Sigma$ be a covariance matrix with eigenvector decomposition $\Sigma = U\Lambda U\T$ (where $\lambda_1 \geq \dotsb \geq \lambda_d$).
+Write $\Sigma = U_0 \Lambda_0 U_0\T + U_1 \Lambda_1 U_1\T$ where $U_0 = [u_1, \dotsc, u_r]$.
+
+	For a perturbed matrix $\hat\Sigma = \Sigma + E$ with $E$ symmetric, define an analogous decomposition $\hat\Sigma = \hat U_0 \hat\Lambda_0 \hat U_0\T + \hat U_1 \hat\Lambda_1 \hat U_1\T$.
+If $\lambda_r - \hat\lambda_{r+1} > 0$, then
+	\[
+	\sint (U_0, \hat U_0)  \leq \frac{\norm{E}_F}{\lambda_r - \hat\lambda_{r+1}}. 
+	\]
+\end{theorem}
+
+\begin{proof}
+	We write
+	\[
+	\norm{\Pi_{U_0}(I - \Pi_{\hat U_0})}_F = \norm{U_0 U_0\T \hat U_1 \hat U_1\T}_F = \norm{U_0\T \hat U_1}_F, 
+	\]
+	and note that 
+	\[
+	\Sigma U_0 = U_0\Lambda U_0\T U_0 + U_1 \Lambda_1 U_1\T U_0 =  U_0\Lambda_0, \quad \hat\Sigma \hat U_1 = \hat U_1 \hat\Lambda_1. 
+	\]
+	We therefore find 
+	\begin{align*}
+	E U_0 &= (E + \Sigma) U_0 - \Sigma U_0 = \hat\Sigma U_0 - U_0 \Lambda_0 \\
+	\hat U_1\T EU_0 &= \hat U_1\T \hat\Sigma U_0 - \hat U_1\T U_0 \Lambda_0 = \hat\Lambda_1 \hat U_1\T U_0 - \hat U_1\T U_0 \Lambda_0. 
+	\end{align*}
+By the triangle inequality we have
+\begin{align*}
+\norm{\hat U_1\T E U_0}_F &\geq \norm{\hat U_1\T U_0 \Lambda_0}_F - \norm{\hat\Lambda_1 \hat U_1\T U_0}_F \\
+&\overset\star\geq \lambda_r \norm{\hat U_1\T U_0} - \hat\lambda_{r+1} \norm{\hat U_1\T U_0}_F = (\lambda_r - \hat\lambda_{r+1}) \norm{U_1\T U_0}_F = (\lambda_r - \hat\lambda_{r+1}) \sint(U_0, \hat U_0).
+\end{align*}
+Noting that $\norm{\hat U_1\T E U_0}_F \leq \norm{E}_F$,
+since
+\[
+\norm{E}_F = \norm{\hat U\T E U} = \norm{\mqty[\hat U_0\T \\ \hat U_1\T ] E \mqty[U_0 & U_1]}_F \geq \norm{\hat U_1\T E U_0}_F. 
+\] yields the result. 
+\end{proof}
+
+The Davis-Kahan theorem is not directly useful since the result depends on the gap $\lambda_r - \hat\lambda_{r+1}$.
+However, the following variant was proved in 2014, whose result depends only on the eigenvalues of $\Sigma$:
+\begin{theorem}
+	If $\lambda_r - \lambda_{r+1} > 0$, then
+	\[
+	\sint (U_0, \hat U_0) \leq \frac{2\min \qty{\sqrt r \norm{E}_\Rm{op}, \norm{E}_F}}{\abs{\lambda_r - \lambda_{r+1}}}. 
+	\]
+\end{theorem}
+
+\subsubsection{The spiked covariance model}
+As an application of the Davis-Kahan theorem, consider the model where
+\[
+\Sigma = I + \theta v v\T, \qquad  v\in S^{d-1}. 
+\]
+Any random vector with covariance $\Sigma$ can be represented as $\eta \sqrt\theta v + \eps$, where $\eta, \eps_1, \dotsc, \eps_d$ are independent, mean 0, variance 1 random variables.
+The samples are a random ``spike'' with scale $\sqrt\theta$ in a fixed direction $v$, with independent noise added to each entry.
+
+The eigenvalues of $\Sigma$ are $1+\theta, 1, \dotsc, 1$, and $v$ is the top eigenvector. Using the variant of the Davis-Kahan theorem and \cref{prop:sint_distance} we find 
+\[
+\min_{z = \pm 1} \norm{v - z\hat v} \leq \sqrt2 \sint(v, \hat v) \leq \frac{\norm{\hat \Sigma - \Sigma}_\Rm{op}}{(1 + \theta) - 1}
+\]
+If we assume that the rows of $X$ are vector sub-Gaussian, we can bound $\norm{\hat\Sigma - \Sigma}_\Rm{op}$ in probability, and we find that 
+\[
+\PP\qty(\min_{z = \pm 1} \norm{v - z\hat v} \geq C \frac{1 + \theta}{\theta} \qty(\frac{d + \delta}{n} \lor \sqrt{\frac{d + \delta}{n}})) \leq e^{-\delta}. 
+\]
+We conclude that if $\frac{1+\theta}{\theta} \sqrt{\frac dn} \to 0$, then $\hat v$ estimates $v$ consistently.
+
+\subsubsection{High-dimensional asymptotics}
+{\color{red} Non-examinable}.
+
+\subsection{Sparse PCA}
+We have seen that in the spiked covariance model, PCA is consistent when $d/n \to 0$, but PCA can fail if $d/n$ is bounded away from zero.
+Considering the case $d/n \to \gamma > 0$, we can ask if there exists a better estimator.
+Minimax lower bounds show that in general, the answer is negative, and therefore we must place some assumptions on $\Sigma$.
+
+Again, we assume $\Sigma = \theta vv\T + I$. Define
+\[
+\norm{v}_0 \ceq \abs{\qty{k \mid v_k \neq 0}} \quad\text{(the amount of nonzero components of $v$)}
+\]
+We will assume that $v$ is spare, i.e., $\norm{v}_0 = k \ll d$. This means that the extra variance in the data along the ``spike'' only involved a few variables. 
+We can now define an estimator similar to best subset selection
+\[
+\hat v^{(0, k')} \ceq \argmax_{w \in S^{d-1}, \norm{w}_0 \leq k'} w\T \hat\Sigma w. 
+\]
+
+\begin{proposition}
+	Let $\hat\Sigma$ be the sample covariance of $n$ i.i.d.\ $N(0, \Sigma)$ random variables. Suppose $k \leq d/2$ and $k \leq k'$, then 
+	\[
+	\PP\qty(\min_{z = \pm 1} \norm{v - z\hat v^{(0, k')}}_2 \geq  C\frac{1+\theta}{\theta} \qty[ \sqrt{\eta_n \lor \eta_n}]) \leq e^{-\delta} \quad\forall \delta > 0
+	\]
+	where 
+	\[
+	\eta_n = \frac{(k + k') \log(de^2/(k+ k')) + \delta}{n}.
+	\]
+\end{proposition}
+Note that the dependence on $d$ is now logarithmic, so morally, we only need $\frac{\max(k, k')}{n} \to 0$ for consistency.
+
+The problem with this estimator is that it is hard to compute. The SCoTLASS estimator relaxes the estimator with 
+\[
+\hat v^{(1, \lambda)} \ceq \argmax_{w \in S^{d-1}, \norm{w}_1 \leq \lambda} w\T \hat\Sigma w.
+\]
+The optimisation problem is not convex (since we are maximising a convex function instead of minimising), but is easier to compute and has similar convergence rates as the estimator $\hat v^{(0, k)}$.
+We can relax the problem even further (see example sheet) which gives an easily computable solution.
+
+It has been shown that sparce PCA has no estimators computable in polynomial time which achieve minimax rates for certain distribution classes, which means that sparce PCA has some inherent computational difficulty.

### 2020-12-01 inverse_problems/C5.tex
almost end of termgit add .!

@@
 \end{theorem}
+
+\begin{definition}
+	In the formulation of Bayes' formula: 
+	\begin{enumerate}
+		\item $Z(y) \ceq g_{Y}(y)$ is called the \emph{model evidence} or \emph{marginal likelihood};
+		\item $L(y\mid u) \ceq g_{Y\mid U =u}(y)$ is called the \emph{(data) likelihood};
+		\item $\mu_0 \ceq \PP(U \in \cdot)$ is called the \emph{prior measure};
+		\item $\mu_\Rm{post} \ceq \PP(U \in \cdot \mid Y = y)$ is called the \emph{posterior measure}. 
+	\end{enumerate}
+\end{definition}
+
+If we assume that $\nu_U \ceq \mu_0$ (we can choose $\nu_U$ freely so long as $\mu_0$ has a probability density w.r.t.\ $\nu_0$), we obtain from Bayes' formula the formulation
+\[
+\dv{\mu_\Rm{post}}{\mu_0}(u) = \frac{L(y \mid u)}{Z(y)} \qquad(\mu_0\text{-a.s.}).
+\] 
+If $\mu_0$ has a density with respect to some other measure $\nu$ (say, Lebesgue measure), then we obtain
+\[
+\dv{\mu_\Rm{post}}{\nu} = \dv{\mu_\Rm{post}}{\mu_0} \dv{\mu_0}{\nu} = \frac{L(y \mid u)}{Z(y)} \dv{\mu_0}{\nu}. 
+\]
+In practice, this means that computing the density of the posterior measure comes down to multiplying the likelihood with the prior measure and normalising accordingly.

### 2020-12-01 inverse_problems/C6.tex
almost end of termgit add .!

\section{Bayesian inverse problems}
\subsection{Bayesian inverse problems}
\begin{definition}
	Let $(\Omega', \FF')$ be some measurable space. We denote the space of probability measures on $(\Omega', \FF')$ by $\Prob(\Omega', \FF')$. Furthermore, if $\nu$ is a $\sigma$-finite measure on $(\Omega', \FF')$, we let $\Prob(\Omega', \FF', \nu)$ denote the space of probability measures $\mu \in \Prob(\Omega', \FF')$ with $\mu \ll\nu$. 
\end{definition}

\begin{definition}
	Let $\mu_0 \in \Prob(\XX, \BB\XX)$ and $L \colon (\XX \times \YY, \BB\XX \otimes \BB\YY) \to (\RR, \BB\RR)$ be a measurable function. Then we define the \emph{Bayesian inverse problem with prior $\mu_0$ and likelihood $L$} to be the problem of finding $\mu\post \in \Prob(\XX, \BB\XX)$ with 
	\[
	\dv{\mu\post}{\mu_0}\qty(u) = \frac{L(f_n \mid u)}{\int_\XX L(f_n \mid u) \dd{\mu_0(u)}} \qquad(u \in \XX ; \mu_0\text{-a.s.}), 
	\]
	for any data set $f_n \in \YY$. 
\end{definition}

\begin{definition}
	Consider the BIP as in the previous definition, let $P \subseteq \Prob(\XX, \BB\XX)$ be a space of probability measures and $d$ a metric on $P$. Then the problem is $\emph{$(P, d)$-well-posed}$ if
	\begin{enumerate}
		\item For all $f_n \in \YY$, there exists a solution $\mu\post \in P$ (existence);
		\item For all $f_n \in \YY$, the solution $\mu\post \in P$ is unique (uniqueness);
		\item The map $f_n \mapsto \mu\post$ is continuous from $(\YY, \BB\YY)$ to $(P, d)$ (stability). 
	\end{enumerate}
\end{definition}

If $\int_\XX L(f_n \mid u) \dd\mu_0(u) \in (0, \infty)$, then existence and uniqueness is clear. This is immediate if $L > 0$ and $L(f_n \mid \cdot) \in L^1$ for all $f_n$.

\subsection{Metrics on spaces of probability measures}
We consider two metrics on spaces of probability measures:
\begin{definition}
	Let $(\Omega', \FF')$ be a measurable space.
	\begin{enumerate}
		\item We define the \emph{total variation distance} on $\Prob(\Omega', \FF')$ by
		\[
		d_\Rm{TV}(\mu, \nu)  \ceq \sup_{F' \in \FF'} \abs{\mu(F') - \nu(F')}. 
		\]
		\item Suppose that $\Omega'$ is a topological space and $\FF'=\BB\Omega'$. Let $(\mu_n) \subseteq \Prob(\Omega', \FF')$  and $\mu \in \Prob(\Omega', \FF')$. We say that $\mu_n$ converges \emph{weakly} to $\mu$ as $n\to\infty$, denoted $\mu_n \cw \mu$, if
		\[
		\int_{\Omega'} g \dd{\mu_n} \to \int_{\Omega'} g \dd{\mu} \quad\text{for all continuous bounded functions } g \colon (\Omega', \BB\Omega') \to (\RR, \BB\RR). 
		\]
	\end{enumerate}
\end{definition}

It is known that the weak convergence topology is metrisable, however, the metric itself is hardly used. We will denote this metric $d_\Rm{LP}$.

\begin{lemma}
	Convergence in TV-distance is stronger than weak convergence. 
\end{lemma}

\begin{lemma} \label{lem:tv_distance_formulation}
	Let $\mu, \nu \in \Prob(\Omega, \FF)$, and let $\rho$ be a $\sigma$-finite measure with $\mu, \nu \ll\rho$. Then we have
	\[
	d_\Rm{TV}(\mu, \nu) = \frac12 \int_\Omega \abs{\dv{\mu}{\rho} - \dv{\nu}{\rho}}\dd{\rho}. 
	\]
\end{lemma}
One can always choose $\rho = \mu + \nu$ in the previous lemma.

\subsection{Stability}
We will assume the following, given a Bayesian inverse problem with prior $\mu_0$ and likelihood $L$:
\begin{enumerate} [{A}1]
	\item $L(\cdot \mid u)$ is strictly positive;
	\item $L(f_n \mid \cdot) \in L^1(\mu_0)$;
	\item There exists $h \in L^1(\mu_0)$ such that $L(f_n'\mid \cdot) \leq h(\cdot)$ for all $f_n' \in \YY$;
	\item $L(\cdot \mid u)$ is continuous.
\end{enumerate}
Note that A1 and A2 were already required for existence and uniqueness. A4 asks for continuity only in the data, not in the parameter space.

\begin{theorem}
	Given a BIP with prior $\mu_0$ and likelihood $L$ which satisfies assumptions (A1)-(A4) and let $P = \Prob(\XX, \BB\XX, \mu_0)$. Then the problem is $(P, d_\Rm{TV})$-well-posed, and therefore also $(P, D_\Rm{LP})$-well-posed. 
\end{theorem}

\begin{proof}
	We have already proved existence and uniqueness. By \cref{lem:tv_distance_formulation}, to prove that $f_n \mapsto \mu\post$ is continuous, it suffices to prove for any $f_n^{(m)} \to f_n$ that
	\[
	\int_\XX \abs{\frac{L(f_n \mid u)}{Z(f_n)} - \frac{L(f_n^{(m)} \mid u)}{Z(f_n^{(m)})}} \dd{\mu_0(u)} \to 0.
	\]
	For this, we first show that the map $f_n \mapsto L(f_n \mid \cdot)$ is continuous as a map from $\YY$ to $L^1(\mu_0)$. Note that
	\[
	\lim_{m\to\infty} \int_\XX \abs{L(f_n^{(m)} \mid u) - L(f_n \mid u)} \dd{\mu_0(u)} = 0
	\]
	since we can exchange limit and integral by the dominated convergence theorem (since $L \in L^1$, the integrand is bounded by $2\abs{L}$), and then apply assumption A4 which guarantees continuity of $L(\cdot \mid u)$.

	Analogously, we can show that the map $f_n \mapsto Z(f_n)$ is continuous, and therefore, the map $f_n \mapsto L(f_n \mid \cdot) / Z(f_n)$ is continuous, from which the claim follows. 
\end{proof}

### 2020-12-01 inverse_problems/C7.tex
almost end of termgit add .!

\section{Function space priors and Monte Carlo}
Recall that a random variable in $\RR^d$ is called Gaussian if and only if every linear combination of its components is a Gaussian random variable on $\RR$. This can be generalised to any separable Banach space $(\XX, \BB\XX)$: 
\begin{definition}
	Let $\mu$ be a probability measure on $\Prob(\XX, \BB\XX)$ and let $U \sim \mu$. We call $\mu$ \emph{Gaussian} if, for all $\ell \in \XX^*$, the random variable $\ang{\ell, U}$ has a normal distribution.

	We define the \emph{mean} of $\mu$ by  
	\[
	a_\mu \in \XX^{**} \colon \ell \mapsto \int_\XX  \ang{\ell, u} \dd{\mu}(u),
	\]
	and the \emph{covariance operator} of $\mu$ by $R_\mu \colon (\XX^*)^2 \to \RR$, where 
	\[
	R_\mu(\ell, \ell') = \int_\XX \qty(\ang{\ell, u} - a_\mu(\ell))(\ang{\ell', u} - a_\mu(\ell'))\dd{\mu(u)}
	\]

	If $\XX$ is a function space, $U$ is called a \emph{Gaussian random field}. 
\end{definition}

This definition is not constructive, and in general, it is nontrivial to construct a Gaussian random field. On a separable Hilbert space however, an explicit construction is possible.

\begin{definition}
	Let $\XX$ be a separable Hilbert space and $C \in \KK(\XX, \XX)$ self-adjoint, 
	and write the eigendecomposition
	\[
	Cx = \sum_{i=1}^\infty \lambda_i \ang{x, \phi_i} \phi_i
	\]
	where $\abs{\lambda_1} \geq \abs{\lambda_2} \geq \dotsb$. Then $C$ is called a \emph{trace class} operator if $(\lambda_i)_{i \in \NN} \in \ell^1$. 
\end{definition}

\begin{proposition}
	Let $\XX$ be a separable Hilbert space and $C$ a positive semi-definite trace class operator with an eigenvalue decomposition as in the previous definition. Letting $m \in \XX$ and $\xi \sim \Rm N(0, 1)^{\otimes\NN}$ (a sequence of independent $\Rm N(0,1)$ random variables), we have that
	\[
	U \ceq m + \sum_{i=1}^\infty \sqrt{\lambda_i} \xi_i \phi_i. 
	\]
	is distributed according to a Gaussian measure with mean $m$ and covariance operator $C$ (in the sence that $R_\mu(x, y) = \ang{Cx, y}$). 
\end{proposition}

\begin{proof}
	Let $k \in \NN$ and $U_k \ceq m + \sum_{i=1}^k \sqrt{\lambda_i} \xi_i \phi_i$, and let $x \in \XX$ and $x_i \ceq \ang{x, \phi_i}$. Then we have 
	\begin{align*}
		\ang{x, U_k} &= \ang{x, m} + \sum_{i=1}^k \sqrt{\lambda_i} \ang{x, \phi_i} \xi_i \\
		&= \ang{x, m} + \sum_{i=1}^k \sqrt{\lambda_i} x_i \xi_i, 
	\end{align*}
and since $\sqrt{\lambda_i}x_i\xi_i \sim \Rm N(0, \lambda_i x_i^2)$, it can be shown that $\ang{x, U_k}$ converges weakly to the distribution $\Rm N(\ang{x, m}, \sum_{i=1}^\infty \lambda_i x_i^2)$ if the sum $\sum_{i=1}^\infty \lambda_ix_i^2$ is finite.

Indeed, the sum is finite: $(\lambda_i) \in \ell^1$ by assumption and $(x_i^2) \in \ell^1$ since $\sum_i x_i^2 = \norm{x}^2 < \infty$. It is known that the pointwise product of two $\ell^1$ sequences lies again in $\ell^1$.

This proves that $U$ has a normal distribution, and we cam compute the mean and covariance  using Fubini's theorem, which proves the claim. 
\end{proof}

\begin{example}
	Let $D = [0, 1]^2$ and $\XX \ceq L^2(D, \BB D, \lambda_2)$ and $\ell > 0$, $\sigma^2 \geq 0$. Then the \emph{exponential covariance function} is defined as
	\[
	c_\Rm{exp}(x, y) \ceq \sigma^2\exp(-\norm{x - y}_2 / \ell),
	\]
	while the \emph{Gaussian covariance function} is defined as
	\[
	c_\Rm{N}(x, y) \ceq \sigma^2 \exp(- \norm{x - y}_2 / (2\ell^2)). 
	\]

	The parameter $\ell$ is called \emph{correlation length} while $\sigma^2$ is called \emph{pointwise variance}. 
\end{example}

### 2020-12-01 inverse_problems/inverse_ES3.tex
almost end of termgit add .!

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[margin=3cm]{geometry}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools, amsmath, amssymb, amsthm, mdframed, bbm, graphicx, float, physics, xcolor, cleveref}

\hypersetup{
    colorlinks   = true, %Colours links instead of ugly boxes
    urlcolor     = blue, %Colour for external hyperlinks
    linkcolor    = blue, %Colour of internal links
    citecolor   = red %Colour of citations
}

% Definition of numbered environments.
% Usage: \begin{theorem} ... \end{theorem}
% Remark has no numbering.
\theoremstyle{plain}
\newtheorem{question}{Question}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

% Question with box around it
\newenvironment{qbox}{\begin{mdframed}\begin{question}}{\end{question}\end{mdframed}}

% A proof with "solution" instead of "proof" and no QED symbol
\newenvironment{solution}{\begin{proof}[Solution]\renewcommand\qedsymbol{}}{\end{proof}}

% Some renewed commands
\renewcommand{\vec}{\mathbf}
\renewcommand{\emptyset}{\varnothing}
\renewcommand{\epsilon}{\varepsilon}
\renewcommand{\theta}{\vartheta}
\renewcommand{\phi}{\varphi}

% Frequently used math alphabets
\newcommand{\Bb}{\mathbb}
\newcommand{\Cal}{\mathcal}
\newcommand{\Bf}{\mathbf}
\newcommand{\Rm}{\mathrm}

% Frequently used letters in the blackboard alphabet
\newcommand{\CC}{\Bb C}
\newcommand{\NN}{\Bb N}
\newcommand{\PP}{\Bb P}
\newcommand{\QQ}{\Bb Q}
\newcommand{\RR}{\Bb R}
\newcommand{\EE}{\Bb E}

% Usage: \ang{...} is equivalent to \langle ... \rangle, while \ang*{...} is equivalent to \left\langle ... \right\rangle
% For other delimiters: use \qty from the physics package (i.e., \qty(...))
\DeclarePairedDelimiter{\ang}{\langle}{\rangle}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

% Frequently used commands
\newcommand{\T}{^\top} % Matrix transpose A\T
\newcommand{\C}{^\complement} % Set complement A\C
\renewcommand{\P}{^\perp}
\newcommand{\D}{^\dagger}
\newcommand\ceq\coloneqq % Definitions :=
\newcommand\pow{\Cal P} % Power sets
\newcommand\eps\epsilon
\newcommand\ind{\mathbbm 1} % Blackboard 1 for indicator functions
\newcommand\restr{\mathord\restriction}
\newcommand\TODO{{\color{red} TODO: }}
\newcommand\clos\overline

% Functions that appear frequently
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\Int}{Int}
\DeclareMathOperator{\Span}{Span}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\newcommand\XX{\Cal X}
\newcommand\YY{\Cal Y}
\newcommand\UU{\Cal U}
\newcommand\VV{\Cal V}
\newcommand\LL{\Cal L}
\newcommand\KK{\Cal K}
\newcommand\BB{\Cal B}
\newcommand\JJ{\Cal J}
\newcommand\LLL{\Bb L}
\newcommand\DD{\Cal D}
\newcommand\FF{\Cal F}
\newcommand\Pp{\Cal P}
\DeclareMathOperator\Dom{\Cal D}
\DeclareMathOperator\Nul{\Cal N}
\DeclareMathOperator\Ran{\Cal R}
\DeclareMathOperator\TV{TV}
\DeclareMathOperator\BV{BV}
\DeclareMathOperator\Prob{Prob}
\newcommand\pro{_\Rm{prod}}

\newcommand\cw\rightharpoonup
\newcommand\cws{\overset{*}{\rightharpoonup}}
\newcommand\rext{\overline\RR}
\newcommand\pt\partial
\DeclareMathOperator\dom{dom}

\title{Inverse Problems --- Example Sheet 3}
\author{Lucas Riedstra}
\begin{document}
\maketitle

\begin{question}
	Let $(\Omega, \FF)$ be a measurable space and $\mu, \nu, \rho$ be $\sigma$-finite measures on $(\Omega, \FF)$. Prove the following statements.
	\begin{enumerate}[(a)]
		\item Let $\nu \ll \mu$ and $a \geq 0$. Then, $a \cdot \nu \ll \mu$ and $\dv{a\cdot \nu}{\mu} = a \dv{\nu}{\mu}$ ($\mu$-a.e.). 
		\item Let $\nu \ll \mu$ and $\rho \ll \mu$. Then, $\nu + \rho \ll \mu$ and $\dv{\nu + \rho}{\mu} = \dv{\nu}{\mu} + \dv{\rho}{\mu}$ ($\mu$-a.e.). 
		\item Let $\rho \ll \nu$ and $\nu \ll \mu$. Then, $\rho \ll \mu$ and $\dv{\rho}{\mu} = \dv{\rho}{\nu} \cdot \dv{\nu}{\mu}$ ($\mu$-a.e.). 
	\end{enumerate}
\end{question}

\begin{proof}
	\begin{enumerate}[(a)]
		\item We have, for $F \in \FF$, that $\mu(F) = 0 \implies \nu(F) = 0 \implies (a\cdot \nu)(F) = 0$. Furthermore, we have
		\[
		(a \cdot \nu)(F) = a \cdot (\nu(F)) = a \cdot \int_F \dv{\nu}{\mu} \dd{\mu} = \int_F \qty(a \dv{\nu}{\mu}) \dd{\mu}, 
		\]
		which proves the Radon-Nikodym derivative of $a\cdot\nu$ is $a \dv{\nu}{\mu}$.

		\item We have, for $F \in \FF$, that $\mu(F) = 0 \implies \nu(F) = 0 \text{ and } \rho(F) = 0$ and therefore also $(\nu + \rho)(F) =0$. Furthermore, we have
		\[
		(\nu + \rho)(F) = \nu(F) + \rho(F) = \int_F\dv{\nu}{\mu} \dd{\mu} + \int_F \dv{\rho}{\mu} \dd{\mu} = \int_F \qty(\dv{\nu}{\mu} + \dv{\rho}{\mu}) \dd{\mu}, 
		\]
		which proves the Radon-Nikodym derivative of $\nu + \rho$ is $\dv{\nu}{\mu} + \dv{\rho}{\mu}$.

		\item We have, for $F \in \FF$, that $\mu(F) = 0\implies \nu(F) = 0 \implies \rho(F) = 0$. Furthermore, we have
		\[
		\rho(F) = \int_F \dv{\rho}{\nu} \dd{\nu} \overset\star= \int_F \dv{\rho}{\nu} \dv{\nu}{\mu} \dd{\mu},
		\]
		which proves the Radon-Nikodym derivative of $\rho$ is $\dv{\rho}{\nu} \dv{\nu}{\mu}$. Here, $\star$ follows from the fact that if a measure $\nu$ has $\mu$-density $g$, then $\int_\XX f \dd{\nu} = \int_\XX fg \dd{\mu}$ for all $f \geq 0$ --- this can be proved first for simple functions, and then extended to nonnegative integrable functions. 
	\end{enumerate}
\end{proof}

\begin{question}
	Let $(\Omega, \FF, \PP)$ be a probability space, and $U, U' \colon (\Omega, \FF) \to (\RR, \BB\RR)$, $Y \colon (\Omega, \FF) \to (\YY, \BB\YY)$ random variables. Moreover, let $U$ and $U'$ be integrable. Prove the following statements:
	\begin{enumerate}[(a)]
		\item Let $c \in \RR$ and $\PP(U = c) = 1$. Then, $\EE[U \mid Y = y] = c$  ($\PP(Y \in \cdot)$-a.s.). 
		\item Let $c \in \RR$. Then $\EE[cU \mid Y = y] = c \EE[U \mid Y = y]$ ($\PP(Y \in \cdot)$-a.s.).
		\item $\EE[U + U' \mid Y = y] = \EE[U \mid Y = y] + \EE[U' \mid Y = y]$ ($\PP(Y \in \cdot)$-a.s.). 
	\end{enumerate}
\end{question}

\begin{proof}
	Let $F \in \BB\YY$.

	\begin{enumerate}[(a)]
		\item We have for all $F$
		\[
		%\int_F \EE[U \mid Y = y] \PP(Y \in \dd{y}) = 
		\int_\qty{Y \in F} U \dd{\PP} = \int_\qty{Y \in F} c \dd{\PP} = c \PP(Y \in F) = \int_F c \PP(Y \in \dd{y}).
		\]
		\item We have for all $F$
		\[
		\int_\qty{Y \in F} cU \dd{\PP} = c \int_\qty{Y \in F} U \dd{\PP} = c \int_F \EE[U \mid Y = y] \PP(Y \in \dd{y}) = \int_F c\EE[U \mid Y = y] \PP(Y \in \dd{y}).
		\]

		\item We have for all $F$
		\begin{align*}
		\int_\qty{Y \in F} \qty(U + U') \dd{\PP} &= \int_\qty{Y \in F} U \dd{\PP} + \int_\qty{Y \in F} U' \dd{\PP} \\
		&= \int_F \EE[U \mid Y = y] \PP(Y \in \dd{y}) + \int_F \EE[U' \mid Y = y] \PP(Y \in \dd{y}) \\
		&= \int_F \qty( \EE[U \mid Y = y] + \EE[U' \mid Y = y]) \PP(Y \in \dd{y}).
		\end{align*}
	\end{enumerate}
\end{proof}

\begin{question}
	Let $a \in \RR\setminus \qty{0}$. We consider the inverse problem $au + n = f_n$, where $u \in \RR$ is the unknown parameter, $n \in \RR$ is measurement noise, and $f_n \in \RR$ is observed data. We assume that the noise and prior are Gaussian, $N \sim \Rm N(0, \gamma^2)$ and $U \sim \Rm N(m_0, \sigma_0^2)$, where $\gamma^2 > 0, \sigma_0^2 > 0$. Assume that the likelihood is given by
	\[
	L(f_n \mid u) \ceq \frac1{\sqrt{2\pi\gamma^2}} \exp(- \frac{(au - f_n)^2}{2\gamma}).
	\]
	\begin{enumerate}
		\item[(i)] Compute the posterior measure $\PP(U \in \cdot \mid aU + N = f_n)$. 
	\end{enumerate}
Next, we assume that we take $N$ independent observations of the data, i.e., we consider the likelihood
\[
L\qty(f_n^{(1:M)} \mid u) \ceq \prod_{i=1}^M \frac{1}{\sqrt{2\pi \gamma^2}} \exp(- \frac{(au - f_n^{(i)})^2}{2\gamma^2}), 
\]
where $f_n^{(i)} \in \RR$. 
\begin{enumerate}
	\item[(ii)] Compute the posterior measure $N(m_M, \sigma_M^2) \ceq \PP(U \in \cdot \mid aU + N = f_n^{(i)} (i = 1, \dotsc, M))$. 
	\item[(iii)] Replace the data $f_n^{(1:M)}$ in the posterior by the random vector
	\[
	F \ceq \mqty(a u\D \\ \vdots \\ a u\D) + \eta,
	\]
	where $\eta \sim \Rm N(0, \gamma^2 I)$ for some $u\D \in \RR$ and study the asymptotic behaviour of $\EE[m_M], m_M, \sigma_M^2$ as $M \to\infty$. How do you explain your findings?
\end{enumerate}
\end{question}

\begin{proof}
	\begin{enumerate}[(i)]
		\item 
		%		We will assume the conditions for Bayes' theorem hold, and now we must compute 
		%		\[
		%		\frac{L(f_n \mid U)}{\int_{-\infty}^\infty L(f_n \mid v) \dd{v}}.
		%		\] 
		We have, with $C = (2\pi\gamma^2)^{-1/2}$, 
		\begin{align*}
			\int_{-\infty}^\infty L(f_n \mid v) \dd{v} &=  C \int_{-\infty}^\infty \exp\qty(- \frac{(av - f_n)^2}{2\gamma^2}) \dd{v} = \frac Ca \int_{-\infty}^\infty \exp(- \frac{(v - f_n)^2}{2\gamma^2}) \dd{v} = \frac1a. 
		\end{align*} 
		Thus we have
		\[
		\dv{\mu_\Rm{post}}{\mu_0} \qty(u)= \frac aC \exp(- \frac{(au - f_n)^2}{2\gamma^2}) = \frac aC \exp(- \frac{(u - f_n/a)^2}{2(\gamma / a)^2}),
		\]
		which is the density of an $N(f_n/a, \gamma^2/a^2)$ distribution.

		By question 1c, we have
		\begin{align*}
			\dv{\mu_\Rm{post}}{\lambda} \qty(u)&= \dv{\mu_\Rm{post}}{\mu_0} \qty(u)\dv{\mu_0}{\lambda}\qty(u),
		\end{align*}
		which is (up to a constant) the product of an $N(f_n/a, \gamma^2/a^2)$ density with an $N(m_0, \sigma_0^2)$ density.

		For this, we can use the following lemma:
		\begin{mdframed}
		\begin{lemma}
			Let $f_{\mu_1, \sigma_1}, f_{\mu_2, \sigma_2}$ be the density functions of $\Rm N(\mu_1, \sigma_1^2), \Rm N(\mu_2, \sigma_2^2)$ distributions respectively. Then the product $f_{\mu_1, \sigma_1} f_{\mu_2, \sigma_2}$ is proportional to an $f_{\mu\pro, \sigma\pro}$ density, where
			\[
			\mu\pro \ceq \frac{\mu_1\sigma_2^2 + \mu_2\sigma_1^2}{\sigma_1^2 + \sigma_2^2}, \quad \sigma\pro^2 \ceq \frac{\sigma_1^2\sigma_2^2}{\sigma_1^2 + \sigma_2^2}. 
			\]
		\end{lemma}

	\begin{proof}
		Since we are discussing proportionality, we only care about the exponents. We have
		\begin{align*}
			\frac{(x - \mu_1)^2}{\sigma_1^2} + \frac{(x - \mu_2)^2}{\sigma_2^2} &= \frac{(\sigma_1^2 + \sigma_2^2) x^2 - 2 (\mu_1\sigma_2^2 + \mu_2 \sigma_1^2)x  +  \mu_1^2\sigma_2^2 + \mu_2^2\sigma_1^2}{\sigma_1^2\sigma_2^2} \\
			&= \frac{x^2 - 2 \frac{\mu_1\sigma_2^2 + \mu_2\sigma_1^2}{\sigma_1^2 + \sigma_2^2} x + \frac{\mu_1^2\sigma_2^2 + \mu_2^2\sigma_1^2}{\sigma_1^2 + \sigma_2^2}}{\frac{\sigma_1^2\sigma_2^2}{\sigma_1^2 + \sigma_2^2}} \\
			&= \frac{\qty(x - \frac{\mu_1\sigma_2^2 + \mu_2\sigma_1^2}{\sigma_1^2 + \sigma_2^2})^2}{\frac{\sigma_1^2\sigma_2^2}{\sigma_1^2 + \sigma_2^2}} + C,
		\end{align*}
	where $C$ is some constant independent of $x$. The claim follows. 
	\end{proof}
\end{mdframed}
Plugging in our values we can compute the posterior density: it is an $\Rm N(\mu\pro, \sigma\pro^2)$ density where
\begin{align*}
	\mu\pro &= \frac{\frac{f_n\sigma_0^2}{a} + \frac{m_0\gamma^2}{a^2}}{\frac{\gamma^2}{a^2} + \sigma_0^2}  = \frac{a f_n \sigma_0^2 + m_0\gamma^2}{\gamma^2 + a^2\sigma_0^2}, \\
	\sigma\pro^2 &= \frac{\frac{\sigma_0^2\gamma^2}{a^2}}{\frac{\gamma^2}{a^2} + \sigma_0^2} = \frac{\sigma_0^2 \gamma^2}{\gamma^2 + a^2\sigma_0^2}. 
\end{align*}%		We can compute the product density, and it is proportional to a Gaussian density with parameters
%		\[
%		\mu_c \ceq \frac{1}{\frac{a^2}{\gamma^2} + \frac{1}{\sigma_0^2}} \qty(\frac{a f_n}{\gamma^2} + \frac{m_0}{\sigma_0^2}), \quad \sigma_c^2 \ceq  \frac{1}{\frac{a^2}{\gamma^2} + \frac{1}{\sigma_0^2}} . 
%		\]
%		Therefore, the product density is an $N(\mu_c, \sigma_c^2)$ density, from which the posterior measure follows.

\item We get similar computations as in the previous part, except that we have to compute the product of $N + 1$ densities, namely $\Rm N(f_n^{(1)}/a, \gamma^2/a^2), \dotsc, \Rm N(f_n^{(N)}/a, \gamma^2/a^2), \Rm N(m_0, \sigma_0^2)$. 
Note that in the previous lemma, if $\sigma_1^2 = \sigma_2^2  = \sigma^2$, the formula gives
\[
\mu\pro = \frac12(\mu_1 + \mu_2), \quad \sigma\pro = \frac{\sigma^2}{2}, 
\]
and this generalises: for $N$ observations we get 
\[
\mu\pro^{(n)} = \frac1n (\mu_1 + \dotsb + \mu_n) \eqqcolon \bar\mu, \quad\sigma\pro^{(n)} = \frac{\sigma^2}{n}. 
\]
This shows that the product of the $\Rm N(f_n^{(1)}/a, \gamma^2/a^2), \dotsc, \Rm N(f_n^{(N)}/a, \gamma^2/a^2)$ densities is proportional to a $\Rm N(\bar{f_n}/a, \gamma^2/(na^2))$ distribution, where $\bar f_n$ is the average of $f_n^{(1)}, \dotsc, f_n^{(n)}$.

When we multiply this with prior density $\Rm N(m_0, \sigma_0^2)$, we get
\begin{align*}
	\mu_M &= \frac{\frac{\bar f_n \sigma_0^2}{a} + \frac{m_0\gamma^2}{na^2}}{\frac{\gamma^2}{na^2} + \sigma_0^2} = \frac{n a \bar f_n\sigma_0^2 + m_0\gamma^2}{\gamma^2 + na^2\sigma_0^2}, \\
	\sigma_M^2 &= \frac{\frac{\sigma_0^2\gamma^2}{na^2}}{\frac{\gamma^2}{na^2} + \sigma_0^2} = \frac{\sigma_0^2\gamma^2}{\gamma^2 + na^2\sigma_0^2}. 
\end{align*}

\item Note that $\bar F = a u\D + \bar\eta$, and we know from elementary probability theory that if $\eta \sim N(0, \gamma^2I)$, then for the average $\bar\eta$ we have $\bar \eta \sim N(0, \gamma^2/n)$. 
We get
\begin{align*}
	\EE[m_M]  &= \EE\qty[ \frac{n a (au\D + \bar\eta) \sigma_0^2 + m_0\gamma^2}{\gamma^2 + na^2\sigma_0^2}] = \frac{na^2 u\D \sigma_0^2 + m_0\gamma^2}{\gamma^2 + na^2\sigma_0^2} \overset{n\to\infty}\to  u\D,
\end{align*}
so in the limit $n \to\infty$, we have $\EE[m_M] \to u\D$, which seems reasonable: the more observations we get, the less our prior assumptions are taken into account.

By the law of large numbers, we have
\[
m_M = \frac{na^2\sigma_0^2 u\D + m_0\gamma^2}{n a^2 \sigma_0^2 + \gamma^2} + \frac{na\sigma_0^2 \bar\eta + m_0\gamma^2}{na^2\sigma_0^2\gamma^2} \to u\D, 
\]
since $\bar\eta \to 0$ as $n\to\infty$ by the law of large numbers.

Finally, since $\sigma_M^2$ is independent of the data (it depends only on the likelihood and the prior), we can simply let $n\to\infty$ in our expression for $\sigma_M^2$ and see $\sigma_M^2 \to 0$, which also makes sense: the more observations we get, the less variance we have. 
	\end{enumerate}
\end{proof}

\begin{question}
	Let $(\Omega, \FF)$ and let $\Prob(\Omega, \FF)$ be the space of probability measures on $(\Omega, \FF)$. 
	\begin{enumerate}[(i)]
		\item Show that $d_{\TV} \colon \Prob(\Omega, \FF)^2 \to [0, \infty) \colon (\mu, \nu) \mapsto \sup_{F \in \FF} \abs{\mu(F) - \nu(F)}$ is a metric on $\Prob(\Omega, \FF)$.

		\item Let $\mu, \nu \in \Prob(\Omega, \FF)$ and $\rho$ be a $\sigma$-finite measure with $\mu, \nu \ll\rho$. Show that
		\[
		d_{\TV} (\mu, \nu) = \frac12\int_\Omega \abs{\dv{\mu}{\rho} - \dv{\nu}{\rho}} \dd{\rho}. 
		\]

		\item Let $\KK \ceq \qty{h \colon (\Omega, \FF) \to (\RR, \BB\RR) \colon \sup_{\omega \in \Omega} \abs{h(\omega)} \leq 1}$ and $\mu, \nu \in \Prob(\Omega, \FF)$. Show that
		\[
		d_\Rm{TV}(\mu, \nu) = \sup_{h \in \KK} \frac12 \abs{\int_\Omega h \dd{\mu} - h\dd{\nu}}. 
		\]

		\item Let $\Omega$ be a topological space and $(\Omega, \FF) \ceq (\Omega, \BB\Omega)$. Let $(\mu_n)_{n\in\NN} \in \Prob(\Omega, \FF)^\NN$ and $\mu \in \Prob(\Omega, \FF)$. Show that 
		\[
		\lim_{n\to\infty} d_\Rm{TV}(\mu_n, \mu) = 0 \implies \mu_n \to \mu \text{ weakly, as } n \to\infty. 
		\]

		\item Show that the converse of (iv) is in general not true. 
	\end{enumerate}
\end{question}

\begin{proof}
	\begin{enumerate}[(i)]
		\item We check the metric definition. It is clear that $d_\Rm{TV}$ is nonnegative and symmetric. Furthermore, we have
		\[
		d_\Rm{TV}(\mu, \nu) = 0 \implies \sup_{F\in \FF} \abs{\mu(F) - \nu(F)} = 0 \implies \mu(F) = \nu(F) \text{ for all } F \in \FF \implies \mu = \nu.
		\]
		Finally, if $\mu, \nu, \rho \in \Prob(\Omega, \FF)$, then by the ``normal'' triangle inequality we have
		\begin{align*}
			d_\Rm{TV}(\mu, \rho) &= \sup_{F \in \FF} \abs{\mu(F) - \rho(F)} = \sup_{F \in \FF} \abs{\mu(F) - \nu(F) + \nu(F) - \rho(F)} \\
			&\leq \sup_{F \in \FF} \qty(\abs{\mu(F) - \nu(F)} + \abs{\nu(F) - \rho(F)}) \\
			&\leq \sup_{F \in \FF} \abs{\mu(F) - \nu(F)} + \sup_{F\in \FF} \abs{\nu(F) - \rho(F)} \\
			&= d_\Rm{TV}(\mu, \nu) + d_\Rm{TV}(\nu, \rho). 
		\end{align*}

	\item Write $A = \qty{x \in \Omega \mid \dv{\mu}{\rho}\qty(x) > \dv{\nu}{\rho}\qty(x)}$ and $B = \qty{x \in \Omega \mid \dv{\mu}{\rho}\qty(x) < \dv{\nu}{\rho}\qty(x)}$. 
	Note that for any $X \subseteq A$ we have $\mu(X) \geq \nu(X)$ while for any $X \subseteq B$ we have $\mu(X) \leq \nu(X)$.

	Also note that $\mu(A) - \nu(A) = \nu(B) - \mu(B)$. 
	Therefore, we have
	\begin{align*}
	d_\Rm{TV}(\mu, \nu) &= \sup_{F \in \FF} \abs{\mu(F) - \nu(F)} \\
	&= \sup_{F \in \FF} \abs{\mu(F \cap A) - \nu(F \cap A) + \mu(F \cap B) - \nu(F \cap B)} \\
	&\leq \sup_{F \in \FF} \max\qty{\mu(F \cap A) - \nu(F \cap A), \nu(F \cap B) - \mu(F \cap B)} \\
	&\leq \sup_{F \in \FF} \max\qty{\mu(A) - \nu(A), \nu(B) - \mu(B)} \\
	&= \mu(A) - \nu(A). 
	\end{align*}

	Looking at the integral in the question, we see
	\begin{align*}
		\frac12\int_\Omega \abs{\dv{\mu}{\rho} - \dv{\nu}{\rho}} \dd{\rho} &=  \frac12 \qty(\int_A \qty(\dv{\mu}{\rho} - \dv{\nu}{\rho}) \dd{\rho} - \int_B \qty(\dv{\mu}{\rho} - \dv{\nu}{\rho}) \dd{\rho}) \\
		&= \frac12\qty(\mu(A) - \nu(A) - \mu(B) + \nu(B)) \\
		&= \mu(A) - \nu(A). 
	\end{align*}
We conclude
\[
d_\Rm{TV}(\mu, \nu) = \mu(A) - \nu(A) = \frac12\int_\Omega \abs{\dv{\mu}{\rho} - \dv{\nu}{\rho}} \dd{\rho}.
\]

\item Let $\rho = \mu + \nu$, then we have $\mu \ll\rho$ and $\nu \ll \rho$, and $\rho$ is $(\sigma$-)finite. Define $A$ and $B$ as in the solution to the previous exercise, then we have for $h \in \KK$ that
\begin{align*}
	\frac12 \abs{\int_\Omega h \dd{\mu} - \int_\Omega h \dd{\nu}} &= \frac12 \abs{\int_\Omega h\qty(\dv{\mu}{\rho} - \dv{\nu}{\rho}) \dd{\rho}} \\
	&\leq \frac12 \int_\Omega \abs{\dv{\mu}{\rho} - \dv{\nu}{\rho}}\dd{\rho} \\
	&= \frac12 \int_A \qty(\dv{\mu}{\rho} - \dv{\nu}{\rho}) \dd{\rho} + \frac12 \int_B \qty(\dv{\nu}{\rho} - \dv{\mu}{\rho}) \dd{\rho} \\
	&= \mu(A) - \nu(A) = d_\Rm{TV}(\mu, \nu). 
\end{align*}
Furthermore, equality can be obtained by letting $h = \ind_A - \ind_B \in \KK$, which concludes the proof.

\item Suppose that $d_\Rm{TV}(\mu_n, \mu) \to 0$, and let $g \colon (\Omega, \FF) \to (\RR, \BB\RR)$ be continuous and bounded. Since $g$ is bounded, without loss of generality we can assume $\sup_{\omega \in \Omega} \abs{g(\omega)} \leq 1$ (otherwise we divide by a constant). Now, by the previous exercise we have
\[
d_\Rm{TV}(\mu_n, \mu) \geq \frac12 \abs{\int_\Omega g \dd{\mu} - \int_\Omega g \dd{\mu_n}}, 
\] 
and since $d_\Rm{TV}(\mu_n, \mu) \to 0$, we conclude that $\abs{\int_\Omega g \dd{\mu} - \int_\Omega g \dd{\mu_n}} \to 0$, or equivalently that $\int_\Omega g \dd{\mu_n} \to \int_\Omega g \dd{\mu}$. Since $g$ was arbitrarily chosen, we conclude that $\mu_n \to \mu$ weakly.

\item Let $\mu_n$ be the measure on $(\RR, \BB\RR)$ corresponding to the uniform distribution on $[-\frac1n, \frac1n]$ with density function $f(x) = \frac n2 \cdot  \ind_{[-\frac1n, \frac1n]}$, and let $\mu \ceq \delta_0$. We claim $\mu_n \to \mu$ weakly.

To prove this claim, let $g \colon (\RR, \BB\RR) \to (\RR, \BB\RR)$ be continuous and bounded, and let $\eps > 0$. Choose $n$ large enough such that, on $[-1/n, 1/n]$, $g$ takes values in $[g(0) - \eps, g(0) + \eps]$. Then we have 
\[
\int_\RR g \dd{\mu_n} = \frac n2 \int_{-1/n}^{1/n} g(x) \dd{x} \in [g(0) - \eps, g(0) + \eps], 
\]
and since $\eps$ was randomly chosen, we conclude $\int_\RR g \dd{\mu_n} \to g(0) = \int_\RR g \dd{\delta_0}$.

However, it is immediate that $d_\Rm{TV}(\mu_n, \mu)$ does not converge to 0, since $\mu_n(\qty{0}) = 0$ for all $n$ while $\mu(\qty{0}) = 1$. 
	\end{enumerate}
\end{proof}

\begin{question}
	Let $\mu$ be a probability measure on $(\RR, \BB\RR)$ with cumulative distribution function (cdf) $F \colon \RR \to [0, 1] \colon x \mapsto \mu((-\infty, x])$, for $x \in \RR$. 
	\begin{enumerate}[(i)]
		\item Let $Q \colon (0, 1) \to \RR$ be the quantile function of $\mu$, i.e., $Q(y) \ceq \inf\qty{x \in \RR \mid F(x) \geq y}$. Moreover, let $U \sim \Rm{Unif}(0, 1) \ceq \lambda_1(\cdot \cap (0, 1))$ be a uniformly distributed random variable on the interval $(0, 1)$. Show that $\PP(Q(U) \in \cdot) = \mu$. (Hint: you may use the fact that probability measures on $(\RR, \BB\RR)$ are uniquely determined by their CDF). 
		\item Derive the quantile function for the exponential distribution, i.e., the distribution with cdf $F(x) \ceq 1 - \exp(-\lambda x)$ for some $\lambda > 0$.

		\item Use the idea from (i) and your quantile function from (ii) to generate independent samples (i.e., realisations of random variables) with $\lambda = 1$. 
		\begin{enumerate}[(a)]
			\item Plot the cdf of the exponential distribution along with the empirical cdf of 
			\[
			M \in \qty{10, 100, 1000, 10000}
			\]
			of your samples. (Hint: you can use the \verb|ecdf| command to obtain a representation of your empirical cdf). What do you observe?
			\item Compute the sample mean $\bar X_M$ of your exponentially distributed samples $X_1, \dotsc, X_m$; for $M = \qty{2^n \mid n = 1, \dotsc, 20}$. What do you observe? 
		\end{enumerate}

		\item Repeat (iii)(b) using the quantile function $Q'(y) \ceq \tan(\pi(y - 0.5))$ ($y \in \RR$). What do you observe?

		\item Let $X_1, \dotsc, X_M \sim \mu$ be i.i.d., and assume that $\Var(X_1)$ is finite. Show that
		\[
		\EE\qty[\qty(\EE[X_1] - \bar X_M)^2] = \frac{\Var(X_1)}{M}. 
		\]
		\item Can you recover the rate from (v) in your experiments in (iii)? What could be the issue in (iv)? 
	\end{enumerate}

\end{question}

\begin{proof}
	\begin{enumerate}[(i)]
		\item Since probability measures on $(\RR, \BB\RR)$ are uniquely determined by their CDF, we must show that the CDF of $\mu$ agrees with the CDF of $\PP(Q(U) \in \cdot)$, i.e., for all $x \in \RR$
		\[
		F(x) = \mu((-\infty, x]) = \PP(Q(U) \in (-\infty, x]). 
		\]
		For this, note that
		\begin{align*}
			\PP(Q(U) \in (-\infty, x]) = \PP(Q(U) \leq x) \overset\star= \PP(U \leq F(x)) = F(x), 
		\end{align*}

	\item For the exponential distribution, the cdf $F$ is an invertible function between $(0, \infty)$ and $(0, 1)$, so the quantile function is just the inverse. We have
	\[	
	y = 1 - e^{-\lambda x} \iff e^{-\lambda x} = 1 - y \iff -\lambda x = \ln(1 - y) \iff x = -\frac{\ln(1 - y)}{\lambda}, 
	\]
	so the quantile function is given by $Q(y) = -\frac{\ln(1 - y)}{\lambda}$.

	\item

	\begin{enumerate}[(a)]
		\item From approximately $M = 1000$, the empirical CDF aligns almost exactly with the true cdf.
		\item Keeps getting closer to 1. 
	\end{enumerate}

\item The means are all over the place and don't seem to converge to anything.

\item We have 
\begin{align*}
	\EE\qty[\EE[X_1]^2 - 2\EE[X_1] \bar X_M + \bar{X_M}^2] &= \EE[X_1]^2 - 2\EE[X_1] \EE[\bar X_M] + \EE[\bar X_M^2] \\
	&= \EE[X_1]^2 - 2\EE[X_1]^2 + \frac1{M^2} \sum_{ij} \EE[X_i X_j] \\
	&= - \EE[X_1]^2 + \frac1M \EE[X_1^2] + \frac{M-1}{M} \EE[X_1]^2 \\
	&= \frac{\EE[X_1^2] - \EE[X_1]^2}{M} = \frac1M \Var(X_1).  
\end{align*}

\item In (iv), the issue is that the distribution does not have finite mean or variance. 
	\end{enumerate}

\end{proof}
\end{document}

### 2020-12-01 inverse_problems/inverse_ES3_handin.tex
almost end of termgit add .!

--- 2020-12-01 inverse_problems/inverse_ES3.tex
@@
 
-\begin{question}
-	Let $(\Omega, \FF)$ be a measurable space and $\mu, \nu, \rho$ be $\sigma$-finite measures on $(\Omega, \FF)$. Prove the following statements.
-	\begin{enumerate}[(a)]
-		\item Let $\nu \ll \mu$ and $a \geq 0$. Then, $a \cdot \nu \ll \mu$ and $\dv{a\cdot \nu}{\mu} = a \dv{\nu}{\mu}$ ($\mu$-a.e.). 
-		\item Let $\nu \ll \mu$ and $\rho \ll \mu$. Then, $\nu + \rho \ll \mu$ and $\dv{\nu + \rho}{\mu} = \dv{\nu}{\mu} + \dv{\rho}{\mu}$ ($\mu$-a.e.). 
-		\item Let $\rho \ll \nu$ and $\nu \ll \mu$. Then, $\rho \ll \mu$ and $\dv{\rho}{\mu} = \dv{\rho}{\nu} \cdot \dv{\nu}{\mu}$ ($\mu$-a.e.). 
-	\end{enumerate}
-\end{question}
-
-\begin{proof}
-	\begin{enumerate}[(a)]
-		\item We have, for $F \in \FF$, that $\mu(F) = 0 \implies \nu(F) = 0 \implies (a\cdot \nu)(F) = 0$. Furthermore, we have
-		\[
-		(a \cdot \nu)(F) = a \cdot (\nu(F)) = a \cdot \int_F \dv{\nu}{\mu} \dd{\mu} = \int_F \qty(a \dv{\nu}{\mu}) \dd{\mu}, 
-		\]
-		which proves the Radon-Nikodym derivative of $a\cdot\nu$ is $a \dv{\nu}{\mu}$.
-
-		\item We have, for $F \in \FF$, that $\mu(F) = 0 \implies \nu(F) = 0 \text{ and } \rho(F) = 0$ and therefore also $(\nu + \rho)(F) =0$. Furthermore, we have
-		\[
-		(\nu + \rho)(F) = \nu(F) + \rho(F) = \int_F\dv{\nu}{\mu} \dd{\mu} + \int_F \dv{\rho}{\mu} \dd{\mu} = \int_F \qty(\dv{\nu}{\mu} + \dv{\rho}{\mu}) \dd{\mu}, 
-		\]
-		which proves the Radon-Nikodym derivative of $\nu + \rho$ is $\dv{\nu}{\mu} + \dv{\rho}{\mu}$.
-
-		\item We have, for $F \in \FF$, that $\mu(F) = 0\implies \nu(F) = 0 \implies \rho(F) = 0$. Furthermore, we have
-		\[
-		\rho(F) = \int_F \dv{\rho}{\nu} \dd{\nu} \overset\star= \int_F \dv{\rho}{\nu} \dv{\nu}{\mu} \dd{\mu},
-		\]
-		which proves the Radon-Nikodym derivative of $\rho$ is $\dv{\rho}{\nu} \dv{\nu}{\mu}$.
-Here, $\star$ follows from the fact that if a measure $\nu$ has $\mu$-density $g$, then $\int_\XX f \dd{\nu} = \int_\XX fg \dd{\mu}$ for all $f \geq 0$ --- this can be proved first for simple functions, and then extended to nonnegative integrable functions.
-	\end{enumerate}
-\end{proof}
-
-\begin{question}
-	Let $(\Omega, \FF, \PP)$ be a probability space, and $U, U' \colon (\Omega, \FF) \to (\RR, \BB\RR)$, $Y \colon (\Omega, \FF) \to (\YY, \BB\YY)$ random variables.
-Moreover, let $U$ and $U'$ be integrable.
-Prove the following statements:
-	\begin{enumerate}[(a)]
-		\item Let $c \in \RR$ and $\PP(U = c) = 1$. Then, $\EE[U \mid Y = y] = c$  ($\PP(Y \in \cdot)$-a.s.). 
-		\item Let $c \in \RR$. Then $\EE[cU \mid Y = y] = c \EE[U \mid Y = y]$ ($\PP(Y \in \cdot)$-a.s.).
-		\item $\EE[U + U' \mid Y = y] = \EE[U \mid Y = y] + \EE[U' \mid Y = y]$ ($\PP(Y \in \cdot)$-a.s.). 
-	\end{enumerate}
-\end{question}
-
-\begin{proof}
-	Let $F \in \BB\YY$.
-
-	\begin{enumerate}[(a)]
-		\item We have for all $F$
-		\[
-		%\int_F \EE[U \mid Y = y] \PP(Y \in \dd{y}) = 
-		\int_\qty{Y \in F} U \dd{\PP} = \int_\qty{Y \in F} c \dd{\PP} = c \PP(Y \in F) = \int_F c \PP(Y \in \dd{y}).
-		\]
-		\item We have for all $F$
-		\[
-		\int_\qty{Y \in F} cU \dd{\PP} = c \int_\qty{Y \in F} U \dd{\PP} = c \int_F \EE[U \mid Y = y] \PP(Y \in \dd{y}) = \int_F c\EE[U \mid Y = y] \PP(Y \in \dd{y}).
-		\]
-
-		\item We have for all $F$
-		\begin{align*}
-		\int_\qty{Y \in F} \qty(U + U') \dd{\PP} &= \int_\qty{Y \in F} U \dd{\PP} + \int_\qty{Y \in F} U' \dd{\PP} \\
-		&= \int_F \EE[U \mid Y = y] \PP(Y \in \dd{y}) + \int_F \EE[U' \mid Y = y] \PP(Y \in \dd{y}) \\
-		&= \int_F \qty( \EE[U \mid Y = y] + \EE[U' \mid Y = y]) \PP(Y \in \dd{y}).
-		\end{align*}
-	\end{enumerate}
-\end{proof}
+\setcounter{question}{2}
 
@@
 
-\begin{question}
-	Let $\mu$ be a probability measure on $(\RR, \BB\RR)$ with cumulative distribution function (cdf) $F \colon \RR \to [0, 1] \colon x \mapsto \mu((-\infty, x])$, for $x \in \RR$. 
-	\begin{enumerate}[(i)]
-		\item Let $Q \colon (0, 1) \to \RR$ be the quantile function of $\mu$, i.e., $Q(y) \ceq \inf\qty{x \in \RR \mid F(x) \geq y}$.
-Moreover, let $U \sim \Rm{Unif}(0, 1) \ceq \lambda_1(\cdot \cap (0, 1))$ be a uniformly distributed random variable on the interval $(0, 1)$.
-Show that $\PP(Q(U) \in \cdot) = \mu$.
-(Hint: you may use the fact that probability measures on $(\RR, \BB\RR)$ are uniquely determined by their CDF).
-		\item Derive the quantile function for the exponential distribution, i.e., the distribution with cdf $F(x) \ceq 1 - \exp(-\lambda x)$ for some $\lambda > 0$.
-
-		\item Use the idea from (i) and your quantile function from (ii) to generate independent samples (i.e., realisations of random variables) with $\lambda = 1$. 
-		\begin{enumerate}[(a)]
-			\item Plot the cdf of the exponential distribution along with the empirical cdf of 
-			\[
-			M \in \qty{10, 100, 1000, 10000}
-			\]
-			of your samples. (Hint: you can use the \verb|ecdf| command to obtain a representation of your empirical cdf). What do you observe?
-			\item Compute the sample mean $\bar X_M$ of your exponentially distributed samples $X_1, \dotsc, X_m$; for $M = \qty{2^n \mid n = 1, \dotsc, 20}$. What do you observe? 
-		\end{enumerate}
-
-		\item Repeat (iii)(b) using the quantile function $Q'(y) \ceq \tan(\pi(y - 0.5))$ ($y \in \RR$). What do you observe?
-
-		\item Let $X_1, \dotsc, X_M \sim \mu$ be i.i.d., and assume that $\Var(X_1)$ is finite. Show that
-		\[
-		\EE\qty[\qty(\EE[X_1] - \bar X_M)^2] = \frac{\Var(X_1)}{M}. 
-		\]
-		\item Can you recover the rate from (v) in your experiments in (iii)? What could be the issue in (iv)? 
-	\end{enumerate}
-
-\end{question}
-
-\begin{proof}
-	\begin{enumerate}[(i)]
-		\item Since probability measures on $(\RR, \BB\RR)$ are uniquely determined by their CDF, we must show that the CDF of $\mu$ agrees with the CDF of $\PP(Q(U) \in \cdot)$, i.e., for all $x \in \RR$
-		\[
-		F(x) = \mu((-\infty, x]) = \PP(Q(U) \in (-\infty, x]). 
-		\]
-		For this, note that
-		\begin{align*}
-			\PP(Q(U) \in (-\infty, x]) = \PP(Q(U) \leq x) \overset\star= \PP(U \leq F(x)) = F(x), 
-		\end{align*}
-
-	\item For the exponential distribution, the cdf $F$ is an invertible function between $(0, \infty)$ and $(0, 1)$, so the quantile function is just the inverse. We have
-	\[	
-	y = 1 - e^{-\lambda x} \iff e^{-\lambda x} = 1 - y \iff -\lambda x = \ln(1 - y) \iff x = -\frac{\ln(1 - y)}{\lambda}, 
-	\]
-	so the quantile function is given by $Q(y) = -\frac{\ln(1 - y)}{\lambda}$.
-
-	\item
-
-	\begin{enumerate}[(a)]
-		\item From approximately $M = 1000$, the empirical CDF aligns almost exactly with the true cdf.
-		\item Keeps getting closer to 1. 
-	\end{enumerate}
-
-\item The means are all over the place and don't seem to converge to anything.
-
-\item We have 
-\begin{align*}
-	\EE\qty[\EE[X_1]^2 - 2\EE[X_1] \bar X_M + \bar{X_M}^2] &= \EE[X_1]^2 - 2\EE[X_1] \EE[\bar X_M] + \EE[\bar X_M^2] \\
-	&= \EE[X_1]^2 - 2\EE[X_1]^2 + \frac1{M^2} \sum_{ij} \EE[X_i X_j] \\
-	&= - \EE[X_1]^2 + \frac1M \EE[X_1^2] + \frac{M-1}{M} \EE[X_1]^2 \\
-	&= \frac{\EE[X_1^2] - \EE[X_1]^2}{M} = \frac1M \Var(X_1).  
-\end{align*}
-
-\item In (iv), the issue is that the distribution does not have finite mean or variance. 
-	\end{enumerate}
-
-\end{proof}
 \end{document}

### 2020-12-01 modern_statistical_methods/C2.tex
almost end of termgit add .!

@@
 
-With the previous proposition, we can write characterise solutions to the Lasso: namely we have
+Using the KKT conditions, we can write characterise solutions to the Lasso: namely we have
 \[
@@
 
+From this proposition, it follows that the vector $\hat\nu$ from the KKT conditions is also unique with $\hat\nu = \frac1{\lambda n} X\T (Y - X\hat\beta)$. 
 \subsubsection{Variable selection}

### 2020-12-01 modern_statistical_methods/C4.tex
almost end of termgit add .!

\section{Multiple testing}
Recall that when we test a hypothesis $H$, we either claim it is non-significant or reject it.

We wish to test null hypotheses $H_1, \dotsc, H_m$, of which $m_0$ are true,
and we let $R$ denote the amount of rejected hypotheses. We then define
\begin{table}[H]
	\centering
	\begin{tabular}{c || c | c | c}
		& Claimed non-significant & Rejected & Total\\ \hline \hline
		True hypotheses & $N_{00}$ & $N_{01}$ & $m_0$\\ \hline
		False hypotheses & $N_{10}$ & $N_{11}$ & $m - m_0$ \\ \hline
		Total & $m - R$ & $R$
	\end{tabular}
\end{table}
Note that the $N_{ij}$ are unobserved random variables.

Furthermore, we let $I_0 \subseteq \qty{1, \dotsc, m}$ denote the set of true hypotheses, and we let $p_i$ denote the $p$-value of the individual hypothesis test satisfying
\[
\PP(p_i \leq \alpha) \leq \alpha \qquad\forall \alpha \in [0, 1],
\]
so the chance of rejecting a true null hypothesis at $\alpha$-level is at most $\alpha$.

\begin{definition}
	The \emph{familywise error rate} or FWER is $\FWER \ceq \PP(N_{01} \geq 1)$, the probability that at least one true hypothesis is rejected. 
\end{definition}

A procedure for testing multiple hypotheses with $\FWER \leq\alpha$ has a chance of at most $\alpha$ to reject at least one true hypothesis. Note that this is a very strong condition.

\begin{example}
	The \emph{Bonferroni correction} is defined by rejecting $H_i$ if the $p$-values satisfy $p_i \leq \frac\alpha m$ for each $i$.

	Note that in this case we have by Markov's theorem
	\[
	\PP(N_{01} \geq 1) \leq \EE[N_{01}] = \EE\qty[\sum_{i \in I_0} \ind_{p_i \leq \frac\alpha m}] = \sum_{i \in I_0} \PP(p_i \leq \frac\alpha m) \leq m_0 \frac\alpha m \leq \alpha, 
	\]
	so indeed the Bonferroni correction has $\FWER \leq \alpha$.
\end{example}

However, the Bonferroni correction is too restrictive

\subsection{Closed testing}
For $I \subseteq \qty{1, \dotsc, m}$, let $H_I$ denote the intersection hypothesis $H_I = \cap_{i \in I} H_i$ (``all $H_i$ are true'').  Furthermore, suppose that for each $I \subseteq \qty{1, \dotsc, m}$, we have a \emph{local test} $\phi_I \in \qty{0, 1}$ such that
\[
\PP_{H_I}(\phi_I = 1) \leq \alpha,
\]
where we reject if $\phi_I= 1$.

\begin{definition}
	A \emph{closed testing procedure} rejects $H_I$ if and only if, for all $J \supseteq I$, $H_J$ is rejected by the local test $\phi_J$. 
\end{definition}

\begin{example}
	If we have hypotheses $H_1, H_2, H_3$, and the local test rejects only $H_1, H_2, H_{12}, H_{13}, H_{123}$, then $H_1$ is rejected but $H_2$ is not, since $H_{23}$ is not rejected. 
\end{example}

\begin{theorem}
	A closed testing procedure has $\FWER \leq\alpha$. 
\end{theorem}

\begin{proof}
	If $I_0$ is empty then $N_{01} =0$ so $\FWER = 0$. Assum therefore that $I_0$ is nonempty. Let
	\[
	A \ceq \qty{N_{01} \geq 1}, \qquad B \ceq \qty{\text{reject $H_{I_0}$ with local test}} = \qty{\phi_{I_0} = 1}.
	\]

	By definition of closed testing, in order to reject any $H_i$ for $i \in I_0$, we need to reject $H_{I_0}$, so $A \subseteq B$. Therefore we have
	$\FWER = \PP(A) \leq \PP(\phi_{I_0}= 1) \leq \alpha$. 
\end{proof}

\begin{example}
	\emph{Holm's procedure} uses the Bonferroni correction to build local tests, so we have
	\[
	\phi_I = 1 \iff \min_{i \in I} p_i \leq \frac\alpha{\abs{I}}. 
	\]
\end{example}
It seems hard to compute Holm's procedure, however, in example sheet 4 we show that it is equivalent to the following procedure. First, we order the p-values to get $p_{(1)} \leq \dotsb \leq p_{(m)}$ with corresponding hypotheses $H_{(1)}, \dotsc, H_{(m)}$. Then, we do the following:
\begin{enumerate}
	\item If $p_{(1)} \leq \frac\alpha m$, reject $H_{(1)}$ and go to the next step. Else, accept $H_{(1)}, \dotsc, H_{(m)}$. 
	\item If $p_{(2)} \leq \frac\alpha{m-1}$, reject $H_{(2)}$ and go to the next step. Else, accept $H_{(2)}, \dotsc, H_{(m)}$. 
	\item[$\vdots$]
	\item[$m$.] If $p_{(m)} \leq \alpha$, reject $H_{(m)}$, else, accept $H_{(m)}$.  
\end{enumerate}
In this formulation, we see that Holm's procedure always rejects at least as many hypotheses as Bonferroni correction, so it has higher power: while Bonferroni rejects the hypotheses $H_{(1)}, \dotsc, H_{(j)}$ that satisfy $p_{(i)} \leq \frac\alpha m$, Holm rejects the hypotheses $H_{(1)}, \dotsc, H_{(j)}$ that satisfy $p_{(i)} \leq \frac{\alpha}{m - i + 1}$.

\subsection{The False Discovery Rate}
As noted earlier, controlling the FWER gives very stringent requirements. Instead, one can try to control the \emph{false discovery rate}:
\begin{definition}
	We define the \emph{false discovery proportion} or FDP by 
	\[
	\FDP \ceq \frac{N_{01}}{\max(R, 1)} = \begin{cases}
		\frac{N_{01}}{R} &\text{if $R > 0$}, \\ 0 &\text{else}. 
	\end{cases}
	\]
	The \emph{false discovery rate} or FDR is defined as $\FDR\ceq \EE[\FDP]$. 
\end{definition}

\begin{example}
	The \emph{Benjamini-Hochberg procedure} attempts to control the FDR at level $\alpha$. The procedure is defined as follows: let
	\[
	\hat k \ceq \max \qty{i \mid p_{(i)} \leq \frac{i\alpha}{m}},
	\]
	and reject $H_{(1)}, \dotsc, H_{(\hat k)}$.

	We can visualise this procedure and compare it to the Bonferroni and Holm procedures with the following graph:
	\begin{figure}[H]
		\centering
		\includegraphics[width=.5\paperwidth]{multiple_testing_plot.pdf}
	\end{figure}
The Bonferroni procedure rejects only the hypotheses under the threshold. The Holm procedure rejects all hypotheses up to the first one that goes above the threshold (even if later hypotheses go below the threshold again). The Benjamini-Hochberg procedure takes the last hypothesis that goes below the threshold and rejects all previous hypotheses (even if some of those are above the threshold).

We see that the Benjamini-Hochberg procedure has more power than the Holm procedure (it rejects more hypotheses). However, since this procedure does not control the FWER, there is less control over type 1 errors.

Under some assumptions, the Benjamini-Hochberg procedure does control the FDR: 
\begin{theorem}
	Assume that $\qty{p_i \mid i \in I_0}$ are independent, and also independent of $\qty{p_i \mid i \notin I_0}$. Then the Benjamini-Hochberg procedure satisfies
	\[
	\FDR\leq \frac{m_0}{m}\alpha \leq\alpha. 
	\]
\end{theorem}

\begin{remark}
	The independence assumption is quite restrictive: for example, if we have a linear model $Y = X\beta^0 + \eps$ and our hypotheses are $H_i = \qty{\beta_i^0 = 0}$, then all $p_i$ depend on all responses $Y_i$, so in general they will not be independent. In the example sheet we consider a version of the procedure that doesn't require the independence. 
\end{remark}

\begin{proof}
	We have
	\begin{align*}
		\FDR &= \EE\qty[ \frac{N_{01}}{\max(R, 1)}] = \sum_{r=1}^m \EE\qty[\frac{N_{01}}{r} \ind_{R = r}].
	\end{align*}
	Note that when $R = r$, we have by the definition of the procedure that a hypothesis $H_i$ is rejected if and only if $p_i \leq \frac{\alpha r}{m}$.  We therefor continue
	\begin{align*}
		\FDR &= \sum_{r=1}^m \frac 1r \EE\qty[ \sum_{i \in I_0} \ind_{p_i \leq \frac{\alpha r}{m}} \ind_{R = r}]  = \sum_{r=1}^m \frac1r \sum_{i \in I_0} \PP(R = r, p_i \leq \frac{\alpha r}{m}) \\
		&= \sum_{r=1}^m  \frac1r \sum_{i \in I_0} \PP(p_i \leq \frac{\alpha r}{m}) \PP(R = r \mid p_i \leq \frac{\alpha r}{m}) \\
		&\leq \sum_{r=1}^m \frac1r \frac{\alpha r}{m} \sum_{i \in I_0} \PP(R = r \mid P_i \leq \frac{\alpha r}{m})  \\
		&= \frac\alpha m \sum_{i \in I_0} \qty(\sum_{r=1}^m \PP(R = r \mid p_i \leq \frac{\alpha r}{m})) \overset\star= \frac\alpha m \sum_{i \in I_0} 1 = \frac{m_0}{m}\alpha \leq \alpha, 
	\end{align*}
	which proves the claim except that we have to justify $\star$. For this, fix any $i, r$ and let $p^{-i}$ be the set of $p$-values excluding $p_i$ and $p^{-i}_{(1)}, \dotsc, p^{-i}_{(m-1)}$ its order statistics. Note that if $p_i \leq \frac{\alpha r}{m}$ and $R = r$, then $p_i \in \qty{p_{(1)}, \dotsc, p_{(m)}}$, so we have
	\begin{align*}
		\PP(R = r \mid p_i \leq \frac{\alpha r}{m}) &= \PP(p_{(r)} \leq \frac{\alpha r}{m}, p_{(s)} > \frac{\alpha s}{m} \ \forall s > r \mid p_i \leq \frac{\alpha r}{m}) \\
		&= \PP(p^{-i}_{(r-1)} \leq \frac{\alpha r}{m}, p^{-i}_{(s-1)} > \frac{\alpha s}{r} \ \forall s > r \mid p_i \leq \frac{\alpha r }{m}) \\
		&= \PP(p^{-i}_{(r-1)} \leq \frac{\alpha r}{m}, p^{-i}_{(s-1)} > \frac{\alpha s}{r} \ \forall s > r) &\text{(by independence)} \\
		&= \PP(\max \qty{1 \leq \ell \leq m : p_{(\ell -1 )}^{-i} \leq \frac{\alpha \ell}{m}} = r) \\
		&\eqqcolon \PP(Z = r),
	\end{align*}
	where we define $p_{(0)}^{-i} \ceq 0$. Noting that $Z$ takes values in $\qty{1, \dotsc, m}$, we have
	\[
	\sum_{r=1}^m \PP(R = r \mid p_i \leq \frac{\alpha r}{m}) = \sum_{r = 1}^m \PP(Z =r) = 1,
	\]
	which concludes the proof.
\end{proof}
\end{example}

\subsection{Inference in high-dimensional regression}
Recall that with \emph{inference} we mean the forming of confidence intervals and the testing of hypotheses. We consider our usual model $Y = X\beta^0 + \eps$ where $\eps \sim \Rm N(0, \sigma^2 I)$.

In the classical situation $n \gg p$, it is known that $\hat\beta^\Rm{OLS} - \beta^0 \sim \Rm N(0, \sigma^2 (X\T X)^{-1})$, and therefore we obtain the ellipsoidal confidence set
\[
\qty{\beta : \frac{\norm{\beta - \hat\beta^\Rm{OLS}}_2^2}{p\hat\sigma^2} \leq t}
\]
where $\hat\sigma$ is an unbiased estimator of $\sigma$, and this confidence set contains the true parameter with probability $\PP(Z \leq t)$, where $Z \sim F_{p, n-p}$. We can use this to test hypotheses such as $H: \beta_j^0 = 0$.

However, when $p  \gg n$, even when $\beta^0$ is assumed to be sparse it is much harder to obtain confidence sets, because the distribution of $\hat\beta_\lambda^L - \beta^0$  is difficult to characterise as there is no explicit formula for $\hat\beta_\lambda^L$. 
Specifically, the dependence on $\beta^0$ and $X$ is nontrivial.

One option is to consider the \emph{debiased Lasso estimator}. 
We start with the KKT conditions (writing $\hat S = \qty{k : \hat\beta_k \neq 0}$): 
\[
\frac1n X\T (Y - X\hat\beta) = \lambda\hat\nu, \quad \norm{\hat\nu}_\infty \leq 1, \quad \hat\nu_{\hat S} = \sign(\hat\beta_{\hat S}). 
\]
Write $\hat\Sigma = X\T X / n$, then we can rewrite the conditions as
\begin{align*}
	\frac1n X\T (X(\beta^0 + \eps) - X\hat\beta) &= \lambda\hat\nu \\
	\hat\Sigma(\beta^0 - \hat\beta) + \frac1nX\T\eps &= \lambda\hat\nu \\
	\hat\Sigma(\hat\beta - \beta^0) + \lambda\hat\nu &= \frac1nX\T \eps. 
\end{align*}
Suppose now that $\hat\Theta$ is an approximate inverse of $\hat\Sigma$ (which we will not yet define, but it must be the case that $\hat\Theta\hat\Sigma - I$ is small in some sense). Multiplying by the left with $\hat\Theta$ we obtain
\begin{align*}
	\hat\Theta\hat\Sigma (\hat\beta - \beta^0) + \lambda\hat\Theta \hat\nu &= \frac1n \hat\Theta X\T\eps \\
	\hat\beta - \beta^0 + (\hat\Theta\hat\Sigma - I)(\hat\beta - \beta^0) + \lambda\hat\Theta\hat\nu &= \frac1n \hat\Theta X\T \eps \\
	\hat\beta + \lambda\hat\Theta\hat\nu - \beta^0 &= \frac1n \hat\Theta X\T \eps + (\hat\Theta\hat\Sigma - I)(\beta^0 - \hat\beta) \\
	\hat\beta + \lambda\hat\Theta\hat\nu - \beta^0 &= \frac1n \hat\Theta X\T\eps + \frac1{\sqrt n} \Delta, 
\end{align*}
where $\Delta \ceq \sqrt n(\hat\Theta\hat\Sigma - I)(\beta^0 - \hat\beta)$. Considering this expression, we can define the \emph{debiased Lasso estimator} by
\[
\hat b \ceq \hat\beta + \lambda \hat\Theta \hat\nu  = \hat\beta  + \Theta X\T (Y - X\hat\beta) / n, 
\]
and the above expression becomes
\begin{equation}\label{eq:debiased_lasso_error}
	\hat b - \beta^0 = \frac1n \hat\Theta X\T\eps + \frac1{\sqrt n} \Delta. 
\end{equation}
This allows us to almost understand the distribution of $\hat b - \beta^0$: we know that $\frac1n \hat\Theta X\T\eps$ is normally distributed, while for an appropriate choice of $\hat\Theta$, the vector $\Delta$ can be made small.

For a matrix $A$, let $A_{j, \cdot}$ denote its $j$th row. We then have
\begin{align*}
	\norm{\Delta}_\infty &= \sqrt n\max_j \abs{(\hat\Theta\hat\Sigma - I)_{j, \cdot}\T (\hat\beta - \beta^0)} \\
	&\leq  \sqrt n \qty(\max_j \norm{(\hat \Theta\hat\Sigma - I)_{j, \cdot}}_\infty)\norm{\hat\beta - \beta^0}_1. 
\end{align*}
We already have bounds on $\norm{\hat\beta - \beta^0}_1$ using either \cref{thm:lasso_slow_rate} or \cref{thm:lasso_fast_rate}. Therefore we focus on the rows of $\hat\Theta\hat\Sigma - I$. Letting $\hat\theta_j \ceq \Theta_{j, \cdot}$, we have for any $\eta \geq 0$ that 
\begin{align}
	\norm{(\hat\Theta\hat\Sigma - I)_{j, \cdot}}_\infty \leq \eta &\iff 
		\abs{(\hat\Theta\hat\Sigma)_{j, j} - 1} \leq \eta \text{ and for $i \neq j$: } \abs{(\hat\Theta\hat\Sigma)_{j, i}} \leq \eta \nonumber \\
		&\iff \abs{X_j\T X \hat\theta_j / n - 1} \leq \eta  \text{ and } \frac1n\norm{X_{-j}\T X \hat\theta_j}_\infty \leq \eta \label{eq:debiased_lasso_bound_conds}
\end{align}
where $X_{-j}$ is the matrix $X$ with the $j$th column excluded. Note that the first inequality concerns the diagonal elements of $\hat\Theta\hat\Sigma$ while the second concerns the off-diagonal elements.

We define
\begin{equation} \label{eq:lasso_regress_column}
\hat\gamma^{(j)} \ceq \argmin_{\gamma \in \RR^{p-1}} \qty{\frac1{2n} \norm{X_j - X_{-j}\gamma}_2^2 + \lambda_j \norm{\gamma}_1}, 
\end{equation}
which is the lasso estimator for regressing $X_j$ onto $X_{-j}$ where $\lambda_j$ is some tuning parameter. Furthermore, we define 
\[
\hat\tau_j^2 = \frac1n \norm{X_j - X_{-j}\hat\gamma^{(j)}}_2^2 + \lambda_j \norm{\gamma^{(j)}}_1 \overset{\Rm{ES}}=  X_j\T (X_j - X_{-j}\hat\gamma^{(j)})/n. 
\]
We now define the matrix $\hat\Theta$ by setting its rows has
\[
\hat\theta_j = - \frac{1}{\hat \tau_j^2} (\hat \gamma_1^{(j)}, \dotsc, \hat\gamma_{j-1}^{(j)}, -1, \hat\gamma_j^{(j)}, \dotsc, \hat\gamma_{p-1}^{(j)})\T  \in \RR^p, 
\]
so that
\[
X\hat\theta_j = -\frac{1}{\hat\tau_j^2} \qty(X_{-j}\hat\gamma^{(j)}  - X_j) = \frac{X_j - X_{-j}\gamma^{(j)}}{X_j\T (X_j - X_{-j}\gamma^{(j)}) / n}. 
\]

We now consider the conditions from \cref{eq:debiased_lasso_bound_conds}.  For the first one, note that $X_j\T X \hat\theta_j / n = 1$, so that conditions is satisfied for any $\eta$. For the second one, letting $\tilde\nu^{(j)}$ denote the vector appearing in the KKT conditions for $\hat\gamma^{(j)}$ (\cref{eq:lasso_regress_column}), we have that
\[
\frac1n X_{-j}\T(X \hat\theta_j) = \frac1{n\hat\tau_j^2} X_{-j}\T (X_j - X_{-j}\hat\gamma^{(j)}) = \frac{\lambda_j \tilde \nu^{(j)}}{\hat\tau_j^2} \implies \frac1n \norm{X_{-j}\T X \hat\theta_j}_\infty \leq \frac{\lambda_j}{\hat\tau_j^2}. 
\]
Note that $\lambda_j$ and $\hat\tau_j$ do not depend on the responses $Y$ for our original problem in any way.  Putting everything together, we find
\[
\norm{\Delta}_\infty\leq \sqrt n \norm{\hat\beta - \beta^0}_1 \cdot \max_j \frac{\lambda_j}{\hat\tau_j^2}.
\]

Therefore, we must only understand how we can make $\lambda_j/\hat\tau_j^2$ small. First we will briefly consider its practical use.

\subsubsection{Using the debiased Lasso in practice}
Under the assumption that $\norm{\Delta}_\infty$ is small, we have by \cref{eq:debiased_lasso_error} that
\[
\sqrt n(\hat b_j  - \beta^0_j) \approx  W_j, 
\]
where $W_j = (\frac1n \hat\Theta X\T\eps)_j \sim \Rm N(0, \sigma^2 (\hat\Theta\hat\Sigma\hat\Theta\T)_{jj}) \eqqcolon \Rm N(0, \sigma^2 d_j)$.

Therefore, an \emph{approximate} $(1 - \alpha)$-level confidence interval for $\beta^0_j$ is given by 
\[
\qty[\hat b_j - z_{\alpha/2} \sigma \sqrt{d_j/n} , \hat b_j + z_{\alpha / 2} \sigma \sqrt{{d_j}/{n}}],
\]  
(\TODO Recap how to construct confidence intervals).  The only unknown in the above expression is $\sigma$, which must be estimated.

### 2020-12-01 modern_statistical_methods/msm_ES3.tex
almost end of termgit add .!

@@
 \end{align*}
-where $\star$ follows from the KKT conditions. This shows that $\theta^*$ maximises $G$ over the objective set.  \TODO Uniqueness of $\theta^*$?
+where $\star$ follows from the KKT conditions.
+This shows that $\theta^*$ maximises $G$ over the objective set.
+Finally, $\theta^*$ is the unique maximum since we are maximising a strictly concave function over a convex set.
 
@@
 	It follows that the solution path is piecewise linear with knots whenever the sign of the Lasso solution changes.
-Since the Lasso solution is 0 for $\lambda$ large enough, there can only be finitely many sign changes, which concludes the proof.
-(\TODO elaborate on last step)
+Since there are $3^p$ possible signs and alternating is not possible, there are finitely many knots.
 \end{proof}

### 2020-12-01 topics_statistical_theory/C4.tex
almost end of termgit add .!

@@
 Therefore, we have now reduced the problem of lower-bounding $\Cal M$ to the problem of upper-bounding $\sup_{T \in \hat T} \frac1M \sum_{j=1}^M P_{\theta_j}(T = j)$, which is a testing problem.
-
+We repeat the main result:
+\begin{equation} \label{eq:red_to_testing}
+	\Cal M \geq g(\eta) \qty{1 - \sup_{T \in \hat\TT} \frac1M\sum_{j=1}^M P_{\theta_j} (T = j)}, 
+\end{equation}
+where $\eta = \frac12 \min_{jk} d(\theta_j, \theta_k)$. 
 \subsection{Divergences}
@@
 \end{theorem}
+
+Let $f \colon (0, \infty) \to \RR$ be convex. Then for any $y > 0$, the function $x \mapsto \frac{f(x) - f(y)}{x - y}$ is increasing on $(y, \infty)$ (this is easy to check). Furthermore, we have
+\[
+\lim_{x \to \infty} \frac{f(x) - f(y)}{x - y} = \lim_{x \to \infty} \frac{f(x)}{x - y} - \lim_{x \to \infty} \frac{f(y)}{x - y} = \lim_{x \to \infty} \frac{f(x)}{x}, 
+\]
+so in particular the limit is independent of $y$ and we can define the \emph{maximal slope} of $f$ by
+\[
+M_f \ceq \lim_{x\to\infty} \frac{f(x)}{x} \in (-\infty, \infty].
+\]
+We define $f(0) \ceq \lim_{x \downarrow 0} f(x) \in (-\infty, \infty]$ (a convex function is continuous on an open interval, so this limit exists). In this case, we have
+\[
+f(x + y) = f(x) + y \frac{f(x+y) - f(x)}{y} \leq f(x) + y M_f \qquad\forall x, y \geq 0. 
+\]
+
+\begin{definition}
+	Given a convex function $f \colon (0, \infty) \to \RR$ with $f(1) = 0$, and probability measures $P, Q$ on a measurable space $(\XX, \AA)$, we define the $f$-divergence 
+	\[
+	\Div_f(P, Q) \ceq \int_\XX f\qty(\dv{P_\Rm{ac}}{Q}) \dd{Q} + P_\Rm{sing}(\XX) \cdot M_f. 
+	\]
+\end{definition}
+
+By Jensen's inequality we have
+\begin{align*}
+	\Div_f(P, Q) &=  \int_\XX f\qty(\dv{P_\Rm{ac}}{Q}) \dd{Q} + P_\Rm{sing}(\XX) \cdot M_f \geq f\qty(\int_\XX \dv{P_\Rm{ac}}{Q} \dd{Q}) + P_\Rm{sing}(\XX) \cdot M_f \\
+	&= f(P_\Rm{ac}(\XX)) + P_\Rm{sing}(\XX) \cdot M_f = f(P_\Rm{ac}(\XX) + P_\Rm{sing}(\XX)) = f(P(\XX)) = f(1) = 0, 
+\end{align*}
+so $f$-divergences are nonnegative.
+
+\begin{example}
+	\begin{enumerate}
+		\item If $f(x) = x\log x$, then $M_f = \infty$. If $P \ll Q$ (i.e., $P_\Rm{sing} = 0$), then we have
+		\[
+		\Div_f(P, Q) = \int_\XX  \dv{P}{Q} \log(\dv{P}{Q}) \dd{Q} = \int_\XX \log(\dv{P}{Q}) \dd{P},
+		\]
+		and otherwise $\Div_f(P, Q) = 0$.
+
+		This divergence is known as the \emph{Kullbach-Leibler} divergence from $Q$ to $P$, denoted $\KL(P, Q)$.
+
+		If $P \ll Q$ and $P$ and $Q$ have densities $p$ and $q$ w.r.t.\ a  measure $\mu$, we have $\KL(P, Q) = \int_\XX p \log(\frac pq) \dd{\mu}$.
+
+		\item If $f(x) = x^2 - 1$, then $M_f = \infty$. If $P \ll Q$ we have
+		\[
+		\Div_f(P, Q) = \int_\XX \qty(\dv{P}{Q})^2 \dd{Q} - \int_\XX \dd{Q} = \int_\XX \qty(\dv{P}{Q})^2 \dd{Q} - 1, 
+		\]
+		and otherwise $\Div_f(P, Q) = \infty$.
+
+		This divergence is known as the $\chi^2$ divergence from $Q$ to $P$, denoted $\chi^2(P, Q)$.
+
+		If $P \ll Q$ and  $P$ and $Q$ have densities $p$ and $q$ w.r.t.\ a measure $\mu$, we have $\chi^2(P, Q) = \int_\XX \frac{p^2}{q} \dd{\mu} - 1$.
+
+		\item If $f(x) = (\sqrt x - 1)^2 = x + 1 - 2\sqrt x$ (note that this is convex since $\sqrt x$ is concave), then $M_f = 1$ and therefore
+		\[
+		\Div_f(P, Q) = \int_\XX \qty(\sqrt{\dv{P_\Rm{ac}}{Q}} - 1)^2 \dd{Q} + P_\Rm{sing}(\XX) \eqqcolon \Hh^2(P, Q),
+		\]
+		the \emph{squared Hellinger distance} between $P$ and $Q$.
+
+		If $P$ and $Q$ have densities $p$ and $q$ w.r.t.\ a $\sigma$-finite measure $\mu$, then $H^2(P, Q) = \int_\XX (\sqrt p - \sqrt q)^2 \dd{\mu}$ (example sheet).
+
+		\item If $f(x) = \frac{\abs{x-1}}{2}$, then $M_f = \frac12$ and we have
+		\begin{align*}
+			\Div_f(P, Q) \overset{\text{\TODO}}{=} \sup_{A \in \AA} \abs{P(A) - Q(A)} \eqqcolon \TV(P, Q),
+		\end{align*}
+	the \emph{total variation} divergence between $P$ and $Q$. 
+	\end{enumerate}
+\end{example}
+
+All $f$-divergences are \emph{jointly convex}: for all $\lambda \in [0, 1]$ we have (see Example Sheet)
+\[
+\Div_f\qty((1 - \lambda) P_1 + \lambda P_2, (1 - \lambda) Q_1 + \lambda Q_2) \leq (1-\lambda) \Div_f(P_1, Q_1) + \lambda \Div_f(Q_1, Q_2)
+\]
+
+\subsection{Le Cam's two point lemma}
+Plugging $M = 1$ into \cref{eq:red_to_testing} yields the trivial result $\Cal M \geq 0$.
+Surprisingly, when we plug in $M = 2$, we obtain Le Cam's two point lemma, which can often provide optimal rates for estimating real-valued parameters (though not optimal constants).
+\begin{lemma}
+	In the set-up of section 4.1, we have for any $\theta_1, \theta_2 \in \Theta$ that 
+	\[
+	\Cal M \geq \frac{g(\eta)}{2} \qty{1 - \TV(P_{\theta_1}, P_{\theta_2})}. 
+	\]
+\end{lemma}
+
+\begin{proof}
+	For $T \in \hat\TT_2$, let $A \ceq T^{-1}(\qty{1})$, then we have by \cref{eq:red_to_testing}
+	\begin{align*}
+		\Cal M &\geq g(\eta) \qty{1 - \sup_{T \in \hat T_2} \frac{P_{\theta_1}(T = 1) + P_{\theta_2}(T = 2)}{2}} = \frac{g(\eta)}{2} \qty{1 - \sup_{T \in \hat\TT_2} \qty(P_{\theta_1}(T = 1) - P_{\theta_2}(T = 1))} \\
+		&\geq \frac{g(\eta)}{2} \qty{ 1 - \TV(P_{\theta_1}, P_{\theta_2})}. 
+	\end{align*}
+\end{proof}
+
+\begin{example}
+	Let $X_1, \dotsc, X_n \iid N(\theta, 1)$ for some $\theta \in \RR$.
+Let $\theta_1 = 0$ and $\theta_2 = c n^{-1/2}$ for some $c > 0$, and let $P_{\theta_j} \ceq N(\theta_j, 1)$ for $j = 1, 2$.
+Then we have (where $\star$ denotes equalities that will be proved on the example sheet):
+	\[
+	\TV(P_{\theta_1}^{\times n}, P_{\theta_2}^{\times n}) \overset\star\leq \sqrt{\frac{\KL(P_0^{\times n}, P_{cn^{-1/2}}^{\times n})}{2}} \overset\star= \sqrt{\frac n2 \KL(P_0, P_{cn^{-1/2}})} \overset\star= \frac c2.
+	\]
+	Plugging this into Le Cam's two point lemma yields (using the squared error loss $L(x, y) = (x-y)^2$ that
+	\[
+	\Cal M = \inf_{\hat\theta \in \hat\Theta} \sup_{\theta \in \RR} \EE_\theta \qty[(\hat\theta(X_1, \dotsc, X_n) - \theta)^2] \geq \sup_{c > 0} \frac{c^2}{8n} \qty(1 - \frac c2) = \frac2{27n}.
+		\]
+
+	In this problem, it can be shown that $\Cal M = 1/n$, so Le Cam's two point lemma does give the optimal rate, but it does not give the optimal constant. 
+\end{example}
+
+\begin{example}
+	\TODO Understand remark 45 in lecture notes
+\end{example}
+
+\subsection{Assouad's lemma}
+\begin{lemma}
+	Let $m \in \NN$, $\Phi \ceq \qty{0, 1}^m$ and $\qty{P_\phi \mid \phi \in \Phi}$ a family of probability measures on $(\XX, \AA)$.
+Write $\phi \sim \phi'$ when $\phi$ and $\phi'$ differ in precisely one coordinate and $\phi \sim_j \phi'$ when this coordinate is the $j$-th.
+
+	Suppose the loss function is of the form 
+	\[
+	L(\phi', \phi) = \sum_{j=1}^m L_j(\phi', \phi) = \sum_{j=1}^m g(d_j(\phi', \phi)), 
+	\]
+	where $d_1, \dotsc, d_m$ are semimetrics satisfying $d_j(\phi', \phi) \geq\alpha_j$ whenever $\phi \sim_j \phi'$, and where $g$ is increasing and  satisfies $g(x+y) \leq A\qty{g(x) + g(y)}$ for all $x, y \geq 0$ and some $A > 0$.
+Then
+	\[
+	\inf_{\hat\phi \in \hat\Phi} \max_{\phi\in\Phi} \EE_\phi L(\hat\phi,\phi) \geq \frac1{2A} \qty{1 - \max_{\phi \sim \phi'} \TV(P_\phi, P_{\phi'})} \sum_{j=1}^m g(\alpha_j). 
+	\]
+
+\end{lemma}
+
+\begin{proof}
+	For any $\phi \in \Phi, j \in [m]$, let $\phi^{[j]}$ be the unique element of $\Phi$ with $\phi \sim_j \phi^{[j]}$.
+Letting $\hat\Phi$ denote the set of measurable functions from $\XX$ to $\Phi$, we have
+\begin{equation} \label{eq:minimax}
+		 \max_{\phi \in \Phi} \EE_\phi L(\hat\phi, \phi) \geq \frac1{2^m} \sum_{\phi \in \Phi} \sum_{j=1}^m \EE_\phi L_j(\hat\phi, \phi) = \frac1{2^{m+1}} \sum_{j=1}^m \sum_{\phi \in \Phi} \qty{\EE_\phi L_j(\hat\phi,\phi) + \EE_{\phi^{[j]}} L_j(\hat\phi, \phi^{[j]})},
+	 	\end{equation}
+ 	where the last equality follows from the fact that in the sum we count every element of $\phi$ twice (so we must divide by 2). By the definition of $L_j$ and the triangle inequality,
+ 	\[
+ 	L_j(\hat\phi, \phi) + L_j(\hat \phi, \phi^{[j]}) \geq \frac1A g(d_j(\hat\phi, \phi) + d_j(\hat\phi, \phi^{[j]})) \geq \frac1A g(d_j(\phi, \phi^{[j]})) \geq \frac{g(\alpha_j)}{A}.
+ 	\]
+ 	If we multiply and divide \cref{eq:minimax} by $\sum_{j=1}^m L_j(\hat\phi, \phi) + L_j(\hat\phi, \phi^{[j]})$, we obtain, writing $\FF$ for the set of measurable functions from $\XX\to [0, 1]$, 
+ 	\begin{align*}
+ 	\max_{\phi \in \Phi} \EE_\phi L(\hat\phi, \phi) &\geq \qty(\sum_{j=1}^m L_j(\hat\phi,\phi) + L_j(\hat\phi, \phi^{[j]}))\frac1{2^{m+1}}  \frac{ \sum_{j=1}^m \sum_{\phi \in \Phi} \qty{\EE_\phi L_j(\hat\phi,\phi) + \EE_{\phi^{[j]}} L_j(\hat\phi, \phi^{[j]})}}{\sum_{j=1}^m L_j(\hat\phi, \phi) + L_j(\hat\phi, \phi^{[j]})} \\
+ 	&\overset{??}\geq \frac{\sum_{j=1}^m g(\alpha_j)}{2^{m+1}A} \sum_{\phi \in \Phi} \inf_{f_1, f_2 \in \FF: f_1 + f_2 =1 } \qty{\EE_\phi(f_1) + \EE_{\phi^{[j]}} (f_2)} \\ 
+ 	&= \frac{\sum_{j=1}^m g(\alpha_j)}{2^{m+1}A}  \sum_{\phi \in \Phi}\qty(1 - \sup_{f \in \FF}\qty{ \EE_\phi f - \EE_{\phi^{[j]}} f}) \\ 
+ 	&\overset\star\geq \frac1{2A} \qty{1 - \max_{\phi \sim \phi'} \TV(P_\phi, P_{\phi'})} \sum_{j=1}^m g(\alpha_j), 
+  	\end{align*}
+  \TODO explain $\star$ (related to the alternative expression for $\TV$ as the divergence of $\abs{x-1}/2$).
+  % where $\star$ follows from the fact that $\TV(P_\phi, P_{\phi'})
+\end{proof}
+
+\begin{example}
+	Let $X_1, \dotsc, X_n \iid N_d(\theta, \Sigma) \eqqcolon P_\theta$ for some $\theta \in \RR^d$, where $\Sigma = \Rm{diag}(\sigma_1^2, \dotsc, \sigma_d^2)$.
+Fix $c > 0$, and for $\phi \in \qty{0, 1}^d$ define $\theta^\phi \in \RR^d$ by $\theta_j^\phi = c\sigma_j n^{-1/2} \ind_{\phi_j = 1}$ for $j \in [d]$.
+
+	If we used the squared error loss $L(\theta, \theta') = \norm{\theta - \theta'}_2^2$, then we have $d_j(\theta, \theta') = \abs{\theta_j - \theta'_j}$, so $\alpha_j = c \sigma_j n^{-1/2}$ and $g(x) = x^2$, and  $(x+y)^2 \leq 2(x^2 + y^2)$ implies $A = 2$.
+
+	Write $\hat\Theta$ for the set of measurable functions from $(\RR^d)^{\times n}$ to $\RR^d$, then Assouad's lemma tells us that, for any $\hat\theta(X_1, \dotsc, X_n) = \hat\theta \in \Theta$, 
+	\begin{align*}
+		\sup_{\theta \in \RR^d} \EE_\theta \qty(\norm{\hat\theta - \theta}^2) &\geq \max_{\phi \in \Phi} \EE_{\theta^\phi} \qty(\norm{\hat\theta - \theta^\phi}^2) \\
+		&\geq \frac14 \qty{1 - \max_{\phi \sim \phi'} \TV(P_{\theta^\phi}, P_{\theta^{\phi'}})} \sum_{j=1}^d \frac{c^2\sigma_j^2}{n} \\
+		&\overset\star\geq \frac{c^2}{4n} \qty{1 - \max_{\phi \sim \phi'} \sqrt{\frac{\KL(P_{\theta^\phi}^{\times n}, P_{\theta^{\phi'}}^{\times n})}{2}}} \sum_{j=1}^d \sigma^2 \\
+		&= \frac{c^2}{4n} \qty(1 - \frac c2) \sum_{j=1}^d \sigma_j^2, 
+	\end{align*}
+and taking the supremum over all $c > 0$ on the right-hand side gives $\frac4{27n} \sum_{j=1}^d \sigma_j^2$.  Here, $\star$ follows from Pinsker's inequality (example sheet).
+
+It can be shown that this is the optimal rate in terms of $n$ and the $\sigma_j$, but again, $4/27$ is not the optimal constant.
+
+\subsection{The data processing inequality}
+If $(\XX, \AA, \mu)$ is a measure space and $(\YY, \BB)$ a measurable space, and $g \colon \XX \to \YY$ is measurable, we denote the pushforward measure by $\mu^g \ceq \mu \circ g^{-1}$. 
+\begin{lemma}[Data processing inequality]
+	Let $(\XX, \AA)$ and $(\YY, \BB)$ be measurable spaces, $P, Q$ probability measures on $\XX$, and $g \colon \XX \to \YY$ measurable. Then for any $f$-divergence $\Div_f$ we have
+	\[
+	\Div_f (P^g, Q^g) \leq \Div_f(P, Q). 
+	\]
+\end{lemma}
+
+\begin{proof}
+	We decompose $P = P\ac + P\sing$ w.r.t.\ $Q$. We will show that $(P\ac)^g \ll Q^g$ with Radon-Nikodym derivative $\gamma$, where
+	\[
+	\gamma(y) \ceq \EE_Q \qty( \dv{P\ac}{Q}(X) \mid g(X) = y). 
+	\]
+	To see this, let $B \in \BB$, then
+	\begin{align*}
+		(P\ac)^g(B) &= P\ac(g^{-1}(B)) = \int_\XX \ind_{g^{-1}(B)} \dv{P\ac}{Q} \dd{Q} = \EE_Q\qty[ \EE_Q \qty[ \ind_{g^{-1}(B)}(X) \dv{P\ac}{Q}\qty(X) \mid g(X)]] \\
+		&\overset\star= \EE_Q \qty[ \ind_{g^{-1}(B)}(X) \EE_Q\qty[\dv{P\ac}{Q}\qty(X) \mid g(X)]] \\
+		&= \int_\XX \ind_{g^{-1}(B)} \cdot (\gamma \circ g) \dd{Q} 
+		\overset\star= \int_\YY \ind_B \cdot \gamma \dd{Q^g} = \int_B \gamma \dd{Q^g}, 
+	\end{align*}
+where $\star$ follows from the transformation theorem. This establishes the claim.
+
+Now, writing $(P\sing)^g = ((P\sing)^g)\ac + ((P\sing)^g)\sing$ (the Lebesgue decomposition w.r.t.\ $Q^g$), we have
+\[
+P^g = (P\ac)^g + (P\sing)^g = (P\ac)^g + ((P\sing)^g)\ac + ((P\sing)^g)\sing, 
+\]
+and since $(P\ac)^g \ll Q^g$, this gives the Lebesgue decomposition of $P^g$ w.r.t.\ $Q^g$, namely $(P^g)\ac = (P\ac)^g$ and $(P^g)\sing = ((P\sing)^g)\sing$.
+We now obtain, using $f(x+y) \leq f(x) + M_f$, that
+\begin{align*}
+	\Div_f(P^g, Q^g) &= \int_\YY f\qty(\dv{(P\ac)^g}{Q^g} + \dv{((P\sing)^g)\ac}{Q^g}) \dd{Q^g} + ((P\sing)^g)\sing(\YY) \cdot M_f \\
+	&\leq \int_\YY f\qty(\dv{(P\ac)^g}{Q^g}) \dd{Q^g} + (P\sing)^g(\YY) \cdot M_f \\
+	&= \int_\YY f \circ \gamma \dd{Q^g} + (P\sing)^g(\YY) \cdot M_f \\
+	&= \int_\XX f\circ\gamma\circ g \dd{Q} + P\sing(\XX) \cdot M_f \\
+	&= \EE_Q\qty[ f\qty(\EE_Q\qty[ \dv{P\ac}{Q}\qty(X) \mid g(X)])] + P\sing(\XX) \cdot M_f \\
+	&\leq \EE_Q \qty[ \EE_Q \qty{ f\qty(\dv{P\ac}{Q} \qty(X) \mid g(X))}] + P\sing(\XX) \cdot M_f \\
+	&= \int_\XX f\qty(\dv{P\ac}{Q}) \dd{Q} + P\sing(\XX) \cdot M_f = \Div_f(P, Q), 
+\end{align*}
+where the last inequality follows from the conditional version of Jensen. 
+\end{proof}
+\end{example}

### 2020-12-01 topics_statistical_theory/topics_ES2.tex
almost end of termgit add .!

@@
 \begin{proof}
-	???
+	?? (problem: $f$ being a density does not mean that $f', f'', \dotsc$ are densities). 
 \end{proof}
@@
 	\end{align*}
-where $Z_i(y) = \int_{-\infty}^\infty y K_h(y - Y_i) \dd{y}$. Assuming that $K$ is a second-order kernel, we have
+where $Z_i(y) = \int_{-\infty}^\infty y K_h(y - Y_i) \dd{y}$. We have
 \begin{align*}
@@
 \end{align*}
+(where we use that $K_h$ is symmetric to get rid of $\int y K_h(y) \dd{y}$),
 and so $\hat m(x)$ is simply the Nadaraya-Watson estimator (the local polynomial estimator with $p = 0$). 

### 2020-12-08 inverse_problems/C7.tex
updaaaaaaaaaaates

@@
 \section{Function space priors and Monte Carlo}
+This section covers two different topics: 
+\begin{enumerate}
+	\item The general definition of a Gaussian distribution on an infinite-dimensional space (useful for constructing priors);
+	\item Description of Monte Carlo methods for approximating the posterior. 
+\end{enumerate}
+\subsection{Gaussian measures}
 Recall that a random variable in $\RR^d$ is called Gaussian if and only if every linear combination of its components is a Gaussian random variable on $\RR$.
@@
 \end{example}
+
+\subsection{Monte Carlo methods}
+Generally, Monte Carlo techniques aim at approximating the integral $\bar g \ceq \int_\XX g \dd{\mu}$ by generating independent samples $U_1, U_2, \dotsc \overset{\Rm{iid}}\sim \mu$ and computing the estimator
+\[
+\bar g \approx \hat g_M \ceq \frac1M \sum_{m=1}^M  g(U_m). 
+\]
+By the strong law of large numbers, we know that $\hat g_m \to \bar g$ for $M \to\infty$, $\PP$-almost surely.
+Furthermore, if $\Var_\mu(g)< \infty$, then we have the following expression for the standard deviation:
+\[
+\sqrt{\EE[(\hat g_m - \bar g)^2]} = \sqrt{\Var_\mu(g)} \cdot M^{-1/2},
+\]
+which gives a convergence rate of $M^{-1/2}$.
+
+This rate is not the best possible rate: if we assume that $g$ is smooth in some sense, much better rates can be obtained.
+However, a large advantage is that the rate $M^{-1/2}$ is independent of the dimension of the domain.
+
+The standard Monte Carlo method does not work for Bayesian inverse problems because in general it is not possible to sample independently from the posterior measure.
+We must therefore sample dependently (there are also other options such as sampling independently from a different measure and correct using weights).
+
+\begin{definition}
+	Let $(U_n)_{n=1}^\infty$ be a sequence of $\XX$-valued random variables. Then $(U_n)$ is called a \emph{Markov chain} if, for any $n \in \NN$ and $u_1, \dotsc, u_n \in \XX$, we have
+	\[
+	\PP(U_{n+1} \in \cdot \mid U_1 = u_1, \dotsc, U_n = u_n) = \PP(U_{n+1} \in \cdot \mid U_n = u_n)
+	\]
+	We call the chain \emph{time-homogeneous} if, for all $u \in \XX, n \in \NN$ we have
+	\[
+	 \PP(U_{k+2} \in \cdot \mid U_{k+1} = u) = \PP(U_2 \in \cdot \mid U_1 = u). 
+	\]
+
+	A time-homogeneous Markov chain can be fully represented by a Markov kernel $K \colon \BB\XX \times \XX \to [0, 1]$ where
+	\[
+	K(B \mid u) \ceq \PP(U_2 \in B \mid U_1 = u)
+	\]
+
+	Let $\mu \in \Prob(\XX, \BB\XX)$ be any probability meaure, then we denote the \emph{composition} of $\mu$ and $K$ by
+	\[
+	\mu K(B) \ceq \int_\XX K(B \mid u) \dd{\mu(u)} \qquad(B \in \BB\XX),
+	\]
+	and we call $\mu$ \emph{stationary} w.r.t.\ $K$ if $\mu K = \mu$. Note that $\mu$ is stationary w.r.t.\ $K$ if and only if $U_1 \sim \mu \implies U_2 \sim \mu$.
+
+	Finally, we say that the Markov kernel $K$ satisfies \emph{detailed balance} w.r.t.\ $\mu' \in \Prob(\XX, \BB\XX)$ if
+	\[
+	\int_B K(A \mid u) \dd{\mu'(u)} = \int_A K(B \mid u) \dd{\mu'(u)} \qquad(A, B \in \BB\XX). 
+	\]
+\end{definition}
+
+\begin{lemma}
+	Let $K \colon \BB \XX \times \XX \to [0, 1]$ be a Markov kernel that satisfies detailed balance w.r.t.\ $\mu \in \Prob(\XX, \BB\XX)$. Then $K$ is stationary w.r.t.\ $\mu$. 
+\end{lemma}
+
+\begin{definition}
+	Let $\mu \in \Prob(\XX, \BB\XX, \nu)$ and let $g \colon (\XX, \BB\XX) \to (\RR, \BB\RR)$ be positive with $g = c \dv{\mu}{\nu}$ for some $c > 0$.
+Moreover, let $Q \colon \XX \times \BB\XX \to [0, 1]$ be a Markov kernel, given by a positive function $q \colon (\XX, \BB\XX)^2 \to (\RR, \BB\RR)$ with
+	\[
+	Q(A \mid u) = \int_A q(u' \mid u) \dd{\nu(u')} \qquad(A \in \BB\XX, u \in \XX). 
+	\]
+	Now the \emph{Metropolis-Hastings} Markov kernel is given by
+	\[
+	K_\Rm{MH}(A \mid u) \ceq \delta(A - u) \int_\XX (1-\alpha(u, u'')) Q(\dd{u''} \mid u) + \int_A \alpha(u, u') Q(\dd{u'} \mid u) \qquad(u \in \XX, A \in \BB\XX),
+	\]
+	where $\alpha(u, u') \ceq 1 \lor \frac{g(u') q(u \mid u')}{g(u) q(u' \mid u)}$. 
+\end{definition}
+
+The \emph{Metropolis-Hastings MCMC method} (MCMC = Markov Chain Monte Carlo) is given as follows:
+\begin{enumerate}
+	\item Start with some initial $U_1 \in \XX$ and set $m = 1$;
+	\item Sample $U^* \sim Q(\cdot \mid U_m)$;
+	\item With probability $\alpha(U_m, U^*)$ set $U_{m+1} \leftarrow U^*$, otherwise set $U_{m+1} \leftarrow U_m$. 
+	\item Increment $m \leftarrow m + 1$ and go to step 2. 
+\end{enumerate}

### 2020-12-08 modern_statistical_methods/C4.tex
updaaaaaaaaaaates

@@
 
-Therefore, we must only understand how we can make $\lambda_j/\hat\tau_j^2$ small. First we will briefly consider its practical use.
+Therefore, we must only understand how we can make $\lambda_j/\hat\tau_j^2$ small. First we will briefly consider the practical use of the debiased lasso.
 
@@
 (\TODO Recap how to construct confidence intervals).  The only unknown in the above expression is $\sigma$, which must be estimated.
+
+\subsubsection{Making the error small}
+We will need the following fact about normally distributed random variables:
+\begin{lemma} \label{lem:multivariate_normal_conditional}
+	Let $\Sigma \in \RR^{p \times p}$ be a covariance matrix with $\Omega \ceq \Sigma^{-1}$.
+
+	Let $x_{-j} \sim N(0, \Sigma_{-j, -j})$ and let $\eps \sim N(0, \Omega_{jj}^{-1})$ where $x_{-j}$ and $\eps$ are independent. Then, defining
+	\[
+	x_j \ceq x_{-j}\T \gamma^{(j)} + \eps, \qquad\text{where } \gamma^{(j)} \ceq - \Omega_{jj}^{-1} \Omega_{-j, j},
+	\]
+	we have $x \sim N(0, \Sigma)$. 
+\end{lemma}
+
+\begin{proof}
+	It is clear that $x$ has a multivariate normal distribution, since it is a linear combination of the independent normal variables $x_{-j}$ and $\eps$.
+It is therefore enough to compute its mean and covariance.
+We clearly have
+	\[
+	\EE[x] = \EE[x_{-j}\T \gamma^{(j)}] + \EE[\eps] = 0.
+	\]
+
+	For the variance, let $j = 1$ without loss of generality. 
+%	Then we can write by blockwise matrix inversion
+%	\[
+%	\Sigma = \mqty[ \Sigma_{jj} & \Sigma_{-j, j}\T \\ \Sigma_{-j, j} & \Sigma_{-j, -j}] \implies \Omega = \mqty[ \Omega_{jj}^{-1} + \Omega_{jj}^{-1} \Sigma_{-j, j}\T ]
+%	\]
+%	
+	Then we have
+	\[
+	\Var(x) = \EE\mqty[x_j^2 & x_j x_{-j}\T  \\
+	x_{-j} x_j & x_{-j} x_{-j}\T] \eqqcolon \mqty[ A & B \\ B\T & \Sigma_{-j, -j}]. 
+	\]
+	We can compute 
+	\begin{align*}
+		A &= \EE[x_j^2] = \gamma^{(j)\top} \Sigma_{-j, -j} \gamma^{(j)} + \Omega_{jj}^{-1} = \Omega_{jj}^{-1} \qty(1 + \Omega_{jj}^{-1} \Omega_{-j, j}\T \Sigma_{-j, -j} \Omega_{-j, j}) \overset\star= \Sigma_{jj},
+	\end{align*}
+where $\star$ follows from blockwise matrix inversion.
+
+Analogously, we can compute 
+\begin{align*}
+	B &= \EE[x_j x_{-j}\T] = \EE[\gamma^{(j)\top} x_{-j} x_{-j}\T] + \EE[\eps] \EE[x_{-j}\T] \\
+	&= \gamma^{(j)\top} \Sigma_{-j, -j} = -\Omega_{jj}^{-1} \Omega_{-j, j} \Sigma_{-j, -j} = \Sigma_{j, -j},
+\end{align*}
+where the last step again follows from blockwise matrix inversion.
+This proves that $x \sim N(0, \Sigma)$. 
+\end{proof}
+
+By this lemma, we know that the $j$-th variable can be interpreted as the response of a linear model in the other variables with coefficients $\gamma^{(j)}$.
+Therefore, $\hat\gamma^{(j)}$ defined in \cref{eq:lasso_regress_column} is the Lasso estimator of $\gamma^{(j)}$ with tuning parameter $\lambda_j$.
+We can use Lasso properties to understand properties of $\hat\gamma^{(j)}$, but for this we need $\gamma^{(j)}$ to be sparse.
+We define
+\[
+s \ceq \norm{\beta^0}_0, \quad s_j \ceq \norm{\gamma^{(j)}}_0, \quad s_\Rm{max} \ceq \max(s, s_1, \dotsc, s_p),
+\]
+where $\norm{v}_0$ is the amount of nonzero entries of $v$.
+
+We assume we have a sequence of models with $n \to \infty$. Recall our setting:
+\begin{enumerate}
+	\item We have $Y = X\beta^0 + \eps$, where $\eps \sim N(0, \sigma^2I)$. We will assume $\sigma$ is constant. 
+	\item We assume that the rows of $X$ are i.i.d.\ $N(0, \Sigma)$ variables, where $\Sigma$ is positive definite with inverse $\Omega$. We let $c$ denote the smallest eigenvalue of $\Sigma$. 
+\end{enumerate}
+\begin{theorem}
+	Suppose $c \geq c_\Rm{min} > 0$ and $\max_j \Sigma_{jj} \leq 1$ for all models in the sequence, and suppose $s_\Rm{max} \sqrt{\log(p) / n} \to 0$.
+Then there exist constants $A_1, A_2$ such that setting $\lambda = \lambda_j = A_1 \sqrt{\log(p)/n}$ we have
+\begin{align*}
+	\sqrt n(\hat b - \beta^0) = W + \Delta \qquad\text{where } W \mid X  \sim N(0, \sigma^2 \hat\Theta\Sigma\hat\Theta\T), 
+\end{align*}
+and as $n, p \to\infty$, 
+\[
+\PP(\norm{\Delta}_\infty > A_2 s \log(p)/\sqrt n) \to 0. 
+\]
+\end{theorem}
+
+\begin{proof}
+	We define an event $\Lambda_n$ by the following properties:
+	\begin{enumerate}
+		\item $\phi^2_{\hat\Sigma, s} \geq c_\Rm{min} / 2$ and $\phi^2_{\hat \Sigma_{-j, -j}, s_j} \geq c_\Rm{min} / 2$ for all $j$ (recall the definitions from \cref{thm:lasso_compatibility_condition_asymptotics});
+
+		\item $2\norm{X\T\eps}_\infty / n \leq \lambda$ and $2\norm{X_{-j}\T\eps}_\infty/n \leq \lambda$ for all $j$;
+		\item $\norm{\eps^{(j)}}_2^2/n \geq \Omega_{jj}^{-1} (1 - 4\sqrt{\log(p)/n})$ for all $j$.
+	\end{enumerate}
+On the third example sheet it is shown that $\PP(\Lambda_n) \to 1$ for sufficiently large $A_1$. We work on $\Lambda_n$ and let $c_1, c_2, \dotsc$ be constants.
+
+By the fast rate (\cref{thm:lasso_fast_rate}), we have
+\[
+\norm{\hat\beta - \beta^0}_1 \leq c_1 s \sqrt{\log(p)/n}. 
+\]
+
+Now, fix any $j$, and note that the maximum eigenvalue of $\Omega$ is at most $c_\Rm{min}^{-1}$, so we have
+\[
+c_\Rm{min}^{-1} \geq \max_{v \in S^{p-1}} v\T \Omega v \geq e_j\T \Omega e_j = \Omega_{jj} \implies  \Omega_{jj}\leq c_\Rm{min}^{-1}. 
+\]
+Furthermore, we have by \cref{lem:multivariate_normal_conditional} that 
+\[
+\Omega_{jj}^{-1} = \Var(X_{ij} \mid X_{i, -j}) \overset?\leq \Var(X_{ij}) \leq \max_j \Sigma_{jj} \leq 1.
+\]
+
+Applying the fast rate (\cref{thm:lasso_fast_rate}) to the lasso estimator $\hat\gamma^{(j)}$, we have
+\begin{align*}
+	\norm{\gamma^{(j)} - \hat\gamma^{(j)}}_1 \leq c_2 s_j \Omega_{jj}^{-1} \sqrt{\log(p)/n} \cdot \frac{1}{\phi^2_{\hat\Sigma_{-j, -j}}},
+\end{align*}
+but since we are working on $\Lambda_n$ we can upper bound the compatibility constant and bound $\Omega_{jj}^{-1} \leq 1$ to obtain
+\[
+\norm{\gamma^{(j)} -\hat\gamma^{(j)}}_1 \leq c_3 s_j \sqrt{\log(p)/n}.
+\]
+
+Now we can lower bound the factor $\hat\tau_j^2$: by definition we have
+\begin{align*}
+	\hat \tau_j^2 &\geq \frac1n \norm{X_j - X_{-j} \hat\gamma^{(j)}}_2^2 \geq \frac1n \norm{e^{(j)}}_2^2 - \frac2n \norm{X_{-j}\T \eps^{(j)}}_\infty \norm{\gamma^{(j)} - \hat\gamma^{(j)}}_1 \\
+	&\overset\star\geq \Omega_{jj}^{-1} (1 - 4 \sqrt{\log(p)/n}) - c_5 \frac{s_\Rm{max} \log(p)}{n} \\
+	&\geq c_\Rm{min}(1 - 4 \sqrt{\log(p)/n}) - c_5 \frac{s_\Rm{max} \log p}{n}, 
+\end{align*}
+and the right hand side goes to 0 for $n \to\infty$ by assumption (here $\star$ follows from the conditions on $\Lambda_n$), so for $n$ large enough we have $\hat\tau_j^2 \geq \frac{c_\Rm{min}}{2}$.
+
+Putting everything together, on $\Lambda_n$ we have
+\begin{align*}
+	\norm{\Delta}_\infty &\leq \lambda \sqrt n \norm{\hat\beta- \beta^0}_1 \max_j \hat\tau_j^{-2} \\
+	&\leq \frac{2}{c_\Rm{min}} \cdot A_1 \sqrt{\log p} \cdot c_1 s \sqrt{\log(p)/n} \\
+	&\leq A_2 s \log(p)/\sqrt n, 
+\end{align*}
+and therefore  $\PP(\norm{\Delta}_\infty > A_2 s \log(p)/\sqrt n) \leq \PP(\Lambda_n\C) \to 0$, which proves the claim.
+\end{proof}

### 2020-12-08 topics_statistical_theory/C1.tex
updaaaaaaaaaaates

@@
 
-Note that the strong law of large numbers immediately implies that $\hat{F_n}(x)$ converges almost surely to $F(x)$ as $n \to \infty$.
+Note that the strong law of large numbers immediately implies that $\hat{F_n}(x)$ converges almost surely to $F(x)$ as $n \to \infty$ for all fixed $x \in \RR$.
 However, the following stronger result states that this convergence holds uniformly in $x$:
@@
 
-\begin{proof}
-    See lecture notes.
 The main idea of the proof is to ``control'' $\hat{F_n}$ in a finite number of points $x_1, \dotsc, x_k$, and then deduce what happens between those points using the fact that distributions are increasing and right-continuous.
 On \href{https://en.wikipedia.org/wiki/Glivenko%E2%80%93Cantelli_theorem#Proof}{Wikipedia},
-        a simplified proof can be found assuming that $F$ is continuous, which still encapsulates the main idea.  
+	a simplified proof can be found assuming that $F$ is continuous, which still encapsulates the main idea.  For the general proof, we need the following fact about quantile functions:
+	\begin{recap}
+		For any distribution function $F$, its \emph{quantile function} is defined as
+		\[
+		F^{-1} \colon (0, 1] \to \RR \cup \qty{\infty} \colon p \mapsto \inf \qty{x \in \RR \mid F(x) \geq p}. 
+		\]
+		Note that since $F$ is right-continuous and non-decreasing, the infimum is well-defined and may be replaced by a minimum, and therefore we always have $F(F^{-1}(p)) \geq p$.
+
+		When necessary, we also define $F^{-1}(0) \ceq \sup\qty{x \in \RR \mid F(x) = 0}$.
+
+	\end{recap}
+
+\begin{proof}
+    Let $\eps > 0$ and choose $k$ such that $\frac1k \leq \eps$. Now set $x_0 \ceq -\infty$ and $x_i \ceq F^{-1}(\frac ik)$. 
+   Then we have 
+    \[
+    F(x_i-) - F(x_{i-1}) \leq \frac ik - \frac{i-1}{k} = \frac1k \leq \eps
+    \]
+    for all $i$. Define $X = \qty{x_1, \dotsc, x_k, x_1-, \dotsc, x_k-}$ (we abuse notation here) and 
+    \[
+    \Omega_{n, \eps} \ceq \qty{\max_{x \in X} \sup_{m \geq n} \abs{\hat F_m(x) - F(x)} \leq \eps}. 
+    \]
+    By a union bound and the strong law of large numbers we have
+    \[
+    \PP_F(\Omega_{n, \eps}\C) \leq \sum_{x \in X} \PP(\sup_{m \geq n} \abs{\hat F_m(x) - F(x)} > \eps) \to 0, 
+    \]
+    since $\hat F_m(x)$ and $\hat F_m(x-)$ are the sample averages of the random variables $\ind_{X \leq x}$ and $\ind_{X < x}$ and therefore converge almost surely to their means $F(x)$ and $F(x-)$.
+
+    Now, fixing $x \in [x_{i-1}, x_i)$ we have for any $n \in \NN$, $m \geq n$
+    \begin{align*}
+    \hat F_m(x) - F(x) &\leq \hat F_m(x_i-) - F(x_{i-1}) \leq \hat F_m(x_i-) - F(x_i-) + F(x_i-) - F(x_{i-1}) \\
+    &\leq \max_{x \in X} \sup_{m \geq n} \abs{\hat F_m(x) - F(x)} + \eps, 
+        \end{align*}
+    and analogously $F(x) - \hat F_n(x) \leq \max_{x \in X}\sup_{m \geq n} \abs{\hat F_m(x) - F(x)} + \eps$.
+
+    Therefore, we have 
+    \begin{align*}
+    	\PP_F\qty(\sup_{m \geq n} \sup_{x \in \RR} \abs{\hat F_m(x) - F(x)} > 2\eps) \leq \PP_F \qty(\max_{x \in X} \sup_{m \geq n} \abs{\hat F_m(x) - F(x)} > \eps) = \PP(\Omega_{n, \eps}\C) \to 0. 
+    \end{align*}
+
+	Noting that $\eps$ was arbitrary, we conclude
+	\begin{align*}
+		\PP_F\qty(\sup_{x \in \RR} \abs{\hat F_n(x) - F(x)} \to 0) &= \PP_F\qty(\forall L \in \NN \ \exists n \in \NN \ \forall m \geq n : \sup_{x \in \RR} \abs{\hat F_n(x) - F(x)} \leq \frac1L) \\
+		&= \PP_F \qty(\bigcap_{L=1}^\infty \bigcup_{n = 1}^\infty  \bigcap_{m = n}^\infty  \qty{\sup_{x \in \RR} \abs{\hat F_n(x) - F(x)} \leq \frac1L}) \\
+		&= \lim_{L \to\infty} \lim_{n \to\infty} \PP_F \qty(\sup_{m \geq n} \sup_{x \in \RR} \abs{\hat F_m(x) - F(x)} \leq \frac 1L) = 1. 
+	\end{align*}
 \end{proof}
@@
 
+\begin{consequence}
 For another consequence, we consider the problem of finding a confidence band for $F$.
@@
 \]
+\end{consequence}
 
-We can say even more.
-
-\begin{recap}
-    For any distribution function $F$, its \emph{quantile function} is defined as
-    \[
-    F^{-1} \colon (0, 1] \to \RR \cup \qty{\infty} \colon p \mapsto \inf \qty{x \in \RR \mid F(x) \geq p}. 
-    \]
-    When necessary, we also define $F^{-1}(0) \ceq \sup\qty{x \in \RR \mid F(x) = 0}$.
-
-    If $U \sim U(0, 1)$ and $X \sim F$, then for any $x \in \RR$ we have
-    \[
-    \PP(F^{-1}(U) \leq x) = \PP(U \leq F(x)) = F(x) = \PP(X \leq x). 
-    \]
-    This can be written simply as $F^{-1}(U) \deq X$. 
-\end{recap}
-
-Let $U_1, \dotsc, U_n \iid U(0, 1)$ with empirical distribution function $\hat{G_n}$, and let $X_1, \dotsc, X_n \iid F$. Then, we have
+\begin{discussion}
+Let $U_1, \dotsc, U_n \iid U(0, 1)$ with empirical distribution function $\hat{G}_n$, and let $X_1, \dotsc, X_n \iid F$. Then, we have
 \begin{align*}
-\hat{G_n}(F(x)) = \frac1n \sum_{i=1}^n \ind_\qty{U_i \leq F(x)} \deq \frac1n \sum_{i=1}^n \ind_\qty{X \leq x} = \hat{F_n}(x), 
+	\hat{G_n}(F(x)) = \frac1n \sum_{i=1}^n \ind_\qty{U_i \leq F(x)} \deq \frac1n \sum_{i=1}^n \ind_\qty{X_i \leq x} = \hat{F_n}(x), 
 \end{align*}
@@
 \]
-with equality if $F$ is continuous. We conclude that if $F$ is continuous, the distribution of $\sup_{x \in \RR} \abs{\hat{F_n}(x) - F(x)}$ does not depend on $F$.
+with equality if $F$ is continuous.
+We conclude that if $F$ is continuous, the distribution of $\sup_{x \in \RR} \abs{\hat{F_n}(x) - F(x)}$ does not depend on $F$, and that continuous functions give a ``worst-case'' scenario for $\sup_{x \in \RR} \abs{\hat F_n(x) - F(x)}$.
+\end{discussion}
 
+\begin{discussion}
 	Other generalisations of \cref{thm:GC} include Uniform Laws of Large Numbers.
@@
 	In \cref{thm:GC}, we showed that $\Cal  G = \qty{ \ind_\qty{\cdot \leq x} \mid x \in \RR}$ satisfies a ULLN. 
+\end{discussion}
 
@@
 \end{recap}
+\begin{discussion}
 Another extension starts with the observation that $\sqrt n \qty(\hat{F_n}(x) - F(x)) \todist N(0, \sigma^2)$, where
@@
 \]
-This can be strengthened by considering $(\sqrt n \qty(\hat{F_n}(x) - F(x)) \todist N(0, \sigma^2) \mid x \in \RR)$ as a stochastic process.
+This can be strengthened by considering $\qty(\sqrt n (\hat{F_n}(x) - F(x)) : x \in \RR)$ as a stochastic process. 
+\end{discussion}
 
 \subsection{Order statistics and quantiles}
+\begin{recap}
+
+	If $U \sim U(0, 1)$ and $X \sim F$, then for any $x \in \RR$ we have
+	\[
+	\PP(F^{-1}(U) \leq x) = \PP(U \leq F(x)) = F(x) = \PP(X \leq x). 
+	\]
+	This can be written simply as $F^{-1}(U) \deq X$. 
+\end{recap}
 \begin{definition}
@@
 \end{definition}
+Note that the sample quantile function is the quantile function of the empirical distribution function.
 
@@
 
-For $p = \frac12$ for example, this proposition tells us that $\hat F_n^{-1}(p) = X_{(\ceil{n/2})}$, the median of the data. We now explore the distribution of $X_{(\ceil{np})}$.
+For example, this proposition tells us that $\hat F_n^{-1}(1/2) = X_{(\ceil{n/2})}$, the median of the data. We now explore the distribution of $X_{(\ceil{np})}$.
 
@@
     Now, by the central limit theorem, the term between brackets converges to a standard $N(0, 1)$ distribution.
-The term $\sqrt{\ceil{np}}{\sqrt n}$ converges to $\sqrt p$ and the term $(\ceil{np} - np)/\sqrt n$ converges to 0, so by Slutsky's lemma, we find
+The term $\sqrt{\ceil{np}}/{\sqrt n}$ converges to $\sqrt p$ and the term $(\ceil{np} - np)/\sqrt n$ converges to 0, so by Slutsky's lemma, we find
     \[
@@
 
-    An analogous calculation shows that $\sqrt n \qty(\frac{W_n}{n} - (1-p)) \to N(0, 1 - p)$.
+    Letting $q \ceq 1 - p$, 
+    an analogous calculation shows that $\sqrt n \qty(\frac{W_n}{n} - q) \to N(0, q)$.
 
-    Now we define $g \colon (0, \infty)^2 \to (0, \infty)$ by $g(x, y) \ceq x / (x + y)$, which is differentiable at $(p, 1 - p)$.
-Note that the distribution of $(V_n, W_n)$ is an $N(0, \smqty(p & 0 \\ 0 & q))$ distribution.
-By the delta method we find
+    Now we define $g \colon (0, \infty)^2 \to (0, \infty)$ by $g(x, y) \ceq x / (x + y)$, which is differentiable at $(p, q)$ with derivative
+    \[
+    \nabla g(x, y) = \mqty[y / (x+y)^2 \\ -x / (x+y)^2] \implies \nabla g(p, q) = \mqty[ q \\ -p]. 
+    \]Note that the distribution of $(V_n, W_n)$ is an $N(0, \smqty(p & 0 \\ 0 & q))$ distribution. By the delta method we find
     \begin{align*}
         \sqrt n \qty(U_{\ceil{np}} - p) &\deq \sqrt n\qty( g\qty(\frac{V_n}{n}, \frac{W_n}{n}) - g(p, q)) \\
-        &\todist g'(p, 1-p) N\qty(0, \mqty(p & 0 \\ 0 & q)) \\
-        &= N\qty(0, g'(p, 1-p) \mqty(p & 0 \\ 0 & q) g'(p, 1-p)\T) \\
-        &= N(0, p(1-p)).
+        &\todist \nabla g(p, q)\T \cdot N\qty(0, \mqty(p & 0 \\ 0 & q)) \\
+        &\deq  N\qty(0, \nabla g(p, q)\T  \mqty(p & 0 \\ 0 & q) \nabla g(p, q)) \\
+        &\deq  N(0, pq),
     \end{align*}
+since $\nabla g(p, q)\T \mqty(p & 0 \\ 0 & q) \nabla g(p, q) = q^2 p + p^2 q = pq(p+q) = pq$.
 \end{proof}
@@
 \end{definition}
+\begin{remark}
 Note that equality holds when $X \sim N(0, \sigma^2)$, since the MGF of an $N(\mu, \sigma^2)$ distribution is given by $t \mapsto \exp(\mu t + \sigma^2 t^2 /2)$. 
+	content...
+\end{remark}
 
@@
 
+\begin{discussion}[Chernoff bounding]
+	Let $X$ be any random variable, then by Markov's inequality, then we have for all $t > 0$ that 
+	\[
+	\PP(X \geq x) = \PP(e^{tX} \geq e^{tx}) \leq e^{-tx} \EE[e^{tX}] = e^{-tx} M_X(t).
+	\]
+	Since the left-hand side is independent of $t$, we can minimise over all $t$ and obtain \[
+	\PP(X \geq x) \leq \inf_{t > 0} e^{-tx} M_X(t), 
+	\]
+	which often gives better results than the ``standard'' Markov bound. This technique is called \emph{Chernoff bounding}, and is very useful if bounds on $M_X(t)$ are known. 
+\end{discussion}
 \begin{proposition}We consider some characterisations of sub-Gaussianity: 
@@
         \item We first consider $\PP(X \geq x)$. 
-        By Markov's inequality, we have for all $t \in \RR$ that
-        \[
-        \PP(X \geq x) = \PP(e^{tX} \geq e^{tx}) \leq e^{-tX} \EE(e^{tX}) \leq e^{-tx + t^2\sigma^2/2}.  
-        \]
-        Since the LHS is independent of $t$, we can take the infimum over $t$ on the RHS and obtain
+        By a Chernoff bound, we have
         \[
@@
         \]
-        since the infimum of $t^2\sigma^2/2 - tx$ is attained at $t = x/\sigma^2$ (this method is called \emph{Chernoff bounding}).
+        since the infimum of $t^2\sigma^2/2 - tx$ is attained at $t = x/\sigma^2$.
 
@@
 \begin{align*}
-	\EE[e^{tX}] &\leq \EE[e^{-tX'}] \EE[e^{tX}] = \EE[e^{t(X - X')}] = \EE \sum_{q=0}^\infty \qty[\frac{t^{2q} (X - X')^{2q}}{(2q)!}] \\
-	&= \sum_{q=0}^\infty \frac{t^{2q} \EE[(X - X')^{2q}]}{(2q)!} \leq \sum_{q=0}^\infty \frac{2^{2q-1} t^{2q} \qty(\EE[X^{2q}] + \EE[(X')^{2q}])}{(2q)!} \\
+	\EE[e^{tX}] &\leq \EE[e^{-tX'}] \EE[e^{tX}] = \EE[e^{t(X - X')}]= \EE \sum_{q=0}^\infty \qty[\frac{t^q (X - X')^{q}}{q!}] 
+	\\
+	&\overset\star= \sum_{q=0}^\infty \frac{t^{2q} \EE[(X - X')^{2q}]}{(2q)!} \leq \sum_{q=0}^\infty \frac{2^{2q-1} t^{2q} \qty(\EE[X^{2q}] + \EE[(X')^{2q}])}{(2q)!} \\
 	&\leq \sum_{q=0}^\infty  \frac{2^{2q - 1}t^{2q} 2 q! C^{2q}}{(2q)!} =\sum_{q=0}^\infty  \frac{(2tC)^{2q}q!}{(2q)!} = \sum_{q=0}^\infty \frac{(2tC)^{2q}}{\prod_{j=1}^q (q+j)} \\
@@
 \end{align*}
+Here, $\star$ follows from Fubini's theorem and the fact that the odd moments of $X - X'$ vanish. 
 This shows that $X$ is sub-Gaussian with parameter $4C^2$. 
@@
 	Let $X_1, \dotsc, X_n$ be independent sub-Gaussian random variables, with $X_i$ having parameter $\sigma_i^2$.
-Then $\bar X$ is sub-Gaussian with parameter $\overline{\sigma^2}$.
+Then $\bar X$ is sub-Gaussian with parameter $\overline{\sigma^2} = \frac1n \sum_{i=1}^n \sigma_i^2$.
 In particular, we have
@@
 
+\begin{example}
 	Note that when $X$ takes values in $[a, b]$, its variance is at most $(b-a)^2$.
 However, when $\Var(X_i) \ll (b_i - a_i)^2$, Hoeffding's inequality can be loose (for example, when $X_i \sim \Rm{Bern}(p_i)$ with $p_i$ small).
-In such circumstances, Bennett's or Bernstein's inequality may give better results.
+In such circumstances, Bennett's or Bernstein's inequality may give better results (example sheet 1, question 4).
+\end{example}
 
@@
 	Let $X_1, \dotsc, X_n$ be independent random variables with $\EE[X_i] = 0$, $\sigma_i^2 \ceq \Var(X_i) < \infty$, and $X_i \leq b$ for some $b > 0$.
-Define  $S \ceq \sum_{i=1}^n X_i$, $\nu \ceq \overline{\Var(X_i)}$ and $\phi \colon \RR \to \RR$ by $\phi(u) \ceq e^u - 1 - u = \sum_{k=2}^\infty \frac{u^k}{k!}$, then for every $t > 0$ we have
+Define  $S \ceq \sum_{i=1}^n X_i$, $\nu \ceq \overline{\sigma^2}$ and $\phi \colon \RR \to \RR$ by $\phi(u) \ceq e^u - 1 - u = \sum_{k=2}^\infty \frac{u^k}{k!}$, then for every $t > 0$ we have
 	\[
@@
 	\[
-	e^{tX_i} \leq 1 + tX_i + X_i^2 \frac{\phi(bt)}{b^2} \implies \EE[e^{tX_i}] \leq 1 + \EE[X_i^2] \frac{\phi(bt)}{t^2} = 1 + \Var(X_i) \frac{\phi(bt)}{b^2}. 
+	e^{tX_i} \leq 1 + tX_i + X_i^2 \frac{\phi(bt)}{b^2} \implies \EE[e^{tX_i}] \leq 1 + \EE[X_i^2] \frac{\phi(bt)}{t^2} = 1 + \sigma_i^2 \frac{\phi(bt)}{b^2}. 
 	\]
@@
 	\begin{align*}
-		\log \EE[e^{tS}] &= \sum_{i=1}^n \log\EE[e^{tX_i}] \leq n \cdot \frac1n\sum_{i=1}^n \log\qty(1 + \Var(X_i) \frac{\phi(bt)}{b^2}) \\
+		\log \EE[e^{tS}] &= \sum_{i=1}^n \log\EE[e^{tX_i}] \leq n \cdot \frac1n\sum_{i=1}^n \log\qty(1 + \sigma_i^2 \frac{\phi(bt)}{b^2}) \\
 		&\overset*\leq n \log(1 + \frac{\nu \phi(bt)}{b^2}) \overset{**}\leq  \frac{n\nu}{b^2}\phi(bt).
@@
 
-\begin{definition}
-	For any $x \in \RR$ we define $x_+ \ceq \max(x, 0)$. 
-\end{definition}
+%\begin{definition}
+%	For any $x \in \RR$ we define $x_+ \ceq \max(x, 0)$. 
+%\end{definition}
 
@@
 \[
-\EE \qty[\frac{t^2X^2}{2} + \sum_{q=3}^\infty \frac{t^q \EE[X_+^q]}{q!}] =  \frac{t^2 \Var[X]}{2} + \sum_{q=3}^\infty \frac{t^q \EE[X_+^q]}{q!} \leq \frac{\sigma^2 t^2}{2} \sum_{q=2}^\infty t^{q-2} c^{q-2} = \frac{\sigma^2t^2}{2(1-ct)}.
+\EE \qty[\frac{t^2X^2}{2} + \sum_{q=3}^\infty \frac{t^q X_+^q}{q!}] =  \frac{t^2 \EE[X^2]}{2} + \sum_{q=3}^\infty \frac{t^q \EE[X_+^q]}{q!} \leq \frac{\sigma^2 t^2}{2} \sum_{q=2}^\infty t^{q-2} c^{q-2} = \frac{\sigma^2t^2}{2(1-ct)}.
 \]
